### 2024-02-12 article_jos.qmd
first commit

---
title: "Improving real-time trend estimates using local parametrisation of local polynomial regression filters"
format:
  pdf:
    documentclass: article
    number-sections: true
    pdf-engine: pdflatex
    keep-tex: true
    toc: false
    include-in-header: [preamble.tex]
    #include-before-body: template/preambule.tex
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
    cite-method: biblatex
linestretch: 2
echo: false
tbl-cap-location: top
fig-cap-location: top
bibliography: biblio.bib
csl: chicago-author-date.csl
---

```{r}
#| include: false
library(kableExtra)
```

# Abstract {-}

:::: {.abstract}
This paper studies and compares real-time estimates of the trend cycle component with moving averages based on local polynomial regression.
In particular, this theory allows to reproduce Henderson's symmetric and Musgrave's asymmetric filters.
This paper describes how they can be extended to include a timeliness criterion to minimise the phase shift.
While asymmetric filters are generally parameterised globally, which can be suboptimal around turning points, this paper proposes a procedure for parameterising them locally.
An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parameterisation reduces the delay in detecting turning points and reduces revisions.
The paper also shows how these results can be easily reproduced using the R package `rjd3filters` dedicated to the manipulation of moving averages.

Keywords: time series, trend-cycle, seasonal adjustment, turning points, R statistical software.

JEL Classification: E32, E37.
::::

# Introduction {-}

Analysis of the economic cycle, and in particular the early detection of turning points in a series, is a major topic in the analysis of economic outlook.
To this end, economic indicators are generally seasonally adjusted.
However, in order to improve their interpretability, it may be necessary to perform additional smoothing in order to reduce noise, and thus analyse the trend-cycle component.
By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series.

A moving average, or linear filter, is a statistical method that consists in applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
For example, the X-13ARIMA seasonal adjustment method uses several moving averages to estimate the main components of a time series. 
Symmetric filters are applied to the center of the series, but when it comes to estimate the most recent points, all of these methods must rely on asymmetric filters. 
For trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-13ARIMA seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
For example, for trend-cycle extraction, X-13ARIMA uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
Since the predicted values are linear combinations of past values, this amounts to applying asymmetric averages at the end of the series.

If the classic asymmetric moving averages have good properties regarding the future revisions induced by the process (see for example @pierce1980SA), they create, by construction, phase shifts that impact the real-time estimation of turning points, introducing time delay in the detection.

Using local polynomial regression, @proietti2008 developed general class of symmetric and asymmetric moving averages.
In particular, it makes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
However, these methods have two drawbacks.
Firstly, the phase shift (i.e. the delay in detecting turning points) is not directly controlled.
Secondly, asymmetric filters depend on a parameter that is generally estimated over the whole series using assumptions that may be wrong locally.
The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
The extensions have been implemented in the {{< fa brands r-project >}} `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.

In @sec-propMM, we describe the general properties of moving averages and the associated quality criteria.
This allows us to understand the foundations behind the construction of moving averages from local polynomial regressions, as well as those behind the two extensions proposed in this article (@sec-nonparamreg).
In a final section (@sec-comparison), all these methods are compared empirically on simulated and real series.

# Some properties on moving averages {#sec-propMM}

Lots of papers describe the definition and the properties of moving averages and linear filters, see for example @ch12HBSA. 
In this section we summarize some of the main results to understand the next sections.

Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}$ such as for all time series $X_t$:
$$
M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
$$

- $p+f+1$ is called the *moving average order*.

- When $p=f$ the moving average is said to be *centered*. 
If we also have $\forall k:\:\theta_{-k} = \theta_k$, the moving average $M_{\boldsymbol\theta}$ is said to be *symmetric*. 
In this case, the quantity $h=p=f$ is called the *bandwidth*.

## Gain and phase shift functions {#subsec:gain-deph}

To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $X_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
The moving average $M_{\boldsymbol\theta}$ transforms $X_t$ into:
$$
Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
$$
The function  $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
The frequency response function can equivalently be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ ou $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
It can be rewritten as:
$$
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
$$
where  $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
This function is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to mesure the phase shift in number of periods.
For all symmetric moving average we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.

To sum up, applying a moving average to a harmonic times series ($X_t=\e^{-i\omega t}$) affects it in in two different ways:

- by multiplying it by an amplitude coefficient $\rho_{\boldsymbol\theta}\left(\omega\right)$ (gain);

- by "shifting" it in time by $\varphi_{\boldsymbol\theta}(\omega)/\omega$ which directly affects the detection of turning points.
When  $\varphi_{\boldsymbol\theta}(\omega)/\omega<0$ the time shift is negative: the turning point is detected with delay.

Fourier decomposition allows to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
Noting $\omega = 2\pi/p$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $p$ periods.
For example, for a monthly series (12 observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12[$, i.e. movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component (undesirable oscillations).

@fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e. when no point in the future is known).
The gain function is greater than 1 on the frequencies associated with the trend-cycle ($[0, 2\pi/12]$): this means that the trend-cycle is well preserved and that short cycles of 1 to 2 years ($[2\pi/24,2\pi/12]$) are even amplified.
However, cycles of 8 to 12 months ($[2\pi/12, 2\pi/8]$), considered undesirable because associated with the irregular, are also amplified: this can lead to the detection of false turning points.
At other frequencies, the gain function is less than 1, but always positive: this means that the series smoothed by this moving average will always contain noise, even if it is attenuated.
Analysis of the phase shift shows that the shorter the cycles, the greater the phase shift: this means that on series smoothed by this moving average, turning points could be detected at the wrong date.

```{r, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(fig.pos = "H")
```

```{r, echo=FALSE, out.width="100%"}
#| label: fig-graphsmusgrave
#| fig-cap: "Coefficients, gain and phase shift function for the Musgrave filter in real time with $I/C=3.5$."
img <- sprintf("img/filters_used/%s.%s", tolower("musgrave"), "pdf")
knitr::include_graphics(img)
```

## Desirable properties of a moving average

To decompose a time series into a seasonal component, a trend-cycle and the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses a succession of moving averages, all with specific constraints.

In our case, we assume that our initial series $y_t$ is seasonally adjusted and can be written as the sum of a trend-cycle, $TC_t$, and an irregular component, $I_t$:
$$
y_t=TC_t+I_t.
$$
The aim will be to build moving averages that best preserve the trend-cycle ($M_{\boldsymbol\theta} (TC_t)\simeq TC_t$) and minimise noise ($M_{\boldsymbol\theta} (I_t)\simeq 0$).

### Trend preservation

Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg).
In order to best preserve the trend-cycles, we try to have moving averages that preserve the polynomial trends.
A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t:M_{\boldsymbol\theta} f(t)=f(t)$.

We have the following properties for the moving average $M_{\boldsymbol\theta}$:

- To preserve a constant trend $X_t=a$ we need:
$$
\forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
$$
The sum of the coefficients of the moving average $\sum_{k=-p}^{+f}\theta_k$ must then be equal to $1$.

- To preserve linear trends $X_t=at+b$ we need:
$$
\forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
$$
Which is equivalent to:
$$
\sum_{k=-p}^{+f}\theta_k=1
\quad\text{et}\quad
\sum_{k=-p}^{+f}k\theta_k=0.
$$
- In general, $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
$$
\sum_{k=-p}^{+f}\theta_k=1
 \text{ and }
\forall j \in \left\llbracket 1,d\right\rrbracket:\:
\sum_{k=-p}^{+f}k^j\theta_k=0.
$$
- If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.

### Noise reduction

All time series are affected by noise, which can blur the extraction of trends and cycles.
This is why we try to reduce this noise (by reducing its variance) while preserving the relevant trends.
The sum of the squares of the coefficients $\sum_{k=-p}^{+f}\theta_k^2$ is the *reduction of variance* ratio.

Indeed, let $\{\varepsilon_t\}$ a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$:
$$
\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
$$

## Real-time estimation and asymmetric moving average {#sec-mmasym}

### Asymmetric moving averages and forecasts {#sec-mmetprev}

For symmetric filters, the phase shift function is equal to zero (modulo $\pi$).
So there is no delay in detecting turning points.
However, they cannot be used in the beginning and in the end of the time series because no past or future value can be used. 
Thus, for real-time estimation, it is needed to build asymmetric moving average that approximate the symmetric moving average.

Several approaches can be used for real-time estimation:

1. Use asymmetric moving averages to take account of the lack of available data;

2. Apply symmetrical filters to extended forecast series.
This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period:\
"*As the first $m$ and last $m$ terms of the series cannot be reached directly by the formula, the series should be graphically extended by m terms at both ends, first plotting the observations on paper as ordinates, and then extending the curve along what seems to be its probable course, and measuring the ordinates of the extended portions.*
*It is not necessary that this extension should coincide with what would be the true course of the curve in those parts. *
*The important point is that the m terms thus added, taken together with the $m+1$ adjacent given terms, should follow a curve whose form is approximately algebraic and of a degree not higher than the third.*"\
This is also the approach used in the X-13ARIMA seasonal adjustment method, which extends which extends the series by one year with an ARIMA model.

Ultimately, the second method is equivalent to using asymmetric moving averages, since forecasts are linear combinations of the past.

Conversely, the implicit forecasts of an asymmetric moving average can be deduced from a reference symmetric moving average. 
This allows to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
Let us denote $v=(v_{-h},\dots, v_{h})$ the reference symmetrical moving average and $w^0,\dots w^{h-1}$ a sequence of asymmetrical moving averages, of order $h+1$ to $2h$ used to estimate the last $h$ points with, for convention, $w_t^q=0$ for $t>q$.
This means that $w^0=(w_{-h}^0,\dots, w_{0}^0)$ is used for estimation in real time (when no points are known in the future), $w^1=(w_{-h}^1,\dots, w_{1}^1)$ for estimation of the penultimate point (when only one point is known in the future), and so on.
Let also note $y_{-h},\dots,y_{0}$ the observed series studied and $y_{1}^*,\dots y_h^*$ the implicit forecast induced by $w^0,\dots w^{h-1}$.
This means that for all $q$ we have:
$$
\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
$$
This is equivalent to:
$$
\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
$$
To sum up, in matrix form, this is equivalent to solving:
$$
\scriptstyle
\begin{pmatrix}
  v_1 & v_2 & \cdots & v_h \\
  v_1 - w_1^1 & v_2 & \cdots & v_h \\
  \vdots & \vdots & \cdots & \vdots \\
   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
\end{pmatrix}
\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
\begin{pmatrix}
  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
  \vdots & \vdots & \cdots & \vdots \\
  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
\end{pmatrix}
\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
$$
This is implemented in the `rjd3filters::implicit_forecast()` function.

As highlighted by @wildischis2004, extending the series through forecasting with an ARIMA model is equivalent to calculating asymmetric filters whose coefficients are optimised in relation to the one-step ahead forecast.
In other words, the aim is to minimise the revisions between the first and last estimates (with the symmetric filter).
However, the phase shift induced by the asymmetric filters is not controlled: we might prefer to have faster detection of turning points and a larger revision rather than just minimising the revisions between the first and last estimates.
Furthermore, since the coefficients of the symmetric filter (and therefore the weight associated with distant forecasts) decrease slowly, we should also be interested in the performance of multi-step ahead forecasting. 
This is why it may be necessary to define alternative criteria for judging the quality of asymmetric moving averages.

### Quality indicators for asymmetric moving averages {#subsec:crit-qual}

For asymmetric filters, most of the criteria come from those defined by @ch15HBSA and @trilemmaWMR2019 to build asymmetric filters. 
They are summarised in Table\ \ref{tab-QC} and can be calculated using the `rjd3filters::diagnostic_matrix()` function.

@ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$).
Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::fst_filter()`.

@trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
A simplified version of this method can be used in {{< fa brands r-project >}} with the function `rjd3filters::dfa_filter()`.

```{r, include=FALSE}
title <- "Quality criteria for a moving average $\\boldsymbol\\theta=(\\theta_k)_{-p\\leq k\\leq f}$ defined by a gain function $\\rho_{\\boldsymbol\\theta}$ and a phase shift function $\\varphi_{\\boldsymbol\\theta}$."
footnotes <- c("$X_g$ criteria from Grun-Rehomme, Guggemos, and Ladiray (2018) and $X_w$ criteria from Wildi and McElroy (2019).",
               "$\\rho_s$ and  $\\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.",
               "$h$ e is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\\frac{1}{2(1-\\cos(x))}$)."
)
```

```{r, echo = FALSE, fig.note = footnotes}
#| label: tab-QC
#| eval: false
criteria_mm <- data.frame(Sigle = c("b_c", "b_l", "b_q",
                                    "F_g", "S_g", "T_g",
                                    "A_w", "T_w", "S_w", "R_w"),
                          Description = c(sprintf("%s bias", c("Constant", "Linear", "Quadratic")),
                                          sprintf("%s (Guggemos)",
                                                  c("Variance reduction / Fidelity", "Smoothness", "Timeliness")),
                                          sprintf("%s (Wildi)",c("Accuracy", "Timeliness", "Smoothness", "Residual"))),
                          Formule = c("\\sum_{k=-p}^{+f}\\theta_{k}-1",
                                      "\\sum_{k=-p}^{+f}k\\theta_{k}",
                                      "\\sum_{k=-p}^{+f}k^{2}\\theta_{k}",
                                      "\\sum_{k=-p}^{+f}\\theta_{k}^{2}",
                                      "\\sum_{j}(\\nabla^{3}\\theta_{j})^{2}",
                                      "\\int_{0}^{\\omega_1}\\rho_{\\boldsymbol\\theta}(\\omega)\\sin(\\varphi_{\\boldsymbol\\theta}(\\omega))^{2}\\ud\\omega",
                                      "2\\int_0^{\\omega_1}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega",
                                      "8\\int_0^{\\omega_1} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega",
                                      "2\\int_{\\omega_1}^{\\pi}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega",
                                      "8\\int_{\\omega_1}^{\\pi} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega"))
criteria_mm[,1] <- sprintf("$%s$", criteria_mm[,1])
criteria_mm[,3] <- sprintf("$%s$", criteria_mm[,3])
library(kableExtra)

criteria_mm %>%
  kbl(align = "c", booktabs = T, row.names = FALSE,
        escape = FALSE,caption = title)  %>%
kable_styling(latex_options = "scale_down") 
    # add_footnote_kable(stop_centering = TRUE)
```

```{=tex}
\begin{table}[!h]
\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
{\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ccc}
\toprule
Code & Description & Formula\\
\midrule
$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
\addlinespace
$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
\bottomrule
\end{tabular}}}

\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}

\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}

\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
\end{table}
```

# Non-parametric regression and local polynomial regression {#sec-nonparamreg}

Many trend-cycle extraction methods are based on non-parametric regressions, which are particularly flexible because they do not assume any predetermined dependency in the predictors. 
In practice, local regressions can be used. 
More specifically, consider a set of points $(x_i,y_i)_{1\leq i\leq n}$. 
Non-parametric regression involves assuming that there exists a function $\mu$, to be estimated, such that $y_i=\mu(x_i)+\varepsilon_i$ with $\varepsilon_i$ an error term.
According to Taylor's theorem, for any point $x_0$, if $\mu$ is differentiable $d$ times, then:
$$
\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
$$ {#eq-taylor}
where $R_d$ is a negligible residual term in the neighbourhood of $x_0$. 
In a neighbourhood $h(x_0)$ around $x_0$, $\mu$ can be approximated by a polynomial of degree $d$. 
The quantity $h(x_0)$ is called the *bandwidth*.
If $\varepsilon_i$ is white noise, then we can estimate by least squares $\mu(x_0)$ using the observations which are in $\left[x_0-h(x_0),x_0+h(x_0)\right]$.

In practice, this means assuming that the trend is locally polynomial. 
Various estimation methods can be used to derive symmetrical and asymmetrical moving averages.  
@GrayThomson1996 propose a complete statistical framework which makes it possible, in particular, to model the error in approximating the trend by local polynomials.
However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::rkhs_filter()`.

## Symmetric moving averages and local polynomial regression

Using @proietti2008's notation, we assume that our time series $y_t$ can be decomposed into
$$
y_t=\mu_t+\varepsilon_t,
$$
where $\mu_t$ is the trend and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
Within a neighbourhood $h$ of $t$, the local trend $\mu_t$ is approximated by a polynomial of degree $d$, such that $\mu_t\simeq m_{t}$ with:
$$
\forall j\in\left\llbracket -h,h\right\rrbracket:\:
y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
$$
The problem of trend extraction is equivalent to estimating $m_t=\beta_0$ (the constant in the previous formula).

In matrix notation:
$$
\underbrace{\begin{pmatrix}y_{t-h}\\
y_{t-(h-1)}\\
\vdots\\
y_{t}\\
\vdots\\
y_{t+(h-1)}\\
y_{t+h}
\end{pmatrix}}_{\boldsymbol y}=\underbrace{\begin{pmatrix}1 & -h & h^{2} & \cdots & (-h)^{d}\\
1 & -(h-1) & (h-1)^{2} & \cdots & (-(h-1))^{d}\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & 0 & 0 & \cdots & 0\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & h-1 & (h-1)^{2} & \cdots & (h-1)^{d}\\
1 & h & h^{2} & \cdots & h^{d}
\end{pmatrix}}_{\boldsymbol X}\underbrace{\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\vdots\\
\vdots\\
\vdots\\
\vdots\\
\beta_{d}
\end{pmatrix}}_{\boldsymbol \beta}+\underbrace{\begin{pmatrix}\varepsilon_{t-h}\\
\varepsilon_{t-(h-1)}\\
\vdots\\
\varepsilon_{t}\\
\vdots\\
\varepsilon_{t+(h-1)}\\
\varepsilon_{t+h}
\end{pmatrix}}_{\boldsymbol \varepsilon}.
$$

To estimate $\boldsymbol \beta$ we need $2h+1\geq d+1$ and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
$$
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
$$
where $\kappa_j$ is a set of weights called *kernel*.
We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
With $\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}$, the estimate of the trend is:
$$
\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
$$ {#eq-mmsym}
To conclude, the estimate of the trend $\hat{m}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
Moreover, $\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}$ so:
$$
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
$$
Hence, the filter $\boldsymbol w$ preserve deterministic polynomial of order $d$.

Regarding parameter selection, the general consensus is that the choice between different kernels is not crucial.
See, for example, @cleveland1996smoothing or @Loader1999.
The only desired constraints on the kernel are that it assigns greater weight to the central estimation ($\kappa_0$) and decreases towards 0 as it moves away from the central estimation. The uniform kernel should therefore be avoided.
It is then preferable to focus on two other parameters:

- the degree of the polynomial, denoted by $d$: if it is too small, there is a risk of biased estimates of the trend-cycle, and if it is too large, there is a risk of excessive variance in the estimates (due to over-adjustment);

- the number of neighbors $2h+1$ (or the window $h$): if it is too small, then there will be too little data for the estimates (resulting in high variance in the estimates), and if it is too large, the polynomial approximation will probably be wrong, leading to biased estimates.

In this paper we will use the Henderson kernel:
$$
\kappa_{j}=\left[1-\frac{j^2}{(h+1)^2}\right]
\left[1-\frac{j^2}{(h+2)^2}\right]
\left[1-\frac{j^2}{(h+3)^2}\right].
$$
However, several other kernels are available in `rjd3filters`.

## Asymmetric moving averages and local polynomial regression {#sec-lppasymf}

As mentioned in @sec-mmetprev, for real-time estimation, several approaches can be used:

1. Apply symmetric filters to the series extended by forecasting $\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket$.

2. Build an asymmetric filter by local polynomial approximation on the available observations ($y_{t}$ for $t\in\left\llbracket n-h,n\right\rrbracket$).

3. Build asymmetric filters that minimise the mean squared error of revision under polynomial trend reproduction constraints.

@proietti2008 show that the first two approaches are equivalent when forecasts are made by polynomial extrapolation of degree $d$.
They are also equivalent to the third approach under the same constraints as those of the symmetric filter.
The third method is called *direct asymmetric filter* (DAF).
This method is used for real-time estimation in the STL seasonal adjustment method (*Seasonal-Trend decomposition based on Loess*, see @cleveland90).
Although DAF filter estimates are unbiased, this is at the cost of greater variance in the estimates.

To solve the problem of the variance of real-time filter estimates, @proietti2008 propose a general method for constructing asymmetric filters that allows a bias-variance trade-off.
This is a generalisation of @musgrave1964set's asymmetric filters (used in the X-13ARIMA seasonal adjustment algorithm).

The initial series is modelled here as:
$$
\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
$$ {#eq-lpgeneralmodel}
where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$.
The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
These constraints are represented by the matrix $\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
The problem is equivalent to finding $v$ which minimises:
$$
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
$$ {#eq-lppasym}
where $\boldsymbol l$ is a vector of Lagrange multipliers.

When $\boldsymbol U=\boldsymbol X$, the constraint is equivalent to preserving polynomials of degree $d$: we find asymmetric direct filters (DAF) when $\boldsymbol D=\boldsymbol K^{-1}$.

When $\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}$, $\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
This filter assumes that, for real-time estimation, the data is generated by a linear process and that the asymmetric filters preserve the constants ($\sum v_i=\sum \theta_i=1$).
These asymmetric filters depend on the ratio $\lvert\delta_1/\sigma\rvert$, which, assuming that the trend is linear and the bias constant, can be linked to the I-C ratio $R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}$: $\delta_1/\sigma=2/(R\sqrt{\pi})$.
This ratio is used in particular in X-13ARIMA to determine the length of the Henderson filter. 
For monthly data:

- If the ratio is large ($3.5< R$) then a 23 term filter is used (to remove more noise).

- If the ratio is small ($R<1$) a 9 term filter is used.

- Otherwise (most of the cases) a 13-term filter is used.

When $\boldsymbol U$ corresponds to the first $d^*+1$ columns of $\boldsymbol X$, $d^*<d$, the constraint is to reproduce polynomial trends of degree $d^*$.
This introduces bias but reduces the variance. 
Thus, @proietti2008 proposes three classes of asymmetric filters:

1. *Linear-Constant* (LC): $y_t$ linear ($d=1$) and $\boldsymbol v$ preserves constant signals ($d^*=0$). 
We obtain Musgrave filters when the Henderson kernel is used and the Henderson filter is used as the symmetric filter.

2. *Quadratic-Linear* (QL): $y_t$ quadratic ($d=2$) and $\boldsymbol v$ preserves linear signals ($d^*=1$).

3. *Cubic-Quadratic* (CQ): $y_t$ cubic ($d=3$) and $\boldsymbol v$ preserves quadratic signals ($d^*=2$).

Appendix [-@sec-an-cof] shows the coefficients, gain and phase shift functions of the four asymmetric filters.

Table\ \ref{tab-criteriaLp} compares the quality criteria of the different methods using the Henderson filter and $h=6$ (a symmetric filter with 13 terms). 
For real-time filters ($q=0$), a more complex asymmetrical filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance. 
This result varies as $q$ increases: for $q=2$, the QL filter has greater timeliness than the LC filter. 
This surprising result highlights the fact that the phase shift is not controlled by @proietti2008's approach.

In terms of revision ($A_w+S_w+T_w+R_w$), the LC and QL filters consistently outperform the CQ and DAF filters.

These theoretical properties are consistent with the empirical results observed in @sec-comparison. Revisions are more important for the CQ and DAF filters, which leads to greater variability in estimations and a longer delay in detecting turning points. 
The globally parameterised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter.
Indeed, to identify a turning point at date $t$, it is necessary to have knowledge of at least two points after the given date to confirm the reversal in trend.

```{r echo = FALSE}
#| label: tab-criteraLp
#| eval: false
#| tab-cap: 
title <- "Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$."

lp_diagnostics <- readRDS("data/lp_diagnostics_henderson.RDS")
lp_diagnostics[,"$ EQM_w $"] <- rowSums(lp_diagnostics[,grep("_w", colnames(lp_diagnostics), fixed = TRUE)])
colnames(lp_diagnostics) <- gsub(" ?\\$ ?","$",colnames(lp_diagnostics))
lp_diagnostics[,1] <- gsub(" ?\\$ ?","$",lp_diagnostics[,1])
groupement <- table(lp_diagnostics[,1])
lp_diagnostics[,-(1:2)] <- round(lp_diagnostics[,-(1:2)],2)
lp_diagnostics[,-1] %>%
  kable(format.args = list(digits = 2,
                           decimal.mark = "."), align = "c", booktabs = T, row.names = FALSE,
        escape = FALSE,caption = title,format = "latex") %>%
  kable_styling(latex_options=c(#"striped",
      "hold_position")) %>%
  pack_rows(index = groupement, escape = FALSE) %>%
  footnote("With $EQM_w=A_w + S_w + T_w + R_w$.")
```

```{=tex}
\begin{table}[!h]
\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
\centering
\begin{tabular}[t]{cccccccccccc}
\toprule
Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
\midrule
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=0$}}\\
\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=1$}}\\
\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=2$}}\\
\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
\bottomrule
\end{tabular}

\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
\end{table}
```

## Extension with the timeliness criterion {#sec-lptimeliness}

One drawback of the previous method is the lack of control over the phase shift. 
However, it is possible to improve the modelling by incorporating in @eq-lppasym the timeliness criterion defined by @ch15HBSA.
This was proposed by Jean Palate, then coded in Java and integrated into `rjd3filters`.

Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
With $\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
$$
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
$$
Furthermore, the objective function $\varphi$ of @eq-lppasym can be rewritten:
\begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*}
Adding the timeliness criterion:
$$
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
$$
where $\alpha_T$ is the weight associated to the timeliness criterion. 
With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
This extension is implemented in the {{< fa brands r-project >}} function `rjd3filters::lp_filter()`.

One drawback is that $T_g(\boldsymbol v)$ is not normalized: the weight $\alpha_T$ has no economic sense.
It is then difficult to calibrate the value of $\alpha_T$ since the value of $\lvert\delta/\sigma\rvert$ must also be defined.
That's why in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parameterisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).

## Local parameterisation of asymmetric filters {#sec-localic}

Asymmetric filters are usually parameterised globally: $\lvert\delta/\sigma\rvert$ estimated on all the data using the IC-ratio or a cross-validation criterion. 
However, we might prefer a local parameterisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
Indeed, although the overall parameterisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetrical filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
For example, with the LC method, global parameterisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up to the turning point.

This is what is proposed in this article, with a local parameterisation of the asymmetric filters by estimating $\delta$ and $\sigma^2$ separately.
Although this does not give an unbiased estimator of the ratio $\lvert\delta/\sigma\rvert$, it does allow the main evolutions to be captured, such as the decay towards 0 before a turning point and the growth after the turning point for the LC method:

- The variance $\sigma^2$ can be estimated using the observed data set and the symmetrical filter $(w_{-p},\dots,w_p)$:
$$
\hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
$$
- The parameter $\delta$ can be estimated by moving average from @eq-mmsym. 
For example, for the LC method we can use the moving average $\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}$ to obtain a local estimate of the slope and for the QL method we can use $\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}$ to obtain a local estimate of the concavity. 
The DAF method then simply allows us to calculate the associated asymmetric moving averages.  
Although a moving average of different length to that used for estimating the trend could be considered, this seems to degrade the results in terms of phase shift (using the same methodology as in @sec-comparison).
Furthermore, for the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
Here, we have chosen to model a trend of degree 2: this reduces the phase shift but slightly increases the revisions linked to the first estimate of the trend-cycle. 
@fig-mmpenteconcac shows the moving averages used.

```{r, echo=FALSE, out.width="90%"}
#| label: fig-mmpenteconcac
#| fig-cap: Moving averages used for real-time estimation of slope and concavity.
img <- sprintf("img/filters_used/mm_penteconcavite.%s", "pdf")
knitr::include_graphics(img)
```

In the empirical applications of @sec-comparison, the final local parameterisation is obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two the local parameterisation:

- @fig-mmpenteconcac-ex-1: using a symmetric filter to estimate $\delta_1$ (final local parameterisation).

- @fig-mmpenteconcac-ex-2: using the asymmetric filter which need two points in the futur to estimate $\delta_1$.
This is closed to the real-time estimates since two points in the future are needed to detect a turning point.

The turning points are clearly detected by the local estimators, but there is much more variance in the estimate of $|\delta_1/\sigma|$ in the real-time estimate.
This will lead to another source of revisions in the intermediate estimates.

```{r, echo=FALSE, out.width="90%"}
#| label: fig-mmpenteconcac-ex
#| fig-cap: "Comparison of the estimates of $|\\delta_1/\\sigma|$ with a local estimator and global estimator (with IC ratio) on a simulated series, the vertical lines being the simulated turning points."
#| fig-subcap: 
#|   - "Symmetric moving average used for $\\hat \\delta_1$."
#|   - "Asymmetric moving average using two futur points ($q=2$) used for $\\hat \\delta_1$."
#| layout-ncol: 1
knitr::include_graphics(c("img/filters_used/mm_penteconcavite_ex.pdf",
                          "img/filters_used/mm_penteconcavite_ex2.pdf"))
```

# Comparison of different methods {#sec-comparison}

The different methods are compared on simulated and real data.

For all series, a symmetrical 13-term filter is used.
These methods are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetrical Henderson filter.
The ARIMA model is determined automatically, using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).

The performance of the different methods is judged on criteria relating to revisions (between two consecutive estimates and in relation to the final estimate) and the number of periods required to detect turning points.

## Simulated series

### Methodology

Following a methodology close to that of @DarneDagum2009, nine monthly series are simulated between January 1960 and December 2020 with different levels of variability. 
Each simulated series $y_t= C_t+ T_t + I_t$ can be written as the sum of three components:

- the cycle $C_t = \rho [\cos (2 \pi t / \lambda) +\sin (2 \pi t / \lambda)]$, $\lambda$ is fixed at 72 (6-year cycles, so there are 19 detectable turning points);

- the trend $T_t = T_{t-1} + \nu_t$ with $\nu_t \sim \mathcal{N}(0, \sigma_\nu^2)$, $\sigma_\nu$ being fixed at $0.08$;

- and the irregular $I_t = e_t$ with $e_t \sim \mathcal{N}(0, \sigma_e^2)$.

For the different simulations, we vary the parameters $\rho$ and $\sigma_e^2$ in order to obtain series with different signal-to-noise ratios (see @fig-graphs-data-simul):

- High signal-to-noise ratio (i.e. low I-C ratio and low variability): $\sigma_e^2=0.2$ and $\rho = 3.0, 3.5$ or $4.0$ (I-C ratio between 0.9 and 0.7);

- Medium signal-to-noise ratio (i.e. medium I-C ratio and medium variability): $\sigma_e^2=0.3$ and $\rho = 1.5,\, 2.0$ or $3.0$ (I-C ratio between 2.3 and 1.4);

- Low signal-to-noise ratio (i.e. high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).

```{r, echo=FALSE, out.width="100%"}
#| label: fig-graphs-data-simul
#| fig-cap: "Simulated series with low ($\\sigma_e^2=0.2$ and $\\rho = 3,5$), medium ($\\sigma_e^2=0.3$ and $\\rho = 2.0$) and high variability ($\\sigma_e^2=0.4$ and $\\rho = 1.0$)."
img <- sprintf("img/simulations/simul_data.%s", "pdf")
knitr::include_graphics(img)
```

For each series and each date, the trend-cycle is estimated using the different methods presented in this report.
Three quality criteria are also calculated:

1. Computation of the phase shift in the detection of turning points. The definition of @Zellner1991 is used to determine the turning points:  
    - An upturn occurs when the economy moves from a phase of recession to a phase of expansion. 
    This is the case at date  $t$ when $y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}$.  
    - A downturn occurs when the economy moves from a phase of expansion to a phase of recession.
This is the case at date  $t$ when $y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}$.  
    It thus takes at least 2 months to detect a turning point.  
    The phase shift is often defined as the number of months required to detect the right turning point.
    Here we use a slightly modified criterion: the phase shift is defined as the number of months needed to detect the right turning point without any future revision.
    Indeed, it can happen that the right turning point is detected by asymmetric filters but is not detected by the final estimate using a symmetric filter (this is the case for 41 reversal points out of the 9 series with asymmetric Musgrave filters) or that there are revisions in successive estimates (this is the case for 7 turning points out of the 9 series with asymmetric Musgrave filters).
  Finally, relatively few turning points are detected at the right date with the final estimate.
  With the 13-term Henderson filter, 18 are correctly detected in series with low variability (out of 57 possible), 11 in series with medium variability and 12 in series with high variability.

2. Computation of two revision criteria: the average of the relative deviations between the $q$^th^ estimate and the last estimate $MAE_{fe}(q)$ and the average of the relative deviations between the $q$^th^ estimate and the $q+1$^th^ estimate $MAE_{qe}(q)$.
$$
MAE_{fe}(q)=\mathbb E\left[
\left|\frac{
y_{t|t+q} -  y_{t|last}
}{
 y_{t|last}
}\right|
\right]
\quad\text{and}\quad
MAE_{qe}(q)=\mathbb E\left[
\left|\frac{
y_{t|t+q} - y_{t|t+q+1}
}{
y_{t|t+q+1}
}\right|
\right].
$$

## Comparison

Excluding for the moment the local parameterisations of the polynomial filters, the linear-constant polynomial filter (LC) seems to give the best results in terms of delay in the detection of turning points (@fig-graphstpsimul).
Performance is relatively close to that obtained by extending the series using an ARIMA model.
However, when variability is low, the LC filter seems to give poorer results and the quadratic-linear polynomial filter (QL) seems to give the best results.

For series with moderate variability, local parameterisation of the LC and QL filters reduces the phase shift.
For series with high variability, the phase shift is only reduced by using the final $\hat\delta$ parameters: real-time estimates seem to add more variance.
For series with low variability, performance seems to be slightly improved only with the LC filter.

```{r, echo=FALSE, out.width="100%", fig.note = footnotes}
#| label: fig-graphstpsimul
#| fig-cap: Distribution of phase shifts on simulated series.
img <- sprintf("img/simulations/phase_shift_simul.%s", "pdf")
knitr::include_graphics(img)
```

In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability (Table\ \ref{tab-simulrev}).
In general, LC filters always minimise revisions (with relatively small effect from the local parameterisation of the filters) and revisions are greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.

For the QL filter, there is a large revision between the second and third estimates: this may be due to the fact that for the second estimate (when one point in the future is known), the QL filter assigns a greater weight to the estimate in $t+1$ than to the estimate in $t$, which creates a discontinuity. 
This revision is considerably reduced by parameterising the filter locally.
For polynomial filters other than the LC filter, the large revisions at the first estimate were to be expected given the coefficient curve: a very large weight is associated with the current observation and there is a strong discontinuity between the moving average used for real-time estimation (when no point in the future is known) and the other moving averages.

Extending the series using an ARIMA model gives revisions with the latest estimates of the same order of magnitude as the LC filter, but slightly larger revisions between consecutive estimates, particularly between the fourth and fifth estimates (as might be expected, as highlighted in @sec-mmetprev).

```{r, echo = FALSE}
#| label: tab-simulrev
#| tbl-cap: "Average of the relative deviations of the revisions for the different filters on simulated series with medium variability."
#| eval: false
rev_table <- readRDS("data/simulations_revisions.RDS")
pack_row_index = c("$MAE_{fe}(q) = \\mathbb E\\\\left[\\\\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\\\\right|\\\\right]$" = nrow(rev_table) / 2,
                   "$MAE_{ce}(q)=\\mathbb E\\\\left[
\\\\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\\\\right|
\\\\right]$" = nrow(rev_table) / 2)
p <- rev_table  %>%
  kable(format.args = list(digits = 2,
                           decimal.mark = "."),
        align = "c", booktabs = T, row.names = FALSE, 
        label = "tab-simulrev",
        escape = FALSE,
        format = "latex") %>%  
  kable_styling(latex_options=c(#"striped",  
    "hold_position")) %>% 
  pack_rows(index = pack_row_index, escape = FALSE) # %>%
    # add_footnote_kable(stop_centering = TRUE)
```

```{=tex}
\begin{table}[!h]
\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
filters on simulated series with medium variability.}
\centering
\begin{tabular}{ccccccc}
\toprule
Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
\right]$}}\\
\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
\bottomrule
\end{tabular}
\end{table}
```

## Real series

The differences between the methods are also illustrated using an example from the FRED-MD database (@fredmd) containing economic series on the United States.
The series studied correspond to the database published in November 2022.
The series studied is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
This point and this series were chosen because the economic turnaround is particularly visible in the raw series (@fig-ce16ov-previmp-lp), because this series is used for dating economic cycles and because the FRED-MD database facilitates the reproducibility of the results (thanks to the availability of series published on past dates).
Studying the series up to January 2020, this series has a medium variability: a symmetrical filter of 13 terms is therefore appropriate.
Appendix [-@ann-ex-r] shows how the code to reproduce the different plots.

@fig-ce16ovlp shows the successive estimates of the trend-cycle using the different methods studied.

For this series, the phase shift is 6 months for the LC and CQ methods and the extension of the series by ARIMA. 
Except for the LC method (where the turning point is first detected in January 2001), the high phase shift is due to revisions in the turning point detected in the intermediate estimates (even if the correct turning point is detected after 2 months, it is no longer detected after 3 months).
It is of two months for the other methods (QL, DAF).

In this example, local parameterisation does not reduce the phase shift, but it does reduce revisions.
The QL and DAF polynomials lead to greater variability in the intermediate estimates, especially in February 2001.

```{r, out.width="95%", fig.show="hold"}
#| label: fig-ce16ovlp
#| fig-cap:  Successive estimates of the trend-cycle in US employment (in logarithms).
series <- "CE16OV"
date <- "fev2001"
img <- sprintf("img/nber/%s_%s_%s.%s",
               tolower(series),
               date,
               c("lp"),
               "pdf")
knitr::include_graphics(img)
```

The quality of the intermediate estimates can also be analysed using the implicit forecasts of the different methods (@fig-ce16ov-previmp-lp). 
As a reminder, these are forecasts of the raw series which, by applying the symmetrical Henderson filter to the extended series, give the same estimates as the asymmetrical moving averages. 
The forecasts of the ARIMA model are naive and do not take the turning point into account, unlike the other methods.
Finally, local parameterisation of the QL filter produces much more consistent forecasts.

```{r, out.width="95%"}
#| label: fig-ce16ov-previmp-lp
#| fig-cap: Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
series <- "CE16OV"
date <- "fev2001"
img <- sprintf("img/nber/%s_%s_prev_imp_%s.%s",
               tolower(series),
               date,
               c("lp"),
               "pdf")
knitr::include_graphics(img)
```

# Conclusion {.unnumbered}

For business cycle analysis, most statisticians use trend-cycle extraction methods, either directly or indirectly.
They are used, for example, to reduce the noise of an indicator in order to improve its analysis, and models, such as forecasting models, usually use seasonally adjusted series based on these methods.

This paper presents the {{< fa brands r-project >}} package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
It also offers various functions for the theoretical analysis of moving averages (gain functions, phase shift, quality criteria, etc.) and judging the quality of the latest estimates (for example with implicit forecasts).
It also makes it easy to combine moving averages and also to integrate custom moving averages into the X-11 seasonal adjustment algorithm (with the `rjd3filters::x11plus_trend()` function).
`rjd3filters` thus facilitates research into the use of moving averages as part of real-time cycle trend estimation.
This is the case, for example, with the local parametrisation of Musgrave moving averages and other asymmetric moving averages linked to local polynomial regression, presented in this article.

By comparing the different methods, we can learn a few lessons about the construction of these moving averages.

During economic downturns, asymmetric filters used as an alternative to extending the series using the ARIMA model can reduce revisions to intermediate estimates of the trend-cycle and enable turning points to be detected more quickly.

At the end of the period, modelling polynomial trends of degree greater than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce variance into the estimates (and therefore more revisions) without allowing faster detection of turning points.
At the end of the period, for estimating the trend-cycle, we can therefore restrict ourselves to methods modelling polynomial trends of degree two or less (linear-constant, LC, and quadratic-linear, QL).
In addition, parameterising polynomial filters locally as suggested in this paper enables turning points to be detected more quickly (especially for the QL filter).
Even when the phase shift is not reduced, local parameterisation is recommended because it reduces revisions and produces intermediate estimates that are more consistent with expected future trends.
However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too short (i.e. if the variability of the series is "low"), retaining polynomial trends of degree one or less (LC method) produces poorer results in terms of detecting turning points.

This study could be extended in many ways.

One possible extension would be to look at the impact of filter length on the detection of turning points.
Asymmetric filters are calibrated using indicators calculated for the estimation of symmetric filters (for example, to automatically determine their length), whereas a local estimate might be preferable.
Furthermore, we have only focused on monthly series with a 13-term symmetric filter, but the results may be different if the symmetric filter studied is longer or shorter and if we study series with other frequencies (quarterly or daily, for example).

Another possibility would be to study the impact of atypical points: moving averages, like any linear operator, are very sensitive to the presence of atypical points. 
To limit their impact, in X-13ARIMA a strong correction for atypical points is performed on the irregular component before applying the filters to extract the trend-cycle.
This leads us to study the impact of these points on the estimation of the trend-cycle and turning points, and also to explore new types of asymmetric filters based on robust methods (such as robust local regressions or moving medians).

\newpage

# References {.unnumbered}

::: {#refs}
:::

\appendix
\newpage

# Coefficients, gain and phase shift functions {#sec-an-cof}

```{r, out.width="100%"}
#| label: fig-graphs-coef-lc
#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-ql
#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-cq
#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$.
img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf")
knitr::include_graphics(img)
```

```{r, out.width="100%"}
#| label: fig-graphs-coef-daf
#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF).
img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf")
knitr::include_graphics(img)
```

# R code example {#ann-ex-r}

In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.

```{r, eval=FALSE, echo=TRUE}
#| eval: false
#| echo: true
#| file: R_ex.R
```

### 2024-02-12 article_jos.tex
first commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=2cm,bottom=2cm,left = 2.5cm,right = 2.5cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{stmaryrd}
\usepackage{multicol}
\usepackage{graphicx}
\usepackage{animate, dsfont, here, xspace}
%\usepackage{tikz}
\usepackage{tikz,pgfplots}
 \pgfplotsset{compat=1.17}
%\includepdf[fitpaper=true, pages=-]{img/pdg.pdf}

\DeclareMathOperator{\e}{e}
\DeclareMathOperator{\Determinant}{det}
\renewcommand{\P}{\mathds{P}} %Apparement \P existe déjà ?
\newcommand\N{\mathds{N}}
\newcommand\Norm{\mathcal{N}}
\newcommand\R{\mathds{R}}
\newcommand\Z{\mathds{Z}}
\newcommand{\determinant}[1]{\Determinant\left(#1\right)}
%\newcommand\C{\mathds{C}}
%\newcommand\Z{\mathds{Z}}

\newcommand\1{\mathds{1}}
\newcommand{\E}[2][]{{\mathds{E}}_{#1}
  \def\temp{#2}\ifx\temp\empty
  \else
    \left[#2\right]
  \fi
}
\newcommand{\V}[2][]{{\mathds{V}}_{#1}
  \def\temp{#2}\ifx\temp\empty
  \else
    \left[#2\right]
  \fi
}
\newcommand\ud{\,\mathrm{d}}
\newcommand{\ps}[2]{\left\langle #1 \,,\, #2 \right\rangle}

% blocks
\usepackage{environ}
\usepackage[tikz]{bclogo}

\tikzstyle{titlestyle} =[draw=black!80,fill=black!20, text=black,
 right=10pt, rounded corners]
\mdfdefinestyle{symmaryboxstyle}{
	linecolor=black!80, backgroundcolor = black!5,
	skipabove=\baselineskip, innertopmargin=\baselineskip,
	innerbottommargin=\baselineskip,
	userdefinedwidth=\textwidth,
	middlelinewidth=1.2pt, roundcorner=5pt,
	skipabove={\dimexpr0.5\baselineskip+\topskip\relax},
	frametitleaboveskip=\dimexpr-\ht\strutbox\relax,
	innerlinewidth=0pt,
}
\newcounter{summarybox}
\NewEnviron{summary_box}[2][true]{%
\refstepcounter{summarybox} % on incrémente le compteur
\begin{mdframed}[style=symmaryboxstyle,
nobreak=#1,
frametitle={%
      \tikz[baseline=(current bounding box.east),outer sep=0pt]
      \node[titlestyle,anchor=east]
    {Encadré \thesummarybox{} --- #2};}
]
\vspace{-0.5em}
\BODY
\end{mdframed}
}

\usepackage{amsthm}
%\theoremstyle{remark}
\newtheorem*{remarque}{Remarque}

\usepackage{mathrsfs}
\usepackage{fontawesome5}

%\usepackage[style=authoryear,%,uniquename=false, uniquelist=false,
%maxbibnames=2]{biblatex-chicago}
% \usepackage[options]{natbib}
% \bibliographystyle{chicago}

%\DefineBibliographyStrings{english}{andothers={et\addabbrvspace alii}}
\usepackage[authordate16,backend=biber,maxcitenames=2]{biblatex-chicago}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{biblatex}
\addbibresource{biblio.bib}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Improving real-time trend estimates using local parametrisation of local polynomial regression filters},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Improving real-time trend estimates using local parametrisation
of local polynomial regression filters}
\author{}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, frame hidden]}{\end{tcolorbox}}\fi

\setstretch{2}
\hypertarget{abstract}{%
\section*{Abstract}\label{abstract}}
\addcontentsline{toc}{section}{Abstract}

This paper studies and compares real-time estimates of the trend cycle
component with moving averages based on local polynomial regression. In
particular, this theory allows to reproduce Henderson's symmetric and
Musgrave's asymmetric filters. This paper describes how they can be
extended to include a timeliness criterion to minimise the phase shift.
While asymmetric filters are generally parameterised globally, which can
be suboptimal around turning points, this paper proposes a procedure for
parameterising them locally. An empirical comparison, based on simulated
and real data, shows that modelling polynomial trends that are too
complex introduces more revisions without reducing the phase shift, and
that local parameterisation reduces the delay in detecting turning
points and reduces revisions. The paper also shows how these results can
be easily reproduced using the R package \texttt{rjd3filters} dedicated
to the manipulation of moving averages.

Keywords: time series, trend-cycle, seasonal adjustment, turning points,
R statistical software.

JEL Classification: E32, E37.

\hypertarget{introduction}{%
\section*{Introduction}\label{introduction}}
\addcontentsline{toc}{section}{Introduction}

Analysis of the economic cycle, and in particular the early detection of
turning points in a series, is a major topic in the analysis of economic
outlook. To this end, economic indicators are generally seasonally
adjusted. However, in order to improve their interpretability, it may be
necessary to perform additional smoothing in order to reduce noise, and
thus analyse the trend-cycle component. By construction, trend-cycle
extraction methods are closely related to seasonal adjustment methods,
since they are generally applied to seasonally adjusted series.

A moving average, or linear filter, is a statistical method that
consists in applying a rolling weighted mean to a times series: for each
date \(t\) it computes a weighted mean of \(p\) past points and \(q\)
future points where \(p,q\geq0\) depends on the moving average. Moving
averages are ubiquitous in business cycle extraction and seasonal
adjustment methods. For example, the X-13ARIMA seasonal adjustment
method uses several moving averages to estimate the main components of a
time series. Symmetric filters are applied to the center of the series,
but when it comes to estimate the most recent points, all of these
methods must rely on asymmetric filters. For trend-cycle extraction, the
most popular symmetric filter is the \textcite{henderson1916note} moving
average, which is used in the X-13ARIMA seasonal adjustment algorithm.

However, for real-time estimates, due to the lack of future
observations, all these methods must rely on asymmetric filters to
estimate the most recent points. For example, for trend-cycle
extraction, X-13ARIMA uses Henderson's symmetric filter and
\textcite{musgrave1964set}'s asymmetric filters on an extended series
using an ARIMA model. Since the predicted values are linear combinations
of past values, this amounts to applying asymmetric averages at the end
of the series.

If the classic asymmetric moving averages have good properties regarding
the future revisions induced by the process (see for example
\textcite{pierce1980SA}), they create, by construction, phase shifts
that impact the real-time estimation of turning points, introducing time
delay in the detection.

Using local polynomial regression, \textcite{proietti2008} developed
general class of symmetric and asymmetric moving averages. In
particular, it makes possible to reproduce Henderson's symmetric filter
and Musgrave's asymmetric filters. However, these methods have two
drawbacks. Firstly, the phase shift (i.e.~the delay in detecting turning
points) is not directly controlled. Secondly, asymmetric filters depend
on a parameter that is generally estimated over the whole series using
assumptions that may be wrong locally. The aim of this study is to
propose extensions to the \textcite{proietti2008} class of filters in
order to take these shortcomings into account. The extensions have been
implemented in the \faIcon{r-project} \texttt{rjd3filters} package
(available at \url{https://github.com/rjdemetra/rjd3filters}) and are
therefore easily reusable.

In Section~\ref{sec-propMM}, we describe the general properties of
moving averages and the associated quality criteria. This allows us to
understand the foundations behind the construction of moving averages
from local polynomial regressions, as well as those behind the two
extensions proposed in this article (Section~\ref{sec-nonparamreg}). In
a final section (Section~\ref{sec-comparison}), all these methods are
compared empirically on simulated and real series.

\hypertarget{sec-propMM}{%
\section{Some properties on moving averages}\label{sec-propMM}}

Lots of papers describe the definition and the properties of moving
averages and linear filters, see for example \textcite{ch12HBSA}. In
this section we summarize some of the main results to understand the
next sections.

Let \(p\) et \(f\) two integers, a moving average
\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
\(\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}\)
such as for all time series \(X_t\): \[
M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
\]

\begin{itemize}
\item
  \(p+f+1\) is called the \emph{moving average order}.
\item
  When \(p=f\) the moving average is said to be \emph{centered}. If we
  also have \(\forall k:\:\theta_{-k} = \theta_k\), the moving average
  \(M_{\boldsymbol\theta}\) is said to be \emph{symmetric}. In this
  case, the quantity \(h=p=f\) is called the \emph{bandwidth}.
\end{itemize}

\hypertarget{subsec:gain-deph}{%
\subsection{Gain and phase shift functions}\label{subsec:gain-deph}}

To interpret the notions of gain and phase shift, it is useful to
illustrate the effects of moving averages on harmonic series
\(X_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
\(M_{\boldsymbol\theta}\) transforms \(X_t\) into: \[
Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
\] The function
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
is called the \emph{transfer function} or \emph{frequency response
function}. The frequency response function can equivalently be defined
as
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
ou
\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
It can be rewritten as: \[
\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
\] where
\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
is the \emph{gain} or \emph{amplitude} function and
\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
\emph{time shift} function. This function is sometimes represented as
\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
to mesure the phase shift in number of periods. For all symmetric moving
average we have
\(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).

To sum up, applying a moving average to a harmonic times series
(\(X_t=\e^{-i\omega t}\)) affects it in in two different ways:

\begin{itemize}
\item
  by multiplying it by an amplitude coefficient
  \(\rho_{\boldsymbol\theta}\left(\omega\right)\) (gain);
\item
  by ``shifting'' it in time by
  \(\varphi_{\boldsymbol\theta}(\omega)/\omega\) which directly affects
  the detection of turning points. When
  \(\varphi_{\boldsymbol\theta}(\omega)/\omega<0\) the time shift is
  negative: the turning point is detected with delay.
\end{itemize}

Fourier decomposition allows to analyze any time series as a sum of
harmonic series, and each component (trend, cycle, seasonal, irregular)
is associated with a set of frequencies. Noting \(\omega = 2\pi/p\), the
harmonic series of frequency \(\omega\) represents a series that repeats
itself every \(p\) periods. For example, for a monthly series (12
observations per year), seasonal movements are those that repeat every
year: they are therefore associated with frequencies \(2\pi/12\) (annual
periodicity), \(2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5\). In this
paper, we consider that the trend-cycle is associated with frequencies
in the interval \([0, 2\pi/12[\), i.e.~movements recurring at least
every 12 months. Even if different frequencies are sometimes used
(e.g.~\([0, 2\pi/36]\) to consider only cycles of at least 36 months),
this has no impact on the results of the study. The other frequencies,
\([2\pi/12, \pi]\) are associated with the irregular component
(undesirable oscillations).

Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
for the asymmetric Musgrave filter (see Section~\ref{sec-lppasymf})
often used for real-time trend-cycle estimation (i.e.~when no point in
the future is known). The gain function is greater than 1 on the
frequencies associated with the trend-cycle (\([0, 2\pi/12]\)): this
means that the trend-cycle is well preserved and that short cycles of 1
to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. However, cycles
of 8 to 12 months (\([2\pi/12, 2\pi/8]\)), considered undesirable
because associated with the irregular, are also amplified: this can lead
to the detection of false turning points. At other frequencies, the gain
function is less than 1, but always positive: this means that the series
smoothed by this moving average will always contain noise, even if it is
attenuated. Analysis of the phase shift shows that the shorter the
cycles, the greater the phase shift: this means that on series smoothed
by this moving average, turning points could be detected at the wrong
date.

\begin{figure}[H]

\caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
function for the Musgrave filter in real time with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}

}

\end{figure}

\hypertarget{desirable-properties-of-a-moving-average}{%
\subsection{Desirable properties of a moving
average}\label{desirable-properties-of-a-moving-average}}

To decompose a time series into a seasonal component, a trend-cycle and
the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses
a succession of moving averages, all with specific constraints.

In our case, we assume that our initial series \(y_t\) is seasonally
adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
an irregular component, \(I_t\): \[
y_t=TC_t+I_t.
\] The aim will be to build moving averages that best preserve the
trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)) and minimise
noise (\(M_{\boldsymbol\theta} (I_t)\simeq 0\)).

\hypertarget{trend-preservation}{%
\subsubsection{Trend preservation}\label{trend-preservation}}

Trend-cycles are generally modelled by local polynomial trends (see
Section~\ref{sec-nonparamreg}). In order to best preserve the
trend-cycles, we try to have moving averages that preserve the
polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
a function of time \(f(t)\) if
\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).

We have the following properties for the moving average
\(M_{\boldsymbol\theta}\):

\begin{itemize}
\item
  To preserve a constant trend \(X_t=a\) we need: \[
  \forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
  \] The sum of the coefficients of the moving average
  \(\sum_{k=-p}^{+f}\theta_k\) must then be equal to \(1\).
\item
  To preserve linear trends \(X_t=at+b\) we need: \[
  \forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
  \] Which is equivalent to: \[
  \sum_{k=-p}^{+f}\theta_k=1
  \quad\text{et}\quad
  \sum_{k=-p}^{+f}k\theta_k=0.
  \]
\item
  In general, \(M_{\boldsymbol\theta}\) preserves polynomials of degree
  \(d\) if and only if: \[
  \sum_{k=-p}^{+f}\theta_k=1
   \text{ and }
  \forall j \in \left\llbracket 1,d\right\rrbracket:\:
  \sum_{k=-p}^{+f}k^j\theta_k=0.
  \]
\item
  If \(M_{\boldsymbol\theta}\) is symmetric (\(p=f\) and
  \(\theta_{-k} = \theta_k\)) and preserves polynomials of degree \(2d\)
  then it also preserves polynomials of degree \(2d+1\).
\end{itemize}

\hypertarget{noise-reduction}{%
\subsubsection{Noise reduction}\label{noise-reduction}}

All time series are affected by noise, which can blur the extraction of
trends and cycles. This is why we try to reduce this noise (by reducing
its variance) while preserving the relevant trends. The sum of the
squares of the coefficients \(\sum_{k=-p}^{+f}\theta_k^2\) is the
\emph{reduction of variance} ratio.

Indeed, let \(\{\varepsilon_t\}\) a sequence of independent random
variables with \(\E{\varepsilon_t}=0\), \(\V{\varepsilon_t}=\sigma^2\):
\[
\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
\]

\hypertarget{sec-mmasym}{%
\subsection{Real-time estimation and asymmetric moving
average}\label{sec-mmasym}}

\hypertarget{sec-mmetprev}{%
\subsubsection{Asymmetric moving averages and
forecasts}\label{sec-mmetprev}}

For symmetric filters, the phase shift function is equal to zero (modulo
\(\pi\)). So there is no delay in detecting turning points. However,
they cannot be used in the beginning and in the end of the time series
because no past or future value can be used. Thus, for real-time
estimation, it is needed to build asymmetric moving average that
approximate the symmetric moving average.

Several approaches can be used for real-time estimation:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Use asymmetric moving averages to take account of the lack of
  available data;
\item
  Apply symmetrical filters to extended forecast series. This method
  seems to date back to \textcite{deforest1877adjustment}, which also
  suggests modelling a polynomial trend of degree three or less at the
  end of the period:\\
  ``\emph{As the first \(m\) and last \(m\) terms of the series cannot
  be reached directly by the formula, the series should be graphically
  extended by m terms at both ends, first plotting the observations on
  paper as ordinates, and then extending the curve along what seems to
  be its probable course, and measuring the ordinates of the extended
  portions.} \emph{It is not necessary that this extension should
  coincide with what would be the true course of the curve in those
  parts. } \emph{The important point is that the m terms thus added,
  taken together with the \(m+1\) adjacent given terms, should follow a
  curve whose form is approximately algebraic and of a degree not higher
  than the third.}''\\
  This is also the approach used in the X-13ARIMA seasonal adjustment
  method, which extends which extends the series by one year with an
  ARIMA model.
\end{enumerate}

Ultimately, the second method is equivalent to using asymmetric moving
averages, since forecasts are linear combinations of the past.

Conversely, the implicit forecasts of an asymmetric moving average can
be deduced from a reference symmetric moving average. This allows to
judge the quality of real-time estimates of the trend-cycle and to
anticipate future revisions when forecasts are far from expected
evolutions.\\
Let us denote \(v=(v_{-h},\dots, v_{h})\) the reference symmetrical
moving average and \(w^0,\dots w^{h-1}\) a sequence of asymmetrical
moving averages, of order \(h+1\) to \(2h\) used to estimate the last
\(h\) points with, for convention, \(w_t^q=0\) for \(t>q\). This means
that \(w^0=(w_{-h}^0,\dots, w_{0}^0)\) is used for estimation in real
time (when no points are known in the future),
\(w^1=(w_{-h}^1,\dots, w_{1}^1)\) for estimation of the penultimate
point (when only one point is known in the future), and so on. Let also
note \(y_{-h},\dots,y_{0}\) the observed series studied and
\(y_{1}^*,\dots y_h^*\) the implicit forecast induced by
\(w^0,\dots w^{h-1}\). This means that for all \(q\) we have: \[
\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
\] This is equivalent to: \[
\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
\] To sum up, in matrix form, this is equivalent to solving: \[
\scriptstyle
\begin{pmatrix}
  v_1 & v_2 & \cdots & v_h \\
  v_1 - w_1^1 & v_2 & \cdots & v_h \\
  \vdots & \vdots & \cdots & \vdots \\
   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
\end{pmatrix}
\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
\begin{pmatrix}
  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
  \vdots & \vdots & \cdots & \vdots \\
  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
\end{pmatrix}
\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
\] This is implemented in the \texttt{rjd3filters::implicit\_forecast()}
function.

As highlighted by \textcite{wildischis2004}, extending the series
through forecasting with an ARIMA model is equivalent to calculating
asymmetric filters whose coefficients are optimised in relation to the
one-step ahead forecast. In other words, the aim is to minimise the
revisions between the first and last estimates (with the symmetric
filter). However, the phase shift induced by the asymmetric filters is
not controlled: we might prefer to have faster detection of turning
points and a larger revision rather than just minimising the revisions
between the first and last estimates. Furthermore, since the
coefficients of the symmetric filter (and therefore the weight
associated with distant forecasts) decrease slowly, we should also be
interested in the performance of multi-step ahead forecasting. This is
why it may be necessary to define alternative criteria for judging the
quality of asymmetric moving averages.

\hypertarget{subsec:crit-qual}{%
\subsubsection{Quality indicators for asymmetric moving
averages}\label{subsec:crit-qual}}

For asymmetric filters, most of the criteria come from those defined by
\textcite{ch15HBSA} and \textcite{trilemmaWMR2019} to build asymmetric
filters. They are summarised in Table~\ref{tab-QC} and can be calculated
using the \texttt{rjd3filters::diagnostic\_matrix()} function.

\textcite{ch15HBSA} propose a general approach to derive linear filters,
based on an optimisation problem involving three criteria:
\emph{Fidelity} (\(F_g\)), \emph{Smoothness} (\(S_g\)) and
\emph{Timeliness} (\(T_g\)). Fidelity can be directly linked to the
reduction in variance created by the filter, and timeliness to the
notion of phase shift, which again we want to be low. This method can be
used in\faIcon{r-project} with the function
\texttt{rjd3filters::fst\_filter()}.

\textcite{trilemmaWMR2019} propose an approach based on the
decomposition of the mean squared error between the symmetric and the
asymmetric filter in four quantities: \emph{Accuracy} (\(A_w\)),
\emph{Timeliness} (\(T_w\)), \emph{Smoothness} (\(S_w\)) and
\emph{Residual} (\(R_w\)). A simplified version of this method can be
used in \faIcon{r-project} with the function
\texttt{rjd3filters::dfa\_filter()}.

\begin{table}[!h]
\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
{\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ccc}
\toprule
Code & Description & Formula\\
\midrule
$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
\addlinespace
$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
\bottomrule
\end{tabular}}}

\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}

\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}

\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
\end{table}

\hypertarget{sec-nonparamreg}{%
\section{Non-parametric regression and local polynomial
regression}\label{sec-nonparamreg}}

Many trend-cycle extraction methods are based on non-parametric
regressions, which are particularly flexible because they do not assume
any predetermined dependency in the predictors. In practice, local
regressions can be used. More specifically, consider a set of points
\((x_i,y_i)_{1\leq i\leq n}\). Non-parametric regression involves
assuming that there exists a function \(\mu\), to be estimated, such
that \(y_i=\mu(x_i)+\varepsilon_i\) with \(\varepsilon_i\) an error
term. According to Taylor's theorem, for any point \(x_0\), if \(\mu\)
is differentiable \(d\) times, then:
\begin{equation}\protect\hypertarget{eq-taylor}{}{
\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
}\label{eq-taylor}\end{equation} where \(R_d\) is a negligible residual
term in the neighbourhood of \(x_0\). In a neighbourhood \(h(x_0)\)
around \(x_0\), \(\mu\) can be approximated by a polynomial of degree
\(d\). The quantity \(h(x_0)\) is called the \emph{bandwidth}. If
\(\varepsilon_i\) is white noise, then we can estimate by least squares
\(\mu(x_0)\) using the observations which are in
\(\left[x_0-h(x_0),x_0+h(x_0)\right]\).

In practice, this means assuming that the trend is locally polynomial.
Various estimation methods can be used to derive symmetrical and
asymmetrical moving averages.\\
\textcite{GrayThomson1996} propose a complete statistical framework
which makes it possible, in particular, to model the error in
approximating the trend by local polynomials. However, as the
specification of this error is generally complex, simpler models may be
preferred, such as that of \textcite{proietti2008}.\\
Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
the trend-cycle but using the theory of Hilbert spaces with reproducing
kernels for estimation, which has the particular advantage of
facilitating the calculation of different moving averages at different
time frequencies. This method can be used in\faIcon{r-project} with the
function \texttt{rjd3filters::rkhs\_filter()}.

\hypertarget{symmetric-moving-averages-and-local-polynomial-regression}{%
\subsection{Symmetric moving averages and local polynomial
regression}\label{symmetric-moving-averages-and-local-polynomial-regression}}

Using \textcite{proietti2008}'s notation, we assume that our time series
\(y_t\) can be decomposed into \[
y_t=\mu_t+\varepsilon_t,
\] where \(\mu_t\) is the trend and
\(\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})\) is the
noise (the series is therefore seasonally adjusted). Within a
neighbourhood \(h\) of \(t\), the local trend \(\mu_t\) is approximated
by a polynomial of degree \(d\), such that \(\mu_t\simeq m_{t}\) with:
\[
\forall j\in\left\llbracket -h,h\right\rrbracket:\:
y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
\] The problem of trend extraction is equivalent to estimating
\(m_t=\beta_0\) (the constant in the previous formula).

In matrix notation: \[
\underbrace{\begin{pmatrix}y_{t-h}\\
y_{t-(h-1)}\\
\vdots\\
y_{t}\\
\vdots\\
y_{t+(h-1)}\\
y_{t+h}
\end{pmatrix}}_{\boldsymbol y}=\underbrace{\begin{pmatrix}1 & -h & h^{2} & \cdots & (-h)^{d}\\
1 & -(h-1) & (h-1)^{2} & \cdots & (-(h-1))^{d}\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & 0 & 0 & \cdots & 0\\
\vdots & \vdots & \vdots & \cdots & \vdots\\
1 & h-1 & (h-1)^{2} & \cdots & (h-1)^{d}\\
1 & h & h^{2} & \cdots & h^{d}
\end{pmatrix}}_{\boldsymbol X}\underbrace{\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\vdots\\
\vdots\\
\vdots\\
\vdots\\
\beta_{d}
\end{pmatrix}}_{\boldsymbol \beta}+\underbrace{\begin{pmatrix}\varepsilon_{t-h}\\
\varepsilon_{t-(h-1)}\\
\vdots\\
\varepsilon_{t}\\
\vdots\\
\varepsilon_{t+(h-1)}\\
\varepsilon_{t+h}
\end{pmatrix}}_{\boldsymbol \varepsilon}.
\]

To estimate \(\boldsymbol \beta\) we need \(2h+1\geq d+1\) and the
estimation is made by weighted least squares (WLS), which is equivalent
to minimising the following objective function: \[
S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
\] where \(\kappa_j\) is a set of weights called \emph{kernel}. We have
\(\kappa_j\geq 0:\kappa_{-j}=\kappa_j\), and with
\(\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})\), the estimate of
\(\boldsymbol \beta\) can be written as
\(\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.\)
With \(\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}\),
the estimate of the trend is:
\begin{equation}\protect\hypertarget{eq-mmsym}{}{
\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
}\label{eq-mmsym}\end{equation} To conclude, the estimate of the trend
\(\hat{m}_{t}\) can be obtained applying the symmetric filter
\(\boldsymbol \theta\) to \(y_t\) (\(\boldsymbol \theta\) is symmetric
due to the symmetry of the kernel weights \(\kappa_j\)). Moreover,
\(\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}\) so: \[
\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
\] Hence, the filter \(\boldsymbol w\) preserve deterministic polynomial
of order \(d\).

Regarding parameter selection, the general consensus is that the choice
between different kernels is not crucial. See, for example,
\textcite{cleveland1996smoothing} or \textcite{Loader1999}. The only
desired constraints on the kernel are that it assigns greater weight to
the central estimation (\(\kappa_0\)) and decreases towards 0 as it
moves away from the central estimation. The uniform kernel should
therefore be avoided. It is then preferable to focus on two other
parameters:

\begin{itemize}
\item
  the degree of the polynomial, denoted by \(d\): if it is too small,
  there is a risk of biased estimates of the trend-cycle, and if it is
  too large, there is a risk of excessive variance in the estimates (due
  to over-adjustment);
\item
  the number of neighbors \(2h+1\) (or the window \(h\)): if it is too
  small, then there will be too little data for the estimates (resulting
  in high variance in the estimates), and if it is too large, the
  polynomial approximation will probably be wrong, leading to biased
  estimates.
\end{itemize}

In this paper we will use the Henderson kernel: \[
\kappa_{j}=\left[1-\frac{j^2}{(h+1)^2}\right]
\left[1-\frac{j^2}{(h+2)^2}\right]
\left[1-\frac{j^2}{(h+3)^2}\right].
\] However, several other kernels are available in \texttt{rjd3filters}.

\hypertarget{sec-lppasymf}{%
\subsection{Asymmetric moving averages and local polynomial
regression}\label{sec-lppasymf}}

As mentioned in Section~\ref{sec-mmetprev}, for real-time estimation,
several approaches can be used:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Apply symmetric filters to the series extended by forecasting
  \(\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket\).
\item
  Build an asymmetric filter by local polynomial approximation on the
  available observations (\(y_{t}\) for
  \(t\in\left\llbracket n-h,n\right\rrbracket\)).
\item
  Build asymmetric filters that minimise the mean squared error of
  revision under polynomial trend reproduction constraints.
\end{enumerate}

\textcite{proietti2008} show that the first two approaches are
equivalent when forecasts are made by polynomial extrapolation of degree
\(d\). They are also equivalent to the third approach under the same
constraints as those of the symmetric filter. The third method is called
\emph{direct asymmetric filter} (DAF). This method is used for real-time
estimation in the STL seasonal adjustment method (\emph{Seasonal-Trend
decomposition based on Loess}, see \textcite{cleveland90}). Although DAF
filter estimates are unbiased, this is at the cost of greater variance
in the estimates.

To solve the problem of the variance of real-time filter estimates,
\textcite{proietti2008} propose a general method for constructing
asymmetric filters that allows a bias-variance trade-off. This is a
generalisation of \textcite{musgrave1964set}'s asymmetric filters (used
in the X-13ARIMA seasonal adjustment algorithm).

The initial series is modelled here as:
\begin{equation}\protect\hypertarget{eq-lpgeneralmodel}{}{
\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
}\label{eq-lpgeneralmodel}\end{equation} where
\([\boldsymbol U,\boldsymbol Z]\) is of full rank and forms a subset of
the columns of \(X\). The objective is to find a filter
\(\boldsymbol v\) that minimises the mean square error of revision (to
the symmetric filter \(\boldsymbol \theta\)) under certain constraints.
These constraints are represented by the matrix
\(\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}\):
\(\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta\) (with
\(\boldsymbol U_p\) the matrix \((h+q+1)\times (d+1)\) which contains
the observations of the matrix \(U\) known during estimation by the
asymmetric filter). The problem is equivalent to finding \(v\) which
minimises: \begin{equation}\protect\hypertarget{eq-lppasym}{}{
\varphi(\boldsymbol v)=
\underbrace{
  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
}_\text{mean square revision error}+
\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
}\label{eq-lppasym}\end{equation} where \(\boldsymbol l\) is a vector of
Lagrange multipliers.

When \(\boldsymbol U=\boldsymbol X\), the constraint is equivalent to
preserving polynomials of degree \(d\): we find asymmetric direct
filters (DAF) when \(\boldsymbol D=\boldsymbol K^{-1}\).

When \(\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}\),
\(\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}\),
\(\boldsymbol \delta=\delta_1\), \(\boldsymbol D=\sigma^2\boldsymbol I\)
and when the symmetric filter is the Henderson filter, we find the
asymmetric Musgrave filters (using the Henderson kernel). This filter
assumes that, for real-time estimation, the data is generated by a
linear process and that the asymmetric filters preserve the constants
(\(\sum v_i=\sum \theta_i=1\)). These asymmetric filters depend on the
ratio \(\lvert\delta_1/\sigma\rvert\), which, assuming that the trend is
linear and the bias constant, can be linked to the I-C ratio
\(R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}\):
\(\delta_1/\sigma=2/(R\sqrt{\pi})\). This ratio is used in particular in
X-13ARIMA to determine the length of the Henderson filter. For monthly
data:

\begin{itemize}
\item
  If the ratio is large (\(3.5< R\)) then a 23 term filter is used (to
  remove more noise).
\item
  If the ratio is small (\(R<1\)) a 9 term filter is used.
\item
  Otherwise (most of the cases) a 13-term filter is used.
\end{itemize}

When \(\boldsymbol U\) corresponds to the first \(d^*+1\) columns of
\(\boldsymbol X\), \(d^*<d\), the constraint is to reproduce polynomial
trends of degree \(d^*\). This introduces bias but reduces the variance.
Thus, \textcite{proietti2008} proposes three classes of asymmetric
filters:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \emph{Linear-Constant} (LC): \(y_t\) linear (\(d=1\)) and
  \(\boldsymbol v\) preserves constant signals (\(d^*=0\)). We obtain
  Musgrave filters when the Henderson kernel is used and the Henderson
  filter is used as the symmetric filter.
\item
  \emph{Quadratic-Linear} (QL): \(y_t\) quadratic (\(d=2\)) and
  \(\boldsymbol v\) preserves linear signals (\(d^*=1\)).
\item
  \emph{Cubic-Quadratic} (CQ): \(y_t\) cubic (\(d=3\)) and
  \(\boldsymbol v\) preserves quadratic signals (\(d^*=2\)).
\end{enumerate}

Appendix \ref{sec-an-cof} shows the coefficients, gain and phase shift
functions of the four asymmetric filters.

Table~\ref{tab-criteriaLp} compares the quality criteria of the
different methods using the Henderson filter and \(h=6\) (a symmetric
filter with 13 terms). For real-time filters (\(q=0\)), a more complex
asymmetrical filter (in terms of polynomial preservation) results in
lower timeliness and higher fidelity/smoothness: reduction in phase
shift comes at the expense of an increase in variance. This result
varies as \(q\) increases: for \(q=2\), the QL filter has greater
timeliness than the LC filter. This surprising result highlights the
fact that the phase shift is not controlled by \textcite{proietti2008}'s
approach.

In terms of revision (\(A_w+S_w+T_w+R_w\)), the LC and QL filters
consistently outperform the CQ and DAF filters.

These theoretical properties are consistent with the empirical results
observed in Section~\ref{sec-comparison}. Revisions are more important
for the CQ and DAF filters, which leads to greater variability in
estimations and a longer delay in detecting turning points. The globally
parameterised QL filter (with \(R\) fixed for all asymmetric filters)
may result in later detection of turning points than the LC filter.
Indeed, to identify a turning point at date \(t\), it is necessary to
have knowledge of at least two points after the given date to confirm
the reversal in trend.

\begin{table}[!h]
\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
\centering
\begin{tabular}[t]{cccccccccccc}
\toprule
Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
\midrule
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=0$}}\\
\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=1$}}\\
\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
\addlinespace[0.3em]
\multicolumn{12}{l}{\textbf{$q=2$}}\\
\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
\bottomrule
\end{tabular}

\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
\end{table}

\hypertarget{sec-lptimeliness}{%
\subsection{Extension with the timeliness
criterion}\label{sec-lptimeliness}}

One drawback of the previous method is the lack of control over the
phase shift. However, it is possible to improve the modelling by
incorporating in Equation~\ref{eq-lppasym} the timeliness criterion
defined by \textcite{ch15HBSA}. This was proposed by Jean Palate, then
coded in Java and integrated into \texttt{rjd3filters}.

Using the same notation as in Section~\ref{sec-lppasymf},
\(\boldsymbol\theta\) represents the symmetric filter and
\(\boldsymbol v\) the asymmetric filter. With
\(\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}\)
and \(\boldsymbol\theta_p\) of the same length of \(\boldsymbol v\), and
\(\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p\), the timeliness
criterion can be expressed as: \[
T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
\] Furthermore, the objective function \(\varphi\) of
Equation~\ref{eq-lppasym} can be rewritten: \begin{align*}
\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
\begin{cases}
\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
\boldsymbol c\text{ a constant independent of }\boldsymbol v
\end{cases}.
\end{align*} Adding the timeliness criterion: \[
\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
\widetilde {\boldsymbol c}\quad\text{with }
\begin{cases}
\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
\end{cases},
\] where \(\alpha_T\) is the weight associated to the timeliness
criterion. With \(\alpha_T=0\) we find \(\varphi(\boldsymbol v)\). This
extension therefore makes it possible to find all the symmetric and
asymmetric filters presented in the previous section but also
generalizes the approach of \textcite{GrayThomson1996}. This extension
is implemented in the \faIcon{r-project} function
\texttt{rjd3filters::lp\_filter()}.

One drawback is that \(T_g(\boldsymbol v)\) is not normalized: the
weight \(\alpha_T\) has no economic sense. It is then difficult to
calibrate the value of \(\alpha_T\) since the value of
\(\lvert\delta/\sigma\rvert\) must also be defined. That's why in this
paper we will not compare this extension to the other methods and we
will focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
However, in future studies, we could imagine a two step calibration: fix
the value of \(\lvert\delta/\sigma\rvert\) to find \(\alpha_T\) by
cross-validation and then use this weight with a local parameterisation
of the ratio \(\lvert\delta/\sigma\rvert\) (as in
Section~\ref{sec-localic}).

\hypertarget{sec-localic}{%
\subsection{Local parameterisation of asymmetric
filters}\label{sec-localic}}

Asymmetric filters are usually parameterised globally:
\(\lvert\delta/\sigma\rvert\) estimated on all the data using the
IC-ratio or a cross-validation criterion. However, we might prefer a
local parameterisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
as a function of time. Indeed, although the overall parameterisation is
generally valid, assuming a constant value of the
\(\lvert\delta/\sigma\rvert\) ratio for all asymmetrical filters does
not appear relevant for real-time estimation, especially during periods
of economic downturn. For example, with the LC method, global
parameterisation means assuming that the slope of the trend is constant,
whereas during economic downturns it tends towards 0 up to the turning
point.

This is what is proposed in this article, with a local parameterisation
of the asymmetric filters by estimating \(\delta\) and \(\sigma^2\)
separately. Although this does not give an unbiased estimator of the
ratio \(\lvert\delta/\sigma\rvert\), it does allow the main evolutions
to be captured, such as the decay towards 0 before a turning point and
the growth after the turning point for the LC method:

\begin{itemize}
\tightlist
\item
  The variance \(\sigma^2\) can be estimated using the observed data set
  and the symmetrical filter \((w_{-p},\dots,w_p)\): \[
  \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
  \]
\item
  The parameter \(\delta\) can be estimated by moving average from
  Equation~\ref{eq-mmsym}. For example, for the LC method we can use the
  moving average
  \(\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}\)
  to obtain a local estimate of the slope and for the QL method we can
  use
  \(\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}\)
  to obtain a local estimate of the concavity. The DAF method then
  simply allows us to calculate the associated asymmetric moving
  averages.\\
  Although a moving average of different length to that used for
  estimating the trend could be considered, this seems to degrade the
  results in terms of phase shift (using the same methodology as in
  Section~\ref{sec-comparison}). Furthermore, for the construction of
  moving averages, the trend can be modelled as being locally of degree
  2 or 3 (this has no impact on the final estimate of concavity). Here,
  we have chosen to model a trend of degree 2: this reduces the phase
  shift but slightly increases the revisions linked to the first
  estimate of the trend-cycle. Figure~\ref{fig-mmpenteconcac} shows the
  moving averages used.
\end{itemize}

\begin{figure}[H]

\caption{\label{fig-mmpenteconcac}Moving averages used for real-time
estimation of slope and concavity.}

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite.pdf}

}

\end{figure}

In the empirical applications of Section~\ref{sec-comparison}, the final
local parameterisation is obtained by estimating \(\delta\) using all
the data (i.e., using the symmetric filters shown in
Figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
of \(\sigma^2\). Figure~\ref{fig-mmpenteconcac-ex} shows an example of
the comparison of estimates of the ratio \(|\delta_1/\sigma|\) with a
global estimator (IC ratio) and two the local parameterisation:

\begin{itemize}
\item
  Figure~\ref{fig-mmpenteconcac-ex-1}: using a symmetric filter to
  estimate \(\delta_1\) (final local parameterisation).
\item
  Figure~\ref{fig-mmpenteconcac-ex-2}: using the asymmetric filter which
  need two points in the futur to estimate \(\delta_1\). This is closed
  to the real-time estimates since two points in the future are needed
  to detect a turning point.
\end{itemize}

The turning points are clearly detected by the local estimators, but
there is much more variance in the estimate of \(|\delta_1/\sigma|\) in
the real-time estimate. This will lead to another source of revisions in
the intermediate estimates.

\begin{figure}

\caption{\label{fig-mmpenteconcac-ex}Comparison of the estimates of
\(|\delta_1/\sigma|\) with a local estimator and global estimator (with
IC ratio) on a simulated series, the vertical lines being the simulated
turning points.}\begin{minipage}[t]{\linewidth}
\subcaption{\label{fig-mmpenteconcac-ex-1}Symmetric moving average used
for \(\hat \delta_1\).}

{\centering

\raisebox{-\height}{

\includegraphics{img/filters_used/mm_penteconcavite_ex.pdf}

}

}

\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}
\subcaption{\label{fig-mmpenteconcac-ex-2}Asymmetric moving average
using two futur points (\(q=2\)) used for \(\hat \delta_1\).}

{\centering

\raisebox{-\height}{

\includegraphics{img/filters_used/mm_penteconcavite_ex2.pdf}

}

}

\end{minipage}%

\end{figure}

\hypertarget{sec-comparison}{%
\section{Comparison of different methods}\label{sec-comparison}}

The different methods are compared on simulated and real data.

For all series, a symmetrical 13-term filter is used. These methods are
also compared with estimates obtained by extending the series with an
ARIMA model, then applying a 13-term symmetrical Henderson filter. The
ARIMA model is determined automatically, using no seasonal lag (the
series being seasonally adjusted) and no external variables (such as
external regressors for correction of atypical points).

The performance of the different methods is judged on criteria relating
to revisions (between two consecutive estimates and in relation to the
final estimate) and the number of periods required to detect turning
points.

\hypertarget{simulated-series}{%
\subsection{Simulated series}\label{simulated-series}}

\hypertarget{methodology}{%
\subsubsection{Methodology}\label{methodology}}

Following a methodology close to that of \textcite{DarneDagum2009}, nine
monthly series are simulated between January 1960 and December 2020 with
different levels of variability. Each simulated series
\(y_t= C_t+ T_t + I_t\) can be written as the sum of three components:

\begin{itemize}
\item
  the cycle
  \(C_t = \rho [\cos (2 \pi t / \lambda) +\sin (2 \pi t / \lambda)]\),
  \(\lambda\) is fixed at 72 (6-year cycles, so there are 19 detectable
  turning points);
\item
  the trend \(T_t = T_{t-1} + \nu_t\) with
  \(\nu_t \sim \mathcal{N}(0, \sigma_\nu^2)\), \(\sigma_\nu\) being
  fixed at \(0.08\);
\item
  and the irregular \(I_t = e_t\) with
  \(e_t \sim \mathcal{N}(0, \sigma_e^2)\).
\end{itemize}

For the different simulations, we vary the parameters \(\rho\) and
\(\sigma_e^2\) in order to obtain series with different signal-to-noise
ratios (see Figure~\ref{fig-graphs-data-simul}):

\begin{itemize}
\item
  High signal-to-noise ratio (i.e.~low I-C ratio and low variability):
  \(\sigma_e^2=0.2\) and \(\rho = 3.0, 3.5\) or \(4.0\) (I-C ratio
  between 0.9 and 0.7);
\item
  Medium signal-to-noise ratio (i.e.~medium I-C ratio and medium
  variability): \(\sigma_e^2=0.3\) and \(\rho = 1.5,\, 2.0\) or \(3.0\)
  (I-C ratio between 2.3 and 1.4);
\item
  Low signal-to-noise ratio (i.e.~high I-C ratio and high variability):
  \(\sigma_e^2=0.4\) and \(\rho = 0.5,\, 0.7\) or \(1.0\) (I-C ratio
  between 8.9 and 5.2).
\end{itemize}

\begin{figure}[H]

\caption{\label{fig-graphs-data-simul}Simulated series with low
(\(\sigma_e^2=0.2\) and \(\rho = 3,5\)), medium (\(\sigma_e^2=0.3\) and
\(\rho = 2.0\)) and high variability (\(\sigma_e^2=0.4\) and
\(\rho = 1.0\)).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/simul_data.pdf}

}

\end{figure}

For each series and each date, the trend-cycle is estimated using the
different methods presented in this report. Three quality criteria are
also calculated:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Computation of the phase shift in the detection of turning points. The
  definition of \textcite{Zellner1991} is used to determine the turning
  points:

  \begin{itemize}
  \tightlist
  \item
    An upturn occurs when the economy moves from a phase of recession to
    a phase of expansion. This is the case at date \(t\) when
    \(y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}\).\\
  \item
    A downturn occurs when the economy moves from a phase of expansion
    to a phase of recession. This is the case at date \(t\) when
    \(y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}\).\\
    It thus takes at least 2 months to detect a turning point.\\
    The phase shift is often defined as the number of months required to
    detect the right turning point. Here we use a slightly modified
    criterion: the phase shift is defined as the number of months needed
    to detect the right turning point without any future revision.
    Indeed, it can happen that the right turning point is detected by
    asymmetric filters but is not detected by the final estimate using a
    symmetric filter (this is the case for 41 reversal points out of the
    9 series with asymmetric Musgrave filters) or that there are
    revisions in successive estimates (this is the case for 7 turning
    points out of the 9 series with asymmetric Musgrave filters).
    Finally, relatively few turning points are detected at the right
    date with the final estimate. With the 13-term Henderson filter, 18
    are correctly detected in series with low variability (out of 57
    possible), 11 in series with medium variability and 12 in series
    with high variability.
  \end{itemize}
\item
  Computation of two revision criteria: the average of the relative
  deviations between the \(q\)\textsuperscript{th} estimate and the last
  estimate \(MAE_{fe}(q)\) and the average of the relative deviations
  between the \(q\)\textsuperscript{th} estimate and the
  \(q+1\)\textsuperscript{th} estimate \(MAE_{qe}(q)\). \[
  MAE_{fe}(q)=\mathbb E\left[
  \left|\frac{
  y_{t|t+q} -  y_{t|last}
  }{
   y_{t|last}
  }\right|
  \right]
  \quad\text{and}\quad
  MAE_{qe}(q)=\mathbb E\left[
  \left|\frac{
  y_{t|t+q} - y_{t|t+q+1}
  }{
  y_{t|t+q+1}
  }\right|
  \right].
  \]
\end{enumerate}

\hypertarget{comparison}{%
\subsection{Comparison}\label{comparison}}

Excluding for the moment the local parameterisations of the polynomial
filters, the linear-constant polynomial filter (LC) seems to give the
best results in terms of delay in the detection of turning points
(Figure~\ref{fig-graphstpsimul}). Performance is relatively close to
that obtained by extending the series using an ARIMA model. However,
when variability is low, the LC filter seems to give poorer results and
the quadratic-linear polynomial filter (QL) seems to give the best
results.

For series with moderate variability, local parameterisation of the LC
and QL filters reduces the phase shift. For series with high
variability, the phase shift is only reduced by using the final
\(\hat\delta\) parameters: real-time estimates seem to add more
variance. For series with low variability, performance seems to be
slightly improved only with the LC filter.

\begin{figure}[H]

\caption{\label{fig-graphstpsimul}Distribution of phase shifts on
simulated series.}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}

}

\end{figure}

In terms of revisions, the variability of the series has little effect
on the respective performances of the different methods, but it does
affect the orders of magnitude, which is why the results are presented
only for series with medium variability (Table~\ref{tab-simulrev}). In
general, LC filters always minimise revisions (with relatively small
effect from the local parameterisation of the filters) and revisions are
greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.

For the QL filter, there is a large revision between the second and
third estimates: this may be due to the fact that for the second
estimate (when one point in the future is known), the QL filter assigns
a greater weight to the estimate in \(t+1\) than to the estimate in
\(t\), which creates a discontinuity. This revision is considerably
reduced by parameterising the filter locally. For polynomial filters
other than the LC filter, the large revisions at the first estimate were
to be expected given the coefficient curve: a very large weight is
associated with the current observation and there is a strong
discontinuity between the moving average used for real-time estimation
(when no point in the future is known) and the other moving averages.

Extending the series using an ARIMA model gives revisions with the
latest estimates of the same order of magnitude as the LC filter, but
slightly larger revisions between consecutive estimates, particularly
between the fourth and fifth estimates (as might be expected, as
highlighted in Section~\ref{sec-mmetprev}).

\begin{table}[!h]
\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
filters on simulated series with medium variability.}
\centering
\begin{tabular}{ccccccc}
\toprule
Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
\midrule
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
\addlinespace[0.3em]
\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
\right]$}}\\
\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{real-series}{%
\subsection{Real series}\label{real-series}}

The differences between the methods are also illustrated using an
example from the FRED-MD database (\textcite{fredmd}) containing
economic series on the United States. The series studied correspond to
the database published in November 2022. The series studied is the level
of employment in the United States (series \texttt{CE160V}, used in
logarithm) around the February 2001 turning point, consistent with the
monthly dating of the turning points. This point and this series were
chosen because the economic turnaround is particularly visible in the
raw series (Figure~\ref{fig-ce16ov-previmp-lp}), because this series is
used for dating economic cycles and because the FRED-MD database
facilitates the reproducibility of the results (thanks to the
availability of series published on past dates). Studying the series up
to January 2020, this series has a medium variability: a symmetrical
filter of 13 terms is therefore appropriate. Appendix
\autocite*{ann-ex-r} shows how the code to reproduce the different
plots.

Figure~\ref{fig-ce16ovlp} shows the successive estimates of the
trend-cycle using the different methods studied.

For this series, the phase shift is 6 months for the LC and CQ methods
and the extension of the series by ARIMA. Except for the LC method
(where the turning point is first detected in January 2001), the high
phase shift is due to revisions in the turning point detected in the
intermediate estimates (even if the correct turning point is detected
after 2 months, it is no longer detected after 3 months). It is of two
months for the other methods (QL, DAF).

In this example, local parameterisation does not reduce the phase shift,
but it does reduce revisions. The QL and DAF polynomials lead to greater
variability in the intermediate estimates, especially in February 2001.

\begin{figure}[H]

\caption{\label{fig-ce16ovlp}Successive estimates of the trend-cycle in
US employment (in logarithms).}

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}

}

\end{figure}

The quality of the intermediate estimates can also be analysed using the
implicit forecasts of the different methods
(Figure~\ref{fig-ce16ov-previmp-lp}). As a reminder, these are forecasts
of the raw series which, by applying the symmetrical Henderson filter to
the extended series, give the same estimates as the asymmetrical moving
averages. The forecasts of the ARIMA model are naive and do not take the
turning point into account, unlike the other methods. Finally, local
parameterisation of the QL filter produces much more consistent
forecasts.

\begin{figure}[H]

\caption{\label{fig-ce16ov-previmp-lp}Implicit forecasts linked to
successive estimates of the trend-cycle in US employment (in logarithms)
using local polynomial methods.}

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}

}

\end{figure}

\hypertarget{conclusion}{%
\section*{Conclusion}\label{conclusion}}
\addcontentsline{toc}{section}{Conclusion}

For business cycle analysis, most statisticians use trend-cycle
extraction methods, either directly or indirectly. They are used, for
example, to reduce the noise of an indicator in order to improve its
analysis, and models, such as forecasting models, usually use seasonally
adjusted series based on these methods.

This paper presents the \faIcon{r-project} package \texttt{rjd3filters},
which implements a several methods to build moving average for real-time
trend-cycle estimation. It also offers various functions for the
theoretical analysis of moving averages (gain functions, phase shift,
quality criteria, etc.) and judging the quality of the latest estimates
(for example with implicit forecasts). It also makes it easy to combine
moving averages and also to integrate custom moving averages into the
X-11 seasonal adjustment algorithm (with the
\texttt{rjd3filters::x11plus\_trend()} function). \texttt{rjd3filters}
thus facilitates research into the use of moving averages as part of
real-time cycle trend estimation. This is the case, for example, with
the local parametrisation of Musgrave moving averages and other
asymmetric moving averages linked to local polynomial regression,
presented in this article.

By comparing the different methods, we can learn a few lessons about the
construction of these moving averages.

During economic downturns, asymmetric filters used as an alternative to
extending the series using the ARIMA model can reduce revisions to
intermediate estimates of the trend-cycle and enable turning points to
be detected more quickly.

At the end of the period, modelling polynomial trends of degree greater
than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce
variance into the estimates (and therefore more revisions) without
allowing faster detection of turning points. At the end of the period,
for estimating the trend-cycle, we can therefore restrict ourselves to
methods modelling polynomial trends of degree two or less
(linear-constant, LC, and quadratic-linear, QL). In addition,
parameterising polynomial filters locally as suggested in this paper
enables turning points to be detected more quickly (especially for the
QL filter). Even when the phase shift is not reduced, local
parameterisation is recommended because it reduces revisions and
produces intermediate estimates that are more consistent with expected
future trends. However, with these methods, the length of the filter
used must be adapted to the variability of the series: if the filter
used is too short (i.e.~if the variability of the series is ``low''),
retaining polynomial trends of degree one or less (LC method) produces
poorer results in terms of detecting turning points.

This study could be extended in many ways.

One possible extension would be to look at the impact of filter length
on the detection of turning points. Asymmetric filters are calibrated
using indicators calculated for the estimation of symmetric filters (for
example, to automatically determine their length), whereas a local
estimate might be preferable. Furthermore, we have only focused on
monthly series with a 13-term symmetric filter, but the results may be
different if the symmetric filter studied is longer or shorter and if we
study series with other frequencies (quarterly or daily, for example).

Another possibility would be to study the impact of atypical points:
moving averages, like any linear operator, are very sensitive to the
presence of atypical points. To limit their impact, in X-13ARIMA a
strong correction for atypical points is performed on the irregular
component before applying the filters to extract the trend-cycle. This
leads us to study the impact of these points on the estimation of the
trend-cycle and turning points, and also to explore new types of
asymmetric filters based on robust methods (such as robust local
regressions or moving medians).

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\printbibliography[heading=none]

\appendix
\newpage

\hypertarget{sec-an-cof}{%
\section{Coefficients, gain and phase shift
functions}\label{sec-an-cof}}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-lc}Coefficients, gain and phase shift
functions for the Linear-Constant (LC) filter with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}

}

\end{figure}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-ql}Coefficients, gain and phase shift
functions for the Quadratic-Linear (QL) filter with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}

}

\end{figure}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-cq}Coefficients, gain and phase shift
functions for the Cubic-Quadratic (CQ) filter with \(I/C=3.5\).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}

}

\end{figure}

\begin{figure}[H]

\caption{\label{fig-graphs-coef-daf}Coefficients, gain and phase shift
functions for the direct asymmetric filter (DAF).}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}

}

\end{figure}

\hypertarget{ann-ex-r}{%
\section{R code example}\label{ann-ex-r}}

In this appendix we present the \faIcon{r-project} code that can be used
to estimate trend-cycle components for the US Employment with local
polynomial filters.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# \# To install rjd3filters}
\CommentTok{\# remotes::install\_github("rjdemetra/rjd3toolkit")}
\CommentTok{\# remotes::install\_github("rjdemetra/rjd3x11plus")}
\CommentTok{\# remotes::install\_github("rjdemetra/rjd3filters")}
\FunctionTok{library}\NormalTok{(rjd3filters)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(patchwork)}
\FunctionTok{library}\NormalTok{(zoo)}
\FunctionTok{library}\NormalTok{(forecast)}
\NormalTok{file }\OtherTok{\textless{}{-}}\StringTok{"https://files.stlouisfed.org/files/htdocs/fred{-}md/monthly/2022{-}11.csv"}
\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(file)}
\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ data[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,] }\CommentTok{\# First line removed: it contains the transformation codes}
\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{ts}\NormalTok{(}\FunctionTok{log}\NormalTok{(data[, }\StringTok{"CE16OV"}\NormalTok{]), }\CommentTok{\# We study US employment}
         \AttributeTok{start =} \DecValTok{1959}\NormalTok{, }\AttributeTok{frequency =} \DecValTok{12}\NormalTok{)}
\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =} \FunctionTok{c}\NormalTok{(}\DecValTok{2001}\NormalTok{, }\DecValTok{9}\NormalTok{))}
\NormalTok{last\_dates }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(y), }\DecValTok{8}\NormalTok{))}
\FunctionTok{names}\NormalTok{(last\_dates) }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(zoo}\SpecialCharTok{::}\FunctionTok{as.yearmon}\NormalTok{(last\_dates))}
\NormalTok{last\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(last\_dates, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =}\NormalTok{ x))}

\CommentTok{\# MA of length 13 is appropriated:}
\FunctionTok{select\_trend\_filter}\NormalTok{(y)}
\CommentTok{\# Final estimates:}
\NormalTok{tc\_f }\OtherTok{\textless{}{-}} \FunctionTok{henderson}\NormalTok{(y, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{)}
\FunctionTok{plot}\NormalTok{(y)}
\FunctionTok{lines}\NormalTok{(tc\_f, }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\NormalTok{forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(}\FunctionTok{ts.union}\NormalTok{(y, tc\_f))}

\CommentTok{\# IC{-}ratio computation}
\NormalTok{icr }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
  \FunctionTok{ic\_ratio}\NormalTok{(x, }\FunctionTok{henderson}\NormalTok{(x, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{))}
\NormalTok{\})}
\NormalTok{lp\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
                         \AttributeTok{endpoints =}\NormalTok{ method,}
                         \AttributeTok{ic =}\NormalTok{ icr[i])}
\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(last\_est[[i]], lp\_coef)}
\NormalTok{  \})}
  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
\NormalTok{  res}
\NormalTok{\})}
\NormalTok{lp\_if }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
                         \AttributeTok{endpoints =}\NormalTok{ method,}
                         \AttributeTok{ic =}\NormalTok{ icr[i])}
    \FunctionTok{implicit\_forecast}\NormalTok{(last\_est[[i]], lp\_coef)}
\NormalTok{  \})}
  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
\NormalTok{  res}
\NormalTok{\})}
\FunctionTok{names}\NormalTok{(lp\_est) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(lp\_if) }\OtherTok{\textless{}{-}}  \FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{)}

\CommentTok{\# Local estimates of IC{-}ratios}
\CommentTok{\# We replicate the direct estimates to have}
\CommentTok{\# estimators of the slope and the concavity}
\NormalTok{gen\_MM }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{p=}\DecValTok{6}\NormalTok{, }\AttributeTok{q=}\NormalTok{p, }\AttributeTok{d=}\DecValTok{2}\NormalTok{)\{}
\NormalTok{  X\_gen }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{d =} \DecValTok{1}\NormalTok{, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{q =}\NormalTok{ p)\{}
    \FunctionTok{sapply}\NormalTok{(}\DecValTok{0}\SpecialCharTok{:}\NormalTok{d, }\ControlFlowTok{function}\NormalTok{(exp) }\FunctionTok{seq}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{p, q)}\SpecialCharTok{\^{}}\NormalTok{exp)}
\NormalTok{  \}}
\NormalTok{  k }\OtherTok{=}\NormalTok{ rjd3filters}\SpecialCharTok{::}\FunctionTok{get\_kernel}\NormalTok{(}\StringTok{"Henderson"}\NormalTok{, }\AttributeTok{h =}\NormalTok{ p)}
\NormalTok{  k }\OtherTok{=} \FunctionTok{c}\NormalTok{(}\FunctionTok{rev}\NormalTok{(k}\SpecialCharTok{$}\NormalTok{coef[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{]), k}\SpecialCharTok{$}\NormalTok{coef[}\FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,q)}\SpecialCharTok{+}\DecValTok{1}\NormalTok{])}
\NormalTok{  K }\OtherTok{=} \FunctionTok{diag}\NormalTok{(k)}
\NormalTok{  X }\OtherTok{=} \FunctionTok{X\_gen}\NormalTok{(}\AttributeTok{d=}\NormalTok{d, }\AttributeTok{p =}\NormalTok{ p, }\AttributeTok{q =}\NormalTok{ q)}
\NormalTok{  e1 }\OtherTok{=}\NormalTok{ e2 }\OtherTok{=}\NormalTok{ e3 }\OtherTok{=} \FunctionTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{, }\AttributeTok{nrow =}\NormalTok{ d}\SpecialCharTok{+}\DecValTok{1}\NormalTok{)}
\NormalTok{  e1[}\DecValTok{1}\NormalTok{] }\OtherTok{=} \DecValTok{1}
\NormalTok{  e2[}\DecValTok{2}\NormalTok{] }\OtherTok{=} \DecValTok{1}
\NormalTok{  e3[}\DecValTok{3}\NormalTok{] }\OtherTok{=} \DecValTok{1}
  \CommentTok{\# Estimator of the constant}
\NormalTok{  M1 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e1)}
  \CommentTok{\# Estimator of the slope}
\NormalTok{  M2 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e2)}
  \CommentTok{\# Estimor of the concavity}
\NormalTok{  M3 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e3)}
\NormalTok{  mm }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\AttributeTok{const =}\NormalTok{ M1, }\AttributeTok{slope =}\NormalTok{ M2, }\AttributeTok{concav =}\NormalTok{ M3)}
  \FunctionTok{lapply}\NormalTok{(mm, moving\_average, }\AttributeTok{lags =} \SpecialCharTok{{-}}\NormalTok{p)}
\NormalTok{\}}
\NormalTok{all\_mm }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{0}\NormalTok{, gen\_MM, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{d =} \DecValTok{2}\NormalTok{)}
\NormalTok{est\_slope }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{slope,}
                            \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"slope"}\NormalTok{))}
\NormalTok{est\_concav }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{concav,}
                             \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"concav"}\NormalTok{))}

\NormalTok{henderson\_f }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{h=}\DecValTok{6}\NormalTok{)}\SpecialCharTok{@}\NormalTok{sfilter}
\NormalTok{lp\_filter2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(icr, }\AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)\{}
\NormalTok{  all\_coef }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(icr, }\ControlFlowTok{function}\NormalTok{(ic)\{}
    \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =}\NormalTok{ h,}
              \AttributeTok{kernel =}\NormalTok{ kernel,}
              \AttributeTok{endpoints =}\NormalTok{ method,}
              \AttributeTok{ic =}\NormalTok{ ic)}
\NormalTok{  \})}
\NormalTok{  rfilters }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{h, }\ControlFlowTok{function}\NormalTok{(i)\{}
\NormalTok{    q}\OtherTok{=}\NormalTok{h }\SpecialCharTok{{-}}\NormalTok{i}
\NormalTok{    all\_coef[[i]][,}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"q=\%i"}\NormalTok{, q)]}
\NormalTok{  \})}
  \FunctionTok{finite\_filters}\NormalTok{(henderson\_f, }\AttributeTok{rfilters =}\NormalTok{ rfilters)}
\NormalTok{\}}
\NormalTok{loc\_lc\_est }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}
\NormalTok{loc\_lc\_if }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}
\NormalTok{loc\_ql\_est }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}
\NormalTok{loc\_ql\_if }\OtherTok{\textless{}{-}}
  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
\NormalTok{  \})}

\DocumentationTok{\#\# Plots}
\CommentTok{\# Plots with all the estimates}
\NormalTok{plot\_est }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, data)}
\NormalTok{  joint\_data }\OtherTok{\textless{}{-}}
    \FunctionTok{window}\NormalTok{(joint\_data,}
           \AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(joint\_data))}

\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =}\NormalTok{ last\_dates,}
               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}

\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
              \AttributeTok{data =}\NormalTok{ data\_legend,}
              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{)  }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
\NormalTok{\}}

\NormalTok{plot\_prevs }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, }\FunctionTok{lapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) \{}
\NormalTok{    first\_date }\OtherTok{\textless{}{-}} \FunctionTok{time}\NormalTok{(x)[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}} \FunctionTok{deltat}\NormalTok{(x)}
    \CommentTok{\# The last observed data is added for readability}
    \FunctionTok{ts}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{end =}\NormalTok{ first\_date), x),}
       \AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{frequency =} \FunctionTok{frequency}\NormalTok{(x))}
\NormalTok{  \}))}

\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =} \FunctionTok{sapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(x), }\DecValTok{1}\NormalTok{)),}
               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}
\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
\NormalTok{    forecast}\SpecialCharTok{::}\FunctionTok{autolayer}\NormalTok{(}
      \FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(y)),}
      \AttributeTok{colour =} \ConstantTok{FALSE}
\NormalTok{    ) }\SpecialCharTok{+}
    \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
              \AttributeTok{data =}\NormalTok{ data\_legend,}
              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{) }\SpecialCharTok{+}
    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
\NormalTok{\}}

\NormalTok{all\_est }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_est, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_est),}
             \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_est))}
\NormalTok{all\_if }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_if, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_if),}
            \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_if))}
\NormalTok{y\_lim }\OtherTok{\textless{}{-}} \ConstantTok{NULL}
\NormalTok{all\_plots\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
  \FunctionTok{names}\NormalTok{(all\_est),}
  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_est}\NormalTok{(all\_est[[x]]) }\SpecialCharTok{+}
    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Trend{-}cycle with \%s"}\NormalTok{, x))) }\SpecialCharTok{+}
    \FunctionTok{coord\_cartesian}\NormalTok{(}\AttributeTok{ylim =}\NormalTok{ y\_lim)}
\NormalTok{)}
\NormalTok{all\_plots\_prev }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
  \FunctionTok{names}\NormalTok{(all\_if),}
  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_prevs}\NormalTok{(all\_if[[x]]) }\SpecialCharTok{+}
    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Implicit forecasts with \%s"}\NormalTok{, x)))}
\NormalTok{)}
\CommentTok{\# Combine all plots:}
\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_est, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_prev, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2024-02-12 supp_material.qmd
first commit

--- 2024-02-12 article_jos.qmd
@@
 ---
-title: "Improving real-time trend estimates using local parametrisation of local polynomial regression filters"
+title: "Supplemental material"
 format:
@@
     pdf-engine: pdflatex
-    keep-tex: true
     toc: false
@@
       - right = 2.5cm
-    cite-method: biblatex
 linestretch: 2
@@
 fig-cap-location: top
-bibliography: biblio.bib
-csl: chicago-author-date.csl
 ---
 
-```{r}
-#| include: false
-library(kableExtra)
-```
-
-# Abstract {-}
-
-:::: {.abstract}
-This paper studies and compares real-time estimates of the trend cycle component with moving averages based on local polynomial regression.
-In particular, this theory allows to reproduce Henderson's symmetric and Musgrave's asymmetric filters.
-This paper describes how they can be extended to include a timeliness criterion to minimise the phase shift.
-While asymmetric filters are generally parameterised globally, which can be suboptimal around turning points, this paper proposes a procedure for parameterising them locally.
-An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parameterisation reduces the delay in detecting turning points and reduces revisions.
-The paper also shows how these results can be easily reproduced using the R package `rjd3filters` dedicated to the manipulation of moving averages.
-
-Keywords: time series, trend-cycle, seasonal adjustment, turning points, R statistical software.
-
-JEL Classification: E32, E37.
-::::
-
-# Introduction {-}
-
-Analysis of the economic cycle, and in particular the early detection of turning points in a series, is a major topic in the analysis of economic outlook.
-To this end, economic indicators are generally seasonally adjusted.
-However, in order to improve their interpretability, it may be necessary to perform additional smoothing in order to reduce noise, and thus analyse the trend-cycle component.
-By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series.
-
-A moving average, or linear filter, is a statistical method that consists in applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
-Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
-For example, the X-13ARIMA seasonal adjustment method uses several moving averages to estimate the main components of a time series. 
-Symmetric filters are applied to the center of the series, but when it comes to estimate the most recent points, all of these methods must rely on asymmetric filters. 
-For trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-13ARIMA seasonal adjustment algorithm.
-
-However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
-For example, for trend-cycle extraction, X-13ARIMA uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
-Since the predicted values are linear combinations of past values, this amounts to applying asymmetric averages at the end of the series.
-
-If the classic asymmetric moving averages have good properties regarding the future revisions induced by the process (see for example @pierce1980SA), they create, by construction, phase shifts that impact the real-time estimation of turning points, introducing time delay in the detection.
-
-Using local polynomial regression, @proietti2008 developed general class of symmetric and asymmetric moving averages.
-In particular, it makes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
-However, these methods have two drawbacks.
-Firstly, the phase shift (i.e. the delay in detecting turning points) is not directly controlled.
-Secondly, asymmetric filters depend on a parameter that is generally estimated over the whole series using assumptions that may be wrong locally.
-The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
-The extensions have been implemented in the {{< fa brands r-project >}} `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.
-
-In @sec-propMM, we describe the general properties of moving averages and the associated quality criteria.
-This allows us to understand the foundations behind the construction of moving averages from local polynomial regressions, as well as those behind the two extensions proposed in this article (@sec-nonparamreg).
-In a final section (@sec-comparison), all these methods are compared empirically on simulated and real series.
-
-# Some properties on moving averages {#sec-propMM}
-
-Lots of papers describe the definition and the properties of moving averages and linear filters, see for example @ch12HBSA. 
-In this section we summarize some of the main results to understand the next sections.
-
-Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}$ such as for all time series $X_t$:
-$$
-M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
-$$
-
-- $p+f+1$ is called the *moving average order*.
-
-- When $p=f$ the moving average is said to be *centered*. 
-If we also have $\forall k:\:\theta_{-k} = \theta_k$, the moving average $M_{\boldsymbol\theta}$ is said to be *symmetric*. 
-In this case, the quantity $h=p=f$ is called the *bandwidth*.
-
-## Gain and phase shift functions {#subsec:gain-deph}
-
-To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $X_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
-The moving average $M_{\boldsymbol\theta}$ transforms $X_t$ into:
-$$
-Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
-= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
-$$
-The function  $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
-The frequency response function can equivalently be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ ou $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
-It can be rewritten as:
-$$
-\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
-$$
-where  $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
-This function is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to mesure the phase shift in number of periods.
-For all symmetric moving average we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.
-
-To sum up, applying a moving average to a harmonic times series ($X_t=\e^{-i\omega t}$) affects it in in two different ways:
-
-- by multiplying it by an amplitude coefficient $\rho_{\boldsymbol\theta}\left(\omega\right)$ (gain);
-
-- by "shifting" it in time by $\varphi_{\boldsymbol\theta}(\omega)/\omega$ which directly affects the detection of turning points.
-When  $\varphi_{\boldsymbol\theta}(\omega)/\omega<0$ the time shift is negative: the turning point is detected with delay.
-
-Fourier decomposition allows to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
-Noting $\omega = 2\pi/p$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $p$ periods.
-For example, for a monthly series (12 observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
-In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12[$, i.e. movements recurring at least every 12 months.
-Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
-The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component (undesirable oscillations).
-
-@fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e.
-when no point in the future is known).
-The gain function is greater than 1 on the frequencies associated with the trend-cycle ($[0, 2\pi/12]$): this means that the trend-cycle is well preserved and that short cycles of 1 to 2 years ($[2\pi/24,2\pi/12]$) are even amplified.
-However, cycles of 8 to 12 months ($[2\pi/12, 2\pi/8]$), considered undesirable because associated with the irregular, are also amplified: this can lead to the detection of false turning points.
-At other frequencies, the gain function is less than 1, but always positive: this means that the series smoothed by this moving average will always contain noise, even if it is attenuated.
-Analysis of the phase shift shows that the shorter the cycles, the greater the phase shift: this means that on series smoothed by this moving average, turning points could be detected at the wrong date.
-
-```{r, include = FALSE, cache = FALSE}
-knitr::opts_chunk$set(fig.pos = "H")
-```
-
-```{r, echo=FALSE, out.width="100%"}
-#| label: fig-graphsmusgrave
-#| fig-cap: "Coefficients, gain and phase shift function for the Musgrave filter in real time with $I/C=3.5$."
-img <- sprintf("img/filters_used/%s.%s", tolower("musgrave"), "pdf")
-knitr::include_graphics(img)
-```
-
-## Desirable properties of a moving average
-
-To decompose a time series into a seasonal component, a trend-cycle and the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses a succession of moving averages, all with specific constraints.
-
-In our case, we assume that our initial series $y_t$ is seasonally adjusted and can be written as the sum of a trend-cycle, $TC_t$, and an irregular component, $I_t$:
-$$
-y_t=TC_t+I_t.
-$$
-The aim will be to build moving averages that best preserve the trend-cycle ($M_{\boldsymbol\theta} (TC_t)\simeq TC_t$) and minimise noise ($M_{\boldsymbol\theta} (I_t)\simeq 0$).
-
-### Trend preservation
-
-Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg).
-In order to best preserve the trend-cycles, we try to have moving averages that preserve the polynomial trends.
-A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t:M_{\boldsymbol\theta} f(t)=f(t)$.
-
-We have the following properties for the moving average $M_{\boldsymbol\theta}$:
-
-- To preserve a constant trend $X_t=a$ we need:
-$$
-\forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
-$$
-The sum of the coefficients of the moving average $\sum_{k=-p}^{+f}\theta_k$ must then be equal to $1$.
-
-- To preserve linear trends $X_t=at+b$ we need:
-$$
-\forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
-$$
-Which is equivalent to:
-$$
-\sum_{k=-p}^{+f}\theta_k=1
-\quad\text{et}\quad
-\sum_{k=-p}^{+f}k\theta_k=0.
-$$
-- In general, $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
-$$
-\sum_{k=-p}^{+f}\theta_k=1
- \text{ and }
-\forall j \in \left\llbracket 1,d\right\rrbracket:\:
-\sum_{k=-p}^{+f}k^j\theta_k=0.
-$$
-- If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.
-
-### Noise reduction
-
-All time series are affected by noise, which can blur the extraction of trends and cycles.
-This is why we try to reduce this noise (by reducing its variance) while preserving the relevant trends.
-The sum of the squares of the coefficients $\sum_{k=-p}^{+f}\theta_k^2$ is the *reduction of variance* ratio.
-
-Indeed, let $\{\varepsilon_t\}$ a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$:
-$$
-\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
-= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
-\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
-$$
-
-## Real-time estimation and asymmetric moving average {#sec-mmasym}
-
-### Asymmetric moving averages and forecasts {#sec-mmetprev}
-
-For symmetric filters, the phase shift function is equal to zero (modulo $\pi$).
-So there is no delay in detecting turning points.
-However, they cannot be used in the beginning and in the end of the time series because no past or future value can be used. 
-Thus, for real-time estimation, it is needed to build asymmetric moving average that approximate the symmetric moving average.
-
-Several approaches can be used for real-time estimation:
-
-1. Use asymmetric moving averages to take account of the lack of available data;
-
-2. Apply symmetrical filters to extended forecast series.
-This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period:\
-"*As the first $m$ and last $m$ terms of the series cannot be reached directly by the formula, the series should be graphically extended by m terms at both ends, first plotting the observations on paper as ordinates, and then extending the curve along what seems to be its probable course, and measuring the ordinates of the extended portions.*
-*It is not necessary that this extension should coincide with what would be the true course of the curve in those parts. *
-*The important point is that the m terms thus added, taken together with the $m+1$ adjacent given terms, should follow a curve whose form is approximately algebraic and of a degree not higher than the third.*"\
-This is also the approach used in the X-13ARIMA seasonal adjustment method, which extends which extends the series by one year with an ARIMA model.
-
-Ultimately, the second method is equivalent to using asymmetric moving averages, since forecasts are linear combinations of the past.
-
-Conversely, the implicit forecasts of an asymmetric moving average can be deduced from a reference symmetric moving average. 
-This allows to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
-Let us denote $v=(v_{-h},\dots, v_{h})$ the reference symmetrical moving average and $w^0,\dots w^{h-1}$ a sequence of asymmetrical moving averages, of order $h+1$ to $2h$ used to estimate the last $h$ points with, for convention, $w_t^q=0$ for $t>q$.
-This means that $w^0=(w_{-h}^0,\dots, w_{0}^0)$ is used for estimation in real time (when no points are known in the future), $w^1=(w_{-h}^1,\dots, w_{1}^1)$ for estimation of the penultimate point (when only one point is known in the future), and so on.
-Let also note $y_{-h},\dots,y_{0}$ the observed series studied and $y_{1}^*,\dots y_h^*$ the implicit forecast induced by $w^0,\dots w^{h-1}$.
-This means that for all $q$ we have:
-$$
-\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
-=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
-$$
-This is equivalent to:
-$$
-\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
-=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
-$$
-To sum up, in matrix form, this is equivalent to solving:
-$$
-\scriptstyle
-\begin{pmatrix}
-  v_1 & v_2 & \cdots & v_h \\
-  v_1 - w_1^1 & v_2 & \cdots & v_h \\
-  \vdots & \vdots & \cdots & \vdots \\
-   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
-\end{pmatrix}
-\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
-\begin{pmatrix}
-  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
-  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
-  \vdots & \vdots & \cdots & \vdots \\
-  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
-\end{pmatrix}
-\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
-$$
-This is implemented in the `rjd3filters::implicit_forecast()` function.
-
-As highlighted by @wildischis2004, extending the series through forecasting with an ARIMA model is equivalent to calculating asymmetric filters whose coefficients are optimised in relation to the one-step ahead forecast.
-In other words, the aim is to minimise the revisions between the first and last estimates (with the symmetric filter).
-However, the phase shift induced by the asymmetric filters is not controlled: we might prefer to have faster detection of turning points and a larger revision rather than just minimising the revisions between the first and last estimates.
-Furthermore, since the coefficients of the symmetric filter (and therefore the weight associated with distant forecasts) decrease slowly, we should also be interested in the performance of multi-step ahead forecasting.
-This is why it may be necessary to define alternative criteria for judging the quality of asymmetric moving averages.
-
-### Quality indicators for asymmetric moving averages {#subsec:crit-qual}
-
-For asymmetric filters, most of the criteria come from those defined by @ch15HBSA and @trilemmaWMR2019 to build asymmetric filters. 
-They are summarised in Table\ \ref{tab-QC} and can be calculated using the `rjd3filters::diagnostic_matrix()` function.
-
-@ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$).
-Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
-This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::fst_filter()`.
-
-@trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
-A simplified version of this method can be used in {{< fa brands r-project >}} with the function `rjd3filters::dfa_filter()`.
-
-```{r, include=FALSE}
-title <- "Quality criteria for a moving average $\\boldsymbol\\theta=(\\theta_k)_{-p\\leq k\\leq f}$ defined by a gain function $\\rho_{\\boldsymbol\\theta}$ and a phase shift function $\\varphi_{\\boldsymbol\\theta}$."
-footnotes <- c("$X_g$ criteria from Grun-Rehomme, Guggemos, and Ladiray (2018) and $X_w$ criteria from Wildi and McElroy (2019).",
-               "$\\rho_s$ and  $\\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.",
-               "$h$ e is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\\frac{1}{2(1-\\cos(x))}$)."
-)
-```
-
-```{r, echo = FALSE, fig.note = footnotes}
-#| label: tab-QC
-#| eval: false
-criteria_mm <- data.frame(Sigle = c("b_c", "b_l", "b_q",
-                                    "F_g", "S_g", "T_g",
-                                    "A_w", "T_w", "S_w", "R_w"),
-                          Description = c(sprintf("%s bias", c("Constant", "Linear", "Quadratic")),
-                                          sprintf("%s (Guggemos)",
-                                                  c("Variance reduction / Fidelity", "Smoothness", "Timeliness")),
-                                          sprintf("%s (Wildi)",c("Accuracy", "Timeliness", "Smoothness", "Residual"))),
-                          Formule = c("\\sum_{k=-p}^{+f}\\theta_{k}-1",
-                                      "\\sum_{k=-p}^{+f}k\\theta_{k}",
-                                      "\\sum_{k=-p}^{+f}k^{2}\\theta_{k}",
-                                      "\\sum_{k=-p}^{+f}\\theta_{k}^{2}",
-                                      "\\sum_{j}(\\nabla^{3}\\theta_{j})^{2}",
-                                      "\\int_{0}^{\\omega_1}\\rho_{\\boldsymbol\\theta}(\\omega)\\sin(\\varphi_{\\boldsymbol\\theta}(\\omega))^{2}\\ud\\omega",
-                                      "2\\int_0^{\\omega_1}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega",
-                                      "8\\int_0^{\\omega_1} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega",
-                                      "2\\int_{\\omega_1}^{\\pi}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega",
-                                      "8\\int_{\\omega_1}^{\\pi} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega"))
-criteria_mm[,1] <- sprintf("$%s$", criteria_mm[,1])
-criteria_mm[,3] <- sprintf("$%s$", criteria_mm[,3])
-library(kableExtra)
-
-criteria_mm %>%
-  kbl(align = "c", booktabs = T, row.names = FALSE,
-        escape = FALSE,caption = title)  %>%
-kable_styling(latex_options = "scale_down") 
-    # add_footnote_kable(stop_centering = TRUE)
-```
-
-```{=tex}
-\begin{table}[!h]
-\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
-{\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{ccc}
-\toprule
-Code & Description & Formula\\
-\midrule
-$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
-$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
-$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
-$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
-$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
-\addlinespace
-$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
-$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-\bottomrule
-\end{tabular}}}
-
-\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}
-
-\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}
-
-\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
-\end{table}
-```
-
-# Non-parametric regression and local polynomial regression {#sec-nonparamreg}
-
-Many trend-cycle extraction methods are based on non-parametric regressions, which are particularly flexible because they do not assume any predetermined dependency in the predictors. 
-In practice, local regressions can be used. 
-More specifically, consider a set of points $(x_i,y_i)_{1\leq i\leq n}$. 
-Non-parametric regression involves assuming that there exists a function $\mu$, to be estimated, such that $y_i=\mu(x_i)+\varepsilon_i$ with $\varepsilon_i$ an error term.
-According to Taylor's theorem, for any point $x_0$, if $\mu$ is differentiable $d$ times, then:
-$$
-\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
-\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
-$$ {#eq-taylor}
-where $R_d$ is a negligible residual term in the neighbourhood of $x_0$. 
-In a neighbourhood $h(x_0)$ around $x_0$, $\mu$ can be approximated by a polynomial of degree $d$. 
-The quantity $h(x_0)$ is called the *bandwidth*.
-If $\varepsilon_i$ is white noise, then we can estimate by least squares $\mu(x_0)$ using the observations which are in $\left[x_0-h(x_0),x_0+h(x_0)\right]$.
-
-In practice, this means assuming that the trend is locally polynomial. 
-Various estimation methods can be used to derive symmetrical and asymmetrical moving averages.  
-@GrayThomson1996 propose a complete statistical framework which makes it possible, in particular, to model the error in approximating the trend by local polynomials.
-However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
-Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
-This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::rkhs_filter()`.
-
-## Symmetric moving averages and local polynomial regression
-
-Using @proietti2008's notation, we assume that our time series $y_t$ can be decomposed into
-$$
-y_t=\mu_t+\varepsilon_t,
-$$
-where $\mu_t$ is the trend and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
-Within a neighbourhood $h$ of $t$, the local trend $\mu_t$ is approximated by a polynomial of degree $d$, such that $\mu_t\simeq m_{t}$ with:
-$$
-\forall j\in\left\llbracket -h,h\right\rrbracket:\:
-y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
-$$
-The problem of trend extraction is equivalent to estimating $m_t=\beta_0$ (the constant in the previous formula).
-
-In matrix notation:
-$$
-\underbrace{\begin{pmatrix}y_{t-h}\\
-y_{t-(h-1)}\\
-\vdots\\
-y_{t}\\
-\vdots\\
-y_{t+(h-1)}\\
-y_{t+h}
-\end{pmatrix}}_{\boldsymbol y}=\underbrace{\begin{pmatrix}1 & -h & h^{2} & \cdots & (-h)^{d}\\
-1 & -(h-1) & (h-1)^{2} & \cdots & (-(h-1))^{d}\\
-\vdots & \vdots & \vdots & \cdots & \vdots\\
-1 & 0 & 0 & \cdots & 0\\
-\vdots & \vdots & \vdots & \cdots & \vdots\\
-1 & h-1 & (h-1)^{2} & \cdots & (h-1)^{d}\\
-1 & h & h^{2} & \cdots & h^{d}
-\end{pmatrix}}_{\boldsymbol X}\underbrace{\begin{pmatrix}\beta_{0}\\
-\beta_{1}\\
-\vdots\\
-\vdots\\
-\vdots\\
-\vdots\\
-\beta_{d}
-\end{pmatrix}}_{\boldsymbol \beta}+\underbrace{\begin{pmatrix}\varepsilon_{t-h}\\
-\varepsilon_{t-(h-1)}\\
-\vdots\\
-\varepsilon_{t}\\
-\vdots\\
-\varepsilon_{t+(h-1)}\\
-\varepsilon_{t+h}
-\end{pmatrix}}_{\boldsymbol \varepsilon}.
-$$
-
-To estimate $\boldsymbol \beta$ we need $2h+1\geq d+1$ and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
-$$
-S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
-$$
-where $\kappa_j$ is a set of weights called *kernel*.
-We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
-With $\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}$, the estimate of the trend is:
-$$
-\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
-$$ {#eq-mmsym}
-To conclude, the estimate of the trend $\hat{m}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
-Moreover, $\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}$ so:
-$$
-\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
-$$
-Hence, the filter $\boldsymbol w$ preserve deterministic polynomial of order $d$.
-
-Regarding parameter selection, the general consensus is that the choice between different kernels is not crucial.
-See, for example, @cleveland1996smoothing or @Loader1999.
-The only desired constraints on the kernel are that it assigns greater weight to the central estimation ($\kappa_0$) and decreases towards 0 as it moves away from the central estimation.
-The uniform kernel should therefore be avoided.
-It is then preferable to focus on two other parameters:
-
-- the degree of the polynomial, denoted by $d$: if it is too small, there is a risk of biased estimates of the trend-cycle, and if it is too large, there is a risk of excessive variance in the estimates (due to over-adjustment);
-
-- the number of neighbors $2h+1$ (or the window $h$): if it is too small, then there will be too little data for the estimates (resulting in high variance in the estimates), and if it is too large, the polynomial approximation will probably be wrong, leading to biased estimates.
-
-In this paper we will use the Henderson kernel:
-$$
-\kappa_{j}=\left[1-\frac{j^2}{(h+1)^2}\right]
-\left[1-\frac{j^2}{(h+2)^2}\right]
-\left[1-\frac{j^2}{(h+3)^2}\right].
-$$
-However, several other kernels are available in `rjd3filters`.
-
-## Asymmetric moving averages and local polynomial regression {#sec-lppasymf}
-
-As mentioned in @sec-mmetprev, for real-time estimation, several approaches can be used:
-
-1. Apply symmetric filters to the series extended by forecasting $\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket$.
-
-2. Build an asymmetric filter by local polynomial approximation on the available observations ($y_{t}$ for $t\in\left\llbracket n-h,n\right\rrbracket$).
-
-3. Build asymmetric filters that minimise the mean squared error of revision under polynomial trend reproduction constraints.
-
-@proietti2008 show that the first two approaches are equivalent when forecasts are made by polynomial extrapolation of degree $d$.
-They are also equivalent to the third approach under the same constraints as those of the symmetric filter.
-The third method is called *direct asymmetric filter* (DAF).
-This method is used for real-time estimation in the STL seasonal adjustment method (*Seasonal-Trend decomposition based on Loess*, see @cleveland90).
-Although DAF filter estimates are unbiased, this is at the cost of greater variance in the estimates.
-
-To solve the problem of the variance of real-time filter estimates, @proietti2008 propose a general method for constructing asymmetric filters that allows a bias-variance trade-off.
-This is a generalisation of @musgrave1964set's asymmetric filters (used in the X-13ARIMA seasonal adjustment algorithm).
-
-The initial series is modelled here as:
-$$
-\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
-\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
-$$ {#eq-lpgeneralmodel}
-where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$.
-The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
-These constraints are represented by the matrix $\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
-The problem is equivalent to finding $v$ which minimises:
-$$
-\varphi(\boldsymbol v)=
-\underbrace{
-  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
-  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
-  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
-}_\text{mean square revision error}+
-\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
-$$ {#eq-lppasym}
-where $\boldsymbol l$ is a vector of Lagrange multipliers.
-
-When $\boldsymbol U=\boldsymbol X$, the constraint is equivalent to preserving polynomials of degree $d$: we find asymmetric direct filters (DAF) when $\boldsymbol D=\boldsymbol K^{-1}$.
-
-When $\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}$, $\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
-This filter assumes that, for real-time estimation, the data is generated by a linear process and that the asymmetric filters preserve the constants ($\sum v_i=\sum \theta_i=1$).
-These asymmetric filters depend on the ratio $\lvert\delta_1/\sigma\rvert$, which, assuming that the trend is linear and the bias constant, can be linked to the I-C ratio $R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}$: $\delta_1/\sigma=2/(R\sqrt{\pi})$.
-This ratio is used in particular in X-13ARIMA to determine the length of the Henderson filter. 
-For monthly data:
-
-- If the ratio is large ($3.5< R$) then a 23 term filter is used (to remove more noise).
-
-- If the ratio is small ($R<1$) a 9 term filter is used.
-
-- Otherwise (most of the cases) a 13-term filter is used.
-
-When $\boldsymbol U$ corresponds to the first $d^*+1$ columns of $\boldsymbol X$, $d^*<d$, the constraint is to reproduce polynomial trends of degree $d^*$.
-This introduces bias but reduces the variance. 
-Thus, @proietti2008 proposes three classes of asymmetric filters:
-
-1. *Linear-Constant* (LC): $y_t$ linear ($d=1$) and $\boldsymbol v$ preserves constant signals ($d^*=0$). 
-We obtain Musgrave filters when the Henderson kernel is used and the Henderson filter is used as the symmetric filter.
-
-2. *Quadratic-Linear* (QL): $y_t$ quadratic ($d=2$) and $\boldsymbol v$ preserves linear signals ($d^*=1$).
-
-3. *Cubic-Quadratic* (CQ): $y_t$ cubic ($d=3$) and $\boldsymbol v$ preserves quadratic signals ($d^*=2$).
-
-Appendix [-@sec-an-cof] shows the coefficients, gain and phase shift functions of the four asymmetric filters.
-
-Table\ \ref{tab-criteriaLp} compares the quality criteria of the different methods using the Henderson filter and $h=6$ (a symmetric filter with 13 terms). 
-For real-time filters ($q=0$), a more complex asymmetrical filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance.
-This result varies as $q$ increases: for $q=2$, the QL filter has greater timeliness than the LC filter. 
-This surprising result highlights the fact that the phase shift is not controlled by @proietti2008's approach.
-
-In terms of revision ($A_w+S_w+T_w+R_w$), the LC and QL filters consistently outperform the CQ and DAF filters.
-
-These theoretical properties are consistent with the empirical results observed in @sec-comparison.
-Revisions are more important for the CQ and DAF filters, which leads to greater variability in estimations and a longer delay in detecting turning points.
-The globally parameterised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter.
-Indeed, to identify a turning point at date $t$, it is necessary to have knowledge of at least two points after the given date to confirm the reversal in trend.
-
-```{r echo = FALSE}
-#| label: tab-criteraLp
-#| eval: false
-#| tab-cap: 
-title <- "Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$."
-
-lp_diagnostics <- readRDS("data/lp_diagnostics_henderson.RDS")
-lp_diagnostics[,"$ EQM_w $"] <- rowSums(lp_diagnostics[,grep("_w", colnames(lp_diagnostics), fixed = TRUE)])
-colnames(lp_diagnostics) <- gsub(" ?\\$ ?","$",colnames(lp_diagnostics))
-lp_diagnostics[,1] <- gsub(" ?\\$ ?","$",lp_diagnostics[,1])
-groupement <- table(lp_diagnostics[,1])
-lp_diagnostics[,-(1:2)] <- round(lp_diagnostics[,-(1:2)],2)
-lp_diagnostics[,-1] %>%
-  kable(format.args = list(digits = 2,
-                           decimal.mark = "."), align = "c", booktabs = T, row.names = FALSE,
-        escape = FALSE,caption = title,format = "latex") %>%
-  kable_styling(latex_options=c(#"striped",
-      "hold_position")) %>%
-  pack_rows(index = groupement, escape = FALSE) %>%
-  footnote("With $EQM_w=A_w + S_w + T_w + R_w$.")
-```
-
-```{=tex}
-\begin{table}[!h]
-\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
-\centering
-\begin{tabular}[t]{cccccccccccc}
-\toprule
-Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=0$}}\\
-\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
-\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=1$}}\\
-\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
-\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=2$}}\\
-\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
-\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
-\bottomrule
-\end{tabular}
-
-\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
-\end{table}
-```
-
-## Extension with the timeliness criterion {#sec-lptimeliness}
-
-One drawback of the previous method is the lack of control over the phase shift. 
-However, it is possible to improve the modelling by incorporating in @eq-lppasym the timeliness criterion defined by @ch15HBSA.
-This was proposed by Jean Palate, then coded in Java and integrated into `rjd3filters`.
-
-Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
-With $\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
-$$
-T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
-\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
-$$
-Furthermore, the objective function $\varphi$ of @eq-lppasym can be rewritten:
-\begin{align*}
-\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
-  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
-  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
-2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
-&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
-\begin{cases}
-\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
-\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
-\boldsymbol c\text{ a constant independent of }\boldsymbol v
-\end{cases}.
-\end{align*}
-Adding the timeliness criterion:
-$$
-\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
-2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
-\widetilde {\boldsymbol c}\quad\text{with }
-\begin{cases}
-\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
-\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
-\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
-\end{cases},
-$$
-where $\alpha_T$ is the weight associated to the timeliness criterion. 
-With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
-This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
-This extension is implemented in the {{< fa brands r-project >}} function `rjd3filters::lp_filter()`.
-
-One drawback is that $T_g(\boldsymbol v)$ is not normalized: the weight $\alpha_T$ has no economic sense.
-It is then difficult to calibrate the value of $\alpha_T$ since the value of $\lvert\delta/\sigma\rvert$ must also be defined.
-That's why in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
-However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parameterisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).
-
-## Local parameterisation of asymmetric filters {#sec-localic}
-
-Asymmetric filters are usually parameterised globally: $\lvert\delta/\sigma\rvert$ estimated on all the data using the IC-ratio or a cross-validation criterion. 
-However, we might prefer a local parameterisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
-Indeed, although the overall parameterisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetrical filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
-For example, with the LC method, global parameterisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up to the turning point.
-
-This is what is proposed in this article, with a local parameterisation of the asymmetric filters by estimating $\delta$ and $\sigma^2$ separately.
-Although this does not give an unbiased estimator of the ratio $\lvert\delta/\sigma\rvert$, it does allow the main evolutions to be captured, such as the decay towards 0 before a turning point and the growth after the turning point for the LC method:
-
-- The variance $\sigma^2$ can be estimated using the observed data set and the symmetrical filter $(w_{-p},\dots,w_p)$:
-$$
-\hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
-$$
-- The parameter $\delta$ can be estimated by moving average from @eq-mmsym. 
-For example, for the LC method we can use the moving average $\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}$ to obtain a local estimate of the slope and for the QL method we can use $\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}$ to obtain a local estimate of the concavity.
-The DAF method then simply allows us to calculate the associated asymmetric moving averages.  
-Although a moving average of different length to that used for estimating the trend could be considered, this seems to degrade the results in terms of phase shift (using the same methodology as in @sec-comparison).
-Furthermore, for the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
-Here, we have chosen to model a trend of degree 2: this reduces the phase shift but slightly increases the revisions linked to the first estimate of the trend-cycle. 
-@fig-mmpenteconcac shows the moving averages used.
-
-```{r, echo=FALSE, out.width="90%"}
-#| label: fig-mmpenteconcac
-#| fig-cap: Moving averages used for real-time estimation of slope and concavity.
-img <- sprintf("img/filters_used/mm_penteconcavite.%s", "pdf")
-knitr::include_graphics(img)
-```
-
-In the empirical applications of @sec-comparison, the final local parameterisation is obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
-@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two the local parameterisation:
-
-- @fig-mmpenteconcac-ex-1: using a symmetric filter to estimate $\delta_1$ (final local parameterisation).
-
-- @fig-mmpenteconcac-ex-2: using the asymmetric filter which need two points in the futur to estimate $\delta_1$.
-This is closed to the real-time estimates since two points in the future are needed to detect a turning point.
-
-The turning points are clearly detected by the local estimators, but there is much more variance in the estimate of $|\delta_1/\sigma|$ in the real-time estimate.
-This will lead to another source of revisions in the intermediate estimates.
-
-```{r, echo=FALSE, out.width="90%"}
-#| label: fig-mmpenteconcac-ex
-#| fig-cap: "Comparison of the estimates of $|\\delta_1/\\sigma|$ with a local estimator and global estimator (with IC ratio) on a simulated series, the vertical lines being the simulated turning points."
-#| fig-subcap: 
-#|   - "Symmetric moving average used for $\\hat \\delta_1$."
-#|   - "Asymmetric moving average using two futur points ($q=2$) used for $\\hat \\delta_1$."
-#| layout-ncol: 1
-knitr::include_graphics(c("img/filters_used/mm_penteconcavite_ex.pdf",
-                          "img/filters_used/mm_penteconcavite_ex2.pdf"))
-```
-
-# Comparison of different methods {#sec-comparison}
-
-The different methods are compared on simulated and real data.
-
-For all series, a symmetrical 13-term filter is used.
-These methods are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetrical Henderson filter.
-The ARIMA model is determined automatically, using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
-
-The performance of the different methods is judged on criteria relating to revisions (between two consecutive estimates and in relation to the final estimate) and the number of periods required to detect turning points.
-
-## Simulated series
-
-### Methodology
-
-Following a methodology close to that of @DarneDagum2009, nine monthly series are simulated between January 1960 and December 2020 with different levels of variability. 
-Each simulated series $y_t= C_t+ T_t + I_t$ can be written as the sum of three components:
-
-- the cycle $C_t = \rho [\cos (2 \pi t / \lambda) +\sin (2 \pi t / \lambda)]$, $\lambda$ is fixed at 72 (6-year cycles, so there are 19 detectable turning points);
-
-- the trend $T_t = T_{t-1} + \nu_t$ with $\nu_t \sim \mathcal{N}(0, \sigma_\nu^2)$, $\sigma_\nu$ being fixed at $0.08$;
-
-- and the irregular $I_t = e_t$ with $e_t \sim \mathcal{N}(0, \sigma_e^2)$.
-
-For the different simulations, we vary the parameters $\rho$ and $\sigma_e^2$ in order to obtain series with different signal-to-noise ratios (see @fig-graphs-data-simul):
-
-- High signal-to-noise ratio (i.e. low I-C ratio and low variability): $\sigma_e^2=0.2$ and $\rho = 3.0, 3.5$ or $4.0$ (I-C ratio between 0.9 and 0.7);
-
-- Medium signal-to-noise ratio (i.e. medium I-C ratio and medium variability): $\sigma_e^2=0.3$ and $\rho = 1.5,\, 2.0$ or $3.0$ (I-C ratio between 2.3 and 1.4);
-
-- Low signal-to-noise ratio (i.e. high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).
-
-```{r, echo=FALSE, out.width="100%"}
-#| label: fig-graphs-data-simul
-#| fig-cap: "Simulated series with low ($\\sigma_e^2=0.2$ and $\\rho = 3,5$), medium ($\\sigma_e^2=0.3$ and $\\rho = 2.0$) and high variability ($\\sigma_e^2=0.4$ and $\\rho = 1.0$)."
-img <- sprintf("img/simulations/simul_data.%s", "pdf")
-knitr::include_graphics(img)
-```
-
-For each series and each date, the trend-cycle is estimated using the different methods presented in this report.
-Three quality criteria are also calculated:
-
-1. Computation of the phase shift in the detection of turning points. The definition of @Zellner1991 is used to determine the turning points:  
-    - An upturn occurs when the economy moves from a phase of recession to a phase of expansion. 
-    This is the case at date  $t$ when $y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}$.  
-    - A downturn occurs when the economy moves from a phase of expansion to a phase of recession.
-This is the case at date  $t$ when $y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}$.  
-    It thus takes at least 2 months to detect a turning point.  
-    The phase shift is often defined as the number of months required to detect the right turning point.
-    Here we use a slightly modified criterion: the phase shift is defined as the number of months needed to detect the right turning point without any future revision.
-    Indeed, it can happen that the right turning point is detected by asymmetric filters but is not detected by the final estimate using a symmetric filter (this is the case for 41 reversal points out of the 9 series with asymmetric Musgrave filters) or that there are revisions in successive estimates (this is the case for 7 turning points out of the 9 series with asymmetric Musgrave filters).
-  Finally, relatively few turning points are detected at the right date with the final estimate.
-  With the 13-term Henderson filter, 18 are correctly detected in series with low variability (out of 57 possible), 11 in series with medium variability and 12 in series with high variability.
-
-2.
-Computation of two revision criteria: the average of the relative deviations between the $q$^th^ estimate and the last estimate $MAE_{fe}(q)$ and the average of the relative deviations between the $q$^th^ estimate and the $q+1$^th^ estimate $MAE_{qe}(q)$.
-$$
-MAE_{fe}(q)=\mathbb E\left[
-\left|\frac{
-y_{t|t+q} -  y_{t|last}
-}{
- y_{t|last}
-}\right|
-\right]
-\quad\text{and}\quad
-MAE_{qe}(q)=\mathbb E\left[
-\left|\frac{
-y_{t|t+q} - y_{t|t+q+1}
-}{
-y_{t|t+q+1}
-}\right|
-\right].
-$$
-
-## Comparison
-
-Excluding for the moment the local parameterisations of the polynomial filters, the linear-constant polynomial filter (LC) seems to give the best results in terms of delay in the detection of turning points (@fig-graphstpsimul).
-Performance is relatively close to that obtained by extending the series using an ARIMA model.
-However, when variability is low, the LC filter seems to give poorer results and the quadratic-linear polynomial filter (QL) seems to give the best results.
-
-For series with moderate variability, local parameterisation of the LC and QL filters reduces the phase shift.
-For series with high variability, the phase shift is only reduced by using the final $\hat\delta$ parameters: real-time estimates seem to add more variance.
-For series with low variability, performance seems to be slightly improved only with the LC filter.
-
-```{r, echo=FALSE, out.width="100%", fig.note = footnotes}
-#| label: fig-graphstpsimul
-#| fig-cap: Distribution of phase shifts on simulated series.
-img <- sprintf("img/simulations/phase_shift_simul.%s", "pdf")
-knitr::include_graphics(img)
-```
-
-In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability (Table\ \ref{tab-simulrev}).
-In general, LC filters always minimise revisions (with relatively small effect from the local parameterisation of the filters) and revisions are greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
-
-For the QL filter, there is a large revision between the second and third estimates: this may be due to the fact that for the second estimate (when one point in the future is known), the QL filter assigns a greater weight to the estimate in $t+1$ than to the estimate in $t$, which creates a discontinuity.
-This revision is considerably reduced by parameterising the filter locally.
-For polynomial filters other than the LC filter, the large revisions at the first estimate were to be expected given the coefficient curve: a very large weight is associated with the current observation and there is a strong discontinuity between the moving average used for real-time estimation (when no point in the future is known) and the other moving averages.
-
-Extending the series using an ARIMA model gives revisions with the latest estimates of the same order of magnitude as the LC filter, but slightly larger revisions between consecutive estimates, particularly between the fourth and fifth estimates (as might be expected, as highlighted in @sec-mmetprev).
-
-```{r, echo = FALSE}
-#| label: tab-simulrev
-#| tbl-cap: "Average of the relative deviations of the revisions for the different filters on simulated series with medium variability."
-#| eval: false
-rev_table <- readRDS("data/simulations_revisions.RDS")
-pack_row_index = c("$MAE_{fe}(q) = \\mathbb E\\\\left[\\\\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\\\\right|\\\\right]$" = nrow(rev_table) / 2,
-                   "$MAE_{ce}(q)=\\mathbb E\\\\left[
-\\\\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\\\\right|
-\\\\right]$" = nrow(rev_table) / 2)
-p <- rev_table  %>%
-  kable(format.args = list(digits = 2,
-                           decimal.mark = "."),
-        align = "c", booktabs = T, row.names = FALSE, 
-        label = "tab-simulrev",
-        escape = FALSE,
-        format = "latex") %>%  
-  kable_styling(latex_options=c(#"striped",  
-    "hold_position")) %>% 
-  pack_rows(index = pack_row_index, escape = FALSE) # %>%
-    # add_footnote_kable(stop_centering = TRUE)
-```
-
-```{=tex}
-\begin{table}[!h]
-\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
-filters on simulated series with medium variability.}
-\centering
-\begin{tabular}{ccccccc}
-\toprule
-Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
-\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
-\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
-\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
-\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
-\right]$}}\\
-\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
-\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
-\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
-\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
-\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
-\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
-\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
-\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
-\bottomrule
-\end{tabular}
-\end{table}
-```
-
-## Real series
-
-The differences between the methods are also illustrated using an example from the FRED-MD database (@fredmd) containing economic series on the United States.
-The series studied correspond to the database published in November 2022.
-The series studied is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
-This point and this series were chosen because the economic turnaround is particularly visible in the raw series (@fig-ce16ov-previmp-lp), because this series is used for dating economic cycles and because the FRED-MD database facilitates the reproducibility of the results (thanks to the availability of series published on past dates).
-Studying the series up to January 2020, this series has a medium variability: a symmetrical filter of 13 terms is therefore appropriate.
-Appendix [-@ann-ex-r] shows how the code to reproduce the different plots.
-
-@fig-ce16ovlp shows the successive estimates of the trend-cycle using the different methods studied.
-
-For this series, the phase shift is 6 months for the LC and CQ methods and the extension of the series by ARIMA. 
-Except for the LC method (where the turning point is first detected in January 2001), the high phase shift is due to revisions in the turning point detected in the intermediate estimates (even if the correct turning point is detected after 2 months, it is no longer detected after 3 months).
-It is of two months for the other methods (QL, DAF).
-
-In this example, local parameterisation does not reduce the phase shift, but it does reduce revisions.
-The QL and DAF polynomials lead to greater variability in the intermediate estimates, especially in February 2001.
-
-```{r, out.width="95%", fig.show="hold"}
-#| label: fig-ce16ovlp
-#| fig-cap:  Successive estimates of the trend-cycle in US employment (in logarithms).
-series <- "CE16OV"
-date <- "fev2001"
-img <- sprintf("img/nber/%s_%s_%s.%s",
-               tolower(series),
-               date,
-               c("lp"),
-               "pdf")
-knitr::include_graphics(img)
-```
-
-The quality of the intermediate estimates can also be analysed using the implicit forecasts of the different methods (@fig-ce16ov-previmp-lp). 
-As a reminder, these are forecasts of the raw series which, by applying the symmetrical Henderson filter to the extended series, give the same estimates as the asymmetrical moving averages. 
-The forecasts of the ARIMA model are naive and do not take the turning point into account, unlike the other methods.
-Finally, local parameterisation of the QL filter produces much more consistent forecasts.
-
-```{r, out.width="95%"}
-#| label: fig-ce16ov-previmp-lp
-#| fig-cap: Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
-series <- "CE16OV"
-date <- "fev2001"
-img <- sprintf("img/nber/%s_%s_prev_imp_%s.%s",
-               tolower(series),
-               date,
-               c("lp"),
-               "pdf")
-knitr::include_graphics(img)
-```
-
-# Conclusion {.unnumbered}
-
-For business cycle analysis, most statisticians use trend-cycle extraction methods, either directly or indirectly.
-They are used, for example, to reduce the noise of an indicator in order to improve its analysis, and models, such as forecasting models, usually use seasonally adjusted series based on these methods.
-
-This paper presents the {{< fa brands r-project >}} package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
-It also offers various functions for the theoretical analysis of moving averages (gain functions, phase shift, quality criteria, etc.) and judging the quality of the latest estimates (for example with implicit forecasts).
-It also makes it easy to combine moving averages and also to integrate custom moving averages into the X-11 seasonal adjustment algorithm (with the `rjd3filters::x11plus_trend()` function).
-`rjd3filters` thus facilitates research into the use of moving averages as part of real-time cycle trend estimation.
-This is the case, for example, with the local parametrisation of Musgrave moving averages and other asymmetric moving averages linked to local polynomial regression, presented in this article.
-
-By comparing the different methods, we can learn a few lessons about the construction of these moving averages.
-
-During economic downturns, asymmetric filters used as an alternative to extending the series using the ARIMA model can reduce revisions to intermediate estimates of the trend-cycle and enable turning points to be detected more quickly.
-
-At the end of the period, modelling polynomial trends of degree greater than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce variance into the estimates (and therefore more revisions) without allowing faster detection of turning points.
-At the end of the period, for estimating the trend-cycle, we can therefore restrict ourselves to methods modelling polynomial trends of degree two or less (linear-constant, LC, and quadratic-linear, QL).
-In addition, parameterising polynomial filters locally as suggested in this paper enables turning points to be detected more quickly (especially for the QL filter).
-Even when the phase shift is not reduced, local parameterisation is recommended because it reduces revisions and produces intermediate estimates that are more consistent with expected future trends.
-However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too short (i.e.
-if the variability of the series is "low"), retaining polynomial trends of degree one or less (LC method) produces poorer results in terms of detecting turning points.
-
-This study could be extended in many ways.
-
-One possible extension would be to look at the impact of filter length on the detection of turning points.
-Asymmetric filters are calibrated using indicators calculated for the estimation of symmetric filters (for example, to automatically determine their length), whereas a local estimate might be preferable.
-Furthermore, we have only focused on monthly series with a 13-term symmetric filter, but the results may be different if the symmetric filter studied is longer or shorter and if we study series with other frequencies (quarterly or daily, for example).
-
-Another possibility would be to study the impact of atypical points: moving averages, like any linear operator, are very sensitive to the presence of atypical points. 
-To limit their impact, in X-13ARIMA a strong correction for atypical points is performed on the irregular component before applying the filters to extract the trend-cycle.
-This leads us to study the impact of these points on the estimation of the trend-cycle and turning points, and also to explore new types of asymmetric filters based on robust methods (such as robust local regressions or moving medians).
-
-\newpage
-
-# References {.unnumbered}
-
-::: {#refs}
-:::
-
 \appendix
-\newpage
 
@@
 
+\newpage
+
 # R code example {#ann-ex-r}

### 2024-02-12 R_ex.R

@@ -1,0 +1,205 @@
+# # To install rjd3filters
+# remotes::install_github("rjdemetra/rjd3toolkit")
+# remotes::install_github("rjdemetra/rjd3x11plus")
+# remotes::install_github("rjdemetra/rjd3filters")
+library(rjd3filters)
+library(ggplot2)
+library(patchwork)
+library(zoo)
+library(forecast)
+file <-"https://files.stlouisfed.org/files/htdocs/fred-md/monthly/2022-11.csv"
+data <- read.csv(file)
+data <- data[-1,] # First line removed: it contains the transformation codes
+y <- ts(log(data[, "CE16OV"]), # We study US employment
+         start = 1959, frequency = 12)
+y <- window(y, end = c(2001, 9))
+last_dates <- c(tail(time(y), 8))
+names(last_dates) <- as.character(zoo::as.yearmon(last_dates))
+last_est <- lapply(last_dates, function(x) window(y, end = x))
+
+# MA of length 13 is appropriated:
+select_trend_filter(y)
+# Final estimates:
+tc_f <- henderson(y, length = 13, musgrave = FALSE)
+plot(y)
+lines(tc_f, col = "red")
+forecast::autoplot(ts.union(y, tc_f))
+
+# IC-ratio computation
+icr <- sapply(last_est, function(x) {
+  ic_ratio(x, henderson(x, length = 13, musgrave = FALSE))
+})
+lp_est <- lapply(c("LC", "QL", "CQ", "DAF"), function(method) {
+  res <- lapply(seq_along(icr), function(i) {
+    lp_coef <- lp_filter(horizon = 6,
+                         kernel = "Henderson",
+                         endpoints = method,
+                         ic = icr[i])
+    rjd3filters::filter(last_est[[i]], lp_coef)
+  })
+  names(res) <- names(last_est)
+  res
+})
+lp_if <- lapply(c("LC", "QL", "CQ", "DAF"), function(method) {
+  res <- lapply(seq_along(icr), function(i) {
+    lp_coef <- lp_filter(horizon = 6,
+                         kernel = "Henderson",
+                         endpoints = method,
+                         ic = icr[i])
+    implicit_forecast(last_est[[i]], lp_coef)
+  })
+  names(res) <- names(last_est)
+  res
+})
[…]

### 2024-02-12 R_filters/plot_coef.R

@@ -1,0 +1,56 @@
+library(rjd3filters)
+library(patchwork)
+library(AQLThesis)
+library(ggplot2)
+
+all_filters <- lapply(c("LC", "QL", "CQ", "DAF"), function(endpoints){
+  lp_filter(
+    horizon = 6,
+    degree = 3,
+    kernel = "Henderson",
+    endpoints = endpoints,
+    ic = 3.5,
+    tweight = 0,
+    passband = pi/12
+  )
+})
+
+
+names(all_filters) <- c("LC", "QL", "CQ", "DAF")
+
+
+Sys.setenv(LANG = "en_US.UTF-8")
+for(i in names(all_filters)){
+  print(i)
+  x <- all_filters[[i]]
+  p <- ggplot_coef(x, q = c(0:6)) / (
+    ggplot_gain(x, q = c(0:6)) +
+      ggplot2::scale_y_continuous(
+        "Gain",
+        breaks = seq(0, 1, by = 0.2))+
+      ggplot2::guides(colour = "none") +
+      ggplot_phase(x, xlim = c(0, 4/12*pi), q = c(0:6))+
+      ggplot2::labs(y = "Phase shift") +
+      ggplot2::guides(colour = "none"))
+  p
+  ggsave(sprintf("img/filters_used/%s.pdf",tolower(i)),
+              width = 8, height = 4.5,
+              plot = p)
+}
+
+x <- all_filters[["LC"]]
+p <- (ggplot_coef(x, q=0) + ggplot2::guides(colour = "none") )/ (
+  ggplot_gain(x, q = 0)  +
+    ggplot2::scale_y_continuous(
+      "Gain",
+      breaks = seq(0, 1, by = 0.2),
+      limits = c(0, NA))+ ggplot2::guides(colour = "none") +
+    ggplot_phase(x, xlim = c(0, pi), q = 0)+
+    ggplot2::labs(y = "Phase shift") + ggplot2::guides(colour = "none")
+)
+p
+
+ggsave("img/filters_used/musgrave.pdf",
+       width = 8, height = 4.5,
+       plot = p)
+

### 2024-02-12 R_fredm/4A_Analyse_est.R

@@ -1,0 +1,42 @@
+# Pour que ce programme puisse tourner, il faut également avoir lancé ceux sous R_local_ic
+source("R_fredm/4_utils.R",encoding = "UTF-8")
+library(patchwork)
+
+
+fst_weights = "weight235"
+fst_degree = 2
+all_tp <- get_all_tp(dossier = "results_fredm/compile_tp_norev/",
+                     fst_weights = "weight235",
+                     fst_degree = 2)
+all_tp_rev <- get_all_tp(dossier = "results_fredm/compile_tp/",
+                     fst_weights = "weight235",
+                     fst_degree = 2)
+detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
+
+
+
+
+Sys.setenv(LANG = "en_US.UTF-8")
+data <- readRDS("data_fredm/byseries/CE16OV.RDS")[["2020"]]
+dput(round(data, 3))
+plot(data)
+all_plots <- get_all_plots(all_tp,
+                           all_tp_rev,
+                           series = "CE16OV", tp_keep ="2001.16666666667",
+                           nb_est = 8)
+wrap_plots(all_plots, ncol =3)
+
+ggsave("img/nber/ce16ov_fev2001_lp.pdf",
+            plot = wrap_plots(all_plots, ncol = 3),
+            width = 8, height = 8)
+
+data <- readRDS("data_fredm/byseries/RETAILx.RDS")[["2020"]]
+all_plots <- get_all_plots(all_tp,
+                           all_tp_rev,
+                           series = "RETAILx", tp_keep ="2007.91666666667",
+                           nb_est = 8)
+wrap_plots(all_plots, ncol = 3)
+
+ggsave("img/nber/retailx_nov2007_lp.pdf",
+            plot = wrap_plots(all_plots, ncol = 3),
+            width = 8, height = 8)

### 2024-02-12 R_fredm/4B_Analyse_implicit_forecast_dt.R

@@ -1,0 +1,60 @@
+# Pour que ce programme puisse tourner, il faut également avoir lancé ceux sous R_local_ic
+source("R_fredm/4_utils.R",encoding = "UTF-8")
+
+library(rjd3filters)
+library(AQLThesis)
+
+Sys.setenv(LANG = "en_US.UTF-8")
+fst_weights <- "weight235"
+fst_degree <- 2
+all_tp <- get_all_tp(dossier = "results_fredm/compile_tp_norev/",
+                     fst_weights = fst_weights,
+                     fst_degree = fst_degree)
+all_tp_rev <- get_all_tp(dossier = "results_fredm/compile_tp/",
+                         fst_weights = "weight235",
+                         fst_degree = 2)
+
+detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
+
+
+series <- "CE16OV"
+tp_keep <- "2001.16666666667"
+all_prevs <- get_all_prevs(series = series, tp_keep = tp_keep,
+                           nb_est = 8, nb_dates_before = 6,
+                           fst_weights = fst_weights,
+                           fst_degree = fst_degree)
+plots <- get_all_plots_prevs(data_prevs = all_prevs,
+                             series = series,
+                             all_tp = all_tp,
+                             all_tp_rev = all_tp_rev,
+                             tp_keep = tp_keep,
+                             dossier = "results_fredm/compile_tp_norev/",
+                             fst_weights = fst_weights,
+                             fst_degree = fst_degree)
+wrap_plots(plots,ncol = 3)
+
+ggsave("img/nber/ce16ov_fev2001_prev_imp_lp.pdf",
+            plot = wrap_plots(plots, ncol = 3),
+            width = 8, height = 8)
+
+
+series <- "RETAILx"
+tp_keep <- "2007.91666666667"
+
+all_prevs <- get_all_prevs(series = series, tp_keep = tp_keep,
+                           nb_est = 8, nb_dates_before = 6,
+                           fst_weights = fst_weights,
+                           fst_degree = fst_degree)
+plots <- get_all_plots_prevs(data_prevs = all_prevs,
+                             series = series,
+                             all_tp = all_tp,
+                             all_tp_rev = all_tp_rev,
+                             tp_keep = tp_keep,
+                             dossier = "results_fredm/compile_tp_norev/",
+                             fst_weights = fst_weights,
[…]

### 2024-02-12 R_fredm/4_utils.R

@@ -1,0 +1,465 @@
+library(dplyr)
+library(ggplot2)
+library(patchwork)
+library(AQLThesis)
+
+length_info <-
+  sapply(list.files("data_fredm/byseriesinfo/",full.names = TRUE),
+         function(f){
+           info <- readRDS(f)
+           info[["2020"]][1]
+         })
+length_info <- data.frame(series = gsub("(.*/)|(.RDS.*)", "",
+                                        names(length_info)),
+                          length = as.numeric(length_info))
+
+troughs_detected <- t(sapply(list.files("results_fredm/compile_tp_norev",pattern = "troughs",full.names = TRUE), function(f){
+  all_t <- readRDS(f)
+  !apply(all_t[,1:35],2,\(x) all(is.na(x)))
+}))
+
+peaks_detected <- t(sapply(list.files("results_fredm/compile_tp_norev",pattern = "peaks",full.names = TRUE), function(f){
+  all_t <- readRDS(f)
+  !apply(all_t[,1:34],2,\(x) all(is.na(x)))
+}))
+
+
+selected_tp <- c("X2001.83333333333", "X2009.41666666667", "X2020.25", "X2001.16666666667",
+                 "X2007.91666666667", "X2020.08333333333")
+not_selected_tp <- c("X1854.91666666667", "X1858.91666666667", "X1861.41666666667",
+                     "X1867.91666666667", "X1870.91666666667", "X1879.16666666667",
+                     "X1885.33333333333", "X1888.25", "X1891.33333333333", "X1894.41666666667",
+                     "X1897.41666666667", "X1900.91666666667", "X1904.58333333333",
+                     "X1908.41666666667", "X1912", "X1914.91666666667", "X1919.16666666667",
+                     "X1921.5", "X1924.5", "X1927.83333333333", "X1933.16666666667",
+                     "X1938.41666666667", "X1945.75", "X1949.75", "X1954.33333333333",
+                     "X1958.25", "X1961.08333333333", "X1970.83333333333", "X1975.16666666667",
+                     "X1980.5", "X1982.83333333333", "X1991.16666666667", "X1857.41666666667",
+                     "X1860.75", "X1865.25", "X1869.41666666667", "X1873.75", "X1882.16666666667",
+                     "X1887.16666666667", "X1890.5", "X1893", "X1895.91666666667",
[…]

### 2024-02-12 R_local_ic/0_Construction_mm.R

@@ -1,0 +1,78 @@
+library(rjd3filters)
+library(ggplot2)
+library(patchwork)
+library(AQLThesis)
+h <- 6
+q <- 6
+X_gen <- function(d = 1, p = 6, q = p){
+  sapply(0:d, function(exp) seq(-p, q)^exp)
+}
+gen_MM <- function(p=6, q=p, d=2){
+  k = rjd3filters::get_kernel("Henderson", h = h)
+  k = c(rev(k$coef[-1]), k$coef[seq(0,q)+1])
+  K = diag(k)
+  X = X_gen(d=d, p = h, q = q)
+  e1 = e2 = e3 = matrix(0, ncol = 1, nrow = d+1)
+  e1[1] = 1
+  e2[2] = 1
+  e3[3] = 1
+  M1 = K %*% X %*% solve(t(X) %*% K %*% X, e1)
+  M2 = K %*% X %*% solve(t(X) %*% K %*% X, e2)
+  M3 = K %*% X %*% solve(t(X) %*% K %*% X, e3)
+  cbind(M1, M2, M3)
+}
+d2 = gen_MM(d=2)
+d3 = gen_MM(d=3)
+ylim_m2 = range(c(d2[,2], d2[,3]))
+plot_coef_ggplot <- function(data){
+  data$date <- factor(rownames(data), levels = rownames(data),ordered = TRUE)
+  dataGraph <- reshape2::melt(data)
+
+  ggplot(data = dataGraph, aes(x = date, y = value, group = variable,
+                               colour = variable)) +
+    geom_line(linewidth = 0.7) +
+    geom_point(size = 1) +
+    theme(panel.background = element_rect(fill = "white", colour = NA),
+          panel.border = element_rect(fill = NA, colour = "grey20"),
+          panel.grid.major = element_line(colour = "grey92"),
+          panel.grid.minor = element_line(colour = "grey92",
+                                          linewidth = 0.25),
+          strip.background = element_rect(fill = "grey85", colour = "grey20"),
+          complete = TRUE, plot.title = element_text(hjust = 0.5),
+          legend.title=element_blank()) +
+    labs(x = NULL, y = "Coefficients")
+}
+
+MM_D3 = data.frame(gen_MM(d=3,q=6)[,2], gen_MM(d=3,q=6)[,3])
[…]

### 2024-02-12 R_simul/1B_plot_data.R

@@ -1,0 +1,34 @@
+library(AQLThesis)
+load_simul <- function(x){
+  data <- readRDS(sprintf("data_simul/byseries/%s.RDS", x))
+  data <- data[[length(data)]]
+  data
+}
+data <- do.call(cbind, lapply(sprintf("%svariability%i", rep(c("low", "medium", "high"), each = 3),
+                                      rep(1:3, 3)),load_simul))
+data <- data[,seq(from = 2, by = 3, length.out = 3)]
+colnames(data) <- paste(c("Low", "Medium", "High"), "variability")
+dataGraph <- data.frame(cbind(time(data), data))
+colnames(dataGraph) <- c("date", colnames(data))
+dataGraph <- reshape2::melt(dataGraph, id="date")
+p <- ggplot(data = dataGraph,
+            aes(x = date, y = value, group = variable, colour = variable)) +
+  geom_line(linewidth=0.7, data = dataGraph) +
+  labs(title = NULL,
+       x = NULL, y = NULL) +
+  scale_x_continuous(breaks = scales::pretty_breaks(n = 10),
+  ) +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 12),
+                     labels = function(x) format(x, decimal.mark = ",")) +
+  theme_bw() +theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
+                                                     colour = "gray80", linetype = "solid"),
+                    legend.justification = c(0,0),
+                    legend.position = c(0,0),
+                    legend.key = element_blank(),
+                    legend.title = element_blank())
+
+ggMultisave("img/simulations/simul_data",
+            plot = p,
+            width = 8, height = 4)
+
+

### 2024-02-12 R_simul/4B_Analyse_tp.R

@@ -1,0 +1,83 @@
+# Pour que ce programme puisse tourner, il faut également avoir lancé ceux sous R_local_ic
+
+source("R_simul/4_utils.R",encoding = "UTF-8")
+library(ggplot2)
+library(scales)
+library(latex2exp)
+
+tp_lp <- merge(readRDS("results_simul/compile_tp_norev/troughs_lp.RDS"),
+                   readRDS("results_simul/compile_tp_norev/peaks_lp.RDS"),
+                   by=c("series","kernel", "method")) %>%
+  select_var()
+
+tp_lic_final <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic_final.RDS"),
+                          readRDS("results_simul/compile_tp_norev/peaks_localic_final.RDS"),
+                          by=c("series", "kernel", "h", "degree", "method"))  %>%
+  dplyr::filter(degree == "d2", h == "h6") %>%
+  select_var() %>% mutate(method = sprintf("%s_localic_final", method)) %>%
+  select(!c(degree, h))
+tp_lic_daf_trunc <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic_daf_trunc.RDS"),
+                              readRDS("results_simul/compile_tp_norev/peaks_localic_daf_trunc.RDS"),
+                              by=c("series", "kernel", "h", "degree", "method")) %>%
+  dplyr::filter(degree == "d2", h == "h6") %>%
+  select_var() %>% mutate(method = sprintf("%s_localic", method)) %>%
+  select(!c(degree, h))
+
+tp_arima <-
+  merge(readRDS("results_simul/compile_tp_norev/troughs_arima.RDS"),
+        readRDS("results_simul/compile_tp_norev/peaks_arima.RDS"),
+        by=c("series","kernel", "method")) %>%
+  select_var()
+fst_weights <- c("weight235")
+
+all_tp <- rbind(tp_lp,
+                tp_lic_final,
+                tp_lic_daf_trunc,
+                tp_arima) %>%
+  mutate(method = factor(method,c("lc", "lc_localic_final", "lc_localic",
+                                  "ql", "ql_localic_final", "ql_localic",
+                                  "cq","daf", "auto_arima"),
+                         ordered = TRUE),
+         variability = factor(variability,
+                              levels = c("lowvariability","mediumvariability","highvariability"),
+                              ordered = TRUE),
+         kernel = tolower(kernel))
+
+
[…]

### 2024-02-12 R_simul/4C_Analyse_tp_by_kernel.R

@@ -1,0 +1,86 @@
+source("R_simul/4_utils.R",encoding = "UTF-8")
+library(ggplot2)
+library(scales)
+library(latex2exp)
+
+tp_lp <- merge(readRDS("results_simul/compile_tp_norev/troughs_lp.RDS"),
+                   readRDS("results_simul/compile_tp_norev/peaks_lp.RDS"),
+                   by=c("series","kernel", "method")) %>%
+  select_var() %>% 
+  mutate(method2 = sprintf("%s_%s", kernel, method))
+tp_lp$method %>% unique()
+apply(!is.na(tp_lp %>% dplyr::filter(!kernel %in% c("uniform", "gaussian"))),2, all)
+tp_lp %>% 
+  group_by(kernel)
+apply(!is.na(tp_lp %>% dplyr::filter(method == "lc")),2, all)
+apply(!is.na(tp_lp %>% dplyr::filter(method == "QL")),2, all)
+apply(!is.na(tp_lp %>% dplyr::filter(method == "CQ")),2, all)
+apply(!is.na(tp_lp %>% dplyr::filter(method == "DAF")),2, all)
+
+# Graphique sur le dephasage
+format_table_tp <- function(x){
+  x %>% 
+    tidyr::pivot_longer(
+      cols = starts_with("x"), 
+      names_to = "name",
+      values_to = "value"
+    ) %>% 
+    unique_series_pivot() %>% 
+    mutate(variability = recode(variability,
+                                lowvariability = "Faible variabilité",
+                                mediumvariability = "Variabilité moyenne",
+                                highvariability = "Forte variabilité")) %>% 
+  na.omit()
+}
+data_tp <- tp_lp %>% 
+  format_table_tp()
+all_p <- lapply(c("lc", "ql", "cq", "daf"), function(m) {
+  ggplot(data_tp %>% 
+           dplyr::filter(method == m),aes(x=kernel, y = value))+ 
+    geom_boxplot() +
+    facet_wrap(vars(variability), ncol = 1) + theme_bw() +
+    labs(y="Déphasage", x = NULL,
+         title = m) + 
+    theme(axis.text.x = element_text( margin = margin(10, 0, 0, 0), vjust = 1))
+})
+names(all_p) <- c("lc", "ql", "cq", "daf")
+
+all_p$lc
+all_p$ql
+all_p$cq
+all_p$daf
+# ggMultisave("DT/img/simulations/phase_shift_simul", 
+#             plot = p,
+#             width = 10, height = 6)
+
[…]

### 2024-02-12 R_simul/5_analyse_revisions.R

@@ -1,0 +1,86 @@
+# Pour que ce programme puisse tourner, il faut également avoir lancé ceux sous R_local_ic
+
+source("R_simul/4_utils.R",encoding = "UTF-8")
+library(ggplot2)
+library(scales)
+
+all_methods <- c("lp", "arima", "localic_daf_trunc", "localic_final")
+
+for (crit in c("ce", "fe")) {
+  for (method in all_methods) {
+    data <- readRDS(sprintf("results_simul/compile_revisions/%s_%s_rev.RDS", method, crit)) %>%
+      select_series() %>%
+      select_mae()
+    if (method == "lp")
+      data <- data %>% dplyr::filter(kernel == "henderson")
+    if (method %in% c("localic_daf_trunc", "localic_final")){
+      suff <- ifelse(method == "localic_final", "_final", "")
+      data <- data  %>% dplyr::filter(degree == "d2", h == "h6") %>%
+        mutate(method = sprintf("%s_localic%s", method,
+                                suff)) %>%
+        select(!c(degree, h))
+    }
+    data <- data %>% select(!kernel)
+    assign(sprintf("rev_%s_%s", crit, method), data)
+  }
+  all_data <- do.call(rbind, mget(sprintf("rev_%s_%s", crit, all_methods))) %>%
+    mutate(method = factor(method,c("lc", "lc_localic_final", "lc_localic",
+                                    "ql", "ql_localic_final", "ql_localic",
+                                    "cq","daf", "auto_arima"),
+                           ordered = TRUE),
+           variability = factor(variability,
+                                levels = c("lowvariability","mediumvariability","highvariability"),
+                                ordered = TRUE))
+  assign(sprintf("all_rev_%s", crit), all_data)
+}
+
+x = all_rev_fe
+normalise_rev <- function(x, ref = "lc", suff = "^(rev|X)"){
+  ref = x[(x$method == "lc"),grep(suff,colnames(x)) ]
+  for(m in unique(x$method)){
+    if(nrow(x[x$method == m,grep(suff,colnames(x))]) > 0){
+      x[x$method == m,grep(suff,colnames(x))] <-
+        x[x$method == m,grep(suff,colnames(x))] / ref
+    }
+  }
+  x
+}
+summarise_ref <- function(x, normalise = FALSE){
+  if(normalise){
+    x = x %>% normalise_rev()
+    digits = 1
+  } else{
+    digits = 2
+  }
+  x %>%
[…]

### 2024-02-12 ex_pente.R

@@ -1,0 +1,110 @@
+# # To install rjd3filters
+# remotes::install_github("rjdemetra/rjd3toolkit")
+# remotes::install_github("rjdemetra/rjd3x11plus")
+# remotes::install_github("rjdemetra/rjd3filters")
+library(rjd3filters)
+library(ggplot2)
+library(patchwork)
+library(zoo)
+library(forecast)
+y <- readRDS("data_simul/byseries/mediumvariability2.RDS")
+
+# Local estimates of IC-ratios
+# We replicate the direct estimates to have
+# estimators of the slope and the concavity
+gen_MM <- function(p=6, q=p, d=2){
+  X_gen <- function(d = 1, p = 6, q = p){
+    sapply(0:d, function(exp) seq(-p, q)^exp)
+  }
+  k = rjd3filters::get_kernel("Henderson", h = p)
+  k = c(rev(k$coef[-1]), k$coef[seq(0,q)+1])
+  K = diag(k)
+  X = X_gen(d=d, p = p, q = q)
+  e1 = e2 = e3 = matrix(0, ncol = 1, nrow = d+1)
+  e1[1] = 1
+  e2[2] = 1
+  e3[3] = 1
+  # Estimator of the constant
+  M1 = K %*% X %*% solve(t(X) %*% K %*% X, e1)
+  # Estimator of the slope
+  M2 = K %*% X %*% solve(t(X) %*% K %*% X, e2)
+  # Estimor of the concavity
+  M3 = K %*% X %*% solve(t(X) %*% K %*% X, e3)
+  mm <- list(const = M1, slope = M2, concav = M3)
+  lapply(mm, moving_average, lags = -p)
+}
+henderson_f = lp_filter()@sfilter
+all_mm <- lapply(6:0, gen_MM, p = 6, d = 2)
+est_slope <- finite_filters(all_mm[[1]]$slope,
+                            lapply(all_mm[-1], `[[`, "slope"))
+
+var_est <- sapply(y, var_estimator, coef = henderson_f)
+slope_est <- y[[length(y)]] * est_slope@sfilter
+slope_est2 <- y[[length(y)]] * est_slope[,"q=2"]
+
+var_est <- ts(var_est, end = end(y[[length(y)]]),
+              frequency = 12)
+slope_est <- ts(slope_est, end = end(y[[length(y)]]),
+                frequency = 12)
+slope_est2 <- ts(slope_est2, end = end(y[[length(y)]]),
[…]

### 2024-02-12 img/Ex_climat_cl4.R

@@ -1,0 +1,317 @@
+# id <- c(manuf ="001585934",
+#         C1 = "001786503",
+#         C3 = "001786504",
+#         C4 = "001786505",
+#         C5 = "001786506"
+#         )
+library(insee)
+library(rjd3filters)
+library(ggplot2)
+library(patchwork)
+library(zoo)
+library(forecast)
+library(AQLThesis)
+# Import des données
+y <- data.frame(insee::get_insee_idbank("001786505"))
+y <- y[order(y[,"TIME_PERIOD"]), ]
+first_date <- as.numeric(strsplit(y$TIME_PERIOD[1], "-")[[1]])
+y <- ts(y$OBS_VALUE, start = first_date, frequency = 12)
+y <- window(y, start = 2010)
+last_dates <- c(tail(time(y), 7))
+names(last_dates) <- as.character(zoo::as.yearmon(last_dates))
+der_est <- lapply(last_dates, function(x) window(y, end = x))
+
+# MM de longueur 13 adaptées : 
+select_trend_filter(y)
+# Estimation finale
+tc_f <- henderson(y, length = 13, musgrave = FALSE) # Estimation finale
+plot(y)
+lines(tc_f, col = "red")
+
+p <- forecast::autoplot(ts.union(y, tc_f)) +
+  theme_bw() +
+  theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
+                                         colour = "gray80", linetype = "solid"),
+        legend.justification = c(0,0),
+        legend.position = c(0,0),
+        legend.key = element_blank(),
+        legend.title = element_blank()) +
+  labs(x = NULL, y = NULL)  +
+  scale_color_discrete(labels =c("Série initiale","Série lissée"))
+p
+ggMultisave("DT/img/ex/tc_finale", 
+            plot = p,
+            width = 8, height = 3)
+
+## Méthode polynomiale
+
+# Calcul des IC-ratios : 
+icr <- sapply(der_est, function(x) {
+  ic_ratio(x, henderson(x, length = 13, musgrave = FALSE))
+})
+lp_est <- lapply(c("LC", "QL", "CQ", "DAF"), function(method) {
+  res <- lapply(seq_along(icr), function(i) {
+    lp_coef <- lp_filter(horizon = 6,
+                         kernel = "Henderson",
+                         endpoints = method,
[…]

### 2024-02-12 article_jos.qmd
numbered section
R logo removed

@@
 
-# Introduction {-}
+# Introduction
 
@@
 The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
-The extensions have been implemented in the {{< fa brands r-project >}} `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.
+The extensions have been implemented in the R `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.
 
@@
 Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
-This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::fst_filter()`.
+This method can be used inR with the function `rjd3filters::fst_filter()`.
 
 @trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
-A simplified version of this method can be used in {{< fa brands r-project >}} with the function `rjd3filters::dfa_filter()`.
+A simplified version of this method can be used in R with the function `rjd3filters::dfa_filter()`.
 
@@
 Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
-This method can be used in{{< fa brands r-project >}} with the function `rjd3filters::rkhs_filter()`.
+This method can be used inR with the function `rjd3filters::rkhs_filter()`.
 
@@
 This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
-This extension is implemented in the {{< fa brands r-project >}} function `rjd3filters::lp_filter()`.
+This extension is implemented in the R function `rjd3filters::lp_filter()`.
 
@@
 
-# Conclusion {.unnumbered}
+# Conclusion
 
@@
 
-This paper presents the {{< fa brands r-project >}} package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
+This paper presents the R package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
 It also offers various functions for the theoretical analysis of moving averages (gain functions, phase shift, quality criteria, etc.) and judging the quality of the latest estimates (for example with implicit forecasts).
@@
 
-In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.
+In this appendix we present the R code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.
 

### 2024-02-12 article_jos.qmd
transpose vector in inline text

@@
 
-Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}$ such as for all time series $X_t$:
+Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $X_t$:
 $$
@@
 We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
-With $\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}$, the estimate of the trend is:
+With $\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}$, the estimate of the trend is:
 $$
@@
 The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
-These constraints are represented by the matrix $\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
+These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
 The problem is equivalent to finding $v$ which minimises:
@@
 
-When $\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}$, $\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
+When $\boldsymbol U=\transp{\begin{pmatrix}1&\cdots&1\end{pmatrix}}$, $\boldsymbol Z=\transp{\begin{pmatrix}-h&\cdots&+h\end{pmatrix}}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
 This filter assumes that, for real-time estimation, the data is generated by a linear process and that the asymmetric filters preserve the constants ($\sum v_i=\sum \theta_i=1$).
@@
 Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
-With $\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
+With $\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
 $$

### 2024-02-12 article_jos.qmd
(a)symmetrical -> (a)symmetric

@@
 
-2. Apply symmetrical filters to extended forecast series.
+2. Apply symmetric filters to extended forecast series.
 This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period:\
@@
 This allows to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
-Let us denote $v=(v_{-h},\dots, v_{h})$ the reference symmetrical moving average and $w^0,\dots w^{h-1}$ a sequence of asymmetrical moving averages, of order $h+1$ to $2h$ used to estimate the last $h$ points with, for convention, $w_t^q=0$ for $t>q$.
+Let us denote $v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average and $w^0,\dots w^{h-1}$ a sequence of asymmetric moving averages, of order $h+1$ to $2h$ used to estimate the last $h$ points with, for convention, $w_t^q=0$ for $t>q$.
 This means that $w^0=(w_{-h}^0,\dots, w_{0}^0)$ is used for estimation in real time (when no points are known in the future), $w^1=(w_{-h}^1,\dots, w_{1}^1)$ for estimation of the penultimate point (when only one point is known in the future), and so on.
@@
 In practice, this means assuming that the trend is locally polynomial. 
-Various estimation methods can be used to derive symmetrical and asymmetrical moving averages.  
+Various estimation methods can be used to derive symmetric and asymmetric moving averages.  
 @GrayThomson1996 propose a complete statistical framework which makes it possible, in particular, to model the error in approximating the trend by local polynomials.
@@
 Table\ \ref{tab-criteriaLp} compares the quality criteria of the different methods using the Henderson filter and $h=6$ (a symmetric filter with 13 terms). 
-For real-time filters ($q=0$), a more complex asymmetrical filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance.
+For real-time filters ($q=0$), a more complex asymmetric filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance.
 This result varies as $q$ increases: for $q=2$, the QL filter has greater timeliness than the LC filter. 
@@
 However, we might prefer a local parameterisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
-Indeed, although the overall parameterisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetrical filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
+Indeed, although the overall parameterisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetric filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
 For example, with the LC method, global parameterisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up to the turning point.
@@
 
-- The variance $\sigma^2$ can be estimated using the observed data set and the symmetrical filter $(w_{-p},\dots,w_p)$:
+- The variance $\sigma^2$ can be estimated using the observed data set and the symmetric filter $(w_{-p},\dots,w_p)$:
 $$
@@
 
-For all series, a symmetrical 13-term filter is used.
-These methods are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetrical Henderson filter.
+For all series, a symmetric 13-term filter is used.
+These methods are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
 The ARIMA model is determined automatically, using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
@@
 This point and this series were chosen because the economic turnaround is particularly visible in the raw series (@fig-ce16ov-previmp-lp), because this series is used for dating economic cycles and because the FRED-MD database facilitates the reproducibility of the results (thanks to the availability of series published on past dates).
-Studying the series up to January 2020, this series has a medium variability: a symmetrical filter of 13 terms is therefore appropriate.
+Studying the series up to January 2020, this series has a medium variability: a symmetric filter of 13 terms is therefore appropriate.
 Appendix [-@ann-ex-r] shows how the code to reproduce the different plots.
@@
 The quality of the intermediate estimates can also be analysed using the implicit forecasts of the different methods (@fig-ce16ov-previmp-lp). 
-As a reminder, these are forecasts of the raw series which, by applying the symmetrical Henderson filter to the extended series, give the same estimates as the asymmetrical moving averages. 
+As a reminder, these are forecasts of the raw series which, by applying the symmetric Henderson filter to the extended series, give the same estimates as the asymmetric moving averages. 
 The forecasts of the ARIMA model are naive and do not take the turning point into account, unlike the other methods.

### 2024-02-12 article_jos.qmd
parameterisation -> parametrisation

@@
 This paper describes how they can be extended to include a timeliness criterion to minimise the phase shift.
-While asymmetric filters are generally parameterised globally, which can be suboptimal around turning points, this paper proposes a procedure for parameterising them locally.
-An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parameterisation reduces the delay in detecting turning points and reduces revisions.
+While asymmetric filters are generally parametrised globally, which can be suboptimal around turning points, this paper proposes a procedure for parametrising them locally.
+An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
 The paper also shows how these results can be easily reproduced using the R package `rjd3filters` dedicated to the manipulation of moving averages.
@@
 Revisions are more important for the CQ and DAF filters, which leads to greater variability in estimations and a longer delay in detecting turning points.
-The globally parameterised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter.
+The globally parametrised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter.
 Indeed, to identify a turning point at date $t$, it is necessary to have knowledge of at least two points after the given date to confirm the reversal in trend.
@@
 That's why in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
-However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parameterisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).
+However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parametrisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).
 
-## Local parameterisation of asymmetric filters {#sec-localic}
+## Local parametrisation of asymmetric filters {#sec-localic}
 
-Asymmetric filters are usually parameterised globally: $\lvert\delta/\sigma\rvert$ estimated on all the data using the IC-ratio or a cross-validation criterion. 
-However, we might prefer a local parameterisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
-Indeed, although the overall parameterisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetric filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
-For example, with the LC method, global parameterisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up to the turning point.
+Asymmetric filters are usually parametrised globally: $\lvert\delta/\sigma\rvert$ estimated on all the data using the IC-ratio or a cross-validation criterion. 
+However, we might prefer a local parametrisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
+Indeed, although the overall parametrisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetric filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
+For example, with the LC method, global parametrisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up to the turning point.
 
-This is what is proposed in this article, with a local parameterisation of the asymmetric filters by estimating $\delta$ and $\sigma^2$ separately.
+This is what is proposed in this article, with a local parametrisation of the asymmetric filters by estimating $\delta$ and $\sigma^2$ separately.
 Although this does not give an unbiased estimator of the ratio $\lvert\delta/\sigma\rvert$, it does allow the main evolutions to be captured, such as the decay towards 0 before a turning point and the growth after the turning point for the LC method:
@@
 
-In the empirical applications of @sec-comparison, the final local parameterisation is obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
-@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two the local parameterisation:
+In the empirical applications of @sec-comparison, the final local parametrisation is obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
+@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two the local parametrisation:
 
-- @fig-mmpenteconcac-ex-1: using a symmetric filter to estimate $\delta_1$ (final local parameterisation).
+- @fig-mmpenteconcac-ex-1: using a symmetric filter to estimate $\delta_1$ (final local parametrisation).
 
@@
 
-Excluding for the moment the local parameterisations of the polynomial filters, the linear-constant polynomial filter (LC) seems to give the best results in terms of delay in the detection of turning points (@fig-graphstpsimul).
+Excluding for the moment the local parametrisations of the polynomial filters, the linear-constant polynomial filter (LC) seems to give the best results in terms of delay in the detection of turning points (@fig-graphstpsimul).
 Performance is relatively close to that obtained by extending the series using an ARIMA model.
@@
 
-For series with moderate variability, local parameterisation of the LC and QL filters reduces the phase shift.
+For series with moderate variability, local parametrisation of the LC and QL filters reduces the phase shift.
 For series with high variability, the phase shift is only reduced by using the final $\hat\delta$ parameters: real-time estimates seem to add more variance.
@@
 In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability (Table\ \ref{tab-simulrev}).
-In general, LC filters always minimise revisions (with relatively small effect from the local parameterisation of the filters) and revisions are greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
+In general, LC filters always minimise revisions (with relatively small effect from the local parametrisation of the filters) and revisions are greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
 
 For the QL filter, there is a large revision between the second and third estimates: this may be due to the fact that for the second estimate (when one point in the future is known), the QL filter assigns a greater weight to the estimate in $t+1$ than to the estimate in $t$, which creates a discontinuity.
-This revision is considerably reduced by parameterising the filter locally.
+This revision is considerably reduced by parametrising the filter locally.
 For polynomial filters other than the LC filter, the large revisions at the first estimate were to be expected given the coefficient curve: a very large weight is associated with the current observation and there is a strong discontinuity between the moving average used for real-time estimation (when no point in the future is known) and the other moving averages.
@@
 
-In this example, local parameterisation does not reduce the phase shift, but it does reduce revisions.
+In this example, local parametrisation does not reduce the phase shift, but it does reduce revisions.
 The QL and DAF polynomials lead to greater variability in the intermediate estimates, especially in February 2001.
@@
 The forecasts of the ARIMA model are naive and do not take the turning point into account, unlike the other methods.
-Finally, local parameterisation of the QL filter produces much more consistent forecasts.
+Finally, local parametrisation of the QL filter produces much more consistent forecasts.
 
@@
 At the end of the period, for estimating the trend-cycle, we can therefore restrict ourselves to methods modelling polynomial trends of degree two or less (linear-constant, LC, and quadratic-linear, QL).
-In addition, parameterising polynomial filters locally as suggested in this paper enables turning points to be detected more quickly (especially for the QL filter).
-Even when the phase shift is not reduced, local parameterisation is recommended because it reduces revisions and produces intermediate estimates that are more consistent with expected future trends.
+In addition, parametrising polynomial filters locally as suggested in this paper enables turning points to be detected more quickly (especially for the QL filter).
+Even when the phase shift is not reduced, local parametrisation is recommended because it reduces revisions and produces intermediate estimates that are more consistent with expected future trends.
 However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too short (i.e.

### 2024-02-12 article_jos.qmd
remove local in title

@@
 ---
-title: "Improving real-time trend estimates using local parametrisation of local polynomial regression filters"
+title: "Improving real-time trend estimates using local parametrisation of polynomial regression filters"
 format:

### 2024-02-12 article_jos.qmd
trend cycle -> trend-cycle

@@
 :::: {.abstract}
-This paper studies and compares real-time estimates of the trend cycle component with moving averages based on local polynomial regression.
+This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
 In particular, this theory allows to reproduce Henderson's symmetric and Musgrave's asymmetric filters.

### 2024-02-12 article_jos.qmd
typos

@@
 This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
-In particular, this theory allows to reproduce Henderson's symmetric and Musgrave's asymmetric filters.
+In particular, this theory allows us to reproduce Henderson's symmetric and Musgrave's asymmetric filters.
 This paper describes how they can be extended to include a timeliness criterion to minimise the phase shift.
@@
 
-A moving average, or linear filter, is a statistical method that consists in applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
+A moving average, or linear filter, is a statistical method that consists of applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
 Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
@@
 
-Using local polynomial regression, @proietti2008 developed general class of symmetric and asymmetric moving averages.
-In particular, it makes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
+Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
+In particular, it becomes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
 However, these methods have two drawbacks.
@@
 
-# Some properties on moving averages {#sec-propMM}
+# Some properties of moving averages {#sec-propMM}
 
-Lots of papers describe the definition and the properties of moving averages and linear filters, see for example @ch12HBSA. 
+Numerous papers describe the definition and the properties of moving averages and linear filters, see for example @ch12HBSA. 
 In this section we summarize some of the main results to understand the next sections.
 
-Let $p$ et $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $X_t$:
+Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $X_t$:
 $$
@@
 The function  $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
-The frequency response function can equivalently be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ ou $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
+The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
 It can be rewritten as:
@@
 where  $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
-This function is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to mesure the phase shift in number of periods.
-For all symmetric moving average we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.
+This function is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure the phase shift in number of periods.
+For all symmetric moving averages we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.
 
@@
 
-Fourier decomposition allows to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
+Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
 Noting $\omega = 2\pi/p$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $p$ periods.
 For example, for a monthly series (12 observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
-In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12[$, i.e. movements recurring at least every 12 months.
+In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e. movements recurring at least every 12 months.
 Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
@@
 The gain function is greater than 1 on the frequencies associated with the trend-cycle ($[0, 2\pi/12]$): this means that the trend-cycle is well preserved and that short cycles of 1 to 2 years ($[2\pi/24,2\pi/12]$) are even amplified.
-However, cycles of 8 to 12 months ($[2\pi/12, 2\pi/8]$), considered undesirable because associated with the irregular, are also amplified: this can lead to the detection of false turning points.
 At other frequencies, the gain function is less than 1, but always positive: this means that the series smoothed by this moving average will always contain noise, even if it is attenuated.
@@
 Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg).
-In order to best preserve the trend-cycles, we try to have moving averages that preserve the polynomial trends.
+In order to best preserve the trend-cycles, we try to design moving averages that preserve the polynomial trends.
 A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t:M_{\boldsymbol\theta} f(t)=f(t)$.
@@
 \sum_{k=-p}^{+f}\theta_k=1
-\quad\text{et}\quad
+\quad\text{and}\quad
 \sum_{k=-p}^{+f}k\theta_k=0.
@@
 
-Indeed, let $\{\varepsilon_t\}$ a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$:
+Indeed, let $\{\varepsilon_t\}$ be a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$:
 $$
@@
 However, they cannot be used in the beginning and in the end of the time series because no past or future value can be used. 
-Thus, for real-time estimation, it is needed to build asymmetric moving average that approximate the symmetric moving average.
+Thus, for real-time estimation, it is needed to build asymmetric moving averages that approximate the symmetric moving average.
 
@@
 *The important point is that the m terms thus added, taken together with the $m+1$ adjacent given terms, should follow a curve whose form is approximately algebraic and of a degree not higher than the third.*"\
-This is also the approach used in the X-13ARIMA seasonal adjustment method, which extends which extends the series by one year with an ARIMA model.
+This is also the approach used in the X-13ARIMA seasonal adjustment method, which extends the series by one year with an ARIMA model.
 
@@
 Conversely, the implicit forecasts of an asymmetric moving average can be deduced from a reference symmetric moving average. 
-This allows to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
+This allows us to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
 Let us denote $v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average and $w^0,\dots w^{h-1}$ a sequence of asymmetric moving averages, of order $h+1$ to $2h$ used to estimate the last $h$ points with, for convention, $w_t^q=0$ for $t>q$.
@@
 $$
-\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
+\underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
 =\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
@@
 $$
-\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
+\quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
 =\sum_{i=-h}^0 (w_i^q-v_i)y_i.
@@
 Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
-This method can be used inR with the function `rjd3filters::fst_filter()`.
+This method can be used in R with the function `rjd3filters::fst_filter()`.
 
@@
 \begin{table}[!h]
-\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
+\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.55$.}
 {\centering
@@
 Various estimation methods can be used to derive symmetric and asymmetric moving averages.  
-@GrayThomson1996 propose a complete statistical framework which makes it possible, in particular, to model the error in approximating the trend by local polynomials.
+@GrayThomson1996 propose a complete statistical framework that makes it possible, in particular, to model the error in approximating the trend by local polynomials.
 However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
-Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
-This method can be used inR with the function `rjd3filters::rkhs_filter()`.
+Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies,
+this method can be used in R with the function `rjd3filters::rkhs_filter()`.
 
@@
 
-To estimate $\boldsymbol \beta$ we need $2h+1\geq d+1$ and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
+To estimate $\boldsymbol \beta$ we need at least $d+1$ observations (thus $2h\geq d$) and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
 $$
@@
 $$
-where $\kappa_j$ is a set of weights called *kernel*.
+where $\kappa_j$ is a set of weights called a *kernel*.
 We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
@@
 $$
-Hence, the filter $\boldsymbol w$ preserve deterministic polynomial of order $d$.
+Hence, the filter $\boldsymbol w$ preserves deterministic polynomials of order $d$.
 
@@
 They are also equivalent to the third approach under the same constraints as those of the symmetric filter.
-The third method is called *direct asymmetric filter* (DAF).
+The third method is called *direct asymmetric filtering* (DAF).
 This method is used for real-time estimation in the STL seasonal adjustment method (*Seasonal-Trend decomposition based on Loess*, see @cleveland90).
@@
 The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
-These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the matrix $(h+q+1)\times (d+1)$ which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
+These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the $(h+q+1)\times (d+1)$ matrix which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
 The problem is equivalent to finding $v$ which minimises:
@@
 #| tab-cap: 
-title <- "Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$."
+title <- "Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.5$."
 
@@
 \begin{table}[!h]
-\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
+\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.55$.}
 \centering
@@
 
-One drawback of the previous method is the lack of control over the phase shift. 
+One drawback of the approach of @proietti2008 is the lack of control over the phase shift. 
 However, it is possible to improve the modelling by incorporating in @eq-lppasym the timeliness criterion defined by @ch15HBSA.
@@
 Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
-With $\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}$ and $\boldsymbol\theta_p$ of the same length of $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
+With $\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}$ and $\boldsymbol\theta_p$ of the same length as $\boldsymbol v$, and $\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p$, the timeliness criterion can be expressed as:
 $$
@@
 2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
-&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
-\begin{cases}
-\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
-\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
-\boldsymbol c\text{ a constant independent of }\boldsymbol v
-\end{cases}.
+&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c,
 \end{align*}
+with $\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p,$ $\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'$ and $\boldsymbol c$ a constant independent of $\boldsymbol v.$
+
 Adding the timeliness criterion:
@@
 2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
-\widetilde {\boldsymbol c}\quad\text{with }
-\begin{cases}
-\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
-\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
-\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
-\end{cases},
+\widetilde {\boldsymbol c},
 $$
-where $\alpha_T$ is the weight associated to the timeliness criterion. 
+where $\alpha_T$ is the weight associated with the timeliness criterion, $\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T$, $\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T$ and $\widetilde {\boldsymbol c}$ a constant independent of $\boldsymbol v.$
+
 With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
@@
 It is then difficult to calibrate the value of $\alpha_T$ since the value of $\lvert\delta/\sigma\rvert$ must also be defined.
-That's why in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
+Hence, in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
 However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parametrisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).
@@
 In the empirical applications of @sec-comparison, the final local parametrisation is obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
-@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two the local parametrisation:
+@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two local parametrisations:
 
@@
 
-- @fig-mmpenteconcac-ex-2: using the asymmetric filter which need two points in the futur to estimate $\delta_1$.
+- @fig-mmpenteconcac-ex-2: using the asymmetric filter which need two points in the future to estimate $\delta_1$.
 This is closed to the real-time estimates since two points in the future are needed to detect a turning point.
@@
 #|   - "Symmetric moving average used for $\\hat \\delta_1$."
-#|   - "Asymmetric moving average using two futur points ($q=2$) used for $\\hat \\delta_1$."
+#|   - "Asymmetric moving average using two future points ($q=2$) used for $\\hat \\delta_1$."
 #| layout-ncol: 1
@@
 The different methods are compared on simulated and real data. 
-
 For all series, a symmetric 13-term filter is used.
@@
 #| label: fig-graphs-data-simul
-#| fig-cap: "Simulated series with low ($\\sigma_e^2=0.2$ and $\\rho = 3,5$), medium ($\\sigma_e^2=0.3$ and $\\rho = 2.0$) and high variability ($\\sigma_e^2=0.4$ and $\\rho = 1.0$)."
+#| fig-cap: "Simulated series with low ($\\sigma_e^2=0.2$ and $\\rho = 3.55$), medium ($\\sigma_e^2=0.3$ and $\\rho = 2.0$) and high variability ($\\sigma_e^2=0.4$ and $\\rho = 1.0$)."
 img <- sprintf("img/simulations/simul_data.%s", "pdf")
@@
 2.
-Computation of two revision criteria: the average of the relative deviations between the $q$^th^ estimate and the last estimate $MAE_{fe}(q)$ and the average of the relative deviations between the $q$^th^ estimate and the $q+1$^th^ estimate $MAE_{qe}(q)$.
+Computation of two revision criteria: the average of the relative deviations between the $q$^th^ estimate and the last estimate $MAE_{fe}(q)$ and the average of the relative deviations between the $q$^th^ estimate and the $q+1$^th^ estimate $MAE_{ce}(q)$.
 $$
@@
 \quad\text{and}\quad
-MAE_{qe}(q)=\mathbb E\left[
+MAE_{ce}(q)=\mathbb E\left[
 \left|\frac{
@@
 Studying the series up to January 2020, this series has a medium variability: a symmetric filter of 13 terms is therefore appropriate.
-Appendix [-@ann-ex-r] shows how the code to reproduce the different plots.
+Appendix [-@sec-ann-ex-r] shows how the code to reproduce the different plots.
 
@@
 
-This paper presents the R package `rjd3filters`, which implements a several methods to build moving average for real-time trend-cycle estimation.
+This paper presents the R package `rjd3filters`, which implements several methods to build moving averages for real-time trend-cycle estimation.
 It also offers various functions for the theoretical analysis of moving averages (gain functions, phase shift, quality criteria, etc.) and judging the quality of the latest estimates (for example with implicit forecasts).
@@
 This study could be extended in many ways.
-
 One possible extension would be to look at the impact of filter length on the detection of turning points.
@@
 
-# R code example {#ann-ex-r}
+# R code example {#sec-ann-ex-r}
 

### 2024-02-12 article_jos.qmd
quote removed

@@
 2. Apply symmetric filters to extended forecast series.
-This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period:\
-"*As the first $m$ and last $m$ terms of the series cannot be reached directly by the formula, the series should be graphically extended by m terms at both ends, first plotting the observations on paper as ordinates, and then extending the curve along what seems to be its probable course, and measuring the ordinates of the extended portions.*
-*It is not necessary that this extension should coincide with what would be the true course of the curve in those parts. *
-*The important point is that the m terms thus added, taken together with the $m+1$ adjacent given terms, should follow a curve whose form is approximately algebraic and of a degree not higher than the third.*"\
+This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
 This is also the approach used in the X-13ARIMA seasonal adjustment method, which extends the series by one year with an ARIMA model.

### 2024-02-12 R_filters/slope_concavity_estimator.R
local_ic folder merge with others

@@ -1,7 +1,6 @@
 library(rjd3filters)
 library(ggplot2)
 library(patchwork)
-library(AQLThesis)
 h <- 6
 q <- 6
 X_gen <- function(d = 1, p = 6, q = p){
@@ -26,8 +25,8 @@ d3 = gen_MM(d=3)
 ylim_m2 = range(c(d2[,2], d2[,3]))
 plot_coef_ggplot <- function(data){
   data$date <- factor(rownames(data), levels = rownames(data),ordered = TRUE)
-  dataGraph <- reshape2::melt(data)
-
+  dataGraph <- reshape2::melt(data, id.vars = "date")
+  
   ggplot(data = dataGraph, aes(x = date, y = value, group = variable,
                                colour = variable)) +
     geom_line(linewidth = 0.7) +
@@ -45,7 +44,7 @@ plot_coef_ggplot <- function(data){
 
 MM_D3 = data.frame(gen_MM(d=3,q=6)[,2], gen_MM(d=3,q=6)[,3])
 rownames(MM_D3) <- rjd3filters:::coefficients_names(-6,6)
-colnames(MM_D3) <- c("Est. pente", "Est. concavité")
+colnames(MM_D3) <- c("Slope estimators", "Concavity estimators")
 plot_coef_ggplot(MM_D3)
 
 MM_d3_pente = data.frame(sapply(0:6, function(x) c(gen_MM(d=3,q=x)[,2], rep(NA, 6-x))))
@@ -63,9 +62,6 @@ colnames(MM_d2_pente) <- sprintf("q=%i",0:6)
 MM_d2_deriv2 = data.frame(sapply(0:6, function(x) c(gen_MM(d=2,q=x)[,3], rep(NA, 6-x))))
 rownames(MM_d2_deriv2) <-  rjd3filters:::coefficients_names(-6,6)
 colnames(MM_d2_deriv2) <- sprintf("q=%i",0:6)
-p = (plot_coef_ggplot(MM_d3_pente) + ggtitle("MM utilisées pour estimer la pente") +
-       theme(legend.position="none")
-) + (plot_coef_ggplot(MM_d3_deriv2) + ggtitle("MM utilisées pour estimer la concavité"))
 
 p21 = plot_coef_ggplot(MM_d2_pente) +
   theme(legend.position="none")

### 2024-02-12 R_fredm/1A_Est_pente_finale.R

@@ -1,0 +1,84 @@
+library(AQLThesis)
+library(future)
+library(rjd3filters)
+plan(multisession)
+if(!dir.exists("data_fredm/byseriespente_final_nber_dt"))
+  dir.create("data_fredm/byseriespente_final_nber_dt")
+
+# Dans ce programme, pour paramétrer les méthodes LC et QL :
+# 1. On prend les MM symétriques finales d'estimation de la pente et polynôme degré 2
+#  (dans gen_MM q = p)
+# 2. la variance est estimée à chaque date sur les données connues
+
+X_gen <- function(d = 1, p = 6, q = p){
+  sapply(0:d, function(exp) seq(-p, q)^exp)
+}
+
+gen_MM <- function(p=6, q=p, d=2){
+  k = rjd3filters::get_kernel("Henderson", h = p)
+  k
+  k = c(rev(k$coef[-1]), k$coef[seq(0,q)+1])
+  k
+  K = diag(k)
+  X = X_gen(d=d, p = p, q = q)
+  e1 = e2 = e3 = matrix(0, ncol = 1, nrow = d+1)
+  e1[1] = 1
+  e2[2] = 1
+  e3[3] = 1
+  M1 = K %*% X %*% solve(t(X) %*% K %*% X, e1)
+  M2 = K %*% X %*% solve(t(X) %*% K %*% X, e2)
+  M3 = K %*% X %*% solve(t(X) %*% K %*% X, e3)
+  cbind(M1, M2, M3)
+}
+
+
+MM = lapply(3:6, function(h){
+  list(pente = list(
+    `d=2` = gen_MM(p=h, d=2)[,2],
+    `d=3` = gen_MM(p = h, d=3)[,2]),
+    `deriv2` = list(
+      `d=2` = gen_MM(p=h, d=2)[,3],
+      `d=3` = gen_MM(p = h, d=3)[,3]),
+       henderson = lp_filter(horizon = h)@sfilter
+    )
+})
+names(MM) <- sprintf("h=%i", 3:6)
+
+
+for(s in list.files("data_fredm/byseries",full.names = TRUE)){
+  print(s)
+  for(h in 6){
+    new_f = sprintf("data_fredm/byseriespente_final_nber_dt/%s_h%i.RDS", gsub(".RDS", "",basename(s)),h)
+    print(new_f)
+    MM_h = MM[[sprintf("h=%i", h)]]
+    hend_filter = lp_filter(horizon = h)[,h+1]
[…]

### 2024-02-12 R_fredm/1B_Est_pente_daf.R

@@ -2,14 +2,8 @@ library(AQLThesis)
 library(future)
 library(rjd3filters)
 plan(multisession)
-if(!dir.exists("data_simul/byseriespente_daf"))
-  dir.create("data_simul/byseriespente_daf")
 if(!dir.exists("data_fredm/byseriespente_daf_nber_dt"))
   dir.create("data_fredm/byseriespente_daf_nber_dt")
-if(!dir.exists("data_simul/byseries")) # création des données si elles n'existent pas
-  source("R_simul/1A_data_creation.R")
-if(!dir.exists("data_fredm/byseries")) # création des données si elles n'existent pas
-  source("R_fredm/0_download_files.R")
 
 # Dans ce programme, pour paramétrer les méthodes LC et QL :
 # 1. On prend les MM asymétriques d'estimation de la pente et polynôme degré 2 et on fait une estimation en temps réel
@@ -62,38 +56,6 @@ MM = list(pente = list(
 MM = lapply(MM, function(x){
   lapply(x, finite_filters, first_to_last = TRUE)
 })
-s = list.files("data_simul/byseries",full.names = TRUE)[1]
-d = 2
-h=6
-for(s in list.files("data_simul/byseries",full.names = TRUE)){
-  new_f = sprintf("data_simul/byseriespente_daf/%s.RDS",
-                  gsub(".RDS", "",basename(s)))
-  print(new_f)
-  hend_filter = lp_filter(horizon = h)@sfilter
-  if(!file.exists(new_f)){
-    data <- readRDS(s)
-    info_fs <- lapply(data, function(x){
-      future({
-        sigma2 <- var_estimator(x, hend_filter)
-        list("LC" = list(
-          `d=2` = tail(rjd3filters::filterfilter(x, MM$pente$`d=2`),6),
-          `d=3` = tail(rjd3filters::filterfilter(x, MM$pente$`d=3`),6),
-          `sigma2` = sigma2
-        ),
-        "QL" = list(
-          `d=2` = tail(rjd3filters::filterfilter(x, MM$deriv2$`d=2`),6),
-          `d=3` = tail(rjd3filters::filterfilter(x, MM$deriv2$`d=3`),6),
-          `sigma2` = sigma2
-        )
-        )
-      }
-      )
-    })
-    info <- lapply(info_fs, value)
-    saveRDS(info, new_f)
-  }
-}
-
 
[…]

### 2024-02-12 R_fredm/2A_Est_final.R

@@ -1,0 +1,89 @@
+if(!dir.exists("results_fredm"))
+  dir.create("results_fredm")
+if(!dir.exists("results_fredm/localic_final"))
+  dir.create("results_fredm/localic_final")
+library(rjd3filters)
+library(AQLThesis)
+library(future)
+plan(multisession)
+
+kernel = "Henderson"
+fs <- list()
+j <- 1
+reload <- FALSE
+lp_filter2 <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
+  all_coef = lapply(icr, function(ic){
+    lp_filter(horizon = h,
+              kernel = kernel,
+              endpoints = method,
+              ic = ic)
+  })
+  sym = all_coef[[1]]@sfilter
+  rfilters = lapply(1:h, function(i){
+    q=h -i
+    all_coef[[i]][,sprintf("q=%i", q)]
+  })
+  finite_filters(sym, rfilters = rfilters)
+}
+
+
+fs <- list()
+j <- 1
+h_retain <- 6
+kernel = "Henderson"
+for (method in c("LC","QL")){
+  print(method)
+  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
+    for(d in 2:3){
+      for(h in h_retain) {
+        name_file <- gsub(".RDS$", "", basename(s))
+        print(name_file)
+        data <- readRDS(s)
+        complement = sprintf("_d%i", d)
+        
+        nom_f_s <- sprintf("results_fredm/localic_final/%s_%s_h%i%s.RDS",
+                           name_file, tolower(method), h, complement)
+        nom_f_s_tp <- 
+          sprintf("results_fredm/localic_final/%s_%s_h%i%s_tp.RDS",
+                  name_file,
+                  tolower(method), h, complement)
+        
+        data_info <- readRDS(sprintf("data_fredm/byseriespente_final_nber_dt/%s_h%i.RDS",
+                                     gsub(".RDS", "",basename(s)),h))
+        if(all(file.exists(nom_f_s_tp),
+               file.exists(nom_f_s)))
+          next;
+        
+        reload <- TRUE
+        fs[[j]] <- future({
+          print(s)
+          series_s <- lapply(names(data), function(nom_d){
+            x <- data[[nom_d]]
+            data_t = data_info[[nom_d]][[method]]
+            ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
+            icr = 2/(sqrt(pi) * ratio)
[…]

### 2024-02-12 R_fredm/2B_Est_dafic.R

@@ -1,0 +1,81 @@
+if(!dir.exists("results_fredm"))
+  dir.create("results_fredm")
+if(!dir.exists("results_fredm/localic_daf"))
+  dir.create("results_fredm/localic_daf")
+library(rjd3filters)
+library(AQLThesis)
+library(future)
+plan(multisession)
+
+lp_filter2 <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
+  all_coef = lapply(icr, function(ic){
+    lp_filter(horizon = h,
+              kernel = kernel,
+              endpoints = method,
+              ic = ic)
+  })
+  sym = all_coef[[1]]@sfilter
+  rfilters = lapply(1:h, function(i){
+    q=h -i
+    all_coef[[i]][,sprintf("q=%i", q)]
+  })
+  finite_filters(sym, rfilters = rfilters)
+}
+
+reload <- FALSE
+fs <- list()
+j <- 1
+kernel = "Henderson"
+for (method in c("LC","QL")){
+  print(method)
+  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
+    for(d in 2:3){
+      name_file <- gsub(".RDS$", "", basename(s))
+      print(name_file)
+      data <- readRDS(s)
+      complement = sprintf("_d%i", d)
+      
+      nom_f_s <- sprintf("results_fredm/localic_daf/%s_%s%s.RDS",
+                         name_file, tolower(method), complement)
+      nom_f_s_tp <- 
+        sprintf("results_fredm/localic_daf/%s_%s%s_tp.RDS",
+                name_file,
+                tolower(method), complement)
+      
+      data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber_dt/%s.RDS",
+                                   gsub(".RDS", "",basename(s))))
+      if(all(file.exists(nom_f_s_tp),
+             file.exists(nom_f_s)))
+        next;
+      
+      reload <- TRUE
+      fs[[j]] <- future({
+        print(s)
+        series_s <- lapply(names(data), function(nom_d){
+          x <- data[[nom_d]]
+          data_t = data_info[[nom_d]][[method]]
+          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
+          icr = 2/(sqrt(pi) * ratio)
+          lp_coef = lp_filter2(ic = icr, method = method, h = h, kernel = kernel)
+          filter(x, lp_coef)
+        })
[…]

### 2024-02-12 R_fredm/2C_Est_dafic_trunc.R

@@ -1,0 +1,82 @@
+if(!dir.exists("results_fredm"))
+  dir.create("results_fredm")
+if(!dir.exists("results_fredm/localic_daf_trunc"))
+  dir.create("results_fredm/localic_daf_trunc")
+library(rjd3filters)
+library(AQLThesis)
+library(future)
+plan(multisession)
+
+lp_filter2 <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
+  all_coef = lapply(icr, function(ic){
+    lp_filter(horizon = h,
+              kernel = kernel,
+              endpoints = method,
+              ic = ic)
+  })
+  sym = all_coef[[1]]@sfilter
+  rfilters = lapply(1:h, function(i){
+    q=h -i
+    all_coef[[i]][,sprintf("q=%i", q)]
+  })
+  finite_filters(sym, rfilters = rfilters)
+}
+
+
+fs <- list()
+j <- 1
+kernel = "Henderson"
+for (method in c("LC","QL")){
+  print(method)
+  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
+    for(d in 2:3){
+      name_file <- gsub(".RDS$", "", basename(s))
+      print(name_file)
+      data <- readRDS(s)
+      complement = sprintf("_d%i", d)
+      
+      nom_f_s <- sprintf("results_fredm/localic_daf_trunc/%s_%s%s.RDS",
+                         name_file, tolower(method), complement)
+      nom_f_s_tp <- 
+        sprintf("results_fredm/localic_daf_trunc/%s_%s%s_tp.RDS",
+                name_file,
+                tolower(method), complement)
+    
+      
+      data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber_dt/%s.RDS",
+                                   gsub(".RDS", "",basename(s))))
+      if(all(file.exists(nom_f_s_tp),
+             file.exists(nom_f_s)))
+        next;
+      
+      reload <- TRUE
+      fs[[j]] <- future({
+        print(s)
+        series_s <- lapply(names(data), function(nom_d){
+          x <- data[[nom_d]]
+          data_t = data_info[[nom_d]][[method]]
+          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
+          icr = 2/(sqrt(pi) * ratio)
+          icr[abs(icr) > 12] <- 12
[…]

### 2024-02-12 R_fredm/3_Summary_tp.R

@@ -1,0 +1,74 @@
+library(AQLThesis)
+if(!dir.exists("results_fredm/compile_revisions"))
+  dir.create("results_fredm/compile_revisions")
+if(!dir.exists("results_fredm/compile_tp"))
+  dir.create("results_fredm/compile_tp")
+if(!dir.exists("results_fredm/compile_tp_norev"))
+  dir.create("results_fredm/compile_tp_norev")
+
+for(dir in c("localic_daf", "localic_daf_trunc", "localic_final")){
+  print(dir)
+  all_files <- list.files(sprintf("results_fredm/%s/", dir),pattern = "_tp",full.names = TRUE)
+  all_tp_lp <- lapply(seq_along(all_files), function(i){
+    print(i)
+    f = all_files[i]
+    compute_time_lag(readRDS(f),                    
+                     peaks = nber_tp_m[,"Peak"],
+                     troughs = nber_tp_m[,"Trough"],
+                     n_ahead_max = 12)
+  })
+  
+  full_names <- gsub("_tp.RDS$", "", basename(all_files))
+  split <- strsplit(full_names, "_")
+  series <- sapply(split, `[`, 1)
+  method <- sapply(split, `[`, 2)
+  if(length(grep("localic_daf", dir)) >0){
+    h <- "h6"
+    degree <- sapply(split, `[`, 3)
+  } else {
+    h <- sapply(split, `[`, 3)
+    degree <- sapply(split, `[`, 4)
+  }
+  degree[is.na(degree)] <- "d3"
+  all_t <- data.frame(t(sapply(all_tp_lp, function(x) x[["troughs"]][["phaseshift"]])),
+                      series, kernel = "henderson", method, h, degree)
+  all_p <- data.frame(t(sapply(all_tp_lp, function(x) x[["peaks"]][["phaseshift"]])),
+                      series, kernel = "henderson", method, h, degree)
+  rownames(all_t) <- rownames(all_p) <- full_names
+  
+  saveRDS(all_t, sprintf("results_fredm/compile_tp/troughs_%s.RDS", dir))
+  saveRDS(all_p, sprintf("results_fredm/compile_tp/peaks_%s.RDS", dir))
+  
+  all_tp_lp_localic <- lapply(seq_along(all_files), function(i){
+    print(i)
+    f = all_files[i]
+    compute_time_lag(readRDS(f),                    
+                     peaks = nber_tp_m[,"Peak"],
[…]

### 2024-02-12 R_simul/1B_Est_pente_daf.R

@@ -1,0 +1,89 @@
+library(AQLThesis)
+library(future)
+library(rjd3filters)
+plan(multisession)
+if(!dir.exists("data_simul/byseriespente_daf"))
+  dir.create("data_simul/byseriespente_daf")
+
+# Dans ce programme, pour paramétrer les méthodes LC et QL :
+# 1. On prend les MM asymétriques d'estimation de la pente et polynôme degré 2 et on fait une estimation en temps réel
+# Pour chaque date on a donc h+1=6+1 moyennes mobiles
+# 2. la variance est estimée à chaque date sur les données connues
+
+X_gen <- function(d = 1, p = 6, q = p){
+  sapply(0:d, function(exp) seq(-p, q)^exp)
+}
+
+gen_MM <- function(p=6, q=p, d=2){
+  k = rjd3filters::get_kernel("Henderson", h = p)
+  k
+  k = c(rev(k$coef[-1]), k$coef[seq(0,q)+1])
+  k
+  K = diag(k)
+  X = X_gen(d=d, p = p, q = q)
+  e1 = e2 = e3 = matrix(0, ncol = 1, nrow = d+1)
+  e1[1] = 1
+  e2[2] = 1
+  e3[3] = 1
+  M1 = K %*% X %*% solve(t(X) %*% K %*% X, e1)
+  M2 = K %*% X %*% solve(t(X) %*% K %*% X, e2)
+  M3 = K %*% X %*% solve(t(X) %*% K %*% X, e3)
+  cbind(M1, M2, M3)
+}
+
+MM = lapply(0:6, function(q){
+  d2 = gen_MM(p=6, q=q, d=2)
+  d3 = gen_MM(p=6, q=q, d=3)
+  add_0 <- matrix(0, ncol = ncol(d2), nrow = 6-q)
+  d2 = rbind(d2, add_0)
+  d3 = rbind(d3, add_0)
+  list(pente = list(
+    `d=2` = d2[,2],
+    `d=3` = d3[,2]),
+    `deriv2` = list(
+      `d=2` = d2[,3],
+      `d=3` = d3[,3]))
+})
+MM = list(pente = list(
+  `d=2` = sapply(MM, function(x) x[[1]][[1]]),
+  `d=3` = sapply(MM, function(x) x[[1]][[2]])
+),
+`deriv2` = list(
+  `d=2` = sapply(MM, function(x) x[[2]][[1]]),
[…]

### 2024-02-12 R_simul/2A_Est_final.R

@@ -1,11 +1,7 @@
 if(!dir.exists("results_simul"))
   dir.create("results_simul")
-if(!dir.exists("results_fredm"))
-  dir.create("results_fredm")
 if(!dir.exists("results_simul/localic_final"))
   dir.create("results_simul/localic_final")
-if(!dir.exists("results_fredm/localic_final"))
-  dir.create("results_fredm/localic_final")
 library(rjd3filters)
 library(AQLThesis)
 library(future)
@@ -101,64 +97,3 @@ if(reload){
   j <- 1
 }
 
-
-fs <- list()
-j <- 1
-h_retain <- 6
-kernel = "Henderson"
-for (method in c("LC","QL")){
-  print(method)
-  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
-    for(d in 2:3){
-      for(h in h_retain) {
-        name_file <- gsub(".RDS$", "", basename(s))
-        print(name_file)
-        data <- readRDS(s)
-        complement = sprintf("_d%i", d)
-        
-        nom_f_s <- sprintf("results_fredm/localic_final/%s_%s_h%i%s.RDS",
-                           name_file, tolower(method), h, complement)
-        nom_f_s_tp <- 
-          sprintf("results_fredm/localic_final/%s_%s_h%i%s_tp.RDS",
-                  name_file,
-                  tolower(method), h, complement)
-        
-        data_info <- readRDS(sprintf("data_fredm/byseriespente_final_nber_dt/%s_h%i.RDS",
-                                     gsub(".RDS", "",basename(s)),h))
-        if(all(file.exists(nom_f_s_tp),
-               file.exists(nom_f_s)))
-          next;
-        
-        reload <- TRUE
-        fs[[j]] <- future({
-          print(s)
-          series_s <- lapply(names(data), function(nom_d){
-            x <- data[[nom_d]]
-            data_t = data_info[[nom_d]][[method]]
-            ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
-            icr = 2/(sqrt(pi) * ratio)
-            lp_coef = lp_filter2(ic = icr, method = method, h = h, kernel = kernel)
-            rjd3filters::filterfilter(x, lp_coef)
-          })
-          names(series_s) <- names(data)
-          
-          saveRDS(series_s, nom_f_s)
-          
-          print("turning points")
[…]

### 2024-02-12 R_simul/2B_Est_dafic.R

@@ -1,11 +1,7 @@
 if(!dir.exists("results_simul"))
   dir.create("results_simul")
-if(!dir.exists("results_fredm"))
-  dir.create("results_fredm")
 if(!dir.exists("results_simul/localic_daf"))
   dir.create("results_simul/localic_daf")
-if(!dir.exists("results_fredm/localic_daf"))
-  dir.create("results_fredm/localic_daf")
 library(rjd3filters)
 library(AQLThesis)
 library(future)
@@ -103,61 +99,3 @@ if(reload){
   j <- 1
 }
 
-
-reload <- FALSE
-fs <- list()
-j <- 1
-kernel = "Henderson"
-for (method in c("LC","QL")){
-  print(method)
-  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
-    for(d in 2:3){
-      name_file <- gsub(".RDS$", "", basename(s))
-      print(name_file)
-      data <- readRDS(s)
-      complement = sprintf("_d%i", d)
-      
-      nom_f_s <- sprintf("results_fredm/localic_daf/%s_%s%s.RDS",
-                         name_file, tolower(method), complement)
-      nom_f_s_tp <- 
-        sprintf("results_fredm/localic_daf/%s_%s%s_tp.RDS",
-                name_file,
-                tolower(method), complement)
-      
-      data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber_dt/%s.RDS",
-                                   gsub(".RDS", "",basename(s))))
-      if(all(file.exists(nom_f_s_tp),
-             file.exists(nom_f_s)))
-        next;
-      
-      reload <- TRUE
-      fs[[j]] <- future({
-        print(s)
-        series_s <- lapply(names(data), function(nom_d){
-          x <- data[[nom_d]]
-          data_t = data_info[[nom_d]][[method]]
-          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
-          icr = 2/(sqrt(pi) * ratio)
-          lp_coef = lp_filter2(ic = icr, method = method, h = h, kernel = kernel)
-          filter(x, lp_coef)
-        })
-        names(series_s) <- names(data)
-        
-        saveRDS(series_s, nom_f_s)
-        
-        print("turning points")
-        tp <- lapply(series_s, turning_points)
-        saveRDS(tp,
-                nom_f_s_tp)
-        
[…]

### 2024-02-12 R_simul/2C_Est_dafic_trunc.R

@@ -1,11 +1,7 @@
 if(!dir.exists("results_simul"))
   dir.create("results_simul")
-if(!dir.exists("results_fredm"))
-  dir.create("results_fredm")
 if(!dir.exists("results_simul/localic_daf_trunc"))
   dir.create("results_simul/localic_daf_trunc")
-if(!dir.exists("results_fredm/localic_daf_trunc"))
-  dir.create("results_fredm/localic_daf_trunc")
 library(rjd3filters)
 library(AQLThesis)
 library(future)
@@ -104,61 +100,3 @@ if(reload){
   j <- 1
 }
 
-
-fs <- list()
-j <- 1
-kernel = "Henderson"
-for (method in c("LC","QL")){
-  print(method)
-  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
-    for(d in 2:3){
-      name_file <- gsub(".RDS$", "", basename(s))
-      print(name_file)
-      data <- readRDS(s)
-      complement = sprintf("_d%i", d)
-      
-      nom_f_s <- sprintf("results_fredm/localic_daf_trunc/%s_%s%s.RDS",
-                         name_file, tolower(method), complement)
-      nom_f_s_tp <- 
-        sprintf("results_fredm/localic_daf_trunc/%s_%s%s_tp.RDS",
-                name_file,
-                tolower(method), complement)
-    
-      
-      data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber_dt/%s.RDS",
-                                   gsub(".RDS", "",basename(s))))
-      if(all(file.exists(nom_f_s_tp),
-             file.exists(nom_f_s)))
-        next;
-      
-      reload <- TRUE
-      fs[[j]] <- future({
-        print(s)
-        series_s <- lapply(names(data), function(nom_d){
-          x <- data[[nom_d]]
-          data_t = data_info[[nom_d]][[method]]
-          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
-          icr = 2/(sqrt(pi) * ratio)
-          icr[abs(icr) > 12] <- 12
-          lp_coef = lp_filter2(ic = icr, method = method, h = h, kernel = kernel)
-          rjd3filters::filterfilter(x, lp_coef)
-        })
-        names(series_s) <- names(data)
-        
-        saveRDS(series_s, nom_f_s)
-        
-        print("turning points")
[…]

### 2024-03-13 R_simul/2D_est_arima_robust.R
robust ARIMA

@@ -1,0 +1,76 @@
+if(!dir.exists("results_simul"))
+  dir.create("results_simul")
+if(!dir.exists("results_simul/arima"))
+  dir.create("results_simul/arima")
+library(rjd3filters)
+library(AQLThesis)
+library(future)
+library(forecast)
+plan(multisession)
+
+list_series <- list.files("data_simul/byseries", full.names = TRUE)
+l = 13
+henderson <- lp_filter(horizon = (l - 1) / 2)@sfilter
+s = list_series[1]
+fs <- list()
+i <- 0
+nyears <- c(4, 6, 8, 10, 12)
+for(y in nyears){
+  for(s in list_series){
+    i <- i+1
+    name_file <- gsub(".RDS$", "", basename(s))
+    print(sprintf("%s nyears = %s", name_file, y))
+    fs[[i]] <- future({
+      print(s)
+      data <- readRDS(s)
+      data_info <- readRDS(sub("byseries", "byseriesinfo", s))
+      nom_f_s <- sprintf("results_simul/arima/%s_ny%s.RDS",
+                         name_file, y)
+      nom_f_s_tp <- 
+        sprintf("results_simul/arima/%s_ny%s_tp.RDS",
+                name_file, y)
+      
+      nom_f_s_rev_fe <- sprintf("results_simul/arima/%s_ny%s_fe_rev.RDS",
+                                name_file, y)
+      nom_f_s_rev_ce <- sprintf("results_simul/arima/%s_ny%s_ce_rev.RDS",
+                                name_file, y)
+      
+      if(all(file.exists(nom_f_s_tp),
+             file.exists(nom_f_s_rev_fe),
+             file.exists(nom_f_s_rev_ce)))
+        next;
+      
+      series_s <- lapply(names(data), function(nom_d){
+        x <- data[[nom_d]]
+        # l = data_info[[nom_d]]["optimal_length"]
+        first_date <- 1 + max(0, length(x) - y * frequency(x))
+        x_arima <- window(x, start = time(x)[first_date])
+        prevs = auto.arima(x_arima, max.Q = 0, max.D = 0, max.P = 0)
+        prevs = forecast(prevs, h=(l-1)/2)
+        y_prevs = ts(c(x, prevs$mean), start = start(x), frequency = frequency(x))
+        window(henderson * y_prevs,
+               start = start(x), end = end(x))
+      })
+      names(series_s) <- names(data)
+      
[…]

### 2024-03-13 R_filters/plot_coef.R
typos

@@ -1,6 +1,6 @@
 library(rjd3filters)
 library(patchwork)
-library(AQLThesis)
+library(ggdemetra3)
 library(ggplot2)
 
 all_filters <- lapply(c("LC", "QL", "CQ", "DAF"), function(endpoints){
@@ -19,7 +19,6 @@ all_filters <- lapply(c("LC", "QL", "CQ", "DAF"), function(endpoints){
 names(all_filters) <- c("LC", "QL", "CQ", "DAF")
 
 
-Sys.setenv(LANG = "en_US.UTF-8")
 for(i in names(all_filters)){
   print(i)
   x <- all_filters[[i]]

### 2024-03-13 R_fredm/1_infobyseries.R

@@ -12,10 +12,10 @@ for(s in list.files("data_fredm/byseries",full.names = TRUE)){
   fs[[i]] <- future({
     data <- readRDS(s)
     info <- lapply(data, function(x){
-      res = rjd3filters::select_trend_filter(x)
+      res = rjd3x11plus::select_trend_filter(x)
       res = c(res[c("length", "icr")],
-              rjd3filters::select_trend_filter(x, length = 9)[1], 
-              rjd3filters::select_trend_filter(x, length = 23)[1])[c(1,3,2,4)]
+              rjd3x11plus::select_trend_filter(x, length = 9)[1], 
+              rjd3x11plus::select_trend_filter(x, length = 23)[1])[c(1,3,2,4)]
       names(res) = c("optimal_length", "icr-9", "icr-13", "icr-23")
       res
       })

### 2024-03-13 R_fredm/4A_Analyse_est.R

@@ -16,7 +16,6 @@ detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
 
 
 
-Sys.setenv(LANG = "en_US.UTF-8")
 data <- readRDS("data_fredm/byseries/CE16OV.RDS")[["2020"]]
 dput(round(data, 3))
 plot(data)

### 2024-03-13 R_fredm/4B_Analyse_implicit_forecast_dt.R

@@ -4,7 +4,6 @@ source("R_fredm/4_utils.R",encoding = "UTF-8")
 library(rjd3filters)
 library(AQLThesis)
 
-Sys.setenv(LANG = "en_US.UTF-8")
 fst_weights <- "weight235"
 fst_degree <- 2
 all_tp <- get_all_tp(dossier = "results_fredm/compile_tp_norev/",

### 2024-03-13 R_simul/2D_est_arima.R

@@ -5,12 +5,13 @@ if(!dir.exists("results_simul/arima"))
 library(rjd3filters)
 library(AQLThesis)
 library(future)
+library(forecast)
 plan(multisession)
 
 list_series <- list.files("data_simul/byseries", full.names = TRUE)
-list_method <- names(arima_f$`h=6`)
 s = list_series[1]
-method = "phase"
+l = 13
+henderson <- lp_filter(horizon = (l - 1) / 2)@sfilter
 fs <- list()
 i <- 0
 for(s in list_series){
@@ -40,11 +41,10 @@ for(s in list_series){
     series_s <- lapply(names(data), function(nom_d){
       x <- data[[nom_d]]
       # l = data_info[[nom_d]]["optimal_length"]
-      l = 13
       prevs = auto.arima(x, max.Q = 0, max.D = 0, max.P = 0)
       prevs = forecast(prevs, h=(l-1)/2)
       y_prevs = ts(c(x, prevs$mean), start = start(x), frequency = frequency(x))
-      window(henderson(y_prevs, musgrave = FALSE,length = l),
+      window(henderson * y_prevs,
              start = start(x), end = end(x))
     })
     names(series_s) <- names(data)

### 2024-03-13 img/Ex_climat_cl4.R

@@ -1,317 +1,0 @@
-# id <- c(manuf ="001585934",
-#         C1 = "001786503",
-#         C3 = "001786504",
-#         C4 = "001786505",
-#         C5 = "001786506"
-#         )
-library(insee)
-library(rjd3filters)
-library(ggplot2)
-library(patchwork)
-library(zoo)
-library(forecast)
-library(AQLThesis)
-# Import des données
-y <- data.frame(insee::get_insee_idbank("001786505"))
-y <- y[order(y[,"TIME_PERIOD"]), ]
-first_date <- as.numeric(strsplit(y$TIME_PERIOD[1], "-")[[1]])
-y <- ts(y$OBS_VALUE, start = first_date, frequency = 12)
-y <- window(y, start = 2010)
-last_dates <- c(tail(time(y), 7))
-names(last_dates) <- as.character(zoo::as.yearmon(last_dates))
-der_est <- lapply(last_dates, function(x) window(y, end = x))
-
-# MM de longueur 13 adaptées : 
-select_trend_filter(y)
-# Estimation finale
-tc_f <- henderson(y, length = 13, musgrave = FALSE) # Estimation finale
-plot(y)
-lines(tc_f, col = "red")
-
-p <- forecast::autoplot(ts.union(y, tc_f)) +
-  theme_bw() +
-  theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
-                                         colour = "gray80", linetype = "solid"),
-        legend.justification = c(0,0),
-        legend.position = c(0,0),
-        legend.key = element_blank(),
-        legend.title = element_blank()) +
-  labs(x = NULL, y = NULL)  +
-  scale_color_discrete(labels =c("Série initiale","Série lissée"))
-p
-ggMultisave("DT/img/ex/tc_finale", 
-            plot = p,
-            width = 8, height = 3)
-
-## Méthode polynomiale
-
-# Calcul des IC-ratios : 
-icr <- sapply(der_est, function(x) {
-  ic_ratio(x, henderson(x, length = 13, musgrave = FALSE))
-})
-lp_est <- lapply(c("LC", "QL", "CQ", "DAF"), function(method) {
-  res <- lapply(seq_along(icr), function(i) {
-    lp_coef <- lp_filter(horizon = 6,
-                         kernel = "Henderson",
-                         endpoints = method,
[…]

### 2024-03-13 article_jos.qmd
typo + simplify preservation of trends

@@
 Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
-Noting $\omega = 2\pi/p$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $p$ periods.
-For example, for a monthly series (12 observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
+Noting $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
+For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
 In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e. movements recurring at least every 12 months.
@@
 
-We have the following properties for the moving average $M_{\boldsymbol\theta}$:
+For the moving average $M_{\boldsymbol\theta}$, it can be shown that:
 
-- To preserve a constant trend $X_t=a$ we need:
-$$
-\forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
-$$
-The sum of the coefficients of the moving average $\sum_{k=-p}^{+f}\theta_k$ must then be equal to $1$.
-
-- To preserve linear trends $X_t=at+b$ we need:
-$$
-\forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
-$$
-Which is equivalent to:
-$$
-\sum_{k=-p}^{+f}\theta_k=1
-\quad\text{and}\quad
-\sum_{k=-p}^{+f}k\theta_k=0.
-$$
-- In general, $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
+- $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
 $$

### 2024-03-13 R_simul/1A_data_creation.R
typo rjd3x11plus

@@ -50,10 +50,10 @@ for(s in list.files("data_simul/byseries",full.names = TRUE)){
     data <- readRDS(s)
     info <- lapply(data, function(x){
         x = data[[5]]
-      res = rjd3filters::select_trend_filter(x)
+      res = rjd3x11plus::select_trend_filter(x)
       res = c(res[c("length", "icr")],
-              rjd3filters::select_trend_filter(x, length = 9)[1], 
-              rjd3filters::select_trend_filter(x, length = 23)[1])[c(1,3,2,4)]
+              rjd3x11plus::select_trend_filter(x, length = 9)[1], 
+              rjd3x11plus::select_trend_filter(x, length = 23)[1])[c(1,3,2,4)]
       names(res) = c("optimal_length", "icr-9", "icr-13", "icr-23")
       res
     })

### 2024-03-28 R_filters/plot_coef.R
nearest neighbour filters

@@ -81,17 +81,29 @@ ggsave("img/filters_used/nn_henderson.pdf",
        width = 8, height = 4.5,
        plot = p)
 
-x <- nn_ma[["d=2"]]
-p <- ggplot_coef(x, q = c(0:6)) / (
-  ggplot_gain(x, q = c(0:6)) +
-    ggplot2::scale_y_continuous(
-      "Gain",
-      breaks = seq(0, 1, by = 0.2))+
-    ggplot2::guides(colour = "none") +
-    ggplot_phase(x, xlim = c(0, 4/12*pi), q = c(0:6))+
-    ggplot2::labs(y = "Phase shift") +
-    ggplot2::guides(colour = "none"))
-p
-ggsave("img/filters_used/nn_lp_d2.pdf",
-       width = 8, height = 4.5,
-       plot = p)
+h_filter <- lp_filter(horizon = 6, kernel = "Henderson")@sfilter
+all_filters <- lapply(c("LC", "QL", "CQ", "DAF"), function(endpoints){
+  finite_filters(h_filter, lapply(1:6, function(i){
+    lp_filter(horizon=6+i, endpoints = endpoints,
+              kernel = "Henderson",
+              ic = 3.5)[,2*i+1]
+  }))
+})
+names(all_filters) <- c("LC", "QL", "CQ", "DAF")
+for(i in names(all_filters)){
+  print(i)
+  x <- all_filters[[i]]
+  p <- ggplot_coef(x, q = c(0:6)) / (
+    ggplot_gain(x, q = c(0:6)) +
+      ggplot2::scale_y_continuous(
+        "Gain",
+        breaks = seq(0, 1, by = 0.2))+
+      ggplot2::guides(colour = "none") +
+      ggplot_phase(x, xlim = c(0, 4/12*pi), q = c(0:6))+
+      ggplot2::labs(y = "Phase shift") +
+      ggplot2::guides(colour = "none"))
+  p
+  ggsave(sprintf("img/filters_used/%s_nn.pdf",tolower(i)),
+         width = 8, height = 4.5,
+         plot = p)
+}

### 2024-03-28 R_filters/slope_concavity_estimator.R
update plots

@@ -1,6 +1,7 @@
 library(rjd3filters)
 library(ggplot2)
 library(patchwork)
+library(ggdemetra3)
 h <- 6
 q <- 6
 X_gen <- function(d = 1, p = 6, q = p){

### 2024-03-28 R_fredm/4A_Analyse_est.R

@@ -2,53 +2,48 @@
 source("R_fredm/4_utils.R",encoding = "UTF-8")
 library(patchwork)
 
-Sys.setlocale(locale = "en_US.UTF-8")
 all_tp <- get_all_tp(dossier = "results_fredm/compile_tp_norev/")
 all_tp_rev <- get_all_tp(dossier = "results_fredm/compile_tp/")
 detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
 
 
-data <- readRDS("data_fredm/byseries/CE16OV.RDS")[["2020"]]
-dput(round(data, 3))
+data <- readRDS("data_fredm/byseries/CE16OV.RDS")[["2022"]]
 plot(data)
 all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
                            series = "CE16OV", tp_keep ="2001.16666666667",
                            nb_est = 8)
-wrap_plots(all_plots, ncol =3)
+wrap_plots(all_plots, ncol = 3) & scale_color_grey()
 
 ggsave("img/nber/ce16ov_fev2001_lp.pdf",
-            plot = wrap_plots(all_plots, ncol = 3),
-            width = 8, height = 8)
+       plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
+       width = 8, height = 8)
 
 all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
                            series = "CE16OV", tp_keep ="2020.25",
                            nb_est = 8)
-wrap_plots(all_plots, ncol =3)
+wrap_plots(all_plots, ncol = 3) & scale_color_grey()
 
 ggsave("img/nber/ce16ov_covid_lp.pdf",
-       plot = wrap_plots(all_plots, ncol = 3),
+       plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 
-data <- readRDS("data_fredm/byseries/RETAILx.RDS")[["2020"]]
+data <- readRDS("data_fredm/byseries/RETAILx.RDS")[["2022"]]
 plot(window(data, start = 2019))
 all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
                            series = "RETAILx", tp_keep ="2007.91666666667",
                            nb_est = 8)
-wrap_plots(all_plots, ncol = 3)
[…]

### 2024-03-28 R_fredm/4B_Analyse_implicit_forecast_dt.R

@@ -21,11 +21,11 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              all_tp_rev = all_tp_rev,
                              tp_keep = tp_keep,
                              dossier = "results_fredm/compile_tp_norev/")
-wrap_plots(plots,ncol = 3)
+wrap_plots(plots,ncol = 3) & scale_color_grey()
 
 ggsave("img/nber/ce16ov_fev2001_prev_imp_lp.pdf",
-            plot = wrap_plots(plots, ncol = 3),
-            width = 8, height = 8)
+       plot = wrap_plots(plots, ncol = 3) & scale_color_grey(),
+       width = 8, height = 8)
 
 series <- "CE16OV"
 tp_keep <- "2020.25"
@@ -37,10 +37,14 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              all_tp_rev = all_tp_rev,
                              tp_keep = tp_keep,
                              dossier = "results_fredm/compile_tp_norev/")
-wrap_plots(plots,ncol = 3)
+wrap_plots(plots,ncol = 3) & 
+  scale_color_grey() & 
+  labs(subtitle = NULL)
 
 ggsave("img/nber/ce16ov_covid_prev_imp_lp.pdf",
-       plot = wrap_plots(plots, ncol = 3),
+       plot = wrap_plots(plots, ncol = 3) & 
+         scale_color_grey() & 
+         labs(subtitle = NULL),
        width = 8, height = 8)
 
 
@@ -62,5 +66,5 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
 wrap_plots(plots, ncol = 3)
 
 ggsave("img/nber/retailx_nov2007_prev_imp_lp.pdf",
-            plot = wrap_plots(plots, ncol = 3),
-            width = 8, height = 8)
+       plot = wrap_plots(plots, ncol = 3),
+       width = 8, height = 8)

### 2024-03-28 R_simul/2A_est_lp_nn.R

@@ -11,6 +11,16 @@ list_series <- list.files("data_simul/byseries", full.names = TRUE)
 list_kernel <- c("Henderson", "Uniform", "Biweight", "Triweight", "Tricube",
                  "Gaussian", "Triangular", "Parabolic")[1]
 
+h_filter <- lp_filter(horizon = 6, kernel = kernel)@sfilter
+lp_filter2_nn <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
+  rfilters = lapply(seq_along(icr), function(i){
+    lp_filter(horizon=h+i, 
+              endpoints = method,
+              kernel = kernel,
+              ic = icr[i])[,2*i+1]
+  })
+  finite_filters(h_filter, rfilters = rfilters)
+}
 fs <- list()
 j <- 1
 reload <- FALSE

### 2024-03-28 R_simul/4A_Analyse_tp.R

@@ -35,14 +35,14 @@ tp_lic_final_nn <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic
   dplyr::filter(degree == "d2", h == "h6") %>%
   select_var() %>% mutate(method = sprintf("%s_localic_final", method)) %>%
   select(!c(degree, h)) %>% 
-  mutate(method = sprintf("%s_localic", method))
+  mutate(method = sprintf("%s_nn", method))
 tp_lic_daf_trunc_nn <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic_daf_trunc_nn.RDS"),
                           readRDS("results_simul/compile_tp_norev/peaks_localic_daf_trunc_nn.RDS"),
                           by=c("series", "kernel", "h", "degree", "method")) %>%
   dplyr::filter(degree == "d2", h == "h6") %>%
   select_var() %>% mutate(method = sprintf("%s_localic", method)) %>%
-  select(!c(degree, h))%>%
-  mutate(method = sprintf("%s_localic", method))
+  select(!c(degree, h)) %>%
+  mutate(method = sprintf("%s_nn", method))
 
 tp_arima <-
   merge(readRDS("results_simul/compile_tp_norev/troughs_arima.RDS"),
@@ -99,9 +99,9 @@ legende <- c(lc = "LC", ql = "QL",
              ql_localic_final = "QL loc. param.\n(final estimates)",
              ql_localic = "QL loc.\nparam.",
              auto_arima = "ARIMA")
-p <- ggplot(data_tp %>% 
+p <- ggplot(all_tp %>% 
               filter(method %in%
-                       names(legende)) ,aes(x=method, y = value))+
+                       names(legende))  %>% format_table_tp() ,aes(x=method, y = value))+
   geom_boxplot() +
   facet_wrap(vars(variability), ncol = 1) + theme_bw() +
   labs(y="Phase shift", x = NULL) +
@@ -109,27 +109,19 @@ p <- ggplot(data_tp %>%
   theme(axis.text.x = element_text( margin = margin(10, 0, 0, 0), vjust = 1))
 p
 
-
 ggsave("img/simulations/phase_shift_simul.pdf",
             plot = p,
             width = 10, height = 6)
-
-legende <- c(lc = "LC",
-             ql = "QL",
[…]

### 2024-03-28 resp_referee.qmd
resp referee

---
title: "Response to referee"
format: pdf
editor: visual
---

# Associated Editor

1. Page 3, line 46: you say that it is "equivalently defined" and then provide the conjugate,
writing $e^{i\omega k}$ instead of $e^{-i\omega k}$. "Equivalence" implies something akin to equality, so I would soften this to "alternatively", since the two expressions are not equal.

2. Page 5, line 53: replace "have" with "design", I think, because you are constructing these
filters, right?

3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function
at $\omega = 0$ yield the conditions. So the results could be stated with references, and the
derivations omitted, and this subsection compactified.

4. Page 7, line 40: why say the series is extended by one year? The extensions are for as many
years as needed by the particular filters (typically, half their length in years).

5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.

6. Table 1: I repeat the referees here. The table needs more explication in the text. What is $\omega_1$? Define $\nabla$, Guggemos, etc.

7. Page 13, equation (4): it is hard to understand what $\theta_f$ and $\theta_p$ are, and what 
$\gamma$ and $\delta$ are.
Add more explanation to the text.

8. Page 15, line 36: should be "same length as $v$"

9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness?
That is, when you display the entries of the matrix T, you need to make a connection between
those values and the concept of timeliness.

10. Page 15, line 53: I don't see why $v$ in this equation should be defined through the case bracket;
everything starting with "with" could be moved to the following text.

11. Page 17, lines 53-55: This discussion needs more clarity; it seems a bit terse.

12. Appendix B: rather than typing out the R code, it would be better to have the files available
as supplementary material. (Or possibly put in a github repo and provide the links in the
text.)

# Referee 1

The authors should provide a clear definition of trend-cycle, $TC_t$,
and irregular, $I_t$. In other words, how is the decomposition
$y_t =TC_t+I_t$ identified? In section 1 they seem to hint at the interpretation
of $TC_t$ as the low-pass component of yt with cutoff frequency
$\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component).
This point should be clarified, I think.
Related questions are: are $y_t$ and its components stochastic or deterministic?
The treatment in section 1.2.1 suggests that $TC_t$ is
an unknown possibly nonlinear deterministic function of time.
The variance inflation factor $\sum_k \theta_k^2$ implies that $I_t$ is white noise.

2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not
very formal. Do you mean that the unconditional variance is minimized?
Or the long run variance?

3. A definition of turning point would also be needed. Are these ’regular’
points at which the slope of the trend changes sign? Or changepoints?

4. The notation should be unified: why not writing $TC_t = \mu(t)$ and
in section 2 using $t$ instead of $x$? The observation index i hould
be replaced by $t$, as in section 2.1. Similarly, why not using ϵt
directly and start section 1.2 by stating that you assume that yt is
generated as $y_t = \mu(t) + \varepsilon_t$, etc.
Secondly, you should clarify the meaning of (3), where you set
$\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix
$D$ for later treatment?
5. You should state that you assume a fixed bandwidth. A nearest
neighbour bandwidth should be interesting to consider: as such
it would contribute more to the reduction of the variance at the
expenses of the phase shift.
6. Is the phase shift related to the notion of bias in local polynomial
smoothing?

Minor points

• Section 1.1. It would perhaps be useful to provide the expression
of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of
the transfer function of the linear filter.
• Section 1.1. The symbol p for the period is already in use from the
previous section, where it denotes the bandwidth and number of
past observations used by the filter.
• Page 4. The definition of the high-pass component (oscillations
in the frequency range from $\pi/6$ to $\pi$ as undesirable oscillations is
controversial.
• Figure 1. In the caption, the ration $I/C$ was never defined before.
The description of the three panels should be better explained in
the main text.
• Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
assumes that the reader is knowledgeable about seasonal adjustment
methods. Perhaps the authors should state that their
intent is to separate the trend-cycle component from the ’noise’
component.
• Page 7, line 38. ’which extends’ is repeated twice.
• Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts
induced by $w_0$, etc.? The forecasts are obtained from an assumed
polynomial trend model?
• Table 1. Some criteria are labelled as Guggemos, not referenced in
the main text.
• Figures 6-7. The labels of the time axes and the months are in
French.

# Referee 2

> The author simulates time series of length 60 years to illustrate the methodology at the end of the series.  It may be worth considering (in addition) shorter series.  Another thought is that the ARIMA model being used to extend the series appears to be modeled using the full series; what if the ARIMA model used to extend the series was instead estimated using just a short span toward the end of the series?

> The author avoids numbering the Introduction and Conclusion; numbering both (and adjusting the numbering of the other sections accordingly) may be better.

> Adding the appropriate items for both R and X-13ARIMA-SEATS to the References may be warranted.  In addition, in the text, it would probably be preferable to use R (the letter itself) instead of the R icon.

> Some of the important results appear in the tables and figures, but the accompanying discussion in the text should still be fleshed out more.  In addition, in Figures 2-4 and 6-7 (and the figures in the supplement), the lines can be distinguished by color, but it may be helpful to find some way to keep them distinguishable without the benefit of color.  It might be the case that there is too much information to display neatly in some of these.

> It is somewhat unclear just how prominent the new functionality being added via the R package rjd3filters is intended to be in this paper.

# Referee 3

> 1. Equivalence between Asymmetric and Symmetric Filters (Page 8, Line 9)
> I suggest that the authors provide a more explicit explanation on page 8, line 9, regarding
the equivalence between asymmetric and symmetric filters. It would be beneficial to
clarify the normalization difference and ensure a clear understanding of the relationship
between the two types of filters.

> 2.  Symbol [[ ]] Meaning (Page 11, Lines 53-54)
> The meaning of the symbol [[ ]] on page 11, lines 53-54, is not clear. I recommend that
the authors explicitly define this symbol or provide a clear explanation to avoid any
ambiguity in its interpretation.

> 3.  Variance Determination in Local Parameterization (Page 17)
> The determination of variance in the local parameterization of asymmetric filters (page
17) could benefit from a more detailed explanation. I recommend that the authors
clearly identify the quantities appearing in the denominator for better understanding.

> 4.  Applicability to High-Frequency Data.
> It would be valuable if the authors could explicitly address the applicability of the
proposed procedure to high-frequency data. If modifications or considerations are
necessary for such data, it should be discussed to provide a comprehensive
understanding of the methodology.

> 5.  Clarity on Phase Shift Computation (Page 20)
> A clearer explanation of how the phase shift is computed on page 20 would enhance the
paper's overall clarity. The authors should specify the method used for determining the
phase shift, particularly in relation to the identification of true turning points based on
symmetric filter estimates.

> 6.  Commentary on Figure 5
> I suggest that the authors include a detailed commentary on Figure 5 in the paper.
Providing explanations for the results shown in the figure would be crucial for readers to
better understand the performance of the filters/methods.

> 7.  Better Description of Results in Table 3
> The meaning of "second and third estimates" in Table 3 on page 22 is not clear. I
recommend that the authors provide a more detailed description or explanation of the
results to improve comprehension.

> 8.  Consideration of Different Series Periodicity
> It would be beneficial if the authors could address the suggestion of considering series
with different periodicities in the real data application. Discussing how the proposed
methodology performs when applied to series with various frequencies would add
depth to the study.

# Referee 4

1) The paper lacks in describing the contribution from the beginning. I had to read the paper a couple of times to clearly identify the contribution. I found the contribution of the paper clearly stated on page 16, but I think I should come before in the text.
2) The paper omits to consider the Covid19 period and how this influences the trend
and cycle decomposition. I understand that this is not the main focus of the paper,
but at least the authors should address this problem in word and possible solution
to address it, e.g. outlier correction, robust distributions, etc.
3) I think a quote from other works can be removed from the paper. For example, I
would remove the quote in point 2) of page 7.

### 2024-03-28 article_jos.qmd
beginning discussion

@@
 
+# Discussion
+## Outliers and COVID-19
+
+Graph + simulations
+
+Enfin, dans cet article et dans ceux associés aux méthodes étudiées, les filtres asymétriques sont appliqués et comparés sur des séries déjà désaisonnalisées ou sans saisonnalité (séries simulées). 
+Les révisions et donc le déphasage sont limités à 6 mois (lorsque le filtre symétrique est de 13 termes) et cela a l'avantage d'isoler les impacts des différents filtres des autres processus inhérents à la désaisonnalisation.
+Il y a toutefois deux inconvénients à cette simplification :
+
+1. D'une part, l'estimation de la série désaisonnalisée dépend de la méthode utilisée pour extraire la tendance-cycle.
+Le choix de la méthode utilisée pour l'estimation de la tendance-cycle peut donc avoir un impact bien au-delà de 6 mois.
+
+2. D'autre part, les moyennes mobiles étant des opérateurs linéaires, ils sont sensibles à la présence de points atypiques.
+L'application directe des méthodes peut donc conduire à des estimations biaisées, du fait de leur présence, alors que les méthodes de désaisonnalisation (comme la méthode X-13ARIMA) ont un module de correction des points atypiques.
+Par ailleurs, comme notamment montré par @dagum1996new, le filtre symétrique final utilisé par X-13ARIMA pour extraire la tendance-cycle (et donc celui indirectement utilisé lorsqu'on applique les méthodes sur les séries désaisonnalisées) laisse passer environ 72 % des cycles de 9 ou 10 mois (généralement associés à du bruit plutôt qu'à la tendance-cycle).
+Les filtres asymétriques finaux amplifient même les cycles de 9 ou 10 mois.
+Cela peut avoir pour conséquence l'introduction d'ondulations indésirables, c'est-à-dire la détection de faux points de retournement.
+Ce problème est réduit par la correction des points atypiques (ces cycles étant considérés comme de l'irrégulier). 
+C'est ainsi que le *Nonlinear Dagum Filter* (NLDF) a été développé et consiste à :
+
+    a. appliquer l'algorithme de correction des points atypiques de X-13ARIMA sur la série désaisonnalisée, puis de la prolonger par un modèle ARIMA ;
+
+    b. effectuer une nouvelle correction des points atypiques en utilisant un seuil bien plus strict et appliquer ensuite le filtre symétrique de 13 termes. 
+    En supposant une distribution normale cela revient à modifier 48 % des valeurs de l'irrégulier.
+
+    Les *cascade linear filter* (CLF), notamment étudiés dans @dagumBianconcini2023, correspondent à une approximation des NLDF en utilisant un filtre de 13 termes et lorsque les prévisions sont obtenus à partir d'un modèle ARIMA(0,1,1) où $\theta=0,40.$
+
+Une piste d'étude serait alors d'étudier plus précisément l'effet des points atypiques sur l'estimation de la tendance-cycle et la détection des points de retournement, mais aussi d'explorer de nouveaux types de filtres asymétriques fondés sur des méthodes robustes (comme les régressions locales robustes, les médianes mobiles, etc.).
+
+## Other frequencies
+
 # Conclusion

### 2024-03-28 R_simul/2A_est_lp_nn.R
dclean

@@ -11,16 +11,6 @@ list_series <- list.files("data_simul/byseries", full.names = TRUE)
 list_kernel <- c("Henderson", "Uniform", "Biweight", "Triweight", "Tricube",
                  "Gaussian", "Triangular", "Parabolic")[1]
 
-h_filter <- lp_filter(horizon = 6, kernel = kernel)@sfilter
-lp_filter2_nn <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
-  rfilters = lapply(seq_along(icr), function(i){
-    lp_filter(horizon=h+i, 
-              endpoints = method,
-              kernel = kernel,
-              ic = icr[i])[,2*i+1]
-  })
-  finite_filters(h_filter, rfilters = rfilters)
-}
 fs <- list()
 j <- 1
 reload <- FALSE

### 2024-04-02 paper/article_jos.tex
paer folder

--- 2024-02-12 article_jos.tex
@@
 \newcommand\Z{\mathds{Z}}
-\newcommand{\determinant}[1]{\Determinant\left(#1\right)}
+\newcommand\transp[1]{{#1}^T}
 %\newcommand\C{\mathds{C}}
@@
 \usepackage{xcolor}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
 \makeatletter
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
 \makeatother
@@
 \addbibresource{biblio.bib}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\hypertarget{abstract}{%
-\section*{Abstract}\label{abstract}}
+\section*{Abstract}\label{abstract}
 \addcontentsline{toc}{section}{Abstract}
@@
 
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
+\section{Introduction}\label{introduction}
 
@@
 order to take these shortcomings into account. The extensions have been
-implemented in the \faIcon{r-project} \texttt{rjd3filters} package
-(available at \url{https://github.com/rjdemetra/rjd3filters}) and are
-therefore easily reusable.
+implemented in the R \texttt{rjd3filters} package (available at
+\url{https://github.com/rjdemetra/rjd3filters}) and are therefore easily
+reusable.
 
@@
 
-\hypertarget{sec-propMM}{%
-\section{Some properties on moving averages}\label{sec-propMM}}
+\section{Some properties on moving averages}\label{sec-propMM}
 
@@
 \(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
-\(\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}\)
+\(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
 such as for all time series \(X_t\): \[
@@
 
-\hypertarget{subsec:gain-deph}{%
-\subsection{Gain and phase shift functions}\label{subsec:gain-deph}}
+\subsection{Gain and phase shift functions}\label{subsec:gain-deph}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}
-
-}
-
-\end{figure}
-
-\hypertarget{desirable-properties-of-a-moving-average}{%
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}
+
+}
+
+\end{figure}%
+
 \subsection{Desirable properties of a moving
-average}\label{desirable-properties-of-a-moving-average}}
+average}\label{desirable-properties-of-a-moving-average}
 
@@
 
-\hypertarget{trend-preservation}{%
-\subsubsection{Trend preservation}\label{trend-preservation}}
+\subsubsection{Trend preservation}\label{trend-preservation}
 
@@
 
-\hypertarget{noise-reduction}{%
-\subsubsection{Noise reduction}\label{noise-reduction}}
+\subsubsection{Noise reduction}\label{noise-reduction}
 
@@
 
-\hypertarget{sec-mmasym}{%
 \subsection{Real-time estimation and asymmetric moving
-average}\label{sec-mmasym}}
-
-\hypertarget{sec-mmetprev}{%
+average}\label{sec-mmasym}
+
 \subsubsection{Asymmetric moving averages and
-forecasts}\label{sec-mmetprev}}
+forecasts}\label{sec-mmetprev}
 
@@
 
-\hypertarget{subsec:crit-qual}{%
 \subsubsection{Quality indicators for asymmetric moving
-averages}\label{subsec:crit-qual}}
+averages}\label{subsec:crit-qual}
 
@@
 notion of phase shift, which again we want to be low. This method can be
-used in\faIcon{r-project} with the function
-\texttt{rjd3filters::fst\_filter()}.
+used inR with the function \texttt{rjd3filters::fst\_filter()}.
 
@@
 \emph{Residual} (\(R_w\)). A simplified version of this method can be
-used in \faIcon{r-project} with the function
-\texttt{rjd3filters::dfa\_filter()}.
-
-\begin{table}[!h]
+used in R with the function \texttt{rjd3filters::dfa\_filter()}.
+
+\begin{table}
+
 \caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
+
+\centering{
+
+[!h]
+
 {\centering
@@
 \textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
-\end{table}
-
-\hypertarget{sec-nonparamreg}{%
+
+}
+
+\end{table}%
+
 \section{Non-parametric regression and local polynomial
-regression}\label{sec-nonparamreg}}
+regression}\label{sec-nonparamreg}
 
@@
 is differentiable \(d\) times, then:
-\begin{equation}\protect\hypertarget{eq-taylor}{}{
+\begin{equation}\phantomsection\label{eq-taylor}{
 \forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
 \frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
-}\label{eq-taylor}\end{equation} where \(R_d\) is a negligible residual
-term in the neighbourhood of \(x_0\). In a neighbourhood \(h(x_0)\)
-around \(x_0\), \(\mu\) can be approximated by a polynomial of degree
-\(d\). The quantity \(h(x_0)\) is called the \emph{bandwidth}. If
-\(\varepsilon_i\) is white noise, then we can estimate by least squares
-\(\mu(x_0)\) using the observations which are in
-\(\left[x_0-h(x_0),x_0+h(x_0)\right]\).
+}\end{equation} where \(R_d\) is a negligible residual term in the
+neighbourhood of \(x_0\). In a neighbourhood \(h(x_0)\) around \(x_0\),
+\(\mu\) can be approximated by a polynomial of degree \(d\). The
+quantity \(h(x_0)\) is called the \emph{bandwidth}. If \(\varepsilon_i\)
+is white noise, then we can estimate by least squares \(\mu(x_0)\) using
+the observations which are in \(\left[x_0-h(x_0),x_0+h(x_0)\right]\).
 
@@
 facilitating the calculation of different moving averages at different
-time frequencies. This method can be used in\faIcon{r-project} with the
-function \texttt{rjd3filters::rkhs\_filter()}.
-
-\hypertarget{symmetric-moving-averages-and-local-polynomial-regression}{%
+time frequencies. This method can be used inR with the function
+\texttt{rjd3filters::rkhs\_filter()}.
+
 \subsection{Symmetric moving averages and local polynomial
-regression}\label{symmetric-moving-averages-and-local-polynomial-regression}}
+regression}\label{symmetric-moving-averages-and-local-polynomial-regression}
 
@@
 \(\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.\)
-With \(\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}\),
+With
+\(\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}\),
 the estimate of the trend is:
-\begin{equation}\protect\hypertarget{eq-mmsym}{}{
+\begin{equation}\phantomsection\label{eq-mmsym}{
 \hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
-}\label{eq-mmsym}\end{equation} To conclude, the estimate of the trend
-\(\hat{m}_{t}\) can be obtained applying the symmetric filter
-\(\boldsymbol \theta\) to \(y_t\) (\(\boldsymbol \theta\) is symmetric
-due to the symmetry of the kernel weights \(\kappa_j\)). Moreover,
+}\end{equation} To conclude, the estimate of the trend \(\hat{m}_{t}\)
+can be obtained applying the symmetric filter \(\boldsymbol \theta\) to
+\(y_t\) (\(\boldsymbol \theta\) is symmetric due to the symmetry of the
+kernel weights \(\kappa_j\)). Moreover,
 \(\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}\) so: \[
@@
 
-\hypertarget{sec-lppasymf}{%
 \subsection{Asymmetric moving averages and local polynomial
-regression}\label{sec-lppasymf}}
+regression}\label{sec-lppasymf}
 
@@
 The initial series is modelled here as:
-\begin{equation}\protect\hypertarget{eq-lpgeneralmodel}{}{
+\begin{equation}\phantomsection\label{eq-lpgeneralmodel}{
 \boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
 \boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
-}\label{eq-lpgeneralmodel}\end{equation} where
-\([\boldsymbol U,\boldsymbol Z]\) is of full rank and forms a subset of
-the columns of \(X\). The objective is to find a filter
-\(\boldsymbol v\) that minimises the mean square error of revision (to
-the symmetric filter \(\boldsymbol \theta\)) under certain constraints.
-These constraints are represented by the matrix
-\(\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}\):
+}\end{equation} where \([\boldsymbol U,\boldsymbol Z]\) is of full rank
+and forms a subset of the columns of \(X\). The objective is to find a
+filter \(\boldsymbol v\) that minimises the mean square error of
+revision (to the symmetric filter \(\boldsymbol \theta\)) under certain
+constraints. These constraints are represented by the matrix
+\(\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}\):
 \(\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta\) (with
@@
 asymmetric filter). The problem is equivalent to finding \(v\) which
-minimises: \begin{equation}\protect\hypertarget{eq-lppasym}{}{
+minimises: \begin{equation}\phantomsection\label{eq-lppasym}{
 \varphi(\boldsymbol v)=
@@
 \underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
-}\label{eq-lppasym}\end{equation} where \(\boldsymbol l\) is a vector of
-Lagrange multipliers.
+}\end{equation} where \(\boldsymbol l\) is a vector of Lagrange
+multipliers.
 
@@
 
-When \(\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}\),
-\(\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}\),
+When \(\boldsymbol U=\transp{\begin{pmatrix}1&\cdots&1\end{pmatrix}}\),
+\(\boldsymbol Z=\transp{\begin{pmatrix}-h&\cdots&+h\end{pmatrix}}\),
 \(\boldsymbol \delta=\delta_1\), \(\boldsymbol D=\sigma^2\boldsymbol I\)
@@
 
-\begin{table}[!h]
+\begin{table}
+
 \caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
+
+\centering{
+
+[!h]
+
 \centering
@@
 \textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
-\end{table}
-
-\hypertarget{sec-lptimeliness}{%
+
+}
+
+\end{table}%
+
 \subsection{Extension with the timeliness
-criterion}\label{sec-lptimeliness}}
+criterion}\label{sec-lptimeliness}
 
@@
 \(\boldsymbol v\) the asymmetric filter. With
-\(\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}\)
+\(\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}\)
 and \(\boldsymbol\theta_p\) of the same length of \(\boldsymbol v\), and
@@
 generalizes the approach of \textcite{GrayThomson1996}. This extension
-is implemented in the \faIcon{r-project} function
-\texttt{rjd3filters::lp\_filter()}.
+is implemented in the R function \texttt{rjd3filters::lp\_filter()}.
 
@@
 
-\hypertarget{sec-localic}{%
 \subsection{Local parameterisation of asymmetric
-filters}\label{sec-localic}}
+filters}\label{sec-localic}
 
@@
 
-{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 IC ratio) on a simulated series, the vertical lines being the simulated
-turning points.}\begin{minipage}[t]{\linewidth}
+turning points.}
+
+\begin{minipage}{\linewidth}
+
 \subcaption{\label{fig-mmpenteconcac-ex-1}Symmetric moving average used
@@
 
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{img/filters_used/mm_penteconcavite_ex.pdf}
-
-}
+\centering{
+
+\includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite_ex.pdf}
 
@@
 \newline
-\begin{minipage}[t]{\linewidth}
+\begin{minipage}{\linewidth}
+
 \subcaption{\label{fig-mmpenteconcac-ex-2}Asymmetric moving average
@@
 
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{img/filters_used/mm_penteconcavite_ex2.pdf}
-
-}
+\centering{
+
+\includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite_ex2.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-comparison}{%
-\section{Comparison of different methods}\label{sec-comparison}}
+\end{figure}%
+
+\section{Comparison of different methods}\label{sec-comparison}
 
@@
 
-\hypertarget{simulated-series}{%
-\subsection{Simulated series}\label{simulated-series}}
-
-\hypertarget{methodology}{%
-\subsubsection{Methodology}\label{methodology}}
+\subsection{Simulated series}\label{simulated-series}
+
+\subsubsection{Methodology}\label{methodology}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/simul_data.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/simul_data.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-\hypertarget{comparison}{%
-\subsection{Comparison}\label{comparison}}
+\subsection{Comparison}\label{comparison}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-\begin{table}[!h]
+\begin{table}
+
 \caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
 filters on simulated series with medium variability.}
+
+\centering{
+
+[!h]
+
 \centering
@@
 \end{tabular}
-\end{table}
-
-\hypertarget{real-series}{%
-\subsection{Real series}\label{real-series}}
+
+}
+
+\end{table}%
+
+\subsection{Real series}\label{real-series}
 
@@
 
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}
-
-}
-
-\end{figure}
-
-\hypertarget{conclusion}{%
-\section*{Conclusion}\label{conclusion}}
-\addcontentsline{toc}{section}{Conclusion}
+\centering{
+
+\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}
+
+}
+
+\end{figure}%
+
+\section{Conclusion}\label{conclusion}
 
@@
 
-This paper presents the \faIcon{r-project} package \texttt{rjd3filters},
-which implements a several methods to build moving average for real-time
-trend-cycle estimation. It also offers various functions for the
-theoretical analysis of moving averages (gain functions, phase shift,
-quality criteria, etc.) and judging the quality of the latest estimates
-(for example with implicit forecasts). It also makes it easy to combine
+This paper presents the R package \texttt{rjd3filters}, which implements
+a several methods to build moving average for real-time trend-cycle
+estimation. It also offers various functions for the theoretical
+analysis of moving averages (gain functions, phase shift, quality
+criteria, etc.) and judging the quality of the latest estimates (for
+example with implicit forecasts). It also makes it easy to combine
 moving averages and also to integrate custom moving averages into the
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
@@
 
-\hypertarget{sec-an-cof}{%
-\section{Coefficients, gain and phase shift
-functions}\label{sec-an-cof}}
+\section{Coefficients, gain and phase shift functions}\label{sec-an-cof}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}
-
-}
-
-\end{figure}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}
-
-}
-
-\end{figure}
-
-\hypertarget{ann-ex-r}{%
-\section{R code example}\label{ann-ex-r}}
-
-In this appendix we present the \faIcon{r-project} code that can be used
-to estimate trend-cycle components for the US Employment with local
-polynomial filters.
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}
+
+}
+
+\end{figure}%
+
+\section{R code example}\label{ann-ex-r}
+
+In this appendix we present the R code that can be used to estimate
+trend-cycle components for the US Employment with local polynomial
+filters.
 

### 2024-04-02 paper/supp_material.qmd
paer folder

--- 2024-02-12 supp_material.qmd
@@
 
+```{r, out.width="100%"}
+#| label: fig-graphs-coef-lc-nn
+#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
+img <- sprintf("img/filters_used/%s_nn.%s", tolower("LC"), "pdf")
+knitr::include_graphics(img)
+```
+
+```{r, out.width="100%"}
+#| label: fig-graphs-coef-ql-nn
+#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
+img <- sprintf("img/filters_used/%s_nn.%s", tolower("QL"), "pdf")
+knitr::include_graphics(img)
+```
+
+```{r, out.width="100%"}
+#| label: fig-graphs-coef-cq-nn
+#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$ and a nearest neighbour bandwidth.
+img <- sprintf("img/filters_used/%s_nn.%s", tolower("CQ"), "pdf")
+knitr::include_graphics(img)
+```
+
+```{r, out.width="100%"}
+#| label: fig-graphs-coef-daf-nn
+#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF) with a nearest neighbour bandwidth.
+img <- sprintf("img/filters_used/%s_nn.%s", tolower("DAF"), "pdf")
+knitr::include_graphics(img)
+```
+
 \newpage
 
-# R code example {#ann-ex-r}
+<!-- # R code example {#ann-ex-r} -->
 
-In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.
+<!-- In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters. -->
 
-```{r, eval=FALSE, echo=TRUE}
-#| eval: false
-#| echo: true
-#| file: R_ex.R
-```
+<!-- ```{r, eval=FALSE, echo=TRUE} -->
+<!-- #| eval: false -->
+<!-- #| echo: true -->
+<!-- #| file: R_ex.R -->
+<!-- ``` -->

### 2024-04-02 R_filters/plot_coef.R

@@ -32,7 +32,7 @@ for(i in names(all_filters)){
       ggplot2::labs(y = "Phase shift") +
       ggplot2::guides(colour = "none"))
   p
-  ggsave(sprintf("img/filters_used/%s.pdf",tolower(i)),
+  ggsave(sprintf("paper/img/filters_used/%s.pdf",tolower(i)),
               width = 8, height = 4.5,
               plot = p)
 }
@@ -49,7 +49,7 @@ p <- (ggplot_coef(x, q=0) + ggplot2::guides(colour = "none") )/ (
 )
 p
 
-ggsave("img/filters_used/musgrave.pdf",
+ggsave("paper/img/filters_used/musgrave.pdf",
        width = 8, height = 4.5,
        plot = p)
 
@@ -77,7 +77,7 @@ p <- ggplot_coef(x, q = c(0:6)) / (
     ggplot2::labs(y = "Phase shift") +
     ggplot2::guides(colour = "none"))
 p
-ggsave("img/filters_used/nn_henderson.pdf",
+ggsave("paper/img/filters_used/nn_henderson.pdf",
        width = 8, height = 4.5,
        plot = p)
 
@@ -103,7 +103,7 @@ for(i in names(all_filters)){
       ggplot2::labs(y = "Phase shift") +
       ggplot2::guides(colour = "none"))
   p
-  ggsave(sprintf("img/filters_used/%s_nn.pdf",tolower(i)),
+  ggsave(sprintf("paper/img/filters_used/%s_nn.pdf",tolower(i)),
          width = 8, height = 4.5,
          plot = p)
 }

### 2024-04-02 R_filters/slope_concavity_estimator.R

@@ -70,5 +70,5 @@ p22 = plot_coef_ggplot(MM_d2_deriv2)
 p = (p21 + ggtitle("Slope estimators")) +
   (p22 + ggtitle("Concavity estimators"))
 p
-ggsave("img/filters_used/mm_penteconcavite.pdf",
+ggsave("paper/img/filters_used/mm_penteconcavite.pdf",
        width = 7,height = 3)

### 2024-04-02 R_fredm/4A_Analyse_est.R

@@ -15,7 +15,7 @@ all_plots <- get_all_plots(all_tp,
                            nb_est = 8)
 wrap_plots(all_plots, ncol = 3) & scale_color_grey()
 
-ggsave("img/nber/ce16ov_fev2001_lp.pdf",
+ggsave("paper/img/nber/ce16ov_fev2001_lp.pdf",
        plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 
@@ -25,7 +25,7 @@ all_plots <- get_all_plots(all_tp,
                            nb_est = 8)
 wrap_plots(all_plots, ncol = 3) & scale_color_grey()
 
-ggsave("img/nber/ce16ov_covid_lp.pdf",
+ggsave("paper/img/nber/ce16ov_covid_lp.pdf",
        plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 
@@ -37,7 +37,7 @@ all_plots <- get_all_plots(all_tp,
                            nb_est = 8)
 wrap_plots(all_plots, ncol = 3) & scale_color_grey()
 
-ggsave("img/nber/retailx_nov2007_lp.pdf",
+ggsave("paper/img/nber/retailx_nov2007_lp.pdf",
        plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 

### 2024-04-02 R_fredm/4B_Analyse_implicit_forecast_dt.R

@@ -23,7 +23,7 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              dossier = "results_fredm/compile_tp_norev/")
 wrap_plots(plots,ncol = 3) & scale_color_grey()
 
-ggsave("img/nber/ce16ov_fev2001_prev_imp_lp.pdf",
+ggsave("paper/img/nber/ce16ov_fev2001_prev_imp_lp.pdf",
        plot = wrap_plots(plots, ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 
@@ -41,7 +41,7 @@ wrap_plots(plots,ncol = 3) &
   scale_color_grey() & 
   labs(subtitle = NULL)
 
-ggsave("img/nber/ce16ov_covid_prev_imp_lp.pdf",
+ggsave("paper/img/nber/ce16ov_covid_prev_imp_lp.pdf",
        plot = wrap_plots(plots, ncol = 3) & 
          scale_color_grey() & 
          labs(subtitle = NULL),
@@ -65,6 +65,6 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              fst_degree = fst_degree)
 wrap_plots(plots, ncol = 3)
 
-ggsave("img/nber/retailx_nov2007_prev_imp_lp.pdf",
+ggsave("paper/img/nber/retailx_nov2007_prev_imp_lp.pdf",
        plot = wrap_plots(plots, ncol = 3),
        width = 8, height = 8)

### 2024-04-02 R_simul/1B_plot_data.R

@@ -27,7 +27,7 @@ p <- ggplot(data = dataGraph,
                     legend.key = element_blank(),
                     legend.title = element_blank())
 
-ggsave("img/simulations/simul_data.pdf",
+ggsave("paper/img/simulations/simul_data.pdf",
        plot = p,
        width = 8, height = 4)
 

### 2024-04-02 R_simul/4A_Analyse_tp.R

@@ -109,7 +109,7 @@ p <- ggplot(all_tp %>%
   theme(axis.text.x = element_text( margin = margin(10, 0, 0, 0), vjust = 1))
 p
 
-ggsave("img/simulations/phase_shift_simul.pdf",
+ggsave("paper/img/simulations/phase_shift_simul.pdf",
             plot = p,
             width = 10, height = 6)
 legende <- c(
@@ -132,7 +132,7 @@ p <- ggplot(data_tp %>%
   scale_x_discrete(labels = legende)
 p
 
-ggsave("img/simulations/phase_shift_simul_nn_lp.pdf",
+ggsave("paper/img/simulations/phase_shift_simul_nn_lp.pdf",
        plot = p,
        width = 10, height = 6)
 
@@ -156,7 +156,7 @@ p <- ggplot(data_tp %>%
   scale_x_discrete(labels = legende)
 p
 
-ggsave("img/simulations/phase_shift_simul_nn_lp_localparam.pdf",
+ggsave("paper/img/simulations/phase_shift_simul_nn_lp_localparam.pdf",
        plot = p,
        width = 10, height = 6)
 
@@ -185,7 +185,7 @@ p <- ggplot(data_tp %>%
   labs(y="Phase shift", x = NULL) +
   scale_x_discrete(labels = legende) 
 p
-ggsave("img/simulations/phase_shift_simul_arima_length.pdf",
+ggsave("paper/img/simulations/phase_shift_simul_arima_length.pdf",
        plot = p,
        width = 10, height = 6)
 
@@ -205,7 +205,7 @@ p_kernel <- ggplot(data_tp_kernel %>% filter(method == "lc"),
   labs(y="Phase shift", x = NULL)
 p_kernel
 
-ggsave("img/simulations/phase_shift_kernel.pdf",
+ggsave("paper/img/simulations/phase_shift_kernel.pdf",
        plot = p_kernel,
        width = 10, height = 6)
 

### 2024-04-02 R_simul/4B_analyse_revisions.R

@@ -60,4 +60,4 @@ rev_table <- rev_tot %>% dplyr::filter(variability == "mediumvariability")%>%
     sprintf("$q=%s$", gsub("rev.q","",x))
   }) %>%
   rename(`Method` = method)
-saveRDS(rev_table, file = "data/simulations_revisions.RDS")
+saveRDS(rev_table, file = "paper/data/simulations_revisions.RDS")

### 2024-04-02 R_simul_quarter/1B_plot_data.R

@@ -27,7 +27,7 @@ p <- ggplot(data = dataGraph,
                     legend.key = element_blank(),
                     legend.title = element_blank())
 p
-ggsave("img/simulations/simul_data_q",
+ggsave("paper/img/simulations/simul_data_q",
        width = 8, height = 4)
 
 

### 2024-04-02 R_simul_quarter/4A_Analyse_tp.R

@@ -77,7 +77,7 @@ p <- ggplot(data_tp %>%
 p
 
 
-ggsave("img/simulations/phase_shift_simul.pdf",
+ggsave("paper/img/simulations/phase_shift_simul.pdf",
             plot = p,
             width = 10, height = 6)
 

### 2024-04-02 R_simul_quarter/4B_analyse_revisions.R

@@ -61,4 +61,4 @@ rev_table <- rev_tot %>% dplyr::filter(variability == "mediumvariability")%>%
   }) %>%
   rename(`Method` = method)
 rev_table
-saveRDS(rev_table, file = "data/simulations_revisions_q.RDS")
+saveRDS(rev_table, file = "paper/data/simulations_revisions_q.RDS")

### 2024-04-02 ex_pente.R

@@ -84,7 +84,7 @@ p <- autoplot(data[,c(1,2)]) +
   )
 
 p
-ggsave("img/filters_used/mm_penteconcavite_ex.pdf",
+ggsave("paper/img/filters_used/mm_penteconcavite_ex.pdf",
             plot = p,
             width = 7,height = 2)
 
@@ -109,6 +109,6 @@ p2 <- autoplot(data[,c(1,3)]) +
   )
 p2
 
-ggsave("img/filters_used/mm_penteconcavite_ex2.pdf",
+ggsave("paper/img/filters_used/mm_penteconcavite_ex2.pdf",
        plot = p2,
        width = 7,height = 2)

### 2024-04-02 paper/resp_referee.qmd
update

--- 2024-03-28 resp_referee.qmd
@@
 format: pdf
-editor: visual
 ---
@@
 
-• Section 1.1. It would perhaps be useful to provide the expression
+- Section 1.1. It would perhaps be useful to provide the expression
 of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of
 the transfer function of the linear filter.
-• Section 1.1. The symbol p for the period is already in use from the
+- Section 1.1. The symbol p for the period is already in use from the
 previous section, where it denotes the bandwidth and number of
 past observations used by the filter.
-• Page 4. The definition of the high-pass component (oscillations
+- Page 4. The definition of the high-pass component (oscillations
 in the frequency range from $\pi/6$ to $\pi$ as undesirable oscillations is
 controversial.
-• Figure 1. In the caption, the ration $I/C$ was never defined before.
+- Figure 1. In the caption, the ration $I/C$ was never defined before.
 The description of the three panels should be better explained in
 the main text.
-• Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
+- Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
 assumes that the reader is knowledgeable about seasonal adjustment
@@
 component.
-• Page 7, line 38. ’which extends’ is repeated twice.
-• Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts
+- Page 7, line 38. ’which extends’ is repeated twice.
+- Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts
 induced by $w_0$, etc.? The forecasts are obtained from an assumed
 polynomial trend model?
-• Table 1. Some criteria are labelled as Guggemos, not referenced in
+- Table 1. Some criteria are labelled as Guggemos, not referenced in
 the main text.
-• Figures 6-7. The labels of the time axes and the months are in
+- Figures 6-7. The labels of the time axes and the months are in
 French.
@@
 > The author avoids numbering the Introduction and Conclusion; numbering both (and adjusting the numbering of the other sections accordingly) may be better.
+
+Those sections are now numbered
 
@@
 
+> 8.  Consideration of Different Series Periodicity
+> It would be beneficial if the authors could address the suggestion of considering series
+with different periodicities in the real data application. Discussing how the proposed
+methodology performs when applied to series with various frequencies would add
+depth to the study.
+
 > 5.  Clarity on Phase Shift Computation (Page 20)
@@
 
-> 8.  Consideration of Different Series Periodicity
-> It would be beneficial if the authors could address the suggestion of considering series
-with different periodicities in the real data application. Discussing how the proposed
-methodology performs when applied to series with various frequencies would add
-depth to the study.
-
 # Referee 4
@@
 I found the contribution of the paper clearly stated on page 16, but I think I should come before in the text.
+
 2) The paper omits to consider the Covid19 period and how this influences the trend
@@
 to address it, e.g. outlier correction, robust distributions, etc.
-3) I think a quote from other works can be removed from the paper. For example, I
+
+> 3) I think a quote from other works can be removed from the paper. For example, I
 would remove the quote in point 2) of page 7.
+
+Removed

### 2024-04-15 paper/article_jos.qmd
small updates

--- 2024-03-28 article_jos.qmd
@@
 This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
-This is also the approach used in the X-13ARIMA seasonal adjustment method, which extends the series by one year with an ARIMA model.
+A similar approach is used in the X-13ARIMA seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters
 
@@
 
-### Quality indicators for asymmetric moving averages {#subsec:crit-qual}
+<!-- ### Quality indicators for asymmetric moving averages {#subsec:crit-qual} -->
 
-For asymmetric filters, most of the criteria come from those defined by @ch15HBSA and @trilemmaWMR2019 to build asymmetric filters. 
-They are summarised in Table\ \ref{tab-QC} and can be calculated using the `rjd3filters::diagnostic_matrix()` function.
+<!-- For asymmetric filters, most of the criteria come from those defined by @ch15HBSA and @trilemmaWMR2019 to build asymmetric filters.  -->
+<!-- They are summarised in @tbl-qc and can be calculated using the `rjd3filters::diagnostic_matrix()` function. -->
 
-@ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$).
-Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low. 
-This method can be used in R with the function `rjd3filters::fst_filter()`.
+<!-- @ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$). -->
+<!-- Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low.  -->
+<!-- This method can be used in R with the function `rjd3filters::fst_filter()`. -->
 
-@trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
-A simplified version of this method can be used in R with the function `rjd3filters::dfa_filter()`.
+<!-- @trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
+-->
+<!-- A simplified version of this method can be used in R with the function `rjd3filters::dfa_filter()`. -->
 
-```{r, include=FALSE}
-title <- "Quality criteria for a moving average $\\boldsymbol\\theta=(\\theta_k)_{-p\\leq k\\leq f}$ defined by a gain function $\\rho_{\\boldsymbol\\theta}$ and a phase shift function $\\varphi_{\\boldsymbol\\theta}$."
-footnotes <- c("$X_g$ criteria from Grun-Rehomme, Guggemos, and Ladiray (2018) and $X_w$ criteria from Wildi and McElroy (2019).",
-               "$\\rho_s$ and  $\\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.",
-               "$h$ e is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\\frac{1}{2(1-\\cos(x))}$)."
-)
-```
+<!-- ```{r, include=FALSE} -->
+<!-- title <- "Quality criteria for a moving average $\\boldsymbol\\theta=(\\theta_k)_{-p\\leq k\\leq f}$ defined by a gain function $\\rho_{\\boldsymbol\\theta}$ and a phase shift function $\\varphi_{\\boldsymbol\\theta}$." -->
+<!-- footnotes <- c("$X_g$ criteria from Grun-Rehomme, Guggemos, and Ladiray (2018) and $X_w$ criteria from Wildi and McElroy (2019).", -->
+<!--                "$\\rho_s$ and  $\\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.", -->
+<!--                "$h$ e is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\\frac{1}{2(1-\\cos(x))}$)." -->
+<!-- ) -->
+<!-- ``` -->
 
-```{r, echo = FALSE, fig.note = footnotes}
-#| label: tab-QC
-#| eval: false
-criteria_mm <- data.frame(Sigle = c("b_c", "b_l", "b_q",
-                                    "F_g", "S_g", "T_g",
-                                    "A_w", "T_w", "S_w", "R_w"),
-                          Description = c(sprintf("%s bias", c("Constant", "Linear", "Quadratic")),
-                                          sprintf("%s (Guggemos)",
-                                                  c("Variance reduction / Fidelity", "Smoothness", "Timeliness")),
-                                          sprintf("%s (Wildi)",c("Accuracy", "Timeliness", "Smoothness", "Residual"))),
-                          Formule = c("\\sum_{k=-p}^{+f}\\theta_{k}-1",
-                                      "\\sum_{k=-p}^{+f}k\\theta_{k}",
-                                      "\\sum_{k=-p}^{+f}k^{2}\\theta_{k}",
-                                      "\\sum_{k=-p}^{+f}\\theta_{k}^{2}",
-                                      "\\sum_{j}(\\nabla^{3}\\theta_{j})^{2}",
-                                      "\\int_{0}^{\\omega_1}\\rho_{\\boldsymbol\\theta}(\\omega)\\sin(\\varphi_{\\boldsymbol\\theta}(\\omega))^{2}\\ud\\omega",
-                                      "2\\int_0^{\\omega_1}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega",
-                                      "8\\int_0^{\\omega_1} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega",
-                                      "2\\int_{\\omega_1}^{\\pi}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega",
-                                      "8\\int_{\\omega_1}^{\\pi} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega"))
-criteria_mm[,1] <- sprintf("$%s$", criteria_mm[,1])
-criteria_mm[,3] <- sprintf("$%s$", criteria_mm[,3])
-library(kableExtra)
+<!-- ::: {#tbl-qc} -->
+<!-- ```{r} -->
+<!-- #| echo: false -->
+<!-- criteria_mm <- data.frame(Sigle = c("b_c", "b_l", "b_q", -->
+<!--                                     "F_g", "S_g", "T_g", -->
+<!--                                     "A_w", "T_w", "S_w", "R_w"), -->
+<!--                           Description = c(sprintf("%s bias", c("Constant", "Linear", "Quadratic")), -->
+<!--                                           sprintf("%s (Grun-Rehomme et al)", -->
+<!--                                                   c("Variance reduction / Fidelity", "Smoothness", "Timeliness")), -->
+<!--                                           sprintf("%s (Wildi and McElroy)",c("Accuracy", "Timeliness", "Smoothness", "Residual"))), -->
+<!--                           Formule = c("\\sum_{k=-p}^{+f}\\theta_{k}-1", -->
+<!--                                       "\\sum_{k=-p}^{+f}k\\theta_{k}", -->
+<!--                                       "\\sum_{k=-p}^{+f}k^{2}\\theta_{k}", -->
+<!--                                       "\\sum_{k=-p}^{+f}\\theta_{k}^{2}", -->
+<!--                                       "\\sum_{j}(\\nabla^{3}\\theta_{j})^{2}", -->
+<!--                                       "\\int_{0}^{\\omega_1}\\rho_{\\boldsymbol\\theta}(\\omega)\\sin(\\varphi_{\\boldsymbol\\theta}(\\omega))^{2}\\ud\\omega", -->
+<!--                                       "2\\int_0^{\\omega_1}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega", -->
+<!--                                       "8\\int_0^{\\omega_1} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega", -->
+<!--                                       "2\\int_{\\omega_1}^{\\pi}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega", -->
+<!--                                       "8\\int_{\\omega_1}^{\\pi} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega")) -->
+<!-- criteria_mm[,1] <- sprintf("$%s$", criteria_mm[,1]) -->
+<!-- criteria_mm[,3] <- sprintf("$%s$", criteria_mm[,3]) -->
 
-criteria_mm %>%
-  kbl(align = "c", booktabs = T, row.names = FALSE,
-        escape = FALSE,caption = title)  %>%
-kable_styling(latex_options = "scale_down") 
-    # add_footnote_kable(stop_centering = TRUE)
-```
+<!-- criteria_mm %>% -->
+<!--   kbl(align = "c", booktabs = T, row.names = FALSE, -->
+<!--         escape = FALSE,caption = title)  %>% -->
+<!-- kable_styling(latex_options = "scale_down")  -->
+<!--     # add_footnote_kable(stop_centering = TRUE) -->
+<!-- ``` -->
 
-```{=tex}
-\begin{table}[!h]
-\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.55$.}
-{\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{ccc}
-\toprule
-Code & Description & Formula\\
-\midrule
-$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
-$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
-$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
-$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
-$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
-\addlinespace
-$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
-$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-\bottomrule
-\end{tabular}}}
+<!-- *Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).* -->
+<!-- *$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.* -->
+<!-- *$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).* -->
 
-\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}
-
-\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}
-
-\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
-\end{table}
-```
+<!-- Quality criteria for a moving average $\boldsymbol\theta=(\theta_k)_{-p\leq k\leq f}$ defined by a gain function $\rho_{\boldsymbol\theta}$ and a phase shift function $\varphi_{\boldsymbol\theta}$.
+-->
+<!-- ::: -->
 
@@
 $$
-\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
+\forall t \::\:TC_t = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
 \frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
@@
 
-Table\ \ref{tab-criteriaLp} compares the quality criteria of the different methods using the Henderson filter and $h=6$ (a symmetric filter with 13 terms). 
-For real-time filters ($q=0$), a more complex asymmetric filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance.
-This result varies as $q$ increases: for $q=2$, the QL filter has greater timeliness than the LC filter. 
-This surprising result highlights the fact that the phase shift is not controlled by @proietti2008's approach.
+<!-- Table\ \ref{tab-criteriaLp} compares the quality criteria of the different methods using the Henderson filter and $h=6$ (a symmetric filter with 13 terms).  -->
+<!-- For real-time filters ($q=0$), a more complex asymmetric filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance.
+-->
+<!-- This result varies as $q$ increases: for $q=2$, the QL filter has greater timeliness than the LC filter.  -->
+<!-- This surprising result highlights the fact that the phase shift is not controlled by @proietti2008's approach. -->
 
-In terms of revision ($A_w+S_w+T_w+R_w$), the LC and QL filters consistently outperform the CQ and DAF filters.
+<!-- In terms of revision ($A_w+S_w+T_w+R_w$), the LC and QL filters consistently outperform the CQ and DAF filters.  -->
 
-These theoretical properties are consistent with the empirical results observed in @sec-comparison.
+<!-- These theoretical properties are consistent with the empirical results observed in @sec-comparison.
 Revisions are more important for the CQ and DAF filters, which leads to greater variability in estimations and a longer delay in detecting turning points.
-The globally parametrised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter.
-Indeed, to identify a turning point at date $t$, it is necessary to have knowledge of at least two points after the given date to confirm the reversal in trend.
+-->
+<!-- The globally parametrised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter. -->
+<!-- Indeed, to identify a turning point at date $t$, it is necessary to have knowledge of at least two points after the given date to confirm the reversal in trend. -->
 
-```{r echo = FALSE}
-#| label: tab-criteraLp
-#| eval: false
-#| tab-cap: 
-title <- "Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.5$."
+<!-- ::: {#tbl-criteraLp} -->
+<!-- ```{r echo = FALSE} -->
+<!-- #| echo: false -->
+<!-- lp_diagnostics <- readRDS("data/lp_diagnostics_henderson.RDS") -->
+<!-- lp_diagnostics[,"$ EQM_w $"] <- rowSums(lp_diagnostics[,grep("_w", colnames(lp_diagnostics), fixed = TRUE)]) -->
+<!-- colnames(lp_diagnostics) <- gsub(" ?\\$ ?","$",colnames(lp_diagnostics)) -->
+<!-- lp_diagnostics[,1] <- gsub(" ?\\$ ?","$",lp_diagnostics[,1]) -->
+<!-- groupement <- table(lp_diagnostics[,1]) -->
+<!-- lp_diagnostics[,-(1:2)] <- round(lp_diagnostics[,-(1:2)],2) -->
+<!-- lp_diagnostics[,-1] %>% -->
+<!--   kable(format.args = list(digits = 2, -->
+<!--                            decimal.mark = "."), align = "c", booktabs = T, row.names = FALSE, -->
+<!--         escape = FALSE,caption = title,format = "latex") %>% -->
+<!--   kable_styling(latex_options=c(#"striped", -->
+<!--       "hold_position")) %>% -->
+<!--   pack_rows(index = groupement, escape = FALSE) %>% -->
+<!--   footnote("With $EQM_w=A_w + S_w + T_w + R_w$.") -->
+<!-- ``` -->
 
-lp_diagnostics <- readRDS("data/lp_diagnostics_henderson.RDS")
-lp_diagnostics[,"$ EQM_w $"] <- rowSums(lp_diagnostics[,grep("_w", colnames(lp_diagnostics), fixed = TRUE)])
-colnames(lp_diagnostics) <- gsub(" ?\\$ ?","$",colnames(lp_diagnostics))
-lp_diagnostics[,1] <- gsub(" ?\\$ ?","$",lp_diagnostics[,1])
-groupement <- table(lp_diagnostics[,1])
-lp_diagnostics[,-(1:2)] <- round(lp_diagnostics[,-(1:2)],2)
-lp_diagnostics[,-1] %>%
-  kable(format.args = list(digits = 2,
-                           decimal.mark = "."), align = "c", booktabs = T, row.names = FALSE,
-        escape = FALSE,caption = title,format = "latex") %>%
-  kable_styling(latex_options=c(#"striped",
-      "hold_position")) %>%
-  pack_rows(index = groupement, escape = FALSE) %>%
-  footnote("With $EQM_w=A_w + S_w + T_w + R_w$.")
-```
+<!-- Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.5$. -->
+<!-- ::: -->
 
-```{=tex}
-\begin{table}[!h]
-\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.55$.}
-\centering
-\begin{tabular}[t]{cccccccccccc}
-\toprule
-Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=0$}}\\
-\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
-\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=1$}}\\
-\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
-\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=2$}}\\
-\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
-\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
-\bottomrule
-\end{tabular}
+<!-- ```{=tex} -->
+<!-- \begin{table}[!h] -->
+<!-- \caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.55$.} -->
+<!-- \centering -->
+<!-- \begin{tabular}[t]{cccccccccccc} -->
+<!-- \toprule -->
+<!-- Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\ -->
+<!-- \midrule -->
+<!-- \addlinespace[0.3em] -->
+<!-- \multicolumn{12}{l}{\textbf{$q=0$}}\\ -->
+<!-- \hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\ -->
+<!-- \hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\ -->
+<!-- \hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\ -->
+<!-- \hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\ -->
+<!-- \addlinespace[0.3em] -->
+<!-- \multicolumn{12}{l}{\textbf{$q=1$}}\\ -->
+<!-- \hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\ -->
+<!-- \hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\ -->
+<!-- \hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\ -->
+<!-- \hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\ -->
+<!-- \addlinespace[0.3em] -->
+<!-- \multicolumn{12}{l}{\textbf{$q=2$}}\\ -->
+<!-- \hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\ -->
+<!-- \hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\ -->
+<!-- \hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\ -->
+<!-- \hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\ -->
+<!-- \bottomrule -->
+<!-- \end{tabular} -->
 
-\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
-\end{table}
-```
+<!-- \textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.} -->
+<!-- \end{table} -->
+<!-- ``` -->
 
@@
 
-2.
-Computation of two revision criteria: the average of the relative deviations between the $q$^th^ estimate and the last estimate $MAE_{fe}(q)$ and the average of the relative deviations between the $q$^th^ estimate and the $q+1$^th^ estimate $MAE_{ce}(q)$.
+2. The average of the relative deviations between to the last estimate (comparison of the $q$^th^ estimate and the last estimate):
 $$
@@
 }\right|
-\right]
-\quad\text{and}\quad
+\right].
+$$
+3. The average of the relative deviations between two consecutive estimates (the $q$^th^ estimate and the $q+1$^th^ estimate):
+$$
 MAE_{ce}(q)=\mathbb E\left[
@@
 ```
-
-# R code example {#sec-ann-ex-r}
-
-In this appendix we present the R code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters.
-
-```{r, eval=FALSE, echo=TRUE}
-#| eval: false
-#| echo: true
-#| file: R_ex.R
-```

### 2024-04-15 paper/resp_referee.qmd
small updates

@@
 
-1. Page 3, line 46: you say that it is "equivalently defined" and then provide the conjugate,
-writing $e^{i\omega k}$ instead of $e^{-i\omega k}$. "Equivalence" implies something akin to equality, so I would soften this to "alternatively", since the two expressions are not equal.
+> 1. Page 3, line 46: you say that it is "equivalently defined" and then provide the conjugate, writing $e^{i\omega k}$ instead of $e^{-i\omega k}$.  
+> "Equivalence" implies something akin to equality, so I would soften this to "alternatively", since the two expressions are not equal.
 
-2. Page 5, line 53: replace "have" with "design", I think, because you are constructing these
-filters, right?
+*It was corrected.*
 
-3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function
-at $\omega = 0$ yield the conditions. So the results could be stated with references, and the
-derivations omitted, and this subsection compactified.
+> 2. Page 5, line 53: replace "have" with "design", I think, because you are constructing these filters, right?
 
-4. Page 7, line 40: why say the series is extended by one year? The extensions are for as many
-years as needed by the particular filters (typically, half their length in years).
+*Indeed, it was corrected.*
 
-5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.
+> 3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function 
+> at $\omega = 0$ yield the conditions. So the results could be stated with references, and the 
+> derivations omitted, and this subsection compactified.
 
-6. Table 1: I repeat the referees here. The table needs more explication in the text. What is $\omega_1$? Define $\nabla$, Guggemos, etc.
+> 4. Page 7, line 40: why say the series is extended by one year? The extensions are for as many 
+> years as needed by the particular filters (typically, half their length in years).
 
-7. Page 13, equation (4): it is hard to understand what $\theta_f$ and $\theta_p$ are, and what 
-$\gamma$ and $\delta$ are.
-Add more explanation to the text.
+> 5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.
 
-8. Page 15, line 36: should be "same length as $v$"
+> 6. Table 1: I repeat the referees here. The table needs more explication in the text. What is $\omega_1$? Define $\nabla$, Guggemos, etc.
 
-9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness?
-That is, when you display the entries of the matrix T, you need to make a connection between
-those values and the concept of timeliness.
+*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed *
 
-10. Page 15, line 53: I don't see why $v$ in this equation should be defined through the case bracket;
-everything starting with "with" could be moved to the following text.
+> 7. Page 13, equation (4): it is hard to understand what $\theta_f$ and $\theta_p$ are, and what $\gamma$ and $\delta$ are. 
+> Add more explanation to the text.
 
-11. Page 17, lines 53-55: This discussion needs more clarity; it seems a bit terse.
+> 8. Page 15, line 36: should be "same length as $v$"
 
-12. Appendix B: rather than typing out the R code, it would be better to have the files available
-as supplementary material. (Or possibly put in a github repo and provide the links in the
-text.)
+> 9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness? 
+> That is, when you display the entries of the matrix T, you need to make a connection between 
+> those values and the concept of timeliness.
+
+> 10. Page 15, line 53: I don't see why $v$ in this equation should be defined through the case bracket; 
+> everything starting with "with" could be moved to the following text.
+
+> 11. Page 17, lines 53-55: This discussion needs more clarity; it seems a bit terse.
+
+> 12. Appendix B: rather than typing out the R code, it would be better to have the files available 
+> as supplementary material. (Or possibly put in a github repo and provide the links in the 
+> text.)
 
@@
 
-The authors should provide a clear definition of trend-cycle, $TC_t$,
-and irregular, $I_t$. In other words, how is the decomposition
-$y_t =TC_t+I_t$ identified? In section 1 they seem to hint at the interpretation
-of $TC_t$ as the low-pass component of yt with cutoff frequency
-$\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component).
-This point should be clarified, I think.
-Related questions are: are $y_t$ and its components stochastic or deterministic?
-The treatment in section 1.2.1 suggests that $TC_t$ is
-an unknown possibly nonlinear deterministic function of time.
-The variance inflation factor $\sum_k \theta_k^2$ implies that $I_t$ is white noise.
+> 1. The authors should provide a clear definition of trend-cycle, $TC_t$, 
+> and irregular, $I_t$. In other words, how is the decomposition 
+> $y_t =TC_t+I_t$ identified? In section 1 they seem to hint at the interpretation 
+> of $TC_t$ as the low-pass component of yt with cutoff frequency 
+> $\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component). 
+> This point should be clarified, I think. 
+> Related questions are: are $y_t$ and its components stochastic or deterministic? 
+> The treatment in section 1.2.1 suggests that $TC_t$ is 
+> an unknown possibly nonlinear deterministic function of time. 
+> The variance inflation factor $\sum_k \theta_k^2$ implies that $I_t$ is white noise.
 
-2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not
-very formal. Do you mean that the unconditional variance is minimized?
-Or the long run variance?
+> 2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not
+> very formal. Do you mean that the unconditional variance is minimized?
+> Or the long run variance?
 
-3. A definition of turning point would also be needed. Are these ’regular’
+> 3. A definition of turning point would also be needed. Are these ’regular’
 points at which the slope of the trend changes sign? Or changepoints?
 
-4. The notation should be unified: why not writing $TC_t = \mu(t)$ and
-in section 2 using $t$ instead of $x$? The observation index i hould
-be replaced by $t$, as in section 2.1. Similarly, why not using ϵt
-directly and start section 1.2 by stating that you assume that yt is
-generated as $y_t = \mu(t) + \varepsilon_t$, etc.
-Secondly, you should clarify the meaning of (3), where you set
-$\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix
-$D$ for later treatment?
-5. You should state that you assume a fixed bandwidth. A nearest
-neighbour bandwidth should be interesting to consider: as such
-it would contribute more to the reduction of the variance at the
-expenses of the phase shift.
-6. Is the phase shift related to the notion of bias in local polynomial
+> 4. The notation should be unified: why not writing $TC_t = \mu(t)$ and
+> in section 2 using $t$ instead of $x$? The observation index i hould
+> be replaced by $t$, as in section 2.1. Similarly, why not using ϵt
+> directly and start section 1.2 by stating that you assume that yt is
+> generated as $y_t = \mu(t) + \varepsilon_t$, etc.
+> Secondly, you should clarify the meaning of (3), where you set
+> $\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix $D$ for later treatment?
+
+> 5. You should state that you assume a fixed bandwidth. 
+A nearest neighbour bandwidth should be interesting to consider: as such it would contribute more to the reduction of the variance at the expenses of the phase shift.
+
+> 6. Is the phase shift related to the notion of bias in local polynomial
 smoothing?
@@
 
-- Section 1.1. It would perhaps be useful to provide the expression
+> - Section 1.1. It would perhaps be useful to provide the expression
 of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of
 the transfer function of the linear filter.
-- Section 1.1. The symbol p for the period is already in use from the
+
+> - Section 1.1. The symbol $p$ for the period is already in use from the
 previous section, where it denotes the bandwidth and number of
 past observations used by the filter.
-- Page 4. The definition of the high-pass component (oscillations
+
+> - Page 4. The definition of the high-pass component (oscillations
 in the frequency range from $\pi/6$ to $\pi$ as undesirable oscillations is
 controversial.
-- Figure 1. In the caption, the ration $I/C$ was never defined before.
+
+> - Figure 1. In the caption, the ration $I/C$ was never defined before.
 The description of the three panels should be better explained in
 the main text.
-- Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
+
+> - Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
 assumes that the reader is knowledgeable about seasonal adjustment
@@
 component.
-- Page 7, line 38. ’which extends’ is repeated twice.
-- Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts
-induced by $w_0$, etc.? The forecasts are obtained from an assumed
-polynomial trend model?
-- Table 1. Some criteria are labelled as Guggemos, not referenced in
-the main text.
-- Figures 6-7. The labels of the time axes and the months are in
+
+> - Page 7, line 38. ’which extends’ is repeated twice.
+
+> - Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts induced by $w_0$, etc.? 
+The forecasts are obtained from an assumed polynomial trend model?
+
+> - Table 1. Some criteria are labelled as Guggemos, not referenced in the main text.
+
+*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed *
+
+> - Figures 6-7. The labels of the time axes and the months are in
 French.
@@
 
-Those sections are now numbered
+*Those sections are now numbered*
 
@@
 
-> 8.  Consideration of Different Series Periodicity
-> It would be beneficial if the authors could address the suggestion of considering series
-with different periodicities in the real data application. Discussing how the proposed
-methodology performs when applied to series with various frequencies would add
-depth to the study.
-
 > 5.  Clarity on Phase Shift Computation (Page 20)
@@
 
+> 8.  Consideration of Different Series Periodicity
+> It would be beneficial if the authors could address the suggestion of considering series
+with different periodicities in the real data application. Discussing how the proposed
+methodology performs when applied to series with various frequencies would add
+depth to the study.
+
 # Referee 4
@@
 
-Removed
+*The quote was removed*

### 2024-04-15 R_fredm/4_utils.R

@@ -357,7 +357,7 @@ get_all_plots <- function(all_tp, all_tp_rev = NULL, series, tp_keep, tp_plot = 
   if (!is.null(all_tp_rev)) {
     all_tp_rev_plot <-
       all_tp_rev %>% dplyr::filter(series %in% !!series) %>%
-      mutate(subtitle = sprintf("\n(1rst detection in %i months)",
+      mutate(subtitle = sprintf("\n(1st detection in %i months)",
                                 round(.data[[tp_keep_col]]))
       )
     all_tp_plot[, "subtitle"] <-
@@ -415,7 +415,7 @@ get_all_plots_prevs <- function(data_prevs,
   if (!is.null(all_tp_rev)) {
     all_tp_rev_plot <-
       all_tp_rev %>% dplyr::filter(series %in% !!series) %>%
-      mutate(subtitle = sprintf("\n(1rst detection in %i months)",
+      mutate(subtitle = sprintf("\n(1st detection in %i months)",
                                 round(.data[[tp_keep_col]]))
       )
     all_tp_plot[, "subtitle"] <-

### 2024-04-17 paper/article_jos.qmd
improvement

@@
     cite-method: biblatex
+crossref:
+  tbl-prefix: table
+  fig-prefix: figure
+  sec-prefix: section
 linestretch: 2
@@
 Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
-For example, the X-13ARIMA seasonal adjustment method uses several moving averages to estimate the main components of a time series. 
+For example, the X-11 decomposition method (used in X-13ARIMA-SEATS) uses several moving averages to estimate the main components of a time series. 
 Symmetric filters are applied to the center of the series, but when it comes to estimate the most recent points, all of these methods must rely on asymmetric filters. 
-For trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-13ARIMA seasonal adjustment algorithm.
+For trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.
 
 However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
-For example, for trend-cycle extraction, X-13ARIMA uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
+For example, for trend-cycle extraction, X-11 uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
 Since the predicted values are linear combinations of past values, this amounts to applying asymmetric averages at the end of the series.
@@
 The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
-The extensions have been implemented in the R `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and are therefore easily reusable.
+The extensions have been implemented in the R `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and can therefore be easily used.
 
@@
 where  $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
-This function is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure the phase shift in number of periods.
+We this have:
+\begin{align*}
+\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
+\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
+\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
+&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
+\end{align*}
+The phase shift is sometimes represented as $\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}$ to measure it in number of periods.
 For all symmetric moving averages we have $\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})$.
@@
 
-To decompose a time series into a seasonal component, a trend-cycle and the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses a succession of moving averages, all with specific constraints.
+To decompose a time series into a seasonal component, a trend-cycle and the irregular, the X-11 decomposition algorithm (used in X-13ARIMA-SEATS) uses a succession of moving averages, all with specific constraints.
 
@@
  \text{ and }
-\forall j \in \left\llbracket 1,d\right\rrbracket:\:
+\forall j \in \left\{1,2,\dots,d\right\}:\:
 \sum_{k=-p}^{+f}k^j\theta_k=0.
@@
 This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
-A similar approach is used in the X-13ARIMA seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters
+A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters
 
@@
 This allows us to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
-Let us denote $v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average and $w^0,\dots w^{h-1}$ a sequence of asymmetric moving averages, of order $h+1$ to $2h$ used to estimate the last $h$ points with, for convention, $w_t^q=0$ for $t>q$.
-This means that $w^0=(w_{-h}^0,\dots, w_{0}^0)$ is used for estimation in real time (when no points are known in the future), $w^1=(w_{-h}^1,\dots, w_{1}^1)$ for estimation of the penultimate point (when only one point is known in the future), and so on.
-Let also note $y_{-h},\dots,y_{0}$ the observed series studied and $y_{1}^*,\dots y_h^*$ the implicit forecast induced by $w^0,\dots w^{h-1}$.
-This means that for all $q$ we have:
+Let us denote $\boldsymbol v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average (used to compute the final estimates) and $\boldsymbol w^0,\dots, \boldsymbol w^{h-1}$ a sequence of asymmetric moving averages used for the intermediate estimates.
+$\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)$ is used for estimation in real time estimate (when 0 point in the future is known), $\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)$ is used for estimation of the penultimate point (when only 1 in the future is known), and so on.
+Let $y_{t-h},\dots,y_{t}$ be the observed series during the last $h$ periods (for which asymmetric moving averages are used to get the intermediate trend-cycle estimates for the dates $t-h$ to $h$). 
+The implicit forecast $y_{t+1}^*,\dots y_{t+h}^*$ induced by $w^0,\dots w^{h-1}$ are the forecasts of the series $y_t$ for which the estimates of the symmetric filter applied to the extended series $\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)$ produce the same estimates as the asymmetric filter applied to the series $\boldsymbol y^*$.
+This means:
+\begin{align*}
+\underbrace{\sum_{i=-h}^0 v_iy_{ti} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
+\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ points after }t} \\
+&=
+\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ points after }t} \\
+&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
+&= \cdots \\
+&=\underbrace{\sum_{i=-h}^0 w_i^{h-1}y_{t+i} + \sum_{i=1}^{h-1} w_i^{h-1}y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^{h-1}\text{ using }h-1\text{ points after }t}
+\end{align*}
+For any integer $q$ between 0 and $h-1$, using for convention $w_t^q=0$ if $t>q$ we then have:
 $$
-\underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
-=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
-$$
-This is equivalent to:
-$$
-\quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
+\sum_{i=1}^h (v_i- w_i^q) y_i^*
 =\sum_{i=-h}^0 (w_i^q-v_i)y_i.
 $$
-To sum up, in matrix form, this is equivalent to solving:
+In matrix form, this is equivalent to solving:
 $$
@@
 $$
-\forall j\in\left\llbracket -h,h\right\rrbracket:\:
+\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
 y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
@@
 \end{pmatrix}}_{\boldsymbol \varepsilon}.
-$$
+$$ {#eq-lpp}
 
-To estimate $\boldsymbol \beta$ we need at least $d+1$ observations (thus $2h\geq d$) and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
+Since $d+1$ parameters are estimated with $\boldsymbol \beta$, we need at least $2h+1$ observations (thus $2h\geq d$) and the estimation is made by weighted least squares (WLS), which is equivalent to minimising the following objective function:
 $$
@@
 where $\kappa_j$ is a set of weights called a *kernel*.
-We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.$
+We have $\kappa_j\geq 0:\kappa_{-j}=\kappa_j$, and with $\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})$, the estimate of $\boldsymbol \beta$ can be written as $\hat{\boldsymbol\beta}=(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\transp{\boldsymbol X}\boldsymbol K\boldsymbol y.$
 With $\boldsymbol e_1=\transp{\begin{pmatrix}1 &0 &\cdots&0 \end{pmatrix}}$, the estimate of the trend is:
 $$
-\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
+\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
 $$ {#eq-mmsym}
 To conclude, the estimate of the trend $\hat{m}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
-Moreover, $\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}$ so:
+Moreover, since $\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}$ we have:
 $$
-\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
+\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\{1,2,\dots,d\right\}:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
 $$
@@
 
-1. Apply symmetric filters to the series extended by forecasting $\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket$.
+1. Apply symmetric filters to the series extended by forecasting $\hat{y}_{n+l\mid n},l\in\left\{1,2,\dots,h\right\}$.
 
-2. Build an asymmetric filter by local polynomial approximation on the available observations ($y_{t}$ for $t\in\left\llbracket n-h,n\right\rrbracket$).
+2. Build an asymmetric filter by local polynomial approximation on the available observations ($y_{t}$ for $t\in\left\{n-h,n-h+1,\dots,n\right\}$).
 
@@
 To solve the problem of the variance of real-time filter estimates, @proietti2008 propose a general method for constructing asymmetric filters that allows a bias-variance trade-off.
-This is a generalisation of @musgrave1964set's asymmetric filters (used in the X-13ARIMA seasonal adjustment algorithm).
+This is a generalisation of @musgrave1964set's asymmetric filters (used in the X-11 seasonal adjustment algorithm).
 
-The initial series is modelled here as:
+Rewriting @eq-lpp:
 $$
-\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
+\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
+\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
 \boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
 $$ {#eq-lpgeneralmodel}
-where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$.
+where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$ and the variance of the error is defined as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ to directly take into account the weights associated to the observations.
 The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
-These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta$ (with $\boldsymbol U_p$ the $(h+q+1)\times (d+1)$ matrix which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
+These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$ (with $\boldsymbol U_p$ the $(h+q+1)\times (d+1)$ matrix which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
 The problem is equivalent to finding $v$ which minimises:
@@
 \underbrace{
-  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
-  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
-  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
+  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
+  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
+  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
 }_\text{mean square revision error}+
-\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
+\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
 $$ {#eq-lppasym}
@@
 These asymmetric filters depend on the ratio $\lvert\delta_1/\sigma\rvert$, which, assuming that the trend is linear and the bias constant, can be linked to the I-C ratio $R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}$: $\delta_1/\sigma=2/(R\sqrt{\pi})$.
-This ratio is used in particular in X-13ARIMA to determine the length of the Henderson filter. 
+This ratio is used in particular in X-11 to determine the length of the Henderson filter. 
 For monthly data:
@@
 
+To measure the phase shift between the input series and the filtered one, @eq-lppasym introduces a criterion, $T_g$, called *timeliness*.
+When a moving average $\boldsymbol v$ is applied, the level input signal is also altered by the gain function. 
+That’s why the timeliness criterion depends on the gain and phase shift functions ($\rho_{\boldsymbol v}$ and $\varphi_{\boldsymbol v}$), the link between both functions being made by a penalty function $f$.
+The criterion measures phase shift for the frequencies in the interval $[\omega_1, \omega_2]$:
+$$
+T_g(\boldsymbol v)=\int_{\omega_{1}}^{\omega_{2}}f(\rho_{\boldsymbol v}(\omega),\varphi_{\boldsymbol v}(\omega))\ud\omega
+$$
+As mentioned in @sec-propMM, in the case of trend-cycle extraction we can set $\omega_1=0$ and for monthly series $\omega_2=2\pi/12.$
+For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show thant in that case the criteria can be written in a quadratic form:
+$$
+T_g(\boldsymbol v) = \transp{\boldsymbol v}\boldsymbol T\boldsymbol v
+$$
+with $T\boldsymbol=\int_{\omega_{1}}^{\omega_{2}}\boldsymbol{Si}(\omega)\transp{\boldsymbol{Si}(\omega)}\ud\omega$ and $\boldsymbol{Si}(\omega)=(\sin(p\omega),\dots, \sin (f\omega))$ if $\boldsymbol v$ uses $p$ points in the past and $f$ points in the future.
+
 Using the same notation as in @sec-lppasymf, $\boldsymbol\theta$ represents the symmetric filter and $\boldsymbol v$ the asymmetric filter.
@@
 $$
-T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
+T_g(\boldsymbol v)=\transp{\boldsymbol v}\boldsymbol T\boldsymbol v=\transp{\boldsymbol g}\boldsymbol T\boldsymbol g+2\transp{\boldsymbol \theta_p}\boldsymbol T\boldsymbol g+\transp{\boldsymbol \theta_p}\boldsymbol T\boldsymbol \theta_p
 \quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
@@
 \begin{align*}
-\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
-  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
-  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
-2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
-&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c,
+\varphi(\boldsymbol v)&=\transp{(\boldsymbol v-\boldsymbol \theta_p)}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
+  \transp{\boldsymbol \theta_f}\boldsymbol D_{f}\boldsymbol \theta_f+
+  [\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}+
+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)\\
+&=\transp{\boldsymbol g}\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)+\boldsymbol c,
 \end{align*}
-with $\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p,$ $\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'$ and $\boldsymbol c$ a constant independent of $\boldsymbol v.$
+with $\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z}_p,$ $\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}$ and $\boldsymbol c$ a constant independent of $\boldsymbol v.$
 
@@
 $$
-\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
-2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
+\widetilde\varphi(\boldsymbol v)=\transp{\boldsymbol g}\widetilde {\boldsymbol Q}\boldsymbol g-
+2\widetilde {\boldsymbol P}\boldsymbol g+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)+
 \widetilde {\boldsymbol c},
 $$
-where $\alpha_T$ is the weight associated with the timeliness criterion, $\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T$, $\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T$ and $\widetilde {\boldsymbol c}$ a constant independent of $\boldsymbol v.$
+where $\alpha_T$ is the weight associated with the timeliness criterion, $\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p} + \alpha_T\boldsymbol T$, $\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}-\alpha_T\boldsymbol \theta_p\boldsymbol T$ and $\widetilde {\boldsymbol c}$ a constant independent of $\boldsymbol v.$
 
@@
 
-- The variance $\sigma^2$ can be estimated using the observed data set and the symmetric filter $(w_{-p},\dots,w_p)$:
+- The variance $\sigma^2$ can be estimated using the observed data set and the symmetric filter $(w_{-h},\dots,w_h)$.
+Indeed, as for example shown by @Loader1999, the variance $\sigma^2$ can be estimating using the normalized residual sum of squares:
 $$
-\hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
+\hat\sigma^2=\frac{1}{(n-2h)-2\nu_1+\nu_2}\sum_{t=h+1}^{n-h}(y_t-\hat \mu_t)^2.
 $$
+Since the symmetric filter is used, only $n-2h$ observations can be used to estimate $\sigma^2$.
+$\nu_1$ and $\nu_2$ are two definitions of degrees of freedom of a local fit (generalization of the number of parameters of a parametric model).
+In the case of local polynomial smoothings, they can be linked to the coefficient of the associated filter: $\nu_1 = (n-h)w_0$ (number of observations multiplied by the central coefficient of the moving average) and $\nu_2 = (n-h)\sum_{i=-h}^{h} w_i^2$ (number of observations multiplied by the sum of squares of coefficients of the moving average).
+Thus, the variance can be estimated by:
+$$
+\hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum_{i=-h}^{h} w_i^2}.
+$$
+
 - The parameter $\delta$ can be estimated by moving average from @eq-mmsym. 
-For example, for the LC method we can use the moving average $\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}$ to obtain a local estimate of the slope and for the QL method we can use $\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}$ to obtain a local estimate of the concavity.
+For example, for the LC method we can use the moving average $\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}$ to obtain a local estimate of the slope and for the QL method we can use $\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}$ to obtain a local estimate of the concavity.
 The DAF method then simply allows us to calculate the associated asymmetric moving averages.  
@@
 
-```{r, echo=FALSE, out.width="100%", fig.note = footnotes}
+```{r, echo=FALSE, out.width="100%"}
 #| label: fig-graphstpsimul
@@
 2. D'autre part, les moyennes mobiles étant des opérateurs linéaires, ils sont sensibles à la présence de points atypiques.
-L'application directe des méthodes peut donc conduire à des estimations biaisées, du fait de leur présence, alors que les méthodes de désaisonnalisation (comme la méthode X-13ARIMA) ont un module de correction des points atypiques.
+L'application directe des méthodes peut donc conduire à des estimations biaisées, du fait de leur présence, alors que les méthodes de désaisonnalisation (comme la méthode X-13ARIMA-SEATS) ont un module de correction des points atypiques.
 Par ailleurs, comme notamment montré par @dagum1996new, le filtre symétrique final utilisé par X-13ARIMA pour extraire la tendance-cycle (et donc celui indirectement utilisé lorsqu'on applique les méthodes sur les séries désaisonnalisées) laisse passer environ 72 % des cycles de 9 ou 10 mois (généralement associés à du bruit plutôt qu'à la tendance-cycle).
@@
 Another possibility would be to study the impact of atypical points: moving averages, like any linear operator, are very sensitive to the presence of atypical points. 
-To limit their impact, in X-13ARIMA a strong correction for atypical points is performed on the irregular component before applying the filters to extract the trend-cycle.
+To limit their impact, in X-11 a strong correction for atypical points is performed on the irregular component before applying the filters to extract the trend-cycle.
 This leads us to study the impact of these points on the estimation of the trend-cycle and turning points, and also to explore new types of asymmetric filters based on robust methods (such as robust local regressions or moving medians).

### 2024-04-17 paper/resp_referee.qmd
improvement

@@
 
-> 4. Page 7, line 40: why say the series is extended by one year? The extensions are for as many 
-> years as needed by the particular filters (typically, half their length in years).
+*This sections was simplified*
+
+> 4. Page 7, line 40: why say the series is extended by one year? 
+> The extensions are for as many years as needed by the particular filters (typically, half their length in years).
 
 > 5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.
+
+The $\forall q$ was omitted and this section was rewritten to make it clearer.
 
@@
 > 9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness? 
-> That is, when you display the entries of the matrix T, you need to make a connection between 
+> That is, when you display the entries of the matrix $T$, you need to make a connection between 
 > those values and the concept of timeliness.
 
+*More details are added to this section to explain how the matrix $T$ is defined by Grun-Rehomme et al. (2018).*
+
 > 10. Page 15, line 53: I don't see why $v$ in this equation should be defined through the case bracket;
-> everything starting with "with" could be moved to the following text.
+>everything starting with "with" could be moved to the following text.
+
+*Indeed, the formula are corrected*
 
@@
 
+*All the code will be in a github repository*
+
 # Referee 1
 
-> 1. The authors should provide a clear definition of trend-cycle, $TC_t$, 
-> and irregular, $I_t$. In other words, how is the decomposition 
-> $y_t =TC_t+I_t$ identified? In section 1 they seem to hint at the interpretation 
-> of $TC_t$ as the low-pass component of yt with cutoff frequency 
-> $\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component). 
+> 1. The authors should provide a clear definition of trend-cycle, $TC_t$, and irregular, $I_t$. In other words, how is the decomposition $y_t =TC_t+I_t$ identified?
+> In section 1 they seem to hint at the interpretation of $TC_t$ as the low-pass component of yt with cutoff frequency $\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component). 
 > This point should be clarified, I think. 
 > Related questions are: are $y_t$ and its components stochastic or deterministic? 
-> The treatment in section 1.2.1 suggests that $TC_t$ is 
-> an unknown possibly nonlinear deterministic function of time. 
+> The treatment in section 1.2.1 suggests that $TC_t$ is an unknown possibly nonlinear deterministic function of time. 
 > The variance inflation factor $\sum_k \theta_k^2$ implies that $I_t$ is white noise.
 
-> 2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not
-> very formal. Do you mean that the unconditional variance is minimized?
+> 2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not very formal. 
+> Do you mean that the unconditional variance is minimized?
 > Or the long run variance?
 
-> 3. A definition of turning point would also be needed. Are these ’regular’
-points at which the slope of the trend changes sign? Or changepoints?
+> 3. A definition of turning point would also be needed. 
+Are these ’regular’ points at which the slope of the trend changes sign? 
+Or changepoints?
 
-> 4. The notation should be unified: why not writing $TC_t = \mu(t)$ and
-> in section 2 using $t$ instead of $x$? The observation index i hould
-> be replaced by $t$, as in section 2.1. Similarly, why not using ϵt
-> directly and start section 1.2 by stating that you assume that yt is
-> generated as $y_t = \mu(t) + \varepsilon_t$, etc.
-> Secondly, you should clarify the meaning of (3), where you set
-> $\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix $D$ for later treatment?
+> 4. The notation should be unified: why not writing $TC_t = \mu(t)$ and in section 2 using $t$ instead of $x$? 
+The observation index $i$ should be replaced by $t$, as in section 2.1. 
+Similarly, why not using $\varepsilon_t$ directly and start section 1.2 by stating that you assume that $y_t$ is generated as $y_t = \mu(t) + \varepsilon_t$, etc.
+Secondly, you should clarify the meaning of (3), where you set
+$\mu(t) = u'_t\gamma + z'_t\delta$ (also you assume that the noise component has possibly a non-scalar covariance matrix, do you need the matrix $D$ for later treatment?)
 
@@
 
-> 6. Is the phase shift related to the notion of bias in local polynomial
-smoothing?
+> 6. Is the phase shift related to the notion of bias in local polynomial smoothing?
 
@@
 
-> - Section 1.1. It would perhaps be useful to provide the expression
-of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of
-the transfer function of the linear filter.
+> - Section 1.1. It would perhaps be useful to provide the expression of the phase angle, $\varphi(\omega)$ in terms of the real and imaginary part of the transfer function of the linear filter.
 
-> - Section 1.1. The symbol $p$ for the period is already in use from the
-previous section, where it denotes the bandwidth and number of
-past observations used by the filter.
+*More details on the definition $\varphi$ and $\rho$ was added using the real part ($\sum_{k=-p}^{+f}\theta_k\cos(\omega k)$) and the imaginary part ($\sum_{k=-p}^{+f}\theta_k\sin(\omega k)$) of the transfer function*
 
-> - Page 4. The definition of the high-pass component (oscillations
-in the frequency range from $\pi/6$ to $\pi$ as undesirable oscillations is
-controversial.
+> - Section 1.1. The symbol $p$ for the period is already in use from the previous section, where it denotes the bandwidth and number of past observations used by the filter.
 
-> - Figure 1. In the caption, the ration $I/C$ was never defined before.
-The description of the three panels should be better explained in
-the main text.
+*The $p$ for the period was replaced by $s$*
 
-> - Section 1.2 makes a reference to X-11 and X-13-ARIMA. The author
-assumes that the reader is knowledgeable about seasonal adjustment
-methods. Perhaps the authors should state that their
-intent is to separate the trend-cycle component from the ’noise’
-component.
+> - Page 4. The definition of the high-pass component (oscillations in the frequency range from $\pi/6$ to $\pi$) as undesirable oscillations is controversial.
+
+> - Figure 1. In the caption, the ratio $I/C$ was never defined before.
+The description of the three panels should be better explained in the main text.
+
+> - Section 1.2 makes a reference to X-11 and X-13-ARIMA. 
+The author assumes that the reader is knowledgeable about seasonal adjustment methods. 
+Perhaps the authors should state that their intent is to separate the trend-cycle component from the ’noise’ component.
 
@@
 
+*Fixed*
+
 > - Page 8. What is the meaning of $y_1^*n\dots, y_h^*$ being the implicit forecasts induced by $w_0$, etc.? 
 The forecasts are obtained from an assumed polynomial trend model?
+
+*More details are added to this section. The definition is more general than the polynomial ind*
 
@@
 
-> - Figures 6-7. The labels of the time axes and the months are in
-French.
+> - Figures 6-7. The labels of the time axes and the months are in French.
+
+*Fixed*
 
@@
 
+Other (non-exhaustive) [only remarks that require a response]
+- P2 L8:  “X-13ARIMA seasonal adjustment method” -- The software is officially X-13ARIMA-SEATS, but the seasonal adjustment method itself is X-11 (applies to subsequent occurrences as well)
+
+- P2 L29:  “good properties” -- This could be less vague.
+
+- P7 L58:  There is a $q$ introduced here via “$w_t^q$ for $t>q$” that shows up later as well, but  itself does not appear to be properly defined otherwise.
+
+- P9:  Author lists a number of quality criteria in Table 1.
+It may be obvious for some of these, but it would be helpful to the reader to indicate whether lower values or higher values are preferable for each, either in the table or in the accompanying exposition.
+
+*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed *
+
+- P11 L22:  “$2h+1 \geq d+1$” -- Why not just say $2h\geq d$?
+
+This sentence was rephrased: $d+1$ refers to the number of parameters to estimate and $2h+1$ the number of observations.
+
+- P14 L33:  More exposition wrt Table 2 may be helpful, as there is a fair amount of information to digest.
+
+*Table 2 was removed*
+
+- P14 L45:  What is the rationale for this combined measure $EQM_w=A_w + S_w + T_w + R_w$?
+Looking at the expressions for these four from Table 1, it is apparent that $A_w$ and $T_w$ sum up (the two have the same integrand, and the resulting sum is the integral from 0 to $\pi$), and same goes for $S_w$ and $R_w$.
+And it should probably be explicitly stated what $EQM_w$ stands for prior to introducing it in the table.
+
+*This table was removed but to answer your question the criteria comes from the decomposition of the mean squared error into four quantities*
+
+- P19 L15:  “January 1960 and December 2020” -- As noted previously, is it worth considering an additional scenario with (simulated) series of shorter length?
+
 # Referee 3
@@
 
+This section was rewrite to make it clearer.
+
 > 2.  Symbol [[ ]] Meaning (Page 11, Lines 53-54)
@@
 
+This symbol was referring to integer interval, it was removed to avoid confusion (for example $[[1,d]]$ was replaced by $\{1,2,\dots, d\}$).
+
 > 3.  Variance Determination in Local Parameterization (Page 17)
@@
 clearly identify the quantities appearing in the denominator for better understanding.
+
+More explanations was added to this paragraph: the formula comes from local regression theory simplified because we are here in the case of local polynomial approximation.
 
@@
 
-2) The paper omits to consider the Covid19 period and how this influences the trend
-and cycle decomposition. I understand that this is not the main focus of the paper,
-but at least the authors should address this problem in word and possible solution
-to address it, e.g. outlier correction, robust distributions, etc.
+2) The paper omits to consider the Covid19 period and how this influences the trend and cycle decomposition. 
+I understand that this is not the main focus of the paper, but at least the authors should address this problem in word and possible solution to address it, e.g.
+outlier correction, robust distributions, etc.
 

### 2024-04-18 paper/article_jos.qmd
improvement

@@
   sec-prefix: section
+  eq-prefix: equation
 linestretch: 2
@@
 Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
-For example, the X-11 decomposition method (used in X-13ARIMA-SEATS) uses several moving averages to estimate the main components of a time series. 
+For example, the X-11 decomposition method [@x11], used in X-13ARIMA-SEATS software [@x13],  uses several moving averages to estimate the main components of a time series. 
 Symmetric filters are applied to the center of the series, but when it comes to estimate the most recent points, all of these methods must rely on asymmetric filters. 
@@
 The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
-The extensions have been implemented in the R `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and can therefore be easily used.
+Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and can therefore be easily used.
+All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).
 
@@
 In practice, local regressions can be used. 
-More specifically, consider a set of points $(x_i,y_i)_{1\leq i\leq n}$. 
-Non-parametric regression involves assuming that there exists a function $\mu$, to be estimated, such that $y_i=\mu(x_i)+\varepsilon_i$ with $\varepsilon_i$ an error term.
-According to Taylor's theorem, for any point $x_0$, if $\mu$ is differentiable $d$ times, then:
+More specifically, consider a set of points $(x_t,y_t)_{1\leq t\leq n}$ and we here consider that $x_t=t$ (no external regressor is used). 
+Non-parametric regression involves assuming that there exists a function $\mu$, to be estimated, such that $y_t=\mu(t)+\varepsilon_t$ with $\varepsilon_t$ an error term.
+According to Taylor's theorem, for any date $t_0$, if $\mu$ is differentiable $d$ times, then:
 $$
-\forall t \::\:TC_t = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
-\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
+\forall t \::\:\underbrace{\mu(t)}_{=TC_t} = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
+\frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
 $$ {#eq-taylor}
-where $R_d$ is a negligible residual term in the neighbourhood of $x_0$. 
-In a neighbourhood $h(x_0)$ around $x_0$, $\mu$ can be approximated by a polynomial of degree $d$. 
-The quantity $h(x_0)$ is called the *bandwidth*.
-If $\varepsilon_i$ is white noise, then we can estimate by least squares $\mu(x_0)$ using the observations which are in $\left[x_0-h(x_0),x_0+h(x_0)\right]$.
+where $R_d$ is a negligible residual term in the neighbourhood of $t_0$. 
+In a neighbourhood $h(t_0)$ around $t_0$, $\mu$ can therefore be approximated by a polynomial of degree $d$. 
+The quantity $h(t_0)$ is called the *bandwidth*.
+If $\varepsilon_t$ is white noise, we can estimate $\mu(t_0)=TC_{t_0}$ by least squares using the observations which are in $\left[t_0-h(t_0),t_0+h(t_0)\right]$. 
+In practice, this means assuming that the trend is locally polynomial.
+We also generally assume that the bandwith is fixed over time: $h(t_0)=h.$
+This is also what is assumed in the seasonal adjustment algorithm X-11 and what will be assumed in this paper.
 
-In practice, this means assuming that the trend is locally polynomial. 
 Various estimation methods can be used to derive symmetric and asymmetric moving averages.  
@@
 $$ {#eq-lpgeneralmodel}
-where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X$ and the variance of the error is defined as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ to directly take into account the weights associated to the observations.
-The objective is to find a filter $\boldsymbol v$ that minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain constraints.
-These constraints are represented by the matrix $\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}$: $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$ (with $\boldsymbol U_p$ the $(h+q+1)\times (d+1)$ matrix which contains the observations of the matrix $U$ known during estimation by the asymmetric filter).
-The problem is equivalent to finding $v$ which minimises:
+where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X.$
+Defining the variance error as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ (to directly take into account the weights associated to the observations) and with $[\boldsymbol U,\boldsymbol Z] = \boldsymbol X$ we this find the model associated to symmetric filters.
+The objective is to find an asymmetric filter $\boldsymbol v$ which uses $q$ points in the future and minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain (polynomial) constraints.
+The matrices $\boldsymbol y$, $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned as follows:
+$$
+\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
+\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
+\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
+\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
+$$
+where $\boldsymbol y_p$ is a matrix $(h+q+1)\times 1$ which denotes the set of available observations (from $y_{t-h}$ to $y_{t+q}$) and $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned in the same way.
+The constraints are defined as $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$: they represents the polynomial preservation constraints we want to impose on the asymmetric filter.
+@proietti2008 show that the problem is equivalent to finding $v$ which minimises:
 $$
@@
 
-When $\boldsymbol U=\boldsymbol X$, the constraint is equivalent to preserving polynomials of degree $d$: we find asymmetric direct filters (DAF) when $\boldsymbol D=\boldsymbol K^{-1}$.
+When $\boldsymbol U=\boldsymbol X$ (and then $\boldsymbol Z = \boldsymbol 0$), the constraint is equivalent to preserving polynomials of degree $d$ and we find asymmetric direct filters (DAF) when $\boldsymbol D=\boldsymbol K^{-1}$.
 
-When $\boldsymbol U=\transp{\begin{pmatrix}1&\cdots&1\end{pmatrix}}$, $\boldsymbol Z=\transp{\begin{pmatrix}-h&\cdots&+h\end{pmatrix}}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters (using the Henderson kernel).
+When $\boldsymbol U=\transp{\begin{pmatrix}1&\cdots&1\end{pmatrix}}$, $\boldsymbol Z=\transp{\begin{pmatrix}-h&\cdots&+h\end{pmatrix}}$, $\boldsymbol \delta=\delta_1$, $\boldsymbol D=\sigma^2\boldsymbol I$ and when the symmetric filter is the Henderson filter, we find the asymmetric Musgrave filters.
 This filter assumes that, for real-time estimation, the data is generated by a linear process and that the asymmetric filters preserve the constants ($\sum v_i=\sum \theta_i=1$).
@@
 
-Asymmetric filters are usually parametrised globally: $\lvert\delta/\sigma\rvert$ estimated on all the data using the IC-ratio or a cross-validation criterion. 
+Asymmetric filters are usually parametrised globally: $\lvert\delta/\sigma\rvert$ is fixed using all the data (with the IC-ratio or a cross-validation criterion). 
 However, we might prefer a local parametrisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
 Indeed, although the overall parametrisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetric filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
-For example, with the LC method, global parametrisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up to the turning point.
+For example, with the LC method, global parametrisation means assuming that the slope of the trend is constant, whereas during economic downturns it tends towards 0 up around the turning point.
 
@@
 
-- The variance $\sigma^2$ can be estimated using the observed data set and the symmetric filter $(w_{-h},\dots,w_h)$.
+- The variance $\sigma^2$ can be estimated using the observed data set and the symmetric filter $(\theta_{-h},\dots,\theta_h)$.
 Indeed, as for example shown by @Loader1999, the variance $\sigma^2$ can be estimating using the normalized residual sum of squares:
 $$
-\hat\sigma^2=\frac{1}{(n-2h)-2\nu_1+\nu_2}\sum_{t=h+1}^{n-h}(y_t-\hat \mu_t)^2.
+\hat\sigma^2=\frac{1}{(n-2h)-2\nu_1+\nu_2}\sum_{t=h+1}^{n-h}(y_t-\widehat{TC}_t)^2.
 $$
@@
 $\nu_1$ and $\nu_2$ are two definitions of degrees of freedom of a local fit (generalization of the number of parameters of a parametric model).
-In the case of local polynomial smoothings, they can be linked to the coefficient of the associated filter: $\nu_1 = (n-h)w_0$ (number of observations multiplied by the central coefficient of the moving average) and $\nu_2 = (n-h)\sum_{i=-h}^{h} w_i^2$ (number of observations multiplied by the sum of squares of coefficients of the moving average).
+In the case of local polynomial smoothings, they can be linked to the coefficient of the associated filter: $\nu_1 = (n-h)\theta_0$ (number of observations multiplied by the central coefficient of the moving average) and $\nu_2 = (n-h)\sum_{i=-h}^{h} \theta_i^2$ (number of observations multiplied by the sum of squares of coefficients of the moving average).
 Thus, the variance can be estimated by:
 $$
-\hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum_{i=-h}^{h} w_i^2}.
+\hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\widehat{TC}_t)^2}{1-2\theta_0^2+\sum_{i=-h}^{h} \theta_i^2}.
 $$
+<!-- The assumption that $\hat\mu(x)$ is unbiased is rarely exactly true, so variance -->
+<!-- estimates and confidence intervals are usually computed at small band- -->
+<!-- widths where bias is small.  -->
 
@@
 The different methods are compared on simulated and real data. 
-For all series, a symmetric 13-term filter is used.
-These methods are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
+For all series, a symmetric 13-term filter is used: we then assume than the bandwidth is fixed, as in the seasonal adjustement method X-11.
+A nearest neighbour bandwidth is also tested (13-term asymmetric filters: real-time estimates uses observation between $y_{t-12}$ and $y_{t}$, the second asymmetric filters uses obversations between $y_{t-11}$ and $y_{t+1}$, etc.) but with no improvement in the results (see supplemental material).
+More complexe methods to locally select bandwidth, like the one proposed by @fan1992variable, were not tested.
+All the methods to create asymmetric filters are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
 The ARIMA model is determined automatically, using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
-
 The performance of the different methods is judged on criteria relating to revisions (between two consecutive estimates and in relation to the final estimate) and the number of periods required to detect turning points.

### 2024-04-18 paper/resp_referee.qmd
improvement

@@
 
+*It was added in the beginning of section 3 and 4 than the bandwidth is fixed. Nearest neighbour bandwidth was added in supplemental material.*
+
 > 6. Is the phase shift related to the notion of bias in local polynomial smoothing?
@@
 
-*Fixed*
+*Corrected*
 
@@
 
+*This was clarified in the article, for example in the introduction the text was changed with "the X-11 decomposition method (used in X-13ARIMA-SEATS)"*
+
 - P2 L29:  “good properties” -- This could be less vague.
@@
 - P7 L58:  There is a $q$ introduced here via “$w_t^q$ for $t>q$” that shows up later as well, but  itself does not appear to be properly defined otherwise.
+
+*This section was rewritten and all the terms are now defined.*
 
@@
 
-This sentence was rephrased: $d+1$ refers to the number of parameters to estimate and $2h+1$ the number of observations.
+*This sentence was rephrased: $d+1$ refers to the number of parameters to estimate and $2h+1$ the number of observations.*
 
@@
 
+*See previous comment: it was added in supplemental material*
+
 # Referee 3
@@
 
-This section was rewrite to make it clearer.
+*This section was rewrite to make it clearer.*
 
@@
 
-This symbol was referring to integer interval, it was removed to avoid confusion (for example $[[1,d]]$ was replaced by $\{1,2,\dots, d\}$).
+*This symbol was referring to integer interval, it was removed to avoid confusion (for example $[[1,d]]$ was replaced by $\{1,2,\dots, d\}$).*
 
@@
 
-More explanations was added to this paragraph: the formula comes from local regression theory simplified because we are here in the case of local polynomial approximation.
+*More explanations was added to this paragraph: the formula comes from local regression theory simplified because we are here in the case of local polynomial approximation.*
 

### 2024-04-18 paper/article_jos.qmd
appendix

@@
 All the methods to create asymmetric filters are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
-The ARIMA model is determined automatically, using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
+The ARIMA model is determined automatically, using @autoarima algorithm (function `forecast::auto.arima()`), using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
+
 The performance of the different methods is judged on criteria relating to revisions (between two consecutive estimates and in relation to the final estimate) and the number of periods required to detect turning points.

### 2024-04-18 paper/supp_material.qmd
update
appendix

@@
 
+\newpage
+
+# Impact of the span used to detect the ARIMA model
+
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_arima_length.pdf")
+```
+
+Distribution of phase shifts on simulated series by number of years used to identify and estimate the AIRMA model.
+:::
+
+# Quarterly data
+
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_q.pdf")
+```
+
+Distribution of phase shifts on quarterly data.
+:::
+
+# Nearest neighbour bandwidth
+
 ```{r, out.width="100%"}
@@
 
-\newpage
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp.pdf")
+```
+
+Distribution of phase shifts using nearest neighbour bandwidth.
+:::
+
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp.pdf")
+```
+
+Distribution of phase shifts using nearest neighbour bandwidth for polynomial filters with local parametrisation.
+:::
 

### 2024-04-21 paper/article_jos.qmd
update nn + correction local param h<6 + imports errors

@@
 $$
+This is implemented in the function `rjd3filters::var_estimator()`.
 <!-- The assumption that $\hat\mu(x)$ is unbiased is rarely exactly true, so variance -->

### 2024-04-21 paper/article_jos.tex
update nn + correction local param h<6 + imports errors

@@
 
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{241,243,245}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
-\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
-
 \providecommand{\tightlist}{%
@@
 \hypersetup{
-  pdftitle={Improving real-time trend estimates using local parametrisation of local polynomial regression filters},
+  pdftitle={Improving real-time trend estimates using local parametrisation of polynomial regression filters},
   colorlinks=true,
@@
 \title{Improving real-time trend estimates using local parametrisation
-of local polynomial regression filters}
+of polynomial regression filters}
 \author{}
@@
 
-This paper studies and compares real-time estimates of the trend cycle
+This paper studies and compares real-time estimates of the trend-cycle
 component with moving averages based on local polynomial regression. In
-particular, this theory allows to reproduce Henderson's symmetric and
+particular, this theory allows us to reproduce Henderson's symmetric and
 Musgrave's asymmetric filters. This paper describes how they can be
 extended to include a timeliness criterion to minimise the phase shift.
-While asymmetric filters are generally parameterised globally, which can
+While asymmetric filters are generally parametrised globally, which can
 be suboptimal around turning points, this paper proposes a procedure for
-parameterising them locally. An empirical comparison, based on simulated
+parametrising them locally. An empirical comparison, based on simulated
 and real data, shows that modelling polynomial trends that are too
 complex introduces more revisions without reducing the phase shift, and
-that local parameterisation reduces the delay in detecting turning
-points and reduces revisions. The paper also shows how these results can
-be easily reproduced using the R package \texttt{rjd3filters} dedicated
-to the manipulation of moving averages.
+that local parametrisation reduces the delay in detecting turning points
+and reduces revisions. The paper also shows how these results can be
+easily reproduced using the R package \texttt{rjd3filters} dedicated to
+the manipulation of moving averages.
 
@@
 A moving average, or linear filter, is a statistical method that
-consists in applying a rolling weighted mean to a times series: for each
+consists of applying a rolling weighted mean to a times series: for each
 date \(t\) it computes a weighted mean of \(p\) past points and \(q\)
@@
 averages are ubiquitous in business cycle extraction and seasonal
-adjustment methods. For example, the X-13ARIMA seasonal adjustment
-method uses several moving averages to estimate the main components of a
-time series. Symmetric filters are applied to the center of the series,
-but when it comes to estimate the most recent points, all of these
-methods must rely on asymmetric filters. For trend-cycle extraction, the
-most popular symmetric filter is the \textcite{henderson1916note} moving
-average, which is used in the X-13ARIMA seasonal adjustment algorithm.
+adjustment methods. For example, the X-11 decomposition method
+\autocite{x11}, used in X-13ARIMA-SEATS software \autocite{x13}, uses
+several moving averages to estimate the main components of a time
+series. Symmetric filters are applied to the center of the series, but
+when it comes to estimate the most recent points, all of these methods
+must rely on asymmetric filters. For trend-cycle extraction, the most
+popular symmetric filter is the \textcite{henderson1916note} moving
+average, which is used in the X-11 seasonal adjustment algorithm.
 
@@
 estimate the most recent points. For example, for trend-cycle
-extraction, X-13ARIMA uses Henderson's symmetric filter and
+extraction, X-11 uses Henderson's symmetric filter and
 \textcite{musgrave1964set}'s asymmetric filters on an extended series
@@
 
-Using local polynomial regression, \textcite{proietti2008} developed
+Using local polynomial regression, \textcite{proietti2008} developed a
 general class of symmetric and asymmetric moving averages. In
-particular, it makes possible to reproduce Henderson's symmetric filter
-and Musgrave's asymmetric filters. However, these methods have two
-drawbacks. Firstly, the phase shift (i.e.~the delay in detecting turning
-points) is not directly controlled. Secondly, asymmetric filters depend
-on a parameter that is generally estimated over the whole series using
-assumptions that may be wrong locally. The aim of this study is to
+particular, it becomes possible to reproduce Henderson's symmetric
+filter and Musgrave's asymmetric filters. However, these methods have
+two drawbacks. Firstly, the phase shift (i.e.~the delay in detecting
+turning points) is not directly controlled. Secondly, asymmetric filters
+depend on a parameter that is generally estimated over the whole series
+using assumptions that may be wrong locally. The aim of this study is to
 propose extensions to the \textcite{proietti2008} class of filters in
 order to take these shortcomings into account. The extensions have been
-implemented in the R \texttt{rjd3filters} package (available at
-\url{https://github.com/rjdemetra/rjd3filters}) and are therefore easily
-reusable.
-
-In Section~\ref{sec-propMM}, we describe the general properties of
+implemented in the statistical software R \autocite{logicielr}
+\texttt{rjd3filters} package (available at
+\url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
+easily used.
+
+In section~\ref{sec-propMM}, we describe the general properties of
 moving averages and the associated quality criteria. This allows us to
@@
 from local polynomial regressions, as well as those behind the two
-extensions proposed in this article (Section~\ref{sec-nonparamreg}). In
-a final section (Section~\ref{sec-comparison}), all these methods are
+extensions proposed in this article (section~\ref{sec-nonparamreg}). In
+a final section (section~\ref{sec-comparison}), all these methods are
 compared empirically on simulated and real series.
 
-\section{Some properties on moving averages}\label{sec-propMM}
-
-Lots of papers describe the definition and the properties of moving
+\section{Some properties of moving averages}\label{sec-propMM}
+
+Numerous papers describe the definition and the properties of moving
 averages and linear filters, see for example \textcite{ch12HBSA}. In
@@
 
-Let \(p\) et \(f\) two integers, a moving average
+Let \(p\) and \(f\) two integers, a moving average
 \(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
@@
 is called the \emph{transfer function} or \emph{frequency response
-function}. The frequency response function can equivalently be defined
+function}. The frequency response function can alternatively be defined
 as
 \(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
-ou
+or
 \(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
@@
 \(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
-\emph{time shift} function. This function is sometimes represented as
+\emph{time shift} function. We this have: \begin{align*}
+\rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
+\left(\sum_{k=-p}^{+f}\theta_k\sin(\omega k)\right)^2} \\
+\tan \varphi_{\boldsymbol\theta}(\omega) &=- \frac{\sum_{k=-p}^{+f}\theta_k\sin(\omega k)}{\sum_{k=-p}^{+f}\theta_k\cos(\omega k)} \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) \ne 0 \\
+&= 0 \text{ if }\sum_{k=-p}^{+f}\theta_k\cos(\omega k) = 0.
+\end{align*} The phase shift is sometimes represented as
 \(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
-to mesure the phase shift in number of periods. For all symmetric moving
-average we have
-\(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).
+to measure it in number of periods. For all symmetric moving averages we
+have \(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).
 
@@
 
-Fourier decomposition allows to analyze any time series as a sum of
+Fourier decomposition allows us to analyze any time series as a sum of
 harmonic series, and each component (trend, cycle, seasonal, irregular)
-is associated with a set of frequencies. Noting \(\omega = 2\pi/p\), the
+is associated with a set of frequencies. Noting \(\omega = 2\pi/s\), the
 harmonic series of frequency \(\omega\) represents a series that repeats
-itself every \(p\) periods. For example, for a monthly series (12
+itself every \(s\) periods. For example, for a monthly series (\(s=12\)
 observations per year), seasonal movements are those that repeat every
@@
 paper, we consider that the trend-cycle is associated with frequencies
-in the interval \([0, 2\pi/12[\), i.e.~movements recurring at least
+in the interval \([0, 2\pi/12]\), i.e.~movements recurring at least
 every 12 months. Even if different frequencies are sometimes used
@@
 
-Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
-for the asymmetric Musgrave filter (see Section~\ref{sec-lppasymf})
+figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
+for the asymmetric Musgrave filter (see section~\ref{sec-lppasymf})
 often used for real-time trend-cycle estimation (i.e.~when no point in
@@
 means that the trend-cycle is well preserved and that short cycles of 1
-to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. However, cycles
-of 8 to 12 months (\([2\pi/12, 2\pi/8]\)), considered undesirable
-because associated with the irregular, are also amplified: this can lead
-to the detection of false turning points. At other frequencies, the gain
-function is less than 1, but always positive: this means that the series
-smoothed by this moving average will always contain noise, even if it is
-attenuated. Analysis of the phase shift shows that the shorter the
-cycles, the greater the phase shift: this means that on series smoothed
-by this moving average, turning points could be detected at the wrong
-date.
+to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. At other
+frequencies, the gain function is less than 1, but always positive: this
+means that the series smoothed by this moving average will always
+contain noise, even if it is attenuated. Analysis of the phase shift
+shows that the shorter the cycles, the greater the phase shift: this
+means that on series smoothed by this moving average, turning points
+could be detected at the wrong date.
 
@@
 To decompose a time series into a seasonal component, a trend-cycle and
-the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses
-a succession of moving averages, all with specific constraints.
+the irregular, the X-11 decomposition algorithm (used in
+X-13ARIMA-SEATS) uses a succession of moving averages, all with specific
+constraints.
 
@@
 Trend-cycles are generally modelled by local polynomial trends (see
-Section~\ref{sec-nonparamreg}). In order to best preserve the
-trend-cycles, we try to have moving averages that preserve the
+section~\ref{sec-nonparamreg}). In order to best preserve the
+trend-cycles, we try to design moving averages that preserve the
 polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
@@
 
-We have the following properties for the moving average
-\(M_{\boldsymbol\theta}\):
+For the moving average \(M_{\boldsymbol\theta}\), it can be shown that:
 
 \begin{itemize}
-\item
-  To preserve a constant trend \(X_t=a\) we need: \[
-  \forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
-  \] The sum of the coefficients of the moving average
-  \(\sum_{k=-p}^{+f}\theta_k\) must then be equal to \(1\).
-\item
-  To preserve linear trends \(X_t=at+b\) we need: \[
-  \forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
-  \] Which is equivalent to: \[
-  \sum_{k=-p}^{+f}\theta_k=1
-  \quad\text{et}\quad
-  \sum_{k=-p}^{+f}k\theta_k=0.
-  \]
-\item
-  In general, \(M_{\boldsymbol\theta}\) preserves polynomials of degree
-  \(d\) if and only if: \[
+\tightlist
+\item
+  \(M_{\boldsymbol\theta}\) preserves polynomials of degree \(d\) if and
+  only if: \[
   \sum_{k=-p}^{+f}\theta_k=1
    \text{ and }
-  \forall j \in \left\llbracket 1,d\right\rrbracket:\:
+  \forall j \in \left\{1,2,\dots,d\right\}:\:
   \sum_{k=-p}^{+f}k^j\theta_k=0.
@@
 
-Indeed, let \(\{\varepsilon_t\}\) a sequence of independent random
+Indeed, let \(\{\varepsilon_t\}\) be a sequence of independent random
 variables with \(\E{\varepsilon_t}=0\), \(\V{\varepsilon_t}=\sigma^2\):
@@
 because no past or future value can be used. Thus, for real-time
-estimation, it is needed to build asymmetric moving average that
+estimation, it is needed to build asymmetric moving averages that
 approximate the symmetric moving average.
@@
 \item
-  Apply symmetrical filters to extended forecast series. This method
-  seems to date back to \textcite{deforest1877adjustment}, which also
-  suggests modelling a polynomial trend of degree three or less at the
-  end of the period:\\
-  ``\emph{As the first \(m\) and last \(m\) terms of the series cannot
-  be reached directly by the formula, the series should be graphically
-  extended by m terms at both ends, first plotting the observations on
-  paper as ordinates, and then extending the curve along what seems to
-  be its probable course, and measuring the ordinates of the extended
-  portions.} \emph{It is not necessary that this extension should
-  coincide with what would be the true course of the curve in those
-  parts. } \emph{The important point is that the m terms thus added,
-  taken together with the \(m+1\) adjacent given terms, should follow a
-  curve whose form is approximately algebraic and of a degree not higher
-  than the third.}''\\
-  This is also the approach used in the X-13ARIMA seasonal adjustment
-  method, which extends which extends the series by one year with an
-  ARIMA model.
+  Apply symmetric filters to extended forecast series. This method seems
+  to date back to \textcite{deforest1877adjustment}, which also suggests
+  modelling a polynomial trend of degree three or less at the end of the
+  period. A similar approach is used in the X-13ARIMA-SEATS seasonal
+  adjustment method: the series is extend by one year (by default) with
+  an ARIMA model to minimise the revisions associated to asymmetric
+  filters
 \end{enumerate}
@@
 Conversely, the implicit forecasts of an asymmetric moving average can
-be deduced from a reference symmetric moving average. This allows to
+be deduced from a reference symmetric moving average. This allows us to
 judge the quality of real-time estimates of the trend-cycle and to
@@
 evolutions.\\
-Let us denote \(v=(v_{-h},\dots, v_{h})\) the reference symmetrical
-moving average and \(w^0,\dots w^{h-1}\) a sequence of asymmetrical
-moving averages, of order \(h+1\) to \(2h\) used to estimate the last
-\(h\) points with, for convention, \(w_t^q=0\) for \(t>q\). This means
-that \(w^0=(w_{-h}^0,\dots, w_{0}^0)\) is used for estimation in real
-time (when no points are known in the future),
-\(w^1=(w_{-h}^1,\dots, w_{1}^1)\) for estimation of the penultimate
-point (when only one point is known in the future), and so on. Let also
-note \(y_{-h},\dots,y_{0}\) the observed series studied and
-\(y_{1}^*,\dots y_h^*\) the implicit forecast induced by
-\(w^0,\dots w^{h-1}\). This means that for all \(q\) we have: \[
-\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
-=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
-\] This is equivalent to: \[
-\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
+Let us denote \(\boldsymbol v=(v_{-h},\dots, v_{h})\) the reference
+symmetric moving average (used to compute the final estimates) and
+\(\boldsymbol w^0,\dots, \boldsymbol w^{h-1}\) a sequence of asymmetric
+moving averages used for the intermediate estimates.
+\(\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)\) is used for estimation in
+real time estimate (when 0 point in the future is known),
+\(\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)\) is used for
+estimation of the penultimate point (when only 1 in the future is
+known), and so on. Let \(y_{t-h},\dots,y_{t}\) be the observed series
+during the last \(h\) periods (for which asymmetric moving averages are
+used to get the intermediate trend-cycle estimates for the dates \(t-h\)
+to \(h\)). The implicit forecast \(y_{t+1}^*,\dots y_{t+h}^*\) induced
+by \(w^0,\dots w^{h-1}\) are the forecasts of the series \(y_t\) for
+which the estimates of the symmetric filter applied to the extended
+series
+\(\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)\)
+produce the same estimates as the asymmetric filter applied to the
+series \(\boldsymbol y^*\). This means: \begin{align*}
+\underbrace{\sum_{i=-h}^0 v_iy_{ti} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
+\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ points after }t} \\
+&=
+\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ points after }t} \\
+&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
+&= \cdots \\
+&=\underbrace{\sum_{i=-h}^0 w_i^{h-1}y_{t+i} + \sum_{i=1}^{h-1} w_i^{h-1}y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^{h-1}\text{ using }h-1\text{ points after }t}
+\end{align*} For any integer \(q\) between 0 and \(h-1\), using for
+convention \(w_t^q=0\) if \(t>q\) we then have: \[
+\sum_{i=1}^h (v_i- w_i^q) y_i^*
 =\sum_{i=-h}^0 (w_i^q-v_i)y_i.
-\] To sum up, in matrix form, this is equivalent to solving: \[
+\] In matrix form, this is equivalent to solving: \[
 \scriptstyle
@@
 
-\subsubsection{Quality indicators for asymmetric moving
-averages}\label{subsec:crit-qual}
-
-For asymmetric filters, most of the criteria come from those defined by
-\textcite{ch15HBSA} and \textcite{trilemmaWMR2019} to build asymmetric
-filters. They are summarised in Table~\ref{tab-QC} and can be calculated
-using the \texttt{rjd3filters::diagnostic\_matrix()} function.
-
-\textcite{ch15HBSA} propose a general approach to derive linear filters,
-based on an optimisation problem involving three criteria:
-\emph{Fidelity} (\(F_g\)), \emph{Smoothness} (\(S_g\)) and
-\emph{Timeliness} (\(T_g\)). Fidelity can be directly linked to the
-reduction in variance created by the filter, and timeliness to the
-notion of phase shift, which again we want to be low. This method can be
-used inR with the function \texttt{rjd3filters::fst\_filter()}.
-
-\textcite{trilemmaWMR2019} propose an approach based on the
-decomposition of the mean squared error between the symmetric and the
-asymmetric filter in four quantities: \emph{Accuracy} (\(A_w\)),
-\emph{Timeliness} (\(T_w\)), \emph{Smoothness} (\(S_w\)) and
-\emph{Residual} (\(R_w\)). A simplified version of this method can be
-used in R with the function \texttt{rjd3filters::dfa\_filter()}.
-
-\begin{table}
-
-\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
-
-\centering{
-
-[!h]
-
-{\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{ccc}
-\toprule
-Code & Description & Formula\\
-\midrule
-$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
-$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
-$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
-$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
-$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
-\addlinespace
-$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
-$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-\bottomrule
-\end{tabular}}}
-
-\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}
-
-\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}
-
-\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
-
-}
-
-\end{table}%
-
 \section{Non-parametric regression and local polynomial
@@
 regressions can be used. More specifically, consider a set of points
-\((x_i,y_i)_{1\leq i\leq n}\). Non-parametric regression involves
-assuming that there exists a function \(\mu\), to be estimated, such
-that \(y_i=\mu(x_i)+\varepsilon_i\) with \(\varepsilon_i\) an error
-term. According to Taylor's theorem, for any point \(x_0\), if \(\mu\)
-is differentiable \(d\) times, then:
+\((x_t,y_t)_{1\leq t\leq n}\) and we here consider that \(x_t=t\) (no
+external regressor is used). Non-parametric regression involves assuming
+that there exists a function \(\mu\), to be estimated, such that
+\(y_t=\mu(t)+\varepsilon_t\) with \(\varepsilon_t\) an error term.
+According to Taylor's theorem, for any date \(t_0\), if \(\mu\) is
+differentiable \(d\) times, then:
 \begin{equation}\phantomsection\label{eq-taylor}{
-\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
-\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
+\forall t \::\:\underbrace{\mu(t)}_{=TC_t} = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
+\frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
 }\end{equation} where \(R_d\) is a negligible residual term in the
-neighbourhood of \(x_0\). In a neighbourhood \(h(x_0)\) around \(x_0\),
-\(\mu\) can be approximated by a polynomial of degree \(d\). The
-quantity \(h(x_0)\) is called the \emph{bandwidth}. If \(\varepsilon_i\)
-is white noise, then we can estimate by least squares \(\mu(x_0)\) using
-the observations which are in \(\left[x_0-h(x_0),x_0+h(x_0)\right]\).
-
-In practice, this means assuming that the trend is locally polynomial.
-Various estimation methods can be used to derive symmetrical and
-asymmetrical moving averages.\\
-\textcite{GrayThomson1996} propose a complete statistical framework
-which makes it possible, in particular, to model the error in
-approximating the trend by local polynomials. However, as the
-specification of this error is generally complex, simpler models may be
-preferred, such as that of \textcite{proietti2008}.\\
+neighbourhood of \(t_0\). In a neighbourhood \(h(t_0)\) around \(t_0\),
+\(\mu\) can therefore be approximated by a polynomial of degree \(d\).
+The quantity \(h(t_0)\) is called the \emph{bandwidth}. If
+\(\varepsilon_t\) is white noise, we can estimate \(\mu(t_0)=TC_{t_0}\)
+by least squares using the observations which are in
+\(\left[t_0-h(t_0),t_0+h(t_0)\right]\). In practice, this means assuming
+that the trend is locally polynomial. We also generally assume that the
+bandwith is fixed over time: \(h(t_0)=h.\) This is also what is assumed
+in the seasonal adjustment algorithm X-11 and what will be assumed in
+this paper.
+
+Various estimation methods can be used to derive symmetric and
+asymmetric moving averages.\\
+\textcite{GrayThomson1996} propose a complete statistical framework that
+makes it possible, in particular, to model the error in approximating
+the trend by local polynomials. However, as the specification of this
+error is generally complex, simpler models may be preferred, such as
+that of \textcite{proietti2008}.\\
 Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
@@
 facilitating the calculation of different moving averages at different
-time frequencies. This method can be used inR with the function
+time frequencies, this method can be used in R with the function
 \texttt{rjd3filters::rkhs\_filter()}.
@@
 \[
-\forall j\in\left\llbracket -h,h\right\rrbracket:\:
+\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
 y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
@@
 
-In matrix notation: \[
+In matrix notation: \begin{equation}\phantomsection\label{eq-lpp}{
 \underbrace{\begin{pmatrix}y_{t-h}\\
@@
 \end{pmatrix}}_{\boldsymbol \varepsilon}.
-\]
-
-To estimate \(\boldsymbol \beta\) we need \(2h+1\geq d+1\) and the
+}\end{equation}
+
+Since \(d+1\) parameters are estimated with \(\boldsymbol \beta\), we
+need at least \(2h+1\) observations (thus \(2h\geq d\)) and the
 estimation is made by weighted least squares (WLS), which is equivalent
@@
 S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
-\] where \(\kappa_j\) is a set of weights called \emph{kernel}. We have
-\(\kappa_j\geq 0:\kappa_{-j}=\kappa_j\), and with
+\] where \(\kappa_j\) is a set of weights called a \emph{kernel}. We
+have \(\kappa_j\geq 0:\kappa_{-j}=\kappa_j\), and with
 \(\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})\), the estimate of
 \(\boldsymbol \beta\) can be written as
-\(\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.\)
+\(\hat{\boldsymbol\beta}=(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\transp{\boldsymbol X}\boldsymbol K\boldsymbol y.\)
 With
@@
 \begin{equation}\phantomsection\label{eq-mmsym}{
-\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
+\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
 }\end{equation} To conclude, the estimate of the trend \(\hat{m}_{t}\)
@@
 \(y_t\) (\(\boldsymbol \theta\) is symmetric due to the symmetry of the
-kernel weights \(\kappa_j\)). Moreover,
-\(\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}\) so: \[
-\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
-\] Hence, the filter \(\boldsymbol w\) preserve deterministic polynomial
-of order \(d\).
+kernel weights \(\kappa_j\)). Moreover, since
+\(\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}\) we have:
+\[
+\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\{1,2,\dots,d\right\}:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
+\] Hence, the filter \(\boldsymbol w\) preserves deterministic
+polynomials of order \(d\).
 
@@
 
-As mentioned in Section~\ref{sec-mmetprev}, for real-time estimation,
+As mentioned in section~\ref{sec-mmetprev}, for real-time estimation,
 several approaches can be used:
@@
   Apply symmetric filters to the series extended by forecasting
-  \(\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket\).
+  \(\hat{y}_{n+l\mid n},l\in\left\{1,2,\dots,h\right\}\).
 \item
@@
   available observations (\(y_{t}\) for
-  \(t\in\left\llbracket n-h,n\right\rrbracket\)).
+  \(t\in\left\{n-h,n-h+1,\dots,n\right\}\)).
 \item
@@
 constraints as those of the symmetric filter. The third method is called
-\emph{direct asymmetric filter} (DAF). This method is used for real-time
-estimation in the STL seasonal adjustment method (\emph{Seasonal-Trend
-decomposition based on Loess}, see \textcite{cleveland90}). Although DAF
-filter estimates are unbiased, this is at the cost of greater variance
-in the estimates.
+\emph{direct asymmetric filtering} (DAF). This method is used for
+real-time estimation in the STL seasonal adjustment method
+(\emph{Seasonal-Trend decomposition based on Loess}, see
+\textcite{cleveland90}). Although DAF filter estimates are unbiased,
+this is at the cost of greater variance in the estimates.
 
@@
 generalisation of \textcite{musgrave1964set}'s asymmetric filters (used
-in the X-13ARIMA seasonal adjustment algorithm).
-
-The initial series is modelled here as:
+in the X-11 seasonal adjustment algorithm).
+
+Rewriting equation~\ref{eq-lpp}:
 \begin{equation}\phantomsection\label{eq-lpgeneralmodel}{
-\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
+\boldsymbol y=\begin{pmatrix}\boldsymbol U &\boldsymbol Z\end{pmatrix}
+\begin{pmatrix}\boldsymbol \gamma \\ \boldsymbol \delta\end{pmatrix}+\boldsymbol \varepsilon,\quad
 \boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
 }\end{equation} where \([\boldsymbol U,\boldsymbol Z]\) is of full rank
-and forms a subset of the columns of \(X\). The objective is to find a
-filter \(\boldsymbol v\) that minimises the mean square error of
-revision (to the symmetric filter \(\boldsymbol \theta\)) under certain
-constraints. These constraints are represented by the matrix
-\(\boldsymbol U=\transp{\begin{pmatrix} \transp{\boldsymbol U_{p}}& \transp{\boldsymbol U_{f}}\end{pmatrix}}\):
-\(\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta\) (with
-\(\boldsymbol U_p\) the matrix \((h+q+1)\times (d+1)\) which contains
-the observations of the matrix \(U\) known during estimation by the
-asymmetric filter). The problem is equivalent to finding \(v\) which
-minimises: \begin{equation}\phantomsection\label{eq-lppasym}{
+and forms a subset of the columns of \(X.\) Defining the variance error
+as \(\boldsymbol D = \sigma^2\boldsymbol K^{-1}\) (to directly take into
+account the weights associated to the observations) and with
+\([\boldsymbol U,\boldsymbol Z] = \boldsymbol X\) we this find the model
+associated to symmetric filters. The objective is to find an asymmetric
+filter \(\boldsymbol v\) which uses \(q\) points in the future and
+minimises the mean square error of revision (to the symmetric filter
+\(\boldsymbol \theta\)) under certain (polynomial) constraints. The
+matrices \(\boldsymbol y\), \(\boldsymbol U\), \(\boldsymbol Z\) and
+\(\boldsymbol \theta\) are partitioned as follows: \[
+\boldsymbol y = \begin{pmatrix} \boldsymbol y_p \\ \boldsymbol y_f \end{pmatrix}, \quad
+\boldsymbol U = \begin{pmatrix} \boldsymbol U_p \\ \boldsymbol U_f \end{pmatrix}, \quad  
+\boldsymbol Z = \begin{pmatrix} \boldsymbol Z_p \\ \boldsymbol Z_f \end{pmatrix}, \quad 
+\boldsymbol \theta = \begin{pmatrix} \boldsymbol \theta_p \\ \boldsymbol \theta_f \end{pmatrix},
+\] where \(\boldsymbol y_p\) is a matrix \((h+q+1)\times 1\) which
+denotes the set of available observations (from \(y_{t-h}\) to
+\(y_{t+q}\)) and \(\boldsymbol U\), \(\boldsymbol Z\) and
+\(\boldsymbol \theta\) are partitioned in the same way. The constraints
+are defined as
+\(\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta\):
+they represents the polynomial preservation constraints we want to
+impose on the asymmetric filter. \textcite{proietti2008} show that the
+problem is equivalent to finding \(v\) which minimises:
+\begin{equation}\phantomsection\label{eq-lppasym}{
 \varphi(\boldsymbol v)=
 \underbrace{
-  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
-  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
-  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
+  \underbrace{\transp{(\boldsymbol v-\boldsymbol \theta_{p})}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
+  \transp{\boldsymbol \theta_{f}}\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
+  \underbrace{[\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}}_{bias^2}
 }_\text{mean square revision error}+
-\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
+\underbrace{2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)}_{\text{constraints}}.
 }\end{equation} where \(\boldsymbol l\) is a vector of Lagrange
@@
 
-When \(\boldsymbol U=\boldsymbol X\), the constraint is equivalent to
-preserving polynomials of degree \(d\): we find asymmetric direct
+When \(\boldsymbol U=\boldsymbol X\) (and then
+\(\boldsymbol Z = \boldsymbol 0\)), the constraint is equivalent to
+preserving polynomials of degree \(d\) and we find asymmetric direct
 filters (DAF) when \(\boldsymbol D=\boldsymbol K^{-1}\).
@@
 and when the symmetric filter is the Henderson filter, we find the
-asymmetric Musgrave filters (using the Henderson kernel). This filter
-assumes that, for real-time estimation, the data is generated by a
-linear process and that the asymmetric filters preserve the constants
+asymmetric Musgrave filters. This filter assumes that, for real-time
+estimation, the data is generated by a linear process and that the
+asymmetric filters preserve the constants
 (\(\sum v_i=\sum \theta_i=1\)). These asymmetric filters depend on the
@@
 \(\delta_1/\sigma=2/(R\sqrt{\pi})\). This ratio is used in particular in
-X-13ARIMA to determine the length of the Henderson filter. For monthly
-data:
+X-11 to determine the length of the Henderson filter. For monthly data:
 
@@
 
-Table~\ref{tab-criteriaLp} compares the quality criteria of the
-different methods using the Henderson filter and \(h=6\) (a symmetric
-filter with 13 terms). For real-time filters (\(q=0\)), a more complex
-asymmetrical filter (in terms of polynomial preservation) results in
-lower timeliness and higher fidelity/smoothness: reduction in phase
-shift comes at the expense of an increase in variance. This result
-varies as \(q\) increases: for \(q=2\), the QL filter has greater
-timeliness than the LC filter. This surprising result highlights the
-fact that the phase shift is not controlled by \textcite{proietti2008}'s
-approach.
-
-In terms of revision (\(A_w+S_w+T_w+R_w\)), the LC and QL filters
-consistently outperform the CQ and DAF filters.
-
-These theoretical properties are consistent with the empirical results
-observed in Section~\ref{sec-comparison}. Revisions are more important
-for the CQ and DAF filters, which leads to greater variability in
-estimations and a longer delay in detecting turning points. The globally
-parameterised QL filter (with \(R\) fixed for all asymmetric filters)
-may result in later detection of turning points than the LC filter.
-Indeed, to identify a turning point at date \(t\), it is necessary to
-have knowledge of at least two points after the given date to confirm
-the reversal in trend.
-
-\begin{table}
-
-\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
-
-\centering{
-
-[!h]
-
-\centering
-\begin{tabular}[t]{cccccccccccc}
-\toprule
-Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=0$}}\\
-\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
-\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=1$}}\\
-\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
-\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=2$}}\\
-\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
-\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
-\bottomrule
-\end{tabular}
-
-\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
-
-}
-
-\end{table}%
-
 \subsection{Extension with the timeliness
@@
 
-One drawback of the previous method is the lack of control over the
-phase shift. However, it is possible to improve the modelling by
-incorporating in Equation~\ref{eq-lppasym} the timeliness criterion
-defined by \textcite{ch15HBSA}. This was proposed by Jean Palate, then
-coded in Java and integrated into \texttt{rjd3filters}.
-
-Using the same notation as in Section~\ref{sec-lppasymf},
+One drawback of the approach of \textcite{proietti2008} is the lack of
+control over the phase shift. However, it is possible to improve the
+modelling by incorporating in equation~\ref{eq-lppasym} the timeliness
+criterion defined by \textcite{ch15HBSA}. This was proposed by Jean
+Palate, then coded in Java and integrated into \texttt{rjd3filters}.
+
+To measure the phase shift between the input series and the filtered
+one, equation~\ref{eq-lppasym} introduces a criterion, \(T_g\), called
+\emph{timeliness}. When a moving average \(\boldsymbol v\) is applied,
+the level input signal is also altered by the gain function. That's why
+the timeliness criterion depends on the gain and phase shift functions
+(\(\rho_{\boldsymbol v}\) and \(\varphi_{\boldsymbol v}\)), the link
+between both functions being made by a penalty function \(f\). The
+criterion measures phase shift for the frequencies in the interval
+\([\omega_1, \omega_2]\): \[
+T_g(\boldsymbol v)=\int_{\omega_{1}}^{\omega_{2}}f(\rho_{\boldsymbol v}(\omega),\varphi_{\boldsymbol v}(\omega))\ud\omega
+\] As mentioned in section~\ref{sec-propMM}, in the case of trend-cycle
+extraction we can set \(\omega_1=0\) and for monthly series
+\(\omega_2=2\pi/12.\) For the penalty function, the authors suggest
+\(f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2\) and show thant in
+that case the criteria can be written in a quadratic form: \[
+T_g(\boldsymbol v) = \transp{\boldsymbol v}\boldsymbol T\boldsymbol v
+\] with
+\(T\boldsymbol=\int_{\omega_{1}}^{\omega_{2}}\boldsymbol{Si}(\omega)\transp{\boldsymbol{Si}(\omega)}\ud\omega\)
+and \(\boldsymbol{Si}(\omega)=(\sin(p\omega),\dots, \sin (f\omega))\) if
+\(\boldsymbol v\) uses \(p\) points in the past and \(f\) points in the
+future.
+
+Using the same notation as in section~\ref{sec-lppasymf},
 \(\boldsymbol\theta\) represents the symmetric filter and
@@
 \(\boldsymbol\theta=\transp{\begin{pmatrix}\transp{\boldsymbol\theta_p}&\transp{\boldsymbol\theta_f}\end{pmatrix}}\)
-and \(\boldsymbol\theta_p\) of the same length of \(\boldsymbol v\), and
+and \(\boldsymbol\theta_p\) of the same length as \(\boldsymbol v\), and
 \(\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p\), the timeliness
 criterion can be expressed as: \[
-T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
+T_g(\boldsymbol v)=\transp{\boldsymbol v}\boldsymbol T\boldsymbol v=\transp{\boldsymbol g}\boldsymbol T\boldsymbol g+2\transp{\boldsymbol \theta_p}\boldsymbol T\boldsymbol g+\transp{\boldsymbol \theta_p}\boldsymbol T\boldsymbol \theta_p
 \quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
 \] Furthermore, the objective function \(\varphi\) of
-Equation~\ref{eq-lppasym} can be rewritten: \begin{align*}
-\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
-  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
-  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
-2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
-&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
-\begin{cases}
-\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
-\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
-\boldsymbol c\text{ a constant independent of }\boldsymbol v
-\end{cases}.
-\end{align*} Adding the timeliness criterion: \[
-\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
-2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
-\widetilde {\boldsymbol c}\quad\text{with }
-\begin{cases}
-\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
-\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
-\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
-\end{cases},
-\] where \(\alpha_T\) is the weight associated to the timeliness
-criterion. With \(\alpha_T=0\) we find \(\varphi(\boldsymbol v)\). This
-extension therefore makes it possible to find all the symmetric and
-asymmetric filters presented in the previous section but also
-generalizes the approach of \textcite{GrayThomson1996}. This extension
-is implemented in the R function \texttt{rjd3filters::lp\_filter()}.
+equation~\ref{eq-lppasym} can be rewritten: \begin{align*}
+\varphi(\boldsymbol v)&=\transp{(\boldsymbol v-\boldsymbol \theta_p)}\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
+  \transp{\boldsymbol \theta_f}\boldsymbol D_{f}\boldsymbol \theta_f+
+  [\transp{\boldsymbol \delta}(\transp{\boldsymbol Z_{p}}\boldsymbol v-\transp{\boldsymbol Z}\boldsymbol \theta)]^{2}+
+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)\\
+&=\transp{\boldsymbol g}\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)+\boldsymbol c,
+\end{align*} with
+\(\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z}_p,\)
+\(\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}\)
+and \(\boldsymbol c\) a constant independent of \(\boldsymbol v.\)
+
+Adding the timeliness criterion: \[
+\widetilde\varphi(\boldsymbol v)=\transp{\boldsymbol g}\widetilde {\boldsymbol Q}\boldsymbol g-
+2\widetilde {\boldsymbol P}\boldsymbol g+2\transp{\boldsymbol l}(\transp{\boldsymbol U_{p}}\boldsymbol v-\transp{\boldsymbol U}\boldsymbol \theta)+
+\widetilde {\boldsymbol c},
+\] where \(\alpha_T\) is the weight associated with the timeliness
+criterion,
+\(\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p} + \alpha_T\boldsymbol T\),
+\(\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}-\alpha_T\boldsymbol \theta_p\boldsymbol T\)
+and \(\widetilde {\boldsymbol c}\) a constant independent of
+\(\boldsymbol v.\)
+
+With \(\alpha_T=0\) we find \(\varphi(\boldsymbol v)\). This extension
+therefore makes it possible to find all the symmetric and asymmetric
+filters presented in the previous section but also generalizes the
+approach of \textcite{GrayThomson1996}. This extension is implemented in
+the R function \texttt{rjd3filters::lp\_filter()}.
 
@@
 calibrate the value of \(\alpha_T\) since the value of
-\(\lvert\delta/\sigma\rvert\) must also be defined. That's why in this
-paper we will not compare this extension to the other methods and we
-will focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
+\(\lvert\delta/\sigma\rvert\) must also be defined. Hence, in this paper
+we will not compare this extension to the other methods and we will
+focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
 However, in future studies, we could imagine a two step calibration: fix
 the value of \(\lvert\delta/\sigma\rvert\) to find \(\alpha_T\) by
-cross-validation and then use this weight with a local parameterisation
+cross-validation and then use this weight with a local parametrisation
 of the ratio \(\lvert\delta/\sigma\rvert\) (as in
-Section~\ref{sec-localic}).
-
-\subsection{Local parameterisation of asymmetric
+section~\ref{sec-localic}).
+
+\subsection{Local parametrisation of asymmetric
 filters}\label{sec-localic}
 
-Asymmetric filters are usually parameterised globally:
-\(\lvert\delta/\sigma\rvert\) estimated on all the data using the
-IC-ratio or a cross-validation criterion. However, we might prefer a
-local parameterisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
-as a function of time. Indeed, although the overall parameterisation is
+Asymmetric filters are usually parametrised globally:
+\(\lvert\delta/\sigma\rvert\) is fixed using all the data (with the
+IC-ratio or a cross-validation criterion). However, we might prefer a
+local parametrisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
+as a function of time. Indeed, although the overall parametrisation is
 generally valid, assuming a constant value of the
-\(\lvert\delta/\sigma\rvert\) ratio for all asymmetrical filters does
-not appear relevant for real-time estimation, especially during periods
-of economic downturn. For example, with the LC method, global
-parameterisation means assuming that the slope of the trend is constant,
-whereas during economic downturns it tends towards 0 up to the turning
-point.
-
-This is what is proposed in this article, with a local parameterisation
+\(\lvert\delta/\sigma\rvert\) ratio for all asymmetric filters does not
+appear relevant for real-time estimation, especially during periods of
+economic downturn. For example, with the LC method, global
+parametrisation means assuming that the slope of the trend is constant,
+whereas during economic downturns it tends towards 0 up around the
+turning point.
+
+This is what is proposed in this article, with a local parametrisation
 of the asymmetric filters by estimating \(\delta\) and \(\sigma^2\)
@@
 \begin{itemize}
-\tightlist
 \item
   The variance \(\sigma^2\) can be estimated using the observed data set
-  and the symmetrical filter \((w_{-p},\dots,w_p)\): \[
-  \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
+  and the symmetric filter \((\theta_{-h},\dots,\theta_h)\). Indeed, as
+  for example shown by \textcite{Loader1999}, the variance \(\sigma^2\)
+  can be estimating using the normalized residual sum of squares: \[
+  \hat\sigma^2=\frac{1}{(n-2h)-2\nu_1+\nu_2}\sum_{t=h+1}^{n-h}(y_t-\widehat{TC}_t)^2.
+  \] Since the symmetric filter is used, only \(n-2h\) observations can
+  be used to estimate \(\sigma^2\). \(\nu_1\) and \(\nu_2\) are two
+  definitions of degrees of freedom of a local fit (generalization of
+  the number of parameters of a parametric model). In the case of local
+  polynomial smoothings, they can be linked to the coefficient of the
+  associated filter: \(\nu_1 = (n-h)\theta_0\) (number of observations
+  multiplied by the central coefficient of the moving average) and
+  \(\nu_2 = (n-h)\sum_{i=-h}^{h} \theta_i^2\) (number of observations
+  multiplied by the sum of squares of coefficients of the moving
+  average). Thus, the variance can be estimated by: \[
+  \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\widehat{TC}_t)^2}{1-2\theta_0^2+\sum_{i=-h}^{h} \theta_i^2}.
   \]
@@
   The parameter \(\delta\) can be estimated by moving average from
-  Equation~\ref{eq-mmsym}. For example, for the LC method we can use the
+  equation~\ref{eq-mmsym}. For example, for the LC method we can use the
   moving average
-  \(\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}\)
+  \(\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}\)
   to obtain a local estimate of the slope and for the QL method we can
   use
-  \(\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}\)
+  \(\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}\)
   to obtain a local estimate of the concavity. The DAF method then
@@
   results in terms of phase shift (using the same methodology as in
-  Section~\ref{sec-comparison}). Furthermore, for the construction of
+  section~\ref{sec-comparison}). Furthermore, for the construction of
   moving averages, the trend can be modelled as being locally of degree
@@
   shift but slightly increases the revisions linked to the first
-  estimate of the trend-cycle. Figure~\ref{fig-mmpenteconcac} shows the
+  estimate of the trend-cycle. figure~\ref{fig-mmpenteconcac} shows the
   moving averages used.
@@
 
-In the empirical applications of Section~\ref{sec-comparison}, the final
-local parameterisation is obtained by estimating \(\delta\) using all
-the data (i.e., using the symmetric filters shown in
-Figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
-of \(\sigma^2\). Figure~\ref{fig-mmpenteconcac-ex} shows an example of
+In the empirical applications of section~\ref{sec-comparison}, the final
+local parametrisation is obtained by estimating \(\delta\) using all the
+data (i.e., using the symmetric filters shown in
+figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
+of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex} shows an example of
 the comparison of estimates of the ratio \(|\delta_1/\sigma|\) with a
-global estimator (IC ratio) and two the local parameterisation:
+global estimator (IC ratio) and two local parametrisations:
 
@@
 \item
-  Figure~\ref{fig-mmpenteconcac-ex-1}: using a symmetric filter to
-  estimate \(\delta_1\) (final local parameterisation).
-\item
-  Figure~\ref{fig-mmpenteconcac-ex-2}: using the asymmetric filter which
-  need two points in the futur to estimate \(\delta_1\). This is closed
+  figure~\ref{fig-mmpenteconcac-ex-1}: using a symmetric filter to
+  estimate \(\delta_1\) (final local parametrisation).
+\item
+  figure~\ref{fig-mmpenteconcac-ex-2}: using the asymmetric filter which
+  need two points in the future to estimate \(\delta_1\). This is closed
   to the real-time estimates since two points in the future are needed
@@
 \subcaption{\label{fig-mmpenteconcac-ex-2}Asymmetric moving average
-using two futur points (\(q=2\)) used for \(\hat \delta_1\).}
+using two future points (\(q=2\)) used for \(\hat \delta_1\).}
 
@@
 
-The different methods are compared on simulated and real data.
-
-For all series, a symmetrical 13-term filter is used. These methods are
-also compared with estimates obtained by extending the series with an
-ARIMA model, then applying a 13-term symmetrical Henderson filter. The
-ARIMA model is determined automatically, using no seasonal lag (the
-series being seasonally adjusted) and no external variables (such as
-external regressors for correction of atypical points).
-
-The performance of the different methods is judged on criteria relating
-to revisions (between two consecutive estimates and in relation to the
-final estimate) and the number of periods required to detect turning
-points.
+The different methods are compared on simulated and real data. For all
+series, a symmetric 13-term filter is used: we then assume than the
+bandwidth is fixed, as in the seasonal adjustement method X-11. A
+nearest neighbour bandwidth is also tested (13-term asymmetric filters:
+real-time estimates uses observation between \(y_{t-12}\) and \(y_{t}\),
+the second asymmetric filters uses obversations between \(y_{t-11}\) and
+\(y_{t+1}\), etc.) but with no improvement in the results (see
+supplemental material). More complexe methods to locally select
+bandwidth, like the one proposed by \textcite{fan1992variable}, were not
+tested. All the methods to create asymmetric filters are also compared
+with estimates obtained by extending the series with an ARIMA model,
+then applying a 13-term symmetric Henderson filter. The ARIMA model is
+determined automatically, using no seasonal lag (the series being
+seasonally adjusted) and no external variables (such as external
+regressors for correction of atypical points). The performance of the
+different methods is judged on criteria relating to revisions (between
+two consecutive estimates and in relation to the final estimate) and the
+number of periods required to detect turning points.
 
@@
 \(\sigma_e^2\) in order to obtain series with different signal-to-noise
-ratios (see Figure~\ref{fig-graphs-data-simul}):
+ratios (see figure~\ref{fig-graphs-data-simul}):
 
@@
 \caption{\label{fig-graphs-data-simul}Simulated series with low
-(\(\sigma_e^2=0.2\) and \(\rho = 3,5\)), medium (\(\sigma_e^2=0.3\) and
+(\(\sigma_e^2=0.2\) and \(\rho = 3.55\)), medium (\(\sigma_e^2=0.3\) and
 \(\rho = 2.0\)) and high variability (\(\sigma_e^2=0.4\) and
@@
 \item
-  Computation of two revision criteria: the average of the relative
-  deviations between the \(q\)\textsuperscript{th} estimate and the last
-  estimate \(MAE_{fe}(q)\) and the average of the relative deviations
-  between the \(q\)\textsuperscript{th} estimate and the
-  \(q+1\)\textsuperscript{th} estimate \(MAE_{qe}(q)\). \[
+  The average of the relative deviations between to the last estimate
+  (comparison of the \(q\)\textsuperscript{th} estimate and the last
+  estimate): \[
   MAE_{fe}(q)=\mathbb E\left[
@@
   }\right|
-  \right]
-  \quad\text{and}\quad
-  MAE_{qe}(q)=\mathbb E\left[
+  \right].
+  \]
+\item
+  The average of the relative deviations between two consecutive
+  estimates (the \(q\)\textsuperscript{th} estimate and the
+  \(q+1\)\textsuperscript{th} estimate): \[
+  MAE_{ce}(q)=\mathbb E\left[
   \left|\frac{
@@
 
-Excluding for the moment the local parameterisations of the polynomial
+Excluding for the moment the local parametrisations of the polynomial
 filters, the linear-constant polynomial filter (LC) seems to give the
 best results in terms of delay in the detection of turning points
-(Figure~\ref{fig-graphstpsimul}). Performance is relatively close to
+(figure~\ref{fig-graphstpsimul}). Performance is relatively close to
 that obtained by extending the series using an ARIMA model. However,
@@
 
-For series with moderate variability, local parameterisation of the LC
+For series with moderate variability, local parametrisation of the LC
 and QL filters reduces the phase shift. For series with high
@@
 general, LC filters always minimise revisions (with relatively small
-effect from the local parameterisation of the filters) and revisions are
+effect from the local parametrisation of the filters) and revisions are
 greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
@@
 \(t\), which creates a discontinuity. This revision is considerably
-reduced by parameterising the filter locally. For polynomial filters
+reduced by parametrising the filter locally. For polynomial filters
 other than the LC filter, the large revisions at the first estimate were
@@
 between the fourth and fifth estimates (as might be expected, as
-highlighted in Section~\ref{sec-mmetprev}).
+highlighted in section~\ref{sec-mmetprev}).
 
@@
 chosen because the economic turnaround is particularly visible in the
-raw series (Figure~\ref{fig-ce16ov-previmp-lp}), because this series is
+raw series (figure~\ref{fig-ce16ov-previmp-lp}), because this series is
 used for dating economic cycles and because the FRED-MD database
@@
 availability of series published on past dates). Studying the series up
-to January 2020, this series has a medium variability: a symmetrical
+to January 2020, this series has a medium variability: a symmetric
 filter of 13 terms is therefore appropriate. Appendix
-\autocite*{ann-ex-r} shows how the code to reproduce the different
+\textbf{?@sec-ann-ex-r} shows how the code to reproduce the different
 plots.
 
-Figure~\ref{fig-ce16ovlp} shows the successive estimates of the
+figure~\ref{fig-ce16ovlp} shows the successive estimates of the
 trend-cycle using the different methods studied.
@@
 
-In this example, local parameterisation does not reduce the phase shift,
+In this example, local parametrisation does not reduce the phase shift,
 but it does reduce revisions. The QL and DAF polynomials lead to greater
@@
 implicit forecasts of the different methods
-(Figure~\ref{fig-ce16ov-previmp-lp}). As a reminder, these are forecasts
-of the raw series which, by applying the symmetrical Henderson filter to
-the extended series, give the same estimates as the asymmetrical moving
+(figure~\ref{fig-ce16ov-previmp-lp}). As a reminder, these are forecasts
+of the raw series which, by applying the symmetric Henderson filter to
+the extended series, give the same estimates as the asymmetric moving
 averages. The forecasts of the ARIMA model are naive and do not take the
 turning point into account, unlike the other methods. Finally, local
-parameterisation of the QL filter produces much more consistent
+parametrisation of the QL filter produces much more consistent
 forecasts.
@@
 \end{figure}%
+
+\section{Discussion}\label{discussion}
+
+\subsection{Outliers and COVID-19}\label{outliers-and-covid-19}
+
+Graph + simulations
+
+Enfin, dans cet article et dans ceux associés aux méthodes étudiées, les
+filtres asymétriques sont appliqués et comparés sur des séries déjà
+désaisonnalisées ou sans saisonnalité (séries simulées). Les révisions
+et donc le déphasage sont limités à 6 mois (lorsque le filtre symétrique
+est de 13 termes) et cela a l'avantage d'isoler les impacts des
+différents filtres des autres processus inhérents à la
+désaisonnalisation. Il y a toutefois deux inconvénients à cette
+simplification :
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  D'une part, l'estimation de la série désaisonnalisée dépend de la
+  méthode utilisée pour extraire la tendance-cycle. Le choix de la
+  méthode utilisée pour l'estimation de la tendance-cycle peut donc
+  avoir un impact bien au-delà de 6 mois.
+\item
+  D'autre part, les moyennes mobiles étant des opérateurs linéaires, ils
+  sont sensibles à la présence de points atypiques. L'application
+  directe des méthodes peut donc conduire à des estimations biaisées, du
+  fait de leur présence, alors que les méthodes de désaisonnalisation
+  (comme la méthode X-13ARIMA-SEATS) ont un module de correction des
+  points atypiques. Par ailleurs, comme notamment montré par
+  \textcite{dagum1996new}, le filtre symétrique final utilisé par
+  X-13ARIMA pour extraire la tendance-cycle (et donc celui indirectement
+  utilisé lorsqu'on applique les méthodes sur les séries
+  désaisonnalisées) laisse passer environ 72 \% des cycles de 9 ou 10
+  mois (généralement associés à du bruit plutôt qu'à la tendance-cycle).
+  Les filtres asymétriques finaux amplifient même les cycles de 9 ou 10
+  mois. Cela peut avoir pour conséquence l'introduction d'ondulations
+  indésirables, c'est-à-dire la détection de faux points de
+  retournement. Ce problème est réduit par la correction des points
+  atypiques (ces cycles étant considérés comme de l'irrégulier). C'est
+  ainsi que le \emph{Nonlinear Dagum Filter} (NLDF) a été développé et
+  consiste à :
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \item
+    appliquer l'algorithme de correction des points atypiques de
+    X-13ARIMA sur la série désaisonnalisée, puis de la prolonger par un
+    modèle ARIMA ;
+  \item
+    effectuer une nouvelle correction des points atypiques en utilisant
+    un seuil bien plus strict et appliquer ensuite le filtre symétrique
+    de 13 termes. En supposant une distribution normale cela revient à
+    modifier 48 \% des valeurs de l'irrégulier.
+  \end{enumerate}
+
+  Les \emph{cascade linear filter} (CLF), notamment étudiés dans
+  \textcite{dagumBianconcini2023}, correspondent à une approximation des
+  NLDF en utilisant un filtre de 13 termes et lorsque les prévisions
+  sont obtenus à partir d'un modèle ARIMA(0,1,1) où \(\theta=0,40.\)
+\end{enumerate}
+
+Une piste d'étude serait alors d'étudier plus précisément l'effet des
+points atypiques sur l'estimation de la tendance-cycle et la détection
+des points de retournement, mais aussi d'explorer de nouveaux types de
+filtres asymétriques fondés sur des méthodes robustes (comme les
+régressions locales robustes, les médianes mobiles, etc.).
+
+\subsection{Other frequencies}\label{other-frequencies}
 
@@
 This paper presents the R package \texttt{rjd3filters}, which implements
-a several methods to build moving average for real-time trend-cycle
+several methods to build moving averages for real-time trend-cycle
 estimation. It also offers various functions for the theoretical
@@
 (linear-constant, LC, and quadratic-linear, QL). In addition,
-parameterising polynomial filters locally as suggested in this paper
+parametrising polynomial filters locally as suggested in this paper
 enables turning points to be detected more quickly (especially for the
 QL filter). Even when the phase shift is not reduced, local
-parameterisation is recommended because it reduces revisions and
-produces intermediate estimates that are more consistent with expected
-future trends. However, with these methods, the length of the filter
-used must be adapted to the variability of the series: if the filter
-used is too short (i.e.~if the variability of the series is ``low''),
-retaining polynomial trends of degree one or less (LC method) produces
-poorer results in terms of detecting turning points.
-
-This study could be extended in many ways.
-
-One possible extension would be to look at the impact of filter length
-on the detection of turning points. Asymmetric filters are calibrated
-using indicators calculated for the estimation of symmetric filters (for
-example, to automatically determine their length), whereas a local
-estimate might be preferable. Furthermore, we have only focused on
-monthly series with a 13-term symmetric filter, but the results may be
-different if the symmetric filter studied is longer or shorter and if we
-study series with other frequencies (quarterly or daily, for example).
+parametrisation is recommended because it reduces revisions and produces
+intermediate estimates that are more consistent with expected future
+trends. However, with these methods, the length of the filter used must
+be adapted to the variability of the series: if the filter used is too
+short (i.e.~if the variability of the series is ``low''), retaining
+polynomial trends of degree one or less (LC method) produces poorer
+results in terms of detecting turning points.
+
+This study could be extended in many ways. One possible extension would
+be to look at the impact of filter length on the detection of turning
+points. Asymmetric filters are calibrated using indicators calculated
+for the estimation of symmetric filters (for example, to automatically
+determine their length), whereas a local estimate might be preferable.
+Furthermore, we have only focused on monthly series with a 13-term
+symmetric filter, but the results may be different if the symmetric
+filter studied is longer or shorter and if we study series with other
+frequencies (quarterly or daily, for example).
 
@@
 moving averages, like any linear operator, are very sensitive to the
-presence of atypical points. To limit their impact, in X-13ARIMA a
-strong correction for atypical points is performed on the irregular
-component before applying the filters to extract the trend-cycle. This
-leads us to study the impact of these points on the estimation of the
-trend-cycle and turning points, and also to explore new types of
-asymmetric filters based on robust methods (such as robust local
-regressions or moving medians).
+presence of atypical points. To limit their impact, in X-11 a strong
+correction for atypical points is performed on the irregular component
+before applying the filters to extract the trend-cycle. This leads us to
+study the impact of these points on the estimation of the trend-cycle
+and turning points, and also to explore new types of asymmetric filters
+based on robust methods (such as robust local regressions or moving
+medians).
 
@@
 
-\section{R code example}\label{ann-ex-r}
-
-In this appendix we present the R code that can be used to estimate
-trend-cycle components for the US Employment with local polynomial
-filters.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# \# To install rjd3filters}
-\CommentTok{\# remotes::install\_github("rjdemetra/rjd3toolkit")}
-\CommentTok{\# remotes::install\_github("rjdemetra/rjd3x11plus")}
-\CommentTok{\# remotes::install\_github("rjdemetra/rjd3filters")}
-\FunctionTok{library}\NormalTok{(rjd3filters)}
-\FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(patchwork)}
-\FunctionTok{library}\NormalTok{(zoo)}
-\FunctionTok{library}\NormalTok{(forecast)}
-\NormalTok{file }\OtherTok{\textless{}{-}}\StringTok{"https://files.stlouisfed.org/files/htdocs/fred{-}md/monthly/2022{-}11.csv"}
-\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(file)}
-\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ data[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,] }\CommentTok{\# First line removed: it contains the transformation codes}
-\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{ts}\NormalTok{(}\FunctionTok{log}\NormalTok{(data[, }\StringTok{"CE16OV"}\NormalTok{]), }\CommentTok{\# We study US employment}
-         \AttributeTok{start =} \DecValTok{1959}\NormalTok{, }\AttributeTok{frequency =} \DecValTok{12}\NormalTok{)}
-\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =} \FunctionTok{c}\NormalTok{(}\DecValTok{2001}\NormalTok{, }\DecValTok{9}\NormalTok{))}
-\NormalTok{last\_dates }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(y), }\DecValTok{8}\NormalTok{))}
-\FunctionTok{names}\NormalTok{(last\_dates) }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(zoo}\SpecialCharTok{::}\FunctionTok{as.yearmon}\NormalTok{(last\_dates))}
-\NormalTok{last\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(last\_dates, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =}\NormalTok{ x))}
-
-\CommentTok{\# MA of length 13 is appropriated:}
-\FunctionTok{select\_trend\_filter}\NormalTok{(y)}
-\CommentTok{\# Final estimates:}
-\NormalTok{tc\_f }\OtherTok{\textless{}{-}} \FunctionTok{henderson}\NormalTok{(y, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{)}
-\FunctionTok{plot}\NormalTok{(y)}
-\FunctionTok{lines}\NormalTok{(tc\_f, }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
-\NormalTok{forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(}\FunctionTok{ts.union}\NormalTok{(y, tc\_f))}
-
-\CommentTok{\# IC{-}ratio computation}
-\NormalTok{icr }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-  \FunctionTok{ic\_ratio}\NormalTok{(x, }\FunctionTok{henderson}\NormalTok{(x, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{))}
-\NormalTok{\})}
-\NormalTok{lp\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
-\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
-\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
-                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
-                         \AttributeTok{endpoints =}\NormalTok{ method,}
-                         \AttributeTok{ic =}\NormalTok{ icr[i])}
-\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(last\_est[[i]], lp\_coef)}
-\NormalTok{  \})}
-  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
-\NormalTok{  res}
-\NormalTok{\})}
-\NormalTok{lp\_if }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
-\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
-\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
-                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
-                         \AttributeTok{endpoints =}\NormalTok{ method,}
-                         \AttributeTok{ic =}\NormalTok{ icr[i])}
-    \FunctionTok{implicit\_forecast}\NormalTok{(last\_est[[i]], lp\_coef)}
-\NormalTok{  \})}
-  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
-\NormalTok{  res}
-\NormalTok{\})}
-\FunctionTok{names}\NormalTok{(lp\_est) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(lp\_if) }\OtherTok{\textless{}{-}}  \FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{)}
-
-\CommentTok{\# Local estimates of IC{-}ratios}
-\CommentTok{\# We replicate the direct estimates to have}
-\CommentTok{\# estimators of the slope and the concavity}
-\NormalTok{gen\_MM }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{p=}\DecValTok{6}\NormalTok{, }\AttributeTok{q=}\NormalTok{p, }\AttributeTok{d=}\DecValTok{2}\NormalTok{)\{}
-\NormalTok{  X\_gen }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{d =} \DecValTok{1}\NormalTok{, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{q =}\NormalTok{ p)\{}
-    \FunctionTok{sapply}\NormalTok{(}\DecValTok{0}\SpecialCharTok{:}\NormalTok{d, }\ControlFlowTok{function}\NormalTok{(exp) }\FunctionTok{seq}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{p, q)}\SpecialCharTok{\^{}}\NormalTok{exp)}
-\NormalTok{  \}}
-\NormalTok{  k }\OtherTok{=}\NormalTok{ rjd3filters}\SpecialCharTok{::}\FunctionTok{get\_kernel}\NormalTok{(}\StringTok{"Henderson"}\NormalTok{, }\AttributeTok{h =}\NormalTok{ p)}
-\NormalTok{  k }\OtherTok{=} \FunctionTok{c}\NormalTok{(}\FunctionTok{rev}\NormalTok{(k}\SpecialCharTok{$}\NormalTok{coef[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{]), k}\SpecialCharTok{$}\NormalTok{coef[}\FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,q)}\SpecialCharTok{+}\DecValTok{1}\NormalTok{])}
-\NormalTok{  K }\OtherTok{=} \FunctionTok{diag}\NormalTok{(k)}
-\NormalTok{  X }\OtherTok{=} \FunctionTok{X\_gen}\NormalTok{(}\AttributeTok{d=}\NormalTok{d, }\AttributeTok{p =}\NormalTok{ p, }\AttributeTok{q =}\NormalTok{ q)}
-\NormalTok{  e1 }\OtherTok{=}\NormalTok{ e2 }\OtherTok{=}\NormalTok{ e3 }\OtherTok{=} \FunctionTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{, }\AttributeTok{nrow =}\NormalTok{ d}\SpecialCharTok{+}\DecValTok{1}\NormalTok{)}
-\NormalTok{  e1[}\DecValTok{1}\NormalTok{] }\OtherTok{=} \DecValTok{1}
-\NormalTok{  e2[}\DecValTok{2}\NormalTok{] }\OtherTok{=} \DecValTok{1}
-\NormalTok{  e3[}\DecValTok{3}\NormalTok{] }\OtherTok{=} \DecValTok{1}
-  \CommentTok{\# Estimator of the constant}
-\NormalTok{  M1 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e1)}
-  \CommentTok{\# Estimator of the slope}
-\NormalTok{  M2 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e2)}
-  \CommentTok{\# Estimor of the concavity}
-\NormalTok{  M3 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e3)}
-\NormalTok{  mm }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\AttributeTok{const =}\NormalTok{ M1, }\AttributeTok{slope =}\NormalTok{ M2, }\AttributeTok{concav =}\NormalTok{ M3)}
-  \FunctionTok{lapply}\NormalTok{(mm, moving\_average, }\AttributeTok{lags =} \SpecialCharTok{{-}}\NormalTok{p)}
-\NormalTok{\}}
-\NormalTok{all\_mm }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{0}\NormalTok{, gen\_MM, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{d =} \DecValTok{2}\NormalTok{)}
-\NormalTok{est\_slope }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{slope,}
-                            \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"slope"}\NormalTok{))}
-\NormalTok{est\_concav }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{concav,}
-                             \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"concav"}\NormalTok{))}
-
-\NormalTok{henderson\_f }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{h=}\DecValTok{6}\NormalTok{)}\SpecialCharTok{@}\NormalTok{sfilter}
-\NormalTok{lp\_filter2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(icr, }\AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)\{}
-\NormalTok{  all\_coef }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(icr, }\ControlFlowTok{function}\NormalTok{(ic)\{}
-    \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =}\NormalTok{ h,}
-              \AttributeTok{kernel =}\NormalTok{ kernel,}
-              \AttributeTok{endpoints =}\NormalTok{ method,}
-              \AttributeTok{ic =}\NormalTok{ ic)}
-\NormalTok{  \})}
-\NormalTok{  rfilters }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{h, }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{    q}\OtherTok{=}\NormalTok{h }\SpecialCharTok{{-}}\NormalTok{i}
-\NormalTok{    all\_coef[[i]][,}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"q=\%i"}\NormalTok{, q)]}
-\NormalTok{  \})}
-  \FunctionTok{finite\_filters}\NormalTok{(henderson\_f, }\AttributeTok{rfilters =}\NormalTok{ rfilters)}
-\NormalTok{\}}
-\NormalTok{loc\_lc\_est }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-\NormalTok{loc\_lc\_if }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-\NormalTok{loc\_ql\_est }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-\NormalTok{loc\_ql\_if }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-
-\DocumentationTok{\#\# Plots}
-\CommentTok{\# Plots with all the estimates}
-\NormalTok{plot\_est }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
-\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, data)}
-\NormalTok{  joint\_data }\OtherTok{\textless{}{-}}
-    \FunctionTok{window}\NormalTok{(joint\_data,}
-           \AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(joint\_data))}
-
-\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
-    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =}\NormalTok{ last\_dates,}
-               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
-               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}
-
-\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
-    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
-    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
-              \AttributeTok{data =}\NormalTok{ data\_legend,}
-              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
-              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{)  }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
-\NormalTok{\}}
-
-\NormalTok{plot\_prevs }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
-\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, }\FunctionTok{lapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    first\_date }\OtherTok{\textless{}{-}} \FunctionTok{time}\NormalTok{(x)[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}} \FunctionTok{deltat}\NormalTok{(x)}
-    \CommentTok{\# The last observed data is added for readability}
-    \FunctionTok{ts}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{end =}\NormalTok{ first\_date), x),}
-       \AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{frequency =} \FunctionTok{frequency}\NormalTok{(x))}
-\NormalTok{  \}))}
-
-\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
-    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =} \FunctionTok{sapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(x), }\DecValTok{1}\NormalTok{)),}
-               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
-               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}
-\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{    forecast}\SpecialCharTok{::}\FunctionTok{autolayer}\NormalTok{(}
-      \FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(y)),}
-      \AttributeTok{colour =} \ConstantTok{FALSE}
-\NormalTok{    ) }\SpecialCharTok{+}
-    \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
-    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
-    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
-              \AttributeTok{data =}\NormalTok{ data\_legend,}
-              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
-              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
-\NormalTok{\}}
-
-\NormalTok{all\_est }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_est, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_est),}
-             \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_est))}
-\NormalTok{all\_if }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_if, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_if),}
-            \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_if))}
-\NormalTok{y\_lim }\OtherTok{\textless{}{-}} \ConstantTok{NULL}
-\NormalTok{all\_plots\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
-  \FunctionTok{names}\NormalTok{(all\_est),}
-  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_est}\NormalTok{(all\_est[[x]]) }\SpecialCharTok{+}
-    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Trend{-}cycle with \%s"}\NormalTok{, x))) }\SpecialCharTok{+}
-    \FunctionTok{coord\_cartesian}\NormalTok{(}\AttributeTok{ylim =}\NormalTok{ y\_lim)}
-\NormalTok{)}
-\NormalTok{all\_plots\_prev }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
-  \FunctionTok{names}\NormalTok{(all\_if),}
-  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_prevs}\NormalTok{(all\_if[[x]]) }\SpecialCharTok{+}
-    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Implicit forecasts with \%s"}\NormalTok{, x)))}
-\NormalTok{)}
-\CommentTok{\# Combine all plots:}
-\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_est, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
-\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_prev, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
 \end{document}

### 2024-04-21 R_fredm/1A_Est_pente_finale.R

@@ -2,8 +2,8 @@ library(AQLThesis)
 library(future)
 library(rjd3filters)
 plan(multisession)
-if(!dir.exists("data_fredm/byseriespente_final_nber_dt"))
-  dir.create("data_fredm/byseriespente_final_nber_dt")
+if(!dir.exists("data_fredm/byseriespente_final_nber"))
+  dir.create("data_fredm/byseriespente_final_nber")
 
 # Dans ce programme, pour paramétrer les méthodes LC et QL :
 # 1. On prend les MM symétriques finales d'estimation de la pente et polynôme degré 2
@@ -32,23 +32,26 @@ gen_MM <- function(p=6, q=p, d=2){
 }
 
 
+
 MM = lapply(3:6, function(h){
   list(pente = list(
-    `d=2` = gen_MM(p=h, d=2)[,2],
-    `d=3` = gen_MM(p = h, d=3)[,2]),
-    `deriv2` = list(
-      `d=2` = gen_MM(p=h, d=2)[,3],
-      `d=3` = gen_MM(p = h, d=3)[,3]),
-       henderson = lp_filter(horizon = h)@sfilter
-    )
+    `d=2` = moving_average(gen_MM(p = 6, d=2)[,2], -6),
+    `d=3` = moving_average(gen_MM(p = 6, d=3)[,2], -6)
+  ),
+  `deriv2` = list(
+    `d=2` = moving_average(gen_MM(p = 6, d=2)[,3], -6),
+    `d=3` = moving_average(gen_MM(p = 6, d=3)[,3], -6)
+  ),
+  henderson = lp_filter(horizon = h)@sfilter
+  )
 })
 names(MM) <- sprintf("h=%i", 3:6)
 
 
 for(s in list.files("data_fredm/byseries",full.names = TRUE)){
   print(s)
-  for(h in 6){
-    new_f = sprintf("data_fredm/byseriespente_final_nber_dt/%s_h%i.RDS", gsub(".RDS", "",basename(s)),h)
+  for(h in 3:6){
+    new_f = sprintf("data_fredm/byseriespente_final_nber/%s_h%i.RDS", gsub(".RDS", "",basename(s)),h)
     print(new_f)
     MM_h = MM[[sprintf("h=%i", h)]]
     hend_filter = lp_filter(horizon = h)[,h+1]
@@ -56,28 +59,24 @@ for(s in list.files("data_fredm/byseries",full.names = TRUE)){
       data <- readRDS(s)
       last_est = data[[length(data)]]
       
[…]

### 2024-04-21 R_fredm/2A_Est_final.R

@@ -29,7 +29,7 @@ lp_filter2 <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
 
 fs <- list()
 j <- 1
-h_retain <- 6
+h_retain <- 3:6
 kernel = "Henderson"
 for (method in c("LC","QL")){
   print(method)
@@ -48,7 +48,7 @@ for (method in c("LC","QL")){
                   name_file,
                   tolower(method), h, complement)
         
-        data_info <- readRDS(sprintf("data_fredm/byseriespente_final_nber_dt/%s_h%i.RDS",
+        data_info <- readRDS(sprintf("data_fredm/byseriespente_final_nber/%s_h%i.RDS",
                                      gsub(".RDS", "",basename(s)),h))
         if(all(file.exists(nom_f_s_tp),
                file.exists(nom_f_s)))
@@ -62,7 +62,7 @@ for (method in c("LC","QL")){
             data_t = data_info[[nom_d]][[method]]
             ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
             icr = 2/(sqrt(pi) * ratio)
-            lp_coef = lp_filter2(ic = icr, method = method, h = h, kernel = kernel)
+            lp_coef = lp_filter2(ic = icr, method = method, h = 6, kernel = kernel)
             rjd3filters::filter(x, lp_coef)
           })
           names(series_s) <- names(data)

### 2024-04-21 R_fredm/2B_Est_dafic.R

@@ -1,81 +1,0 @@
-if(!dir.exists("results_fredm"))
-  dir.create("results_fredm")
-if(!dir.exists("results_fredm/localic_daf"))
-  dir.create("results_fredm/localic_daf")
-library(rjd3filters)
-library(AQLThesis)
-library(future)
-plan(multisession)
-
-lp_filter2 <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
-  all_coef = lapply(icr, function(ic){
-    lp_filter(horizon = h,
-              kernel = kernel,
-              endpoints = method,
-              ic = ic)
-  })
-  sym = all_coef[[1]]@sfilter
-  rfilters = lapply(1:h, function(i){
-    q=h -i
-    all_coef[[i]][,sprintf("q=%i", q)]
-  })
-  finite_filters(sym, rfilters = rfilters)
-}
-
-reload <- FALSE
-fs <- list()
-j <- 1
-kernel = "Henderson"
-for (method in c("LC","QL")){
-  print(method)
-  for(s in list.files("data_fredm/byseries", full.names = TRUE)){
-    for(d in 2:3){
-      name_file <- gsub(".RDS$", "", basename(s))
-      print(name_file)
-      data <- readRDS(s)
-      complement = sprintf("_d%i", d)
-      
-      nom_f_s <- sprintf("results_fredm/localic_daf/%s_%s%s.RDS",
-                         name_file, tolower(method), complement)
-      nom_f_s_tp <- 
-        sprintf("results_fredm/localic_daf/%s_%s%s_tp.RDS",
-                name_file,
-                tolower(method), complement)
-      
-      data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber_dt/%s.RDS",
-                                   gsub(".RDS", "",basename(s))))
-      if(all(file.exists(nom_f_s_tp),
-             file.exists(nom_f_s)))
-        next;
-      
-      reload <- TRUE
-      fs[[j]] <- future({
-        print(s)
-        series_s <- lapply(names(data), function(nom_d){
-          x <- data[[nom_d]]
-          data_t = data_info[[nom_d]][[method]]
-          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
-          icr = 2/(sqrt(pi) * ratio)
-          lp_coef = lp_filter2(icr = icr, method = method, h = h, kernel = kernel)
-          filter(x, lp_coef)
-        })
[…]

### 2024-04-21 R_fredm/2C_Est_dafic_trunc.R

@@ -30,48 +30,49 @@ for (method in c("LC","QL")){
   print(method)
   for(s in list.files("data_fredm/byseries", full.names = TRUE)){
     for(d in 2:3){
-      name_file <- gsub(".RDS$", "", basename(s))
-      print(name_file)
-      data <- readRDS(s)
-      complement = sprintf("_d%i", d)
-      
-      nom_f_s <- sprintf("results_fredm/localic_daf_trunc/%s_%s%s.RDS",
-                         name_file, tolower(method), complement)
-      nom_f_s_tp <- 
-        sprintf("results_fredm/localic_daf_trunc/%s_%s%s_tp.RDS",
-                name_file,
-                tolower(method), complement)
-    
-      
-      data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber_dt/%s.RDS",
-                                   gsub(".RDS", "",basename(s))))
-      if(all(file.exists(nom_f_s_tp),
-             file.exists(nom_f_s)))
-        next;
-      
-      reload <- TRUE
-      fs[[j]] <- future({
-        print(s)
-        series_s <- lapply(names(data), function(nom_d){
-          x <- data[[nom_d]]
-          data_t = data_info[[nom_d]][[method]]
-          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
-          icr = 2/(sqrt(pi) * ratio)
-          icr[abs(icr) > 12] <- 12
-          lp_coef = lp_filter2(icr = icr, method = method, h = h, kernel = kernel)
-          rjd3filters::filter(x, lp_coef)
+      for(h in 3:6) {
+        name_file <- gsub(".RDS$", "", basename(s))
+        print(name_file)
+        data <- readRDS(s)
+        complement = sprintf("_d%i", d)
+        
+        nom_f_s <- sprintf("results_fredm/localic_daf_trunc/%s_%s_h%i%s.RDS",
+                           name_file, tolower(method), h, complement)
+        nom_f_s_tp <- 
+          sprintf("results_fredm/localic_daf_trunc/%s_%s_h%i%s_tp.RDS",
+                  name_file,
+                  tolower(method), h, complement)
+        
+        data_info <- readRDS(sprintf("data_fredm/byseriespente_daf_nber/%s_h%i.RDS",
+                                     gsub(".RDS", "",basename(s)),h))
+        if(all(file.exists(nom_f_s_tp),
+               file.exists(nom_f_s)))
+          next;
[…]

### 2024-04-21 R_fredm/3A_summary_of_tp.R

@@ -2,7 +2,6 @@ library(AQLThesis)
 if(!dir.exists("results_fredm/compile_tp"))
   dir.create("results_fredm/compile_tp")
 
-
 all_files <- list.files("results_fredm/lp/",pattern = "_tp",full.names = TRUE)
 
 all_tp_lp <- lapply(seq_along(all_files), function(i){
@@ -59,29 +58,39 @@ saveRDS(all_t, "results_fredm/compile_tp/troughs_arima.RDS")
 saveRDS(all_p, "results_fredm/compile_tp/peaks_arima.RDS")
 
 
-all_files <- list.files("results_fredm/fst",
-                        pattern = "_tp",full.names = TRUE)
-all_tp <- lapply(seq_along(all_files), function(i){
-  print(i)
-  f = all_files[i]
-  compute_time_lag(readRDS(f),                    
-                   peaks = nber_tp_m[,"Peak"],
-                   troughs = nber_tp_m[,"Trough"],
-                   type = "first_detection",
-                   n_ahead_max = 12)
-})
+for(dir in c("localic_daf_trunc", "localic_final")){
+  print(dir)
+  all_files <- list.files(sprintf("results_fredm/%s/", dir),pattern = "_tp",full.names = TRUE)
+  all_tp_lp <- lapply(seq_along(all_files), function(i){
+    print(i)
+    f = all_files[i]
+    compute_time_lag(readRDS(f),                    
+                     peaks = nber_tp_m[,"Peak"],
+                     troughs = nber_tp_m[,"Trough"],
+                     n_ahead_max = 12)
+  })
+  
+  full_names <- gsub("_tp.RDS$", "", basename(all_files))
+  split <- strsplit(full_names, "_")
+  series <- sapply(split, `[`, 1)
+  method <- sapply(split, `[`, 2)
+  h <- sapply(split, `[`, 3)
+  degree <- sapply(split, `[`, 4)
+  # if(length(grep("localic_daf", dir)) >0){
+  #   h <- "h6"
+  #   degree <- sapply(split, `[`, 3)
+  # } else {
+  #   h <- sapply(split, `[`, 3)
+  #   degree <- sapply(split, `[`, 4)
+  # }
+  degree[is.na(degree)] <- "d3"
+  all_t <- data.frame(t(sapply(all_tp_lp, function(x) x[["troughs"]][["phaseshift"]])),
[…]

### 2024-04-21 R_fredm/3B_summary_of_tp_norev.R

@@ -63,3 +63,41 @@ saveRDS(all_t, "results_fredm/compile_tp_norev/troughs_arima.RDS")
 saveRDS(all_p, "results_fredm/compile_tp_norev/peaks_arima.RDS")
 
 
+
+for(dir in c("localic_daf_trunc", "localic_final")){
+  print(dir)
+  all_files <- list.files(sprintf("results_fredm/%s/", dir),pattern = "_tp",full.names = TRUE)
+  all_tp_lp <- lapply(seq_along(all_files), function(i){
+    print(i)
+    f = all_files[i]
+    compute_time_lag(data = readRDS(f),                    
+                     peaks = nber_tp_m[,"Peak"],
+                     troughs = nber_tp_m[,"Trough"],
+                     type = "no_revisions",
+                     n_ahead_max = 12)
+  })
+  
+  full_names <- gsub("_tp.RDS$", "", basename(all_files))
+  split <- strsplit(full_names, "_")
+  series <- sapply(split, `[`, 1)
+  method <- sapply(split, `[`, 2)
+  h <- sapply(split, `[`, 3)
+  degree <- sapply(split, `[`, 4)
+  # if(length(grep("localic_daf", dir)) >0){
+  #   h <- "h6"
+  #   degree <- sapply(split, `[`, 3)
+  # } else {
+  #   h <- sapply(split, `[`, 3)
+  #   degree <- sapply(split, `[`, 4)
+  # }
+  degree[is.na(degree)] <- "d3"
+  
+  all_t <- data.frame(t(sapply(all_tp_lp, function(x) x[["troughs"]][["phaseshift"]])),
+                      series, kernel = "henderson", method, h, degree)
+  all_p <- data.frame(t(sapply(all_tp_lp, function(x) x[["peaks"]][["phaseshift"]])),
+                      series, kernel = "henderson", method, h, degree)
+  rownames(all_t) <- rownames(all_p) <- full_names
+  
+  saveRDS(all_t, sprintf("results_fredm/compile_tp_norev/troughs_%s.RDS", dir))
+  saveRDS(all_p, sprintf("results_fredm/compile_tp_norev/peaks_%s.RDS", dir))
+}

### 2024-04-21 R_fredm/3_Summary_tp.R

@@ -1,74 +1,0 @@
-library(AQLThesis)
-if(!dir.exists("results_fredm/compile_revisions"))
-  dir.create("results_fredm/compile_revisions")
-if(!dir.exists("results_fredm/compile_tp"))
-  dir.create("results_fredm/compile_tp")
-if(!dir.exists("results_fredm/compile_tp_norev"))
-  dir.create("results_fredm/compile_tp_norev")
-
-for(dir in c("localic_daf", "localic_daf_trunc", "localic_final")){
-  print(dir)
-  all_files <- list.files(sprintf("results_fredm/%s/", dir),pattern = "_tp",full.names = TRUE)
-  all_tp_lp <- lapply(seq_along(all_files), function(i){
-    print(i)
-    f = all_files[i]
-    compute_time_lag(readRDS(f),                    
-                     peaks = nber_tp_m[,"Peak"],
-                     troughs = nber_tp_m[,"Trough"],
-                     n_ahead_max = 12)
-  })
-  
-  full_names <- gsub("_tp.RDS$", "", basename(all_files))
-  split <- strsplit(full_names, "_")
-  series <- sapply(split, `[`, 1)
-  method <- sapply(split, `[`, 2)
-  if(length(grep("localic_daf", dir)) >0){
-    h <- "h6"
-    degree <- sapply(split, `[`, 3)
-  } else {
-    h <- sapply(split, `[`, 3)
-    degree <- sapply(split, `[`, 4)
-  }
-  degree[is.na(degree)] <- "d3"
-  all_t <- data.frame(t(sapply(all_tp_lp, function(x) x[["troughs"]][["phaseshift"]])),
-                      series, kernel = "henderson", method, h, degree)
-  all_p <- data.frame(t(sapply(all_tp_lp, function(x) x[["peaks"]][["phaseshift"]])),
-                      series, kernel = "henderson", method, h, degree)
-  rownames(all_t) <- rownames(all_p) <- full_names
-  
-  saveRDS(all_t, sprintf("results_fredm/compile_tp/troughs_%s.RDS", dir))
-  saveRDS(all_p, sprintf("results_fredm/compile_tp/peaks_%s.RDS", dir))
-  
-  all_tp_lp_localic <- lapply(seq_along(all_files), function(i){
-    print(i)
-    f = all_files[i]
-    compute_time_lag(readRDS(f),                    
-                     peaks = nber_tp_m[,"Peak"],
[…]

### 2024-04-21 R_fredm/4A_Analyse_est.R

@@ -6,27 +6,33 @@ all_tp <- get_all_tp(dossier = "results_fredm/compile_tp_norev/")
 all_tp_rev <- get_all_tp(dossier = "results_fredm/compile_tp/")
 detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
 
-
 data <- readRDS("data_fredm/byseries/CE16OV.RDS")[["2022"]]
 plot(data)
 all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
                            series = "CE16OV", tp_keep ="2001.16666666667",
                            nb_est = 8)
-wrap_plots(all_plots, ncol = 3) & scale_color_grey()
+legende <- c(lc = "LC",
+             lc_localic_final_h6_d2 = "LC loc. param.\n(final estimates)",
+             lc_localic_h6_d2 = "LC loc.\nparam.",
+             ql = "QL", ql_localic_final_h6_d2 = "QL loc. param.\n(final estimates)",
+             ql_localic_h6_d2 = "QL loc.\nparam.",
+             cq = "CQ", daf = "DAF",
+             auto_arima = "ARIMA")
+wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey()
 
 ggsave("paper/img/nber/ce16ov_fev2001_lp.pdf",
-       plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
+       plot = wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 
 all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
                            series = "CE16OV", tp_keep ="2020.25",
                            nb_est = 8)
-wrap_plots(all_plots, ncol = 3) & scale_color_grey()
+wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey()
 
 ggsave("paper/img/nber/ce16ov_covid_lp.pdf",
-       plot = wrap_plots(all_plots, ncol = 3) & scale_color_grey(),
+       plot = wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey(),
        width = 8, height = 8)
 
 data <- readRDS("data_fredm/byseries/RETAILx.RDS")[["2022"]]
@@ -35,15 +41,9 @@ all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
[…]

### 2024-04-21 R_fredm/4B_Analyse_fred.R

@@ -1,113 +1,0 @@
-source("R_fredm/4_utils.R",encoding = "UTF-8")
-library(patchwork)
-
-all_tp <- merge(readRDS("results_fredm/compile_tp_norev/troughs_lp.RDS"),
-                readRDS("results_fredm/compile_tp_norev/peaks_lp.RDS"),
-                by=c("series","kernel", "method")) %>%
-  select_var()
-
-all_tp_arima <- 
-  merge(readRDS("results_fredm/compile_tp_norev/troughs_arima.RDS"),
-        readRDS("results_fredm/compile_tp_norev/peaks_arima.RDS"),
-        by=c("series","kernel", "method")) %>%
-  select_var() %>% mutate(article = "arima")
-
-all_tp_lic <- 
-  merge(readRDS("results_fredm/compile_tp_norev/troughs_lp_localic.RDS"),
-        readRDS("results_fredm/compile_tp_norev/peaks_lp_localic.RDS"),
-        by=c("series","h", "degree", "method")) %>%
-  select_var() %>% 
-  mutate(article = "localic",
-         kernel = "henderson",
-         method = paste(method,h,degree,sep="_"))
-all_tp_lic <- all_tp_lic[,colnames(all_tp_rkhs)]
-
-
-all_tp <- all_tp_rkhs %>%
-  rbind(all_tp_arima, all_tp_lic) %>% 
-  mutate(method = factor(method,levels = c("lc","ql","cq","daf", "frf", "gain", "phase","auto_arima", "lc_h3_d2", "lc_h3_d3", "ql_h3_d3", "lc_h4_d2", "lc_h4_d3", 
-                                           "ql_h4_d3", "lc_h5_d2", "lc_h5_d3", "ql_h5_d3", "lc_h6_d2", "lc_h6_d3", 
-                                           "ql_h6_d3"),
-                         ordered = TRUE),
-         kernel = tolower(kernel))
-all_rev_fe <- all_rev_rkhs_fe %>% 
-  mutate(method = factor(method,levels = c("lc","ql","cq","daf", "frf", "gain", "phase"),
-                         ordered = TRUE),
-         kernel = tolower(kernel))
-all_rev_ce <- all_rev_rkhs_ce %>% 
-  mutate(method = factor(method,levels = c("lc","ql","cq","daf", "frf", "gain", "phase"),
-                         ordered = TRUE),
-         kernel = tolower(kernel))
-
-
[…]

### 2024-04-21 R_fredm/4B_Analyse_implicit_forecast.R

@@ -10,17 +10,17 @@ all_tp_rev <- get_all_tp(dossier = "results_fredm/compile_tp/")
 
 detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
 
-
+all_tp |> dplyr::filter(series == "CE16OV")
 series <- "CE16OV"
 tp_keep <- "2001.16666666667"
 all_prevs <- get_all_prevs(series = series, tp_keep = tp_keep,
-                           nb_est = 8, nb_dates_before = 6)
+                           nb_est = 8, nb_dates_before = 6,
+                           d_localic = 2, h_localic = 6)
 plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              series = series,
                              all_tp = all_tp,
                              all_tp_rev = all_tp_rev,
-                             tp_keep = tp_keep,
-                             dossier = "results_fredm/compile_tp_norev/")
+                             tp_keep = tp_keep)
 wrap_plots(plots,ncol = 3) & scale_color_grey()
 
 ggsave("paper/img/nber/ce16ov_fev2001_prev_imp_lp.pdf",
@@ -30,13 +30,13 @@ ggsave("paper/img/nber/ce16ov_fev2001_prev_imp_lp.pdf",
 series <- "CE16OV"
 tp_keep <- "2020.25"
 all_prevs <- get_all_prevs(series = series, tp_keep = tp_keep,
-                           nb_est = 8, nb_dates_before = 6)
+                           nb_est = 8, nb_dates_before = 6,
+                           d_localic = 2, h_localic = 6)
 plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              series = series,
                              all_tp = all_tp,
                              all_tp_rev = all_tp_rev,
-                             tp_keep = tp_keep,
-                             dossier = "results_fredm/compile_tp_norev/")
+                             tp_keep = tp_keep)
 wrap_plots(plots,ncol = 3) & 
   scale_color_grey() & 
   labs(subtitle = NULL)
@@ -52,17 +52,13 @@ series <- "RETAILx"
 tp_keep <- "2007.91666666667"
 
 all_prevs <- get_all_prevs(series = series, tp_keep = tp_keep,
-                           nb_est = 8, nb_dates_before = 6,
-                           fst_weights = fst_weights,
-                           fst_degree = fst_degree)
+                           nb_est = 8, nb_dates_before = 6)
 plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              series = series,
[…]

### 2024-04-21 R_fredm/4_utils.R

@@ -72,8 +72,7 @@ select_mae <- function(x){
 extract_est_data <- function(method = "lc", kernel = "henderson", nb_est = 10,
                              series = "RETAILx",
                              tp_date = 2020.25,
-                             nb_dates_before = 6,
-                             fst_degree = 2){
+                             nb_dates_before = 6){
   sep = "_"
   if(method %in% c("lc","ql", "cq", "daf")){
     dir <- "lp"
@@ -85,18 +84,12 @@ extract_est_data <- function(method = "lc", kernel = "henderson", nb_est = 10,
     } else if (length(grep("_localic", method)) >0) {
       if (length(grep("_final", method)) >0) {
         dir <- "localic_final"
-        full_name <- gsub("localic_final", "h6_d2", method)
+        full_name <- gsub("localic_final_", "", method)
       } else {
         dir <- "localic_daf_trunc"
-        full_name <- gsub("localic", "d2", method)
+        full_name <- gsub("localic_", "", method)
       }
-    } else if (length(grep("^weight", method)) > 0 ) {
-      dir <- "fst"
-      full_name <- sprintf("degree%s_%s", fst_degree, method)
-    } else {
-      dir <- "rkhs"
-      full_name <- method
-    }
+    } 
   }
   file = sprintf("results_fredm/%s/%s%s%s.RDS", dir, series, sep, full_name)
   data <- readRDS(file)
@@ -125,21 +118,44 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
     select_var()
 
   tp_lic_final <-
-    merge(readRDS(sprintf("%stroughs_localic_daf_trunc.RDS", dossier)),
-          readRDS(sprintf("%speaks_localic_daf_trunc.RDS", dossier)),
+    merge(readRDS(sprintf("%stroughs_localic_final.RDS", dossier)),
+          readRDS(sprintf("%speaks_localic_final.RDS", dossier)),
           by=c("series","kernel", "method", "degree", "h")) %>%
-    dplyr::filter(degree == "d2", h == "h6") %>%
-    select_var() %>% mutate(method = sprintf("%s_localic_final", method)) %>%
+    # dplyr::filter(degree == "d2", h == "h6") %>%
+    select_var() %>% 
+    mutate(method = sprintf("%s_localic_final_%s_%s", method, h, degree)) %>%
     select(!c(degree, h))
 
   tp_lic_final_daf <-
     merge(readRDS(sprintf("%stroughs_localic_daf_trunc.RDS", dossier)),
[…]

### 2024-04-21 R_simul/1B_Est_pente_final.R

@@ -32,12 +32,14 @@ gen_MM <- function(p=6, q=p, d=2){
 
 MM = lapply(3:6, function(h){
   list(pente = list(
-    `d=2` = gen_MM(p=h, d=2)[,2],
-    `d=3` = gen_MM(p = h, d=3)[,2]),
-    `deriv2` = list(
-      `d=2` = gen_MM(p=h, d=2)[,3],
-      `d=3` = gen_MM(p = h, d=3)[,3]),
-    henderson = lp_filter(horizon = h)@sfilter
+    `d=2` = moving_average(gen_MM(p = 6, d=2)[,2], -6),
+    `d=3` = moving_average(gen_MM(p = 6, d=3)[,2], -6)
+  ),
+  `deriv2` = list(
+    `d=2` = moving_average(gen_MM(p = 6, d=2)[,3], -6),
+    `d=3` = moving_average(gen_MM(p = 6, d=3)[,3], -6)
+  ),
+  henderson = lp_filter(horizon = h)@sfilter
   )
 })
 names(MM) <- sprintf("h=%i", 3:6)
@@ -54,21 +56,21 @@ for(s in list.files("data_simul/byseries",full.names = TRUE)){
     if(!file.exists(new_f)){
       data <- readRDS(s)
       last_est = data[[length(data)]]
-      pente_d2 = zoo::na.locf(moving_average(MM_h$pente[[sprintf("d=%i",2)]], -h) * last_est)
-      pente_d3 = zoo::na.locf(moving_average(MM_h$pente[[sprintf("d=%i",3)]], -h) * last_est)
-      courbure_d2 = zoo::na.locf(moving_average(MM_h$deriv2[[sprintf("d=%i",2)]], -h) * last_est)
-      courbure_d3 = zoo::na.locf(moving_average(MM_h$deriv2[[sprintf("d=%i",3)]], -h) * last_est)
+      pente_d2 = zoo::na.locf(MM_h$pente[[sprintf("d=%i",2)]] * last_est)
+      pente_d3 = zoo::na.locf(MM_h$pente[[sprintf("d=%i",3)]] * last_est)
+      courbure_d2 = zoo::na.locf(MM_h$deriv2[[sprintf("d=%i",2)]] * last_est)
+      courbure_d3 = zoo::na.locf(MM_h$deriv2[[sprintf("d=%i",3)]] * last_est)
       
       info <- lapply(data, function(x){
         sigma2 <- var_estimator(x, MM_h[["henderson"]])
         list("LC" = list(
[…]

### 2024-04-21 R_simul/1B_Est_pente_final_NN.R

@@ -1,9 +1,7 @@
 library(AQLThesis)
-library(future)
 library(rjd3filters)
-plan(multisession)
-if(!dir.exists("data_fredm/byseriespente_daf_nber"))
-  dir.create("data_fredm/byseriespente_daf_nber")
+if(!dir.exists("data_simul/byseriespente_nn"))
+  dir.create("data_simul/byseriespente_nn")
 
 # Dans ce programme, pour paramétrer les méthodes LC et QL :
 # 1. On prend les MM asymétriques d'estimation de la pente et polynôme degré 2 et on fait une estimation en temps réel
@@ -32,35 +30,35 @@ gen_MM <- function(p=6, q=p, d=2){
 }
 
 MM = lapply(0:6, function(q){
-  d2 = gen_MM(p=6, q=q, d=2)
-  d3 = gen_MM(p=6, q=q, d=3)
-  add_0 <- matrix(0, ncol = ncol(d2), nrow = 6-q)
-  d2 = rbind(d2, add_0)
-  d3 = rbind(d3, add_0)
+  d2 = gen_MM(p=12-q, q=q, d=2)
+  d3 = gen_MM(p=12-q, q=q, d=3)
+  
   list(pente = list(
-    `d=2` = d2[,2],
-    `d=3` = d3[,2]),
-    `deriv2` = list(
-      `d=2` = d2[,3],
-      `d=3` = d3[,3]))
+    `d=2` = moving_average(d2[,2], -(12-q)) ,
+    `d=3` = moving_average(d3[,2], -(12-q))
+  ),
+  `deriv2` = list(
+    `d=2` = moving_average(d2[,3], -(12-q)),
+    `d=3` = moving_average(d3[,3], -(12-q))
+  ))
 })
 MM = list(pente = list(
-  `d=2` = sapply(MM, function(x) x[[1]][[1]]),
-  `d=3` = sapply(MM, function(x) x[[1]][[2]])
+  `d=2` = finite_filters(MM[[7]][[1]][[1]], sapply(MM[-7], function(x) x[[1]][[1]]), first_to_last = TRUE),
+  `d=3` = finite_filters(MM[[7]][[1]][[2]], sapply(MM[-7], function(x) x[[1]][[2]]), first_to_last = TRUE)
 ),
 `deriv2` = list(
-  `d=2` = sapply(MM, function(x) x[[2]][[1]]),
-  `d=3` = sapply(MM, function(x) x[[2]][[2]])
[…]

### 2024-04-21 R_simul/1C_Est_pente_daf.R

@@ -56,28 +56,31 @@ MM = lapply(MM, function(x){
 })
 s = list.files("data_simul/byseries",full.names = TRUE)[1]
 d = 2
-h=6
-hend_filter = lp_filter(horizon = h)@sfilter
+h = 6
 for(s in list.files("data_simul/byseries",full.names = TRUE)){
-  new_f = sprintf("data_simul/byseriespente_daf/%s.RDS",
-                  gsub(".RDS", "",basename(s)))
-  print(new_f)
-  if(!file.exists(new_f)){
-    data <- readRDS(s)
-    info <- lapply(data, function(x){
-      sigma2 <- var_estimator(x, hend_filter)
-      list("LC" = list(
-        `d=2` = as.numeric(tail(rjd3filters::filter(x, MM$pente$`d=2`),6)),
-        `d=3` = as.numeric(tail(rjd3filters::filter(x, MM$pente$`d=3`),6)),
-        `sigma2` = sigma2
-      ),
-      "QL" = list(
-        `d=2` = as.numeric(tail(rjd3filters::filter(x, MM$deriv2$`d=2`),6)),
-        `d=3` = as.numeric(tail(rjd3filters::filter(x, MM$deriv2$`d=3`),6)),
-        `sigma2` = sigma2
-      )
-      )
-    })
-    saveRDS(info, new_f)
+  print(s)
+  for(h in 3:6){
+    new_f = sprintf("data_simul/byseriespente_daf/%s_h%i.RDS",
+                    gsub(".RDS", "",basename(s)),h)
+    print(new_f)
+    hend_filter <- lp_filter(horizon = h)@sfilter
+    if(!file.exists(new_f)){
+      data <- readRDS(s)
+      info <- lapply(data, function(x){
+        sigma2 <- var_estimator(x, hend_filter)
+        list("LC" = list(
+          `d=2` = as.numeric(tail(rjd3filters::filter(x, MM$pente$`d=2`),6)),
+          `d=3` = as.numeric(tail(rjd3filters::filter(x, MM$pente$`d=3`),6)),
+          `sigma2` = sigma2
+        ),
+        "QL" = list(
+          `d=2` = as.numeric(tail(rjd3filters::filter(x, MM$deriv2$`d=2`),6)),
+          `d=3` = as.numeric(tail(rjd3filters::filter(x, MM$deriv2$`d=3`),6)),
+          `sigma2` = sigma2
+        )
+        )
+      })
+      saveRDS(info, new_f)
+    }
   }
 }

### 2024-04-21 R_simul/2B_Est_final.R

@@ -64,7 +64,7 @@ for (method in c("LC","QL")){
             data_t = data_info[[nom_d]][[method]]
             ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
             icr = 2/(sqrt(pi) * ratio)
-            lp_coef = lp_filter2(icr = icr, method = method, h = h, kernel = kernel)
+            lp_coef = lp_filter2(icr = icr, method = method, h = 6, kernel = kernel)
             res <- rjd3filters::filter(x, lp_coef)
             
             res[1:h] <- NA

### 2024-04-21 R_simul/2B_Est_final_NN.R

@@ -45,7 +45,7 @@ for (method in c("LC","QL")){
         nom_f_s_rev_ce <- sprintf("results_simul/localic_final_nn/%s_%s_h%i%s_ce_rev.RDS",
                                   name_file,
                                   tolower(method), h, complement)
-        data_info <- readRDS(sprintf("data_simul/byseriespente_final/%s_h%i.RDS",
+        data_info <- readRDS(sprintf("data_simul/byseriespente_nn/%s_h%i.RDS",
                                      gsub(".RDS", "",basename(s)),h))
         
         if(all(file.exists(nom_f_s_tp),
@@ -61,7 +61,7 @@ for (method in c("LC","QL")){
             data_t = data_info[[nom_d]][[method]]
             ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
             icr = 2/(sqrt(pi) * ratio)
-            lp_coef = lp_filter2_nn(icr = icr, method = method, h = h, kernel = kernel)
+            lp_coef = lp_filter2_nn(icr = icr, method = method, h = 6, kernel = kernel)
             res <- rjd3filters::filter(x, lp_coef)
             
             res[1:h] <- NA

### 2024-04-21 R_simul/2C_Est_dafic_trunc.R

@@ -35,67 +35,68 @@ for (method in c("LC","QL")){
   print(method)
   for(s in list.files("data_simul/byseries", full.names = TRUE)){
     for(d in 2:3){
-      name_file <- gsub(".RDS$", "", basename(s))
-      print(name_file)
-      data <- readRDS(s)
-      complement = sprintf("_d%i", d)
-      
-      nom_f_s <- sprintf("results_simul/localic_daf_trunc/%s_%s%s.RDS",
-                         name_file, tolower(method), complement)
-      nom_f_s_tp <- 
-        sprintf("results_simul/localic_daf_trunc/%s_%s%s_tp.RDS",
-                name_file,
-                tolower(method), complement)
-      
-      nom_f_s_rev_fe <- sprintf("results_simul/localic_daf_trunc/%s_%s%s_fe_rev.RDS",
-                                name_file,
-                                tolower(method), complement)
-      nom_f_s_rev_ce <- sprintf("results_simul/localic_daf_trunc/%s_%s%s_ce_rev.RDS",
-                                name_file,
-                                tolower(method), complement)
-      data_info <- readRDS(sprintf("data_simul/byseriespente_daf/%s.RDS",
-                                   gsub(".RDS", "",basename(s))))
-      
-      if(all(file.exists(nom_f_s_tp),
-             file.exists(nom_f_s_rev_fe),
-             file.exists(nom_f_s_rev_ce)))
-        next;
-      
-      reload <- TRUE
-      fs[[j]] <- future({
-        print(s)
-        series_s <- lapply(names(data), function(nom_d){
-          x <- data[[nom_d]]
-          data_t = data_info[[nom_d]][[method]]
-          ratio = data_t[[sprintf("d=%i", d)]] / sqrt(data_t[["sigma2"]])
-          icr = 2/(sqrt(pi) * ratio)
-          icr[abs(icr) > 12] <- 12
-          lp_coef = lp_filter2(icr = icr, method = method, h = h, kernel = kernel)
-          res <- rjd3filters::filter(x, lp_coef)
+      for(h in 3:6) {
+        name_file <- gsub(".RDS$", "", basename(s))
+        print(name_file)
+        data <- readRDS(s)
+        complement = sprintf("_d%i", d)
+        
+        nom_f_s <- sprintf("results_simul/localic_daf_trunc/%s_%s_h%i%s.RDS",
+                           name_file, tolower(method), h, complement)
+        nom_f_s_tp <- 
[…]

### 2024-04-21 R_simul/2C_Est_dafic_trunc_nn.R

@@ -1,104 +1,0 @@
-if(!dir.exists("results_simul"))
-  dir.create("results_simul")
-if(!dir.exists("results_simul/localic_daf_trunc_nn"))
-  dir.create("results_simul/localic_daf_trunc_nn")
-library(rjd3filters)
-library(AQLThesis)
-library(future)
-plan(multisession)
-
-kernel = "Henderson"
-method = "LC"
-s = list.files("data_simul/byseries",full.names = TRUE)[1]
-
-d = 2
-h = 6
-
-h_filter <- lp_filter(horizon = h, kernel = kernel)@sfilter
-lp_filter2_nn <- function(icr, method = "LC", h = 6, kernel = "Henderson"){
-  rfilters = lapply(seq_along(icr), function(i){
-    lp_filter(horizon=h+i, 
-              endpoints = method,
-              kernel = kernel,
-              ic = icr[i])[,2*i+1]
-  })
-  finite_filters(h_filter, rfilters = rfilters)
-}
-fs <- list()
-j <- 1
-reload <- FALSE
-l <- 6
-for (method in c("LC","QL")){
-  print(method)
-  for(s in list.files("data_simul/byseries", full.names = TRUE)){
-    for(d in 2:3){
-      name_file <- gsub(".RDS$", "", basename(s))
-      print(name_file)
-      data <- readRDS(s)
-      complement = sprintf("_d%i", d)
-      
-      nom_f_s <- sprintf("results_simul/localic_daf_trunc_nn/%s_%s%s.RDS",
-                         name_file, tolower(method), complement)
-      nom_f_s_tp <- 
-        sprintf("results_simul/localic_daf_trunc_nn/%s_%s%s_tp.RDS",
-                name_file,
-                tolower(method), complement)
-      
-      nom_f_s_rev_fe <- sprintf("results_simul/localic_daf_trunc_nn/%s_%s%s_fe_rev.RDS",
-                                name_file,
-                                tolower(method), complement)
-      nom_f_s_rev_ce <- sprintf("results_simul/localic_daf_trunc_nn/%s_%s%s_ce_rev.RDS",
-                                name_file,
-                                tolower(method), complement)
-      data_info <- readRDS(sprintf("data_simul/byseriespente_daf/%s.RDS",
-                                   gsub(".RDS", "",basename(s))))
-      
-      if(all(file.exists(nom_f_s_tp),
-             file.exists(nom_f_s_rev_fe),
[…]

### 2024-04-21 R_simul/3A_summary_of_tp.R

@@ -125,13 +125,15 @@ for(dir in c("localic_daf_trunc", "localic_final")){
   split <- strsplit(full_names, "_")
   series <- sapply(split, `[`, 1)
   method <- sapply(split, `[`, 2)
-  if(length(grep("localic_daf", dir)) >0){
-    h <- "h6"
-    degree <- sapply(split, `[`, 3)
-  } else {
-    h <- sapply(split, `[`, 3)
-    degree <- sapply(split, `[`, 4)
-  }
+  h <- sapply(split, `[`, 3)
+  degree <- sapply(split, `[`, 4)
+  # if(length(grep("localic_daf", dir)) >0){
+  #   h <- "h6"
+  #   degree <- sapply(split, `[`, 3)
+  # } else {
+  #   h <- sapply(split, `[`, 3)
+  #   degree <- sapply(split, `[`, 4)
+  # }
   degree[is.na(degree)] <- "d3"
   all_t <- data.frame(t(sapply(all_tp_lp, function(x) x[["troughs"]][["phaseshift"]])),
                       series, kernel = "henderson", method, h, degree)
@@ -144,33 +146,39 @@ for(dir in c("localic_daf_trunc", "localic_final")){
   
   
   all_files <- list.files(sprintf("results_simul/%s_nn/", dir),pattern = "_tp",full.names = TRUE)
-  all_tp_lp <- lapply(seq_along(all_files), function(i){
-    print(i)
-    f = all_files[i]
-    compute_time_lag(readRDS(f),
-                     peaks = tp$downturn,
-                     troughs = tp$upturn)
-  })
-  
-  full_names <- gsub("_tp.RDS$", "", basename(all_files))
-  split <- strsplit(full_names, "_")
-  series <- sapply(split, `[`, 1)
-  method <- sapply(split, `[`, 2)
-  if(length(grep("localic_daf", dir)) >0){
-    h <- "h6"
-    degree <- sapply(split, `[`, 3)
-  } else {
+  if (length(all_files) > 0){
+    
+    all_tp_lp <- lapply(seq_along(all_files), function(i){
+      print(i)
+      f = all_files[i]
+      compute_time_lag(readRDS(f),
+                       peaks = tp$downturn,
+                       troughs = tp$upturn)
+    })
+    
+    full_names <- gsub("_tp.RDS$", "", basename(all_files))
+    split <- strsplit(full_names, "_")
+    series <- sapply(split, `[`, 1)
[…]

### 2024-04-21 R_simul/3B_summary_of_tp_norev.R

@@ -126,13 +126,15 @@ for(dir in c("localic_daf_trunc", "localic_final")){
   split <- strsplit(full_names, "_")
   series <- sapply(split, `[`, 1)
   method <- sapply(split, `[`, 2)
-  if(length(grep("localic_daf", dir)) >0){
-    h <- "h6"
-    degree <- sapply(split, `[`, 3)
-  } else {
-    h <- sapply(split, `[`, 3)
-    degree <- sapply(split, `[`, 4)
-  }
+  h <- sapply(split, `[`, 3)
+  degree <- sapply(split, `[`, 4)
+  # if(length(grep("localic_daf", dir)) >0){
+  #   h <- "h6"
+  #   degree <- sapply(split, `[`, 3)
+  # } else {
+  #   h <- sapply(split, `[`, 3)
+  #   degree <- sapply(split, `[`, 4)
+  # }
   degree[is.na(degree)] <- "d3"
   
   all_t <- data.frame(t(sapply(all_tp_lp, function(x) x[["troughs"]][["phaseshift"]])),
@@ -146,6 +148,7 @@ for(dir in c("localic_daf_trunc", "localic_final")){
   
   
   all_files <- list.files(sprintf("results_simul/%s_nn/", dir),pattern = "_tp",full.names = TRUE)
+  if (length(all_files) > 0){
   all_tp_lp <- lapply(seq_along(all_files), function(i){
     print(i)
     f = all_files[i]
@@ -159,13 +162,15 @@ for(dir in c("localic_daf_trunc", "localic_final")){
   split <- strsplit(full_names, "_")
   series <- sapply(split, `[`, 1)
   method <- sapply(split, `[`, 2)
-  if(length(grep("localic_daf", dir)) >0){
-    h <- "h6"
-    degree <- sapply(split, `[`, 3)
-  } else {
-    h <- sapply(split, `[`, 3)
-    degree <- sapply(split, `[`, 4)
-  }
+  h <- sapply(split, `[`, 3)
+  degree <- sapply(split, `[`, 4)
+  # if(length(grep("localic_daf", dir)) >0){
+  #   h <- "h6"
+  #   degree <- sapply(split, `[`, 3)
+  # } else {
+  #   h <- sapply(split, `[`, 3)
+  #   degree <- sapply(split, `[`, 4)
+  # }
   degree[is.na(degree)] <- "d3"
   
[…]

### 2024-04-21 R_simul/3C_summary_of_rev.R

@@ -127,13 +127,16 @@ for(dir in c("localic_daf_trunc", "localic_final")){
       split <- strsplit(full_names, "_")
       series <- sapply(split, `[`, 1)
       method <- sapply(split, `[`, 2)
-      if(length(grep("localic_daf", dir)) >0){
-        h <- "h6"
-        degree <- sapply(split, `[`, 3)
-      } else {
-        h <- sapply(split, `[`, 3)
-        degree <- sapply(split, `[`, 4)
-      }
+      
+      h <- sapply(split, `[`, 3)
+      degree <- sapply(split, `[`, 4)
+      # if(length(grep("localic_daf", dir)) >0){
+      #   h <- "h6"
+      #   degree <- sapply(split, `[`, 3)
+      # } else {
+      #   h <- sapply(split, `[`, 3)
+      #   degree <- sapply(split, `[`, 4)
+      # }
       degree[is.na(degree)] <- "d3"
       
       data$series <- series
@@ -148,7 +151,7 @@ for(dir in c("localic_daf_trunc", "localic_final")){
     saveRDS(all_rev, sprintf("results_simul/compile_revisions/%s%s.RDS",dir, suffix))
     
     all_files <- list.files(sprintf("results_simul/%s_nn/", dir),pattern = suffix,full.names = TRUE)
-    
+    if (length(all_files) > 0){
     all_rev <- lapply(seq_along(all_files), function(i){
       print(i)
       f = all_files[i]
@@ -162,13 +165,15 @@ for(dir in c("localic_daf_trunc", "localic_final")){
       split <- strsplit(full_names, "_")
       series <- sapply(split, `[`, 1)
       method <- sapply(split, `[`, 2)
-      if(length(grep("localic_daf", dir)) >0){
-        h <- "h6"
-        degree <- sapply(split, `[`, 3)
-      } else {
-        h <- sapply(split, `[`, 3)
-        degree <- sapply(split, `[`, 4)
-      }
+      h <- sapply(split, `[`, 3)
+      degree <- sapply(split, `[`, 4)
+      # if(length(grep("localic_daf", dir)) >0){
+      #   h <- "h6"
+      #   degree <- sapply(split, `[`, 3)
+      # } else {
+      #   h <- sapply(split, `[`, 3)
+      #   degree <- sapply(split, `[`, 4)
+      # }
       degree[is.na(degree)] <- "d3"
       
       data$series <- series
[…]

### 2024-04-21 R_simul/4A_Analyse_tp.R

@@ -6,42 +6,34 @@ library(scales)
 library(latex2exp)
 
 tp_lp <- merge(readRDS("results_simul/compile_tp_norev/troughs_lp.RDS"),
-                   readRDS("results_simul/compile_tp_norev/peaks_lp.RDS"),
-                   by=c("series","kernel", "method")) %>%
+               readRDS("results_simul/compile_tp_norev/peaks_lp.RDS"),
+               by=c("series","kernel", "method")) %>%
   select_var()
 
 tp_lic_final <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic_final.RDS"),
-                          readRDS("results_simul/compile_tp_norev/peaks_localic_final.RDS"),
-                          by=c("series", "kernel", "h", "degree", "method"))  %>%
-  dplyr::filter(degree == "d2", h == "h6") %>%
-  select_var() %>% mutate(method = sprintf("%s_localic_final", method)) %>%
-  select(!c(degree, h))
-tp_lic_daf_trunc <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic_daf_trunc.RDS"),
-                              readRDS("results_simul/compile_tp_norev/peaks_localic_daf_trunc.RDS"),
-                              by=c("series", "kernel", "h", "degree", "method")) %>%
-  dplyr::filter(degree == "d2", h == "h6") %>%
-  select_var() %>% mutate(method = sprintf("%s_localic", method)) %>%
+                      readRDS("results_simul/compile_tp_norev/peaks_localic_final.RDS"),
+                      by=c("series", "kernel", "h", "degree", "method"))  %>%
+  # dplyr::filter() %>%
+  # mutate(method = ifelse(degree == "d2", h == "h6"))
+  select_var() %>% 
+  mutate(method = sprintf("%s_localic_final_%s_%s", method, degree, h)) %>%
   select(!c(degree, h))
 
-tp_lp_nn <- merge(readRDS("results_simul/compile_tp_norev/troughs_lp_nn.RDS"),
-                  readRDS("results_simul/compile_tp_norev/peaks_lp_nn.RDS"),
-                  by=c("series","kernel", "method")) %>%
+tp_lic_daf_trunc <- merge(
[…]

### 2024-04-21 R_simul/4B_analyse_revisions.R

@@ -15,9 +15,8 @@ for (crit in c("ce", "fe")) {
       data <- data %>% dplyr::filter(kernel == "henderson")
     if (method %in% c("localic_daf_trunc", "localic_final")){
       suff <- ifelse(method == "localic_final", "_final", "")
-      data <- data  %>% dplyr::filter(degree == "d2", h == "h6") %>%
-        mutate(method = sprintf("%s_localic%s", method,
-                                suff)) %>%
+      data <- data %>%
+        mutate(method = sprintf("%s_localic%s_%s_%s", method, suff, degree, h)) %>%
         select(!c(degree, h))
     }
     if (method == "arima") {
@@ -28,36 +27,122 @@ for (crit in c("ce", "fe")) {
     assign(sprintf("rev_%s_%s", crit, method), data)
   }
   all_data <- do.call(rbind, mget(sprintf("rev_%s_%s", crit, all_methods))) %>%
-    mutate(method = factor(method,c("lc", "lc_localic_final", "lc_localic",
-                                    "ql", "ql_localic_final", "ql_localic",
-                                    "cq","daf", "auto_arima"),
-                           ordered = TRUE),
-           variability = factor(variability,
+    mutate(variability = factor(variability,
                                 levels = c("lowvariability","mediumvariability","highvariability"),
                                 ordered = TRUE))
   assign(sprintf("all_rev_%s", crit), all_data)
 }
 
 x = all_rev_fe
-rev_tot = rbind(all_rev_fe %>% summarise_ref(),
-                all_rev_ce %>% summarise_ref())
-rev_rel = rbind(all_rev_fe %>% summarise_ref(normalise = TRUE),
-                all_rev_ce %>% summarise_ref(normalise = TRUE))
+rev_tot = rbind(all_rev_fe  %>% summarise_ref()|> mutate(crit = "1fe"),
+                all_rev_ce %>% summarise_ref()|> mutate(crit = "2ce"))
+rev_rel = rbind(all_rev_fe %>% summarise_ref(normalise = TRUE)|> mutate(crit = "1fe"),
[…]

### 2024-04-21 R_simul_quarter/1B_Est_pente_final.R

@@ -32,10 +32,10 @@ gen_MM <- function(p=2, q=p, d=2){
 
 MM = lapply(2:3, function(h){
   list(pente = list(
-    `d=2` = gen_MM(p = h, d=2)[,2]
+    `d=2` = gen_MM(p = 3, d=2)[,2]
     ),
     `deriv2` = list(
-      `d=2` = gen_MM(p = h, d=2)[,3]
+      `d=2` = gen_MM(p = 3, d=2)[,3]
       ),
     henderson = lp_filter(horizon = h)@sfilter
   )

### 2024-04-21 R_simul_quarter/4A_Analyse_tp.R

@@ -77,7 +77,7 @@ p <- ggplot(data_tp %>%
 p
 
 
-ggsave("paper/img/simulations/phase_shift_simul.pdf",
+ggsave("paper/img/simulations/phase_shift_simul_q.pdf",
             plot = p,
             width = 10, height = 6)
 

### 2024-04-23 paper/article_jos.qmd
update arima + paper

@@
   eq-prefix: equation
+
 linestretch: 2
@@
 
-# Some properties of moving averages {#sec-propMM}
+# Local trend-cycle models and moving averages {#sec-propMM}
+
+The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $g_t$ and a an erratic component $\varepsilon_t$ (called irregular):
+$$
+f(y_t)=g_t+\varepsilon_t
+$$
+where $f$ denotes some appropriate transformation (usually logarithmic or no transformation).
+To simplify further notations, $y_t$ will denote the transformed observed series.
+The noise component $\varepsilon_t$ is usually assumed to be a white noise.
+Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $g_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
+This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
+$$
+g_t = \sum_{j=0}^d\beta_jt^j+\xi_t
+$$
+with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
+Event if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
+Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
+The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
+The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
 
@@
 
-Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $X_t$:
-$$
-M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
+Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $y_t$:
+$$
+M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
 $$
@@
 
-To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $X_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
-The moving average $M_{\boldsymbol\theta}$ transforms $X_t$ into:
-$$
-Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
-= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
+To interpret the notions of gain and phase shift, it is useful to illustrate the effects of moving averages on harmonic series $y_t=\e^{-i\omega t}$ with $\omega\in[0,\pi]$.
+The moving average $M_{\boldsymbol\theta}$ transforms $y_t$ into:
+$$
+Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
+= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
 $$
@@
 
-To sum up, applying a moving average to a harmonic times series ($X_t=\e^{-i\omega t}$) affects it in in two different ways:
+To sum up, applying a moving average to a harmonic times series ($y_t=\e^{-i\omega t}$) affects it in in two different ways:
 
@@
 
+::: {#fig-graphsmusgrave}
 ```{r, echo=FALSE, out.width="100%"}
-#| label: fig-graphsmusgrave
-#| fig-cap: "Coefficients, gain and phase shift function for the Musgrave filter in real time with $I/C=3.5$."
 img <- sprintf("img/filters_used/%s.%s", tolower("musgrave"), "pdf")
@@
 
+Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg). 
+In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.
+
+Coefficients, gain and phase shift function for the Musgrave filter used in X-11 in real time (last estimate) for monthly series when the bandwidth is 6.
+:::
+
+## Trend-cycle estimates and moving average
+
+Les méthodes présentées dans cette étude permettent de lisser les séries afin d'extraire la composante tendance-cycle pour faciliter l'analyse des retournements conjoncturels dans le cycle classique (également appelé cycle des affaires).
+On ne cherche donc pas à estimer séparément la tendance (qui représente les évolutions de long terme) et le cycle (qui représente les évolutions cycliques autour de la tendance) puisque ces composantes peuvent être difficiles à séparer (quelle définition prendre pour distinguer les cycles courts et des cycles longs ?).
+Par ailleurs, la décomposition entre tendance et cycle, associée à des méthodes de filtrage de type Hodrick-Prescott ou Baxter-King, sont plutôt utilisées pour analyser les points de retournement dans le cycle de croissance (voir @ferrara2009caracterisationcycles pour une description des différents cycles économiques).
+Ces méthodes de décomposition tendance et cycle nécessitent d'avoir des séries longues (afin de distinguer cycles courts et tendance de long terme) alors que les méthodes utilisées dans cet article utilisent peu d'observations (en général 13 pour des séries mensuelles).
+
 ## Desirable properties of a moving average
@@
 
-In our case, we assume that our initial series $y_t$ is seasonally adjusted and can be written as the sum of a trend-cycle, $TC_t$, and an irregular component, $I_t$:
-$$
-y_t=TC_t+I_t.
-$$
-The aim will be to build moving averages that best preserve the trend-cycle ($M_{\boldsymbol\theta} (TC_t)\simeq TC_t$) and minimise noise ($M_{\boldsymbol\theta} (I_t)\simeq 0$).
-
-### Trend preservation
-
-Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg).
-In order to best preserve the trend-cycles, we try to design moving averages that preserve the polynomial trends.
+In this paper, we assume that our initial series $y_t$ is seasonally adjusted and can be written as the sum of a trend-cycle, $TC_t$, and an irregular component, $\varepsilon_t$:
+$$
+y_t=TC_t+\varepsilon_t.
+$$
+The aim will be to build moving averages that best preserve the trend-cycle ($M_{\boldsymbol\theta} (TC_t)\simeq TC_t$).
+
+Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg) and, in order to best preserve the trend-cycles, we thus design moving averages that preserve the polynomial trends.
 A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t:M_{\boldsymbol\theta} f(t)=f(t)$.
@@
 
-### Noise reduction
-
-All time series are affected by noise, which can blur the extraction of trends and cycles.
-This is why we try to reduce this noise (by reducing its variance) while preserving the relevant trends.
-The sum of the squares of the coefficients $\sum_{k=-p}^{+f}\theta_k^2$ is the *reduction of variance* ratio.
-
-Indeed, let $\{\varepsilon_t\}$ be a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$:
-$$
-\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
-= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
-\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
-$$
+<!-- ### Noise reduction -->
+
+<!-- All time series are affected by noise, which can blur the extraction of trends and cycles. -->
+<!-- This is why we try to reduce this noise (by reducing its variance) while preserving the relevant trends. -->
+<!-- The sum of the squares of the coefficients $\sum_{k=-p}^{+f}\theta_k^2$ is the *reduction of variance* ratio. -->
+
+<!-- Indeed, let $\{\varepsilon_t\}$ be a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$: -->
+<!-- $$ -->
+<!-- \V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}} -->
+<!-- = \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}= -->
+<!-- \sigma^2\sum_{k=-p}^{+f} \theta_k^2. -->
+<!-- $$ -->
 
@@
 This is implemented in the function `rjd3filters::var_estimator()`.
-<!-- The assumption that $\hat\mu(x)$ is unbiased is rarely exactly true, so variance -->
-<!-- estimates and confidence intervals are usually computed at small band- -->
-<!-- widths where bias is small.  -->
+This estimator assumes that $\widehat{TC}_t$ is unbiased, so variance estimates intervals are usually computed at small bandwidths ($h$).
+In this paper we study monthly series with small bandwidth ($h=6$), and the variance is estimated using local polynomial of degree 3 (Henderson filter), the biased of $\widehat{TC}_t$ is thus already small and using smaller bandwidths have almost no impact on the estimates (see Supplemental materials).
 
@@
 Furthermore, for the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
-Here, we have chosen to model a trend of degree 2: this reduces the phase shift but slightly increases the revisions linked to the first estimate of the trend-cycle. 
+Here, we have chosen to model a trend of degree 2: this slightly reduces the phase shift (see Supplemental material) but slightly increases the revisions linked to the first estimate of the trend-cycle.
 @fig-mmpenteconcac shows the moving averages used.
@@
 
-In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability (Table\ \ref{tab-simulrev}).
+In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability (@tbl-simulrev).
 In general, LC filters always minimise revisions (with relatively small effect from the local parametrisation of the filters) and revisions are greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
@@
 
-```{r, echo = FALSE}
-#| label: tab-simulrev
-#| tbl-cap: "Average of the relative deviations of the revisions for the different filters on simulated series with medium variability."
-#| eval: false
+::: {#tbl-simulrev}
+```{r}
+#| echo: false
 rev_table <- readRDS("data/simulations_revisions.RDS")
@@
 \\\\right]$" = nrow(rev_table) / 2)
-p <- rev_table  %>%
+rev_table  %>%
   kable(format.args = list(digits = 2,
@@
         align = "c", booktabs = T, row.names = FALSE, 
-        label = "tab-simulrev",
-        escape = FALSE,
-        format = "latex") %>%  
+        escape = FALSE,) %>%  
   kable_styling(latex_options=c(#"striped",  
@@
 ```
-
-```{=tex}
-\begin{table}[!h]
-\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
-filters on simulated series with medium variability.}
-\centering
-\begin{tabular}{ccccccc}
-\toprule
-Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
-\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
-\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
-\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
-\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
-\right]$}}\\
-\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
-\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
-\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
-\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
-\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
-\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
-\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
-\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
-\bottomrule
-\end{tabular}
-\end{table}
-```
+Lecture: the estimates for $q=0$ correspond to the first estimates of the trend-cycle (real-time estimates when 0 point in the future is known), $q=1$ correspond to the second estimates of the trend-cycle (when 1 point in the future is known).
+Since the final estimates are obtain using a symmetric filter of bandwidth 6, $q=6$ corresponds to the final estimates of the trend-cycle (when 6 points in the future are known).
+$MAE_{fe}(0)$ compares the revision between the first estimate ($q=0$) and the last estimate ($q=6$) and $MAE_{ce}(0)$ compares the revision between the first estimate ($q=0$) and the second estimate ($q=1$).
+
+Average of the relative deviations of the revisions for the different filters on simulated series with medium variability.
+:::
+
+<!-- ```{=tex} -->
+<!-- \begin{table}[!h] -->
+<!-- \caption{\label{tbl-simulrev} Average of the relative deviations of the revisions for the different -->
+<!-- filters on simulated series with medium variability.} -->
+<!-- \centering -->
+<!-- \begin{tabular}{ccccccc} -->
+<!-- \toprule -->
+<!-- Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\ -->
+<!-- \midrule -->
+<!-- \addlinespace[0.3em] -->
+<!-- \multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\ -->
+<!-- \hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
+<!-- \hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
+<!-- \hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
+<!-- \hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\ -->
+<!-- \hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
+<!-- \hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\ -->
+<!-- \hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\ -->
+<!-- \hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\ -->
+<!-- \hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
+<!-- \addlinespace[0.3em] -->
+<!-- \multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[ -->
+<!-- \left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right| -->
+<!-- \right]$}}\\ -->
+<!-- \hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\ -->
+<!-- \hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\ -->
+<!-- \hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\ -->
+<!-- \hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\ -->
+<!-- \hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\ -->
+<!-- \hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\ -->
+<!-- \hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\ -->
+<!-- \hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\ -->
+<!-- \hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\ -->
+<!-- \bottomrule -->
+<!-- \end{tabular} -->
+<!-- \end{table} -->
+<!-- ``` -->
 
@@
 Studying the series up to January 2020, this series has a medium variability: a symmetric filter of 13 terms is therefore appropriate.
-Appendix [-@sec-ann-ex-r] shows how the code to reproduce the different plots.
 

### 2024-04-23 paper/resp_referee.qmd
update arima + paper

@@
 
-> 3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function 
-> at $\omega = 0$ yield the conditions. So the results could be stated with references, and the 
-> derivations omitted, and this subsection compactified.
+> 3. Section 1.2.1: these results are well-known; the derivatives of the frequency response function at $\omega = 0$ yield the conditions. 
+So the results could be stated with references, and the derivations omitted, and this subsection compactified.
 
@@
 
-The $\forall q$ was omitted and this section was rewritten to make it clearer.
+*The $\forall q$ was omitted and this section was rewritten to make it clearer.*
 
@@
 
+*Fixed*
+
 > 9. Page 15, line 40: I couldn't find the matrix $T$ defined. How is it connected to timeliness? 
-> That is, when you display the entries of the matrix $T$, you need to make a connection between 
-> those values and the concept of timeliness.
+That is, when you display the entries of the matrix $T$, you need to make a connection between those values and the concept of timeliness.
 
@@
 
-> 12. Appendix B: rather than typing out the R code, it would be better to have the files available 
-> as supplementary material. (Or possibly put in a github repo and provide the links in the 
-> text.)
+> 12. Appendix B: rather than typing out the R code, it would be better to have the files available as supplementary material. 
+(Or possibly put in a github repo and provide the links in the text.)
 
@@
 > 1. The authors should provide a clear definition of trend-cycle, $TC_t$, and irregular, $I_t$. In other words, how is the decomposition $y_t =TC_t+I_t$ identified?
-> In section 1 they seem to hint at the interpretation of $TC_t$ as the low-pass component of yt with cutoff frequency $\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component). 
+> In section 1 they seem to hint at the interpretation of $TC_t$ as the low-pass component of $y_t$ with cutoff frequency $\pi/6$ (and correspondingly they interpret $I_t$ as the high-pass component).
 > This point should be clarified, I think. 
@@
 
+The modelled of the input time series is now describe more precisely in section 2 (previous section 1). $TC_t$ is here considered as deterministic
+
 > 2. The idea of minimizing the noise component ($M_\theta(I_t)\simeq 0$) is not very formal. 
@@
 > Or the long run variance?
+
+*Since section 1.3.2 was removed (not central to the paper and simplify the reading), the section 1.2.2 and the reference to the minimization of the noise component (through the minimisation of the variance is it is modelled as a white noise) is removed*
 
@@
 
-*It was added in the beginning of section 3 and 4 than the bandwidth is fixed. Nearest neighbour bandwidth was added in supplemental material.*
+*It was added in the beginning of new sections 3 and 4 than the bandwidth is fixed.
+Nearest neighbour bandwidth was added in supplemental material but it doesn't improve the results in therm of phase shift.
+I can be explain that the bi*
 
 > 6. Is the phase shift related to the notion of bias in local polynomial smoothing?
+
+In local polynomial smoothing, the notion of bias is associated to the approximation error of the smoothed series by a polynomial degree of order $d$.
+It can thus be writer as
 
@@
 The description of the three panels should be better explained in the main text.
+
+*A lecture note is added*
 
@@
 Another thought is that the ARIMA model being used to extend the series appears to be modeled using the full series; what if the ARIMA model used to extend the series was instead estimated using just a short span toward the end of the series?
+*Indeed it is in general not relevant to identify an ARIMA model on a monthly time series of length 60 years.*
+In our case, since the same data generation process remains the same, it could be relevant and leads to better results in terms of delay to detect turning points (in supplement materials, a plot is added comparing the results with different spans to identify/estimate the ARIMA model).
+However, on the main text, we now only used 12 years to identify the ARIMA model, to be closer to what would be done on real series.
 
@@
 In addition, in the text, it would probably be preferable to use R (the letter itself) instead of the R icon.
+
+*Bibliography added and the R icon was removed*
 
@@
 > 1. Equivalence between Asymmetric and Symmetric Filters (Page 8, Line 9)
-> I suggest that the authors provide a more explicit explanation on page 8, line 9, regarding
-the equivalence between asymmetric and symmetric filters. It would be beneficial to
-clarify the normalization difference and ensure a clear understanding of the relationship
-between the two types of filters.
+> I suggest that the authors provide a more explicit explanation on page 8, line 9, regarding the equivalence between asymmetric and symmetric filters. 
+It would be beneficial to clarify the normalization difference and ensure a clear understanding of the relationship between the two types of filters.
 
@@
 > 2.  Symbol [[ ]] Meaning (Page 11, Lines 53-54)
-> The meaning of the symbol [[ ]] on page 11, lines 53-54, is not clear. I recommend that
-the authors explicitly define this symbol or provide a clear explanation to avoid any
-ambiguity in its interpretation.
+> The meaning of the symbol [[ ]] on page 11, lines 53-54, is not clear. 
+I recommend that the authors explicitly define this symbol or provide a clear explanation to avoid any ambiguity in its interpretation.
 
@@
 > 3.  Variance Determination in Local Parameterization (Page 17)
-> The determination of variance in the local parameterization of asymmetric filters (page
-17) could benefit from a more detailed explanation. I recommend that the authors
-clearly identify the quantities appearing in the denominator for better understanding.
+> The determination of variance in the local parameterization of asymmetric filters (page 17) could benefit from a more detailed explanation. 
+I recommend that the authors clearly identify the quantities appearing in the denominator for better understanding.
 
@@
 > 4.  Applicability to High-Frequency Data.
-> It would be valuable if the authors could explicitly address the applicability of the
-proposed procedure to high-frequency data. If modifications or considerations are
-necessary for such data, it should be discussed to provide a comprehensive
-understanding of the methodology.
+> It would be valuable if the authors could explicitly address the applicability of the proposed procedure to high-frequency data. 
+If modifications or considerations are necessary for such data, it should be discussed to provide a comprehensive understanding of the methodology.
 
 > 5.  Clarity on Phase Shift Computation (Page 20)
-> A clearer explanation of how the phase shift is computed on page 20 would enhance the
-paper's overall clarity. The authors should specify the method used for determining the
-phase shift, particularly in relation to the identification of true turning points based on
-symmetric filter estimates.
+> A clearer explanation of how the phase shift is computed on page 20 would enhance the paper's overall clarity. 
+The authors should specify the method used for determining the phase shift, particularly in relation to the identification of true turning points based on symmetric filter estimates.
 
@@
 > I suggest that the authors include a detailed commentary on Figure 5 in the paper.
-Providing explanations for the results shown in the figure would be crucial for readers to
-better understand the performance of the filters/methods.
+Providing explanations for the results shown in the figure would be crucial for readers to better understand the performance of the filters/methods.
 
 > 7.  Better Description of Results in Table 3
-> The meaning of "second and third estimates" in Table 3 on page 22 is not clear. I
-recommend that the authors provide a more detailed description or explanation of the
-results to improve comprehension.
+> The meaning of "second and third estimates" in Table 3 on page 22 is not clear. 
+I recommend that the authors provide a more detailed description or explanation of the results to improve comprehension.
+
+*A lecture note is added in the tables and more explainations are added to the text*
 
 > 8.  Consideration of Different Series Periodicity
-> It would be beneficial if the authors could address the suggestion of considering series
-with different periodicities in the real data application. Discussing how the proposed
-methodology performs when applied to series with various frequencies would add
-depth to the study.
+> It would be beneficial if the authors could address the suggestion of considering series with different periodicities in the real data application.
+Discussing how the proposed methodology performs when applied to series with various frequencies would add depth to the study.
 
@@
 
-1) The paper lacks in describing the contribution from the beginning. 
+> 1) The paper lacks in describing the contribution from the beginning. 
 I had to read the paper a couple of times to clearly identify the contribution.
@@
 
-2) The paper omits to consider the Covid19 period and how this influences the trend and cycle decomposition. 
+> 2) The paper omits to consider the Covid19 period and how this influences the trend and cycle decomposition. 
 I understand that this is not the main focus of the paper, but at least the authors should address this problem in word and possible solution to address it, e.g.

### 2024-04-23 paper/supp_material.qmd
update arima + paper

@@
 
+\newpage
+
+#
+
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_final_lc.pdf")
+```
+
+Distribution of phase shifts associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$ in function of the length of the henderson filter used to estimate $\sigma$ ($h$) and the degree of the trend to estimate $\delta$ ($d$)/
+:::
+
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_lc.pdf")
+```
+
+Distribution of phase shifts associated to the local parametrisation of the linear-constant (LC) filter using real time estimates of $\delta/\sigma$ in function of the length of the henderson filter used to estimate $\sigma$ ($h$) and the degree of the trend to estimate $\delta$ ($d$)/
+:::
+
+::: {#fig-graphs-impact-span}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_ql.pdf")
+```
+
+Distribution of phase shifts associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$ in function of the length of the henderson filter used to estimate $\sigma$ ($h$) and the degree of the trend to estimate $\delta$ ($d$)/
+:::
+
+\newpage
 # Quarterly data
@@
 
+\newpage
 # Nearest neighbour bandwidth
@@
 ```
+Note:
 

### 2024-04-23 R_filters/slope_concavity_estimator.R

@@ -69,6 +69,7 @@ p21 = plot_coef_ggplot(MM_d2_pente) +
 p22 = plot_coef_ggplot(MM_d2_deriv2)
 p = (p21 + ggtitle("Slope estimators")) +
   (p22 + ggtitle("Concavity estimators"))
-p
+p <- p & scale_color_grey() & labs (y = NULL)
 ggsave("paper/img/filters_used/mm_penteconcavite.pdf",
-       width = 7,height = 3)
+       width = 7*1.2,height = 3)
+

### 2024-04-23 R_fredm/2A_Est_final.R

@@ -31,6 +31,10 @@ fs <- list()
 j <- 1
 h_retain <- 3:6
 kernel = "Henderson"
+d <- 2
+s = list.files("data_fredm/byseries", full.names = TRUE)[1]
+h <- 6
+method <- "LC"
 for (method in c("LC","QL")){
   print(method)
   for(s in list.files("data_fredm/byseries", full.names = TRUE)){

### 2024-04-23 R_fredm/2D_est_arima.R

@@ -13,6 +13,7 @@ list_series <- list.files("data_fredm/byseries", full.names = TRUE)
 s = list_series[1]
 fs <- list()
 i <- 0
+nyears <- 12
 for(s in list_series){
   name_file <- gsub(".RDS$", "", basename(s))
   print(name_file)
@@ -37,12 +38,14 @@ for(s in list_series){
     
     series_s <- lapply(names(data), function(nom_d){
       x <- data[[nom_d]]
+      first_date <- 1 + max(0, length(x) - nyears * frequency(x))
+      x_arima <- window(x, start = time(x)[first_date])
       l = data_info[[nom_d]]["optimal_length"]
       l = 13
-      prevs = auto.arima(x, max.Q = 0, max.D = 0, max.P = 0) %>% 
+      prevs = auto.arima(x_arima, max.Q = 0, max.D = 0, max.P = 0) %>% 
         forecast(h=(l-1)/2)
       y_prevs = ts(c(x, prevs$mean), start = start(x), frequency = frequency(x))
-      window(henderson(y_prevs, musgrave = FALSE,length = l),
+      window(rjd3x11plus::henderson(y_prevs, musgrave = FALSE,length = l),
              start = start(x), end = end(x))
     })
     names(series_s) <- names(data)

### 2024-04-23 R_fredm/4B_Analyse_implicit_forecast.R

@@ -57,10 +57,11 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              series = series,
                              all_tp = all_tp,
                              all_tp_rev = all_tp_rev,
-                             tp_keep = tp_keep,
-                             dossier = "results_fredm/compile_tp_norev/")
+                             tp_keep = tp_keep)
 wrap_plots(plots, ncol = 3)
 
 ggsave("paper/img/nber/retailx_nov2007_prev_imp_lp.pdf",
-       plot = wrap_plots(plots, ncol = 3),
+       plot = wrap_plots(plots, ncol = 3)& 
+         scale_color_grey() & 
+         labs(subtitle = NULL),
        width = 8, height = 8)

### 2024-04-23 R_fredm/4_utils.R

@@ -171,7 +171,7 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
 
 
 get_all_prevs <- function(series, tp_keep, nb_est = 10, nb_dates_before = 6,
-                          d_localic = 2, h_localic = 6){
+                          d_localic = 2, h_localic = 6, nyears_arima = 12){
   s <- sprintf("data_fredm/byseries/%s.RDS", series)
   tp_date <- as.numeric(tp_keep)
   data <- readRDS(s)
@@ -184,7 +184,9 @@ get_all_prevs <- function(series, tp_keep, nb_est = 10, nb_dates_before = 6,
   colnames(data_merge) <- as.character(zoo::as.yearmon(as.numeric(names(data))))
 
   arima_prevs <- do.call(ts.union, lapply(data, function(y){
-    prevs = forecast::auto.arima(y, max.Q = 0, max.D = 0, max.P = 0) %>%
+    first_date <- 1 + max(0, length(y) - nyears * frequency(y))
+    x_arima <- window(y, start = time(y)[first_date])
+    prevs = forecast::auto.arima(x_arima, max.Q = 0, max.D = 0, max.P = 0) %>%
       forecast::forecast(6)
     ts(c(tail(y,1), prevs$mean), start = tail(time(y),1),
        frequency = frequency(y))

### 2024-04-23 R_simul/4A_Analyse_tp.R

@@ -28,6 +28,11 @@ tp_lic_daf_trunc <- merge(
   select(!c(degree, h))
 
 
+tp_lp_nn <- merge(readRDS("results_simul/compile_tp_norev/troughs_lp_nn.RDS"),
+                  readRDS("results_simul/compile_tp_norev/peaks_lp_nn.RDS"),
+                  by=c("series","kernel", "method")) %>%
+  select_var() %>% 
+  mutate(method = sprintf("%s_nn", method))
 tp_lic_final_nn <- merge(readRDS("results_simul/compile_tp_norev/troughs_localic_final_nn.RDS"),
                          readRDS("results_simul/compile_tp_norev/peaks_localic_final_nn.RDS"),
                          by=c("series", "kernel", "h", "degree", "method"))  %>%
@@ -97,7 +102,7 @@ legende <- c(lc = "LC", ql = "QL",
              lc_localic_d2_h6 = "LC loc.\nparam.",
              ql_localic_final_d2_h6 = "QL loc. param.\n(final estimates)",
              ql_localic_d2_h6 = "QL loc.\nparam.",
-             auto_arima = "ARIMA")
+             auto_arima_ny12 = "ARIMA")
 p <- ggplot(all_tp %>% 
               dplyr::filter(method %in%
                               names(legende))  %>% 

### 2024-04-24 paper/article_jos.qmd
update plots + articles

@@
 This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
-In particular, this theory allows us to reproduce Henderson's symmetric and Musgrave's asymmetric filters.
-This paper describes how they can be extended to include a timeliness criterion to minimise the phase shift.
-While asymmetric filters are generally parametrised globally, which can be suboptimal around turning points, this paper proposes a procedure for parametrising them locally.
+In particular, this theory allows us to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
+This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates (for the last estimates): first including a timeliness criterion to minimise the phase shift;
+second, with procedure for parametrising asymmetric filters locally While they are generally parametrised globally, which can be suboptimal around turning points.
 An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
@@
 In particular, it becomes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
-However, these methods have two drawbacks.
-Firstly, the phase shift (i.e. the delay in detecting turning points) is not directly controlled.
-Secondly, asymmetric filters depend on a parameter that is generally estimated over the whole series using assumptions that may be wrong locally.
-The aim of this study is to propose extensions to the @proietti2008 class of filters in order to take these shortcomings into account.
+The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to two drawbacks of those methods.
+Firstly,  adding, in the optimisation process, a timeliness criterin to directly control the phase shift (i.e. the delay in detecting turning points).
+Secondly, calibrating asymmetric filter locally, while they are generally calibrated globally  using thus assumptions that may be wrong locally.
 Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and can therefore be easily used.
@@
 
-The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $g_t$ and a an erratic component $\varepsilon_t$ (called irregular):
-$$
-f(y_t)=g_t+\varepsilon_t
+The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and a an erratic component $\varepsilon_t$ (called irregular):
+$$
+f(y_t)=\mu_t+\varepsilon_t
 $$
@@
 The noise component $\varepsilon_t$ is usually assumed to be a white noise.
-Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $g_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
+Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
 This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
 $$
-g_t = \sum_{j=0}^d\beta_jt^j+\xi_t
+\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
 $$
@@
 The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
+The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e.
+when no future point is known), using moving averages build on local polynomial smoothing techniques.
 
@@
 Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
-The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component (undesirable oscillations).
-
-@fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e.
+The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component.
+
+@Fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e.
 when no point in the future is known).
@@
 
-## Trend-cycle estimates and moving average
-
-Les méthodes présentées dans cette étude permettent de lisser les séries afin d'extraire la composante tendance-cycle pour faciliter l'analyse des retournements conjoncturels dans le cycle classique (également appelé cycle des affaires).
-On ne cherche donc pas à estimer séparément la tendance (qui représente les évolutions de long terme) et le cycle (qui représente les évolutions cycliques autour de la tendance) puisque ces composantes peuvent être difficiles à séparer (quelle définition prendre pour distinguer les cycles courts et des cycles longs ?).
-Par ailleurs, la décomposition entre tendance et cycle, associée à des méthodes de filtrage de type Hodrick-Prescott ou Baxter-King, sont plutôt utilisées pour analyser les points de retournement dans le cycle de croissance (voir @ferrara2009caracterisationcycles pour une description des différents cycles économiques).
-Ces méthodes de décomposition tendance et cycle nécessitent d'avoir des séries longues (afin de distinguer cycles courts et tendance de long terme) alors que les méthodes utilisées dans cet article utilisent peu d'observations (en général 13 pour des séries mensuelles).
-
 ## Desirable properties of a moving average
@@
 
-In this paper, we assume that our initial series $y_t$ is seasonally adjusted and can be written as the sum of a trend-cycle, $TC_t$, and an irregular component, $\varepsilon_t$:
+In this paper, we assume that our initial series $y_t$ is seasonally adjusted and can be written as the sum of a trend-cycle, $TC_t$, and an irregular (noise) component, $\varepsilon_t$:
 $$
@@
 
-### Asymmetric moving averages and forecasts {#sec-mmetprev}
-
 For symmetric filters, the phase shift function is equal to zero (modulo $\pi$).
@@
 This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
-A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters
+A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
+Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).
 
@@
 However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.  
-Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies,
-this method can be used in R with the function `rjd3filters::rkhs_filter()`.
-
-## Symmetric moving averages and local polynomial regression
+Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
+See @inseeDTM202401 for a more detailed comparison of the different recent methods for trend-cycle extraction included in the `rjd3filters` package
+
+In this paper we will focus on @proietti2008 approach. 
+@Sec-sym-lp and @sec-lppasymf describes their approach to build symmetric moving averages (used for the final estimates of the trend-cycle component) and asymmetric moving averages (used for real-time estimates).
+This paper proposed two extensions of their approach to build asymetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parameterisation of a parameter usually parameterised globally (@sec-lptimeliness), which is used in the simulation.
+
+## Symmetric moving averages and local polynomial regression {#sec-sym-lp}
 
@@
 
-As mentioned in @sec-mmetprev, for real-time estimation, several approaches can be used:
+As mentioned in @sec-mmasym, for real-time estimation, several approaches can be used:
 
@@
 where $[\boldsymbol U,\boldsymbol Z]$ is of full rank and forms a subset of the columns of $X.$
-Defining the variance error as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ (to directly take into account the weights associated to the observations) and with $[\boldsymbol U,\boldsymbol Z] = \boldsymbol X$ we this find the model associated to symmetric filters.
+Defining the variance error as $\boldsymbol D = \sigma^2\boldsymbol K^{-1}$ (to directly take into account the weights associated to the observations) and with $[\boldsymbol U,\boldsymbol Z] = \boldsymbol X$ we thus find the model associated to symmetric filters.
 The objective is to find an asymmetric filter $\boldsymbol v$ which uses $q$ points in the future and minimises the mean square error of revision (to the symmetric filter $\boldsymbol \theta$) under certain (polynomial) constraints.
@@
 
-When $\boldsymbol U=\boldsymbol X$ (and then $\boldsymbol Z = \boldsymbol 0$), the constraint is equivalent to preserving polynomials of degree $d$ and we find asymmetric direct filters (DAF) when $\boldsymbol D=\boldsymbol K^{-1}$.
+When $\boldsymbol U=\boldsymbol X$ (and then $\boldsymbol Z = \boldsymbol 0$), the constraint is equivalent to preserving polynomials of degree $d$ and we find asymmetric direct filters (DAF) when $\boldsymbol D=\sigma^2\boldsymbol K^{-1}$.
 
@@
 Furthermore, for the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
-Here, we have chosen to model a trend of degree 2: this slightly reduces the phase shift (see Supplemental material) but slightly increases the revisions linked to the first estimate of the trend-cycle.
+In this paper, we have chosen to model a trend of degree 2: this slightly reduces the phase shift (see Supplemental material) but slightly increases the revisions linked to the first estimate of the trend-cycle.
 @fig-mmpenteconcac shows the moving averages used.
@@
 
-In the empirical applications of @sec-comparison, the final local parametrisation is obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
+To distinguish the gain from the local parametrisation of the noise biased associated to the real-time estimates of the slope and concavity, in the empirical applications of @sec-comparison we also compares the results to the final local parametrisation obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
 @fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two local parametrisations:
@@
 
-The turning points are clearly detected by the local estimators, but there is much more variance in the estimate of $|\delta_1/\sigma|$ in the real-time estimate.
-This will lead to another source of revisions in the intermediate estimates.
+The turning points are clearly detected by the local estimators (changement in the slope), when $|\delta_1/\sigma|$ is estimated in the real-time estimate.
+However, the real-time local estimates of $|\delta_1/\sigma|$ are noisy: the varability of this estimates, used to build asymmetric filters, will then lead to another source of revisions in the intermediate estimates.
 
@@
 
+There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series rjd3filters`.
+
 # Comparison of different methods {#sec-comparison}
@@
 All the methods to create asymmetric filters are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
-The ARIMA model is determined automatically, using @autoarima algorithm (function `forecast::auto.arima()`), using no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
+The ARIMA model is determined automatically, using @autoarima algorithm (function `forecast::auto.arima()`) on the last 12 years, with no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
 
@@
 - Low signal-to-noise ratio (i.e. high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).
+
+The series associa
 
@@
 
-1. Computation of the phase shift in the detection of turning points. The definition of @Zellner1991 is used to determine the turning points:  
+1. Computation of the phase shift in the detection of turning points. 
+In this paper, we focus on turning points associated to business cycle: it is defined as the succession of phases of economic recession and expansion, delimited by peaks (highest level of activity) and troughs (lowest level of activity), see for example @ferrara for a description of the different economic cycles.
+In this paper, following @Zellner1991, we used an accepted definition for turning-points:  
     - An upturn occurs when the economy moves from a phase of recession to a phase of expansion. 
-    This is the case at date  $t$ when $y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}$.  
+    This is the case at date  $t$ when $TC_{t-3}\geq TC_{t-2}\geq TC_{t-1}<TC_t\leq TC_{t+1}$.  
     - A downturn occurs when the economy moves from a phase of expansion to a phase of recession.
-This is the case at date  $t$ when $y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}$.  
-    It thus takes at least 2 months to detect a turning point.  
-    The phase shift is often defined as the number of months required to detect the right turning point.
-    Here we use a slightly modified criterion: the phase shift is defined as the number of months needed to detect the right turning point without any future revision.
-    Indeed, it can happen that the right turning point is detected by asymmetric filters but is not detected by the final estimate using a symmetric filter (this is the case for 41 reversal points out of the 9 series with asymmetric Musgrave filters) or that there are revisions in successive estimates (this is the case for 7 turning points out of the 9 series with asymmetric Musgrave filters).
+This is the case at date  $t$ when $TC_{t-3}\leq TC_{t-2}\leq TC_{t-1}>TC_t\geq TC_{t+1}$.
+
+  Lets denote $TC_{t|t'}$ the estimate of the trend-cycle at date $t$ using the data up to date $t'$.
+  The phase shift is often defined as the number of months required to detect the right turning point: it would then be equal to two if, in the case of a downturn,  $TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}$ (two observations, at dates $t$ and $t+1$, are needed to qualify the date $t$ as turning point).
+  Note that, in that case, the phase shift could be equivalently set to 1 (only one observation after $t$ is needed to define $t$ as a turning point), without any difference on the comparison of the different methods.\
+  In this paper we use a slightly modified criterion: the phase shift is defined as the number of months needed to detect the right turning point without any future revision.
+  Thus, in the case of a downturn, the phase shift is two if $TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}$ and, for all dates $t'$ higher than $t+1$: $TC_{t-3|t'}\leq TC_{t-2|t'}\leq TC_{t-1|t'}>TC_{t|t'}\geq TC_{t+1|t'}$.
+Since a 13-term Henderson filter is used for the final estimates of the trend-cycle, the phase shift is at most height (the final estimate of $TC_{t+1}$ is compute a the date $t+7$).
+  This definition is used because it can happen that the right turning point is detected by asymmetric filters but is not detected by the final estimate using a symmetric filter (this is the case for 41 reversal points out of the 9 series with asymmetric Musgrave filters) or that there are revisions in successive estimates (this is the case for 7 turning points out of the 9 series with asymmetric Musgrave filters).
   Finally, relatively few turning points are detected at the right date with the final estimate.
@@
 \left|\frac{
-y_{t|t+q} -  y_{t|last}
+TC_{t|t+q} -  TC_{t|last}
 }{
- y_{t|last}
+ TC_{t|last}
 }\right|
@@
 $$
+
 3. The average of the relative deviations between two consecutive estimates (the $q$^th^ estimate and the $q+1$^th^ estimate):
@@
 \left|\frac{
-y_{t|t+q} - y_{t|t+q+1}
+TC_{t|t+q} - TC_{t|t+q+1}
 }{
-y_{t|t+q+1}
+TC_{t|t+q+1}
 }\right|
@@
 
-Extending the series using an ARIMA model gives revisions with the latest estimates of the same order of magnitude as the LC filter, but slightly larger revisions between consecutive estimates, particularly between the fourth and fifth estimates (as might be expected, as highlighted in @sec-mmetprev).
+Extending the series using an ARIMA model gives revisions with the latest estimates of the same order of magnitude as the LC filter, but slightly larger revisions between consecutive estimates, particularly between the fourth and fifth estimates (as might be expected, as highlighted in @sec-mmasym).
 
@@
 rev_table <- readRDS("data/simulations_revisions.RDS")
-pack_row_index = c("$MAE_{fe}(q) = \\mathbb E\\\\left[\\\\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\\\\right|\\\\right]$" = nrow(rev_table) / 2,
+pack_row_index = c("$MAE_{fe}(q) = \\mathbb E\\\\left[\\\\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\\\\right|\\\\right]$" = nrow(rev_table) / 2,
                    "$MAE_{ce}(q)=\\mathbb E\\\\left[
-\\\\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\\\\right|
+\\\\left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\\\\right|
 \\\\right]$" = nrow(rev_table) / 2)
@@
 <!-- \addlinespace[0.3em] -->
-<!-- \multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\ -->
+<!-- \multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\right|\right]$}}\\ -->
 <!-- \hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
@@
 <!-- \multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[ -->
-<!-- \left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right| -->
+<!-- \left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\right| -->
 <!-- \right]$}}\\ -->
@@
 
-Graph + simulations
-
-Enfin, dans cet article et dans ceux associés aux méthodes étudiées, les filtres asymétriques sont appliqués et comparés sur des séries déjà désaisonnalisées ou sans saisonnalité (séries simulées). 
-Les révisions et donc le déphasage sont limités à 6 mois (lorsque le filtre symétrique est de 13 termes) et cela a l'avantage d'isoler les impacts des différents filtres des autres processus inhérents à la désaisonnalisation.
-Il y a toutefois deux inconvénients à cette simplification :
-
-1. D'une part, l'estimation de la série désaisonnalisée dépend de la méthode utilisée pour extraire la tendance-cycle.
-Le choix de la méthode utilisée pour l'estimation de la tendance-cycle peut donc avoir un impact bien au-delà de 6 mois.
-
-2. D'autre part, les moyennes mobiles étant des opérateurs linéaires, ils sont sensibles à la présence de points atypiques.
-L'application directe des méthodes peut donc conduire à des estimations biaisées, du fait de leur présence, alors que les méthodes de désaisonnalisation (comme la méthode X-13ARIMA-SEATS) ont un module de correction des points atypiques.
-Par ailleurs, comme notamment montré par @dagum1996new, le filtre symétrique final utilisé par X-13ARIMA pour extraire la tendance-cycle (et donc celui indirectement utilisé lorsqu'on applique les méthodes sur les séries désaisonnalisées) laisse passer environ 72 % des cycles de 9 ou 10 mois (généralement associés à du bruit plutôt qu'à la tendance-cycle).
-Les filtres asymétriques finaux amplifient même les cycles de 9 ou 10 mois.
-Cela peut avoir pour conséquence l'introduction d'ondulations indésirables, c'est-à-dire la détection de faux points de retournement.
-Ce problème est réduit par la correction des points atypiques (ces cycles étant considérés comme de l'irrégulier). 
-C'est ainsi que le *Nonlinear Dagum Filter* (NLDF) a été développé et consiste à :
-
-    a. appliquer l'algorithme de correction des points atypiques de X-13ARIMA sur la série désaisonnalisée, puis de la prolonger par un modèle ARIMA ;
-
-    b. effectuer une nouvelle correction des points atypiques en utilisant un seuil bien plus strict et appliquer ensuite le filtre symétrique de 13 termes. 
-    En supposant une distribution normale cela revient à modifier 48 % des valeurs de l'irrégulier.
-
-    Les *cascade linear filter* (CLF), notamment étudiés dans @dagumBianconcini2023, correspondent à une approximation des NLDF en utilisant un filtre de 13 termes et lorsque les prévisions sont obtenus à partir d'un modèle ARIMA(0,1,1) où $\theta=0,40.$
-
-Une piste d'étude serait alors d'étudier plus précisément l'effet des points atypiques sur l'estimation de la tendance-cycle et la détection des points de retournement, mais aussi d'explorer de nouveaux types de filtres asymétriques fondés sur des méthodes robustes (comme les régressions locales robustes, les médianes mobiles, etc.).
+In this article and in those associated to trend-cycle extraction techniques, the asymmetrical filters are applied and compared on series already seasonally adjusted or without seasonality (simulated series).
+The revisions and therefore the phase shift are limited to 6 months (when the symmetrical filter is 13 terms), and this has the advantage of isolating the impacts of the different filters from the other processes inherent in seasonal adjustment.
+There are, however, two drawbacks to this simplification:
+
+1. Firstly, the estimation of the seasonally-adjusted series depends on the method used to extract the trend-cycle.
+The choice of method used to estimate the trend-cycle can therefore have an impact well beyond 6 months.
+
+2. On the other hand, as moving averages are linear operators, they are sensitive to the presence of atypical points.
+Direct application of the methods can therefore lead to biased estimates, due to their presence, whereas seasonal adjustment methods (such as the X-13ARIMA-SEATS method) have a correction module for atypical points.
+Furthermore, as shown in particular by @dagum1996new, the final symmetrical filter used by X-13ARIMA to extract the trend-cycle (and therefore the one indirectly used when applying the methods to seasonally-adjusted series) misses around 72% of 9- or 10-month cycles (generally associated with noise rather than the trend-cycle).
+The final asymmetrical filters even amplify 9 or 10-month cycles.
+This can result in the introduction of undesirable ripples, i.e. the detection of false turning points.
+This problem is reduced by correcting atypical points (these cycles being considered irregular). 
+This is how the *Nonlinear Dagum Filter* (NLDF) was developed:
+
+    a. applying the X-13ARIMA atypical point correction algorithm to the seasonally adjusted series, then extending it with an ARIMA model;
+
+    b. perform a new atypical point correction using a much stricter threshold and then apply the 13-term symmetrical filter. 
+    Assuming a normal distribution, this amounts to modifying 48% of the irregular values.
+
+    The *cascade linear filter* (CLF), studied in particular in @dagumBianconcini2023, corresponds to an approximation of the NLDF using a 13-term filter and when the forecasts are obtained from an ARIMA(0,1,1) model where $\theta=0.40.$.
+
+One possible direction of study would then be to investigate more precisely the effect of atypical points on trend-cycle estimation and the detection of turning points, but also to explore new types of asymmetric filters based on robust methods (such as robust local regressions, moving medians, etc.).
+
+```{r, out.width="95%", fig.show="hold"}
+#| label: fig-ce16ovcovid
+#| fig-cap:  Successive estimates of the trend-cycle in US employment (in logarithms) during the COVID-19.
+series <- "CE16OV"
+date <- "covid"
+img <- sprintf("img/nber/%s_%s_%s.%s",
+               tolower(series),
+               date,
+               c("lp"),
+               "pdf")
+knitr::include_graphics(img)
+```
+
+```{r, out.width="95%"}
+#| label: fig-ce16ov-previmp-covid
+#| fig-cap: Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
+series <- "CE16OV"
+date <- "covid"
+img <- sprintf("img/nber/%s_%s_prev_imp_%s.%s",
+               tolower(series),
+               date,
+               c("lp"),
+               "pdf")
+knitr::include_graphics(img)
+```
 
@@
 
-\appendix
-\newpage
-
-# Coefficients, gain and phase shift functions {#sec-an-cof}
-
-```{r, out.width="100%"}
-#| label: fig-graphs-coef-lc
-#| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$.
-img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf")
-knitr::include_graphics(img)
-```
-
-```{r, out.width="100%"}
-#| label: fig-graphs-coef-ql
-#| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$.
-img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf")
-knitr::include_graphics(img)
-```
-
-```{r, out.width="100%"}
-#| label: fig-graphs-coef-cq
-#| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$.
-img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf")
-knitr::include_graphics(img)
-```
-
-```{r, out.width="100%"}
-#| label: fig-graphs-coef-daf
-#| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF).
-img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf")
-knitr::include_graphics(img)
-```
+<!-- \appendix -->
+<!-- \newpage -->
+
+<!-- # Coefficients, gain and phase shift functions {#sec-an-cof} -->
+
+<!-- ```{r, out.width="100%"} -->
+<!-- #| label: fig-graphs-coef-lc -->
+<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$. -->
+<!-- img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf") -->
+<!-- knitr::include_graphics(img) -->
+<!-- ``` -->
+
+<!-- ```{r, out.width="100%"} -->
+<!-- #| label: fig-graphs-coef-ql -->
+<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$. -->
+<!-- img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf") -->
+<!-- knitr::include_graphics(img) -->
+<!-- ``` -->
+
+<!-- ```{r, out.width="100%"} -->
+<!-- #| label: fig-graphs-coef-cq -->
+<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$. -->
+<!-- img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf") -->
+<!-- knitr::include_graphics(img) -->
+<!-- ``` -->
+
+<!-- ```{r, out.width="100%"} -->
+<!-- #| label: fig-graphs-coef-daf -->
+<!-- #| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF). -->
+<!-- img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf") -->
+<!-- knitr::include_graphics(img) -->
+<!-- ``` -->

### 2024-04-24 paper/resp_referee.qmd
update plots + articles

@@
 
+*The final filter used in X-11 is much longer than one year.*
+*The series are thus just partially extended (and asymmetric filters are still used): this avoid problems associated to long horizon forecasting (outliers/shifts).*
+*This was clarified in the text.*
+
 > 5. Page 8, line 10: omit the $\forall q$ since you have stated this qualification in the previous line.
@@
 > Add more explanation to the text.
+
+*More explanation are added to this section*
 
@@
 
+*All the section was rewritten to be clearer*
+
 > 12. Appendix B: rather than typing out the R code, it would be better to have the files available as supplementary material. 
@@
 
-The modelled of the input time series is now describe more precisely in section 2 (previous section 1). $TC_t$ is here considered as deterministic
+*The modelled of the input time series is now describe more precisely in section 2 (previous section 1). $TC_t$ is here considered as deterministic*
 
@@
 
+*The definition was clarified.*
+*In this paper we focus on turning points associated to business-cycles: they are defined as changepoints.*
+*However, the method usually used to detect turning point makes no difference between regular points at which the slope of the trend changes sign and changepoints.*
+*It is then possible to detect "false turning points" (undesirable ripples): that's why in the simulation we only focus on the simulated turning points and in real data application we look at turning points which are officially dated (for the US by the NBER).*
+
 > 4. The notation should be unified: why not writing $TC_t = \mu(t)$ and in section 2 using $t$ instead of $x$? 
@@
 
+*The notation is simplified and unified and the equations/notations clarified.*
+*The matrix $D$ is needed to see the equivalence with symmetric filter/DAF (clarified in the text).*
+
 > 5. You should state that you assume a fixed bandwidth. 
@@
 
-*It was added in the beginning of new sections 3 and 4 than the bandwidth is fixed.
-Nearest neighbour bandwidth was added in supplemental material but it doesn't improve the results in therm of phase shift.
-I can be explain that the bi*
+*It was added in the beginning of new sections 3 and 4 than the bandwidth is fixed.* 
+*Nearest neighbour bandwidth was added in supplemental material but it doesn't improve the results in therm of phase shift.* 
+*One explanation could be that the bias is higher with LC/QL methodes (we model linear trend using higher bandwidth) and for other methods (e.g.
+DAF) that the revisions to the final symmetric filters are higher;*
 
@@
 
-In local polynomial smoothing, the notion of bias is associated to the approximation error of the smoothed series by a polynomial degree of order $d$.
-It can thus be writer as
+*In local polynomial smoothing, the notion of bias is associated to the approximation error of the smoothed series by a polynomial degree of order $d$.*
+*If the moving average $\theta$ preserve a polynomial of degree $d$, the bias at the date $t$ is then:*
+$$
+\mathbb E[\hat \mu(t)] - \mu(t)=\sum_{j=d+1}^{+\infty}\frac{\mu^{(j)}(t)}{j!}\sum_{k=-h}^k\theta_k k^j
+$$
+*It is then not directly related to the phase shift.*
 
@@
 
+*The term "undersirable oscillations" was removed*
+
 > - Figure 1. In the caption, the ratio $I/C$ was never defined before.
@@
 Perhaps the authors should state that their intent is to separate the trend-cycle component from the ’noise’ component.
+
+*Reference are added and the goal of the paper was clarified in new section 2 (previous section 1)*
 
@@
 Another thought is that the ARIMA model being used to extend the series appears to be modeled using the full series; what if the ARIMA model used to extend the series was instead estimated using just a short span toward the end of the series?
+
 *Indeed it is in general not relevant to identify an ARIMA model on a monthly time series of length 60 years.*
-In our case, since the same data generation process remains the same, it could be relevant and leads to better results in terms of delay to detect turning points (in supplement materials, a plot is added comparing the results with different spans to identify/estimate the ARIMA model).
-However, on the main text, we now only used 12 years to identify the ARIMA model, to be closer to what would be done on real series.
+*In our case, since the same data generation process remains the same, it could be relevant and leads to better results in terms of delay to detect turning points (in supplement materials, a plot is added comparing the results with different spans to identify/estimate the ARIMA model).*
+*However, on the main text, we now only used 12 years to identify the ARIMA model, to be closer to what would be done on real series.*
+*Since other methods are local, it is not a problem to have long time series.*
 
@@
 > Some of the important results appear in the tables and figures, but the accompanying discussion in the text should still be fleshed out more.
-In addition, in Figures 2-4 and 6-7 (and the figures in the supplement), the lines can be distinguished by color, but it may be helpful to find some way to keep them distinguishable without the benefit of color.
+
+> In addition, in Figures 2-4 and 6-7 (and the figures in the supplement), the lines can be distinguished by color, but it may be helpful to find some way to keep them distinguishable without the benefit of color.
 It might be the case that there is too much information to display neatly in some of these.
+
+*Figures are now black and white a scaled gray legend.*
+*In some of them, there are indeed a lot of information with several labels, but the goal is not to be able to read all the details of all the lines but to see the general behavior of the filters (for example if the implicit forecasts seem relevant or not)*
 
@@
 
+*We try to clarify in the section associated to the proposed methods: all the functions mentionned are already in rjd3filters.*
+*The local parametrisation is not directly in the package but can be easily implemented (especially because the code is available): the variance estimator is in the package but not the moving-averages associated to slope/concavity estimators (which can be easily derived)*
+
 Other (non-exhaustive) [only remarks that require a response]
-- P2 L8:  “X-13ARIMA seasonal adjustment method” -- The software is officially X-13ARIMA-SEATS, but the seasonal adjustment method itself is X-11 (applies to subsequent occurrences as well)
+
+> - P2 L8:  “X-13ARIMA seasonal adjustment method” -- The software is officially X-13ARIMA-SEATS, but the seasonal adjustment method itself is X-11 (applies to subsequent occurrences as well)
 
@@
 
-- P2 L29:  “good properties” -- This could be less vague.
+> - P2 L29:  “good properties” -- This could be less vague.
 
-- P7 L58:  There is a $q$ introduced here via “$w_t^q$ for $t>q$” that shows up later as well, but  itself does not appear to be properly defined otherwise.
+*TODO*
+
+> - P7 L58:  There is a $q$ introduced here via “$w_t^q$ for $t>q$” that shows up later as well, but  itself does not appear to be properly defined otherwise.
 
@@
 
-- P9:  Author lists a number of quality criteria in Table 1.
+> - P9:  Author lists a number of quality criteria in Table 1.
 It may be obvious for some of these, but it would be helpful to the reader to indicate whether lower values or higher values are preferable for each, either in the table or in the accompanying exposition.
@@
 
-- P11 L22:  “$2h+1 \geq d+1$” -- Why not just say $2h\geq d$?
+> - P11 L22:  “$2h+1 \geq d+1$” -- Why not just say $2h\geq d$?
 
@@
 
-- P14 L33:  More exposition wrt Table 2 may be helpful, as there is a fair amount of information to digest.
+> - P14 L33:  More exposition wrt Table 2 may be helpful, as there is a fair amount of information to digest.
 
@@
 
-- P14 L45:  What is the rationale for this combined measure $EQM_w=A_w + S_w + T_w + R_w$?
-Looking at the expressions for these four from Table 1, it is apparent that $A_w$ and $T_w$ sum up (the two have the same integrand, and the resulting sum is the integral from 0 to $\pi$), and same goes for $S_w$ and $R_w$.
-And it should probably be explicitly stated what $EQM_w$ stands for prior to introducing it in the table.
+> - P14 L45:  What is the rationale for this combined measure $EQM_w=A_w + S_w + T_w + R_w$?  
+> Looking at the expressions for these four from Table 1, it is apparent that $A_w$ and $T_w$ sum up (the two have the same integrand, and the resulting sum is the integral from 0 to $\pi$), and same goes for $S_w$ and $R_w$.
+> And it should probably be explicitly stated what $EQM_w$ stands for prior to introducing it in the table.
 
@@
 
-- P19 L15:  “January 1960 and December 2020” -- As noted previously, is it worth considering an additional scenario with (simulated) series of shorter length?
+> - P19 L15:  “January 1960 and December 2020” -- As noted previously, is it worth considering an additional scenario with (simulated) series of shorter length?
 
@@
 
+*Quarterly series with simulated data are added in supplemental material, a new section was added to discuss the use of the methods with different periodicities*
+
 > 5.  Clarity on Phase Shift Computation (Page 20)
@@
 
+*The definition of phase-shift was clarified in*
+
 > 6.  Commentary on Figure 5
@@
 Providing explanations for the results shown in the figure would be crucial for readers to better understand the performance of the filters/methods.
+
+*TODO*
 
@@
 
+*Quarterly series with simulated data are added in supplemental material, a new section was added to discuss the use of the methods with different periodicities*
+
 # Referee 4
@@
 
+*We try to clarify the contribution in the abtract, introduction and in the beginning of sections 2 and 3*
+
 > 2) The paper omits to consider the Covid19 period and how this influences the trend and cycle decomposition. 
@@
 
-> 3) I think a quote from other works can be removed from the paper. For example, I
-would remove the quote in point 2) of page 7.
+*A new section was added with an example around Covid19 and some possible solutions for outliers*
+
+> 3) I think a quote from other works can be removed from the paper. 
+For example, I would remove the quote in point 2) of page 7.
 

### 2024-04-24 paper/supp_material.qmd
update plots + articles

@@
 \newpage
+
 # Quarterly data
@@
 
-::: {#fig-graphs-impact-span}
+::: {#fig-nn-lp}
 ```{r, echo=FALSE, out.width="100%"}
@@
 
-::: {#fig-graphs-impact-span}
+::: {#fig-nn-localparam}
 ```{r, echo=FALSE, out.width="100%"}
-knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp.pdf")
+knitr::include_graphics("img/simulations/phase_shift_simul_nn_localparam.pdf")
 ```
-Note:
+Note: nearest bandwidth moving average are used to estimate slope and concavity.
+We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.
+Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).
 

### 2024-04-24 R_filters/plot_coef.R

@@ -31,7 +31,7 @@ for(i in names(all_filters)){
       ggplot_phase(x, xlim = c(0, 4/12*pi), q = c(0:6))+
       ggplot2::labs(y = "Phase shift") +
       ggplot2::guides(colour = "none"))
-  p
+  p <- p & ggplot2::scale_color_grey()
   ggsave(sprintf("paper/img/filters_used/%s.pdf",tolower(i)),
               width = 8, height = 4.5,
               plot = p)
@@ -76,6 +76,7 @@ p <- ggplot_coef(x, q = c(0:6)) / (
     ggplot_phase(x, xlim = c(0, 4/12*pi), q = c(0:6))+
     ggplot2::labs(y = "Phase shift") +
     ggplot2::guides(colour = "none"))
+p <- p & ggplot2::scale_color_grey()
 p
 ggsave("paper/img/filters_used/nn_henderson.pdf",
        width = 8, height = 4.5,
@@ -102,7 +103,7 @@ for(i in names(all_filters)){
       ggplot_phase(x, xlim = c(0, 4/12*pi), q = c(0:6))+
       ggplot2::labs(y = "Phase shift") +
       ggplot2::guides(colour = "none"))
-  p
+  p <- p & ggplot2::scale_color_grey()
   ggsave(sprintf("paper/img/filters_used/%s_nn.pdf",tolower(i)),
          width = 8, height = 4.5,
          plot = p)

### 2024-04-24 R_fredm/4B_Analyse_implicit_forecast.R

@@ -10,7 +10,6 @@ all_tp_rev <- get_all_tp(dossier = "results_fredm/compile_tp/")
 
 detected_tp <- readRDS("results_fredm/compile_tp_norev/detected_tp_lp.RDS")
 
-all_tp |> dplyr::filter(series == "CE16OV")
 series <- "CE16OV"
 tp_keep <- "2001.16666666667"
 all_prevs <- get_all_prevs(series = series, tp_keep = tp_keep,
@@ -37,14 +36,11 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              all_tp = all_tp,
                              all_tp_rev = all_tp_rev,
                              tp_keep = tp_keep)
-wrap_plots(plots,ncol = 3) & 
-  scale_color_grey() & 
-  labs(subtitle = NULL)
-
+wrap_plots(plots,ncol = 3) &
+  scale_color_grey()
 ggsave("paper/img/nber/ce16ov_covid_prev_imp_lp.pdf",
-       plot = wrap_plots(plots, ncol = 3) & 
-         scale_color_grey() & 
-         labs(subtitle = NULL),
+       plot = wrap_plots(plots, ncol = 3) &
+         scale_color_grey(),
        width = 8, height = 8)
 
 
@@ -61,7 +57,7 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
 wrap_plots(plots, ncol = 3)
 
 ggsave("paper/img/nber/retailx_nov2007_prev_imp_lp.pdf",
-       plot = wrap_plots(plots, ncol = 3)& 
-         scale_color_grey() & 
+       plot = wrap_plots(plots, ncol = 3)&
+         scale_color_grey() &
          labs(subtitle = NULL),
        width = 8, height = 8)

### 2024-04-24 R_fredm/4_utils.R

@@ -89,7 +89,7 @@ extract_est_data <- function(method = "lc", kernel = "henderson", nb_est = 10,
         dir <- "localic_daf_trunc"
         full_name <- gsub("localic_", "", method)
       }
-    } 
+    }
   }
   file = sprintf("results_fredm/%s/%s%s%s.RDS", dir, series, sep, full_name)
   data <- readRDS(file)
@@ -122,7 +122,7 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
           readRDS(sprintf("%speaks_localic_final.RDS", dossier)),
           by=c("series","kernel", "method", "degree", "h")) %>%
     # dplyr::filter(degree == "d2", h == "h6") %>%
-    select_var() %>% 
+    select_var() %>%
     mutate(method = sprintf("%s_localic_final_%s_%s", method, h, degree)) %>%
     select(!c(degree, h))
 
@@ -131,26 +131,26 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
           readRDS(sprintf("%speaks_localic_daf_trunc.RDS", dossier)),
           by=c("series","kernel", "method", "degree", "h")) %>%
     # dplyr::filter(degree == "d2", h == "h6") %>%
-    select_var() %>% 
+    select_var() %>%
     mutate(method = sprintf("%s_localic_%s_%s", method, h, degree)) %>%
     select(!c(degree, h))
 
   order_methods  <- c(
     "lc",
     "lc_localic_final_h6_d2",
-    "lc_localic_final_h3_d2", "lc_localic_final_h4_d2", "lc_localic_final_h5_d2", 
-    "lc_localic_final_h3_d3", "lc_localic_final_h4_d3", "lc_localic_final_h5_d3", 
+    "lc_localic_final_h3_d2", "lc_localic_final_h4_d2", "lc_localic_final_h5_d2",
+    "lc_localic_final_h3_d3", "lc_localic_final_h4_d3", "lc_localic_final_h5_d3",
     "lc_localic_final_h6_d3",
-    "lc_localic_h3_d2", "lc_localic_h4_d2", "lc_localic_h5_d2", 
-    "lc_localic_h6_d2", "lc_localic_h3_d3", "lc_localic_h4_d3", "lc_localic_h5_d3", 
-    "lc_localic_h6_d3", 
[…]

### 2024-04-24 R_simul/1B_plot_data.R

@@ -20,13 +20,16 @@ p <- ggplot(data = dataGraph,
   ) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 12),
                      labels = function(x) format(x, decimal.mark = ".")) +
-  theme_bw() +theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
-                                                     colour = "gray80", linetype = "solid"),
-                    legend.justification = c(0,0),
-                    legend.position = c(0,0),
-                    legend.key = element_blank(),
-                    legend.title = element_blank())
-
+  theme_bw() +
+  theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
+                                         colour = "gray80", linetype = "solid"),
+        legend.position = "inside",
+        legend.justification = c
+        (0,0),
+        legend.key = element_blank(),
+        legend.title = element_blank()) +
+  scale_color_grey()
+p
 ggsave("paper/img/simulations/simul_data.pdf",
        plot = p,
        width = 8, height = 4)

### 2024-04-24 R_simul_quarter/1A_data_creation.R

@@ -23,22 +23,6 @@ series_q = list(
   lowvariability3 = simulated_tci(time,sigma_nu = 0.08,sigma_e = 0.40,lambda = 72/4,rho = 4)
 )
 series_q = lapply(series_q,ts, start = start, frequency = frequency)
-do.call(rbind, lapply(series_q, function(x){
-  x <- x[,"tc"]
-  res = rjd3x11plus::select_trend_filter(x, length = 5)
-  res = c(res[c("length", "icr")],
-          rjd3x11plus::select_trend_filter(x, length = 7)[1])
-  names(res) = c("optimal_length", "icr-5", "icr-7")
-  res
-}))
-forecast::autoplot(do.call(cbind, lapply(series_q[c(2,5,8)],`[`, ,"tc")))  +
-  theme_bw() +theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
-                                                     colour = "gray80", linetype = "solid"),
-                    legend.justification = c(0,0),
-                    legend.position = c(0,0),
-                    legend.key = element_blank(),
-                    legend.title = element_blank()) +
-  labs(x = NULL, y = NULL)
 
 tp = turning_points(series_q[[1]][,"cycle"])
 first_date = time(series_q[[9]][,"cycle"])[25]

### 2024-04-24 R_simul_quarter/1B_plot_data.R

@@ -1,33 +1,0 @@
-library(AQLThesis)
-load_simul <- function(x){
-  data <- readRDS(sprintf("data_simul_quarter/byseries/%s.RDS", x))
-  data <- data[[length(data)]]
-  data
-}
-data <- do.call(cbind, lapply(sprintf("%svariability%i", rep(c("low", "medium", "high"), each = 3),
-                                      rep(1:3, 3)),load_simul))
-data <- data[,seq(from = 2, by = 3, length.out = 3)]
-colnames(data) <- paste(c("Low", "Medium", "High"), "variability")
-dataGraph <- data.frame(cbind(time(data), data))
-colnames(dataGraph) <- c("date", colnames(data))
-dataGraph <- reshape2::melt(dataGraph, id="date")
-p <- ggplot(data = dataGraph,
-            aes(x = date, y = value, group = variable, colour = variable)) +
-  geom_line(linewidth=0.7, data = dataGraph) +
-  labs(title = NULL,
-       x = NULL, y = NULL) +
-  scale_x_continuous(breaks = scales::pretty_breaks(n = 10),
-  ) +
-  scale_y_continuous(breaks = scales::pretty_breaks(n = 12),
-                     labels = function(x) format(x, decimal.mark = ",")) +
-  theme_bw() +theme(legend.background = element_rect(fill = alpha('gray99', 0.4),
-                                                     colour = "gray80", linetype = "solid"),
-                    legend.justification = c(0,0),
-                    legend.position = c(0,0),
-                    legend.key = element_blank(),
-                    legend.title = element_blank())
-p
-ggsave("paper/img/simulations/simul_data_q",
-       width = 8, height = 4)
-
-

### 2024-04-24 ex_pente.R

@@ -66,8 +66,8 @@ tp = readRDS("data_simul/tp_simul1.RDS")
 p <- autoplot(data[,c(1,2)]) +
   geom_vline(data = data.frame(xintercept = c(tp$downturn,
                                               tp$upturn)),
-             aes(xintercept = xintercept)) +
-  scale_color_discrete(
+             aes(xintercept = xintercept),alpha = 0.5) +
+  scale_color_grey(
     breaks = c("icr", "icr_local"),
     labels = c(
       latex2exp::TeX("Global $|\\delta_1/\\sigma|"),
@@ -76,12 +76,16 @@ p <- autoplot(data[,c(1,2)]) +
   labs(x = NULL, y = NULL) +
   theme_bw() +
   theme(legend.justification = c(0,1),
-        legend.position = c(0,1),
+        legend.position = "inside",
         legend.key = element_blank(),
         legend.title = element_blank(),
         legend.background = element_rect(fill = alpha('gray99', 0.8),
-                                         colour = "gray80", linetype = "solid")
-  )
+                                         colour = "gray80", linetype = "solid"),
+        panel.grid.minor = element_blank(),
+        panel.grid.major = element_blank()
+  ) +
+  coord_cartesian(ylim = c(0, 1.8), expand = FALSE,
+                  xlim = c(1960, 2022))
 
 p
 ggsave("paper/img/filters_used/mm_penteconcavite_ex.pdf",
@@ -91,8 +95,9 @@ ggsave("paper/img/filters_used/mm_penteconcavite_ex.pdf",
 p2 <- autoplot(data[,c(1,3)]) +
   geom_vline(data = data.frame(xintercept = c(tp$downturn,
                                               tp$upturn)),
-             aes(xintercept = xintercept)) +
-  scale_color_discrete(
+             aes(xintercept = xintercept),
+             alpha = 0.5) +
+  scale_color_grey(
     breaks = c("icr", "icr_local2"),
     labels = c(
       latex2exp::TeX("Global $|\\delta_1/\\sigma|"),
@@ -105,8 +110,12 @@ p2 <- autoplot(data[,c(1,3)]) +
         legend.key = element_blank(),
         legend.title = element_blank(),
         legend.background = element_rect(fill = alpha('gray99', 0.8),
-                                         colour = "gray80", linetype = "solid")
-  )
+                                         colour = "gray80", linetype = "solid"),
[…]

### 2024-04-25 paper/article_jos.qmd
update article

@@
 This allows us to understand the foundations behind the construction of moving averages from local polynomial regressions, as well as those behind the two extensions proposed in this article (@sec-nonparamreg).
-In a final section (@sec-comparison), all these methods are compared empirically on simulated and real series.
+In @sec-comparison, all these methods are compared empirically on simulated and real series.
 
@@
 with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
-Event if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
+Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
 Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
@@
 
-Appendix [-@sec-an-cof] shows the coefficients, gain and phase shift functions of the four asymmetric filters.
+Supplemental materials shows the coefficients, gain and phase shift functions of the four asymmetric filters.
 
@@
 
-The series associa
-
 ```{r, echo=FALSE, out.width="100%"}
@@
 
+The simulated series are of length 60 years, which is often not realistic for economic time series.
+However, since all the methods are local, the results are not affected by the length of the series.
+The length of the series would only have an impact for the identification and the estimation of the ARIMA model.
+In this case, since the same data generating process is used during the 60 years, it could be relevant to identify the ARIMA model using all the data.
+Even if it would improve the results in term of phase shift (see supplemental material) we prefer to only use the last 12 years to identify and estimate the ARIMA model, to be closer to what would have been done with a real time series.
+
 ## Comparison
@@
 
+::: {#fig-graphstpsimul}
+
 ```{r, echo=FALSE, out.width="100%"}
-#| label: fig-graphstpsimul
-#| fig-cap: Distribution of phase shifts on simulated series.
-img <- sprintf("img/simulations/phase_shift_simul.%s", "pdf")
-knitr::include_graphics(img)
-```
+knitr::include_graphics("img/simulations/phase_shift_simul.pdf")
+```
+Lecture: this plots represent box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).\
+For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally (using), 50% of the turning point are detected with a phase shift of at most 4 months.
+For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).
+
+Distribution of phase shift on simulated series.
+:::
 
@@
 
-# Discussion
-## Outliers and COVID-19
+## Discussion
+### Outliers and COVID-19
 
@@
 
-## Other frequencies
+### Other frequencies
 
@@
 :::
-
-<!-- \appendix -->
-<!-- \newpage -->
-
-<!-- # Coefficients, gain and phase shift functions {#sec-an-cof} -->
-
-<!-- ```{r, out.width="100%"} -->
-<!-- #| label: fig-graphs-coef-lc -->
-<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Linear-Constant (LC) filter with $I/C=3.5$. -->
-<!-- img <- sprintf("img/filters_used/%s.%s", tolower("LC"), "pdf") -->
-<!-- knitr::include_graphics(img) -->
-<!-- ``` -->
-
-<!-- ```{r, out.width="100%"} -->
-<!-- #| label: fig-graphs-coef-ql -->
-<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Quadratic-Linear (QL) filter with $I/C=3.5$. -->
-<!-- img <- sprintf("img/filters_used/%s.%s", tolower("QL"), "pdf") -->
-<!-- knitr::include_graphics(img) -->
-<!-- ``` -->
-
-<!-- ```{r, out.width="100%"} -->
-<!-- #| label: fig-graphs-coef-cq -->
-<!-- #| fig-cap: Coefficients, gain and phase shift functions for the Cubic-Quadratic (CQ) filter with $I/C=3.5$. -->
-<!-- img <- sprintf("img/filters_used/%s.%s", tolower("CQ"), "pdf") -->
-<!-- knitr::include_graphics(img) -->
-<!-- ``` -->
-
-<!-- ```{r, out.width="100%"} -->
-<!-- #| label: fig-graphs-coef-daf -->
-<!-- #| fig-cap: Coefficients, gain and phase shift functions for the direct asymmetric filter (DAF). -->
-<!-- img <- sprintf("img/filters_used/%s.%s", tolower("DAF"), "pdf") -->
-<!-- knitr::include_graphics(img) -->
-<!-- ``` -->

### 2024-04-25 paper/article_jos.tex
update article

@@
 particular, this theory allows us to reproduce Henderson's symmetric and
-Musgrave's asymmetric filters. This paper describes how they can be
-extended to include a timeliness criterion to minimise the phase shift.
-While asymmetric filters are generally parametrised globally, which can
-be suboptimal around turning points, this paper proposes a procedure for
-parametrising them locally. An empirical comparison, based on simulated
-and real data, shows that modelling polynomial trends that are too
-complex introduces more revisions without reducing the phase shift, and
-that local parametrisation reduces the delay in detecting turning points
-and reduces revisions. The paper also shows how these results can be
-easily reproduced using the R package \texttt{rjd3filters} dedicated to
-the manipulation of moving averages.
+Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal
+adjustment algorithm. This paper proposes two extensions of local
+polynomial filters for real-time trend-cycle estimates (for the last
+estimates): first including a timeliness criterion to minimise the phase
+shift; second, with procedure for parametrising asymmetric filters
+locally While they are generally parametrised globally, which can be
+suboptimal around turning points. An empirical comparison, based on
+simulated and real data, shows that modelling polynomial trends that are
+too complex introduces more revisions without reducing the phase shift,
+and that local parametrisation reduces the delay in detecting turning
+points and reduces revisions. The paper also shows how these results can
+be easily reproduced using the R package \texttt{rjd3filters} dedicated
+to the manipulation of moving averages.
 
@@
 particular, it becomes possible to reproduce Henderson's symmetric
-filter and Musgrave's asymmetric filters. However, these methods have
-two drawbacks. Firstly, the phase shift (i.e.~the delay in detecting
-turning points) is not directly controlled. Secondly, asymmetric filters
-depend on a parameter that is generally estimated over the whole series
-using assumptions that may be wrong locally. The aim of this study is to
-propose extensions to the \textcite{proietti2008} class of filters in
-order to take these shortcomings into account. The extensions have been
-implemented in the statistical software R \autocite{logicielr}
-\texttt{rjd3filters} package (available at
-\url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
-easily used.
+filter and Musgrave's asymmetric filters. The aim of this study is to
+propose two extensions to the \textcite{proietti2008} class of filters
+in order to two drawbacks of those methods. Firstly, adding, in the
+optimisation process, a timeliness criterin to directly control the
+phase shift (i.e.~the delay in detecting turning points). Secondly,
+calibrating asymmetric filter locally, while they are generally
+calibrated globally using thus assumptions that may be wrong locally.
+Those extensions are implemented in the statistical software R
+\autocite{logicielr} throw the \texttt{rjd3filters} package (available
+at \url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
+easily used. All the code associated to this paper is available at
+\url{https://github.com/XXXX} (anonymized linked before the manuscript
+is accepted).
 
@@
 extensions proposed in this article (section~\ref{sec-nonparamreg}). In
-a final section (section~\ref{sec-comparison}), all these methods are
-compared empirically on simulated and real series.
-
-\section{Some properties of moving averages}\label{sec-propMM}
+section~\ref{sec-comparison}, all these methods are compared empirically
+on simulated and real series.
+
+\section{Local trend-cycle models and moving averages}\label{sec-propMM}
+
+The basic assumption used in time series decomposition methods is that
+the input observed time series, \(y_t\) can be decomposed into a signal
+component \(\mu_t\) and a an erratic component \(\varepsilon_t\) (called
+irregular): \[
+f(y_t)=\mu_t+\varepsilon_t
+\] where \(f\) denotes some appropriate transformation (usually
+logarithmic or no transformation). To simplify further notations,
+\(y_t\) will denote the transformed observed series. The noise component
+\(\varepsilon_t\) is usually assumed to be a white noise. Assuming that
+the input time series is seasonally adjusted (or without seasonality),
+the signal \(\mu_t\) represents the trend and cycle components, here
+estimated jointly and called trend-cycle \(TC_t\). This component can be
+locally approximated by a polynomial degree (see
+section~\ref{sec-nonparamreg}): \[
+\mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
+\] with \(\xi_t\) a stochastic process uncorrelated with
+\(\varepsilon_t\). Even if some papers model \(\xi_t\) and
+\(\varepsilon_t\) separately \autocite[see for
+example][]{GrayThomson2002}, an usual assumption, adopted in this paper,
+is to is to gather \(\xi_t\) and \(\varepsilon_t\). Thus, the
+trend-cycle \(TC_t\) is considered as deterministic and modelled as a
+polynomial trend of degree \(d\). The coefficients
+\((\beta_0,\dots,\beta_k)\) can be estimated by weighted least squares.
+The estimate \(\hat \beta_0\) provides the trend-cycle estimate
+\(\hat TC_t\) and it can be shown that this equivalent to apply of
+moving average (see section~\ref{sec-nonparamreg}). The goal of this
+paper is to present two new methods to separate the trend-cycle
+component from the noise component in real-time (i.e.~when no future
+point is known), using moving averages build on local polynomial
+smoothing techniques.
 
@@
 \(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
-such as for all time series \(X_t\): \[
-M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
+such as for all time series \(y_t\): \[
+M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
 \]
@@
 illustrate the effects of moving averages on harmonic series
-\(X_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
-\(M_{\boldsymbol\theta}\) transforms \(X_t\) into: \[
-Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
-= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
+\(y_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
+\(M_{\boldsymbol\theta}\) transforms \(y_t\) into: \[
+Y_t = M_{\boldsymbol\theta}y_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
+= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot y_t.
 \] The function
@@
 To sum up, applying a moving average to a harmonic times series
-(\(X_t=\e^{-i\omega t}\)) affects it in in two different ways:
+(\(y_t=\e^{-i\omega t}\)) affects it in in two different ways:
 
@@
 this has no impact on the results of the study. The other frequencies,
-\([2\pi/12, \pi]\) are associated with the irregular component
-(undesirable oscillations).
-
-figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
+\([2\pi/12, \pi]\) are associated with the irregular component.
+
+Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
 for the asymmetric Musgrave filter (see section~\ref{sec-lppasymf})
@@
 
-\begin{figure}[H]
+\begin{figure}
 
 \caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
-function for the Musgrave filter in real time with \(I/C=3.5\).}
+function for the Musgrave filter used in X-11 in real time (last
+estimate) for monthly series when the bandwidth is 6.}
 
@@
 \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}
+
+Note: the Musgrave filter depends on a ratio \(I/C\) (see
+section~\ref{sec-nonparamreg}). In X-11, for monthly series, when the
+bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used,
+what is used in most cases), the ratio is fixed to 3.5 and this value is
+used in this figure.
 
@@
 
-In our case, we assume that our initial series \(y_t\) is seasonally
+In this paper, we assume that our initial series \(y_t\) is seasonally
 adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
-an irregular component, \(I_t\): \[
-y_t=TC_t+I_t.
+an irregular (noise) component, \(\varepsilon_t\): \[
+y_t=TC_t+\varepsilon_t.
 \] The aim will be to build moving averages that best preserve the
-trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)) and minimise
-noise (\(M_{\boldsymbol\theta} (I_t)\simeq 0\)).
-
-\subsubsection{Trend preservation}\label{trend-preservation}
+trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)).
 
 Trend-cycles are generally modelled by local polynomial trends (see
-section~\ref{sec-nonparamreg}). In order to best preserve the
-trend-cycles, we try to design moving averages that preserve the
+section~\ref{sec-nonparamreg}) and, in order to best preserve the
+trend-cycles, we thus design moving averages that preserve the
 polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
@@
 
-\subsubsection{Noise reduction}\label{noise-reduction}
-
-All time series are affected by noise, which can blur the extraction of
-trends and cycles. This is why we try to reduce this noise (by reducing
-its variance) while preserving the relevant trends. The sum of the
-squares of the coefficients \(\sum_{k=-p}^{+f}\theta_k^2\) is the
-\emph{reduction of variance} ratio.
-
-Indeed, let \(\{\varepsilon_t\}\) be a sequence of independent random
-variables with \(\E{\varepsilon_t}=0\), \(\V{\varepsilon_t}=\sigma^2\):
-\[
-\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
-= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
-\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
-\]
-
 \subsection{Real-time estimation and asymmetric moving
 average}\label{sec-mmasym}
-
-\subsubsection{Asymmetric moving averages and
-forecasts}\label{sec-mmetprev}
 
@@
   an ARIMA model to minimise the revisions associated to asymmetric
-  filters
+  filters. Since the final filter used for the seasonal adjustment
+  extraction is much longer than one year (in most of the cases, for
+  monthly series, the bandwidth is 78), the input series is just
+  partially extended (and asymmetric filters are still used) to prevent
+  from instability over long periods (e.g., outliers or model shifts).
 \end{enumerate}
@@
 facilitating the calculation of different moving averages at different
-time frequencies, this method can be used in R with the function
-\texttt{rjd3filters::rkhs\_filter()}.
+time frequencies. See \textcite{inseeDTM202401} for a more detailed
+comparison of the different recent methods for trend-cycle extraction
+included in the \texttt{rjd3filters} package
+
+In this paper we will focus on \textcite{proietti2008} approach.
+Section~\ref{sec-sym-lp} and section~\ref{sec-lppasymf} describes their
+approach to build symmetric moving averages (used for the final
+estimates of the trend-cycle component) and asymmetric moving averages
+(used for real-time estimates). This paper proposed two extensions of
+their approach to build asymetric moving averages: a first one adding a
+criteria to control the phase shift (section~\ref{sec-localic}), not
+used in the simulation (section~\ref{sec-comparison}) but which could be
+used in future studied; a second one using a local parameterisation of a
+parameter usually parameterised globally
+(section~\ref{sec-lptimeliness}), which is used in the simulation.
 
 \subsection{Symmetric moving averages and local polynomial
-regression}\label{symmetric-moving-averages-and-local-polynomial-regression}
+regression}\label{sec-sym-lp}
 
@@
 
-As mentioned in section~\ref{sec-mmetprev}, for real-time estimation,
+As mentioned in section~\ref{sec-mmasym}, for real-time estimation,
 several approaches can be used:
@@
 account the weights associated to the observations) and with
-\([\boldsymbol U,\boldsymbol Z] = \boldsymbol X\) we this find the model
+\([\boldsymbol U,\boldsymbol Z] = \boldsymbol X\) we thus find the model
 associated to symmetric filters. The objective is to find an asymmetric
@@
 preserving polynomials of degree \(d\) and we find asymmetric direct
-filters (DAF) when \(\boldsymbol D=\boldsymbol K^{-1}\).
+filters (DAF) when \(\boldsymbol D=\sigma^2\boldsymbol K^{-1}\).
 
@@
 
-Appendix \ref{sec-an-cof} shows the coefficients, gain and phase shift
+Supplemental materials shows the coefficients, gain and phase shift
 functions of the four asymmetric filters.
@@
   \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\widehat{TC}_t)^2}{1-2\theta_0^2+\sum_{i=-h}^{h} \theta_i^2}.
-  \]
+  \] This is implemented in the function
+  \texttt{rjd3filters::var\_estimator()}. This estimator assumes that
+  \(\widehat{TC}_t\) is unbiased, so variance estimates intervals are
+  usually computed at small bandwidths (\(h\)). In this paper we study
+  monthly series with small bandwidth (\(h=6\)), and the variance is
+  estimated using local polynomial of degree 3 (Henderson filter), the
+  biased of \(\widehat{TC}_t\) is thus already small and using smaller
+  bandwidths have almost no impact on the estimates (see Supplemental
+  materials).
 \item
@@
   moving averages, the trend can be modelled as being locally of degree
-  2 or 3 (this has no impact on the final estimate of concavity). Here,
-  we have chosen to model a trend of degree 2: this reduces the phase
-  shift but slightly increases the revisions linked to the first
-  estimate of the trend-cycle. figure~\ref{fig-mmpenteconcac} shows the
-  moving averages used.
+  2 or 3 (this has no impact on the final estimate of concavity). In
+  this paper, we have chosen to model a trend of degree 2: this slightly
+  reduces the phase shift (see Supplemental material) but slightly
+  increases the revisions linked to the first estimate of the
+  trend-cycle. figure~\ref{fig-mmpenteconcac} shows the moving averages
+  used.
 \end{itemize}
@@
 
-In the empirical applications of section~\ref{sec-comparison}, the final
-local parametrisation is obtained by estimating \(\delta\) using all the
-data (i.e., using the symmetric filters shown in
-figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
-of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex} shows an example of
-the comparison of estimates of the ratio \(|\delta_1/\sigma|\) with a
-global estimator (IC ratio) and two local parametrisations:
+To distinguish the gain from the local parametrisation of the noise
+biased associated to the real-time estimates of the slope and concavity,
+in the empirical applications of section~\ref{sec-comparison} we also
+compares the results to the final local parametrisation obtained by
+estimating \(\delta\) using all the data (i.e., using the symmetric
+filters shown in figure~\ref{fig-mmpenteconcac}), while maintaining a
+real-time estimate of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex}
+shows an example of the comparison of estimates of the ratio
+\(|\delta_1/\sigma|\) with a global estimator (IC ratio) and two local
+parametrisations:
 
@@
 
-The turning points are clearly detected by the local estimators, but
-there is much more variance in the estimate of \(|\delta_1/\sigma|\) in
-the real-time estimate. This will lead to another source of revisions in
-the intermediate estimates.
+The turning points are clearly detected by the local estimators
+(changement in the slope), when \(|\delta_1/\sigma|\) is estimated in
+the real-time estimate. However, the real-time local estimates of
+\(|\delta_1/\sigma|\) are noisy: the varability of this estimates, used
+to build asymmetric filters, will then lead to another source of
+revisions in the intermediate estimates.
 
@@
 \end{figure}%
+
+There is no function in \texttt{rjd3filters} to directly compute those
+moving averages but they can be easily computed using matrix computation
+formula and applied to the series rjd3filters`.
 
@@
 then applying a 13-term symmetric Henderson filter. The ARIMA model is
-determined automatically, using no seasonal lag (the series being
-seasonally adjusted) and no external variables (such as external
-regressors for correction of atypical points). The performance of the
-different methods is judged on criteria relating to revisions (between
-two consecutive estimates and in relation to the final estimate) and the
-number of periods required to detect turning points.
+determined automatically, using \textcite{autoarima} algorithm (function
+\texttt{forecast::auto.arima()}) on the last 12 years, with no seasonal
+lag (the series being seasonally adjusted) and no external variables
+(such as external regressors for correction of atypical points).
+
+The performance of the different methods is judged on criteria relating
+to revisions (between two consecutive estimates and in relation to the
+final estimate) and the number of periods required to detect turning
+points.
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Computation of the phase shift in the detection of turning points. The
-  definition of \textcite{Zellner1991} is used to determine the turning
-  points:
+\item
+  Computation of the phase shift in the detection of turning points. In
+  this paper, we focus on turning points associated to business cycle:
+  it is defined as the succession of phases of economic recession and
+  expansion, delimited by peaks (highest level of activity) and troughs
+  (lowest level of activity), see for example \textcite{ferrara} for a
+  description of the different economic cycles. In this paper, following
+  \textcite{Zellner1991}, we used an accepted definition for
+  turning-points:
 
@@
     a phase of expansion. This is the case at date \(t\) when
-    \(y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}\).\\
+    \(TC_{t-3}\geq TC_{t-2}\geq TC_{t-1}<TC_t\leq TC_{t+1}\).\\
   \item
@@
     to a phase of recession. This is the case at date \(t\) when
-    \(y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}\).\\
-    It thus takes at least 2 months to detect a turning point.\\
-    The phase shift is often defined as the number of months required to
-    detect the right turning point. Here we use a slightly modified
-    criterion: the phase shift is defined as the number of months needed
-    to detect the right turning point without any future revision.
-    Indeed, it can happen that the right turning point is detected by
+    \(TC_{t-3}\leq TC_{t-2}\leq TC_{t-1}>TC_t\geq TC_{t+1}\).
+  \end{itemize}
+
+  Lets denote \(TC_{t|t'}\) the estimate of the trend-cycle at date
+  \(t\) using the data up to date \(t'\). The phase shift is often
+  defined as the number of months required to detect the right turning
+  point: it would then be equal to two if, in the case of a downturn,
+  \(TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}\)
+  (two observations, at dates \(t\) and \(t+1\), are needed to qualify
+  the date \(t\) as turning point). Note that, in that case, the phase
+  shift could be equivalently set to 1 (only one observation after \(t\)
+  is needed to define \(t\) as a turning point), without any difference
+  on the comparison of the different methods.\\
+  In this paper we use a slightly modified criterion: the phase shift is
+  defined as the number of months needed to detect the right turning
+  point without any future revision. Thus, in the case of a downturn,
+  the phase shift is two if
+  \(TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}\)
+  and, for all dates \(t'\) higher than \(t+1\):
+  \(TC_{t-3|t'}\leq TC_{t-2|t'}\leq TC_{t-1|t'}>TC_{t|t'}\geq TC_{t+1|t'}\).
+  Since a 13-term Henderson filter is used for the final estimates of
+  the trend-cycle, the phase shift is at most height (the final estimate
+  of \(TC_{t+1}\) is compute a the date \(t+7\)). This definition is
+  used because it can happen that the right turning point is detected by
   asymmetric filters but is not detected by the final estimate using a
-    symmetric filter (this is the case for 41 reversal points out of the
-    9 series with asymmetric Musgrave filters) or that there are
-    revisions in successive estimates (this is the case for 7 turning
-    points out of the 9 series with asymmetric Musgrave filters).
-    Finally, relatively few turning points are detected at the right
-    date with the final estimate. With the 13-term Henderson filter, 18
-    are correctly detected in series with low variability (out of 57
-    possible), 11 in series with medium variability and 12 in series
-    with high variability.
-  \end{itemize}
+  symmetric filter (this is the case for 41 reversal points out of the 9
+  series with asymmetric Musgrave filters) or that there are revisions
+  in successive estimates (this is the case for 7 turning points out of
+  the 9 series with asymmetric Musgrave filters). Finally, relatively
+  few turning points are detected at the right date with the final
+  estimate. With the 13-term Henderson filter, 18 are correctly detected
+  in series with low variability (out of 57 possible), 11 in series with
+  medium variability and 12 in series with high variability.
 \item
@@
   \left|\frac{
-  y_{t|t+q} -  y_{t|last}
+  TC_{t|t+q} -  TC_{t|last}
   }{
-   y_{t|last}
+   TC_{t|last}
   }\right|
@@
   \left|\frac{
-  y_{t|t+q} - y_{t|t+q+1}
+  TC_{t|t+q} - TC_{t|t+q+1}
   }{
-  y_{t|t+q+1}
+  TC_{t|t+q+1}
   }\right|
@@
 \end{enumerate}
+
+The simulated series are of length 60 years, which is often not
+realistic for economic time series. However, since all the methods are
+local, the results are not affected by the length of the series. The
+length of the series would only have an impact for the identification
+and the estimation of the ARIMA model. In this case, since the same data
+generating process is used during the 60 years, it could be relevant to
+identify the ARIMA model using all the data. Even if it would improve
+the results in term of phase shift (see supplemental material) we prefer
+to only use the last 12 years to identify and estimate the ARIMA model,
+to be closer to what would have been done with a real time series.
 
@@
 
-\begin{figure}[H]
-
-\caption{\label{fig-graphstpsimul}Distribution of phase shifts on
+\begin{figure}
+
+\caption{\label{fig-graphstpsimul}Distribution of phase shift on
 simulated series.}
@@
 \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}
+
+Lecture: this plots represent box plots of the phase shift: the box
+represents the interquartile range, the central horizontal line the
+median the dot the outliers and the whisker the range of the data (if no
+outlier).\\
+For simulated series with medium variability (when a symmetric filter of
+length 13 is relevant), with the linear constant (LC) method, 50\% of
+the turning point are detected with a phase shift of at most 5 months
+whereas when the LC filter is parametrised locally (using), 50\% of the
+turning point are detected with a phase shift of at most 4 months. For
+the direct asymmetric filter (DAF), 25\% of the turning are detected
+with a phase shift of at least 7 months (compares to 5 for the LC
+method).
 
@@
 affect the orders of magnitude, which is why the results are presented
-only for series with medium variability (Table~\ref{tab-simulrev}). In
+only for series with medium variability (table~\ref{tbl-simulrev}). In
 general, LC filters always minimise revisions (with relatively small
@@
 between the fourth and fifth estimates (as might be expected, as
-highlighted in section~\ref{sec-mmetprev}).
-
-\begin{table}
-
-\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
-filters on simulated series with medium variability.}
-
-\centering{
-
-[!h]
-
-\centering
-\begin{tabular}{ccccccc}
+highlighted in section~\ref{sec-mmasym}).
+
+\begin{longtable}[t]{ccccccc}
+
+\caption{\label{tbl-simulrev}Average of the relative deviations of the
+revisions for the different filters on simulated series with medium
+variability.}
+
+\tabularnewline
+
 \toprule
-Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
+Method & \$q=0\$ & \$q=1\$ & \$q=2\$ & \$q=3\$ & \$q=4\$ & \$q=5\$\\
 \midrule
 \addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
+\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\right|\right]$}}\\
 \hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
@@
 \multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
-\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
+\left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\right|
 \right]$}}\\
@@
 \hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
+\hspace{1em}ARIMA & 0.23 & 0.10 & 0.02 & 0.03 & 0.25 & 0.01\\
 \bottomrule
-\end{tabular}
-
-}
-
-\end{table}%
+
+\end{longtable}
+
+Lecture: the estimates for \(q=0\) correspond to the first estimates of
+the trend-cycle (real-time estimates when 0 point in the future is
+known), \(q=1\) correspond to the second estimates of the trend-cycle
+(when 1 point in the future is known). Since the final estimates are
+obtain using a symmetric filter of bandwidth 6, \(q=6\) corresponds to
+the final estimates of the trend-cycle (when 6 points in the future are
+known). \(MAE_{fe}(0)\) compares the revision between the first estimate
+(\(q=0\)) and the last estimate (\(q=6\)) and \(MAE_{ce}(0)\) compares
+the revision between the first estimate (\(q=0\)) and the second
+estimate (\(q=1\)).
 
@@
 to January 2020, this series has a medium variability: a symmetric
-filter of 13 terms is therefore appropriate. Appendix
-\textbf{?@sec-ann-ex-r} shows how the code to reproduce the different
-plots.
+filter of 13 terms is therefore appropriate.
 
@@
 
-Graph + simulations
-
-Enfin, dans cet article et dans ceux associés aux méthodes étudiées, les
-filtres asymétriques sont appliqués et comparés sur des séries déjà
-désaisonnalisées ou sans saisonnalité (séries simulées). Les révisions
-et donc le déphasage sont limités à 6 mois (lorsque le filtre symétrique
-est de 13 termes) et cela a l'avantage d'isoler les impacts des
-différents filtres des autres processus inhérents à la
-désaisonnalisation. Il y a toutefois deux inconvénients à cette
-simplification :
+In this article and in those associated to trend-cycle extraction
+techniques, the asymmetrical filters are applied and compared on series
+already seasonally adjusted or without seasonality (simulated series).
+The revisions and therefore the phase shift are limited to 6 months
+(when the symmetrical filter is 13 terms), and this has the advantage of
+isolating the impacts of the different filters from the other processes
+inherent in seasonal adjustment. There are, however, two drawbacks to
+this simplification:
 
@@
 \item
-  D'une part, l'estimation de la série désaisonnalisée dépend de la
-  méthode utilisée pour extraire la tendance-cycle. Le choix de la
-  méthode utilisée pour l'estimation de la tendance-cycle peut donc
-  avoir un impact bien au-delà de 6 mois.
-\item
-  D'autre part, les moyennes mobiles étant des opérateurs linéaires, ils
-  sont sensibles à la présence de points atypiques. L'application
-  directe des méthodes peut donc conduire à des estimations biaisées, du
-  fait de leur présence, alors que les méthodes de désaisonnalisation
-  (comme la méthode X-13ARIMA-SEATS) ont un module de correction des
-  points atypiques. Par ailleurs, comme notamment montré par
-  \textcite{dagum1996new}, le filtre symétrique final utilisé par
-  X-13ARIMA pour extraire la tendance-cycle (et donc celui indirectement
-  utilisé lorsqu'on applique les méthodes sur les séries
-  désaisonnalisées) laisse passer environ 72 \% des cycles de 9 ou 10
-  mois (généralement associés à du bruit plutôt qu'à la tendance-cycle).
-  Les filtres asymétriques finaux amplifient même les cycles de 9 ou 10
-  mois. Cela peut avoir pour conséquence l'introduction d'ondulations
-  indésirables, c'est-à-dire la détection de faux points de
-  retournement. Ce problème est réduit par la correction des points
-  atypiques (ces cycles étant considérés comme de l'irrégulier). C'est
-  ainsi que le \emph{Nonlinear Dagum Filter} (NLDF) a été développé et
-  consiste à :
+  Firstly, the estimation of the seasonally-adjusted series depends on
+  the method used to extract the trend-cycle. The choice of method used
+  to estimate the trend-cycle can therefore have an impact well beyond 6
+  months.
+\item
+  On the other hand, as moving averages are linear operators, they are
+  sensitive to the presence of atypical points. Direct application of
+  the methods can therefore lead to biased estimates, due to their
+  presence, whereas seasonal adjustment methods (such as the
+  X-13ARIMA-SEATS method) have a correction module for atypical points.
+  Furthermore, as shown in particular by \textcite{dagum1996new}, the
+  final symmetrical filter used by X-13ARIMA to extract the trend-cycle
+  (and therefore the one indirectly used when applying the methods to
+  seasonally-adjusted series) misses around 72\% of 9- or 10-month
+  cycles (generally associated with noise rather than the trend-cycle).
+  The final asymmetrical filters even amplify 9 or 10-month cycles. This
+  can result in the introduction of undesirable ripples, i.e.~the
+  detection of false turning points. This problem is reduced by
+  correcting atypical points (these cycles being considered irregular).
+  This is how the \emph{Nonlinear Dagum Filter} (NLDF) was developed:
 
@@
   \item
-    appliquer l'algorithme de correction des points atypiques de
-    X-13ARIMA sur la série désaisonnalisée, puis de la prolonger par un
-    modèle ARIMA ;
-  \item
-    effectuer une nouvelle correction des points atypiques en utilisant
-    un seuil bien plus strict et appliquer ensuite le filtre symétrique
-    de 13 termes. En supposant une distribution normale cela revient à
-    modifier 48 \% des valeurs de l'irrégulier.
+    applying the X-13ARIMA atypical point correction algorithm to the
+    seasonally adjusted series, then extending it with an ARIMA model;
+  \item
+    perform a new atypical point correction using a much stricter
+    threshold and then apply the 13-term symmetrical filter. Assuming a
+    normal distribution, this amounts to modifying 48\% of the irregular
+    values.
   \end{enumerate}
 
-  Les \emph{cascade linear filter} (CLF), notamment étudiés dans
-  \textcite{dagumBianconcini2023}, correspondent à une approximation des
-  NLDF en utilisant un filtre de 13 termes et lorsque les prévisions
-  sont obtenus à partir d'un modèle ARIMA(0,1,1) où \(\theta=0,40.\)
+  The \emph{cascade linear filter} (CLF), studied in particular in
+  \textcite{dagumBianconcini2023}, corresponds to an approximation of
+  the NLDF using a 13-term filter and when the forecasts are obtained
+  from an ARIMA(0,1,1) model where \(\theta=0.40.\).
 \end{enumerate}
 
-Une piste d'étude serait alors d'étudier plus précisément l'effet des
-points atypiques sur l'estimation de la tendance-cycle et la détection
-des points de retournement, mais aussi d'explorer de nouveaux types de
-filtres asymétriques fondés sur des méthodes robustes (comme les
-régressions locales robustes, les médianes mobiles, etc.).
+One possible direction of study would then be to investigate more
+precisely the effect of atypical points on trend-cycle estimation and
+the detection of turning points, but also to explore new types of
+asymmetric filters based on robust methods (such as robust local
+regressions, moving medians, etc.).
+
+\begin{figure}[H]
+
+\caption{\label{fig-ce16ovcovid}Successive estimates of the trend-cycle
+in US employment (in logarithms) during the COVID-19.}
+
+\centering{
+
+\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_covid_lp.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\caption{\label{fig-ce16ov-previmp-covid}Implicit forecasts linked to
+successive estimates of the trend-cycle in US employment (in logarithms)
+using local polynomial methods.}
+
+\centering{
+
+\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_covid_prev_imp_lp.pdf}
+
+}
+
+\end{figure}%
 
@@
 
-\appendix
-\newpage
-
-\section{Coefficients, gain and phase shift functions}\label{sec-an-cof}
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-lc}Coefficients, gain and phase shift
-functions for the Linear-Constant (LC) filter with \(I/C=3.5\).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}
-
-}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-ql}Coefficients, gain and phase shift
-functions for the Quadratic-Linear (QL) filter with \(I/C=3.5\).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}
-
-}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-cq}Coefficients, gain and phase shift
-functions for the Cubic-Quadratic (CQ) filter with \(I/C=3.5\).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}
-
-}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-daf}Coefficients, gain and phase shift
-functions for the direct asymmetric filter (DAF).}
-
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}
-
-}
-
-\end{figure}%
-
 \end{document}

### 2024-04-25 paper/supp_material.qmd
update article

--- 2024-04-18 paper/supp_material.qmd
@@
 
+# Different specifications for the estimation of the slope and concavity
+
+::: {#fig-graphs-lc-deg-final}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_final_lc.pdf")
+```
+
+Distribution of phase shift associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
+:::
+
+::: {#fig-graphs-lc-deg}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_lc.pdf")
+```
+
+Distribution of phase shift associated to the local parametrisation of the linear-constant (LC) filter using the final estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
+:::
+
+::: {#fig-graphs-ql-deg-final}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_final_ql.pdf")
+```
+
+Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
+:::
+
+::: {#fig-graphs-ql-deg}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_local_ql.pdf")
+```
+
+Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
+:::
+
+\newpage
+
 # Impact of the span used to detect the ARIMA model
@@
 
-Distribution of phase shifts on simulated series by number of years used to identify and estimate the AIRMA model.
+Distribution of phase shift on simulated series by number of years used to identify and estimate the ARIMA model.
 :::
 
-# Quarterly data
-
-::: {#fig-graphs-impact-span}
-```{r, echo=FALSE, out.width="100%"}
-knitr::include_graphics("img/simulations/phase_shift_simul_q.pdf")
-```
-
-Distribution of phase shifts on quarterly data.
-:::
-
+\newpage
 # Nearest neighbour bandwidth
@@
 
-Distribution of phase shifts using nearest neighbour bandwidth.
+Distribution of phase shift using nearest neighbour bandwidth.
 :::
 
-::: {#fig-graphs-impact-span}
+::: {#fig-nn-localparam}
 ```{r, echo=FALSE, out.width="100%"}
-knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp.pdf")
+knitr::include_graphics("img/simulations/phase_shift_simul_nn_lp_localparam.pdf")
+```
+Note: nearest bandwidth moving average are used to estimate slope and concavity.
+We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.
+Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).
+
+Distribution of phase shift using nearest neighbour bandwidth for polynomial filters with local parametrisation.
+:::
+
+\newpage
+
+# Quarterly data
+
+::: {#fig-graphs-quarterly}
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics("img/simulations/phase_shift_simul_q.pdf")
 ```
 
-Distribution of phase shifts using nearest neighbour bandwidth for polynomial filters with local parametrisation.
+Distribution of phase shift on quarterly data.
 :::
-
-<!-- # R code example {#ann-ex-r} -->
-
-<!-- In this appendix we present the {{< fa brands r-project >}} code that can be used to estimate trend-cycle components for the US Employment with local polynomial filters. -->
-
-<!-- ```{r, eval=FALSE, echo=TRUE} -->
-<!-- #| eval: false -->
-<!-- #| echo: true -->
-<!-- #| file: R_ex.R -->
-<!-- ``` -->

### 2024-04-25 R_simul/4A_Analyse_tp_rev.R

@@ -1,0 +1,109 @@
+source("R_simul/4_utils.R",encoding = "UTF-8")
+library(ggplot2)
+library(scales)
+library(latex2exp)
+
+tp_lp <- merge(readRDS("results_simul/compile_tp/troughs_lp.RDS"),
+               readRDS("results_simul/compile_tp/peaks_lp.RDS"),
+               by=c("series","kernel", "method")) %>%
+  select_var()
+
+tp_lic_final <- merge(readRDS("results_simul/compile_tp/troughs_localic_final.RDS"),
+                      readRDS("results_simul/compile_tp/peaks_localic_final.RDS"),
+                      by=c("series", "kernel", "h", "degree", "method"))  %>%
+  # dplyr::filter() %>%
+  # mutate(method = ifelse(degree == "d2", h == "h6"))
+  select_var() %>%
+  mutate(method = sprintf("%s_localic_final_%s_%s", method, degree, h)) %>%
+  select(!c(degree, h))
+
+tp_lic_daf_trunc <- merge(
+  readRDS("results_simul/compile_tp/troughs_localic_daf_trunc.RDS"),
+  readRDS("results_simul/compile_tp/peaks_localic_daf_trunc.RDS"),
+  by=c("series", "kernel", "h", "degree", "method")) %>%
+  select_var() %>%
+  mutate(method = sprintf("%s_localic_%s_%s", method, degree, h)) %>%
+  select(!c(degree, h))
+
+
+tp_lp_nn <- merge(readRDS("results_simul/compile_tp/troughs_lp_nn.RDS"),
+                  readRDS("results_simul/compile_tp/peaks_lp_nn.RDS"),
+                  by=c("series","kernel", "method")) %>%
+  select_var() %>%
+  mutate(method = sprintf("%s_nn", method))
+tp_lic_final_nn <- merge(readRDS("results_simul/compile_tp/troughs_localic_final_nn.RDS"),
+                         readRDS("results_simul/compile_tp/peaks_localic_final_nn.RDS"),
+                         by=c("series", "kernel", "h", "degree", "method"))  %>%
+  dplyr::filter(degree == "d2", h == "h6") %>%
+  select_var() %>% mutate(method = sprintf("%s_localic_final", method)) %>%
+  select(!c(degree, h)) %>%
+  mutate(method = sprintf("%s_nn", method))
[…]

### 2024-04-29 paper/article_jos.qmd
V2.1

@@
 format:
+  docx:
+    toc: false
   pdf:
@@
     toc: false
+    fig-pos: 'H'
     include-in-header: [preamble.tex]
@@
 This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
-In particular, this theory allows us to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
-This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates (for the last estimates): first including a timeliness criterion to minimise the phase shift;
-second, with procedure for parametrising asymmetric filters locally While they are generally parametrised globally, which can be suboptimal around turning points.
+It enables to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
+This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates: first including a timeliness criterion to minimise the phase shift;
+second, with procedure for parametrising asymmetric filters locally while they are generally parametrised globally, which can be suboptimal around turning points.
 An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
-The paper also shows how these results can be easily reproduced using the R package `rjd3filters` dedicated to the manipulation of moving averages.
+The results are reproducible and all the methods can be easily applied using the R package `rjd3filters`.
 
@@
 
-Analysis of the economic cycle, and in particular the early detection of turning points in a series, is a major topic in the analysis of economic outlook.
-To this end, economic indicators are generally seasonally adjusted.
-However, in order to improve their interpretability, it may be necessary to perform additional smoothing in order to reduce noise, and thus analyse the trend-cycle component.
+Analysis of the economic cycle, and in particular the early detection of turning points, is a major topic in the analysis of economic outlook.
+To achieve this, economic indicators are generally seasonally adjusted.
+However, in order to improve their interpretability, it may be necessary to perform additional smoothing to reduce noise, thereby to analyse the trend-cycle component.
 By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series. 
+Thus, even though trend-cycle extraction methods are generally applied to seasonally adjusted series, the estimation of these series also depends on trend-cycle estimation methods.
 
 A moving average, or linear filter, is a statistical method that consists of applying a rolling weighted mean to a times series: for each date $t$ it computes a weighted mean of $p$ past points and $q$ future points where $p,q\geq0$ depends on the moving average.
-Moving averages are ubiquitous in business cycle extraction and seasonal adjustment methods.
-For example, the X-11 decomposition method [@x11], used in X-13ARIMA-SEATS software [@x13],  uses several moving averages to estimate the main components of a time series. 
-Symmetric filters are applied to the center of the series, but when it comes to estimate the most recent points, all of these methods must rely on asymmetric filters. 
-For trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.
+They are ubiquitous in business cycle extraction and seasonal adjustment methods.
+For example, the X-11 seasonal adjustment method [@ladiray2011seasonal], used in X-13ARIMA-SEATS software [@x13], uses several moving averages to estimate the main components of a time series. 
+Symmetric filters are applied to the centre of the series and for trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.
 
@@
 For example, for trend-cycle extraction, X-11 uses Henderson's symmetric filter and @musgrave1964set's asymmetric filters on an extended series using an ARIMA model.
-Since the predicted values are linear combinations of past values, this amounts to applying asymmetric averages at the end of the series.
+Since the predicted values are linear combinations of past values, this amounts to applying asymmetric moving averages at the end of the series.
 
@@
 Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
-In particular, it becomes possible to reproduce Henderson's symmetric filter and Musgrave's asymmetric filters.
-The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to two drawbacks of those methods.
-Firstly,  adding, in the optimisation process, a timeliness criterin to directly control the phase shift (i.e. the delay in detecting turning points).
+In particular, this class encompass Henderson's symmetric filter and Musgrave's asymmetric filters.
+The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods.
+Firstly,  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
 Secondly, calibrating asymmetric filter locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
-Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package (available at <https://github.com/rjdemetra/rjd3filters>) and can therefore be easily used.
+Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
 All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).
@@
 
-The basic assumption  used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and a an erratic component $\varepsilon_t$ (called irregular):
+The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
 $$
@@
 with $\xi_t$ a stochastic process uncorrelated with $\varepsilon_t$.
-Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], an usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
+Even if some papers model $\xi_t$ and $\varepsilon_t$ separately [see for example @GrayThomson2002], a usual assumption, adopted in this paper, is to is to gather $\xi_t$ and $\varepsilon_t$.
 Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
@@
 The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
-The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e.
-when no future point is known), using moving averages build on local polynomial smoothing techniques.
+The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.
 
@@
 
-- When $p=f$ the moving average is said to be *centered*. 
+- When $p=f$ the moving average is said to be *centred*. 
 If we also have $\forall k:\:\theta_{-k} = \theta_k$, the moving average $M_{\boldsymbol\theta}$ is said to be *symmetric*. 
@@
 For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
-In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e. movements recurring at least every 12 months.
+In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e., movements recurring at least every 12 months.
 Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
@@
 
-@Fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e.
-when no point in the future is known).
+@Fig-graphsmusgrave shows the gain and phase shift function for the asymmetric Musgrave filter (see @sec-lppasymf) often used for real-time trend-cycle estimation (i.e., when no point in the future is known).
 The gain function is greater than 1 on the frequencies associated with the trend-cycle ($[0, 2\pi/12]$): this means that the trend-cycle is well preserved and that short cycles of 1 to 2 years ($[2\pi/24,2\pi/12]$) are even amplified.
@@
 
-Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg). 
-In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.
+\justifying
+*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified of 10% ($G_{\boldsymbol\theta}(\pi/6)=1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*\
+*Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg).* 
+*In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.*
 
@@
 This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
-A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extend by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
+A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extended by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
 Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).
@@
 @Sec-sym-lp and @sec-lppasymf describes their approach to build symmetric moving averages (used for the final estimates of the trend-cycle component) and asymmetric moving averages (used for real-time estimates).
-This paper proposed two extensions of their approach to build asymetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parameterisation of a parameter usually parameterised globally (@sec-lptimeliness), which is used in the simulation.
+This paper proposed two extensions of their approach to build asymmetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parameterisation of a parameter usually parameterised globally (@sec-lptimeliness), which is used in the simulation.
 
@@
 $$
-y_t=\mu_t+\varepsilon_t,
-$$
-where $\mu_t$ is the trend and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
+y_t=TC_t+\varepsilon_t,
+$$
+where $TC_t$ is the trend-cycle and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
 Within a neighbourhood $h$ of $t$, the local trend $\mu_t$ is approximated by a polynomial of degree $d$, such that $\mu_t\simeq m_{t}$ with:
@@
 
-- the number of neighbors $2h+1$ (or the window $h$): if it is too small, then there will be too little data for the estimates (resulting in high variance in the estimates), and if it is too large, the polynomial approximation will probably be wrong, leading to biased estimates.
+- the number of neighbours $2h+1$ (or the window $h$): if it is too small, then there will be too little data for the estimates (resulting in high variance in the estimates), and if it is too large, the polynomial approximation will probably be wrong, leading to biased estimates.
 
@@
 where $\boldsymbol y_p$ is a matrix $(h+q+1)\times 1$ which denotes the set of available observations (from $y_{t-h}$ to $y_{t+q}$) and $\boldsymbol U$, $\boldsymbol Z$ and $\boldsymbol \theta$ are partitioned in the same way.
-The constraints are defined as $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$: they represents the polynomial preservation constraints we want to impose on the asymmetric filter.
+The constraints are defined as $\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta$: they represent the polynomial preservation constraints we want to impose on the asymmetric filter.
+$\boldsymbol Z$ is then associated to the polynomial constraints of the symmetric filter not imposed on asymmetric filter.
 @proietti2008 show that the problem is equivalent to finding $v$ which minimises:
@@
 
-- If the ratio is large ($3.5< R$) then a 23 term filter is used (to remove more noise).
-
-- If the ratio is small ($R<1$) a 9 term filter is used.
+- If the ratio is large ($3.5< R$) then a 23-term filter is used (to remove more noise).
+
+- If the ratio is small ($R<1$) a 9-term filter is used.
 
@@
 As mentioned in @sec-propMM, in the case of trend-cycle extraction we can set $\omega_1=0$ and for monthly series $\omega_2=2\pi/12.$
-For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show thant in that case the criteria can be written in a quadratic form:
+For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show than in that case the criteria can be written in a quadratic form:
 $$
@@
 Hence, in this paper we will not compare this extension to the other methods and we will focus on the local calibration of $\lvert\delta/\sigma\rvert$.
-However, in future studies, we could imagine a two step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parametrisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).
+However, in future studies, we could imagine a two-step calibration: fix the value of $\lvert\delta/\sigma\rvert$ to find $\alpha_T$ by cross-validation and then use this weight with a local parametrisation of the ratio $\lvert\delta/\sigma\rvert$ (as in @sec-localic).
 
@@
 
-The turning points are clearly detected by the local estimators (changement in the slope), when $|\delta_1/\sigma|$ is estimated in the real-time estimate.
-However, the real-time local estimates of $|\delta_1/\sigma|$ are noisy: the varability of this estimates, used to build asymmetric filters, will then lead to another source of revisions in the intermediate estimates.
+The turning points are clearly detected by the local estimators (slope tends towards zero) when $|\delta_1/\sigma|$ is estimated in the real-time.
+However, the real-time local estimates of $|\delta_1/\sigma|$ are noisy: the variability of these estimates, used to build asymmetric filters, will then lead to another source of revisions in the intermediate estimates.
 
@@
 
-There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series rjd3filters`.
+There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series using `rjd3filters`.
 
@@
 The different methods are compared on simulated and real data. 
-For all series, a symmetric 13-term filter is used: we then assume than the bandwidth is fixed, as in the seasonal adjustement method X-11.
-A nearest neighbour bandwidth is also tested (13-term asymmetric filters: real-time estimates uses observation between $y_{t-12}$ and $y_{t}$, the second asymmetric filters uses obversations between $y_{t-11}$ and $y_{t+1}$, etc.) but with no improvement in the results (see supplemental material).
+For all series, a symmetric 13-term filter is used: we then assume than the bandwidth is fixed, as in the seasonal adjustment method X-11.
+A nearest neighbour bandwidth is also tested (13-term asymmetric filters: real-time filter uses observations between $y_{t-12}$ and $y_{t}$, the second asymmetric filters uses observations between $y_{t-11}$ and $y_{t+1}$, etc.) but with no improvement in the results (see supplemental material).
 More complexe methods to locally select bandwidth, like the one proposed by @fan1992variable, were not tested.
@@
 
-- High signal-to-noise ratio (i.e. low I-C ratio and low variability): $\sigma_e^2=0.2$ and $\rho = 3.0, 3.5$ or $4.0$ (I-C ratio between 0.9 and 0.7);
-
-- Medium signal-to-noise ratio (i.e. medium I-C ratio and medium variability): $\sigma_e^2=0.3$ and $\rho = 1.5,\, 2.0$ or $3.0$ (I-C ratio between 2.3 and 1.4);
-
-- Low signal-to-noise ratio (i.e. high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).
+- High signal-to-noise ratio (i.e., low I-C ratio and low variability): $\sigma_e^2=0.2$ and $\rho = 3.0, 3.5$ or $4.0$ (I-C ratio between 0.9 and 0.7);
+
+- Medium signal-to-noise ratio (i.e., medium I-C ratio and medium variability): $\sigma_e^2=0.3$ and $\rho = 1.5,\, 2.0$ or $3.0$ (I-C ratio between 2.3 and 1.4);
+
+- Low signal-to-noise ratio (i.e., high I-C ratio and high variability): $\sigma_e^2=0.4$ and $\rho = 0.5,\, 0.7$ or $1.0$ (I-C ratio between 8.9 and 5.2).
 
@@
 
-   Lets denote $TC_{t|t'}$ the estimate of the trend-cycle at date $t$ using the data up to date $t'$.
+   Let's denote $TC_{t|t'}$ the estimate of the trend-cycle at date $t$ using the data up to date $t'$.
   The phase shift is often defined as the number of months required to detect the right turning point: it would then be equal to two if, in the case of a downturn,  $TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}$ (two observations, at dates $t$ and $t+1$, are needed to qualify the date $t$ as turning point).
@@
   Thus, in the case of a downturn, the phase shift is two if $TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}$ and, for all dates $t'$ higher than $t+1$: $TC_{t-3|t'}\leq TC_{t-2|t'}\leq TC_{t-1|t'}>TC_{t|t'}\geq TC_{t+1|t'}$.
-Since a 13-term Henderson filter is used for the final estimates of the trend-cycle, the phase shift is at most height (the final estimate of $TC_{t+1}$ is compute a the date $t+7$).
+Since a 13-term Henderson filter is used for the final estimates of the trend-cycle, the phase shift is at most height (the final estimate of $TC_{t+1}$ is compute at the date $t+7$).
   This definition is used because it can happen that the right turning point is detected by asymmetric filters but is not detected by the final estimate using a symmetric filter (this is the case for 41 reversal points out of the 9 series with asymmetric Musgrave filters) or that there are revisions in successive estimates (this is the case for 7 turning points out of the 9 series with asymmetric Musgrave filters).
@@
 ```
-Lecture: this plots represent box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).\
-For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally (using), 50% of the turning point are detected with a phase shift of at most 4 months.
-For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).
+
+\justifying
+*Lecture: this plots represent box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).*\
+*For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally (using), 50% of the turning point are detected with a phase shift of at most 4 months.*
+*For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).*
 
@@
         align = "c", booktabs = T, row.names = FALSE, 
-        escape = FALSE,) %>%  
-  kable_styling(latex_options=c(#"striped",  
-    "hold_position")) %>% 
+        escape = FALSE,format = "latex") %>%  
+  # kable_styling(latex_options=c(#"striped",  
+  #   "hold_position")) %>% 
   pack_rows(index = pack_row_index, escape = FALSE) # %>%
@@
 ```
-Lecture: the estimates for $q=0$ correspond to the first estimates of the trend-cycle (real-time estimates when 0 point in the future is known), $q=1$ correspond to the second estimates of the trend-cycle (when 1 point in the future is known).
-Since the final estimates are obtain using a symmetric filter of bandwidth 6, $q=6$ corresponds to the final estimates of the trend-cycle (when 6 points in the future are known).
-$MAE_{fe}(0)$ compares the revision between the first estimate ($q=0$) and the last estimate ($q=6$) and $MAE_{ce}(0)$ compares the revision between the first estimate ($q=0$) and the second estimate ($q=1$).
+
+\justifying
+*Lecture: the estimates for $q=0$ correspond to the first estimates of the trend-cycle (real-time estimates when 0 point in the future is known), $q=1$ correspond to the second estimates of the trend-cycle (when 1 point in the future is known).*
+*Since the final estimates are obtain using a symmetric filter of bandwidth 6, $q=6$ corresponds to the final estimates of the trend-cycle (when 6 points in the future are known).*
+*$MAE_{fe}(0)$ compares the revision between the first estimate ($q=0$) and the last estimate ($q=6$) and $MAE_{ce}(0)$ compares the revision between the first estimate ($q=0$) and the second estimate ($q=1$).*
 
@@
 The differences between the methods are also illustrated using an example from the FRED-MD database (@fredmd) containing economic series on the United States. 
+This database facilitates the reproducibility of the results thanks to the availability of series published on past dates.
 The series studied correspond to the database published in November 2022.
 The series studied is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
-This point and this series were chosen because the economic turnaround is particularly visible in the raw series (@fig-ce16ov-previmp-lp), because this series is used for dating economic cycles and because the FRED-MD database facilitates the reproducibility of the results (thanks to the availability of series published on past dates).
+This turning point was chosen because it is particularly visible in the raw series (@fig-ce16ov-previmp-lp) and this series was select because it is used for dating economic cycles.
 Studying the series up to January 2020, this series has a medium variability: a symmetric filter of 13 terms is therefore appropriate.
@@
 @fig-ce16ovlp shows the successive estimates of the trend-cycle using the different methods studied.
-
 For this series, the phase shift is 6 months for the LC and CQ methods and the extension of the series by ARIMA. 
@@
 
+::: {#fig-ce16ovlp}
 ```{r, out.width="95%", fig.show="hold"}
-#| label: fig-ce16ovlp
-#| fig-cap:  Successive estimates of the trend-cycle in US employment (in logarithms).
 series <- "CE16OV"
@@
 
+\justifying
+*Lecture: the dotted vertical line corresponds to the date of the turning point (February 2001).*
+*The curve "Feb 2001" corresponds to the estimates of the trend-cycle using the data observed until February 2001.*
+*Since a 13-term Henderson filter is used for the final estimates, for the curve "Feb 2001" estimates from September 2001 to February 2001 are intermediate estimates.*
+
+Successive estimates of the trend-cycle in US employment (in logarithms).
+:::
+
 The quality of the intermediate estimates can also be analysed using the implicit forecasts of the different methods (@fig-ce16ov-previmp-lp). 
@@
 
+::: {#fig-ce16ov-previmp-lp}
 ```{r, out.width="95%"}
-#| label: fig-ce16ov-previmp-lp
-#| fig-cap: Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
 series <- "CE16OV"
@@
 
+\justifying
+*Lecture: the dotted vertical line corresponds to the date of the turning point (February 2001).*
+*The curve "Mar 2001" corresponds to the implicit forecasts associated to the asymmetric filters used for the trend-cycle estimations using the data observed until March 2001.*
+
+Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
+:::
+
 ## Discussion
-### Outliers and COVID-19
-
-In this article and in those associated to trend-cycle extraction techniques, the asymmetrical filters are applied and compared on series already seasonally adjusted or without seasonality (simulated series).
-The revisions and therefore the phase shift are limited to 6 months (when the symmetrical filter is 13 terms), and this has the advantage of isolating the impacts of the different filters from the other processes inherent in seasonal adjustment.
+
+In the section we discuss the applicability of the presented methods to other frequencies (@sec-other-freq) and some limits of the presented methods (for example sensibility to outliers like COVID-19, @sec-out-covid) .
+
+### Other frequencies {#sec-other-freq}
+
+In this paper, we have focused on monthly series, but the methods presented can be applied to other frequencies.
+In general, for lower frequencies, the bandwidth of the symmetric filter reduced.
+For example, for quarterly series, X-11 uses a 5-term symmetric filter (in general) or 7-term symmetric filter for series with high variability.
+Mechanically, since fewer points are used, there will be less differences between methods.
+For example, on simulated series the results are similar in terms of phase shift but revisions are reduced with local parametrisation (see supplemental material).
+
+Since X-13ARIMA-SEATS can only be applied for series with frequency lower than monthly, for higher frequencies (weekly, daily, etc.), there is no reference on the length of the filters to uses.
+The length of the filters can be determined using local polynomial methods [see for example @Loader1999 for a description of bandwidth selection approaches] but it will surely be necessary to use higher bandwidth than for monthly series.
+The larger the window, the greater the bias in assuming that the trend is locally of degree 1: the LC method will then be certainly less appropriate than other methods (QL, CQ or DAF).
+Moreover, since the bias is higher, the local parametrisation of the filters will be more important.
+
+### Outliers and COVID-19 {#sec-out-covid}
+
+In this article and in those associated to trend-cycle extraction techniques, the moving averages are applied and compared on series already seasonally adjusted or without seasonality (simulated series).
+The revisions and the phase shift are limited to 8 months (when the symmetric filter is 13 terms), and this has the advantage of isolating the impacts of the different filters from the other processes inherent in seasonal adjustment.
 There are, however, two drawbacks to this simplification:
 
-1. Firstly, the estimation of the seasonally-adjusted series depends on the method used to extract the trend-cycle.
+1. The estimation of the seasonally-adjusted series depends on the method used to extract the trend-cycle.
 The choice of method used to estimate the trend-cycle can therefore have an impact well beyond 6 months.
 
-2. On the other hand, as moving averages are linear operators, they are sensitive to the presence of atypical points.
+2. As moving averages are linear operators, they are sensitive to the presence of atypical points.
 Direct application of the methods can therefore lead to biased estimates, due to their presence, whereas seasonal adjustment methods (such as the X-13ARIMA-SEATS method) have a correction module for atypical points.
-Furthermore, as shown in particular by @dagum1996new, the final symmetrical filter used by X-13ARIMA to extract the trend-cycle (and therefore the one indirectly used when applying the methods to seasonally-adjusted series) misses around 72% of 9- or 10-month cycles (generally associated with noise rather than the trend-cycle).
-The final asymmetrical filters even amplify 9 or 10-month cycles.
-This can result in the introduction of undesirable ripples, i.e. the detection of false turning points.
-This problem is reduced by correcting atypical points (these cycles being considered irregular). 
-This is how the *Nonlinear Dagum Filter* (NLDF) was developed:
-
-    a. applying the X-13ARIMA atypical point correction algorithm to the seasonally adjusted series, then extending it with an ARIMA model;
-
-    b. perform a new atypical point correction using a much stricter threshold and then apply the 13-term symmetrical filter. 
+Furthermore, as shown in particular by @dagum1996new, the final symmetric filter used by X-13ARIMA-SEATS to extract the trend-cycle (and therefore the one indirectly used when applying the methods to seasonally-adjusted series), cycles of length 9 and 10 months (generally associated with noise rather than the trend-cycle) are reduced by just 38%.
+The final asymmetric filters even amplify 9 or 10-month cycles.
+This can result in the introduction of undesirable ripples, i.e., the detection of false turning points.
+This problem is reduced by correcting atypical points and the *Nonlinear Dagum Filter* (NLDF) was defined in this perspective:
+
+    a. applying the X-11 atypical point correction algorithm [see for example @ladiray2011seasonal for a description] to the seasonally adjusted series, then extending it with an ARIMA model;
+
+    b. perform a new atypical point correction using a much stricter threshold and then apply the 13-term symmetric filter. 
     Assuming a normal distribution, this amounts to modifying 48% of the irregular values.
@@
 
-One possible direction of study would then be to investigate more precisely the effect of atypical points on trend-cycle estimation and the detection of turning points, but also to explore new types of asymmetric filters based on robust methods (such as robust local regressions, moving medians, etc.).
-
+[@Fig-ce16ovcovid; @fig-ce16ov-previmp-covid] show the successive estimates of the trend-cycle and the associated implicit forecast for the US employement during the COVID-19.
+First, we observe that while the real turning point is in April 2020, the 13-term Henderson filter produces final estimates of the trend-cycles, biased by the presence of a huge outlier, which lead to a turning point detected in June 2020.
+Second, as in 2001, the CQ and DAF methods lead to lots of variability in the intermediate estimates.
+For the QL method, the local parametrisation seems to slightly reduce the revisions.
+For the LC method, local parametrisation has few impacts on the estimates of the trend-cycle but the associated implicit forecasts are not plausible: the final estimate of the slope is biased by the outlier.
+Third, the ARIMA model can lead to unrealistic forecasts.
+
+<!-- One possible direction of study would then be to investigate more precisely the effect of atypical points on trend-cycle estimation and the detection of turning points, but also to explore new types of asymmetric filters based on robust methods (such as robust local regressions, moving medians, etc.).
+-->
+
+::: {#fig-ce16ovcovid}
 ```{r, out.width="95%", fig.show="hold"}
-#| label: fig-ce16ovcovid
-#| fig-cap:  Successive estimates of the trend-cycle in US employment (in logarithms) during the COVID-19.
 series <- "CE16OV"
@@
 
+\justifying
+*Lecture: the dotted vertical line corresponds to the date of the turning point (April 2020) but the detected one with the 13-term Henderson filter is June 2020.*
+*The curve "Apr 2020" corresponds to the estimates of the trend-cycle using the data observed until April 2020.*
+
+Successive estimates of the trend-cycle in US employment (in logarithms) during the COVID-19.
+:::
+
+::: {#fig-ce16ov-previmp-covid} 
 ```{r, out.width="95%"}
-#| label: fig-ce16ov-previmp-covid
-#| fig-cap: Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
 series <- "CE16OV"
@@
 
-### Other frequencies
+\justifying
+*Lecture: the curve "Apr 2020" corresponds to the implicit forecasts associated to the asymmetric filters used for the trend-cycle estimations using the data observed until April 2001.*
+
+Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods during the COVID-19.
+:::
+
+To sum up, trend-cycle estimates are biased by the presence of outliers which can also impact the detection of turning-points.
+Several approach could be used to limit their impact of to directly take them into account:
+
+- Adjustment of outliers prior to filtering, for example with a RegARIMA model or other correction modules like the one.
+
+- Increase the bandwidth of the filter used to extract the trend-cycle.
+
+- Since all the filter studied in this paper are equivalent to a local polynomial model, an external regressor can be added to the linear model to handle the outlier (the matrix $\boldsymbol X$ would then be modified).
+
+- Use of robust methods to estimate the trend-cycle, such as robust local regressions or moving medians.
 
@@
 Even when the phase shift is not reduced, local parametrisation is recommended because it reduces revisions and produces intermediate estimates that are more consistent with expected future trends.
-However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too short (i.e.
-if the variability of the series is "low"), retaining polynomial trends of degree one or less (LC method) produces poorer results in terms of detecting turning points.
+However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too short (i.e., if the variability of the series is "low"), retaining polynomial trends of degree one or less (LC method) produces poorer results in terms of detecting turning points.
 

### 2024-04-29 paper/resp_referee.qmd
update article
V2.1

@@
 
-*The modelled of the input time series is now describe more precisely in section 2 (previous section 1). $TC_t$ is here considered as deterministic*
+*The modelled of the input time series is now describe more precisely in section 2 (previous section 1).* 
+*$TC_t$ is here considered as deterministic*
 
@@
 *One explanation could be that the bias is higher with LC/QL methodes (we model linear trend using higher bandwidth) and for other methods (e.g.
-DAF) that the revisions to the final symmetric filters are higher;*
+DAF) that the revisions to the final symmetric filters are higher.*
 
@@
 $$
-\mathbb E[\hat \mu(t)] - \mu(t)=\sum_{j=d+1}^{+\infty}\frac{\mu^{(j)}(t)}{j!}\sum_{k=-h}^k\theta_k k^j
+\mathbb E[\hat \mu(t)] - \mu(t)=\sum_{j=d+1}^{+\infty}\frac{\mu^{(j)}(t)}{j!}\sum_{k=-h}^h\theta_k k^j
 $$
@@
 
-*More details are added to this section. The definition is more general than the polynomial ind*
+*More details were added to this section to clarify the definition: the forecasts are obtained so that the filtered extended series (input + forecast) with the symmetric filter is equal to the filtered extended series with the asymmetric filters.*
+*It doesn't correspond to the forecasts of the polynomial model.*
+*The definition is not restricted to moving averages build with local polynomial, the forecasts do not assumed polynomial trend model.*
 
@@
 
-*TODO*
+*A reading note is added to figure 5 to improve its reading*
 

### 2024-04-29 paper/supp_material.qmd
V2.1

@@
 \appendix
+
+```{r}
+#| include: false
+library(kableExtra)
+
+```
 
@@
 ```
-Note: nearest bandwidth moving average are used to estimate slope and concavity.
-We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.
-Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).
+
+\justifying
+*Note: nearest bandwidth moving average are used to estimate slope and concavity.*
+*We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.*
+*Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).*
 
@@
 :::
+
+::: {#tbl-simulrev-q}
+```{r}
+#| echo: false
+rev_table <- readRDS("data/simulations_revisions_q.RDS")
+pack_row_index = c("$MAE_{fe}(q) = \\mathbb E\\\\left[\\\\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\\\\right|\\\\right]$" = nrow(rev_table) / 2,
+                   "$MAE_{ce}(q)=\\mathbb E\\\\left[
+\\\\left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\\\\right|
+\\\\right]$" = nrow(rev_table) / 2)
+rev_table  %>%
+  kable(format.args = list(digits = 2,
+                           decimal.mark = "."),
+        align = "c", booktabs = T, row.names = FALSE, 
+        escape = FALSE,format = "latex") %>%  
+  # kable_styling(latex_options=c(#"striped",  
+  #   "hold_position")) %>% 
+  pack_rows(index = pack_row_index, escape = FALSE) # %>%
+    # add_footnote_kable(stop_centering = TRUE)
+```
+
+Average of the relative deviations of the revisions for the different filters on simulated quarterly series with medium variability.
+:::

### 2024-04-29 R_fredm/4A_Analyse_est.R

@@ -27,12 +27,17 @@ ggsave("paper/img/nber/ce16ov_fev2001_lp.pdf",
 
 all_plots <- get_all_plots(all_tp,
                            all_tp_rev,
-                           series = "CE16OV", tp_keep ="2020.25",
-                           nb_est = 8)
-wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey()
+                           series = "CE16OV", 
+                           tp_keep ="2020.25", tp_plot ="2020.25",
+                           nb_est = 8,
+                           vline = FALSE)
+p <- wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey() &
+  geom_vline(xintercept = detected_tp[detected_tp$series == series,paste0("X", "2020.25")], 
+             linetype = "dotted") &
+  labs(subtitle = NULL)
 
 ggsave("paper/img/nber/ce16ov_covid_lp.pdf",
-       plot = wrap_plots(all_plots[names(legende)], ncol = 3) & scale_color_grey(),
+       plot = p,
        width = 8, height = 8)
 
 data <- readRDS("data_fredm/byseries/RETAILx.RDS")[["2022"]]

### 2024-04-29 R_fredm/4B_Analyse_implicit_forecast.R

@@ -35,12 +35,15 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              series = series,
                              all_tp = all_tp,
                              all_tp_rev = all_tp_rev,
-                             tp_keep = tp_keep)
+                             tp_keep = tp_keep,
+                             vline = FALSE)
 wrap_plots(plots,ncol = 3) &
-  scale_color_grey()
+  scale_color_grey() &
+  labs(subtitle = NULL)
 ggsave("paper/img/nber/ce16ov_covid_prev_imp_lp.pdf",
        plot = wrap_plots(plots, ncol = 3) &
-         scale_color_grey(),
+         scale_color_grey() &
+         labs(subtitle = NULL),
        width = 8, height = 8)
 
 
@@ -53,7 +56,8 @@ plots <- get_all_plots_prevs(data_prevs = all_prevs,
                              series = series,
                              all_tp = all_tp,
                              all_tp_rev = all_tp_rev,
-                             tp_keep = tp_keep)
+                             tp_keep = tp_keep,
+                             vline = FALSE)
 wrap_plots(plots, ncol = 3)
 
 ggsave("paper/img/nber/retailx_nov2007_prev_imp_lp.pdf",

### 2024-04-29 R_fredm/4_utils.R

@@ -110,13 +110,13 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
                  by=c("series","kernel", "method")) %>%
     select_var() %>%
     dplyr::filter(kernel =="henderson")
-
+  
   tp_arima <-
     merge(readRDS(sprintf("%stroughs_arima.RDS", dossier)),
           readRDS(sprintf("%speaks_arima.RDS", dossier)),
           by=c("series","kernel", "method")) %>%
     select_var()
-
+  
   tp_lic_final <-
     merge(readRDS(sprintf("%stroughs_localic_final.RDS", dossier)),
           readRDS(sprintf("%speaks_localic_final.RDS", dossier)),
@@ -125,7 +125,7 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
     select_var() %>%
     mutate(method = sprintf("%s_localic_final_%s_%s", method, h, degree)) %>%
     select(!c(degree, h))
-
+  
   tp_lic_final_daf <-
     merge(readRDS(sprintf("%stroughs_localic_daf_trunc.RDS", dossier)),
           readRDS(sprintf("%speaks_localic_daf_trunc.RDS", dossier)),
@@ -134,7 +134,7 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
     select_var() %>%
     mutate(method = sprintf("%s_localic_%s_%s", method, h, degree)) %>%
     select(!c(degree, h))
-
+  
   order_methods  <- c(
     "lc",
     "lc_localic_final_h6_d2",
@@ -155,7 +155,7 @@ get_all_tp <- function(dossier = "results_fredm/compile_tp_norev/") {
     "cq",
     "daf",
     "auto_arima"
-    )
+  )
   all_tp <-
     rbind(tp_lp,
           tp_arima,
@@ -178,11 +178,11 @@ get_all_prevs <- function(series, tp_keep, nb_est = 10, nb_dates_before = 6,
   data <- data[names(data) >= tp_keep][1:nb_est]
   data_info <- readRDS(sub("byseries", "byseriesinfo", s))
   data_info <- data_info[names(data_info) >= tp_keep][1:nb_est]
-
+  
   data_merge <- do.call(ts.union, data)
   data_merge <- window(data_merge, start = tp_date - nb_dates_before/frequency(data_merge))
[…]

### 2024-04-29 R_simul/4A_Analyse_tp.R

@@ -56,21 +56,20 @@ tp_ner_neigh_hend <-
 
 
 order_methods  <- c(
-  "lc", "lc_nn",
+  "lc", 
   "lc_localic_final_d2_h3", "lc_localic_final_d2_h4", "lc_localic_final_d2_h5", "lc_localic_final_d2_h6",
   "lc_localic_final_d3_h3", "lc_localic_final_d3_h4", "lc_localic_final_d3_h5",
-  "lc_localic_final_d3_h6","lc_localic_final_nn",
+  "lc_localic_final_d3_h6",
   "lc_localic_d2_h3", "lc_localic_d2_h4", "lc_localic_d2_h5",
   "lc_localic_d2_h6", "lc_localic_d3_h3", "lc_localic_d3_h4", "lc_localic_d3_h5",
-  "lc_localic_d3_h6",
-  "ql", "ql_nn",
+  "lc_localic_d3_h6","lc_nn", "lc_localic_final_nn",
+  "ql", 
   "ql_localic_final_d2_h3", "ql_localic_final_d2_h4", "ql_localic_final_d2_h5", "ql_localic_final_d2_h6",
   "ql_localic_final_d3_h3", "ql_localic_final_d3_h4", "ql_localic_final_d3_h5",
   "ql_localic_final_d3_h6",
-  "ql_localic_final_nn",
   "ql_localic_d2_h3", "ql_localic_d2_h4", "ql_localic_d2_h5",
   "ql_localic_d2_h6", "ql_localic_d3_h3", "ql_localic_d3_h4", "ql_localic_d3_h5",
-  "ql_localic_d3_h6",
+  "ql_localic_d3_h6","ql_nn","ql_localic_final_nn",
   "cq", "cq_nn",
   "daf", "daf_nn",
   "auto_arima_ny2", "auto_arima_ny4", "auto_arima_ny6",

### 2024-04-29 R_simul_quarter/4A_Analyse_tp.R

@@ -65,13 +65,9 @@ legende <- c(lc = "LC", ql = "QL",
              lc_localic = "LC loc.\nparam.",
              ql_localic_final = "QL loc. param.\n(final estimates)",
              ql_localic = "QL loc.\nparam.",
-             auto_arima_ny8 = "ARIMA")
-all_tp %>%
-  filter(method %in%
-           names(legende),
-         kernel == "henderson") |>
-  pull(method) |>
-  unique()
+             auto_arima_ny12 = "ARIMA")
+
+
 p <- ggplot(all_tp %>%
               filter(method %in%
                        names(legende),

### 2024-04-29 R_simul_quarter/4B_analyse_revisions.R

@@ -21,7 +21,7 @@ for (crit in c("ce", "fe")) {
         select(!c(degree, h))
     }
     if (method == "arima") {
-      data <- data  %>% dplyr::filter(ny == "All") %>% 
+      data <- data  %>% dplyr::filter(ny == "12") %>% 
         select(!c(ny))
     }
     data <- data %>% select(!kernel)

### 2024-04-29 R_simul_quarter/4_utils.R

@@ -43,7 +43,7 @@ format_table_tp <- function(x, kernel = "henderson"){
       names_to = "name",
       values_to = "value"
     )%>% dplyr::filter(kernel %in% kernel) %>%
-    unique_series_pivot() %>%
+    # unique_series_pivot() %>%
     mutate(variability = recode(variability,
                                 lowvariability = "Low variability",
                                 mediumvariability = "Medium  variability",
@@ -76,6 +76,6 @@ summarise_ref <- function(x, normalise = FALSE){
       .fns = list(Mean = \(x) round(mean(x),digits)),
       .names = "{col}"
     )) %>%
-    select(!c(rev.q3:rev.q10, length)) %>%
+    select(!c(rev.q2:rev.q10, length)) %>%
     data.frame()
 }

### 2024-04-29 R_filters/plot_coef.R
grey musgrave plot

@@ -47,8 +47,8 @@ p <- (ggplot_coef(x, q=0) + ggplot2::guides(colour = "none") )/ (
     ggplot_phase(x, xlim = c(0, pi), q = 0)+
     ggplot2::labs(y = "Phase shift") + ggplot2::guides(colour = "none")
 )
+p <- p & scale_color_grey()
 p
-
 ggsave("paper/img/filters_used/musgrave.pdf",
        width = 8, height = 4.5,
        plot = p)

### 2024-04-29 paper/article_jos.qmd
fignote musgrave

@@
 \justifying
-*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified of 10% ($G_{\boldsymbol\theta}(\pi/6)=1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*\
+*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified by 10% ($G_{\boldsymbol\theta}(\pi/6)\simeq1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*
+*8-month cycles (associated to the irregular) are only reduced by 6% ($G_{\boldsymbol\theta}(\pi/4)\simeq0.94$).*\
 *Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg).* 

### 2024-04-30 paper/article_jos.qmd
v1

--- 2024-04-29 paper/article_jos.qmd
@@
 
+lang: en-GB  
 linestretch: 2
@@
 :::: {.abstract}
-This paper studies and compares real-time estimates of the trend-cycle component with moving averages based on local polynomial regression.
+This paper examines and compares real-time estimates of the trend-cycle component using moving averages constructed with local polynomial regression.
 It enables to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
 This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates: first including a timeliness criterion to minimise the phase shift;
-second, with procedure for parametrising asymmetric filters locally while they are generally parametrised globally, which can be suboptimal around turning points.
+second with procedure for parametrising asymmetric filters locally while they are generally parametrised globally, which can be suboptimal around turning points.
 An empirical comparison, based on simulated and real data, shows that modelling polynomial trends that are too complex introduces more revisions without reducing the phase shift, and that local parametrisation reduces the delay in detecting turning points and reduces revisions.
@@
 By construction, trend-cycle extraction methods are closely related to seasonal adjustment methods, since they are generally applied to seasonally adjusted series.
-Thus, even though trend-cycle extraction methods are generally applied to seasonally adjusted series, the estimation of these series also depends on trend-cycle estimation methods.
 
@@
 Symmetric filters are applied to the centre of the series and for trend-cycle extraction, the most popular symmetric filter is the @henderson1916note moving average, which is used in the X-11 seasonal adjustment algorithm.
-
 However, for real-time estimates, due to the lack of future observations, all these methods must rely on asymmetric filters to estimate the most recent points. 
@@
 Since the predicted values are linear combinations of past values, this amounts to applying asymmetric moving averages at the end of the series.
-
-If the classic asymmetric moving averages have good properties regarding the future revisions induced by the process (see for example @pierce1980SA), they create, by construction, phase shifts that impact the real-time estimation of turning points, introducing time delay in the detection.
+If the classic asymmetric moving averages have good properties regarding the future revisions induced by the process (for example @pierce1980SA shows similar performances in terms of revisions of X-11 compared to theoretically superior procedure based on ARIMA forecast models), they create, by construction, phase shifts that impact the real-time estimation of turning points, introducing time delay in the detection.
 
@@
 The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods.
-Firstly,  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
-Secondly, calibrating asymmetric filter locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
+Firstly  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
+Secondly calibrating asymmetric filters locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
 Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
@@
 
-The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$ can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
+The basic assumption used in  time series decomposition methods is that the input observed time series, $y_t$, can be decomposed into a signal component $\mu_t$ and an erratic component $\varepsilon_t$ (called irregular):
 $$
@@
 The noise component $\varepsilon_t$ is usually assumed to be a white noise.
-Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend and cycle components, here estimated jointly and called trend-cycle $TC_t$.
-This component can be locally approximated by a polynomial degree (see @sec-nonparamreg):
+Assuming that the input time series is seasonally adjusted (or without seasonality), the signal $\mu_t$ represents the trend (sustained and systematic variations over a long period of time) and cycle (cyclical movements superimposed upon the long-term trend) components, here estimated jointly and called trend-cycle $TC_t$.
+This component can be locally approximated by a polynomial degree (@sec-nonparamreg):
 $$
@@
 The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
-The estimate $\hat \beta_0$ provides the trend-cycle estimate $\hat TC_t$ and it can be shown that this equivalent to apply of moving average (see @sec-nonparamreg).
+The estimate $\hat \beta_0$ provides the trend-cycle estimate $\widehat{TC}_t$ and it can be shown that this equivalent to apply of moving average (@sec-nonparamreg).
 The goal of this paper is to present two new methods to separate the trend-cycle component from the noise component in real-time (i.e., when no future point is known), using moving averages build on local polynomial smoothing techniques.
@@
 
-Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ or $M$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $y_t$:
+Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $y_t$:
 $$
@@
 where $\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert$ is the *gain* or *amplitude* function and $\varphi_{\boldsymbol\theta}(\omega)$ is the *phase shift* or *time shift* function.
-We this have:
+We thus have:
 \begin{align*}
@@
 
-Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
-Noting $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
+Fourier decomposition allows us to analyse any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
+With $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
 For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
@@
 \justifying
-*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified of 10% ($G_{\boldsymbol\theta}(\pi/6)=1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*\
-*Note: the Musgrave filter depends on a ratio $I/C$ (see @sec-nonparamreg).* 
+*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified by 10% ($G_{\boldsymbol\theta}(\pi/6)\simeq1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*
+*8-month cycles (associated to the irregular) are only reduced by 6% ($G_{\boldsymbol\theta}(\pi/4)\simeq0.94$).*\
+*Note: the Musgrave filter depends on a ratio denoted $I/C$ (see @sec-nonparamreg).* 
 *In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.*
 
-Coefficients, gain and phase shift function for the Musgrave filter used in X-11 in real time (last estimate) for monthly series when the bandwidth is 6.
+Coefficients, gain and phase shift function for the Musgrave filter used in X-11 in real-time (last estimate) for monthly series when the bandwidth is 6.
 :::
@@
 Trend-cycles are generally modelled by local polynomial trends (see @sec-nonparamreg) and, in order to best preserve the trend-cycles, we thus design moving averages that preserve the polynomial trends.
-A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t:M_{\boldsymbol\theta} f(t)=f(t)$.
-
-For the moving average $M_{\boldsymbol\theta}$, it can be shown that:
+A moving average $M_{\boldsymbol\theta}$ preserves a function of time $f(t)$ if $\forall t,M_{\boldsymbol\theta} f(t)=f(t)$.
+
+For a moving average $M_{\boldsymbol\theta}$, it can be shown that:
 
@@
 \sum_{k=-p}^{+f}\theta_k=1
- \text{ and }
-\forall j \in \left\{1,2,\dots,d\right\}:\:
+ \text{(preserves constant) and }
+\forall j \in \left\{1,2,\dots,d\right\},\:
 \sum_{k=-p}^{+f}k^j\theta_k=0.
@@
 
-2. Apply symmetric filters to extended forecast series.
+2. Apply symmetric filters to an extended forecast series (which is equivalent to using asymmetric moving averages, since forecasts are linear combinations of the past).
 This method seems to date back to @deforest1877adjustment, which also suggests modelling a polynomial trend of degree three or less at the end of the period.
@@
 
-Ultimately, the second method is equivalent to using asymmetric moving averages, since forecasts are linear combinations of the past.
-
-Conversely, the implicit forecasts of an asymmetric moving average can be deduced from a reference symmetric moving average. 
+Conversely, the *implicit forecasts* of an asymmetric moving average can be deduced from a reference symmetric moving average. 
 This allows us to judge the quality of real-time estimates of the trend-cycle and to anticipate future revisions when forecasts are far from expected evolutions.\
 Let us denote $\boldsymbol v=(v_{-h},\dots, v_{h})$ the reference symmetric moving average (used to compute the final estimates) and $\boldsymbol w^0,\dots, \boldsymbol w^{h-1}$ a sequence of asymmetric moving averages used for the intermediate estimates.
-$\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)$ is used for estimation in real time estimate (when 0 point in the future is known), $\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)$ is used for estimation of the penultimate point (when only 1 in the future is known), and so on.
-Let $y_{t-h},\dots,y_{t}$ be the observed series during the last $h$ periods (for which asymmetric moving averages are used to get the intermediate trend-cycle estimates for the dates $t-h$ to $h$). 
-The implicit forecast $y_{t+1}^*,\dots y_{t+h}^*$ induced by $w^0,\dots w^{h-1}$ are the forecasts of the series $y_t$ for which the estimates of the symmetric filter applied to the extended series $\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)$ produce the same estimates as the asymmetric filter applied to the series $\boldsymbol y^*$.
+$\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)$ is used for estimation in real-time estimate (when 0 point in the future is known), $\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)$ is used for estimation of the penultimate point (when only 1 in the future is known), and so on.
+Let $y_{t-h},\dots,y_{t}$ be the observed series during the last $h$ periods (for which asymmetric moving averages are used to get the intermediate trend-cycle estimates for the dates $t-h$ to $t$). 
+The *implicit forecast* $y_{t+1}^*,\dots y_{t+h}^*$ induced by $w^0,\dots w^{h-1}$ are the forecasts of the series $y_t$ for which the estimates of the symmetric filter applied to the extended series $\boldsymbol y^* = (y_{t-h}, \dots, y_t, y_{t+1}^*,\dots y_{t+h}^*)$ produce the same estimates as the asymmetric filter applied to the series $\boldsymbol y^*$.
 This means:
 \begin{align*}
-\underbrace{\sum_{i=-h}^0 v_iy_{ti} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
-\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ points after }t} \\
+\underbrace{\sum_{i=-h}^0 v_iy_{t+i} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
+\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ point after }t} \\
 &=
-\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ points after }t} \\
-&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
+\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ point after }t} \\
+&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}^{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
 &= \cdots \\
@@
 $$
-\forall t \::\:\underbrace{\mu(t)}_{=TC_t} = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
+\forall t,\:\mu(t) = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
 \frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
@@
 In practice, this means assuming that the trend is locally polynomial.
-We also generally assume that the bandwith is fixed over time: $h(t_0)=h.$
+We also generally assume that the bandwidth is fixed over time: $h(t_0)=h.$
 This is also what is assumed in the seasonal adjustment algorithm X-11 and what will be assumed in this paper.
@@
 Various estimation methods can be used to derive symmetric and asymmetric moving averages.
-@GrayThomson1996 propose a complete statistical framework that makes it possible, in particular, to model the error in approximating the trend by local polynomials.
+For example, @GrayThomson1996 propose a complete statistical framework that makes it possible, in particular, to model the error in approximating the trend by local polynomials.
 However, as the specification of this error is generally complex, simpler models may be preferred, such as that of @proietti2008.
-Finally, @dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
-See @inseeDTM202401 for a more detailed comparison of the different recent methods for trend-cycle extraction included in the `rjd3filters` package
+@dagumbianconcini2008 propose a similar modelling of the trend-cycle but using the theory of Hilbert spaces with reproducing kernels for estimation, which has the particular advantage of facilitating the calculation of different moving averages at different time frequencies.
+See @inseeDTM202401 for a more detailed comparison of the different recent methods for trend-cycle extraction included in the `rjd3filters` package.
 
@@
 @Sec-sym-lp and @sec-lppasymf describes their approach to build symmetric moving averages (used for the final estimates of the trend-cycle component) and asymmetric moving averages (used for real-time estimates).
-This paper proposed two extensions of their approach to build asymmetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parameterisation of a parameter usually parameterised globally (@sec-lptimeliness), which is used in the simulation.
+This paper proposed two extensions of their approach to build asymmetric moving averages: a first one adding a criteria to control the phase shift (@sec-localic), not used in the simulation (@sec-comparison) but which could be used in future studied; a second one using a local parametrisation of a parameter usually parametrised globally (@sec-lptimeliness), which is used in the simulation.
 
@@
 
-Using @proietti2008's notation, we assume that our time series $y_t$ can be decomposed into
+Using @proietti2008's framework, we assume that our time series $y_t$ can be decomposed into
 $$
@@
 where $TC_t$ is the trend-cycle and $\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})$ is the noise (the series is therefore seasonally adjusted).
-Within a neighbourhood $h$ of $t$, the local trend $\mu_t$ is approximated by a polynomial of degree $d$, such that $\mu_t\simeq m_{t}$ with:
-$$
-\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
+Within a neighbourhood $h$ of $t$, the local trend $TC_t$ is approximated by a polynomial of degree $d$, such that $TC_t\simeq m_{t}$ with:
+$$
+\forall j\in\left\{-h,-h+1,\dots,h\right\},\:
 y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
 $$
-The problem of trend extraction is equivalent to estimating $m_t=\beta_0$ (the constant in the previous formula).
+The problem of trend extraction is equivalent to estimating $m_t=TC_t=\beta_0$ (the constant in the previous formula).
 
@@
 $$
-\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
+\hat{m}_{t}=\widehat{TC}_t=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
 $$ {#eq-mmsym}
-To conclude, the estimate of the trend $\hat{m}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
+To conclude, the estimate of the trend $\widehat{TC}_{t}$ can be obtained applying the symmetric filter $\boldsymbol \theta$ to $y_t$ ($\boldsymbol \theta$ is symmetric due to the symmetry of the kernel weights $\kappa_j$).
 Moreover, since $\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}$ we have:
@@
 See, for example, @cleveland1996smoothing or @Loader1999.
-The only desired constraints on the kernel are that it assigns greater weight to the central estimation ($\kappa_0$) and decreases towards 0 as it moves away from the central estimation.
+The only desired constraints on the kernel are that it assigns greater weight to the central estimation ($\kappa_0$) and decreases towards 0 as it moves away from the central observation.
 The uniform kernel should therefore be avoided.
@@
 1. *Linear-Constant* (LC): $y_t$ linear ($d=1$) and $\boldsymbol v$ preserves constant signals ($d^*=0$). 
-We obtain Musgrave filters when the Henderson kernel is used and the Henderson filter is used as the symmetric filter.
+We obtain Musgrave filters when the Henderson filter is used for the symmetric filter.
 
@@
 
-To measure the phase shift between the input series and the filtered one, @eq-lppasym introduces a criterion, $T_g$, called *timeliness*.
+To measure the phase shift between the input series and the filtered one, @ch15HBSA introduces a criterion, $T_g$, called *timeliness*.
 When a moving average $\boldsymbol v$ is applied, the level input signal is also altered by the gain function. 
@@
 As mentioned in @sec-propMM, in the case of trend-cycle extraction we can set $\omega_1=0$ and for monthly series $\omega_2=2\pi/12.$
-For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show than in that case the criteria can be written in a quadratic form:
+For the penalty function, the authors suggest $f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2$ and show that in that case the criteria can be written in a quadratic form:
 $$
@@
 $$
-where $\alpha_T$ is the weight associated with the timeliness criterion, $\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p} + \alpha_T\boldsymbol T$, $\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}-\alpha_T\boldsymbol \theta_p\boldsymbol T$ and $\widetilde {\boldsymbol c}$ a constant independent of $\boldsymbol v.$
+where $\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p} + \alpha_T\boldsymbol T$, $\alpha_T$ is the weight associated with the timeliness criterion, $\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}-\alpha_T\boldsymbol \theta_p\boldsymbol T$ and $\widetilde {\boldsymbol c}$ a constant independent of $\boldsymbol v.$
 
 With $\alpha_T=0$ we find $\varphi(\boldsymbol v)$.
-This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section but also generalizes the approach of @GrayThomson1996.
-This extension is implemented in the R function `rjd3filters::lp_filter()`.
+This extension therefore makes it possible to find all the symmetric and asymmetric filters presented in the previous section.
+It is implemented in the R function `rjd3filters::lp_filter()`.
 
@@
 Asymmetric filters are usually parametrised globally: $\lvert\delta/\sigma\rvert$ is fixed using all the data (with the IC-ratio or a cross-validation criterion). 
-However, we might prefer a local parametrisation: ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
+However, we might prefer a local parametrisation: a ratio $\lvert\delta/\sigma\rvert$ which varies as a function of time.
 Indeed, although the overall parametrisation is generally valid, assuming a constant value of the $\lvert\delta/\sigma\rvert$ ratio for all asymmetric filters does not appear relevant for real-time estimation, especially during periods of economic downturn.
@@
 This is what is proposed in this article, with a local parametrisation of the asymmetric filters by estimating $\delta$ and $\sigma^2$ separately.
-Although this does not give an unbiased estimator of the ratio $\lvert\delta/\sigma\rvert$, it does allow the main evolutions to be captured, such as the decay towards 0 before a turning point and the growth after the turning point for the LC method:
+Although this does not give an unbiased estimator of the ratio $\lvert\delta/\sigma\rvert$, it does allow the main evolutions to be captured, such as the decay towards 0 before a turning point and the growth after the turning point for the LC method.
+$\sigma^2$ and $\delta$ are estimated as follows:
 
@@
 $\nu_1$ and $\nu_2$ are two definitions of degrees of freedom of a local fit (generalization of the number of parameters of a parametric model).
-In the case of local polynomial smoothings, they can be linked to the coefficient of the associated filter: $\nu_1 = (n-h)\theta_0$ (number of observations multiplied by the central coefficient of the moving average) and $\nu_2 = (n-h)\sum_{i=-h}^{h} \theta_i^2$ (number of observations multiplied by the sum of squares of coefficients of the moving average).
+In the case of local polynomial smoothing, they can be linked to the coefficient of the associated filter: $\nu_1 = (n-2h)\theta_0$ (number of observations multiplied by the central coefficient of the moving average) and $\nu_2 = (n-2h)\sum_{i=-h}^{h} \theta_i^2$ (number of observations multiplied by the sum of squares of coefficients of the moving average).
 Thus, the variance can be estimated by:
@@
 This estimator assumes that $\widehat{TC}_t$ is unbiased, so variance estimates intervals are usually computed at small bandwidths ($h$).
-In this paper we study monthly series with small bandwidth ($h=6$), and the variance is estimated using local polynomial of degree 3 (Henderson filter), the biased of $\widehat{TC}_t$ is thus already small and using smaller bandwidths have almost no impact on the estimates (see Supplemental materials).
+In this paper we study monthly series with small bandwidth ($h=6$), and the variance is estimated using local polynomial of degree 3 (Henderson filter), the biased of $\widehat{TC}_t$ is thus already small and using smaller bandwidths have almost no impact on the estimates (see supplemental materials).
 
@@
 The DAF method then simply allows us to calculate the associated asymmetric moving averages.  
-Although a moving average of different length to that used for estimating the trend could be considered, this seems to degrade the results in terms of phase shift (using the same methodology as in @sec-comparison).
-Furthermore, for the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
-In this paper, we have chosen to model a trend of degree 2: this slightly reduces the phase shift (see Supplemental material) but slightly increases the revisions linked to the first estimate of the trend-cycle.
-@fig-mmpenteconcac shows the moving averages used.
+For the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
+In this paper, we have chosen to model a trend of degree 2: this slightly reduces the phase shift (see supplemental material) but slightly increases the revisions linked to the first estimate of the trend-cycle.
+@Fig-mmpenteconcac shows the moving averages used.
 
@@
 
-To distinguish the gain from the local parametrisation of the noise biased associated to the real-time estimates of the slope and concavity, in the empirical applications of @sec-comparison we also compares the results to the final local parametrisation obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
-@fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two local parametrisations:
+To distinguish the gain from the local parametrisation of the noise associated to the real-time estimates of the slope and concavity, in the empirical applications of @sec-comparison we also compares the results to the final local parametrisation obtained by estimating $\delta$ using all the data (i.e., using the symmetric filters shown in @fig-mmpenteconcac), while maintaining a real-time estimate of $\sigma^2$.
+@Fig-mmpenteconcac-ex shows an example of the comparison of estimates of the ratio $|\delta_1/\sigma|$ with a global estimator (IC ratio) and two local parametrisations:
 
@@
 A nearest neighbour bandwidth is also tested (13-term asymmetric filters: real-time filter uses observations between $y_{t-12}$ and $y_{t}$, the second asymmetric filters uses observations between $y_{t-11}$ and $y_{t+1}$, etc.) but with no improvement in the results (see supplemental material).
-More complexe methods to locally select bandwidth, like the one proposed by @fan1992variable, were not tested.
+More complex methods to locally select bandwidth, like the one proposed by @fan1992variable, were not tested.
 All the methods to create asymmetric filters are also compared with estimates obtained by extending the series with an ARIMA model, then applying a 13-term symmetric Henderson filter.
 The ARIMA model is determined automatically, using @autoarima algorithm (function `forecast::auto.arima()`) on the last 12 years, with no seasonal lag (the series being seasonally adjusted) and no external variables (such as external regressors for correction of atypical points).
-
 The performance of the different methods is judged on criteria relating to revisions (between two consecutive estimates and in relation to the final estimate) and the number of periods required to detect turning points.
@@
 
-For each series and each date, the trend-cycle is estimated using the different methods presented in this report.
-Three quality criteria are also calculated:
-
-1. Computation of the phase shift in the detection of turning points. 
+For each series and each date, the trend-cycle is estimated using the different methods presented in this paper
+Three quality criteria are computed:
+
+1. The *phase shift* in the detection of turning points. 
 In this paper, we focus on turning points associated to business cycle: it is defined as the succession of phases of economic recession and expansion, delimited by peaks (highest level of activity) and troughs (lowest level of activity), see for example @ferrara for a description of the different economic cycles.
@@
   Note that, in that case, the phase shift could be equivalently set to 1 (only one observation after $t$ is needed to define $t$ as a turning point), without any difference on the comparison of the different methods.\
-  In this paper we use a slightly modified criterion: the phase shift is defined as the number of months needed to detect the right turning point without any future revision.
+  In this paper we use a slightly modified criterion: the *phase shift* is defined as the number of months needed to detect the right turning point without any future revision.
   Thus, in the case of a downturn, the phase shift is two if $TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}$ and, for all dates $t'$ higher than $t+1$: $TC_{t-3|t'}\leq TC_{t-2|t'}\leq TC_{t-1|t'}>TC_{t|t'}\geq TC_{t+1|t'}$.
@@
 In this case, since the same data generating process is used during the 60 years, it could be relevant to identify the ARIMA model using all the data.
-Even if it would improve the results in term of phase shift (see supplemental material) we prefer to only use the last 12 years to identify and estimate the ARIMA model, to be closer to what would have been done with a real time series.
+Even if it would improve the results in term of phase shift (see supplemental material) we prefer to only use the last 12 years to identify and estimate the ARIMA model, to be closer to what would have been done with a real-time series.
 
@@
 *Lecture: this plots represent box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).*\
-*For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally (using), 50% of the turning point are detected with a phase shift of at most 4 months.*
+*For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally, 50% of the turning point are detected with a phase shift of at most 4 months.*
 *For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).*
@@
 
-Extending the series using an ARIMA model gives revisions with the latest estimates of the same order of magnitude as the LC filter, but slightly larger revisions between consecutive estimates, particularly between the fourth and fifth estimates (as might be expected, as highlighted in @sec-mmasym).
+Extending the series using an ARIMA model gives revisions with the latest estimates of the same order of magnitude as the LC filter, but slightly larger revisions between consecutive estimates, particularly between the fourth and fifth estimates (as might be expected as highlighted in @sec-mmasym).
 
@@
 The series studied correspond to the database published in November 2022.
-The series studied is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
+It is the level of employment in the United States (series `CE160V`, used in logarithm) around the February 2001 turning point, consistent with the monthly dating of the turning points.
 This turning point was chosen because it is particularly visible in the raw series (@fig-ce16ov-previmp-lp) and this series was select because it is used for dating economic cycles.
@@
 
-@fig-ce16ovlp shows the successive estimates of the trend-cycle using the different methods studied.
+@Fig-ce16ovlp shows the successive estimates of the trend-cycle using the different methods studied.
 For this series, the phase shift is 6 months for the LC and CQ methods and the extension of the series by ARIMA. 
-Except for the LC method (where the turning point is first detected in January 2001), the high phase shift is due to revisions in the turning point detected in the intermediate estimates (even if the correct turning point is detected after 2 months, it is no longer detected after 3 months).
 It is of two months for the other methods (QL, DAF).
-
 In this example, local parametrisation does not reduce the phase shift, but it does reduce revisions.
@@
 
-In this paper, we have focused on monthly series, but the methods presented can be applied to other frequencies.
+In this paper, we focused on monthly series, but the methods presented can be applied to other frequencies.
 In general, for lower frequencies, the bandwidth of the symmetric filter reduced.
 For example, for quarterly series, X-11 uses a 5-term symmetric filter (in general) or 7-term symmetric filter for series with high variability.
-Mechanically, since fewer points are used, there will be less differences between methods.
+Mechanically, since fewer points are used, there will be less differences between methods in term of phase shift.
 For example, on simulated series the results are similar in terms of phase shift but revisions are reduced with local parametrisation (see supplemental material).
@@
 Since X-13ARIMA-SEATS can only be applied for series with frequency lower than monthly, for higher frequencies (weekly, daily, etc.), there is no reference on the length of the filters to uses.
-The length of the filters can be determined using local polynomial methods [see for example @Loader1999 for a description of bandwidth selection approaches] but it will surely be necessary to use higher bandwidth than for monthly series.
+The length of the filters can be determined using local polynomial methods, for example minimising a cross-validation statistics (using the function `rjd3filters::cross_validation()`) or with more complex methods as describe by @Loader1999, but it will surely be necessary to use higher bandwidth than for monthly series.
 The larger the window, the greater the bias in assuming that the trend is locally of degree 1: the LC method will then be certainly less appropriate than other methods (QL, CQ or DAF).
-Moreover, since the bias is higher, the local parametrisation of the filters will be more important.
+Moreover, since the bias is higher, the local parametrisation of the filters will surely have more impact.
 
@@
 In this article and in those associated to trend-cycle extraction techniques, the moving averages are applied and compared on series already seasonally adjusted or without seasonality (simulated series).
-The revisions and the phase shift are limited to 8 months (when the symmetric filter is 13 terms), and this has the advantage of isolating the impacts of the different filters from the other processes inherent in seasonal adjustment.
+The revisions and the phase shift are limited to 8 months (when the symmetric filter is 13 terms) and this has the advantage of isolating the impacts of the different filters from the other processes inherent in seasonal adjustment.
 There are, however, two drawbacks to this simplification:
@@
 1. The estimation of the seasonally-adjusted series depends on the method used to extract the trend-cycle.
-The choice of method used to estimate the trend-cycle can therefore have an impact well beyond 6 months.
+The choice of method used to estimate the trend-cycle can therefore have an impact well beyond 6 months (when a 13-term Henderson filter is used).
 
@@
 Direct application of the methods can therefore lead to biased estimates, due to their presence, whereas seasonal adjustment methods (such as the X-13ARIMA-SEATS method) have a correction module for atypical points.
-Furthermore, as shown in particular by @dagum1996new, the final symmetric filter used by X-13ARIMA-SEATS to extract the trend-cycle (and therefore the one indirectly used when applying the methods to seasonally-adjusted series), cycles of length 9 and 10 months (generally associated with noise rather than the trend-cycle) are reduced by just 38%.
-The final asymmetric filters even amplify 9 or 10-month cycles.
+Furthermore, as shown in particular by @dagum1996new, the final symmetric filter used by X-13ARIMA-SEATS to extract the trend-cycle (and therefore the one indirectly used when applying the methods to seasonally-adjusted series), reduces by just 38% cycles of length 9 and 10 months (generally associated with noise rather than the trend-cycle).
+The final asymmetric filters even amplify 9 and 10-month cycles.
 This can result in the introduction of undesirable ripples, i.e., the detection of false turning points.
@@
 
-[@Fig-ce16ovcovid; @fig-ce16ov-previmp-covid] show the successive estimates of the trend-cycle and the associated implicit forecast for the US employement during the COVID-19.
+@Fig-ce16ovcovid and @fig-ce16ov-previmp-covid show the successive estimates of the trend-cycle and the associated implicit forecast for the US employment during the COVID-19.
 First, we observe that while the real turning point is in April 2020, the 13-term Henderson filter produces final estimates of the trend-cycles, biased by the presence of a huge outlier, which lead to a turning point detected in June 2020.
@@
 For the QL method, the local parametrisation seems to slightly reduce the revisions.
-For the LC method, local parametrisation has few impacts on the estimates of the trend-cycle but the associated implicit forecasts are not plausible: the final estimate of the slope is biased by the outlier.
+For the LC method, local parametrisation has few impact on the estimates of the trend-cycle but the associated implicit forecasts are not plausible: the final estimate of the slope is biased by the outlier.
 Third, the ARIMA model can lead to unrealistic forecasts.
@@
 \justifying
-*Lecture: the dotted vertical line corresponds to the date of the turning point (April 2020) but the detected one with the 13-term Henderson filter is June 2020.*
+*Lecture: the dotted vertical line corresponds to the date of the turning point (June 2020) detected with the 13-term Henderson filter but the real one is June 2020.*
 *The curve "Apr 2020" corresponds to the estimates of the trend-cycle using the data observed until April 2020.*
@@
 \justifying
-*Lecture: the curve "Apr 2020" corresponds to the implicit forecasts associated to the asymmetric filters used for the trend-cycle estimations using the data observed until April 2001.*
+*Lecture: the curve "Apr 2020" corresponds to the implicit forecasts associated to the asymmetric filters used for the trend-cycle estimations using the data observed until April 2020.*
 
@@
 
-- Adjustment of outliers prior to filtering, for example with a RegARIMA model or other correction modules like the one.
-
-- Increase the bandwidth of the filter used to extract the trend-cycle.
+- Adjustment of outliers prior to filtering, for example with a RegARIMA model or other correction modules like the one used in X-11.
+
+- Increase the bandwidth of the filter used to extract the trend-cycle (it will smooth the trend-cycle by giving less weights to the atypical points).
 
 - Since all the filter studied in this paper are equivalent to a local polynomial model, an external regressor can be added to the linear model to handle the outlier (the matrix $\boldsymbol X$ would then be modified).
+In the case of an additive outlier, this is equivalent to set to 0 the coefficient of the moving average associated to the date of the outlier (using $y_{t-h}$ to $y_{t+h}$ for the trend-cycle estimates, if the outlier is at the date $t$ the central coefficient is set to 0) and renormalise the sum of the coefficients to 1.
 
@@
 
-By comparing the different methods, we can learn a few lessons about the construction of these moving averages.
-
-During economic downturns, asymmetric filters used as an alternative to extending the series using the ARIMA model can reduce revisions to intermediate estimates of the trend-cycle and enable turning points to be detected more quickly.
-
+By comparing the different methods, we can learn a few lessons about the construction of these moving averages.\
+During economic downturns, asymmetric filters used as an alternative to extending the series using the ARIMA model can reduce revisions to intermediate estimates of the trend-cycle and enable turning points to be detected more quickly.\
 At the end of the period, modelling polynomial trends of degree greater than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce variance into the estimates (and therefore more revisions) without allowing faster detection of turning points.
-At the end of the period, for estimating the trend-cycle, we can therefore restrict ourselves to methods modelling polynomial trends of degree two or less (linear-constant, LC, and quadratic-linear, QL).
-In addition, parametrising polynomial filters locally as suggested in this paper enables turning points to be detected more quickly (especially for the QL filter).
+For real-time estimates of the trend-cycle, we can therefore restrict ourselves to methods modelling polynomial trends of degree two or less (linear-constant, LC, and quadratic-linear, QL).
+In addition, parametrising polynomial filters locally as proposed in this paper enables turning points to be detected more quickly (especially for the QL filter).
 Even when the phase shift is not reduced, local parametrisation is recommended because it reduces revisions and produces intermediate estimates that are more consistent with expected future trends.
-However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too short (i.e., if the variability of the series is "low"), retaining polynomial trends of degree one or less (LC method) produces poorer results in terms of detecting turning points.
+However, with these methods, the length of the filter used must be adapted to the variability of the series: if the filter used is too long (i.e., if the variability of the series is "low"), retaining polynomial trends of degree one or less (LC method) produces poorer results in terms of detecting turning points.
 
@@
 Asymmetric filters are calibrated using indicators calculated for the estimation of symmetric filters (for example, to automatically determine their length), whereas a local estimate might be preferable.
-Furthermore, we have only focused on monthly series with a 13-term symmetric filter, but the results may be different if the symmetric filter studied is longer or shorter and if we study series with other frequencies (quarterly or daily, for example).
-
-Another possibility would be to study the impact of atypical points: moving averages, like any linear operator, are very sensitive to the presence of atypical points. 
+Furthermore, we have only focused on monthly series with a 13-term symmetric filter, but the results may be different if the symmetric filter studied is longer or shorter and if we study series with other frequencies (weekly or daily, for example).
+
+Another possibility could be to study the impact of atypical points: moving averages, like any linear operator, are highly sensitive to the presence of atypical points. 
 To limit their impact, in X-11 a strong correction for atypical points is performed on the irregular component before applying the filters to extract the trend-cycle.
-This leads us to study the impact of these points on the estimation of the trend-cycle and turning points, and also to explore new types of asymmetric filters based on robust methods (such as robust local regressions or moving medians).
+This leads to examining the impact of these outliers on the estimation of the trend-cycle and turning points, and also to explore new types of asymmetric filters based on robust methods (such as robust local regressions or moving medians).
 

### 2024-04-30 paper/article_jos.tex
v1

@@
 \usepackage{graphicx}
+\usepackage{ragged2e}
 \usepackage{animate, dsfont, here, xspace}
@@
 \ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{british}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
@@
   pdftitle={Improving real-time trend estimates using local parametrisation of polynomial regression filters},
+  pdflang={en-GB},
   colorlinks=true,
@@
 
-This paper studies and compares real-time estimates of the trend-cycle
-component with moving averages based on local polynomial regression. In
-particular, this theory allows us to reproduce Henderson's symmetric and
-Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal
-adjustment algorithm. This paper proposes two extensions of local
-polynomial filters for real-time trend-cycle estimates (for the last
-estimates): first including a timeliness criterion to minimise the phase
-shift; second, with procedure for parametrising asymmetric filters
-locally While they are generally parametrised globally, which can be
-suboptimal around turning points. An empirical comparison, based on
-simulated and real data, shows that modelling polynomial trends that are
-too complex introduces more revisions without reducing the phase shift,
-and that local parametrisation reduces the delay in detecting turning
-points and reduces revisions. The paper also shows how these results can
-be easily reproduced using the R package \texttt{rjd3filters} dedicated
-to the manipulation of moving averages.
+This paper examines and compares real-time estimates of the trend-cycle
+component using moving averages constructed with local polynomial
+regression. It enables to reproduce Henderson's symmetric and Musgrave's
+asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment
+algorithm. This paper proposes two extensions of local polynomial
+filters for real-time trend-cycle estimates: first including a
+timeliness criterion to minimise the phase shift; second with procedure
+for parametrising asymmetric filters locally while they are generally
+parametrised globally, which can be suboptimal around turning points. An
+empirical comparison, based on simulated and real data, shows that
+modelling polynomial trends that are too complex introduces more
+revisions without reducing the phase shift, and that local
+parametrisation reduces the delay in detecting turning points and
+reduces revisions. The results are reproducible and all the methods can
+be easily applied using the R package \texttt{rjd3filters}.
 
@@
 Analysis of the economic cycle, and in particular the early detection of
-turning points in a series, is a major topic in the analysis of economic
-outlook. To this end, economic indicators are generally seasonally
-adjusted. However, in order to improve their interpretability, it may be
-necessary to perform additional smoothing in order to reduce noise, and
-thus analyse the trend-cycle component. By construction, trend-cycle
-extraction methods are closely related to seasonal adjustment methods,
-since they are generally applied to seasonally adjusted series.
+turning points, is a major topic in the analysis of economic outlook. To
+achieve this, economic indicators are generally seasonally adjusted.
+However, in order to improve their interpretability, it may be necessary
+to perform additional smoothing to reduce noise, thereby to analyse the
+trend-cycle component. By construction, trend-cycle extraction methods
+are closely related to seasonal adjustment methods, since they are
+generally applied to seasonally adjusted series.
 
@@
 date \(t\) it computes a weighted mean of \(p\) past points and \(q\)
-future points where \(p,q\geq0\) depends on the moving average. Moving
-averages are ubiquitous in business cycle extraction and seasonal
-adjustment methods. For example, the X-11 decomposition method
-\autocite{x11}, used in X-13ARIMA-SEATS software \autocite{x13}, uses
-several moving averages to estimate the main components of a time
-series. Symmetric filters are applied to the center of the series, but
-when it comes to estimate the most recent points, all of these methods
-must rely on asymmetric filters. For trend-cycle extraction, the most
-popular symmetric filter is the \textcite{henderson1916note} moving
-average, which is used in the X-11 seasonal adjustment algorithm.
-
-However, for real-time estimates, due to the lack of future
-observations, all these methods must rely on asymmetric filters to
-estimate the most recent points. For example, for trend-cycle
-extraction, X-11 uses Henderson's symmetric filter and
-\textcite{musgrave1964set}'s asymmetric filters on an extended series
-using an ARIMA model. Since the predicted values are linear combinations
-of past values, this amounts to applying asymmetric averages at the end
-of the series.
-
-If the classic asymmetric moving averages have good properties regarding
-the future revisions induced by the process (see for example
-\textcite{pierce1980SA}), they create, by construction, phase shifts
-that impact the real-time estimation of turning points, introducing time
-delay in the detection.
+future points where \(p,q\geq0\) depends on the moving average. They are
+ubiquitous in business cycle extraction and seasonal adjustment methods.
+For example, the X-11 seasonal adjustment method
+\autocite{ladiray2011seasonal}, used in X-13ARIMA-SEATS software
+\autocite{x13}, uses several moving averages to estimate the main
+components of a time series. Symmetric filters are applied to the centre
+of the series and for trend-cycle extraction, the most popular symmetric
+filter is the \textcite{henderson1916note} moving average, which is used
+in the X-11 seasonal adjustment algorithm. However, for real-time
+estimates, due to the lack of future observations, all these methods
+must rely on asymmetric filters to estimate the most recent points. For
+example, for trend-cycle extraction, X-11 uses Henderson's symmetric
+filter and \textcite{musgrave1964set}'s asymmetric filters on an
+extended series using an ARIMA model. Since the predicted values are
+linear combinations of past values, this amounts to applying asymmetric
+moving averages at the end of the series. If the classic asymmetric
+moving averages have good properties regarding the future revisions
+induced by the process (for example \textcite{pierce1980SA} shows
+similar performances in terms of revisions of X-11 compared to
+theoretically superior procedure based on ARIMA forecast models), they
+create, by construction, phase shifts that impact the real-time
+estimation of turning points, introducing time delay in the detection.
 
@@
 general class of symmetric and asymmetric moving averages. In
-particular, it becomes possible to reproduce Henderson's symmetric
-filter and Musgrave's asymmetric filters. The aim of this study is to
-propose two extensions to the \textcite{proietti2008} class of filters
-in order to two drawbacks of those methods. Firstly, adding, in the
-optimisation process, a timeliness criterin to directly control the
-phase shift (i.e.~the delay in detecting turning points). Secondly,
-calibrating asymmetric filter locally, while they are generally
+particular, this class encompass Henderson's symmetric filter and
+Musgrave's asymmetric filters. The aim of this study is to propose two
+extensions to the \textcite{proietti2008} class of filters in order to
+take account of two drawbacks of those methods. Firstly adding, in the
+optimisation process, a timeliness criterion to directly control the
+phase shift (i.e., the delay in detecting turning points). Secondly
+calibrating asymmetric filters locally, while they are generally
 calibrated globally using thus assumptions that may be wrong locally.
 Those extensions are implemented in the statistical software R
-\autocite{logicielr} throw the \texttt{rjd3filters} package (available
-at \url{https://github.com/rjdemetra/rjd3filters}) and can therefore be
-easily used. All the code associated to this paper is available at
-\url{https://github.com/XXXX} (anonymized linked before the manuscript
-is accepted).
+\autocite{logicielr} throw the \texttt{rjd3filters} package
+\autocite{rjd3filters} and can therefore be easily used. All the code
+associated to this paper is available at \url{https://github.com/XXXX}
+(anonymized linked before the manuscript is accepted).
 
@@
 The basic assumption used in time series decomposition methods is that
-the input observed time series, \(y_t\) can be decomposed into a signal
-component \(\mu_t\) and a an erratic component \(\varepsilon_t\) (called
+the input observed time series, \(y_t\), can be decomposed into a signal
+component \(\mu_t\) and an erratic component \(\varepsilon_t\) (called
 irregular): \[
@@
 the input time series is seasonally adjusted (or without seasonality),
-the signal \(\mu_t\) represents the trend and cycle components, here
-estimated jointly and called trend-cycle \(TC_t\). This component can be
-locally approximated by a polynomial degree (see
-section~\ref{sec-nonparamreg}): \[
+the signal \(\mu_t\) represents the trend (sustained and systematic
+variations over a long period of time) and cycle (cyclical movements
+superimposed upon the long-term trend) components, here estimated
+jointly and called trend-cycle \(TC_t\). This component can be locally
+approximated by a polynomial degree (section~\ref{sec-nonparamreg}): \[
 \mu_t = \sum_{j=0}^d\beta_jt^j+\xi_t
@@
 \(\varepsilon_t\) separately \autocite[see for
-example][]{GrayThomson2002}, an usual assumption, adopted in this paper,
+example][]{GrayThomson2002}, a usual assumption, adopted in this paper,
 is to is to gather \(\xi_t\) and \(\varepsilon_t\). Thus, the
@@
 The estimate \(\hat \beta_0\) provides the trend-cycle estimate
-\(\hat TC_t\) and it can be shown that this equivalent to apply of
-moving average (see section~\ref{sec-nonparamreg}). The goal of this
-paper is to present two new methods to separate the trend-cycle
-component from the noise component in real-time (i.e.~when no future
-point is known), using moving averages build on local polynomial
-smoothing techniques.
+\(\widehat{TC}_t\) and it can be shown that this equivalent to apply of
+moving average (section~\ref{sec-nonparamreg}). The goal of this paper
+is to present two new methods to separate the trend-cycle component from
+the noise component in real-time (i.e., when no future point is known),
+using moving averages build on local polynomial smoothing techniques.
 
@@
 Let \(p\) and \(f\) two integers, a moving average
-\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
+\(M_{\boldsymbol\theta}\) is defined by a set of coefficients
 \(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
@@
 \item
-  When \(p=f\) the moving average is said to be \emph{centered}. If we
+  When \(p=f\) the moving average is said to be \emph{centred}. If we
   also have \(\forall k:\:\theta_{-k} = \theta_k\), the moving average
@@
 \(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
-\emph{time shift} function. We this have: \begin{align*}
+\emph{time shift} function. We thus have: \begin{align*}
 \rho_{\boldsymbol\theta}(\omega) &=\sqrt{\left(\sum_{k=-p}^{+f}\theta_k\cos(\omega k)\right)^2+
@@
 
-Fourier decomposition allows us to analyze any time series as a sum of
+Fourier decomposition allows us to analyse any time series as a sum of
 harmonic series, and each component (trend, cycle, seasonal, irregular)
-is associated with a set of frequencies. Noting \(\omega = 2\pi/s\), the
+is associated with a set of frequencies. With \(\omega = 2\pi/s\), the
 harmonic series of frequency \(\omega\) represents a series that repeats
@@
 paper, we consider that the trend-cycle is associated with frequencies
-in the interval \([0, 2\pi/12]\), i.e.~movements recurring at least
+in the interval \([0, 2\pi/12]\), i.e., movements recurring at least
 every 12 months. Even if different frequencies are sometimes used
@@
 for the asymmetric Musgrave filter (see section~\ref{sec-lppasymf})
-often used for real-time trend-cycle estimation (i.e.~when no point in
+often used for real-time trend-cycle estimation (i.e., when no point in
 the future is known). The gain function is greater than 1 on the
@@
 
-\begin{figure}
+\begin{figure}[H]
 
 \caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
-function for the Musgrave filter used in X-11 in real time (last
+function for the Musgrave filter used in X-11 in real-time (last
 estimate) for monthly series when the bandwidth is 6.}
@@
 
-Note: the Musgrave filter depends on a ratio \(I/C\) (see
-section~\ref{sec-nonparamreg}). In X-11, for monthly series, when the
-bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used,
-what is used in most cases), the ratio is fixed to 3.5 and this value is
-used in this figure.
+\justifying
+
+\emph{Lecture: on monthly times series, 12-month cycles (associated to
+the frequency \(2\pi/12=\pi/6\)) are amplified by 10\%
+(\(G_{\boldsymbol\theta}(\pi/6)\simeq1.1\)) and associated turning
+points are detected with a delay of 1 months
+(\(\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1\)).}
+\emph{8-month cycles (associated to the irregular) are only reduced by
+6\% (\(G_{\boldsymbol\theta}(\pi/4)\simeq0.94\)).}\\
+\emph{Note: the Musgrave filter depends on a ratio denoted \(I/C\) (see
+section~\ref{sec-nonparamreg}).} \emph{In X-11, for monthly series, when
+the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is
+used, what is used in most cases), the ratio is fixed to 3.5 and this
+value is used in this figure.}
 
@@
 a function of time \(f(t)\) if
-\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).
-
-For the moving average \(M_{\boldsymbol\theta}\), it can be shown that:
+\(\forall t,M_{\boldsymbol\theta} f(t)=f(t)\).
+
+For a moving average \(M_{\boldsymbol\theta}\), it can be shown that:
 
@@
   \sum_{k=-p}^{+f}\theta_k=1
-   \text{ and }
-  \forall j \in \left\{1,2,\dots,d\right\}:\:
+   \text{(preserves constant) and }
+  \forall j \in \left\{1,2,\dots,d\right\},\:
   \sum_{k=-p}^{+f}k^j\theta_k=0.
@@
 \item
-  Apply symmetric filters to extended forecast series. This method seems
-  to date back to \textcite{deforest1877adjustment}, which also suggests
-  modelling a polynomial trend of degree three or less at the end of the
-  period. A similar approach is used in the X-13ARIMA-SEATS seasonal
-  adjustment method: the series is extend by one year (by default) with
-  an ARIMA model to minimise the revisions associated to asymmetric
-  filters. Since the final filter used for the seasonal adjustment
-  extraction is much longer than one year (in most of the cases, for
-  monthly series, the bandwidth is 78), the input series is just
-  partially extended (and asymmetric filters are still used) to prevent
-  from instability over long periods (e.g., outliers or model shifts).
+  Apply symmetric filters to an extended forecast series (which is
+  equivalent to using asymmetric moving averages, since forecasts are
+  linear combinations of the past). This method seems to date back to
+  \textcite{deforest1877adjustment}, which also suggests modelling a
+  polynomial trend of degree three or less at the end of the period. A
+  similar approach is used in the X-13ARIMA-SEATS seasonal adjustment
+  method: the series is extended by one year (by default) with an ARIMA
+  model to minimise the revisions associated to asymmetric filters.
+  Since the final filter used for the seasonal adjustment extraction is
+  much longer than one year (in most of the cases, for monthly series,
+  the bandwidth is 78), the input series is just partially extended (and
+  asymmetric filters are still used) to prevent from instability over
+  long periods (e.g., outliers or model shifts).
 \end{enumerate}
 
-Ultimately, the second method is equivalent to using asymmetric moving
-averages, since forecasts are linear combinations of the past.
-
-Conversely, the implicit forecasts of an asymmetric moving average can
-be deduced from a reference symmetric moving average. This allows us to
-judge the quality of real-time estimates of the trend-cycle and to
-anticipate future revisions when forecasts are far from expected
+Conversely, the \emph{implicit forecasts} of an asymmetric moving
+average can be deduced from a reference symmetric moving average. This
+allows us to judge the quality of real-time estimates of the trend-cycle
+and to anticipate future revisions when forecasts are far from expected
 evolutions.\\
@@
 \(\boldsymbol w^0=(w_{-h}^0, \dots, w_{0}^0)\) is used for estimation in
-real time estimate (when 0 point in the future is known),
+real-time estimate (when 0 point in the future is known),
 \(\boldsymbol w^1=(w_{-h}^1, \dots, w_{0}^1, w_{1}^1)\) is used for
@@
 used to get the intermediate trend-cycle estimates for the dates \(t-h\)
-to \(h\)). The implicit forecast \(y_{t+1}^*,\dots y_{t+h}^*\) induced
-by \(w^0,\dots w^{h-1}\) are the forecasts of the series \(y_t\) for
-which the estimates of the symmetric filter applied to the extended
+to \(t\)). The \emph{implicit forecast} \(y_{t+1}^*,\dots y_{t+h}^*\)
+induced by \(w^0,\dots w^{h-1}\) are the forecasts of the series \(y_t\)
+for which the estimates of the symmetric filter applied to the extended
 series
@@
 series \(\boldsymbol y^*\). This means: \begin{align*}
-\underbrace{\sum_{i=-h}^0 v_iy_{ti} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
-\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ points after }t} \\
+\underbrace{\sum_{i=-h}^0 v_iy_{t+i} + \sum_{i=1}^h v_iy_{t+i}^*}_{\text{smoothing by }\boldsymbol v\text{ of the extended series}} &=
+\underbrace{\sum_{i=-h}^0 w_i^0y_{t+i}}_{\text{smoothing by }\boldsymbol w^0\text{ using }0\text{ point after }t} \\
 &=
-\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ points after }t} \\
-&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
+\underbrace{\sum_{i=-h}^0 w_i^1y_{t+i} + w_1^1y_{t+1}^*}_{\text{smoothing by }\boldsymbol w^1\text{ using }1\text{ point after }t} \\
+&= \underbrace{\sum_{i=-h}^0 w_i^2y_{t+i} + \sum_{i=1}^{2}w_i^2y_{t+i}^*}_{\text{smoothing by }\boldsymbol w^2\text{ using }2\text{ points after }t} \\
 &= \cdots \\
@@
 \begin{equation}\phantomsection\label{eq-taylor}{
-\forall t \::\:\underbrace{\mu(t)}_{=TC_t} = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
+\forall t,\:\mu(t) = \underbrace{\mu(t_0)}_{=TC_{t_0}} + \mu'(t_0)(t-t_0)+\dots +
 \frac{\mu^{(d)}(t_0)}{d!}(t-t_0)^d+R_d(t),
@@
 that the trend is locally polynomial. We also generally assume that the
-bandwith is fixed over time: \(h(t_0)=h.\) This is also what is assumed
+bandwidth is fixed over time: \(h(t_0)=h.\) This is also what is assumed
 in the seasonal adjustment algorithm X-11 and what will be assumed in
@@
 Various estimation methods can be used to derive symmetric and
-asymmetric moving averages.\\
-\textcite{GrayThomson1996} propose a complete statistical framework that
-makes it possible, in particular, to model the error in approximating
-the trend by local polynomials. However, as the specification of this
-error is generally complex, simpler models may be preferred, such as
-that of \textcite{proietti2008}.\\
-Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
-the trend-cycle but using the theory of Hilbert spaces with reproducing
-kernels for estimation, which has the particular advantage of
-facilitating the calculation of different moving averages at different
-time frequencies. See \textcite{inseeDTM202401} for a more detailed
-comparison of the different recent methods for trend-cycle extraction
-included in the \texttt{rjd3filters} package
+asymmetric moving averages. For example, \textcite{GrayThomson1996}
+propose a complete statistical framework that makes it possible, in
+particular, to model the error in approximating the trend by local
+polynomials. However, as the specification of this error is generally
+complex, simpler models may be preferred, such as that of
+\textcite{proietti2008}. \textcite{dagumbianconcini2008} propose a
+similar modelling of the trend-cycle but using the theory of Hilbert
+spaces with reproducing kernels for estimation, which has the particular
+advantage of facilitating the calculation of different moving averages
+at different time frequencies. See \textcite{inseeDTM202401} for a more
+detailed comparison of the different recent methods for trend-cycle
+extraction included in the \texttt{rjd3filters} package.
 
@@
 (used for real-time estimates). This paper proposed two extensions of
-their approach to build asymetric moving averages: a first one adding a
+their approach to build asymmetric moving averages: a first one adding a
 criteria to control the phase shift (section~\ref{sec-localic}), not
 used in the simulation (section~\ref{sec-comparison}) but which could be
-used in future studied; a second one using a local parameterisation of a
-parameter usually parameterised globally
+used in future studied; a second one using a local parametrisation of a
+parameter usually parametrised globally
 (section~\ref{sec-lptimeliness}), which is used in the simulation.
@@
 
-Using \textcite{proietti2008}'s notation, we assume that our time series
-\(y_t\) can be decomposed into \[
-y_t=\mu_t+\varepsilon_t,
-\] where \(\mu_t\) is the trend and
+Using \textcite{proietti2008}'s framework, we assume that our time
+series \(y_t\) can be decomposed into \[
+y_t=TC_t+\varepsilon_t,
+\] where \(TC_t\) is the trend-cycle and
 \(\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})\) is the
 noise (the series is therefore seasonally adjusted). Within a
-neighbourhood \(h\) of \(t\), the local trend \(\mu_t\) is approximated
-by a polynomial of degree \(d\), such that \(\mu_t\simeq m_{t}\) with:
-\[
-\forall j\in\left\{-h,-h+1,\dots,h\right\}:\:
+neighbourhood \(h\) of \(t\), the local trend \(TC_t\) is approximated
+by a polynomial of degree \(d\), such that \(TC_t\simeq m_{t}\) with: \[
+\forall j\in\left\{-h,-h+1,\dots,h\right\},\:
 y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
 \] The problem of trend extraction is equivalent to estimating
-\(m_t=\beta_0\) (the constant in the previous formula).
+\(m_t=TC_t=\beta_0\) (the constant in the previous formula).
 
@@
 \begin{equation}\phantomsection\label{eq-mmsym}{
-\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
-}\end{equation} To conclude, the estimate of the trend \(\hat{m}_{t}\)
-can be obtained applying the symmetric filter \(\boldsymbol \theta\) to
-\(y_t\) (\(\boldsymbol \theta\) is symmetric due to the symmetry of the
-kernel weights \(\kappa_j\)). Moreover, since
+\hat{m}_{t}=\widehat{TC}_t=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\transp{\boldsymbol \theta}\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
+}\end{equation} To conclude, the estimate of the trend
+\(\widehat{TC}_{t}\) can be obtained applying the symmetric filter
+\(\boldsymbol \theta\) to \(y_t\) (\(\boldsymbol \theta\) is symmetric
+due to the symmetry of the kernel weights \(\kappa_j\)). Moreover, since
 \(\transp{\boldsymbol X}\boldsymbol \theta=\boldsymbol e_{1}\) we have:
@@
 the central estimation (\(\kappa_0\)) and decreases towards 0 as it
-moves away from the central estimation. The uniform kernel should
+moves away from the central observation. The uniform kernel should
 therefore be avoided. It is then preferable to focus on two other
@@
 \item
-  the number of neighbors \(2h+1\) (or the window \(h\)): if it is too
+  the number of neighbours \(2h+1\) (or the window \(h\)): if it is too
   small, then there will be too little data for the estimates (resulting
@@
 \(\transp{\boldsymbol U_p}\boldsymbol v=\transp{\boldsymbol U}\boldsymbol \theta\):
-they represents the polynomial preservation constraints we want to
-impose on the asymmetric filter. \textcite{proietti2008} show that the
-problem is equivalent to finding \(v\) which minimises:
+they represent the polynomial preservation constraints we want to impose
+on the asymmetric filter. \(\boldsymbol Z\) is then associated to the
+polynomial constraints of the symmetric filter not imposed on asymmetric
+filter. \textcite{proietti2008} show that the problem is equivalent to
+finding \(v\) which minimises:
 \begin{equation}\phantomsection\label{eq-lppasym}{
@@
 \item
-  If the ratio is large (\(3.5< R\)) then a 23 term filter is used (to
+  If the ratio is large (\(3.5< R\)) then a 23-term filter is used (to
   remove more noise).
 \item
-  If the ratio is small (\(R<1\)) a 9 term filter is used.
+  If the ratio is small (\(R<1\)) a 9-term filter is used.
 \item
@@
   \(\boldsymbol v\) preserves constant signals (\(d^*=0\)). We obtain
-  Musgrave filters when the Henderson kernel is used and the Henderson
-  filter is used as the symmetric filter.
+  Musgrave filters when the Henderson filter is used for the symmetric
+  filter.
 \item
@@
 To measure the phase shift between the input series and the filtered
-one, equation~\ref{eq-lppasym} introduces a criterion, \(T_g\), called
+one, \textcite{ch15HBSA} introduces a criterion, \(T_g\), called
 \emph{timeliness}. When a moving average \(\boldsymbol v\) is applied,
@@
 \(\omega_2=2\pi/12.\) For the penalty function, the authors suggest
-\(f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2\) and show thant in
+\(f\colon(\rho,\varphi)\mapsto\rho^2\sin(\varphi)^2\) and show that in
 that case the criteria can be written in a quadratic form: \[
@@
 \widetilde {\boldsymbol c},
-\] where \(\alpha_T\) is the weight associated with the timeliness
-criterion,
+\] where
 \(\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p} + \alpha_T\boldsymbol T\),
+\(\alpha_T\) is the weight associated with the timeliness criterion,
 \(\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\transp{\boldsymbol \delta}\transp{\boldsymbol Z_p}-\alpha_T\boldsymbol \theta_p\boldsymbol T\)
@@
 therefore makes it possible to find all the symmetric and asymmetric
-filters presented in the previous section but also generalizes the
-approach of \textcite{GrayThomson1996}. This extension is implemented in
-the R function \texttt{rjd3filters::lp\_filter()}.
+filters presented in the previous section. It is implemented in the R
+function \texttt{rjd3filters::lp\_filter()}.
 
@@
 focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
-However, in future studies, we could imagine a two step calibration: fix
+However, in future studies, we could imagine a two-step calibration: fix
 the value of \(\lvert\delta/\sigma\rvert\) to find \(\alpha_T\) by
@@
 IC-ratio or a cross-validation criterion). However, we might prefer a
-local parametrisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
-as a function of time. Indeed, although the overall parametrisation is
-generally valid, assuming a constant value of the
+local parametrisation: a ratio \(\lvert\delta/\sigma\rvert\) which
+varies as a function of time. Indeed, although the overall
+parametrisation is generally valid, assuming a constant value of the
 \(\lvert\delta/\sigma\rvert\) ratio for all asymmetric filters does not
@@
 to be captured, such as the decay towards 0 before a turning point and
-the growth after the turning point for the LC method:
+the growth after the turning point for the LC method. \(\sigma^2\) and
+\(\delta\) are estimated as follows:
 
@@
   the number of parameters of a parametric model). In the case of local
-  polynomial smoothings, they can be linked to the coefficient of the
-  associated filter: \(\nu_1 = (n-h)\theta_0\) (number of observations
+  polynomial smoothing, they can be linked to the coefficient of the
+  associated filter: \(\nu_1 = (n-2h)\theta_0\) (number of observations
   multiplied by the central coefficient of the moving average) and
-  \(\nu_2 = (n-h)\sum_{i=-h}^{h} \theta_i^2\) (number of observations
+  \(\nu_2 = (n-2h)\sum_{i=-h}^{h} \theta_i^2\) (number of observations
   multiplied by the sum of squares of coefficients of the moving
@@
   biased of \(\widehat{TC}_t\) is thus already small and using smaller
-  bandwidths have almost no impact on the estimates (see Supplemental
+  bandwidths have almost no impact on the estimates (see supplemental
   materials).
@@
   averages.\\
-  Although a moving average of different length to that used for
-  estimating the trend could be considered, this seems to degrade the
-  results in terms of phase shift (using the same methodology as in
-  section~\ref{sec-comparison}). Furthermore, for the construction of
-  moving averages, the trend can be modelled as being locally of degree
-  2 or 3 (this has no impact on the final estimate of concavity). In
-  this paper, we have chosen to model a trend of degree 2: this slightly
-  reduces the phase shift (see Supplemental material) but slightly
-  increases the revisions linked to the first estimate of the
-  trend-cycle. figure~\ref{fig-mmpenteconcac} shows the moving averages
-  used.
+  For the construction of moving averages, the trend can be modelled as
+  being locally of degree 2 or 3 (this has no impact on the final
+  estimate of concavity). In this paper, we have chosen to model a trend
+  of degree 2: this slightly reduces the phase shift (see supplemental
+  material) but slightly increases the revisions linked to the first
+  estimate of the trend-cycle. Figure~\ref{fig-mmpenteconcac} shows the
+  moving averages used.
 \end{itemize}
@@
 To distinguish the gain from the local parametrisation of the noise
-biased associated to the real-time estimates of the slope and concavity,
-in the empirical applications of section~\ref{sec-comparison} we also
-compares the results to the final local parametrisation obtained by
-estimating \(\delta\) using all the data (i.e., using the symmetric
-filters shown in figure~\ref{fig-mmpenteconcac}), while maintaining a
-real-time estimate of \(\sigma^2\). figure~\ref{fig-mmpenteconcac-ex}
-shows an example of the comparison of estimates of the ratio
+associated to the real-time estimates of the slope and concavity, in the
+empirical applications of section~\ref{sec-comparison} we also compares
+the results to the final local parametrisation obtained by estimating
+\(\delta\) using all the data (i.e., using the symmetric filters shown
+in figure~\ref{fig-mmpenteconcac}), while maintaining a real-time
+estimate of \(\sigma^2\). Figure~\ref{fig-mmpenteconcac-ex} shows an
+example of the comparison of estimates of the ratio
 \(|\delta_1/\sigma|\) with a global estimator (IC ratio) and two local
@@
 
-The turning points are clearly detected by the local estimators
-(changement in the slope), when \(|\delta_1/\sigma|\) is estimated in
-the real-time estimate. However, the real-time local estimates of
-\(|\delta_1/\sigma|\) are noisy: the varability of this estimates, used
-to build asymmetric filters, will then lead to another source of
+The turning points are clearly detected by the local estimators (slope
+tends towards zero) when \(|\delta_1/\sigma|\) is estimated in the
+real-time. However, the real-time local estimates of
+\(|\delta_1/\sigma|\) are noisy: the variability of these estimates,
+used to build asymmetric filters, will then lead to another source of
 revisions in the intermediate estimates.
@@
 moving averages but they can be easily computed using matrix computation
-formula and applied to the series rjd3filters`.
+formula and applied to the series using \texttt{rjd3filters}.
 
@@
 series, a symmetric 13-term filter is used: we then assume than the
-bandwidth is fixed, as in the seasonal adjustement method X-11. A
-nearest neighbour bandwidth is also tested (13-term asymmetric filters:
-real-time estimates uses observation between \(y_{t-12}\) and \(y_{t}\),
-the second asymmetric filters uses obversations between \(y_{t-11}\) and
+bandwidth is fixed, as in the seasonal adjustment method X-11. A nearest
+neighbour bandwidth is also tested (13-term asymmetric filters:
+real-time filter uses observations between \(y_{t-12}\) and \(y_{t}\),
+the second asymmetric filters uses observations between \(y_{t-11}\) and
 \(y_{t+1}\), etc.) but with no improvement in the results (see
-supplemental material). More complexe methods to locally select
+supplemental material). More complex methods to locally select
 bandwidth, like the one proposed by \textcite{fan1992variable}, were not
@@
 lag (the series being seasonally adjusted) and no external variables
-(such as external regressors for correction of atypical points).
-
-The performance of the different methods is judged on criteria relating
-to revisions (between two consecutive estimates and in relation to the
+(such as external regressors for correction of atypical points). The
+performance of the different methods is judged on criteria relating to
+revisions (between two consecutive estimates and in relation to the
 final estimate) and the number of periods required to detect turning
@@
 \item
-  High signal-to-noise ratio (i.e.~low I-C ratio and low variability):
+  High signal-to-noise ratio (i.e., low I-C ratio and low variability):
   \(\sigma_e^2=0.2\) and \(\rho = 3.0, 3.5\) or \(4.0\) (I-C ratio
@@
 \item
-  Medium signal-to-noise ratio (i.e.~medium I-C ratio and medium
+  Medium signal-to-noise ratio (i.e., medium I-C ratio and medium
   variability): \(\sigma_e^2=0.3\) and \(\rho = 1.5,\, 2.0\) or \(3.0\)
@@
 \item
-  Low signal-to-noise ratio (i.e.~high I-C ratio and high variability):
+  Low signal-to-noise ratio (i.e., high I-C ratio and high variability):
   \(\sigma_e^2=0.4\) and \(\rho = 0.5,\, 0.7\) or \(1.0\) (I-C ratio
@@
 For each series and each date, the trend-cycle is estimated using the
-different methods presented in this report. Three quality criteria are
-also calculated:
+different methods presented in this paper Three quality criteria are
+computed:
 
@@
 \item
-  Computation of the phase shift in the detection of turning points. In
-  this paper, we focus on turning points associated to business cycle:
-  it is defined as the succession of phases of economic recession and
+  The \emph{phase shift} in the detection of turning points. In this
+  paper, we focus on turning points associated to business cycle: it is
+  defined as the succession of phases of economic recession and
   expansion, delimited by peaks (highest level of activity) and troughs
@@
 
-  Lets denote \(TC_{t|t'}\) the estimate of the trend-cycle at date
+  Let's denote \(TC_{t|t'}\) the estimate of the trend-cycle at date
   \(t\) using the data up to date \(t'\). The phase shift is often
@@
   on the comparison of the different methods.\\
-  In this paper we use a slightly modified criterion: the phase shift is
-  defined as the number of months needed to detect the right turning
-  point without any future revision. Thus, in the case of a downturn,
-  the phase shift is two if
+  In this paper we use a slightly modified criterion: the \emph{phase
+  shift} is defined as the number of months needed to detect the right
+  turning point without any future revision. Thus, in the case of a
+  downturn, the phase shift is two if
   \(TC_{t-3|t+1}\leq TC_{t-2|t+1}\leq TC_{t-1|t+1}>TC_{t|t+1}\geq TC_{t+1|t+1}\)
@@
   the trend-cycle, the phase shift is at most height (the final estimate
-  of \(TC_{t+1}\) is compute a the date \(t+7\)). This definition is
+  of \(TC_{t+1}\) is compute at the date \(t+7\)). This definition is
   used because it can happen that the right turning point is detected by
@@
 to only use the last 12 years to identify and estimate the ARIMA model,
-to be closer to what would have been done with a real time series.
+to be closer to what would have been done with a real-time series.
 
@@
 
-Lecture: this plots represent box plots of the phase shift: the box
-represents the interquartile range, the central horizontal line the
+\justifying
+
+\emph{Lecture: this plots represent box plots of the phase shift: the
+box represents the interquartile range, the central horizontal line the
 median the dot the outliers and the whisker the range of the data (if no
-outlier).\\
-For simulated series with medium variability (when a symmetric filter of
-length 13 is relevant), with the linear constant (LC) method, 50\% of
-the turning point are detected with a phase shift of at most 5 months
-whereas when the LC filter is parametrised locally (using), 50\% of the
-turning point are detected with a phase shift of at most 4 months. For
-the direct asymmetric filter (DAF), 25\% of the turning are detected
-with a phase shift of at least 7 months (compares to 5 for the LC
-method).
+outlier).}\\
+\emph{For simulated series with medium variability (when a symmetric
+filter of length 13 is relevant), with the linear constant (LC) method,
+50\% of the turning point are detected with a phase shift of at most 5
+months whereas when the LC filter is parametrised locally, 50\% of the
+turning point are detected with a phase shift of at most 4 months.}
+\emph{For the direct asymmetric filter (DAF), 25\% of the turning are
+detected with a phase shift of at least 7 months (compares to 5 for the
+LC method).}
 
@@
 slightly larger revisions between consecutive estimates, particularly
-between the fourth and fifth estimates (as might be expected, as
+between the fourth and fifth estimates (as might be expected as
 highlighted in section~\ref{sec-mmasym}).
 
-\begin{longtable}[t]{ccccccc}
+\begin{table}
 
@@
 
-\tabularnewline
-
+\centering{
+
+\begin{tabular}{ccccccc}
 \toprule
-Method & \$q=0\$ & \$q=1\$ & \$q=2\$ & \$q=3\$ & \$q=4\$ & \$q=5\$\\
+Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
 \midrule
@@
 \bottomrule
-
-\end{longtable}
-
-Lecture: the estimates for \(q=0\) correspond to the first estimates of
-the trend-cycle (real-time estimates when 0 point in the future is
-known), \(q=1\) correspond to the second estimates of the trend-cycle
-(when 1 point in the future is known). Since the final estimates are
-obtain using a symmetric filter of bandwidth 6, \(q=6\) corresponds to
-the final estimates of the trend-cycle (when 6 points in the future are
-known). \(MAE_{fe}(0)\) compares the revision between the first estimate
-(\(q=0\)) and the last estimate (\(q=6\)) and \(MAE_{ce}(0)\) compares
-the revision between the first estimate (\(q=0\)) and the second
-estimate (\(q=1\)).
+\end{tabular}
+
+\justifying
+
+\emph{Lecture: the estimates for \(q=0\) correspond to the first
+estimates of the trend-cycle (real-time estimates when 0 point in the
+future is known), \(q=1\) correspond to the second estimates of the
+trend-cycle (when 1 point in the future is known).} \emph{Since the
+final estimates are obtain using a symmetric filter of bandwidth 6,
+\(q=6\) corresponds to the final estimates of the trend-cycle (when 6
+points in the future are known).} \emph{\(MAE_{fe}(0)\) compares the
+revision between the first estimate (\(q=0\)) and the last estimate
+(\(q=6\)) and \(MAE_{ce}(0)\) compares the revision between the first
+estimate (\(q=0\)) and the second estimate (\(q=1\)).}
+
+}
+
+\end{table}%
 
@@
 example from the FRED-MD database (\textcite{fredmd}) containing
-economic series on the United States. The series studied correspond to
-the database published in November 2022. The series studied is the level
-of employment in the United States (series \texttt{CE160V}, used in
-logarithm) around the February 2001 turning point, consistent with the
-monthly dating of the turning points. This point and this series were
-chosen because the economic turnaround is particularly visible in the
-raw series (figure~\ref{fig-ce16ov-previmp-lp}), because this series is
-used for dating economic cycles and because the FRED-MD database
-facilitates the reproducibility of the results (thanks to the
-availability of series published on past dates). Studying the series up
-to January 2020, this series has a medium variability: a symmetric
-filter of 13 terms is therefore appropriate.
-
-figure~\ref{fig-ce16ovlp} shows the successive estimates of the
-trend-cycle using the different methods studied.
-
-For this series, the phase shift is 6 months for the LC and CQ methods
-and the extension of the series by ARIMA. Except for the LC method
-(where the turning point is first detected in January 2001), the high
-phase shift is due to revisions in the turning point detected in the
-intermediate estimates (even if the correct turning point is detected
-after 2 months, it is no longer detected after 3 months). It is of two
-months for the other methods (QL, DAF).
-
-In this example, local parametrisation does not reduce the phase shift,
-but it does reduce revisions. The QL and DAF polynomials lead to greater
-variability in the intermediate estimates, especially in February 2001.
+economic series on the United States. This database facilitates the
+reproducibility of the results thanks to the availability of series
+published on past dates. The series studied correspond to the database
+published in November 2022. It is the level of employment in the United
+States (series \texttt{CE160V}, used in logarithm) around the February
+2001 turning point, consistent with the monthly dating of the turning
+points. This turning point was chosen because it is particularly visible
+in the raw series (figure~\ref{fig-ce16ov-previmp-lp}) and this series
+was select because it is used for dating economic cycles. Studying the
+series up to January 2020, this series has a medium variability: a
+symmetric filter of 13 terms is therefore appropriate.
+
+Figure~\ref{fig-ce16ovlp} shows the successive estimates of the
+trend-cycle using the different methods studied. For this series, the
+phase shift is 6 months for the LC and CQ methods and the extension of
+the series by ARIMA. It is of two months for the other methods (QL,
+DAF). In this example, local parametrisation does not reduce the phase
+shift, but it does reduce revisions. The QL and DAF polynomials lead to
+greater variability in the intermediate estimates, especially in
+February 2001.
 
@@
 \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}
+
+\justifying
+
+\emph{Lecture: the dotted vertical line corresponds to the date of the
+turning point (February 2001).} \emph{The curve ``Feb 2001'' corresponds
+to the estimates of the trend-cycle using the data observed until
+February 2001.} \emph{Since a 13-term Henderson filter is used for the
+final estimates, for the curve ``Feb 2001'' estimates from September
+2001 to February 2001 are intermediate estimates.}
 
@@
 
+\justifying
+
+\emph{Lecture: the dotted vertical line corresponds to the date of the
+turning point (February 2001).} \emph{The curve ``Mar 2001'' corresponds
+to the implicit forecasts associated to the asymmetric filters used for
+the trend-cycle estimations using the data observed until March 2001.}
+
 }
@@
 
-\section{Discussion}\label{discussion}
-
-\subsection{Outliers and COVID-19}\label{outliers-and-covid-19}
+\subsection{Discussion}\label{discussion}
+
+In the section we discuss the applicability of the presented methods to
+other frequencies (section~\ref{sec-other-freq}) and some limits of the
+presented methods (for example sensibility to outliers like COVID-19,
+section~\ref{sec-out-covid}) .
+
+\subsubsection{Other frequencies}\label{sec-other-freq}
+
+In this paper, we focused on monthly series, but the methods presented
+can be applied to other frequencies. In general, for lower frequencies,
+the bandwidth of the symmetric filter reduced. For example, for
+quarterly series, X-11 uses a 5-term symmetric filter (in general) or
+7-term symmetric filter for series with high variability. Mechanically,
+since fewer points are used, there will be less differences between
+methods in term of phase shift. For example, on simulated series the
+results are similar in terms of phase shift but revisions are reduced
+with local parametrisation (see supplemental material).
+
+Since X-13ARIMA-SEATS can only be applied for series with frequency
+lower than monthly, for higher frequencies (weekly, daily, etc.), there
+is no reference on the length of the filters to uses. The length of the
+filters can be determined using local polynomial methods, for example
+minimising a cross-validation statistics (using the function
+\texttt{rjd3filters::cross\_validation()}) or with more complex methods
+as describe by \textcite{Loader1999}, but it will surely be necessary to
+use higher bandwidth than for monthly series. The larger the window, the
+greater the bias in assuming that the trend is locally of degree 1: the
+LC method will then be certainly less appropriate than other methods
+(QL, CQ or DAF). Moreover, since the bias is higher, the local
+parametrisation of the filters will surely have more impact.
+
+\subsubsection{Outliers and COVID-19}\label{sec-out-covid}
 
 In this article and in those associated to trend-cycle extraction
-techniques, the asymmetrical filters are applied and compared on series
+techniques, the moving averages are applied and compared on series
 already seasonally adjusted or without seasonality (simulated series).
-The revisions and therefore the phase shift are limited to 6 months
-(when the symmetrical filter is 13 terms), and this has the advantage of
-isolating the impacts of the different filters from the other processes
-inherent in seasonal adjustment. There are, however, two drawbacks to
-this simplification:
+The revisions and the phase shift are limited to 8 months (when the
+symmetric filter is 13 terms) and this has the advantage of isolating
+the impacts of the different filters from the other processes inherent
+in seasonal adjustment. There are, however, two drawbacks to this
+simplification:
 
@@
 \item
-  Firstly, the estimation of the seasonally-adjusted series depends on
-  the method used to extract the trend-cycle. The choice of method used
-  to estimate the trend-cycle can therefore have an impact well beyond 6
-  months.
-\item
-  On the other hand, as moving averages are linear operators, they are
-  sensitive to the presence of atypical points. Direct application of
-  the methods can therefore lead to biased estimates, due to their
-  presence, whereas seasonal adjustment methods (such as the
-  X-13ARIMA-SEATS method) have a correction module for atypical points.
-  Furthermore, as shown in particular by \textcite{dagum1996new}, the
-  final symmetrical filter used by X-13ARIMA to extract the trend-cycle
-  (and therefore the one indirectly used when applying the methods to
-  seasonally-adjusted series) misses around 72\% of 9- or 10-month
-  cycles (generally associated with noise rather than the trend-cycle).
-  The final asymmetrical filters even amplify 9 or 10-month cycles. This
-  can result in the introduction of undesirable ripples, i.e.~the
-  detection of false turning points. This problem is reduced by
-  correcting atypical points (these cycles being considered irregular).
-  This is how the \emph{Nonlinear Dagum Filter} (NLDF) was developed:
+  The estimation of the seasonally-adjusted series depends on the method
+  used to extract the trend-cycle. The choice of method used to estimate
+  the trend-cycle can therefore have an impact well beyond 6 months
+  (when a 13-term Henderson filter is used).
+\item
+  As moving averages are linear operators, they are sensitive to the
+  presence of atypical points. Direct application of the methods can
+  therefore lead to biased estimates, due to their presence, whereas
+  seasonal adjustment methods (such as the X-13ARIMA-SEATS method) have
+  a correction module for atypical points. Furthermore, as shown in
+  particular by \textcite{dagum1996new}, the final symmetric filter used
+  by X-13ARIMA-SEATS to extract the trend-cycle (and therefore the one
+  indirectly used when applying the methods to seasonally-adjusted
+  series), reduces by just 38\% cycles of length 9 and 10 months
+  (generally associated with noise rather than the trend-cycle). The
+  final asymmetric filters even amplify 9 and 10-month cycles. This can
+  result in the introduction of undesirable ripples, i.e., the detection
+  of false turning points. This problem is reduced by correcting
+  atypical points and the \emph{Nonlinear Dagum Filter} (NLDF) was
+  defined in this perspective:
 
@@
   \item
-    applying the X-13ARIMA atypical point correction algorithm to the
+    applying the X-11 atypical point correction algorithm \autocite[see
+    for example][ for a description]{ladiray2011seasonal} to the
     seasonally adjusted series, then extending it with an ARIMA model;
@@
     perform a new atypical point correction using a much stricter
-    threshold and then apply the 13-term symmetrical filter. Assuming a
+    threshold and then apply the 13-term symmetric filter. Assuming a
     normal distribution, this amounts to modifying 48\% of the irregular
@@
 
-One possible direction of study would then be to investigate more
-precisely the effect of atypical points on trend-cycle estimation and
-the detection of turning points, but also to explore new types of
-asymmetric filters based on robust methods (such as robust local
-regressions, moving medians, etc.).
+Figure~\ref{fig-ce16ovcovid} and figure~\ref{fig-ce16ov-previmp-covid}
+show the successive estimates of the trend-cycle and the associated
+implicit forecast for the US employment during the COVID-19. First, we
+observe that while the real turning point is in April 2020, the 13-term
+Henderson filter produces final estimates of the trend-cycles, biased by
+the presence of a huge outlier, which lead to a turning point detected
+in June 2020. Second, as in 2001, the CQ and DAF methods lead to lots of
+variability in the intermediate estimates. For the QL method, the local
+parametrisation seems to slightly reduce the revisions. For the LC
+method, local parametrisation has few impact on the estimates of the
+trend-cycle but the associated implicit forecasts are not plausible: the
+final estimate of the slope is biased by the outlier. Third, the ARIMA
+model can lead to unrealistic forecasts.
 
@@
 
+\justifying
+
+\emph{Lecture: the dotted vertical line corresponds to the date of the
+turning point (June 2020) detected with the 13-term Henderson filter but
+the real one is June 2020.} \emph{The curve ``Apr 2020'' corresponds to
+the estimates of the trend-cycle using the data observed until April
+2020.}
+
 }
@@
 successive estimates of the trend-cycle in US employment (in logarithms)
-using local polynomial methods.}
+using local polynomial methods during the COVID-19.}
 
@@
 
+\justifying
+
+\emph{Lecture: the curve ``Apr 2020'' corresponds to the implicit
+forecasts associated to the asymmetric filters used for the trend-cycle
+estimations using the data observed until April 2020.}
+
 }
@@
 
-\subsection{Other frequencies}\label{other-frequencies}
+To sum up, trend-cycle estimates are biased by the presence of outliers
+which can also impact the detection of turning-points. Several approach
+could be used to limit their impact of to directly take them into
+account:
+
+\begin{itemize}
+\item
+  Adjustment of outliers prior to filtering, for example with a RegARIMA
+  model or other correction modules like the one used in X-11.
+\item
+  Increase the bandwidth of the filter used to extract the trend-cycle
+  (it will smooth the trend-cycle by giving less weights to the atypical
+  points).
+\item
+  Since all the filter studied in this paper are equivalent to a local
+  polynomial model, an external regressor can be added to the linear
+  model to handle the outlier (the matrix \(\boldsymbol X\) would then
+  be modified). In the case of an additive outlier, this is equivalent
+  to set to 0 the coefficient of the moving average associated to the
+  date of the outlier (using \(y_{t-h}\) to \(y_{t+h}\) for the
+  trend-cycle estimates, if the outlier is at the date \(t\) the central
+  coefficient is set to 0) and renormalise the sum of the coefficients
+  to 1.
+\item
+  Use of robust methods to estimate the trend-cycle, such as robust
+  local regressions or moving medians.
+\end{itemize}
 
@@
 By comparing the different methods, we can learn a few lessons about the
-construction of these moving averages.
-
+construction of these moving averages.\\
 During economic downturns, asymmetric filters used as an alternative to
@@
 intermediate estimates of the trend-cycle and enable turning points to
-be detected more quickly.
-
+be detected more quickly.\\
 At the end of the period, modelling polynomial trends of degree greater
@@
 variance into the estimates (and therefore more revisions) without
-allowing faster detection of turning points. At the end of the period,
-for estimating the trend-cycle, we can therefore restrict ourselves to
-methods modelling polynomial trends of degree two or less
-(linear-constant, LC, and quadratic-linear, QL). In addition,
-parametrising polynomial filters locally as suggested in this paper
-enables turning points to be detected more quickly (especially for the
-QL filter). Even when the phase shift is not reduced, local
-parametrisation is recommended because it reduces revisions and produces
-intermediate estimates that are more consistent with expected future
-trends. However, with these methods, the length of the filter used must
-be adapted to the variability of the series: if the filter used is too
-short (i.e.~if the variability of the series is ``low''), retaining
-polynomial trends of degree one or less (LC method) produces poorer
-results in terms of detecting turning points.
+allowing faster detection of turning points. For real-time estimates of
+the trend-cycle, we can therefore restrict ourselves to methods
+modelling polynomial trends of degree two or less (linear-constant, LC,
+and quadratic-linear, QL). In addition, parametrising polynomial filters
+locally as proposed in this paper enables turning points to be detected
+more quickly (especially for the QL filter). Even when the phase shift
+is not reduced, local parametrisation is recommended because it reduces
+revisions and produces intermediate estimates that are more consistent
+with expected future trends. However, with these methods, the length of
+the filter used must be adapted to the variability of the series: if the
+filter used is too long (i.e., if the variability of the series is
+``low''), retaining polynomial trends of degree one or less (LC method)
+produces poorer results in terms of detecting turning points.
 
@@
 filter studied is longer or shorter and if we study series with other
-frequencies (quarterly or daily, for example).
-
-Another possibility would be to study the impact of atypical points:
-moving averages, like any linear operator, are very sensitive to the
+frequencies (weekly or daily, for example).
+
+Another possibility could be to study the impact of atypical points:
+moving averages, like any linear operator, are highly sensitive to the
 presence of atypical points. To limit their impact, in X-11 a strong
 correction for atypical points is performed on the irregular component
-before applying the filters to extract the trend-cycle. This leads us to
-study the impact of these points on the estimation of the trend-cycle
-and turning points, and also to explore new types of asymmetric filters
-based on robust methods (such as robust local regressions or moving
-medians).
+before applying the filters to extract the trend-cycle. This leads to
+examining the impact of these outliers on the estimation of the
+trend-cycle and turning points, and also to explore new types of
+asymmetric filters based on robust methods (such as robust local
+regressions or moving medians).
 

### 2024-04-30 paper/resp_referee.qmd
v1

@@
 title: "Response to referee"
+# format: docx
 format: pdf
+lang: en-GB  
 ---
+
+We would like to thanks all the reviewers and the associate editor for the comments, suggestions and the time spent proofreading the paper.
 
@@
 *Nearest neighbour bandwidth was added in supplemental material but it doesn't improve the results in therm of phase shift.* 
-*One explanation could be that the bias is higher with LC/QL methodes (we model linear trend using higher bandwidth) and for other methods (e.g.
+*One explanation could be that the bias is higher with LC/QL methods (we model linear trend using higher bandwidth) and for other methods (e.g.
 DAF) that the revisions to the final symmetric filters are higher.*
@@
 
-*The term "undersirable oscillations" was removed*
+*The term "undesirable oscillations" was removed*
 
@@
 It may be worth considering (in addition) shorter series. 
-Another thought is that the ARIMA model being used to extend the series appears to be modeled using the full series; what if the ARIMA model used to extend the series was instead estimated using just a short span toward the end of the series?
+Another thought is that the ARIMA model being used to extend the series appears to be modelled using the full series; what if the ARIMA model used to extend the series was instead estimated using just a short span toward the end of the series?
 
@@
 *Figures are now black and white a scaled gray legend.*
-*In some of them, there are indeed a lot of information with several labels, but the goal is not to be able to read all the details of all the lines but to see the general behavior of the filters (for example if the implicit forecasts seem relevant or not)*
+*In some of them, there are indeed a lot of information with several labels, but the goal is not to be able to read all the details of all the lines but to see the general behaviour of the filters (for example if the implicit forecasts seem relevant or not)*
 
@@
 
-*We try to clarify in the section associated to the proposed methods: all the functions mentionned are already in rjd3filters.*
+*We try to clarify in the section associated to the proposed methods: all the functions mentioned are already in rjd3filters.*
 *The local parametrisation is not directly in the package but can be easily implemented (especially because the code is available): the variance estimator is in the package but not the moving-averages associated to slope/concavity estimators (which can be easily derived)*
@@
 
-*TODO*
+*More explanation with the reference in parentheses*
 
@@
 
-*A lecture note is added in the tables and more explainations are added to the text*
+*A lecture note is added in the tables and more explanations are added to the text*
 
@@
 
-*We try to clarify the contribution in the abtract, introduction and in the beginning of sections 2 and 3*
+*We try to clarify the contribution in the abstract, introduction and in the beginning of sections 2 and 3*
 

### 2024-04-30 paper/supp_material.qmd
v1

@@
 
-Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
+Distribution of phase shift associated to the local parametrisation of the quadratic-linear (QL) filter using real-time estimates of $\delta/\sigma$, by the bandwidth $h$ of the henderson filter used to estimate $\sigma$ and the degree $d$ of the trend to estimate $\delta$.
 :::
@@
 *We thus do not distinguish local parametrisation and final local parametrisation which asumptions might not be plausible.*
-*Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).*
+*Indeed, using the same framework as for fixed bandwidth, the final local parametrisation for real-time estimates would implie modelling a polynomial of degree 2 or 3 other 25 months (which doesn't seem plausible).*
 

### 2024-04-30 paper/article_jos.qmd
submitted article

@@
 
-In the section we discuss the applicability of the presented methods to other frequencies (@sec-other-freq) and some limits of the presented methods (for example sensibility to outliers like COVID-19, @sec-out-covid) .
+In the section we discuss the applicability of the presented methods to other frequencies (@sec-other-freq) and some limits of the presented methods (for example sensitivity to outliers like COVID-19, @sec-out-covid) .
 

### 2024-04-30 paper/article_jos.tex
submitted article

@@
 other frequencies (section~\ref{sec-other-freq}) and some limits of the
-presented methods (for example sensibility to outliers like COVID-19,
+presented methods (for example sensitivity to outliers like COVID-19,
 section~\ref{sec-out-covid}) .

### 2024-04-30 paper/resp_referee.qmd
submitted article

@@
 # format: docx
-format: pdf
+format: 
+  pdf:
+    geometry:
+      - top=2cm
+      - bottom=2cm
+      - left = 2.5cm
+      - right = 2.5cm
 lang: en-GB  
@@
 # Associated Editor
+
+Thank you for reviewing our paper and thank you for your helpful comments and suggestions.
+Below are our responses (*in italics*) to your comments and suggestions:
 
@@
 
-*This sections was simplified*
+*This sections was simplified.*
 
@@
 
-*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed *
+*The section 1.3.2 was removed: it's not central to the paper, and it would have made the reading more cumbersome to go into the details of the notations. Table 2 has also been removed.*
 
@@
 
-*More explanation are added to this section*
+*More explanation are added to this section.*
 
@@
 
-*Fixed*
+*Fixed.*
 
@@
 
-*Indeed, the formula are corrected*
+*Indeed, the formula are corrected.*
 
@@
 
-*All the section was rewritten to be clearer*
+*All the section was rewritten to be clearer.*
 
@@
 
-*All the code will be in a github repository*
+*All the code will be in a github repository.*
 
 # Referee 1
+
+Thank you for reviewing our paper and thank you for your helpful comments and suggestions.
+Below are our responses (*in italics*) to your comments and suggestions:
 
@@
 
-*The modelled of the input time series is now describe more precisely in section 2 (previous section 1).* 
-*$TC_t$ is here considered as deterministic*
+*The model of the input time series is now describe more precisely in section 2 (previous section 1).* 
+*$TC_t$ is here considered as deterministic (clarified in the text).*
 
@@
 *The notation is simplified and unified and the equations/notations clarified.*
-*The matrix $D$ is needed to see the equivalence with symmetric filter/DAF (clarified in the text).*
+*The matrix $D$ is needed to show the equivalence with symmetric filter/DAF (clarified in the text).*
 
@@
 
-*It was added in the beginning of new sections 3 and 4 than the bandwidth is fixed.* 
+*It was added in the beginning of new sections 3 and 4 that the bandwidth is fixed.* 
 *Nearest neighbour bandwidth was added in supplemental material but it doesn't improve the results in therm of phase shift.* 
@@
 
-*A lecture note is added*
+*A lecture note is added below the figure to explain how to read the panels.*
+*The ratio $I/C$ is no longer mentioned in the caption but some details are added in the figure note.*
 
@@
 
+Thank you for reviewing our paper and thank you for your helpful comments and suggestions.
+Below are our responses (*in italics*) to your comments and suggestions:
+
 > The author simulates time series of length 60 years to illustrate the methodology at the end of the series.
@@
 
-*Those sections are now numbered*
+*Those sections are now numbered.*
 
@@
 
-*Bibliography added and the R icon was removed*
+*Bibliography added and the R icon was removed.*
 
 > Some of the important results appear in the tables and figures, but the accompanying discussion in the text should still be fleshed out more.
+
+*Lecture notes were added to all tables and figures.*
 
@@
 
-*Figures are now black and white a scaled gray legend.*
-*In some of them, there are indeed a lot of information with several labels, but the goal is not to be able to read all the details of all the lines but to see the general behaviour of the filters (for example if the implicit forecasts seem relevant or not)*
+*Figures are now black and white with a scaled grey legend.*
+*In some of them, there are indeed a lot of information with several labels, but the goal is not to be able to read all the details of all the lines but to see the general behaviour of the filters (for example if the implicit forecasts seem relevant or not).*
 
@@
 *We try to clarify in the section associated to the proposed methods: all the functions mentioned are already in rjd3filters.*
-*The local parametrisation is not directly in the package but can be easily implemented (especially because the code is available): the variance estimator is in the package but not the moving-averages associated to slope/concavity estimators (which can be easily derived)*
+*The local parametrisation is not directly in the package but can be easily implemented (especially because the code is available): the variance estimator is in the package but not the moving-averages associated to slope/concavity estimators (which can be easily derived).*
 
@@
 
-*This was clarified in the article, for example in the introduction the text was changed with "the X-11 decomposition method (used in X-13ARIMA-SEATS)"*
+*This was clarified in the article, for example in the introduction the text was changed with "the X-11 decomposition method (used in X-13ARIMA-SEATS)".*
 
@@
 
-*More explanation with the reference in parentheses*
+*More explanation are added with a clearer contribution of the reference to Pierce (1980).*
 
@@
 
-*Table 2 was removed*
+*Table 2 was removed.*
 
@@
 
-*This table was removed but to answer your question the criteria comes from the decomposition of the mean squared error into four quantities*
+*This table was removed but to answer your question the criteria comes from the decomposition of the mean squared error into four quantities.*
 
@@
 
-*See previous comment: it was added in supplemental material*
+*See previous comment: it was added in supplemental material.*
 
 # Referee 3
+
+Thank you for reviewing our paper and thank you for your helpful comments and suggestions.
+Below are our responses (*in italics*) to your comments and suggestions:
 
@@
 
-*This section was rewrite to make it clearer.*
+*This section was rewriten to make it clearer.*
 
@@
 
-*Quarterly series with simulated data are added in supplemental material, a new section was added to discuss the use of the methods with different periodicities*
+*Quarterly series with simulated data are added in supplemental material, a new section was added to discuss the use of the methods with different periodicities.*
 
@@
 
-*The definition of phase-shift was clarified in*
+*The corresponding text was rewritten to clarify the definition of phase-shift used in this paper.*
 
@@
 
-*A reading note is added to figure 5 to improve its reading*
+*A reading note is added to figure 5 to improve its reading.*
 
@@
 
-*A lecture note is added in the tables and more explanations are added to the text*
+*A lecture note is added in the table and more explanations are added to the text.*
 
@@
 
-*Quarterly series with simulated data are added in supplemental material, a new section was added to discuss the use of the methods with different periodicities*
+*Quarterly series with simulated data are added in supplemental material, a new section (just before the conclusion) was added to discuss the use of the methods with different periodicities.*
 
 # Referee 4
+
+Thank you for reviewing our paper and thank you for your helpful comments and suggestions.
+Below are our responses (*in italics*) to your comments and suggestions:
 
@@
 
-*We try to clarify the contribution in the abstract, introduction and in the beginning of sections 2 and 3*
+*We try to clarify the contribution in the abstract, introduction and in the beginning of sections 2 and 3.*
 
@@
 
-*A new section was added with an example around Covid19 and some possible solutions for outliers*
+*A new section was added with an example around Covid19 and some possible solutions for outliers.*
 
@@
 
-*The quote was removed*
+*The quote was removed.*

### 2024-05-01 paper/article_jos.qmd
cross_validation() -> cv()

@@
 Since X-13ARIMA-SEATS can only be applied for series with frequency lower than monthly, for higher frequencies (weekly, daily, etc.), there is no reference on the length of the filters to uses.
-The length of the filters can be determined using local polynomial methods, for example minimising a cross-validation statistics (using the function `rjd3filters::cross_validation()`) or with more complex methods as describe by @Loader1999, but it will surely be necessary to use higher bandwidth than for monthly series.
+The length of the filters can be determined using local polynomial methods, for example minimising a cross-validation statistics (using the function `rjd3filters::cv()`) or with more complex methods as describe by @Loader1999, but it will surely be necessary to use higher bandwidth than for monthly series.
 The larger the window, the greater the bias in assuming that the trend is locally of degree 1: the LC method will then be certainly less appropriate than other methods (QL, CQ or DAF).

### 2024-08-13 paper/article_jos.qmd
doi and typos

@@
 This paper examines and compares real-time estimates of the trend-cycle component using moving averages constructed with local polynomial regression.
-It enables to reproduce Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
+It enables the reproduction of Henderson's symmetric and Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment algorithm.
 This paper proposes two extensions of local polynomial filters for real-time trend-cycle estimates: first including a timeliness criterion to minimise the phase shift;
@@
 In particular, this class encompass Henderson's symmetric filter and Musgrave's asymmetric filters.
-The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods.
-Firstly  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points).
-Secondly calibrating asymmetric filters locally, while they are generally calibrated globally using thus assumptions that may be wrong locally.
-Those extensions are implemented in the statistical software R [@logicielr] throw the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
-All the code associated to this paper is available at <https://github.com/XXXX> (anonymized linked before the manuscript is accepted).
+The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods: 
+firstly  adding, in the optimisation process, a timeliness criterion to directly control the phase shift (i.e., the delay in detecting turning points);
+secondly calibrating asymmetric filters locally, while they are generally calibrated globally using assumptions that may be wrong locally.
+Those extensions are implemented in the statistical software R [@logicielr] through the `rjd3filters` package [@rjd3filters] and can therefore be easily used.
+All the code associated to this paper is available at <https://github.com/AQLT/joffstats2024>.
 
@@
 
-Numerous papers describe the definition and the properties of moving averages and linear filters, see for example @ch12HBSA. 
+Numerous papers describe the definition and the properties of moving averages and linear filters; see for example @ch12HBSA. 
 In this section we summarize some of the main results to understand the next sections.
 
-Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such as for all time series $y_t$:
+Let $p$ and $f$ two integers, a moving average $M_{\boldsymbol\theta}$ is defined by a set of coefficients $\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}$ such that for all time series $y_t$:
 $$
@@
 
-To sum up, applying a moving average to a harmonic times series ($y_t=\e^{-i\omega t}$) affects it in in two different ways:
+To sum up, applying a moving average to a harmonic times series ($y_t=\e^{-i\omega t}$) affects it in two different ways:
 
@@
 
+Observed in @fig-graphsmusgrave that for monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified by 10% ($G_{\boldsymbol\theta}(\pi/6)\simeq1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).
+8-month cycles (associated to the irregular) are only reduced by 6% ($G_{\boldsymbol\theta}(\pi/4)\simeq0.94$).
+Note that the Musgrave filter depends on a ratio denoted $I/C$ (see @sec-nonparamreg).
+In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.
+
 ::: {#fig-graphsmusgrave}
@@
 ```
-
-\justifying
-*Lecture: on monthly times series, 12-month cycles (associated to the frequency $2\pi/12=\pi/6$) are amplified by 10% ($G_{\boldsymbol\theta}(\pi/6)\simeq1.1$) and associated turning points are detected with a delay of 1 months ($\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1$).*
-*8-month cycles (associated to the irregular) are only reduced by 6% ($G_{\boldsymbol\theta}(\pi/4)\simeq0.94$).*\
-*Note: the Musgrave filter depends on a ratio denoted $I/C$ (see @sec-nonparamreg).* 
-*In X-11, for monthly series, when the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used, what is used in most cases), the ratio is fixed to 3.5 and this value is used in this figure.*
 
@@
 A similar approach is used in the X-13ARIMA-SEATS seasonal adjustment method: the series is extended by one year (by default) with an ARIMA model to minimise the revisions associated to asymmetric filters.
-Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent from instability over long periods (e.g., outliers or model shifts).
+Since the final filter used for the seasonal adjustment extraction is much longer than one year (in most of the cases, for monthly series, the bandwidth is 78), the input series is just partially extended (and asymmetric filters are still used) to prevent instability over long periods (e.g., outliers or model shifts).
 
@@
 
-Excluding for the moment the local parametrisations of the polynomial filters, the linear-constant polynomial filter (LC) seems to give the best results in terms of delay in the detection of turning points (@fig-graphstpsimul).
+@Fig-graphstpsimul represents box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).
+For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally, 50% of the turning point are detected with a phase shift of at most 4 months.
+For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).
+
+Excluding for the moment the local parametrisations of the polynomial filters, the linear-constant polynomial filter (LC) seems to give the best results in terms of delay in the detection of turning points.
 Performance is relatively close to that obtained by extending the series using an ARIMA model.
@@
 
-\justifying
-*Lecture: this plots represent box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).*\
-*For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally, 50% of the turning point are detected with a phase shift of at most 4 months.*
-*For the direct asymmetric filter (DAF), 25% of the turning are detected with a phase shift of at least 7 months (compares to 5 for the LC method).*
-
 Distribution of phase shift on simulated series.
@@
 
-In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability (@tbl-simulrev).
+Observe @tbl-simulrev the average of the relative deviations with medium variability. 
+The estimates for $q=0$ correspond to the first estimates of the trend-cycle (real-time estimates when 0 point in the future is known), $q=1$ correspond to the second estimates of the trend-cycle (when 1 point in the future is known).
+Since the final estimates are obtain using a symmetric filter of bandwidth 6, $q=6$ corresponds to the final estimates of the trend-cycle (when 6 points in the future are known).
+$MAE_{fe}(0)$ compares the revision between the first estimate ($q=0$) and the last estimate ($q=6$) and $MAE_{ce}(0)$ compares the revision between the first estimate ($q=0$) and the second estimate ($q=1$).
+
+In terms of revisions, the variability of the series has little effect on the respective performances of the different methods, but it does affect the orders of magnitude, which is why the results are presented only for series with medium variability.
 In general, LC filters always minimise revisions (with relatively small effect from the local parametrisation of the filters) and revisions are greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
@@
 
-\justifying
-*Lecture: the estimates for $q=0$ correspond to the first estimates of the trend-cycle (real-time estimates when 0 point in the future is known), $q=1$ correspond to the second estimates of the trend-cycle (when 1 point in the future is known).*
-*Since the final estimates are obtain using a symmetric filter of bandwidth 6, $q=6$ corresponds to the final estimates of the trend-cycle (when 6 points in the future are known).*
-*$MAE_{fe}(0)$ compares the revision between the first estimate ($q=0$) and the last estimate ($q=6$) and $MAE_{ce}(0)$ compares the revision between the first estimate ($q=0$) and the second estimate ($q=1$).*
-
 Average of the relative deviations of the revisions for the different filters on simulated series with medium variability.
 :::
-
-<!-- ```{=tex} -->
-<!-- \begin{table}[!h] -->
-<!-- \caption{\label{tbl-simulrev} Average of the relative deviations of the revisions for the different -->
-<!-- filters on simulated series with medium variability.} -->
-<!-- \centering -->
-<!-- \begin{tabular}{ccccccc} -->
-<!-- \toprule -->
-<!-- Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\ -->
-<!-- \midrule -->
-<!-- \addlinespace[0.3em] -->
-<!-- \multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\right|\right]$}}\\ -->
-<!-- \hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
-<!-- \hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
-<!-- \hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
-<!-- \hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\ -->
-<!-- \hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
-<!-- \hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\ -->
-<!-- \hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\ -->
-<!-- \hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\ -->
-<!-- \hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\ -->
-<!-- \addlinespace[0.3em] -->
-<!-- \multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[ -->
-<!-- \left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\right| -->
-<!-- \right]$}}\\ -->
-<!-- \hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\ -->
-<!-- \hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\ -->
-<!-- \hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\ -->
-<!-- \hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\ -->
-<!-- \hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\ -->
-<!-- \hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\ -->
-<!-- \hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\ -->
-<!-- \hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\ -->
-<!-- \hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\ -->
-<!-- \bottomrule -->
-<!-- \end{tabular} -->
-<!-- \end{table} -->
-<!-- ``` -->
 
@@
 
-\justifying
-*Lecture: the dotted vertical line corresponds to the date of the turning point (February 2001).*
-*The curve "Feb 2001" corresponds to the estimates of the trend-cycle using the data observed until February 2001.*
-*Since a 13-term Henderson filter is used for the final estimates, for the curve "Feb 2001" estimates from September 2001 to February 2001 are intermediate estimates.*
-
 Successive estimates of the trend-cycle in US employment (in logarithms). 
+The dotted vertical line corresponds to the date of the turning point (February 2001).
+The curve "Feb 2001" corresponds to the estimates of the trend-cycle using the data observed until February 2001.
+Since a 13-term Henderson filter is used for the final estimates, for the curve "Feb 2001" estimates from September 2001 to February 2001 are intermediate estimates.
 :::
@@
 
-\justifying
-*Lecture: the dotted vertical line corresponds to the date of the turning point (February 2001).*
-*The curve "Mar 2001" corresponds to the implicit forecasts associated to the asymmetric filters used for the trend-cycle estimations using the data observed until March 2001.*
-
 Implicit forecasts linked to successive estimates of the trend-cycle in US employment (in logarithms) using local polynomial methods.
+The dotted vertical line corresponds to the date of the turning point (February 2001).
+The curve "Mar 2001" corresponds to the implicit forecasts associated to the asymmetric filters used for the trend-cycle estimations using the data observed until March 2001.
 :::
@@
 Since X-13ARIMA-SEATS can only be applied for series with frequency lower than monthly, for higher frequencies (weekly, daily, etc.), there is no reference on the length of the filters to uses.
-The length of the filters can be determined using local polynomial methods, for example minimising a cross-validation statistics (using the function `rjd3filters::cv()`) or with more complex methods as describe by @Loader1999, but it will surely be necessary to use higher bandwidth than for monthly series.
+The length of the filters can be determined using local polynomial methods, for example minimising a statistic such as cross-validation (`rjd3filters::cv()`), CP (`rjd3filters::cp()`) or Rice's T (`rjd3filters::rt()`), or with more complex methods as describe by @Loader1999, but it will surely be necessary to use higher bandwidth than for monthly series.
 The larger the window, the greater the bias in assuming that the trend is locally of degree 1: the LC method will then be certainly less appropriate than other methods (QL, CQ or DAF).
@@
 For the QL method, the local parametrisation seems to slightly reduce the revisions.
-For the LC method, local parametrisation has few impact on the estimates of the trend-cycle but the associated implicit forecasts are not plausible: the final estimate of the slope is biased by the outlier.
+For the LC method, local parametrisation has little impact on the estimates of the trend-cycle but the associated implicit forecasts are not plausible: the final estimate of the slope is biased by the outlier.
 Third, the ARIMA model can lead to unrealistic forecasts.
@@
 
-- Since all the filter studied in this paper are equivalent to a local polynomial model, an external regressor can be added to the linear model to handle the outlier (the matrix $\boldsymbol X$ would then be modified).
-In the case of an additive outlier, this is equivalent to set to 0 the coefficient of the moving average associated to the date of the outlier (using $y_{t-h}$ to $y_{t+h}$ for the trend-cycle estimates, if the outlier is at the date $t$ the central coefficient is set to 0) and renormalise the sum of the coefficients to 1.
+- Since all the filters studied in this paper are equivalent to a local polynomial model, an external regressor can be added to the linear model to handle the outlier (the matrix $\boldsymbol X$ would then be modified).
+In the case of an additive outlier, this is equivalent to setting to 0 the coefficient of the moving average associated to the date of the outlier (using $y_{t-h}$ to $y_{t+h}$ for the trend-cycle estimates, if the outlier is at the date $t$ the central coefficient is set to 0) and renormalise the sum of the coefficients to 1.
 

### 2024-08-18 paper/article_jos.qmd
biblio + clarification

@@
 
-Using local polynomial regression, @proietti2008 developed a general class of symmetric and asymmetric moving averages.
-In particular, this class encompass Henderson's symmetric filter and Musgrave's asymmetric filters.
+In the last decades, numerous authors proposed methods for the construction of alternative moving averages to derive trend-cycle estimates.
+In particular, @trilemmaWMR2019 put forth a model-based approach based on the mean squared error decomposition, while @proietti2008, @dagumbianconcini2008 and @ch15HBSA proposed non-parametric methods based, respectively, on local polynomial regression, Reproducing Kernel Hilbert Space (RKHS) theory and minimising of a weighted sum of moving average quality criteria.
+All of the aforementioned approaches encompass both Henderson's symmetric filter and Musgrave's asymmetric filters.
+More recently, @inseeDTM202401 conducted a comparative analysis of these methods by describing a general unifying framework to derive moving averages; @dagumBianconcini2023 provided statistical tests to assess the main properties of the filters defined in terms of revisions and the detection of turning points.
+
 The aim of this study is to propose two extensions to the @proietti2008 class of filters in order to take account of two drawbacks of those methods: 
@@
 
-<!-- ### Quality indicators for asymmetric moving averages {#subsec:crit-qual} -->
-
-<!-- For asymmetric filters, most of the criteria come from those defined by @ch15HBSA and @trilemmaWMR2019 to build asymmetric filters.  -->
-<!-- They are summarised in @tbl-qc and can be calculated using the `rjd3filters::diagnostic_matrix()` function. -->
-
-<!-- @ch15HBSA propose a general approach to derive linear filters, based on an optimisation problem involving three criteria: *Fidelity* ($F_g$), *Smoothness* ($S_g$) and *Timeliness* ($T_g$). -->
-<!-- Fidelity can be directly linked to the reduction in variance created by the filter, and timeliness to the notion of phase shift, which again we want to be low.  -->
-<!-- This method can be used in R with the function `rjd3filters::fst_filter()`. -->
-
-<!-- @trilemmaWMR2019 propose an approach based on the decomposition of the mean squared error between the symmetric and the asymmetric filter in four quantities: *Accuracy* ($A_w$), *Timeliness* ($T_w$), *Smoothness* ($S_w$) and *Residual* ($R_w$).
--->
-<!-- A simplified version of this method can be used in R with the function `rjd3filters::dfa_filter()`. -->
-
-<!-- ```{r, include=FALSE} -->
-<!-- title <- "Quality criteria for a moving average $\\boldsymbol\\theta=(\\theta_k)_{-p\\leq k\\leq f}$ defined by a gain function $\\rho_{\\boldsymbol\\theta}$ and a phase shift function $\\varphi_{\\boldsymbol\\theta}$." -->
-<!-- footnotes <- c("$X_g$ criteria from Grun-Rehomme, Guggemos, and Ladiray (2018) and $X_w$ criteria from Wildi and McElroy (2019).", -->
-<!--                "$\\rho_s$ and  $\\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.", -->
-<!--                "$h$ e is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\\frac{1}{2(1-\\cos(x))}$)." -->
-<!-- ) -->
-<!-- ``` -->
-
-<!-- ::: {#tbl-qc} -->
-<!-- ```{r} -->
-<!-- #| echo: false -->
-<!-- criteria_mm <- data.frame(Sigle = c("b_c", "b_l", "b_q", -->
-<!--                                     "F_g", "S_g", "T_g", -->
-<!--                                     "A_w", "T_w", "S_w", "R_w"), -->
-<!--                           Description = c(sprintf("%s bias", c("Constant", "Linear", "Quadratic")), -->
-<!--                                           sprintf("%s (Grun-Rehomme et al)", -->
-<!--                                                   c("Variance reduction / Fidelity", "Smoothness", "Timeliness")), -->
-<!--                                           sprintf("%s (Wildi and McElroy)",c("Accuracy", "Timeliness", "Smoothness", "Residual"))), -->
-<!--                           Formule = c("\\sum_{k=-p}^{+f}\\theta_{k}-1", -->
-<!--                                       "\\sum_{k=-p}^{+f}k\\theta_{k}", -->
-<!--                                       "\\sum_{k=-p}^{+f}k^{2}\\theta_{k}", -->
-<!--                                       "\\sum_{k=-p}^{+f}\\theta_{k}^{2}", -->
-<!--                                       "\\sum_{j}(\\nabla^{3}\\theta_{j})^{2}", -->
-<!--                                       "\\int_{0}^{\\omega_1}\\rho_{\\boldsymbol\\theta}(\\omega)\\sin(\\varphi_{\\boldsymbol\\theta}(\\omega))^{2}\\ud\\omega", -->
-<!--                                       "2\\int_0^{\\omega_1}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega", -->
-<!--                                       "8\\int_0^{\\omega_1} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega", -->
-<!--                                       "2\\int_{\\omega_1}^{\\pi}\\left(\\rho_{s}(\\omega)-\\rho_{\\boldsymbol\\theta}(\\omega)\\right)^{2}h(\\omega)\\ud\\omega", -->
-<!--                                       "8\\int_{\\omega_1}^{\\pi} \\rho_{s}(\\omega)\\rho_{\\boldsymbol\\theta}(\\omega)\\sin^{2}\\left(\\frac{\\varphi_s(\\omega)-\\varphi_{\\boldsymbol\\theta}(\\omega)}{2}\\right)h(\\omega)\\ud\\omega")) -->
-<!-- criteria_mm[,1] <- sprintf("$%s$", criteria_mm[,1]) -->
-<!-- criteria_mm[,3] <- sprintf("$%s$", criteria_mm[,3]) -->
-
-<!-- criteria_mm %>% -->
-<!--   kbl(align = "c", booktabs = T, row.names = FALSE, -->
-<!--         escape = FALSE,caption = title)  %>% -->
-<!-- kable_styling(latex_options = "scale_down")  -->
-<!--     # add_footnote_kable(stop_centering = TRUE) -->
-<!-- ``` -->
-
-<!-- *Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).* -->
-<!-- *$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.* -->
-<!-- *$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).* -->
-
-<!-- Quality criteria for a moving average $\boldsymbol\theta=(\theta_k)_{-p\leq k\leq f}$ defined by a gain function $\rho_{\boldsymbol\theta}$ and a phase shift function $\varphi_{\boldsymbol\theta}$.
--->
-<!-- ::: -->
-
 # Non-parametric regression and local polynomial regression {#sec-nonparamreg}
@@
 
-- If the ratio is large ($3.5< R$) then a 23-term filter is used (to remove more noise).
+- If the ratio is large ($3.5< R$) then a 23-term filter is used (to remove more noise) and the ratio $R=4.5$ is used in X-11 to define the Musgrave filter.
 
-- If the ratio is small ($R<1$) a 9-term filter is used.
+- If the ratio is small ($R<1$) a 9-term filter is used and the ratio $R=1$ is used in X-11 to define the Musgrave filter.
 
-- Otherwise (most of the cases) a 13-term filter is used.
+- Otherwise (most of the cases) a 13-term filter is used and the ratio $R=3.5$ is used in X-11 to define the Musgrave filter.
 
@@
 Supplemental materials shows the coefficients, gain and phase shift functions of the four asymmetric filters.
-
-<!-- Table\ \ref{tab-criteriaLp} compares the quality criteria of the different methods using the Henderson filter and $h=6$ (a symmetric filter with 13 terms).  -->
-<!-- For real-time filters ($q=0$), a more complex asymmetric filter (in terms of polynomial preservation) results in lower timeliness and higher fidelity/smoothness: reduction in phase shift comes at the expense of an increase in variance.
--->
-<!-- This result varies as $q$ increases: for $q=2$, the QL filter has greater timeliness than the LC filter.  -->
-<!-- This surprising result highlights the fact that the phase shift is not controlled by @proietti2008's approach. -->
-
-<!-- In terms of revision ($A_w+S_w+T_w+R_w$), the LC and QL filters consistently outperform the CQ and DAF filters.  -->
-
-<!-- These theoretical properties are consistent with the empirical results observed in @sec-comparison.
-Revisions are more important for the CQ and DAF filters, which leads to greater variability in estimations and a longer delay in detecting turning points.
--->
-<!-- The globally parametrised QL filter (with $R$ fixed for all asymmetric filters) may result in later detection of turning points than the LC filter. -->
-<!-- Indeed, to identify a turning point at date $t$, it is necessary to have knowledge of at least two points after the given date to confirm the reversal in trend. -->
-
-<!-- ::: {#tbl-criteraLp} -->
-<!-- ```{r echo = FALSE} -->
-<!-- #| echo: false -->
-<!-- lp_diagnostics <- readRDS("data/lp_diagnostics_henderson.RDS") -->
-<!-- lp_diagnostics[,"$ EQM_w $"] <- rowSums(lp_diagnostics[,grep("_w", colnames(lp_diagnostics), fixed = TRUE)]) -->
-<!-- colnames(lp_diagnostics) <- gsub(" ?\\$ ?","$",colnames(lp_diagnostics)) -->
-<!-- lp_diagnostics[,1] <- gsub(" ?\\$ ?","$",lp_diagnostics[,1]) -->
-<!-- groupement <- table(lp_diagnostics[,1]) -->
-<!-- lp_diagnostics[,-(1:2)] <- round(lp_diagnostics[,-(1:2)],2) -->
-<!-- lp_diagnostics[,-1] %>% -->
-<!--   kable(format.args = list(digits = 2, -->
-<!--                            decimal.mark = "."), align = "c", booktabs = T, row.names = FALSE, -->
-<!--         escape = FALSE,caption = title,format = "latex") %>% -->
-<!--   kable_styling(latex_options=c(#"striped", -->
-<!--       "hold_position")) %>% -->
-<!--   pack_rows(index = groupement, escape = FALSE) %>% -->
-<!--   footnote("With $EQM_w=A_w + S_w + T_w + R_w$.") -->
-<!-- ``` -->
-
-<!-- Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.5$. -->
-<!-- ::: -->
-
-<!-- ```{=tex} -->
-<!-- \begin{table}[!h] -->
-<!-- \caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3.55$.} -->
-<!-- \centering -->
-<!-- \begin{tabular}[t]{cccccccccccc} -->
-<!-- \toprule -->
-<!-- Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\ -->
-<!-- \midrule -->
-<!-- \addlinespace[0.3em] -->
-<!-- \multicolumn{12}{l}{\textbf{$q=0$}}\\ -->
-<!-- \hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\ -->
-<!-- \hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\ -->
-<!-- \hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\ -->
-<!-- \hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\ -->
-<!-- \addlinespace[0.3em] -->
-<!-- \multicolumn{12}{l}{\textbf{$q=1$}}\\ -->
-<!-- \hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\ -->
-<!-- \hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\ -->
-<!-- \hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\ -->
-<!-- \hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\ -->
-<!-- \addlinespace[0.3em] -->
-<!-- \multicolumn{12}{l}{\textbf{$q=2$}}\\ -->
-<!-- \hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\ -->
-<!-- \hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\ -->
-<!-- \hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\ -->
-<!-- \hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\ -->
-<!-- \bottomrule -->
-<!-- \end{tabular} -->
-
-<!-- \textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.} -->
-<!-- \end{table} -->
-<!-- ``` -->
 
@@
 The DAF method then simply allows us to calculate the associated asymmetric moving averages.
+To avoid unrealistic estimates of $|\delta/\sigma|=2/(|R|\sqrt{\pi})$ in real-time, with $R$ the I-C ratio defined in @sec-lppasymf, the value of $|R|$ is truncated to a maximum of 12 (the higher value used in X-11 being 4.5).
 For the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).

### 2024-08-18 R_simul/1C_Est_final_slope_nn.R
doc

@@ -1,12 +1,7 @@
 library(AQLThesis)
 library(rjd3filters)
-if(!dir.exists("data_simul/byseriespente_daf"))
-  dir.create("data_simul/byseriespente_daf")
-
-# Dans ce programme, pour paramétrer les méthodes LC et QL :
-# 1. On prend les MM asymétriques d'estimation de la pente et polynôme degré 2 et on fait une estimation en temps réel
-# Pour chaque date on a donc h+1=6+1 moyennes mobiles
-# 2. la variance est estimée à chaque date sur les données connues
+if(!dir.exists("data_simul/byseriespente_nn"))
+  dir.create("data_simul/byseriespente_nn")
 
 X_gen <- function(d = 1, p = 6, q = p){
   sapply(0:d, function(exp) seq(-p, q)^exp)
@@ -30,37 +25,35 @@ gen_MM <- function(p=6, q=p, d=2){
 }
 
 MM = lapply(0:6, function(q){
-  d2 = gen_MM(p=6, q=q, d=2)
-  d3 = gen_MM(p=6, q=q, d=3)
-  add_0 <- matrix(0, ncol = ncol(d2), nrow = 6-q)
-  d2 = rbind(d2, add_0)
-  d3 = rbind(d3, add_0)
+  d2 = gen_MM(p=12-q, q=q, d=2)
+  d3 = gen_MM(p=12-q, q=q, d=3)
+  
   list(pente = list(
-    `d=2` = d2[,2],
-    `d=3` = d3[,2]),
-    `deriv2` = list(
-      `d=2` = d2[,3],
-      `d=3` = d3[,3]))
+    `d=2` = moving_average(d2[,2], -(12-q)) ,
+    `d=3` = moving_average(d3[,2], -(12-q))
+  ),
+  `deriv2` = list(
+    `d=2` = moving_average(d2[,3], -(12-q)),
+    `d=3` = moving_average(d3[,3], -(12-q))
+  ))
 })
 MM = list(pente = list(
-  `d=2` = sapply(MM, function(x) x[[1]][[1]]),
-  `d=3` = sapply(MM, function(x) x[[1]][[2]])
+  `d=2` = finite_filters(MM[[7]][[1]][[1]], sapply(MM[-7], function(x) x[[1]][[1]]), first_to_last = TRUE),
+  `d=3` = finite_filters(MM[[7]][[1]][[2]], sapply(MM[-7], function(x) x[[1]][[2]]), first_to_last = TRUE)
 ),
[…]

### 2024-08-18 R_simul/1D_Est_real_time_slope.R

@@ -1,12 +1,7 @@
 library(AQLThesis)
 library(rjd3filters)
-if(!dir.exists("data_simul/byseriespente_nn"))
-  dir.create("data_simul/byseriespente_nn")
-
-# Dans ce programme, pour paramétrer les méthodes LC et QL :
-# 1. On prend les MM asymétriques d'estimation de la pente et polynôme degré 2 et on fait une estimation en temps réel
-# Pour chaque date on a donc h+1=6+1 moyennes mobiles
-# 2. la variance est estimée à chaque date sur les données connues
+if(!dir.exists("data_simul/byseriespente_daf"))
+  dir.create("data_simul/byseriespente_daf")
 
 X_gen <- function(d = 1, p = 6, q = p){
   sapply(0:d, function(exp) seq(-p, q)^exp)
@@ -30,35 +25,37 @@ gen_MM <- function(p=6, q=p, d=2){
 }
 
 MM = lapply(0:6, function(q){
-  d2 = gen_MM(p=12-q, q=q, d=2)
-  d3 = gen_MM(p=12-q, q=q, d=3)
-  
+  d2 = gen_MM(p=6, q=q, d=2)
+  d3 = gen_MM(p=6, q=q, d=3)
+  add_0 <- matrix(0, ncol = ncol(d2), nrow = 6-q)
+  d2 = rbind(d2, add_0)
+  d3 = rbind(d3, add_0)
   list(pente = list(
-    `d=2` = moving_average(d2[,2], -(12-q)) ,
-    `d=3` = moving_average(d3[,2], -(12-q))
-  ),
-  `deriv2` = list(
-    `d=2` = moving_average(d2[,3], -(12-q)),
-    `d=3` = moving_average(d3[,3], -(12-q))
-  ))
+    `d=2` = d2[,2],
+    `d=3` = d3[,2]),
+    `deriv2` = list(
+      `d=2` = d2[,3],
+      `d=3` = d3[,3]))
 })
 MM = list(pente = list(
-  `d=2` = finite_filters(MM[[7]][[1]][[1]], sapply(MM[-7], function(x) x[[1]][[1]]), first_to_last = TRUE),
-  `d=3` = finite_filters(MM[[7]][[1]][[2]], sapply(MM[-7], function(x) x[[1]][[2]]), first_to_last = TRUE)
+  `d=2` = sapply(MM, function(x) x[[1]][[1]]),
+  `d=3` = sapply(MM, function(x) x[[1]][[2]])
 ),
[…]

### 2024-08-18 R_simul/1E_Example_slope_concavity.R

@@ -1,7 +1,7 @@
 # # To install rjd3filters
-# remotes::install_github("rjdemetra/rjd3toolkit")
-# remotes::install_github("rjdemetra/rjd3x11plus")
-# remotes::install_github("rjdemetra/rjd3filters")
+# remotes::install_github("rjdverse/rjd3toolkit")
+# remotes::install_github("rjdverse/rjd3x11plus")
+# remotes::install_github("rjdverse/rjd3filters")
 library(rjd3filters)
 library(ggplot2)
 library(patchwork)

### 2024-08-22 paper/article_jos.qmd
final version

@@
 For example, for the LC method we can use the moving average $\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}$ to obtain a local estimate of the slope and for the QL method we can use $\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\transp{\boldsymbol X}\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}$ to obtain a local estimate of the concavity.
-The DAF method then simply allows us to calculate the associated asymmetric moving averages.
-To avoid unrealistic estimates of $|\delta/\sigma|=2/(|R|\sqrt{\pi})$ in real-time, with $R$ the I-C ratio defined in @sec-lppasymf, the value of $|R|$ is truncated to a maximum of 12 (the higher value used in X-11 being 4.5).
+The DAF method then simply allows us to calculate the associated asymmetric moving averages.\
+To avoid unrealistic estimates of $|\delta/\sigma|=2/(|R|\sqrt{\pi})$ in real-time, where $R$ is the I-C ratio defined in @sec-lppasymf, the value of $|R|$ is truncated to a maximum of 12 (the higher value used in X-11 being 4.5), which slighly reduces revisions and phase shift.\
 For the construction of moving averages, the trend can be modelled as being locally of degree 2 or 3 (this has no impact on the final estimate of concavity).
@@
 There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series using `rjd3filters`.
+See for example the file `Simplified_example.R` available at <https://github.com/AQLT/joffstats2024>.
 

### 2024-08-22 paper/article_jos.tex
final version

@@
 component using moving averages constructed with local polynomial
-regression. It enables to reproduce Henderson's symmetric and Musgrave's
-asymmetric filters used in the X-13ARIMA-SEATS seasonal adjustment
-algorithm. This paper proposes two extensions of local polynomial
-filters for real-time trend-cycle estimates: first including a
-timeliness criterion to minimise the phase shift; second with procedure
-for parametrising asymmetric filters locally while they are generally
-parametrised globally, which can be suboptimal around turning points. An
-empirical comparison, based on simulated and real data, shows that
-modelling polynomial trends that are too complex introduces more
+regression. It enables the reproduction of Henderson's symmetric and
+Musgrave's asymmetric filters used in the X-13ARIMA-SEATS seasonal
+adjustment algorithm. This paper proposes two extensions of local
+polynomial filters for real-time trend-cycle estimates: first including
+a timeliness criterion to minimise the phase shift; second with
+procedure for parametrising asymmetric filters locally while they are
+generally parametrised globally, which can be suboptimal around turning
+points. An empirical comparison, based on simulated and real data, shows
+that modelling polynomial trends that are too complex introduces more
 revisions without reducing the phase shift, and that local
@@
 
-Using local polynomial regression, \textcite{proietti2008} developed a
-general class of symmetric and asymmetric moving averages. In
-particular, this class encompass Henderson's symmetric filter and
-Musgrave's asymmetric filters. The aim of this study is to propose two
-extensions to the \textcite{proietti2008} class of filters in order to
-take account of two drawbacks of those methods. Firstly adding, in the
-optimisation process, a timeliness criterion to directly control the
-phase shift (i.e., the delay in detecting turning points). Secondly
-calibrating asymmetric filters locally, while they are generally
-calibrated globally using thus assumptions that may be wrong locally.
-Those extensions are implemented in the statistical software R
-\autocite{logicielr} throw the \texttt{rjd3filters} package
-\autocite{rjd3filters} and can therefore be easily used. All the code
-associated to this paper is available at \url{https://github.com/XXXX}
-(anonymized linked before the manuscript is accepted).
+In the last decades, numerous authors proposed methods for the
+construction of alternative moving averages to derive trend-cycle
+estimates. In particular, \textcite{trilemmaWMR2019} put forth a
+model-based approach based on the mean squared error decomposition,
+while \textcite{proietti2008}, \textcite{dagumbianconcini2008} and
+\textcite{ch15HBSA} proposed non-parametric methods based, respectively,
+on local polynomial regression, Reproducing Kernel Hilbert Space (RKHS)
+theory and minimising of a weighted sum of moving average quality
+criteria. All of the aforementioned approaches encompass both
+Henderson's symmetric filter and Musgrave's asymmetric filters. More
+recently, \textcite{inseeDTM202401} conducted a comparative analysis of
+these methods by describing a general unifying framework to derive
+moving averages; \textcite{dagumBianconcini2023} provided statistical
+tests to assess the main properties of the filters defined in terms of
+revisions and the detection of turning points.
+
+The aim of this study is to propose two extensions to the
+\textcite{proietti2008} class of filters in order to take account of two
+drawbacks of those methods: firstly adding, in the optimisation process,
+a timeliness criterion to directly control the phase shift (i.e., the
+delay in detecting turning points); secondly calibrating asymmetric
+filters locally, while they are generally calibrated globally using
+assumptions that may be wrong locally. Those extensions are implemented
+in the statistical software R \autocite{logicielr} through the
+\texttt{rjd3filters} package \autocite{rjd3filters} and can therefore be
+easily used. All the code associated to this paper is available at
+\url{https://github.com/AQLT/joffstats2024}.
 
@@
 Numerous papers describe the definition and the properties of moving
-averages and linear filters, see for example \textcite{ch12HBSA}. In
+averages and linear filters; see for example \textcite{ch12HBSA}. In
 this section we summarize some of the main results to understand the
@@
 \(\boldsymbol\theta=\transp{\begin{pmatrix}\theta_{-p}&\cdots&\theta_{f}\end{pmatrix}}\)
-such as for all time series \(y_t\): \[
+such that for all time series \(y_t\): \[
 M_{\boldsymbol\theta}(y_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
@@
 To sum up, applying a moving average to a harmonic times series
-(\(y_t=\e^{-i\omega t}\)) affects it in in two different ways:
+(\(y_t=\e^{-i\omega t}\)) affects it in two different ways:
 
@@
 
+Observed in figure~\ref{fig-graphsmusgrave} that for monthly times
+series, 12-month cycles (associated to the frequency \(2\pi/12=\pi/6\))
+are amplified by 10\% (\(G_{\boldsymbol\theta}(\pi/6)\simeq1.1\)) and
+associated turning points are detected with a delay of 1 months
+(\(\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1\)). 8-month
+cycles (associated to the irregular) are only reduced by 6\%
+(\(G_{\boldsymbol\theta}(\pi/4)\simeq0.94\)). Note that the Musgrave
+filter depends on a ratio denoted \(I/C\) (see
+section~\ref{sec-nonparamreg}). In X-11, for monthly series, when the
+bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is used,
+what is used in most cases), the ratio is fixed to 3.5 and this value is
+used in this figure.
+
 \begin{figure}[H]
@@
 \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}
-
-\justifying
-
-\emph{Lecture: on monthly times series, 12-month cycles (associated to
-the frequency \(2\pi/12=\pi/6\)) are amplified by 10\%
-(\(G_{\boldsymbol\theta}(\pi/6)\simeq1.1\)) and associated turning
-points are detected with a delay of 1 months
-(\(\varphi_{\boldsymbol\theta}(\pi/6)/(\pi/6)\simeq -1\)).}
-\emph{8-month cycles (associated to the irregular) are only reduced by
-6\% (\(G_{\boldsymbol\theta}(\pi/4)\simeq0.94\)).}\\
-\emph{Note: the Musgrave filter depends on a ratio denoted \(I/C\) (see
-section~\ref{sec-nonparamreg}).} \emph{In X-11, for monthly series, when
-the bandwidth is 6 (i.e.: a symmetric Henderson filter of order 13 is
-used, what is used in most cases), the ratio is fixed to 3.5 and this
-value is used in this figure.}
 
@@
   the bandwidth is 78), the input series is just partially extended (and
-  asymmetric filters are still used) to prevent from instability over
-  long periods (e.g., outliers or model shifts).
+  asymmetric filters are still used) to prevent instability over long
+  periods (e.g., outliers or model shifts).
 \end{enumerate}
@@
   If the ratio is large (\(3.5< R\)) then a 23-term filter is used (to
-  remove more noise).
-\item
-  If the ratio is small (\(R<1\)) a 9-term filter is used.
-\item
-  Otherwise (most of the cases) a 13-term filter is used.
+  remove more noise) and the ratio \(R=4.5\) is used in X-11 to define
+  the Musgrave filter.
+\item
+  If the ratio is small (\(R<1\)) a 9-term filter is used and the ratio
+  \(R=1\) is used in X-11 to define the Musgrave filter.
+\item
+  Otherwise (most of the cases) a 13-term filter is used and the ratio
+  \(R=3.5\) is used in X-11 to define the Musgrave filter.
 \end{itemize}
@@
   averages.\\
+  To avoid unrealistic estimates of
+  \(|\delta/\sigma|=2/(|R|\sqrt{\pi})\) in real-time, where \(R\) is the
+  I-C ratio defined in section~\ref{sec-lppasymf}, the value of \(|R|\)
+  is truncated to a maximum of 12 (the higher value used in X-11 being
+  4.5), which slighly reduces revisions and phase shift.\\
   For the construction of moving averages, the trend can be modelled as
@@
 moving averages but they can be easily computed using matrix computation
-formula and applied to the series using \texttt{rjd3filters}.
+formula and applied to the series using \texttt{rjd3filters}. See for
+example the file \texttt{Simplified\_example.R} available at
+\url{https://github.com/AQLT/joffstats2024}.
 
@@
 
+Figure~\ref{fig-graphstpsimul} represents box plots of the phase shift:
+the box represents the interquartile range, the central horizontal line
+the median the dot the outliers and the whisker the range of the data
+(if no outlier). For simulated series with medium variability (when a
+symmetric filter of length 13 is relevant), with the linear constant
+(LC) method, 50\% of the turning point are detected with a phase shift
+of at most 5 months whereas when the LC filter is parametrised locally,
+50\% of the turning point are detected with a phase shift of at most 4
+months. For the direct asymmetric filter (DAF), 25\% of the turning are
+detected with a phase shift of at least 7 months (compares to 5 for the
+LC method).
+
 Excluding for the moment the local parametrisations of the polynomial
 filters, the linear-constant polynomial filter (LC) seems to give the
-best results in terms of delay in the detection of turning points
-(figure~\ref{fig-graphstpsimul}). Performance is relatively close to
-that obtained by extending the series using an ARIMA model. However,
-when variability is low, the LC filter seems to give poorer results and
-the quadratic-linear polynomial filter (QL) seems to give the best
-results.
+best results in terms of delay in the detection of turning points.
+Performance is relatively close to that obtained by extending the series
+using an ARIMA model. However, when variability is low, the LC filter
+seems to give poorer results and the quadratic-linear polynomial filter
+(QL) seems to give the best results.
 
@@
 
-\justifying
-
-\emph{Lecture: this plots represent box plots of the phase shift: the
-box represents the interquartile range, the central horizontal line the
-median the dot the outliers and the whisker the range of the data (if no
-outlier).}\\
-\emph{For simulated series with medium variability (when a symmetric
-filter of length 13 is relevant), with the linear constant (LC) method,
-50\% of the turning point are detected with a phase shift of at most 5
-months whereas when the LC filter is parametrised locally, 50\% of the
-turning point are detected with a phase shift of at most 4 months.}
-\emph{For the direct asymmetric filter (DAF), 25\% of the turning are
-detected with a phase shift of at least 7 months (compares to 5 for the
-LC method).}
-
 }
@@
 \end{figure}%
+
+Observe table~\ref{tbl-simulrev} the average of the relative deviations
+with medium variability. The estimates for \(q=0\) correspond to the
+first estimates of the trend-cycle (real-time estimates when 0 point in
+the future is known), \(q=1\) correspond to the second estimates of the
+trend-cycle (when 1 point in the future is known). Since the final
+estimates are obtain using a symmetric filter of bandwidth 6, \(q=6\)
+corresponds to the final estimates of the trend-cycle (when 6 points in
+the future are known). \(MAE_{fe}(0)\) compares the revision between the
+first estimate (\(q=0\)) and the last estimate (\(q=6\)) and
+\(MAE_{ce}(0)\) compares the revision between the first estimate
+(\(q=0\)) and the second estimate (\(q=1\)).
 
@@
 affect the orders of magnitude, which is why the results are presented
-only for series with medium variability (table~\ref{tbl-simulrev}). In
-general, LC filters always minimise revisions (with relatively small
-effect from the local parametrisation of the filters) and revisions are
-greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
+only for series with medium variability. In general, LC filters always
+minimise revisions (with relatively small effect from the local
+parametrisation of the filters) and revisions are greater with
+cubic-quadratic (CQ) and direct (DAF) polynomial filters.
 
@@
 
-\justifying
-
-\emph{Lecture: the estimates for \(q=0\) correspond to the first
-estimates of the trend-cycle (real-time estimates when 0 point in the
-future is known), \(q=1\) correspond to the second estimates of the
-trend-cycle (when 1 point in the future is known).} \emph{Since the
-final estimates are obtain using a symmetric filter of bandwidth 6,
-\(q=6\) corresponds to the final estimates of the trend-cycle (when 6
-points in the future are known).} \emph{\(MAE_{fe}(0)\) compares the
-revision between the first estimate (\(q=0\)) and the last estimate
-(\(q=6\)) and \(MAE_{ce}(0)\) compares the revision between the first
-estimate (\(q=0\)) and the second estimate (\(q=1\)).}
-
 }
@@
 \caption{\label{fig-ce16ovlp}Successive estimates of the trend-cycle in
-US employment (in logarithms).}
-
-\centering{
-
-\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}
-
-\justifying
-
-\emph{Lecture: the dotted vertical line corresponds to the date of the
-turning point (February 2001).} \emph{The curve ``Feb 2001'' corresponds
-to the estimates of the trend-cycle using the data observed until
-February 2001.} \emph{Since a 13-term Henderson filter is used for the
+US employment (in logarithms). The dotted vertical line corresponds to
+the date of the turning point (February 2001). The curve ``Feb 2001''
+corresponds to the estimates of the trend-cycle using the data observed
+until February 2001. Since a 13-term Henderson filter is used for the
 final estimates, for the curve ``Feb 2001'' estimates from September
 2001 to February 2001 are intermediate estimates.}
+
+\centering{
+
+\includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}
 
@@
 successive estimates of the trend-cycle in US employment (in logarithms)
-using local polynomial methods.}
+using local polynomial methods. The dotted vertical line corresponds to
+the date of the turning point (February 2001). The curve ``Mar 2001''
+corresponds to the implicit forecasts associated to the asymmetric
+filters used for the trend-cycle estimations using the data observed
+until March 2001.}
 
@@
 \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}
-
-\justifying
-
-\emph{Lecture: the dotted vertical line corresponds to the date of the
-turning point (February 2001).} \emph{The curve ``Mar 2001'' corresponds
-to the implicit forecasts associated to the asymmetric filters used for
-the trend-cycle estimations using the data observed until March 2001.}
 
@@
 filters can be determined using local polynomial methods, for example
-minimising a cross-validation statistics (using the function
-\texttt{rjd3filters::cross\_validation()}) or with more complex methods
-as describe by \textcite{Loader1999}, but it will surely be necessary to
-use higher bandwidth than for monthly series. The larger the window, the
-greater the bias in assuming that the trend is locally of degree 1: the
-LC method will then be certainly less appropriate than other methods
-(QL, CQ or DAF). Moreover, since the bias is higher, the local
+minimising a statistic such as cross-validation
+(\texttt{rjd3filters::cv()}), CP (\texttt{rjd3filters::cp()}) or Rice's
+T (\texttt{rjd3filters::rt()}), or with more complex methods as describe
+by \textcite{Loader1999}, but it will surely be necessary to use higher
+bandwidth than for monthly series. The larger the window, the greater
+the bias in assuming that the trend is locally of degree 1: the LC
+method will then be certainly less appropriate than other methods (QL,
+CQ or DAF). Moreover, since the bias is higher, the local
 parametrisation of the filters will surely have more impact.
@@
 parametrisation seems to slightly reduce the revisions. For the LC
-method, local parametrisation has few impact on the estimates of the
+method, local parametrisation has little impact on the estimates of the
 trend-cycle but the associated implicit forecasts are not plausible: the
@@
 \item
-  Since all the filter studied in this paper are equivalent to a local
+  Since all the filters studied in this paper are equivalent to a local
   polynomial model, an external regressor can be added to the linear
@@
   be modified). In the case of an additive outlier, this is equivalent
-  to set to 0 the coefficient of the moving average associated to the
-  date of the outlier (using \(y_{t-h}\) to \(y_{t+h}\) for the
+  to setting to 0 the coefficient of the moving average associated to
+  the date of the outlier (using \(y_{t-h}\) to \(y_{t+h}\) for the
   trend-cycle estimates, if the outlier is at the date \(t\) the central

### 2024-08-22 paper/resp_referee2.qmd
final version

---
title: "Response to referee (JOS-2023-0137.R1)"
format: 
  pdf:
    geometry:
      - top=2cm
      - bottom=2cm
      - left = 2.5cm
      - right = 2.5cm
lang: en-GB  
---

We would like to thanks all the reviewers and the associate editor for the comments, suggestions and the time spent proofreading the paper.

# Associated Editor

All the comments and suggestions were taken into account and the paper was improved accordingly.
The "Lecture" was also removed for the table and put in the text.

# Referee 3

Thank you for reviewing our paper.
However, it seems that you might have read the previous version of the paper since the reference to the pages/lines are not valid, the title has been changed and almost all the detailed comments have been adressed in the revised version of the paper.
Below are our responses (*in italics*) to your comments and suggestions.

## General comments

> While the manuscript is comprehensive and methodologically sound, some sections could benefit from additional clarification to enhance readability and understanding. 
The literature review should be expanded to better contextualize this work within the broader field of time series analysis and trend estimation. 
Additionally, validating key assumptions, such as the proportional odds assumption, needs to be more thoroughly addressed.
Comparative analysis with existing methods would provide a more balanced view of the proposed approach's strengths and limitations.

*The manuscript was simplified and clarified and a literature review has been added in the introduction.*
*The proposed approach was compared to the classical parametrisation of local polynomial filters (the "global parametrisation" corresponds to the current usual approach, the LC method corresponds to the Musgrave filter used in the X-11 decomposition method).*
*The proportional odds assumption (same slope in all the sample for the LC/Musgrave filter, same concavity for the QL filter) is the usual assumption made in the current methods but the aim of the proposed approach is to release this assumption (local parametrisation).*
*Even if the proposed approach could be compared to other recent methods mentioned in the bibliography, it would have complexify the readability of the paper (and considerably increased the number of pages, which we believe would not be relevent for a research paper).*
*Therefore, we decided to limit the comparison to the classical parametrisation used in X-11 (Musgrave=LC with "global parametrisation") and to the other local polynomial filters (filters for which we proposed extensions).*
*However, we have added in the bibliography a working paper with a broader comparison of the existing methods and their implementation in `rjd3filters`.*

## Detailed Comments

> 1. The introduction effectively outlines the importance of real-time trend estimation but could provide a broader overview of related methods and how this research fits within the existing literature.

*The following has been added to the introduction, mentioning the recent literature trend-cycle estimates using moving averages:*\
*In the last decades, numerous authors proposed methods for the construction of alternative moving averages to derive trend-cycle estimates. In particular, Wildi and McElroy (2019) put forth a model-based approach based on the mean squared error decomposition, while Proietti and Luati (2008), Dagum and Bianconcini (2008) and Grun-Rehomme, Guggemos, and Ladiray (2018) proposed non-parametric methods based, respectively, on local polynomial regression, Reproducing Kernel Hilbert Space (RKHS) theory and minimising of a weighted sum of moving average quality criteria. All of the aforementioned approaches encompass both Henderson’s symmetric filter and Musgrave’s asymmetric filters. More recently, Quartier-la-Tente (2024) conducted a comparative analysis of these methods by describing a general unifying framework to derive moving averages; Dagum and Bianconcini (2023) provided statistical tests to assess the main properties of the filters defined in terms of revisions and the detection of turning points.*

> 2. The methodology section is robust but dense. More intuitive explanations or graphical representations of key concepts could enhance accessibility for readers.

*Adressed in the first revision of the paper: this section was rewriten in the and simplified to make it clearer (unused sections were removed).*

> 2.  Some symbols and notations, such as the symbol [[ ]] on page 11, lines 53-54, are not clearly defined. Providing explicit definitions or explanations would prevent any ambiguity.

*Adressed in the first revision of the paper: This symbol was referring to integer interval, it was removed to avoid confusion (for example $[[1,d]]$ was replaced by $\{1,2,\dots, d\}$).*

> 4. It would be helpful to explain how phase shift is computed more clearly. More details on the method used for determining phase shift, particularly in relation to the identification of true turning points, would also be beneficial.

*Adressed in the first revision of the paper: The corresponding text was rewritten to clarify the definition of phase-shift used in this paper.*

> 5. Variance Determination: The determination of variance in the local parameterization of asymmetric filters (page 17) could be more detailed. Clearly identifying the quantities appearing in the denominator would improve understanding.

*Adressed in the first revision of the paper: More explanations was added to this paragraph: the formula comes from local regression theory simplified because we are here in the case of local polynomial approximation.*

> 6. Applicability to High-Frequency Data: The manuscript should explicitly address the proposed procedure's applicability to high-frequency data. Discussing any necessary modifications or considerations for such data would provide a more comprehensive understanding of the methodology.

*Adressed in the first revision of the paper: Quarterly series with simulated data are added in supplemental material, a new section was added to discuss the use of the methods with different periodicities (high-frequency).*

# Other modifications

- In section 3.4 (description of the local parameterisation approach), a was added to clarify an implementation choice which wasn't mentionned in the previous version of the paper: for the real time estimates of the slope/concavity, since the moving averages used can lead to a lot of variability (DAF filters), the I/C ratio computed using the slope/concavity + variance estimated was truncated to a maximum of 12 (the higher value used in X-11 being 4.5). The following sentence was added to the paper:\
*To avoid unrealistic estimates of $|\delta/\sigma|=2/(|R|\sqrt{\pi})$ in real-time, where $R$ is the I-C ratio defined in section 3.2, the value of $|R|$ is truncated to a maximum of 12 (the higher value used in X-11 being 4.5), which slighly reduces revisions and phase shift.*

- At the end of section 3.4 (description of the local parameterisation approach), the sentence in italic was added to refer to a specific program available in the code:\
"There is no function in `rjd3filters` to directly compute those moving averages but they can be easily computed using matrix computation formula and applied to the series using `rjd3filters`.
*See for example the file `Simplified_example.R` available at <https://github.com/AQLT/joffstats2024>.*"

- All the URL are now available (including the URL to the code).

- DOI was added for all the article when they are available, when URL is used, the date of access was added, Dates and places of conferences are given for proceedings.

### 2024-08-22 paper/supp_material.qmd
final version

@@
     toc: false
+    keep-tex: true
     include-in-header: [preamble.tex]

### 2024-08-22 paper/supp_material.tex
final version

--- 2024-02-12 article_jos.tex
@@
 \makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
-\makeatother
 \ifLuaTeX
@@
 \fi
-\usepackage[]{biblatex}
-\addbibresource{biblio.bib}
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \hypersetup{
-  pdftitle={Improving real-time trend estimates using local parametrisation of local polynomial regression filters},
+  pdftitle={Supplemental material},
   colorlinks=true,
@@
 
-\title{Improving real-time trend estimates using local parametrisation
-of local polynomial regression filters}
+\title{Supplemental material}
 \author{}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\hypertarget{abstract}{%
-\section*{Abstract}\label{abstract}}
-\addcontentsline{toc}{section}{Abstract}
-
-This paper studies and compares real-time estimates of the trend cycle
-component with moving averages based on local polynomial regression. In
-particular, this theory allows to reproduce Henderson's symmetric and
-Musgrave's asymmetric filters. This paper describes how they can be
-extended to include a timeliness criterion to minimise the phase shift.
-While asymmetric filters are generally parameterised globally, which can
-be suboptimal around turning points, this paper proposes a procedure for
-parameterising them locally. An empirical comparison, based on simulated
-and real data, shows that modelling polynomial trends that are too
-complex introduces more revisions without reducing the phase shift, and
-that local parameterisation reduces the delay in detecting turning
-points and reduces revisions. The paper also shows how these results can
-be easily reproduced using the R package \texttt{rjd3filters} dedicated
-to the manipulation of moving averages.
-
-Keywords: time series, trend-cycle, seasonal adjustment, turning points,
-R statistical software.
-
-JEL Classification: E32, E37.
-
-\hypertarget{introduction}{%
-\section*{Introduction}\label{introduction}}
-\addcontentsline{toc}{section}{Introduction}
-
-Analysis of the economic cycle, and in particular the early detection of
-turning points in a series, is a major topic in the analysis of economic
-outlook. To this end, economic indicators are generally seasonally
-adjusted. However, in order to improve their interpretability, it may be
-necessary to perform additional smoothing in order to reduce noise, and
-thus analyse the trend-cycle component. By construction, trend-cycle
-extraction methods are closely related to seasonal adjustment methods,
-since they are generally applied to seasonally adjusted series.
-
-A moving average, or linear filter, is a statistical method that
-consists in applying a rolling weighted mean to a times series: for each
-date \(t\) it computes a weighted mean of \(p\) past points and \(q\)
-future points where \(p,q\geq0\) depends on the moving average. Moving
-averages are ubiquitous in business cycle extraction and seasonal
-adjustment methods. For example, the X-13ARIMA seasonal adjustment
-method uses several moving averages to estimate the main components of a
-time series. Symmetric filters are applied to the center of the series,
-but when it comes to estimate the most recent points, all of these
-methods must rely on asymmetric filters. For trend-cycle extraction, the
-most popular symmetric filter is the \textcite{henderson1916note} moving
-average, which is used in the X-13ARIMA seasonal adjustment algorithm.
-
-However, for real-time estimates, due to the lack of future
-observations, all these methods must rely on asymmetric filters to
-estimate the most recent points. For example, for trend-cycle
-extraction, X-13ARIMA uses Henderson's symmetric filter and
-\textcite{musgrave1964set}'s asymmetric filters on an extended series
-using an ARIMA model. Since the predicted values are linear combinations
-of past values, this amounts to applying asymmetric averages at the end
-of the series.
-
-If the classic asymmetric moving averages have good properties regarding
-the future revisions induced by the process (see for example
-\textcite{pierce1980SA}), they create, by construction, phase shifts
-that impact the real-time estimation of turning points, introducing time
-delay in the detection.
-
-Using local polynomial regression, \textcite{proietti2008} developed
-general class of symmetric and asymmetric moving averages. In
-particular, it makes possible to reproduce Henderson's symmetric filter
-and Musgrave's asymmetric filters. However, these methods have two
-drawbacks. Firstly, the phase shift (i.e.~the delay in detecting turning
-points) is not directly controlled. Secondly, asymmetric filters depend
-on a parameter that is generally estimated over the whole series using
-assumptions that may be wrong locally. The aim of this study is to
-propose extensions to the \textcite{proietti2008} class of filters in
-order to take these shortcomings into account. The extensions have been
-implemented in the \faIcon{r-project} \texttt{rjd3filters} package
-(available at \url{https://github.com/rjdemetra/rjd3filters}) and are
-therefore easily reusable.
-
-In Section~\ref{sec-propMM}, we describe the general properties of
-moving averages and the associated quality criteria. This allows us to
-understand the foundations behind the construction of moving averages
-from local polynomial regressions, as well as those behind the two
-extensions proposed in this article (Section~\ref{sec-nonparamreg}). In
-a final section (Section~\ref{sec-comparison}), all these methods are
-compared empirically on simulated and real series.
-
-\hypertarget{sec-propMM}{%
-\section{Some properties on moving averages}\label{sec-propMM}}
-
-Lots of papers describe the definition and the properties of moving
-averages and linear filters, see for example \textcite{ch12HBSA}. In
-this section we summarize some of the main results to understand the
-next sections.
-
-Let \(p\) et \(f\) two integers, a moving average
-\(M_{\boldsymbol\theta}\) or \(M\) is defined by a set of coefficients
-\(\boldsymbol\theta=\begin{pmatrix}\theta_{-p}\\\vdots\\\theta_{f}\end{pmatrix}\)
-such as for all time series \(X_t\): \[
-M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}.
-\]
-
-\begin{itemize}
-\item
-  \(p+f+1\) is called the \emph{moving average order}.
-\item
-  When \(p=f\) the moving average is said to be \emph{centered}. If we
-  also have \(\forall k:\:\theta_{-k} = \theta_k\), the moving average
-  \(M_{\boldsymbol\theta}\) is said to be \emph{symmetric}. In this
-  case, the quantity \(h=p=f\) is called the \emph{bandwidth}.
-\end{itemize}
-
-\hypertarget{subsec:gain-deph}{%
-\subsection{Gain and phase shift functions}\label{subsec:gain-deph}}
-
-To interpret the notions of gain and phase shift, it is useful to
-illustrate the effects of moving averages on harmonic series
-\(X_t=\e^{-i\omega t}\) with \(\omega\in[0,\pi]\). The moving average
-\(M_{\boldsymbol\theta}\) transforms \(X_t\) into: \[
-Y_t = M_{\boldsymbol\theta}X_t = \sum_{k=-p}^{+f} \theta_k \e^{-i \omega (t+k)}
-= \left(\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\right)\cdot X_t.
-\] The function
-\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
-is called the \emph{transfer function} or \emph{frequency response
-function}. The frequency response function can equivalently be defined
-as
-\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
-ou
-\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
-It can be rewritten as: \[
-\Gamma_{\boldsymbol\theta}(\omega) = \rho_{\boldsymbol\theta}(\omega)\e^{i\varphi_{\boldsymbol\theta}(\omega)},
-\] where
-\(\rho_{\boldsymbol\theta}(\omega)=G_{\boldsymbol\theta}(\omega)=\lvert\Gamma_{\boldsymbol\theta}(\omega)\rvert\)
-is the \emph{gain} or \emph{amplitude} function and
-\(\varphi_{\boldsymbol\theta}(\omega)\) is the \emph{phase shift} or
-\emph{time shift} function. This function is sometimes represented as
-\(\phi_{\boldsymbol\theta}(\omega)=\frac{\varphi_{\boldsymbol\theta}(\omega)}{\omega}\)
-to mesure the phase shift in number of periods. For all symmetric moving
-average we have
-\(\varphi_{\boldsymbol\theta}(\omega)\equiv 0 \;(modulo\;{\pi})\).
-
-To sum up, applying a moving average to a harmonic times series
-(\(X_t=\e^{-i\omega t}\)) affects it in in two different ways:
-
-\begin{itemize}
-\item
-  by multiplying it by an amplitude coefficient
-  \(\rho_{\boldsymbol\theta}\left(\omega\right)\) (gain);
-\item
-  by ``shifting'' it in time by
-  \(\varphi_{\boldsymbol\theta}(\omega)/\omega\) which directly affects
-  the detection of turning points. When
-  \(\varphi_{\boldsymbol\theta}(\omega)/\omega<0\) the time shift is
-  negative: the turning point is detected with delay.
-\end{itemize}
-
-Fourier decomposition allows to analyze any time series as a sum of
-harmonic series, and each component (trend, cycle, seasonal, irregular)
-is associated with a set of frequencies. Noting \(\omega = 2\pi/p\), the
-harmonic series of frequency \(\omega\) represents a series that repeats
-itself every \(p\) periods. For example, for a monthly series (12
-observations per year), seasonal movements are those that repeat every
-year: they are therefore associated with frequencies \(2\pi/12\) (annual
-periodicity), \(2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5\). In this
-paper, we consider that the trend-cycle is associated with frequencies
-in the interval \([0, 2\pi/12[\), i.e.~movements recurring at least
-every 12 months. Even if different frequencies are sometimes used
-(e.g.~\([0, 2\pi/36]\) to consider only cycles of at least 36 months),
-this has no impact on the results of the study. The other frequencies,
-\([2\pi/12, \pi]\) are associated with the irregular component
-(undesirable oscillations).
-
-Figure~\ref{fig-graphsmusgrave} shows the gain and phase shift function
-for the asymmetric Musgrave filter (see Section~\ref{sec-lppasymf})
-often used for real-time trend-cycle estimation (i.e.~when no point in
-the future is known). The gain function is greater than 1 on the
-frequencies associated with the trend-cycle (\([0, 2\pi/12]\)): this
-means that the trend-cycle is well preserved and that short cycles of 1
-to 2 years (\([2\pi/24,2\pi/12]\)) are even amplified. However, cycles
-of 8 to 12 months (\([2\pi/12, 2\pi/8]\)), considered undesirable
-because associated with the irregular, are also amplified: this can lead
-to the detection of false turning points. At other frequencies, the gain
-function is less than 1, but always positive: this means that the series
-smoothed by this moving average will always contain noise, even if it is
-attenuated. Analysis of the phase shift shows that the shorter the
-cycles, the greater the phase shift: this means that on series smoothed
-by this moving average, turning points could be detected at the wrong
-date.
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphsmusgrave}Coefficients, gain and phase shift
-function for the Musgrave filter in real time with \(I/C=3.5\).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/musgrave.pdf}
-
-}
-
-\end{figure}
-
-\hypertarget{desirable-properties-of-a-moving-average}{%
-\subsection{Desirable properties of a moving
-average}\label{desirable-properties-of-a-moving-average}}
-
-To decompose a time series into a seasonal component, a trend-cycle and
-the irregular, the X-11 decomposition algorithm (used in X-13ARIMA) uses
-a succession of moving averages, all with specific constraints.
-
-In our case, we assume that our initial series \(y_t\) is seasonally
-adjusted and can be written as the sum of a trend-cycle, \(TC_t\), and
-an irregular component, \(I_t\): \[
-y_t=TC_t+I_t.
-\] The aim will be to build moving averages that best preserve the
-trend-cycle (\(M_{\boldsymbol\theta} (TC_t)\simeq TC_t\)) and minimise
-noise (\(M_{\boldsymbol\theta} (I_t)\simeq 0\)).
-
-\hypertarget{trend-preservation}{%
-\subsubsection{Trend preservation}\label{trend-preservation}}
-
-Trend-cycles are generally modelled by local polynomial trends (see
-Section~\ref{sec-nonparamreg}). In order to best preserve the
-trend-cycles, we try to have moving averages that preserve the
-polynomial trends. A moving average \(M_{\boldsymbol\theta}\) preserves
-a function of time \(f(t)\) if
-\(\forall t:M_{\boldsymbol\theta} f(t)=f(t)\).
-
-We have the following properties for the moving average
-\(M_{\boldsymbol\theta}\):
-
-\begin{itemize}
-\item
-  To preserve a constant trend \(X_t=a\) we need: \[
-  \forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
-  \] The sum of the coefficients of the moving average
-  \(\sum_{k=-p}^{+f}\theta_k\) must then be equal to \(1\).
-\item
-  To preserve linear trends \(X_t=at+b\) we need: \[
-  \forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
-  \] Which is equivalent to: \[
-  \sum_{k=-p}^{+f}\theta_k=1
-  \quad\text{et}\quad
-  \sum_{k=-p}^{+f}k\theta_k=0.
-  \]
-\item
-  In general, \(M_{\boldsymbol\theta}\) preserves polynomials of degree
-  \(d\) if and only if: \[
-  \sum_{k=-p}^{+f}\theta_k=1
-   \text{ and }
-  \forall j \in \left\llbracket 1,d\right\rrbracket:\:
-  \sum_{k=-p}^{+f}k^j\theta_k=0.
-  \]
-\item
-  If \(M_{\boldsymbol\theta}\) is symmetric (\(p=f\) and
-  \(\theta_{-k} = \theta_k\)) and preserves polynomials of degree \(2d\)
-  then it also preserves polynomials of degree \(2d+1\).
-\end{itemize}
-
-\hypertarget{noise-reduction}{%
-\subsubsection{Noise reduction}\label{noise-reduction}}
-
-All time series are affected by noise, which can blur the extraction of
-trends and cycles. This is why we try to reduce this noise (by reducing
-its variance) while preserving the relevant trends. The sum of the
-squares of the coefficients \(\sum_{k=-p}^{+f}\theta_k^2\) is the
-\emph{reduction of variance} ratio.
-
-Indeed, let \(\{\varepsilon_t\}\) a sequence of independent random
-variables with \(\E{\varepsilon_t}=0\), \(\V{\varepsilon_t}=\sigma^2\):
-\[
-\V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}}
-= \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}=
-\sigma^2\sum_{k=-p}^{+f} \theta_k^2.
-\]
-
-\hypertarget{sec-mmasym}{%
-\subsection{Real-time estimation and asymmetric moving
-average}\label{sec-mmasym}}
-
-\hypertarget{sec-mmetprev}{%
-\subsubsection{Asymmetric moving averages and
-forecasts}\label{sec-mmetprev}}
-
-For symmetric filters, the phase shift function is equal to zero (modulo
-\(\pi\)). So there is no delay in detecting turning points. However,
-they cannot be used in the beginning and in the end of the time series
-because no past or future value can be used. Thus, for real-time
-estimation, it is needed to build asymmetric moving average that
-approximate the symmetric moving average.
-
-Several approaches can be used for real-time estimation:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Use asymmetric moving averages to take account of the lack of
-  available data;
-\item
-  Apply symmetrical filters to extended forecast series. This method
-  seems to date back to \textcite{deforest1877adjustment}, which also
-  suggests modelling a polynomial trend of degree three or less at the
-  end of the period:\\
-  ``\emph{As the first \(m\) and last \(m\) terms of the series cannot
-  be reached directly by the formula, the series should be graphically
-  extended by m terms at both ends, first plotting the observations on
-  paper as ordinates, and then extending the curve along what seems to
-  be its probable course, and measuring the ordinates of the extended
-  portions.} \emph{It is not necessary that this extension should
-  coincide with what would be the true course of the curve in those
-  parts. } \emph{The important point is that the m terms thus added,
-  taken together with the \(m+1\) adjacent given terms, should follow a
-  curve whose form is approximately algebraic and of a degree not higher
-  than the third.}''\\
-  This is also the approach used in the X-13ARIMA seasonal adjustment
-  method, which extends which extends the series by one year with an
-  ARIMA model.
-\end{enumerate}
-
-Ultimately, the second method is equivalent to using asymmetric moving
-averages, since forecasts are linear combinations of the past.
-
-Conversely, the implicit forecasts of an asymmetric moving average can
-be deduced from a reference symmetric moving average. This allows to
-judge the quality of real-time estimates of the trend-cycle and to
-anticipate future revisions when forecasts are far from expected
-evolutions.\\
-Let us denote \(v=(v_{-h},\dots, v_{h})\) the reference symmetrical
-moving average and \(w^0,\dots w^{h-1}\) a sequence of asymmetrical
-moving averages, of order \(h+1\) to \(2h\) used to estimate the last
-\(h\) points with, for convention, \(w_t^q=0\) for \(t>q\). This means
-that \(w^0=(w_{-h}^0,\dots, w_{0}^0)\) is used for estimation in real
-time (when no points are known in the future),
-\(w^1=(w_{-h}^1,\dots, w_{1}^1)\) for estimation of the penultimate
-point (when only one point is known in the future), and so on. Let also
-note \(y_{-h},\dots,y_{0}\) the observed series studied and
-\(y_{1}^*,\dots y_h^*\) the implicit forecast induced by
-\(w^0,\dots w^{h-1}\). This means that for all \(q\) we have: \[
-\forall q, \quad \underbrace{\sum_{i=-h}^0 v_iy_i + \sum_{i=1}^h v_iy_i^*}_{\text{smoothing by }v\text{ of the extended series}}
-=\underbrace{\sum_{i=-h}^0 w_i^qy_i + \sum_{i=1}^h w_i^qy_i^*}_{\text{smoothing by }w^q\text{ of the extended series}}.
-\] This is equivalent to: \[
-\forall q, \quad \sum_{i=1}^h (v_i- w_i^q) y_i^*
-=\sum_{i=-h}^0 (w_i^q-v_i)y_i.
-\] To sum up, in matrix form, this is equivalent to solving: \[
-\scriptstyle
-\begin{pmatrix}
-  v_1 & v_2 & \cdots & v_h \\
-  v_1 - w_1^1 & v_2 & \cdots & v_h \\
-  \vdots & \vdots & \cdots & \vdots \\
-   v_1 - w_1^{h-1} & v_2-w_2^{h-1} & \cdots & v_h
-\end{pmatrix}
-\begin{pmatrix}y_1^* \\ \vdots \\ y_h^*\end{pmatrix}=
-\begin{pmatrix}
-  w_{-h}^0 - v_{-h} & w_{-(h-1)}^0 - v_{-(h-1)} & \cdots & w_{0}^0 - v_{0} \\
-  w_{-h}^1 - v_{-h} & w_{-(h-1)}^1 - v_{-(h-1)} & \cdots & w_{0}^1 - v_{0} \\
-  \vdots & \vdots & \cdots & \vdots \\
-  w_{-h}^{h-1} - v_{-h} & w_{-(h-1)}^{h-1} - v_{-(h-1)} & \cdots & w_{0}^{h-1} - v_{0}
-\end{pmatrix}
-\begin{pmatrix}y_{-h} \\ \vdots \\ y_0\end{pmatrix}.
-\] This is implemented in the \texttt{rjd3filters::implicit\_forecast()}
-function.
-
-As highlighted by \textcite{wildischis2004}, extending the series
-through forecasting with an ARIMA model is equivalent to calculating
-asymmetric filters whose coefficients are optimised in relation to the
-one-step ahead forecast. In other words, the aim is to minimise the
-revisions between the first and last estimates (with the symmetric
-filter). However, the phase shift induced by the asymmetric filters is
-not controlled: we might prefer to have faster detection of turning
-points and a larger revision rather than just minimising the revisions
-between the first and last estimates. Furthermore, since the
-coefficients of the symmetric filter (and therefore the weight
-associated with distant forecasts) decrease slowly, we should also be
-interested in the performance of multi-step ahead forecasting. This is
-why it may be necessary to define alternative criteria for judging the
-quality of asymmetric moving averages.
-
-\hypertarget{subsec:crit-qual}{%
-\subsubsection{Quality indicators for asymmetric moving
-averages}\label{subsec:crit-qual}}
-
-For asymmetric filters, most of the criteria come from those defined by
-\textcite{ch15HBSA} and \textcite{trilemmaWMR2019} to build asymmetric
-filters. They are summarised in Table~\ref{tab-QC} and can be calculated
-using the \texttt{rjd3filters::diagnostic\_matrix()} function.
-
-\textcite{ch15HBSA} propose a general approach to derive linear filters,
-based on an optimisation problem involving three criteria:
-\emph{Fidelity} (\(F_g\)), \emph{Smoothness} (\(S_g\)) and
-\emph{Timeliness} (\(T_g\)). Fidelity can be directly linked to the
-reduction in variance created by the filter, and timeliness to the
-notion of phase shift, which again we want to be low. This method can be
-used in\faIcon{r-project} with the function
-\texttt{rjd3filters::fst\_filter()}.
-
-\textcite{trilemmaWMR2019} propose an approach based on the
-decomposition of the mean squared error between the symmetric and the
-asymmetric filter in four quantities: \emph{Accuracy} (\(A_w\)),
-\emph{Timeliness} (\(T_w\)), \emph{Smoothness} (\(S_w\)) and
-\emph{Residual} (\(R_w\)). A simplified version of this method can be
-used in \faIcon{r-project} with the function
-\texttt{rjd3filters::dfa\_filter()}.
-
-\begin{table}[!h]
-\caption{\label{tab-QC} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
-{\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{ccc}
+\appendix
+
+\section{Coefficients, gain and phase shift functions}\label{sec-an-cof}
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-lc}Coefficients, gain and phase shift
+functions for the Linear-Constant (LC) filter with \(I/C=3.5\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-ql}Coefficients, gain and phase shift
+functions for the Quadratic-Linear (QL) filter with \(I/C=3.5\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-cq}Coefficients, gain and phase shift
+functions for the Cubic-Quadratic (CQ) filter with \(I/C=3.5\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-daf}Coefficients, gain and phase shift
+functions for the direct asymmetric filter (DAF).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
+
+\section{Different specifications for the estimation of the slope and
+concavity}\label{different-specifications-for-the-estimation-of-the-slope-and-concavity}
+
+\begin{figure}
+
+\caption{\label{fig-graphs-lc-deg-final}Distribution of phase shift
+associated to the local parametrisation of the linear-constant (LC)
+filter using the final estimates of \(\delta/\sigma\), by the bandwidth
+\(h\) of the henderson filter used to estimate \(\sigma\) and the degree
+\(d\) of the trend to estimate \(\delta\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_local_final_lc.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-lc-deg}Distribution of phase shift associated
+to the local parametrisation of the linear-constant (LC) filter using
+the final estimates of \(\delta/\sigma\), by the bandwidth \(h\) of the
+henderson filter used to estimate \(\sigma\) and the degree \(d\) of the
+trend to estimate \(\delta\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_local_lc.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-ql-deg-final}Distribution of phase shift
+associated to the local parametrisation of the quadratic-linear (QL)
+filter using real-time estimates of \(\delta/\sigma\), by the bandwidth
+\(h\) of the henderson filter used to estimate \(\sigma\) and the degree
+\(d\) of the trend to estimate \(\delta\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_local_final_ql.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-ql-deg}Distribution of phase shift associated
+to the local parametrisation of the quadratic-linear (QL) filter using
+real-time estimates of \(\delta/\sigma\), by the bandwidth \(h\) of the
+henderson filter used to estimate \(\sigma\) and the degree \(d\) of the
+trend to estimate \(\delta\).}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_local_ql.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
+
+\section{Impact of the span used to detect the ARIMA
+model}\label{impact-of-the-span-used-to-detect-the-arima-model}
+
+\begin{figure}
+
+\caption{\label{fig-graphs-impact-span}Distribution of phase shift on
+simulated series by number of years used to identify and estimate the
+ARIMA model.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_arima_length.pdf}
+
+}
+
+\end{figure}%
+
+\newpage
+
+\section{Nearest neighbour bandwidth}\label{nearest-neighbour-bandwidth}
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-lc-nn}Coefficients, gain and phase shift
+functions for the Linear-Constant (LC) filter with \(I/C=3.5\) and a
+nearest neighbour bandwidth.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc_nn.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-ql-nn}Coefficients, gain and phase shift
+functions for the Quadratic-Linear (QL) filter with \(I/C=3.5\) and a
+nearest neighbour bandwidth.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql_nn.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-cq-nn}Coefficients, gain and phase shift
+functions for the Cubic-Quadratic (CQ) filter with \(I/C=3.5\) and a
+nearest neighbour bandwidth.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq_nn.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-graphs-coef-daf-nn}Coefficients, gain and phase
+shift functions for the direct asymmetric filter (DAF) with a nearest
+neighbour bandwidth.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf_nn.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-nn-lp}Distribution of phase shift using nearest
+neighbour bandwidth.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_nn_lp.pdf}
+
+}
+
+\end{figure}%
+
+\begin{figure}
+
+\caption{\label{fig-nn-localparam}Distribution of phase shift using
+nearest neighbour bandwidth for polynomial filters with local
+parametrisation.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_nn_lp_localparam.pdf}
+
+\justifying
+
+\emph{Note: nearest bandwidth moving average are used to estimate slope
+and concavity.} \emph{We thus do not distinguish local parametrisation
+and final local parametrisation which asumptions might not be
+plausible.} \emph{Indeed, using the same framework as for fixed
+bandwidth, the final local parametrisation for real-time estimates would
+implie modelling a polynomial of degree 2 or 3 other 25 months (which
+doesn't seem plausible).}
+
+}
+
+\end{figure}%
+
+\newpage
+
+\section{Quarterly data}\label{quarterly-data}
+
+\begin{figure}
+
+\caption{\label{fig-graphs-quarterly}Distribution of phase shift on
+quarterly data.}
+
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul_q.pdf}
+
+}
+
+\end{figure}%
+
+\begin{table}
+
+\caption{\label{tbl-simulrev-q}Average of the relative deviations of the
+revisions for the different filters on simulated quarterly series with
+medium variability.}
+
+\centering{
+
+\begin{tabular}{ccc}
 \toprule
-Code & Description & Formula\\
-\midrule
-$b_c$ & Constant bias & $\sum_{k=-p}^{+f}\theta_{k}-1$\\
-$b_l$ & Linear bias & $\sum_{k=-p}^{+f}k\theta_{k}$\\
-$b_q$ & Quadratic bias & $\sum_{k=-p}^{+f}k^{2}\theta_{k}$\\
-$F_g$ & Variance reduction / Fidelity (Guggemos) & $\sum_{k=-p}^{+f}\theta_{k}^{2}$\\
-$S_g$ & Smoothness (Guggemos) & $\sum_{j}(\nabla^{3}\theta_{j})^{2}$\\
-\addlinespace
-$T_g$ & Timeliness (Guggemos) & $\int_{0}^{\omega_1}\rho_{\boldsymbol\theta}(\omega)\sin(\varphi_{\boldsymbol\theta}(\omega))^{2}\ud\omega$\\
-$A_w$ & Accuracy (Wildi) & $2\int_0^{\omega_1}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$T_w$ & Timeliness (Wildi) & $8\int_0^{\omega_1} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-$S_w$ & Smoothness (Wildi) & $2\int_{\omega_1}^{\pi}\left(\rho_{s}(\omega)-\rho_{\boldsymbol\theta}(\omega)\right)^{2}h(\omega)\ud\omega$\\
-$R_w$ & Residual (Wildi) & $8\int_{\omega_1}^{\pi} \rho_{s}(\omega)\rho_{\boldsymbol\theta}(\omega)\sin^{2}\left(\frac{\varphi_s(\omega)-\varphi_{\boldsymbol\theta}(\omega)}{2}\right)h(\omega)\ud\omega$\\
-\bottomrule
-\end{tabular}}}
-
-\textit{Note: $X_g$ criteria from Grun-Rehomme et al (2018) and $X_w$ criteria from Wildi and McElroy (2019).}
-
-\textit{$\rho_s$ and $\varphi_s$ represent the gain and phase shift function of the symmetric Henderson filter.}
-
-\textit{$h$ is the spectral density of the input series, considered to be that of white noise ($h_{WN}(x)=1$) or a random walk ($h_{RW}(x)=\frac{1}{2(1-\cos(x))}$).}
-\end{table}
-
-\hypertarget{sec-nonparamreg}{%
-\section{Non-parametric regression and local polynomial
-regression}\label{sec-nonparamreg}}
-
-Many trend-cycle extraction methods are based on non-parametric
-regressions, which are particularly flexible because they do not assume
-any predetermined dependency in the predictors. In practice, local
-regressions can be used. More specifically, consider a set of points
-\((x_i,y_i)_{1\leq i\leq n}\). Non-parametric regression involves
-assuming that there exists a function \(\mu\), to be estimated, such
-that \(y_i=\mu(x_i)+\varepsilon_i\) with \(\varepsilon_i\) an error
-term. According to Taylor's theorem, for any point \(x_0\), if \(\mu\)
-is differentiable \(d\) times, then:
-\begin{equation}\protect\hypertarget{eq-taylor}{}{
-\forall x \::\:\mu(x) = \mu(x_0) + \mu'(x_0)(x-x_0)+\dots +
-\frac{\mu^{(d)}(x_0)}{d!}(x-x_0)^d+R_d(x),
-}\label{eq-taylor}\end{equation} where \(R_d\) is a negligible residual
-term in the neighbourhood of \(x_0\). In a neighbourhood \(h(x_0)\)
-around \(x_0\), \(\mu\) can be approximated by a polynomial of degree
-\(d\). The quantity \(h(x_0)\) is called the \emph{bandwidth}. If
-\(\varepsilon_i\) is white noise, then we can estimate by least squares
-\(\mu(x_0)\) using the observations which are in
-\(\left[x_0-h(x_0),x_0+h(x_0)\right]\).
-
-In practice, this means assuming that the trend is locally polynomial.
-Various estimation methods can be used to derive symmetrical and
-asymmetrical moving averages.\\
-\textcite{GrayThomson1996} propose a complete statistical framework
-which makes it possible, in particular, to model the error in
-approximating the trend by local polynomials. However, as the
-specification of this error is generally complex, simpler models may be
-preferred, such as that of \textcite{proietti2008}.\\
-Finally, \textcite{dagumbianconcini2008} propose a similar modelling of
-the trend-cycle but using the theory of Hilbert spaces with reproducing
-kernels for estimation, which has the particular advantage of
-facilitating the calculation of different moving averages at different
-time frequencies. This method can be used in\faIcon{r-project} with the
-function \texttt{rjd3filters::rkhs\_filter()}.
-
-\hypertarget{symmetric-moving-averages-and-local-polynomial-regression}{%
-\subsection{Symmetric moving averages and local polynomial
-regression}\label{symmetric-moving-averages-and-local-polynomial-regression}}
-
-Using \textcite{proietti2008}'s notation, we assume that our time series
-\(y_t\) can be decomposed into \[
-y_t=\mu_t+\varepsilon_t,
-\] where \(\mu_t\) is the trend and
-\(\varepsilon_{t}\overset{i.i.d}{\sim}\mathcal{N}(0,\sigma^{2})\) is the
-noise (the series is therefore seasonally adjusted). Within a
-neighbourhood \(h\) of \(t\), the local trend \(\mu_t\) is approximated
-by a polynomial of degree \(d\), such that \(\mu_t\simeq m_{t}\) with:
-\[
-\forall j\in\left\llbracket -h,h\right\rrbracket:\:
-y_{t+j}=m_{t+j}+\varepsilon_{t+j},\quad m_{t+j}=\sum_{i=0}^{d}\beta_{i}j^{i}.
-\] The problem of trend extraction is equivalent to estimating
-\(m_t=\beta_0\) (the constant in the previous formula).
-
-In matrix notation: \[
-\underbrace{\begin{pmatrix}y_{t-h}\\
-y_{t-(h-1)}\\
-\vdots\\
-y_{t}\\
-\vdots\\
-y_{t+(h-1)}\\
-y_{t+h}
-\end{pmatrix}}_{\boldsymbol y}=\underbrace{\begin{pmatrix}1 & -h & h^{2} & \cdots & (-h)^{d}\\
-1 & -(h-1) & (h-1)^{2} & \cdots & (-(h-1))^{d}\\
-\vdots & \vdots & \vdots & \cdots & \vdots\\
-1 & 0 & 0 & \cdots & 0\\
-\vdots & \vdots & \vdots & \cdots & \vdots\\
-1 & h-1 & (h-1)^{2} & \cdots & (h-1)^{d}\\
-1 & h & h^{2} & \cdots & h^{d}
-\end{pmatrix}}_{\boldsymbol X}\underbrace{\begin{pmatrix}\beta_{0}\\
-\beta_{1}\\
-\vdots\\
-\vdots\\
-\vdots\\
-\vdots\\
-\beta_{d}
-\end{pmatrix}}_{\boldsymbol \beta}+\underbrace{\begin{pmatrix}\varepsilon_{t-h}\\
-\varepsilon_{t-(h-1)}\\
-\vdots\\
-\varepsilon_{t}\\
-\vdots\\
-\varepsilon_{t+(h-1)}\\
-\varepsilon_{t+h}
-\end{pmatrix}}_{\boldsymbol \varepsilon}.
-\]
-
-To estimate \(\boldsymbol \beta\) we need \(2h+1\geq d+1\) and the
-estimation is made by weighted least squares (WLS), which is equivalent
-to minimising the following objective function: \[
-S(\hat{\beta}_{0},\dots,\hat{\beta}_{d})=\sum_{j=-h}^{h}\kappa_{j}(y_{t+j}-\hat{\beta}_{0}-\hat{\beta}_{1}j-\dots-\hat{\beta}_{d}j^{d})^{2}.
-\] where \(\kappa_j\) is a set of weights called \emph{kernel}. We have
-\(\kappa_j\geq 0:\kappa_{-j}=\kappa_j\), and with
-\(\boldsymbol K=diag(\kappa_{-h},\dots,\kappa_{h})\), the estimate of
-\(\boldsymbol \beta\) can be written as
-\(\hat{\boldsymbol\beta}=(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol X'\boldsymbol K\boldsymbol y.\)
-With \(\boldsymbol e_1=\begin{pmatrix}1 \\0 \\\vdots\\0 \end{pmatrix}\),
-the estimate of the trend is:
-\begin{equation}\protect\hypertarget{eq-mmsym}{}{
-\hat{m}_{t}=\boldsymbol e_{1}\hat{\boldsymbol \beta}=\boldsymbol \theta'\boldsymbol y=\sum_{j=-h}^{h}\theta_{j}y_{t-j}\text{ with }\boldsymbol \theta=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{1}.
-}\label{eq-mmsym}\end{equation} To conclude, the estimate of the trend
-\(\hat{m}_{t}\) can be obtained applying the symmetric filter
-\(\boldsymbol \theta\) to \(y_t\) (\(\boldsymbol \theta\) is symmetric
-due to the symmetry of the kernel weights \(\kappa_j\)). Moreover,
-\(\boldsymbol X'\boldsymbol \theta=\boldsymbol e_{1}\) so: \[
-\sum_{j=-h}^{h}\theta_{j}=1,\quad\forall r\in\left\llbracket 1,d\right\rrbracket:\sum_{j=-h}^{h}j^{r}\theta_{j}=0.
-\] Hence, the filter \(\boldsymbol w\) preserve deterministic polynomial
-of order \(d\).
-
-Regarding parameter selection, the general consensus is that the choice
-between different kernels is not crucial. See, for example,
-\textcite{cleveland1996smoothing} or \textcite{Loader1999}. The only
-desired constraints on the kernel are that it assigns greater weight to
-the central estimation (\(\kappa_0\)) and decreases towards 0 as it
-moves away from the central estimation. The uniform kernel should
-therefore be avoided. It is then preferable to focus on two other
-parameters:
-
-\begin{itemize}
-\item
-  the degree of the polynomial, denoted by \(d\): if it is too small,
-  there is a risk of biased estimates of the trend-cycle, and if it is
-  too large, there is a risk of excessive variance in the estimates (due
-  to over-adjustment);
-\item
-  the number of neighbors \(2h+1\) (or the window \(h\)): if it is too
-  small, then there will be too little data for the estimates (resulting
-  in high variance in the estimates), and if it is too large, the
-  polynomial approximation will probably be wrong, leading to biased
-  estimates.
-\end{itemize}
-
-In this paper we will use the Henderson kernel: \[
-\kappa_{j}=\left[1-\frac{j^2}{(h+1)^2}\right]
-\left[1-\frac{j^2}{(h+2)^2}\right]
-\left[1-\frac{j^2}{(h+3)^2}\right].
-\] However, several other kernels are available in \texttt{rjd3filters}.
-
-\hypertarget{sec-lppasymf}{%
-\subsection{Asymmetric moving averages and local polynomial
-regression}\label{sec-lppasymf}}
-
-As mentioned in Section~\ref{sec-mmetprev}, for real-time estimation,
-several approaches can be used:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  Apply symmetric filters to the series extended by forecasting
-  \(\hat{y}_{n+l\mid n},l\in\left\llbracket 1,h\right\rrbracket\).
-\item
-  Build an asymmetric filter by local polynomial approximation on the
-  available observations (\(y_{t}\) for
-  \(t\in\left\llbracket n-h,n\right\rrbracket\)).
-\item
-  Build asymmetric filters that minimise the mean squared error of
-  revision under polynomial trend reproduction constraints.
-\end{enumerate}
-
-\textcite{proietti2008} show that the first two approaches are
-equivalent when forecasts are made by polynomial extrapolation of degree
-\(d\). They are also equivalent to the third approach under the same
-constraints as those of the symmetric filter. The third method is called
-\emph{direct asymmetric filter} (DAF). This method is used for real-time
-estimation in the STL seasonal adjustment method (\emph{Seasonal-Trend
-decomposition based on Loess}, see \textcite{cleveland90}). Although DAF
-filter estimates are unbiased, this is at the cost of greater variance
-in the estimates.
-
-To solve the problem of the variance of real-time filter estimates,
-\textcite{proietti2008} propose a general method for constructing
-asymmetric filters that allows a bias-variance trade-off. This is a
-generalisation of \textcite{musgrave1964set}'s asymmetric filters (used
-in the X-13ARIMA seasonal adjustment algorithm).
-
-The initial series is modelled here as:
-\begin{equation}\protect\hypertarget{eq-lpgeneralmodel}{}{
-\boldsymbol y=\boldsymbol U \boldsymbol \gamma+\boldsymbol Z\boldsymbol \delta+\boldsymbol \varepsilon,\quad
-\boldsymbol \varepsilon\sim\mathcal{N}(0,\boldsymbol D).
-}\label{eq-lpgeneralmodel}\end{equation} where
-\([\boldsymbol U,\boldsymbol Z]\) is of full rank and forms a subset of
-the columns of \(X\). The objective is to find a filter
-\(\boldsymbol v\) that minimises the mean square error of revision (to
-the symmetric filter \(\boldsymbol \theta\)) under certain constraints.
-These constraints are represented by the matrix
-\(\boldsymbol U=\begin{pmatrix}\boldsymbol U_{p}\\\boldsymbol U_{f}\end{pmatrix}\):
-\(\boldsymbol U_p'\boldsymbol v=\boldsymbol U'\boldsymbol \theta\) (with
-\(\boldsymbol U_p\) the matrix \((h+q+1)\times (d+1)\) which contains
-the observations of the matrix \(U\) known during estimation by the
-asymmetric filter). The problem is equivalent to finding \(v\) which
-minimises: \begin{equation}\protect\hypertarget{eq-lppasym}{}{
-\varphi(\boldsymbol v)=
-\underbrace{
-  \underbrace{(\boldsymbol v-\boldsymbol \theta_{p})'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_{p})+
-  \boldsymbol \theta_{f}'\boldsymbol D_{f}\boldsymbol \theta_{f}}_\text{revision error variance}+
-  \underbrace{[\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}}_{bias^2}
-}_\text{mean square revision error}+
-\underbrace{2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)}_{\text{constraints}}.
-}\label{eq-lppasym}\end{equation} where \(\boldsymbol l\) is a vector of
-Lagrange multipliers.
-
-When \(\boldsymbol U=\boldsymbol X\), the constraint is equivalent to
-preserving polynomials of degree \(d\): we find asymmetric direct
-filters (DAF) when \(\boldsymbol D=\boldsymbol K^{-1}\).
-
-When \(\boldsymbol U=\begin{pmatrix}1\\\vdots\\1\end{pmatrix}\),
-\(\boldsymbol Z=\begin{pmatrix}-h\\\vdots\\+h\end{pmatrix}\),
-\(\boldsymbol \delta=\delta_1\), \(\boldsymbol D=\sigma^2\boldsymbol I\)
-and when the symmetric filter is the Henderson filter, we find the
-asymmetric Musgrave filters (using the Henderson kernel). This filter
-assumes that, for real-time estimation, the data is generated by a
-linear process and that the asymmetric filters preserve the constants
-(\(\sum v_i=\sum \theta_i=1\)). These asymmetric filters depend on the
-ratio \(\lvert\delta_1/\sigma\rvert\), which, assuming that the trend is
-linear and the bias constant, can be linked to the I-C ratio
-\(R=\frac{\bar{I}}{\bar{C}}=\frac{\sum\lvert I_t-I_{t-1}\rvert}{\sum\lvert C_t-C_{t-1}\rvert}\):
-\(\delta_1/\sigma=2/(R\sqrt{\pi})\). This ratio is used in particular in
-X-13ARIMA to determine the length of the Henderson filter. For monthly
-data:
-
-\begin{itemize}
-\item
-  If the ratio is large (\(3.5< R\)) then a 23 term filter is used (to
-  remove more noise).
-\item
-  If the ratio is small (\(R<1\)) a 9 term filter is used.
-\item
-  Otherwise (most of the cases) a 13-term filter is used.
-\end{itemize}
-
-When \(\boldsymbol U\) corresponds to the first \(d^*+1\) columns of
-\(\boldsymbol X\), \(d^*<d\), the constraint is to reproduce polynomial
-trends of degree \(d^*\). This introduces bias but reduces the variance.
-Thus, \textcite{proietti2008} proposes three classes of asymmetric
-filters:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\item
-  \emph{Linear-Constant} (LC): \(y_t\) linear (\(d=1\)) and
-  \(\boldsymbol v\) preserves constant signals (\(d^*=0\)). We obtain
-  Musgrave filters when the Henderson kernel is used and the Henderson
-  filter is used as the symmetric filter.
-\item
-  \emph{Quadratic-Linear} (QL): \(y_t\) quadratic (\(d=2\)) and
-  \(\boldsymbol v\) preserves linear signals (\(d^*=1\)).
-\item
-  \emph{Cubic-Quadratic} (CQ): \(y_t\) cubic (\(d=3\)) and
-  \(\boldsymbol v\) preserves quadratic signals (\(d^*=2\)).
-\end{enumerate}
-
-Appendix \ref{sec-an-cof} shows the coefficients, gain and phase shift
-functions of the four asymmetric filters.
-
-Table~\ref{tab-criteriaLp} compares the quality criteria of the
-different methods using the Henderson filter and \(h=6\) (a symmetric
-filter with 13 terms). For real-time filters (\(q=0\)), a more complex
-asymmetrical filter (in terms of polynomial preservation) results in
-lower timeliness and higher fidelity/smoothness: reduction in phase
-shift comes at the expense of an increase in variance. This result
-varies as \(q\) increases: for \(q=2\), the QL filter has greater
-timeliness than the LC filter. This surprising result highlights the
-fact that the phase shift is not controlled by \textcite{proietti2008}'s
-approach.
-
-In terms of revision (\(A_w+S_w+T_w+R_w\)), the LC and QL filters
-consistently outperform the CQ and DAF filters.
-
-These theoretical properties are consistent with the empirical results
-observed in Section~\ref{sec-comparison}. Revisions are more important
-for the CQ and DAF filters, which leads to greater variability in
-estimations and a longer delay in detecting turning points. The globally
-parameterised QL filter (with \(R\) fixed for all asymmetric filters)
-may result in later detection of turning points than the LC filter.
-Indeed, to identify a turning point at date \(t\), it is necessary to
-have knowledge of at least two points after the given date to confirm
-the reversal in trend.
-
-\begin{table}[!h]
-\caption{\label{tab-criteriaLp} Quality criteria for asymmetric filters ($q=0,1,2$) computed by local polynomials using the Henderson kernel with $h=6$ and $R=3,5$.}
-\centering
-\begin{tabular}[t]{cccccccccccc}
-\toprule
-Method & $b_c$ & $b_l$ & $b_q$ & $F_g$ & $S_g$ & $T_g \times 10^{-3}$ & $A_w$ & $S_w$ & $T_w$ & $R_w$ & $EQM_w$\\
+Method & $q=0$ & $q=1$\\
 \midrule
 \addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=0$}}\\
-\hspace{1em}LC & 0 & -0.41 & -2.16 & 0.39 & 1.27 & 30.34 & 0.10 & 0.49 & 0.41 & 0.55 & 1.54\\
-\hspace{1em}QL & 0 & 0.00 & -0.47 & 0.71 & 5.15 & 0.05 & 0.07 & 1.89 & 0.00 & 0.11 & 2.07\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.91 & 11.94 & 0.01 & 0.02 & 2.23 & 0.00 & 0.10 & 2.35\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.94 & 14.20 & 0.00 & 0.01 & 2.18 & 0.00 & 0.10 & 2.29\\
+\multicolumn{3}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(TC_{t|t+q} -  TC_{t|last})/TC_{t|last}\right|\right]$}}\\
+\hspace{1em}LC & 0.76 & 0.11\\
+\hspace{1em}LC local param. (final estimates) & 0.73 & 0.11\\
+\hspace{1em}LC local param. & 0.71 & 0.12\\
+\hspace{1em}QL & 0.93 & 0.46\\
+\hspace{1em}QL local param. (final estimates) & 0.58 & 0.12\\
+\hspace{1em}QL local param. & 0.97 & 0.33\\
+\hspace{1em}CQ & 1.15 & 1.02\\
+\hspace{1em}DAF & 1.15 & 0.43\\
+\hspace{1em}ARIMA & 0.49 & 0.26\\
 \addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=1$}}\\
-\hspace{1em}LC & 0 & -0.12 & -0.52 & 0.27 & 0.43 & 4.80 & 0.01 & 0.12 & 0.06 & 0.11 & 0.30\\
-\hspace{1em}QL & 0 & 0.00 & -0.06 & 0.29 & 0.71 & 0.69 & 0.00 & 0.19 & 0.01 & 0.04 & 0.25\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.57 & 0.16 & 0.02 & 0.58 & 0.00 & 0.06 & 0.66\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.41 & 0.37 & 0.06 & 0.02 & 0.76 & 0.00 & 0.06 & 0.84\\
-\addlinespace[0.3em]
-\multicolumn{12}{l}{\textbf{$q=2$}}\\
-\hspace{1em}LC & 0 & 0.00 & 1.08 & 0.20 & 0.08 & 0.35 & 0.01 & 0.01 & 0.00 & 0.01 & 0.04\\
-\hspace{1em}QL & 0 & 0.00 & 0.03 & 0.22 & 0.05 & 2.08 & 0.00 & 0.01 & 0.02 & 0.07 & 0.10\\
-\hspace{1em}CQ & 0 & 0.00 & 0.00 & 0.37 & 0.66 & 0.13 & 0.02 & 0.56 & 0.00 & 0.06 & 0.64\\
-\hspace{1em}DAF & 0 & 0.00 & 0.00 & 0.40 & 0.77 & 0.02 & 0.02 & 0.68 & 0.00 & 0.05 & 0.74\\
+\multicolumn{3}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
+\left|(TC_{t|t+q} - TC_{t|t+q+1})/TC_{t|t+q+1}\right|
+\right]$}}\\
+\hspace{1em}LC & 0.36 & 0.11\\
+\hspace{1em}LC local param. (final estimates) & 0.32 & 0.11\\
+\hspace{1em}LC local param. & 0.34 & 0.12\\
+\hspace{1em}QL & 1.75 & 0.46\\
+\hspace{1em}QL local param. (final estimates) & 0.23 & 0.12\\
+\hspace{1em}QL local param. & 0.40 & 0.33\\
+\hspace{1em}CQ & 0.07 & 1.02\\
+\hspace{1em}DAF & 0.75 & 0.43\\
+\hspace{1em}ARIMA & 0.27 & 0.26\\
 \bottomrule
@@
 
-\textit{Note: With $EQM_w=A_w + S_w + T_w + R_w$.}
-\end{table}
-
-\hypertarget{sec-lptimeliness}{%
-\subsection{Extension with the timeliness
-criterion}\label{sec-lptimeliness}}
-
-One drawback of the previous method is the lack of control over the
-phase shift. However, it is possible to improve the modelling by
-incorporating in Equation~\ref{eq-lppasym} the timeliness criterion
-defined by \textcite{ch15HBSA}. This was proposed by Jean Palate, then
-coded in Java and integrated into \texttt{rjd3filters}.
-
-Using the same notation as in Section~\ref{sec-lppasymf},
-\(\boldsymbol\theta\) represents the symmetric filter and
-\(\boldsymbol v\) the asymmetric filter. With
-\(\boldsymbol\theta=\begin{pmatrix}\boldsymbol\theta_p\\\boldsymbol\theta_f\end{pmatrix}\)
-and \(\boldsymbol\theta_p\) of the same length of \(\boldsymbol v\), and
-\(\boldsymbol g=\boldsymbol v-\boldsymbol \theta_p\), the timeliness
-criterion can be expressed as: \[
-T_g(\boldsymbol v)=\boldsymbol v'\boldsymbol T\boldsymbol v=\boldsymbol g'\boldsymbol T\boldsymbol g+2\boldsymbol \theta_p'\boldsymbol T\boldsymbol g+\boldsymbol \theta_p'\boldsymbol T\boldsymbol \theta_p
-\quad\text{ with }\boldsymbol T\text{ a symmetric matrix}.
-\] Furthermore, the objective function \(\varphi\) of
-Equation~\ref{eq-lppasym} can be rewritten: \begin{align*}
-\varphi(\boldsymbol v)&=(\boldsymbol v-\boldsymbol \theta_p)'\boldsymbol D_{p}(\boldsymbol v-\boldsymbol \theta_p)+
-  \boldsymbol \theta_f'\boldsymbol D_{f}\boldsymbol \theta_f+
-  [\boldsymbol \delta'(\boldsymbol Z_{p}'\boldsymbol v-\boldsymbol Z'\boldsymbol \theta)]^{2}+
-2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)\\
-&=\boldsymbol g'\boldsymbol Q\boldsymbol g-2\boldsymbol P\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+\boldsymbol c\quad\text{ with }\boldsymbol v
-\begin{cases}
-\boldsymbol Q=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p \\
-\boldsymbol P=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'\\
-\boldsymbol c\text{ a constant independent of }\boldsymbol v
-\end{cases}.
-\end{align*} Adding the timeliness criterion: \[
-\widetilde\varphi(\boldsymbol v)=\boldsymbol g'\widetilde {\boldsymbol Q}\boldsymbol g-
-2\widetilde {\boldsymbol P}\boldsymbol g+2\boldsymbol l'(\boldsymbol U_{p}'\boldsymbol v-\boldsymbol U'\boldsymbol \theta)+
-\widetilde {\boldsymbol c}\quad\text{with }
-\begin{cases}
-\widetilde {\boldsymbol Q}=\boldsymbol D_p+\boldsymbol Z_p\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z'_p + \alpha_T\boldsymbol T\\
-\widetilde {\boldsymbol P}=\boldsymbol \theta_f\boldsymbol Z_f\boldsymbol \delta\boldsymbol \delta'\boldsymbol Z_p'-\alpha_T\boldsymbol \theta_p\boldsymbol T\\
-\widetilde {\boldsymbol c}\text{ a constant independent of }\boldsymbol v
-\end{cases},
-\] where \(\alpha_T\) is the weight associated to the timeliness
-criterion. With \(\alpha_T=0\) we find \(\varphi(\boldsymbol v)\). This
-extension therefore makes it possible to find all the symmetric and
-asymmetric filters presented in the previous section but also
-generalizes the approach of \textcite{GrayThomson1996}. This extension
-is implemented in the \faIcon{r-project} function
-\texttt{rjd3filters::lp\_filter()}.
-
-One drawback is that \(T_g(\boldsymbol v)\) is not normalized: the
-weight \(\alpha_T\) has no economic sense. It is then difficult to
-calibrate the value of \(\alpha_T\) since the value of
-\(\lvert\delta/\sigma\rvert\) must also be defined. That's why in this
-paper we will not compare this extension to the other methods and we
-will focus on the local calibration of \(\lvert\delta/\sigma\rvert\).
-However, in future studies, we could imagine a two step calibration: fix
-the value of \(\lvert\delta/\sigma\rvert\) to find \(\alpha_T\) by
-cross-validation and then use this weight with a local parameterisation
-of the ratio \(\lvert\delta/\sigma\rvert\) (as in
-Section~\ref{sec-localic}).
-
-\hypertarget{sec-localic}{%
-\subsection{Local parameterisation of asymmetric
-filters}\label{sec-localic}}
-
-Asymmetric filters are usually parameterised globally:
-\(\lvert\delta/\sigma\rvert\) estimated on all the data using the
-IC-ratio or a cross-validation criterion. However, we might prefer a
-local parameterisation: ratio \(\lvert\delta/\sigma\rvert\) which varies
-as a function of time. Indeed, although the overall parameterisation is
-generally valid, assuming a constant value of the
-\(\lvert\delta/\sigma\rvert\) ratio for all asymmetrical filters does
-not appear relevant for real-time estimation, especially during periods
-of economic downturn. For example, with the LC method, global
-parameterisation means assuming that the slope of the trend is constant,
-whereas during economic downturns it tends towards 0 up to the turning
-point.
-
-This is what is proposed in this article, with a local parameterisation
-of the asymmetric filters by estimating \(\delta\) and \(\sigma^2\)
-separately. Although this does not give an unbiased estimator of the
-ratio \(\lvert\delta/\sigma\rvert\), it does allow the main evolutions
-to be captured, such as the decay towards 0 before a turning point and
-the growth after the turning point for the LC method:
-
-\begin{itemize}
-\tightlist
-\item
-  The variance \(\sigma^2\) can be estimated using the observed data set
-  and the symmetrical filter \((w_{-p},\dots,w_p)\): \[
-  \hat\sigma^2=\frac{1}{n-2h}\sum_{t=h+1}^{n-h}\frac{(y_t-\hat \mu_t)^2}{1-2w_0^2+\sum w_i^2}.
-  \]
-\item
-  The parameter \(\delta\) can be estimated by moving average from
-  Equation~\ref{eq-mmsym}. For example, for the LC method we can use the
-  moving average
-  \(\boldsymbol \theta_2=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{2}\)
-  to obtain a local estimate of the slope and for the QL method we can
-  use
-  \(\boldsymbol \theta_3=\boldsymbol K\boldsymbol X(\boldsymbol X'\boldsymbol K\boldsymbol X)^{-1}\boldsymbol e_{3}\)
-  to obtain a local estimate of the concavity. The DAF method then
-  simply allows us to calculate the associated asymmetric moving
-  averages.\\
-  Although a moving average of different length to that used for
-  estimating the trend could be considered, this seems to degrade the
-  results in terms of phase shift (using the same methodology as in
-  Section~\ref{sec-comparison}). Furthermore, for the construction of
-  moving averages, the trend can be modelled as being locally of degree
-  2 or 3 (this has no impact on the final estimate of concavity). Here,
-  we have chosen to model a trend of degree 2: this reduces the phase
-  shift but slightly increases the revisions linked to the first
-  estimate of the trend-cycle. Figure~\ref{fig-mmpenteconcac} shows the
-  moving averages used.
-\end{itemize}
-
-\begin{figure}[H]
-
-\caption{\label{fig-mmpenteconcac}Moving averages used for real-time
-estimation of slope and concavity.}
-
-{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{img/filters_used/mm_penteconcavite.pdf}
-
-}
-
-\end{figure}
-
-In the empirical applications of Section~\ref{sec-comparison}, the final
-local parameterisation is obtained by estimating \(\delta\) using all
-the data (i.e., using the symmetric filters shown in
-Figure~\ref{fig-mmpenteconcac}), while maintaining a real-time estimate
-of \(\sigma^2\). Figure~\ref{fig-mmpenteconcac-ex} shows an example of
-the comparison of estimates of the ratio \(|\delta_1/\sigma|\) with a
-global estimator (IC ratio) and two the local parameterisation:
-
-\begin{itemize}
-\item
-  Figure~\ref{fig-mmpenteconcac-ex-1}: using a symmetric filter to
-  estimate \(\delta_1\) (final local parameterisation).
-\item
-  Figure~\ref{fig-mmpenteconcac-ex-2}: using the asymmetric filter which
-  need two points in the futur to estimate \(\delta_1\). This is closed
-  to the real-time estimates since two points in the future are needed
-  to detect a turning point.
-\end{itemize}
-
-The turning points are clearly detected by the local estimators, but
-there is much more variance in the estimate of \(|\delta_1/\sigma|\) in
-the real-time estimate. This will lead to another source of revisions in
-the intermediate estimates.
-
-\begin{figure}
-
-\caption{\label{fig-mmpenteconcac-ex}Comparison of the estimates of
-\(|\delta_1/\sigma|\) with a local estimator and global estimator (with
-IC ratio) on a simulated series, the vertical lines being the simulated
-turning points.}\begin{minipage}[t]{\linewidth}
-\subcaption{\label{fig-mmpenteconcac-ex-1}Symmetric moving average used
-for \(\hat \delta_1\).}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{img/filters_used/mm_penteconcavite_ex.pdf}
-
-}
-
-}
-
-\end{minipage}%
-\newline
-\begin{minipage}[t]{\linewidth}
-\subcaption{\label{fig-mmpenteconcac-ex-2}Asymmetric moving average
-using two futur points (\(q=2\)) used for \(\hat \delta_1\).}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{img/filters_used/mm_penteconcavite_ex2.pdf}
-
-}
-
-}
-
-\end{minipage}%
-
-\end{figure}
-
-\hypertarget{sec-comparison}{%
-\section{Comparison of different methods}\label{sec-comparison}}
-
-The different methods are compared on simulated and real data.
-
-For all series, a symmetrical 13-term filter is used. These methods are
-also compared with estimates obtained by extending the series with an
-ARIMA model, then applying a 13-term symmetrical Henderson filter. The
-ARIMA model is determined automatically, using no seasonal lag (the
-series being seasonally adjusted) and no external variables (such as
-external regressors for correction of atypical points).
-
-The performance of the different methods is judged on criteria relating
-to revisions (between two consecutive estimates and in relation to the
-final estimate) and the number of periods required to detect turning
-points.
-
-\hypertarget{simulated-series}{%
-\subsection{Simulated series}\label{simulated-series}}
-
-\hypertarget{methodology}{%
-\subsubsection{Methodology}\label{methodology}}
-
-Following a methodology close to that of \textcite{DarneDagum2009}, nine
-monthly series are simulated between January 1960 and December 2020 with
-different levels of variability. Each simulated series
-\(y_t= C_t+ T_t + I_t\) can be written as the sum of three components:
-
-\begin{itemize}
-\item
-  the cycle
-  \(C_t = \rho [\cos (2 \pi t / \lambda) +\sin (2 \pi t / \lambda)]\),
-  \(\lambda\) is fixed at 72 (6-year cycles, so there are 19 detectable
-  turning points);
-\item
-  the trend \(T_t = T_{t-1} + \nu_t\) with
-  \(\nu_t \sim \mathcal{N}(0, \sigma_\nu^2)\), \(\sigma_\nu\) being
-  fixed at \(0.08\);
-\item
-  and the irregular \(I_t = e_t\) with
-  \(e_t \sim \mathcal{N}(0, \sigma_e^2)\).
-\end{itemize}
-
-For the different simulations, we vary the parameters \(\rho\) and
-\(\sigma_e^2\) in order to obtain series with different signal-to-noise
-ratios (see Figure~\ref{fig-graphs-data-simul}):
-
-\begin{itemize}
-\item
-  High signal-to-noise ratio (i.e.~low I-C ratio and low variability):
-  \(\sigma_e^2=0.2\) and \(\rho = 3.0, 3.5\) or \(4.0\) (I-C ratio
-  between 0.9 and 0.7);
-\item
-  Medium signal-to-noise ratio (i.e.~medium I-C ratio and medium
-  variability): \(\sigma_e^2=0.3\) and \(\rho = 1.5,\, 2.0\) or \(3.0\)
-  (I-C ratio between 2.3 and 1.4);
-\item
-  Low signal-to-noise ratio (i.e.~high I-C ratio and high variability):
-  \(\sigma_e^2=0.4\) and \(\rho = 0.5,\, 0.7\) or \(1.0\) (I-C ratio
-  between 8.9 and 5.2).
-\end{itemize}
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-data-simul}Simulated series with low
-(\(\sigma_e^2=0.2\) and \(\rho = 3,5\)), medium (\(\sigma_e^2=0.3\) and
-\(\rho = 2.0\)) and high variability (\(\sigma_e^2=0.4\) and
-\(\rho = 1.0\)).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/simul_data.pdf}
-
-}
-
-\end{figure}
-
-For each series and each date, the trend-cycle is estimated using the
-different methods presented in this report. Three quality criteria are
-also calculated:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Computation of the phase shift in the detection of turning points. The
-  definition of \textcite{Zellner1991} is used to determine the turning
-  points:
-
-  \begin{itemize}
-  \tightlist
-  \item
-    An upturn occurs when the economy moves from a phase of recession to
-    a phase of expansion. This is the case at date \(t\) when
-    \(y_{t-3}\geq y_{t-2}\geq y_{t-1}<y_t\leq y_{t+1}\).\\
-  \item
-    A downturn occurs when the economy moves from a phase of expansion
-    to a phase of recession. This is the case at date \(t\) when
-    \(y_{t-3}\leq y_{t-2}\leq y_{t-1}>y_t\geq y_{t+1}\).\\
-    It thus takes at least 2 months to detect a turning point.\\
-    The phase shift is often defined as the number of months required to
-    detect the right turning point. Here we use a slightly modified
-    criterion: the phase shift is defined as the number of months needed
-    to detect the right turning point without any future revision.
-    Indeed, it can happen that the right turning point is detected by
-    asymmetric filters but is not detected by the final estimate using a
-    symmetric filter (this is the case for 41 reversal points out of the
-    9 series with asymmetric Musgrave filters) or that there are
-    revisions in successive estimates (this is the case for 7 turning
-    points out of the 9 series with asymmetric Musgrave filters).
-    Finally, relatively few turning points are detected at the right
-    date with the final estimate. With the 13-term Henderson filter, 18
-    are correctly detected in series with low variability (out of 57
-    possible), 11 in series with medium variability and 12 in series
-    with high variability.
-  \end{itemize}
-\item
-  Computation of two revision criteria: the average of the relative
-  deviations between the \(q\)\textsuperscript{th} estimate and the last
-  estimate \(MAE_{fe}(q)\) and the average of the relative deviations
-  between the \(q\)\textsuperscript{th} estimate and the
-  \(q+1\)\textsuperscript{th} estimate \(MAE_{qe}(q)\). \[
-  MAE_{fe}(q)=\mathbb E\left[
-  \left|\frac{
-  y_{t|t+q} -  y_{t|last}
-  }{
-   y_{t|last}
-  }\right|
-  \right]
-  \quad\text{and}\quad
-  MAE_{qe}(q)=\mathbb E\left[
-  \left|\frac{
-  y_{t|t+q} - y_{t|t+q+1}
-  }{
-  y_{t|t+q+1}
-  }\right|
-  \right].
-  \]
-\end{enumerate}
-
-\hypertarget{comparison}{%
-\subsection{Comparison}\label{comparison}}
-
-Excluding for the moment the local parameterisations of the polynomial
-filters, the linear-constant polynomial filter (LC) seems to give the
-best results in terms of delay in the detection of turning points
-(Figure~\ref{fig-graphstpsimul}). Performance is relatively close to
-that obtained by extending the series using an ARIMA model. However,
-when variability is low, the LC filter seems to give poorer results and
-the quadratic-linear polynomial filter (QL) seems to give the best
-results.
-
-For series with moderate variability, local parameterisation of the LC
-and QL filters reduces the phase shift. For series with high
-variability, the phase shift is only reduced by using the final
-\(\hat\delta\) parameters: real-time estimates seem to add more
-variance. For series with low variability, performance seems to be
-slightly improved only with the LC filter.
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphstpsimul}Distribution of phase shifts on
-simulated series.}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/simulations/phase_shift_simul.pdf}
-
-}
-
-\end{figure}
-
-In terms of revisions, the variability of the series has little effect
-on the respective performances of the different methods, but it does
-affect the orders of magnitude, which is why the results are presented
-only for series with medium variability (Table~\ref{tab-simulrev}). In
-general, LC filters always minimise revisions (with relatively small
-effect from the local parameterisation of the filters) and revisions are
-greater with cubic-quadratic (CQ) and direct (DAF) polynomial filters.
-
-For the QL filter, there is a large revision between the second and
-third estimates: this may be due to the fact that for the second
-estimate (when one point in the future is known), the QL filter assigns
-a greater weight to the estimate in \(t+1\) than to the estimate in
-\(t\), which creates a discontinuity. This revision is considerably
-reduced by parameterising the filter locally. For polynomial filters
-other than the LC filter, the large revisions at the first estimate were
-to be expected given the coefficient curve: a very large weight is
-associated with the current observation and there is a strong
-discontinuity between the moving average used for real-time estimation
-(when no point in the future is known) and the other moving averages.
-
-Extending the series using an ARIMA model gives revisions with the
-latest estimates of the same order of magnitude as the LC filter, but
-slightly larger revisions between consecutive estimates, particularly
-between the fourth and fifth estimates (as might be expected, as
-highlighted in Section~\ref{sec-mmetprev}).
-
-\begin{table}[!h]
-\caption{\label{tab-simulrev} Average of the relative deviations of the revisions for the different
-filters on simulated series with medium variability.}
-\centering
-\begin{tabular}{ccccccc}
-\toprule
-Method & $q=0$ & $q=1$ & $q=2$ & $q=3$ & $q=4$ & $q=5$\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{fe}(q) = \mathbb E\left[\left|(y_{t|t+q} -  y_{t|last})/y_{t|last}\right|\right]$}}\\
-\hspace{1em}LC & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}LC local param. (final estimates) & 0.19 & 0.09 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}LC local param. & 0.29 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}QL & 0.33 & 0.10 & 0.04 & 0.04 & 0.03 & 0.01\\
-\hspace{1em}QL local param. (final estimates) & 0.21 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}QL local param. & 0.30 & 0.10 & 0.04 & 0.03 & 0.03 & 0.01\\
-\hspace{1em}CQ & 0.45 & 0.13 & 0.13 & 0.09 & 0.06 & 0.02\\
-\hspace{1em}DAF & 0.47 & 0.15 & 0.15 & 0.09 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.22 & 0.10 & 0.03 & 0.03 & 0.03 & 0.01\\
-\addlinespace[0.3em]
-\multicolumn{7}{l}{\textbf{$MAE_{ce}(q)=\mathbb E\left[
-\left|(y_{t|t+q} - y_{t|t+q+1})/y_{t|t+q+1}\right|
-\right]$}}\\
-\hspace{1em}LC & 0.19 & 0.10 & 0.02 & 0.01 & 0.07 & 0.01\\
-\hspace{1em}LC local param. (final estimates) & 0.20 & 0.10 & 0.03 & 0.01 & 0.05 & 0.01\\
-\hspace{1em}LC local param. & 0.24 & 0.11 & 0.03 & 0.01 & 0.05 & 0.01\\
-\hspace{1em}QL & 0.29 & 3.46 & 0.00 & 0.03 & 0.04 & 0.01\\
-\hspace{1em}QL local param. (final estimates) & 0.31 & 0.11 & 0.02 & 0.01 & 0.04 & 0.01\\
-\hspace{1em}QL local param. & 0.24 & 0.16 & 0.00 & 0.03 & 0.04 & 0.01\\
-\hspace{1em}CQ & 0.43 & 0.02 & 0.10 & 0.07 & 0.05 & 0.02\\
-\hspace{1em}DAF & 0.66 & 0.24 & 0.11 & 0.14 & 0.06 & 0.02\\
-\hspace{1em}ARIMA & 0.21 & 0.13 & 0.02 & 0.02 & 0.25 & 0.01\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{real-series}{%
-\subsection{Real series}\label{real-series}}
-
-The differences between the methods are also illustrated using an
-example from the FRED-MD database (\textcite{fredmd}) containing
-economic series on the United States. The series studied correspond to
-the database published in November 2022. The series studied is the level
-of employment in the United States (series \texttt{CE160V}, used in
-logarithm) around the February 2001 turning point, consistent with the
-monthly dating of the turning points. This point and this series were
-chosen because the economic turnaround is particularly visible in the
-raw series (Figure~\ref{fig-ce16ov-previmp-lp}), because this series is
-used for dating economic cycles and because the FRED-MD database
-facilitates the reproducibility of the results (thanks to the
-availability of series published on past dates). Studying the series up
-to January 2020, this series has a medium variability: a symmetrical
-filter of 13 terms is therefore appropriate. Appendix
-\autocite*{ann-ex-r} shows how the code to reproduce the different
-plots.
-
-Figure~\ref{fig-ce16ovlp} shows the successive estimates of the
-trend-cycle using the different methods studied.
-
-For this series, the phase shift is 6 months for the LC and CQ methods
-and the extension of the series by ARIMA. Except for the LC method
-(where the turning point is first detected in January 2001), the high
-phase shift is due to revisions in the turning point detected in the
-intermediate estimates (even if the correct turning point is detected
-after 2 months, it is no longer detected after 3 months). It is of two
-months for the other methods (QL, DAF).
-
-In this example, local parameterisation does not reduce the phase shift,
-but it does reduce revisions. The QL and DAF polynomials lead to greater
-variability in the intermediate estimates, especially in February 2001.
-
-\begin{figure}[H]
-
-\caption{\label{fig-ce16ovlp}Successive estimates of the trend-cycle in
-US employment (in logarithms).}
-
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_lp.pdf}
-
-}
-
-\end{figure}
-
-The quality of the intermediate estimates can also be analysed using the
-implicit forecasts of the different methods
-(Figure~\ref{fig-ce16ov-previmp-lp}). As a reminder, these are forecasts
-of the raw series which, by applying the symmetrical Henderson filter to
-the extended series, give the same estimates as the asymmetrical moving
-averages. The forecasts of the ARIMA model are naive and do not take the
-turning point into account, unlike the other methods. Finally, local
-parameterisation of the QL filter produces much more consistent
-forecasts.
-
-\begin{figure}[H]
-
-\caption{\label{fig-ce16ov-previmp-lp}Implicit forecasts linked to
-successive estimates of the trend-cycle in US employment (in logarithms)
-using local polynomial methods.}
-
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{img/nber/ce16ov_fev2001_prev_imp_lp.pdf}
-
-}
-
-\end{figure}
-
-\hypertarget{conclusion}{%
-\section*{Conclusion}\label{conclusion}}
-\addcontentsline{toc}{section}{Conclusion}
-
-For business cycle analysis, most statisticians use trend-cycle
-extraction methods, either directly or indirectly. They are used, for
-example, to reduce the noise of an indicator in order to improve its
-analysis, and models, such as forecasting models, usually use seasonally
-adjusted series based on these methods.
-
-This paper presents the \faIcon{r-project} package \texttt{rjd3filters},
-which implements a several methods to build moving average for real-time
-trend-cycle estimation. It also offers various functions for the
-theoretical analysis of moving averages (gain functions, phase shift,
-quality criteria, etc.) and judging the quality of the latest estimates
-(for example with implicit forecasts). It also makes it easy to combine
-moving averages and also to integrate custom moving averages into the
-X-11 seasonal adjustment algorithm (with the
-\texttt{rjd3filters::x11plus\_trend()} function). \texttt{rjd3filters}
-thus facilitates research into the use of moving averages as part of
-real-time cycle trend estimation. This is the case, for example, with
-the local parametrisation of Musgrave moving averages and other
-asymmetric moving averages linked to local polynomial regression,
-presented in this article.
-
-By comparing the different methods, we can learn a few lessons about the
-construction of these moving averages.
-
-During economic downturns, asymmetric filters used as an alternative to
-extending the series using the ARIMA model can reduce revisions to
-intermediate estimates of the trend-cycle and enable turning points to
-be detected more quickly.
-
-At the end of the period, modelling polynomial trends of degree greater
-than three (cubic-quadratic, CQ, and direct, DAF) seems to introduce
-variance into the estimates (and therefore more revisions) without
-allowing faster detection of turning points. At the end of the period,
-for estimating the trend-cycle, we can therefore restrict ourselves to
-methods modelling polynomial trends of degree two or less
-(linear-constant, LC, and quadratic-linear, QL). In addition,
-parameterising polynomial filters locally as suggested in this paper
-enables turning points to be detected more quickly (especially for the
-QL filter). Even when the phase shift is not reduced, local
-parameterisation is recommended because it reduces revisions and
-produces intermediate estimates that are more consistent with expected
-future trends. However, with these methods, the length of the filter
-used must be adapted to the variability of the series: if the filter
-used is too short (i.e.~if the variability of the series is ``low''),
-retaining polynomial trends of degree one or less (LC method) produces
-poorer results in terms of detecting turning points.
-
-This study could be extended in many ways.
-
-One possible extension would be to look at the impact of filter length
-on the detection of turning points. Asymmetric filters are calibrated
-using indicators calculated for the estimation of symmetric filters (for
-example, to automatically determine their length), whereas a local
-estimate might be preferable. Furthermore, we have only focused on
-monthly series with a 13-term symmetric filter, but the results may be
-different if the symmetric filter studied is longer or shorter and if we
-study series with other frequencies (quarterly or daily, for example).
-
-Another possibility would be to study the impact of atypical points:
-moving averages, like any linear operator, are very sensitive to the
-presence of atypical points. To limit their impact, in X-13ARIMA a
-strong correction for atypical points is performed on the irregular
-component before applying the filters to extract the trend-cycle. This
-leads us to study the impact of these points on the estimation of the
-trend-cycle and turning points, and also to explore new types of
-asymmetric filters based on robust methods (such as robust local
-regressions or moving medians).
-
-\newpage
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\printbibliography[heading=none]
-
-\appendix
-\newpage
-
-\hypertarget{sec-an-cof}{%
-\section{Coefficients, gain and phase shift
-functions}\label{sec-an-cof}}
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-lc}Coefficients, gain and phase shift
-functions for the Linear-Constant (LC) filter with \(I/C=3.5\).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/lc.pdf}
-
-}
-
-\end{figure}
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-ql}Coefficients, gain and phase shift
-functions for the Quadratic-Linear (QL) filter with \(I/C=3.5\).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/ql.pdf}
-
-}
-
-\end{figure}
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-cq}Coefficients, gain and phase shift
-functions for the Cubic-Quadratic (CQ) filter with \(I/C=3.5\).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/cq.pdf}
-
-}
-
-\end{figure}
-
-\begin{figure}[H]
-
-\caption{\label{fig-graphs-coef-daf}Coefficients, gain and phase shift
-functions for the direct asymmetric filter (DAF).}
-
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{img/filters_used/daf.pdf}
-
-}
-
-\end{figure}
-
-\hypertarget{ann-ex-r}{%
-\section{R code example}\label{ann-ex-r}}
-
-In this appendix we present the \faIcon{r-project} code that can be used
-to estimate trend-cycle components for the US Employment with local
-polynomial filters.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# \# To install rjd3filters}
-\CommentTok{\# remotes::install\_github("rjdemetra/rjd3toolkit")}
-\CommentTok{\# remotes::install\_github("rjdemetra/rjd3x11plus")}
-\CommentTok{\# remotes::install\_github("rjdemetra/rjd3filters")}
-\FunctionTok{library}\NormalTok{(rjd3filters)}
-\FunctionTok{library}\NormalTok{(ggplot2)}
-\FunctionTok{library}\NormalTok{(patchwork)}
-\FunctionTok{library}\NormalTok{(zoo)}
-\FunctionTok{library}\NormalTok{(forecast)}
-\NormalTok{file }\OtherTok{\textless{}{-}}\StringTok{"https://files.stlouisfed.org/files/htdocs/fred{-}md/monthly/2022{-}11.csv"}
-\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(file)}
-\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ data[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,] }\CommentTok{\# First line removed: it contains the transformation codes}
-\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{ts}\NormalTok{(}\FunctionTok{log}\NormalTok{(data[, }\StringTok{"CE16OV"}\NormalTok{]), }\CommentTok{\# We study US employment}
-         \AttributeTok{start =} \DecValTok{1959}\NormalTok{, }\AttributeTok{frequency =} \DecValTok{12}\NormalTok{)}
-\NormalTok{y }\OtherTok{\textless{}{-}} \FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =} \FunctionTok{c}\NormalTok{(}\DecValTok{2001}\NormalTok{, }\DecValTok{9}\NormalTok{))}
-\NormalTok{last\_dates }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(y), }\DecValTok{8}\NormalTok{))}
-\FunctionTok{names}\NormalTok{(last\_dates) }\OtherTok{\textless{}{-}} \FunctionTok{as.character}\NormalTok{(zoo}\SpecialCharTok{::}\FunctionTok{as.yearmon}\NormalTok{(last\_dates))}
-\NormalTok{last\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(last\_dates, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{window}\NormalTok{(y, }\AttributeTok{end =}\NormalTok{ x))}
-
-\CommentTok{\# MA of length 13 is appropriated:}
-\FunctionTok{select\_trend\_filter}\NormalTok{(y)}
-\CommentTok{\# Final estimates:}
-\NormalTok{tc\_f }\OtherTok{\textless{}{-}} \FunctionTok{henderson}\NormalTok{(y, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{)}
-\FunctionTok{plot}\NormalTok{(y)}
-\FunctionTok{lines}\NormalTok{(tc\_f, }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
-\NormalTok{forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(}\FunctionTok{ts.union}\NormalTok{(y, tc\_f))}
-
-\CommentTok{\# IC{-}ratio computation}
-\NormalTok{icr }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-  \FunctionTok{ic\_ratio}\NormalTok{(x, }\FunctionTok{henderson}\NormalTok{(x, }\AttributeTok{length =} \DecValTok{13}\NormalTok{, }\AttributeTok{musgrave =} \ConstantTok{FALSE}\NormalTok{))}
-\NormalTok{\})}
-\NormalTok{lp\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
-\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
-\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
-                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
-                         \AttributeTok{endpoints =}\NormalTok{ method,}
-                         \AttributeTok{ic =}\NormalTok{ icr[i])}
-\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(last\_est[[i]], lp\_coef)}
-\NormalTok{  \})}
-  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
-\NormalTok{  res}
-\NormalTok{\})}
-\NormalTok{lp\_if }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{), }\ControlFlowTok{function}\NormalTok{(method) \{}
-\NormalTok{  res }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\FunctionTok{seq\_along}\NormalTok{(icr), }\ControlFlowTok{function}\NormalTok{(i) \{}
-\NormalTok{    lp\_coef }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =} \DecValTok{6}\NormalTok{,}
-                         \AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{,}
-                         \AttributeTok{endpoints =}\NormalTok{ method,}
-                         \AttributeTok{ic =}\NormalTok{ icr[i])}
-    \FunctionTok{implicit\_forecast}\NormalTok{(last\_est[[i]], lp\_coef)}
-\NormalTok{  \})}
-  \FunctionTok{names}\NormalTok{(res) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(last\_est)}
-\NormalTok{  res}
-\NormalTok{\})}
-\FunctionTok{names}\NormalTok{(lp\_est) }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(lp\_if) }\OtherTok{\textless{}{-}}  \FunctionTok{c}\NormalTok{(}\StringTok{"LC"}\NormalTok{, }\StringTok{"QL"}\NormalTok{, }\StringTok{"CQ"}\NormalTok{, }\StringTok{"DAF"}\NormalTok{)}
-
-\CommentTok{\# Local estimates of IC{-}ratios}
-\CommentTok{\# We replicate the direct estimates to have}
-\CommentTok{\# estimators of the slope and the concavity}
-\NormalTok{gen\_MM }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{p=}\DecValTok{6}\NormalTok{, }\AttributeTok{q=}\NormalTok{p, }\AttributeTok{d=}\DecValTok{2}\NormalTok{)\{}
-\NormalTok{  X\_gen }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(}\AttributeTok{d =} \DecValTok{1}\NormalTok{, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{q =}\NormalTok{ p)\{}
-    \FunctionTok{sapply}\NormalTok{(}\DecValTok{0}\SpecialCharTok{:}\NormalTok{d, }\ControlFlowTok{function}\NormalTok{(exp) }\FunctionTok{seq}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{p, q)}\SpecialCharTok{\^{}}\NormalTok{exp)}
-\NormalTok{  \}}
-\NormalTok{  k }\OtherTok{=}\NormalTok{ rjd3filters}\SpecialCharTok{::}\FunctionTok{get\_kernel}\NormalTok{(}\StringTok{"Henderson"}\NormalTok{, }\AttributeTok{h =}\NormalTok{ p)}
-\NormalTok{  k }\OtherTok{=} \FunctionTok{c}\NormalTok{(}\FunctionTok{rev}\NormalTok{(k}\SpecialCharTok{$}\NormalTok{coef[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{]), k}\SpecialCharTok{$}\NormalTok{coef[}\FunctionTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{,q)}\SpecialCharTok{+}\DecValTok{1}\NormalTok{])}
-\NormalTok{  K }\OtherTok{=} \FunctionTok{diag}\NormalTok{(k)}
-\NormalTok{  X }\OtherTok{=} \FunctionTok{X\_gen}\NormalTok{(}\AttributeTok{d=}\NormalTok{d, }\AttributeTok{p =}\NormalTok{ p, }\AttributeTok{q =}\NormalTok{ q)}
-\NormalTok{  e1 }\OtherTok{=}\NormalTok{ e2 }\OtherTok{=}\NormalTok{ e3 }\OtherTok{=} \FunctionTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{, }\AttributeTok{ncol =} \DecValTok{1}\NormalTok{, }\AttributeTok{nrow =}\NormalTok{ d}\SpecialCharTok{+}\DecValTok{1}\NormalTok{)}
-\NormalTok{  e1[}\DecValTok{1}\NormalTok{] }\OtherTok{=} \DecValTok{1}
-\NormalTok{  e2[}\DecValTok{2}\NormalTok{] }\OtherTok{=} \DecValTok{1}
-\NormalTok{  e3[}\DecValTok{3}\NormalTok{] }\OtherTok{=} \DecValTok{1}
-  \CommentTok{\# Estimator of the constant}
-\NormalTok{  M1 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e1)}
-  \CommentTok{\# Estimator of the slope}
-\NormalTok{  M2 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e2)}
-  \CommentTok{\# Estimor of the concavity}
-\NormalTok{  M3 }\OtherTok{=}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X }\SpecialCharTok{\%*\%} \FunctionTok{solve}\NormalTok{(}\FunctionTok{t}\NormalTok{(X) }\SpecialCharTok{\%*\%}\NormalTok{ K }\SpecialCharTok{\%*\%}\NormalTok{ X, e3)}
-\NormalTok{  mm }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\AttributeTok{const =}\NormalTok{ M1, }\AttributeTok{slope =}\NormalTok{ M2, }\AttributeTok{concav =}\NormalTok{ M3)}
-  \FunctionTok{lapply}\NormalTok{(mm, moving\_average, }\AttributeTok{lags =} \SpecialCharTok{{-}}\NormalTok{p)}
-\NormalTok{\}}
-\NormalTok{all\_mm }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}\DecValTok{6}\SpecialCharTok{:}\DecValTok{0}\NormalTok{, gen\_MM, }\AttributeTok{p =} \DecValTok{6}\NormalTok{, }\AttributeTok{d =} \DecValTok{2}\NormalTok{)}
-\NormalTok{est\_slope }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{slope,}
-                            \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"slope"}\NormalTok{))}
-\NormalTok{est\_concav }\OtherTok{\textless{}{-}} \FunctionTok{finite\_filters}\NormalTok{(all\_mm[[}\DecValTok{1}\NormalTok{]]}\SpecialCharTok{$}\NormalTok{concav,}
-                             \FunctionTok{lapply}\NormalTok{(all\_mm[}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{], }\StringTok{\textasciigrave{}}\AttributeTok{[[}\StringTok{\textasciigrave{}}\NormalTok{, }\StringTok{"concav"}\NormalTok{))}
-
-\NormalTok{henderson\_f }\OtherTok{\textless{}{-}} \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{h=}\DecValTok{6}\NormalTok{)}\SpecialCharTok{@}\NormalTok{sfilter}
-\NormalTok{lp\_filter2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(icr, }\AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)\{}
-\NormalTok{  all\_coef }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(icr, }\ControlFlowTok{function}\NormalTok{(ic)\{}
-    \FunctionTok{lp\_filter}\NormalTok{(}\AttributeTok{horizon =}\NormalTok{ h,}
-              \AttributeTok{kernel =}\NormalTok{ kernel,}
-              \AttributeTok{endpoints =}\NormalTok{ method,}
-              \AttributeTok{ic =}\NormalTok{ ic)}
-\NormalTok{  \})}
-\NormalTok{  rfilters }\OtherTok{=} \FunctionTok{lapply}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{h, }\ControlFlowTok{function}\NormalTok{(i)\{}
-\NormalTok{    q}\OtherTok{=}\NormalTok{h }\SpecialCharTok{{-}}\NormalTok{i}
-\NormalTok{    all\_coef[[i]][,}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"q=\%i"}\NormalTok{, q)]}
-\NormalTok{  \})}
-  \FunctionTok{finite\_filters}\NormalTok{(henderson\_f, }\AttributeTok{rfilters =}\NormalTok{ rfilters)}
-\NormalTok{\}}
-\NormalTok{loc\_lc\_est }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-\NormalTok{loc\_lc\_if }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_slope }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_slope }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_slope }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"LC"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-\NormalTok{loc\_ql\_est }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-\NormalTok{    rjd3filters}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-\NormalTok{loc\_ql\_if }\OtherTok{\textless{}{-}}
-  \FunctionTok{lapply}\NormalTok{(last\_est, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    est\_loc\_concav }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{tail}\NormalTok{(est\_concav }\SpecialCharTok{*}\NormalTok{ x, }\DecValTok{6}\NormalTok{))}
-\NormalTok{    sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{var\_estimator}\NormalTok{(x, henderson\_f)}
-\NormalTok{    icr }\OtherTok{=} \DecValTok{2}\SpecialCharTok{/}\NormalTok{(}\FunctionTok{sqrt}\NormalTok{(pi) }\SpecialCharTok{*}\NormalTok{ (est\_loc\_concav }\SpecialCharTok{/} \FunctionTok{sqrt}\NormalTok{(sigma2)))}
-\NormalTok{    lp\_coef }\OtherTok{=} \FunctionTok{lp\_filter2}\NormalTok{(}\AttributeTok{ic =}\NormalTok{ icr,}
-                         \AttributeTok{method =} \StringTok{"QL"}\NormalTok{, }\AttributeTok{h =} \DecValTok{6}\NormalTok{, }\AttributeTok{kernel =} \StringTok{"Henderson"}\NormalTok{)}
-    \FunctionTok{implicit\_forecast}\NormalTok{(x, lp\_coef)}
-\NormalTok{  \})}
-
-\DocumentationTok{\#\# Plots}
-\CommentTok{\# Plots with all the estimates}
-\NormalTok{plot\_est }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
-\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, data)}
-\NormalTok{  joint\_data }\OtherTok{\textless{}{-}}
-    \FunctionTok{window}\NormalTok{(joint\_data,}
-           \AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(joint\_data))}
-
-\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
-    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =}\NormalTok{ last\_dates,}
-               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
-               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}
-
-\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data) }\SpecialCharTok{+} \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
-    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
-    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
-              \AttributeTok{data =}\NormalTok{ data\_legend,}
-              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
-              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{)  }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
-\NormalTok{\}}
-
-\NormalTok{plot\_prevs }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{ (data, }\AttributeTok{nperiod =} \DecValTok{6}\NormalTok{) \{}
-\NormalTok{  joint\_data }\OtherTok{\textless{}{-}} \FunctionTok{do.call}\NormalTok{(ts.union, }\FunctionTok{lapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) \{}
-\NormalTok{    first\_date }\OtherTok{\textless{}{-}} \FunctionTok{time}\NormalTok{(x)[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}} \FunctionTok{deltat}\NormalTok{(x)}
-    \CommentTok{\# The last observed data is added for readability}
-    \FunctionTok{ts}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{end =}\NormalTok{ first\_date), x),}
-       \AttributeTok{start =}\NormalTok{ first\_date, }\AttributeTok{frequency =} \FunctionTok{frequency}\NormalTok{(x))}
-\NormalTok{  \}))}
-
-\NormalTok{  data\_legend }\OtherTok{\textless{}{-}}
-    \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =} \FunctionTok{sapply}\NormalTok{(data, }\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{tail}\NormalTok{(}\FunctionTok{time}\NormalTok{(x), }\DecValTok{1}\NormalTok{)),}
-               \AttributeTok{y =} \FunctionTok{sapply}\NormalTok{(data, tail, }\DecValTok{1}\NormalTok{),}
-               \AttributeTok{label =} \FunctionTok{colnames}\NormalTok{(joint\_data))}
-\NormalTok{  forecast}\SpecialCharTok{::}\FunctionTok{autoplot}\NormalTok{(joint\_data, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
-\NormalTok{    forecast}\SpecialCharTok{::}\FunctionTok{autolayer}\NormalTok{(}
-      \FunctionTok{window}\NormalTok{(y, }\AttributeTok{start =}\NormalTok{ last\_dates[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{{-}}\NormalTok{ nperiod }\SpecialCharTok{*} \FunctionTok{deltat}\NormalTok{(y)),}
-      \AttributeTok{colour =} \ConstantTok{FALSE}
-\NormalTok{    ) }\SpecialCharTok{+}
-    \FunctionTok{theme\_bw}\NormalTok{() }\SpecialCharTok{+}
-    \FunctionTok{scale\_x\_continuous}\NormalTok{(}\AttributeTok{labels =}\NormalTok{ zoo}\SpecialCharTok{::}\NormalTok{as.yearmon) }\SpecialCharTok{+}
-    \FunctionTok{geom\_text}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y, }\AttributeTok{label =}\NormalTok{ label, }\AttributeTok{colour =}\NormalTok{ label),}
-              \AttributeTok{data =}\NormalTok{ data\_legend,}
-              \AttributeTok{check\_overlap =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{hjust =} \DecValTok{0}\NormalTok{, }\AttributeTok{nudge\_x =} \FloatTok{0.01}\NormalTok{,}
-              \AttributeTok{size =} \DecValTok{2}\NormalTok{, }\AttributeTok{inherit.aes =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"none"}\NormalTok{) }\SpecialCharTok{+}
-    \FunctionTok{labs}\NormalTok{(}\AttributeTok{x =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{y =} \ConstantTok{NULL}\NormalTok{)}
-\NormalTok{\}}
-
-\NormalTok{all\_est }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_est, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_est),}
-             \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_est))}
-\NormalTok{all\_if }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(lp\_if, }\FunctionTok{list}\NormalTok{(}\StringTok{"LC local param."} \OtherTok{=}\NormalTok{ loc\_lc\_if),}
-            \FunctionTok{list}\NormalTok{(}\StringTok{"QL local param."} \OtherTok{=}\NormalTok{ loc\_ql\_if))}
-\NormalTok{y\_lim }\OtherTok{\textless{}{-}} \ConstantTok{NULL}
-\NormalTok{all\_plots\_est }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
-  \FunctionTok{names}\NormalTok{(all\_est),}
-  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_est}\NormalTok{(all\_est[[x]]) }\SpecialCharTok{+}
-    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Trend{-}cycle with \%s"}\NormalTok{, x))) }\SpecialCharTok{+}
-    \FunctionTok{coord\_cartesian}\NormalTok{(}\AttributeTok{ylim =}\NormalTok{ y\_lim)}
-\NormalTok{)}
-\NormalTok{all\_plots\_prev }\OtherTok{\textless{}{-}} \FunctionTok{lapply}\NormalTok{(}
-  \FunctionTok{names}\NormalTok{(all\_if),}
-  \ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{plot\_prevs}\NormalTok{(all\_if[[x]]) }\SpecialCharTok{+}
-    \FunctionTok{ggtitle}\NormalTok{(latex2exp}\SpecialCharTok{::}\FunctionTok{TeX}\NormalTok{(}\FunctionTok{sprintf}\NormalTok{(}\StringTok{"Implicit forecasts with \%s"}\NormalTok{, x)))}
-\NormalTok{)}
-\CommentTok{\# Combine all plots:}
-\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_est, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
-\FunctionTok{wrap\_plots}\NormalTok{(all\_plots\_prev, }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+}
+
+\end{table}%
 

### 2024-08-22 paper/article_jos.qmd
vf

@@
     cite-method: biblatex
+    output-file: main_document
 crossref:
@@
 ```
+
+::: {.content-visible when-format="docx"}
+$$
+\DeclareMathOperator{\e}{e}
+\newcommand\transp[1]{{#1}^T}
+\newcommand\1{{\mathbb 1}}
+\newcommand{\E}[1]{\mathbb{E}\left[#1\right]}
+\newcommand\ud{\,\mathrm{d}}
+$$
+:::
 
@@
 $$
-The function $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
-The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}$.
+The function $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}$ is called the *transfer function* or *frequency response function*.
+The frequency response function can alternatively be defined as $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{i \omega k}$ or $\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{2\pi i \omega k}$.
 It can be rewritten as:
@@
 - If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.
-
-<!-- ### Noise reduction -->
-
-<!-- All time series are affected by noise, which can blur the extraction of trends and cycles. -->
-<!-- This is why we try to reduce this noise (by reducing its variance) while preserving the relevant trends. -->
-<!-- The sum of the squares of the coefficients $\sum_{k=-p}^{+f}\theta_k^2$ is the *reduction of variance* ratio. -->
-
-<!-- Indeed, let $\{\varepsilon_t\}$ be a sequence of independent random variables with $\E{\varepsilon_t}=0$, $\V{\varepsilon_t}=\sigma^2$: -->
-<!-- $$ -->
-<!-- \V{M_{\boldsymbol\theta}\varepsilon_t}=\V{\sum_{k=-p}^{+f} \theta_k \varepsilon_{t+k}} -->
-<!-- = \sum_{k=-p}^{+f} \theta_k^2 \V{\varepsilon_{t+k}}= -->
-<!-- \sigma^2\sum_{k=-p}^{+f} \theta_k^2. -->
-<!-- $$ -->
 

### 2024-08-22 paper/main_document.tex
vf

--- 2024-08-22 paper/article_jos.tex
@@
 \] The function
-\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{-i \omega k}\)
+\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{-i \omega k}\)
 is called the \emph{transfer function} or \emph{frequency response
@@
 as
-\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{i \omega k}\)
+\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{i \omega k}\)
 or
-\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k e^{2\pi i \omega k}\).
+\(\Gamma_{\boldsymbol\theta}(\omega)=\sum_{k=-p}^{+f} \theta_k \e^{2\pi i \omega k}\).
 It can be rewritten as: \[

### 2024-10-04 paper/article_jos.qmd
clarification

@@
 
-@Fig-graphstpsimul represents box plots of the phase shift: the box represents the interquartile range, the central horizontal line the median the dot the outliers and the whisker the range of the data (if no outlier).
+@Fig-graphstpsimul represents box plots of the phase shift: the box represents the interquartile range, the central horizontal line represents the median, the dot represents the outliers and the whisker the range of the data (if no outlier).
 For simulated series with medium variability (when a symmetric filter of length 13 is relevant), with the linear constant (LC) method, 50% of the turning point are detected with a phase shift of at most 5 months whereas when the LC filter is parametrised locally, 50% of the turning point are detected with a phase shift of at most 4 months.

### 2024-10-05 paper/article_jos.qmd
typos

@@
 Thus, the trend-cycle $TC_t$ is considered as deterministic and modelled as a polynomial trend of degree $d$.
-The coefficients $(\beta_0,\dots,\beta_k)$ can be estimated by weighted least squares.
+The coefficients $(\beta_0,\dots,\beta_d)$ can be estimated by weighted least squares.
 The estimate $\hat \beta_0$ provides the trend-cycle estimate $\widehat{TC}_t$ and it can be shown that this equivalent to apply of moving average (@sec-nonparamreg).
@@
 
-For symmetric filters, the phase shift function is equal to zero (modulo $\pi$).
+For symmetric filters, the phase shift function is equal to 0 (modulo $\pi$).
 So there is no delay in detecting turning points.
@@
 $$
-Hence, the filter $\boldsymbol w$ preserves deterministic polynomials of order $d$.
+Hence, the filter $\boldsymbol \theta$ preserves deterministic polynomials of order $d$.
 
@@
 #|   - "Symmetric moving average used for $\\hat \\delta_1$."
-#|   - "Asymmetric moving average using two future points ($q=2$) used for $\\hat \\delta_1$."
+#|   - "Asymmetric moving average using two future points ($q=2$) for $\\hat \\delta_1$."
 #| layout-ncol: 1
